### 2021-04-06 README.Rmd
Added readme

---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wagglefit

[![R-CMD-check](https://github.com/joseph-palmer/wagglefit/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/joseph-palmer/wagglefit/actions/workflows/check-standard.yaml)
[![lint](https://github.com/joseph-palmer/wagglefit/actions/workflows/lint.yaml/badge.svg)](https://github.com/joseph-palmer/wagglefit/actions/workflows/lint.yaml)
[![pkgdown](https://github.com/joseph-palmer/wagglefit/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/joseph-palmer/wagglefit/actions/workflows/pkgdown.yaml)
[![test-coverage](https://github.com/joseph-palmer/wagglefit/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/joseph-palmer/wagglefit/actions/workflows/test-coverage.yaml)

The goal of wagglefit is to allow for easy waggle dance analysis.

## Installation

You can install the released version of wagglefit from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("wagglefit")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("joseph-palmer/wagglefit")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(wagglefit)
## basic example code
```

### 2021-06-04 R/plotting.R
finished plotting and utils tests

@@ -125,6 +125,7 @@ make_full_plot <- function(x, model_result_list) {
 #' @export
 #'
 make_results_tibble <- function(result) {
+  . <- NULL
   all_tibble <- as_tibble(result$all)
   all_tibble$parameter <- c("p", "ls", "ln", "qn", "a")
   scout_tibble <- as_tibble(result$scout)
@@ -139,7 +140,7 @@ make_results_tibble <- function(result) {
   result_tibble <- result_tibble %>%
     pivot_wider(names_from = .data$parameter, values_from = .data$est) %>%
     mutate(
-      AIC = calc_aic(rowSums(!(is.na(.data$.))) - 2, .data$fmax)
+      AIC = calc_aic(rowSums(!(is.na(.))) - 2, .data$fmax)
     ) %>%
     mutate(
       p = ifelse(

### 2021-06-04 tests/testthat/test-plotting.R

@@ -42,7 +42,30 @@ test_make_base_plot <- function(x) {
 
 test_make_full_plot <- function(x, model_result_list) {
   plt <- make_full_plot(x, model_result_list)
-  print(plot)
+  expected_xlab <- "Foraging distance (Km)"
+  expected_ylab <- "Ln cumulative probability"
+  test_that("make_full_plot returns a ggplot object with the correct labels", {
+    expect_true(ggplot2::is.ggplot(plt))
+    expect_equal(expected_xlab, plt$labels$x)
+    expect_equal(expected_ylab, plt$labels$y)
+  })
+}
+
+test_make_results_tibble <- function(result) {
+  expected <- tibble::tibble(
+    fmax = c(-256, -256, -256),
+    data_name = c("all", "scout", "recruit"),
+    p = c(0.5, 1, 0),
+    ls = c(1.3, 1.3, NA),
+    ln = c(1.3, NA, 1.3),
+    qn = c(2.2, 2.2, 2.2),
+    a = c(0.5, 0.5, 0.5),
+    AIC = c(522, 518, 518)
+  )
+  actual <- make_results_tibble(result)
+  test_that("make_results_tibble returns the expected result", {
+    expect_identical(expected, actual)
+  })
 }
 
 plotting_tests <- function() {
@@ -60,14 +83,28 @@ plotting_tests <- function() {
   a <- 0.5
   param_est <- c(p, ls, ln, qn, a)
   example_result <- list(
-    est = param_est,
-    data_name = "all"
+    "all" = list(
+      est = param_est,
+      fmax = -256,
+      data_name = "all"
+    ),
+    "scout" = list(
+      est = c(ls, qn, a),
+      fmax = -256,
+      data_name = "scout"
+    ),
+    "recruit" = list(
+      est = c(ln, qn, a),
+      fmax = -256,
+      data_name = "recruit"
+    )
   )
 
   test_inverse_cdf(x)
   test_make_ccdf_plot_data(x, param_est, "all", 10)
   test_make_base_plot(x)
   test_make_full_plot(x, example_result)
+  test_make_results_tibble(example_result)
 }
 
 # run tests

### 2021-06-08 analysis/analysis.Rmd
created Rmd document for making manuscript figures

# Honeybee foraging analysis
## Joseph Palmer
---

### Synopsis

This document contains the analysis of models fit to the data using `run_all_models()`

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
library(cowplot)
library(dplyr)
library(tibble)
source("analysis/fit_data.R")
```

```{r}

path <- "analysis/results_version1"
data <- get_data()

sites <- purrr::map_chr(
  list.files(path),
  ~ {
    return(gsub(".rds", "", .x))
  }
)

plots_data <- purrr::map(
  sites,
  ~ {
    subdat <- data %>% dplyr::filter(.data$site == .x)
    df <- readRDS(paste0(path, "/", .x, ".rds"))[-3]
    results_tibble <- make_results_tibble(df) %>%
      rowwise() %>%
      mutate(
        ks_p = calc_ks(
          subdat$foraging_distance, c(p, ls, ln, qn, a),
          data_name, TRUE
        ),
        ks_d = calc_ks(
          subdat$foraging_distance, c(p, ls, ln, qn, a),
          data_name, FALSE
        ),
        site = .x
      ) %>%
      select(
        site,
        everything()
      )
    ret <- list(
      site_name = list(
        results_tibble,
        make_full_plot(subdat$foraging_distance, df)
      )
    )
    return(ret)
  }
)
```

```{r}
plots <- plots_data %>%
  purrr::map(
    ~ {
      .x[[1]][[2]]
    }
  )

all_data <- plots_data %>%
  purrr::map(
    ~ {
      .x[[1]][[1]]
    }
  ) %>%
  bind_rows()

aic_plot <- all_data %>%
  group_by(site) %>%
  slice(which.min(AIC)) %>%
  select(data_name) %>%
  group_by(data_name) %>%
  summarise(lowest_AIC = n()) %>%
  ggplot(aes(x = data_name, y = lowest_AIC)) +
  geom_bar(stat = "identity", fill = "#0084ff") +
  labs(x = "Model", y = "Number of lowest AIC sites")
```

### 2021-06-15 analysis/analysis.Rmd
added new figures, markdown and more plots

@@
 library(ggplot2)
+theme_set(
+  theme_classic() +
+    theme(
+      text = element_text(family = "URWHelvetica", size = 8)
+    )
+)
 library(cowplot)
@@
       mutate(
-        ks_p = calc_ks(
+        ks_p = calc_ks_boot(
           subdat$foraging_distance, c(p, ls, ln, qn, a),
@@
           subdat$foraging_distance, c(p, ls, ln, qn, a),
-          data_name, FALSE
+          data_name, TRUE
         ),
@@
       )
+    print(paste(.x, length(subdat$foraging_distance)))
+    if (.x == "STU") {
+      ret <- list(
+        site_name = list(
+          results_tibble,
+          make_full_plot(
+            subdat$foraging_distance, df,
+            subplot_coords = c(0.5, 6, -6, -2.5)
+          )
+        )
+      )
+    } else {
       ret <- list(
@@
       )
+    }
     return(ret)
@@
 ```{r}
+
 plots <- plots_data %>%
@@
     ~ {
-      .x[[1]][[2]]
+      site <- unique(.x[[1]][[1]]$site)
+      ret <- list(.x[[1]][[2]])
+      names(ret) <- site
+      return(ret)
     }
@@
   ggplot(aes(x = data_name, y = lowest_AIC)) +
-  geom_bar(stat = "identity", fill = "#0084ff") +
-  labs(x = "Model", y = "Number of lowest AIC sites")
+  geom_bar(stat = "identity") +
+  labs(x = "Model", y = "Number of lowest AIC sites") +
+  scale_y_continuous(breaks = seq(0, 20, by = 2))
+
+ggsave(
+  plot = aic_plot,
+  filename = "analysis/figures/AIC_plot.png",
+  width = 86,
+  height = 90,
+  units = "mm",
+  dpi = 400
+)
+
+ks_plot_dist <- all_data %>%
+  ggplot(aes(x = ks_p)) +
+  geom_histogram(bins = 10, binwidth = 0.1, col = "white") +
+  geom_vline(xintercept = 0.05, color = "red", linetype = "dashed") +
+  scale_y_continuous(breaks = seq(0, 12, by = 2)) +
+  labs(x = "KS P value") +
+  facet_wrap(~data_name, nrow = 3)
+
+stats_plot <- plot_grid(
+  aic_plot,
+  ks_plot_dist,
+  ncol = 2,
+  labels = c("A", "B")
+)
 ```
+
+Model fitting results to real data
+```{r}
+
+ggsave(
+  plot = plots[[20]][[1]],
+  filename = "analysis/figures/zsl_plot.png",
+  width = 86,
+  height = 90,
+  units = "mm",
+  dpi = 400
+)
+
+plots[[20]][[1]]
+
+ggsave(
+  plot = stats_plot,
+  filename = "analysis/figures/model_fits.png",
+  width = 178,
+  height = 200,
+  units = "mm",
+  dpi = 400
+)
+
+stats_plot
+
+ggsave(
+  plot = ks_plot_dist,
+  filename = "analysis/figures/sites_ks.png",
+  width = 86,
+  height = 180,
+  units = "mm",
+  dpi = 400
+)
+```
+
+### Map plots
+
+```{r}
+
+library(ggplot2)
+library(ggsn)
+library(sf)
+
+full_data_path <- "analysis/data/FullHBForagingData.csv"
+data_raw <- tibble(read.csv(full_data_path))
+
+map_data <- data_raw %>%
+  select(
+    site, lat, lon
+  ) %>%
+  distinct()
+
+map_data <- all_data %>%
+  group_by(site) %>%
+  slice(which.min(AIC)) %>%
+  select(site, data_name) %>%
+  left_join(map_data, on = "site")
+
+locations <- st_as_sf(
+  map_data,
+  coords = c("lon", "lat"), crs = 4326
+)
+
+base <- st_read("analysis/shapefiles/english_region_region.shp")
+
+# location of sites box
+locations
+selected_sites <- filter(locations, site %in% c("STU", "ZSL"))
+points_area <- st_as_sfc(st_bbox(locations))
+
+# cropped
+croped_data <- st_crop(
+  base,
+  xmin = 450000,
+  xmax = 650000,
+  ymin = 91000,
+  ymax = 230000
+)
+
+zoom_plt <- ggplot() +
+  geom_sf(data = croped_data, fill = "#e5f5f9") +
+  geom_sf(data = locations, aes(colour = data_name), size = 1) +
+  # geom_sf_text(
+  #   data = selected_sites,
+  #   aes(label = site, family = "URWHelvetica"),
+  #   size = 2.5,
+  #   nudge_x = -5000,
+  #   nudge_y = 3000
+  # ) +
+  theme_void() +
+  theme(
+    text = element_text(family = "URWHelvetica", size = 8),
+    legend.position = "bottom"
+  )
+
+plt <- ggplot() +
+  geom_sf(data = base, fill = "white") +
+  geom_sf(data = points_area, fill = NA, colour = "blue") +
+  theme_void() +
+  theme(
+    text = element_text(family = "URWHelvetica", size = 8),
+    legend.position = "none"
+  )
+
+merge_plot <- ggdraw() +
+  draw_plot(ggplotGrob(zoom_plt)) +
+  draw_plot(plt, x = 0.03, y = 0.3, width = 0.35, height = 0.35)
+
+merge_plot
+
+ggsave(
+  plot = merge_plot,
+  file = "analysis/figures/map_aic.png",
+  width = 86,
+  height = 90,
+  units = "mm",
+  dpi = 400
+)
+
+ggsave(
+  plot = plots[[20]][[1]],
+  file = "analysis/figures/ZSL.png",
+  width = 86,
+  height = 90,
+  units = "mm",
+  dpi = 400
+)
+
+ggsave(
+  plot = plots[[17]][[1]],
+  file = "analysis/figures/STU.png",
+  width = 86,
+  height = 90,
+  units = "mm",
+  dpi = 400
+)
+```

### 2021-06-15 R/utils.R

@@ -116,3 +116,30 @@ calc_ks <- function(x, param_est, model_type, pvalue = TRUE) {
     return(ks$statistic)
   }
 }
+
+#' Run a bootstapped KS test on the data
+#'
+#' @description Calulctate the ks test stat and p value for a given model
+#' @param x doubleArray Foraging distance data
+#' @param param_est doubleArray The parameter estimates for a model
+#' @param model_type characterArray the name of the model ran
+#' @param pvalue Bool To return the p value (TRUE) or D statistic (FALSE).
+#' Defaults to TRUE.
+#' @importFrom stats ks.test
+#' @importFrom rlang .data
+#' @importFrom Matching ks.boot
+#' @concept utility
+calc_ks_boot <- function(x, param_est, model_type, pvalue = TRUE) {
+  if (model_type != "all") {
+    param_est <- param_est[-1]
+  }
+  param_est <- param_est[!is.na(param_est)]
+  ccdf_data <- inverse_ccdf(x)
+  ccdf_model <- model_ccdf(x, param_est, model_number_from_model(model_type))
+  if (all(is.nan(ccdf_model))) {
+    print(paste(model_type, "has bad results"))
+    return(NA)
+  }
+  ks <- ks.boot(ccdf_data$prob, ccdf_model)
+  return(ks$ks.boot.pvalue)
+}

### 2021-06-15 R/plotting.R
added bootstrapped ks test and updated plotting

@@ -49,7 +49,8 @@ make_ccdf_plot_data <- function(x, param_est, model, npoints = 100) {
 #' Creates ccdf plot of data
 #'
 #' @description Creates a ccdf plot for the data provided
-#' @param x The foraging distance to plot
+#' @param x doubleArray The foraging distance to plot
+#' @param logit Bool to log the probabilities or not.
 #' @return ggplot plot of cumulative probability for foraging distances in Km
 #' @importFrom ggplot2 ggplot aes_string geom_point geom_line theme_set theme
 #' theme_classic element_text labs
@@ -58,17 +59,24 @@ make_ccdf_plot_data <- function(x, param_est, model, npoints = 100) {
 #' @importFrom rlang .data
 #' @export
 #'
-make_base_plot <- function(x) {
+make_base_plot <- function(x, logit = TRUE) {
   theme_set(
     theme_classic() +
       theme(
         text = element_text(family = "URWHelvetica", size = 8)
       )
   )
-  plt <- inverse_ccdf(x) %>%
-    ggplot(aes(x = .data$sd, y = log(.data$prob))) +
-    geom_point() +
-    labs(x = "Foraging distance (Km)", y = "Ln cumulative probability")
+  if (logit) {
+    plt <- inverse_ccdf(x) %>%
+      ggplot(aes(x = .data$sd, y = log(.data$prob))) +
+      geom_point() +
+      labs(x = "Foraging distance (Km)", y = "Ln cumulative probability")
+  } else {
+    plt <- inverse_ccdf(x) %>%
+      ggplot(aes(x = .data$sd, y = .data$prob)) +
+      geom_point() +
+      labs(x = "Foraging distance (Km)", y = "Ln cumulative probability")
+  }
   return(plt)
 }
 
@@ -78,16 +86,21 @@ make_base_plot <- function(x) {
 #' @param x The foraging distance to plot
 #' @param model_result_list namedlist The return of `fit`, a list of parameter
 #' estimates ($est) and the data name ($data_name).
+#' @param logit Bool To log the data or not. Defaults to TRUE
+#' @param subplot_coords doubleArray Coordinates for location of inner histogram
 #' @return ggplot plot of cumulative probability of foraging distances in Km
 #' along with the model fits to this data.
-#' @importFrom ggplot2 ggplot aes geom_line geom_line theme_set theme
-#' theme_classic element_text
[…]

### 2021-06-18 analysis/agent_based_sim.Rmd
added simulation python code and linked with R

# Honeybee foraging simulation
## Joseph Palmer
---

The code for the honeybee foraging simulation is writen in python and is contained in the `Simulation` directory (specifically `Simulation/Code`). The results of the simulation (as output csv files) are stored in `Results/Results`. Generating the results can be done by running the script `run.py` in `Simulation/Code`. Once you have ran the simulation you can run through this Rmd doc to generate the figures. To run the python code you will need python3 installed, along with the dependencies listed in **init.py?**.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(family = "URWHelvetica", size = 8)
    )
)
library(dplyr)
library(tibble)
library(Matching)
devtools::load_all()
```

### Extracting the results

```{r, run_sim}

recruit <- read.csv(
  "Simulation/Results/recruit_distribution_v2.csv",
  header = FALSE
)
scout <- read.csv(
  "Simulation/Results/scout_distribution_v2.csv",
  header = FALSE
)

recruit$model <- "Recruit"
scout$model <- "Scout"

print(paste("Recruit:", length(recruit$V1), "Scout:", length(scout$V1)))
```

### Comparison of foraging distance patterns between foraging strategies

For the foraging distances simulated under our different scenarios we evaluate the hypothesis that these two distributions are different and attempt to describe the observed patterns.

To start with describing the obseved pattern we first fit an exponential to the data. The MLE of an exponential distribution is derived as the reciprocal of the data mean.

```{r}

get_exp_prediction <- function(x, n = 100) {
  rate <- 1 / mean(x)
  pred_x <- seq(0, max(x), length.out = n)
  exp_prediction <- purrr::map_dbl(
    pred_x,
    ~ {
      1 - exp(-rate * .x)
    }
  )
  pred <- tibble(x = pred_x, y = 1 - exp_prediction)
  return(pred)
}

scout_cdf <- inverse_ccdf(scout$V1 - min(scout$V1))
pred <- get_exp_prediction(scout_cdf$sd)

# log plot
ggplot(scout_cdf, aes(x = sd, y = log(prob))) +
  geom_point() +
  geom_line(aes(x = x, y = log(y)), pred, colour = "blue") +
  labs(
    x = "Foraging distance (km)",
    y = "count"
  )

# nonlog plot
ggplot(scout_cdf, aes(x = sd, y = prob)) +
  geom_point() +
  geom_line(aes(x = x, y = y), pred, colour = "blue") +
  labs(
    x = "Foraging distance (km)",
    y = "count"
  )

# assessing fit of exponential using sampled KS test
actual_p <- ks.test(scout_cdf$sd, "pexp", 1 / mean(scout_cdf$sd))$p.value
simulated_p <- purrr::map_dbl(
  seq_len(10000),
  ~ {
    z <- rexp(length(scout_cdf$sd), 1 / mean(scout_cdf$sd))
    ks.test(z, "pexp", 1 / mean(z))$p.value
  }
)

ggplot(tibble(x = simulated_p), aes(x = x)) +
  geom_histogram() +
  geom_vline(xintercept = actual_p)

mean(simulated_p <= actual_p)
```

```{r}
# recruit
recruit_cdf <- inverse_ccdf(recruit$V1 - min(recruit$V1))
pred <- get_exp_prediction(recruit_cdf$sd)

# log plot
ggplot(recruit_cdf, aes(x = sd, y = log(prob))) +
  geom_point() +
  geom_line(aes(x = x, y = log(y)), pred, colour = "blue") +
  labs(
    x = "Foraging distance (km)",
    y = "count"
  )

# nonlog plot
ggplot(recruit_cdf, aes(x = sd, y = prob)) +
  geom_point() +
  geom_line(aes(x = x, y = y), pred, colour = "blue") +
  labs(
    x = "Foraging distance (km)",
    y = "count"
  )

# assessing fit of exponential using sampled KS test
actual_p <- ks.test(scout_cdf$sd, "pexp", 1 / mean(scout_cdf$sd))$p.value
simulated_p <- purrr::map_dbl(
  seq_len(1000),
  ~ {
    z <- rexp(length(scout_cdf$sd), 1 / mean(scout_cdf$sd))
    ks.test(z, "pexp", 1 / mean(z))$p.value
  }
)

ggplot(tibble(x = simulated_p), aes(x = x)) +
  geom_histogram() +
  geom_vline(xintercept = actual_p)

mean(simulated_p <= actual_p)
```

### Comparison of the two foraging distance distributions

```{r}
ks <- ks.boot(scout_cdf$sd, recruit_cdf$sd)
print(ks)
```

The distributions are significantly different to each other (Bootstrapped two-sided Kolmogorov-Smirnov test: d = 0.1039, p < 0.001). Whilst the KS test can determine if two data sets stem from different distributions, the stat will also show a different if the data come from the same distribution with different parameter estimates (e.g. exponential with different rates).

```{r}
# testing

x <- inverse_ccdf(rexp(10000, 1.8))
y <- inverse_ccdf(rnorm(1000, mean = 5, sd = 1))
pred <- get_exp_prediction(x$sd)

ggplot(x, aes(x = sd, y = log(prob))) +
  geom_point() +
  geom_line(aes(x = x, y = log(y)), pred, colour = "blue") +
  labs(
    x = "Foraging distance (km)",
    y = "count"
  )

ks.test(x$sd, "pexp", 1 / mean(x$sd))

# assessing fit of exponential using KS test with simulation
actual_d <- ks.test(x$sd, "pexp", 1 / mean(x$sd))$p.value
simulated_d <- purrr::map_dbl(
  seq_len(1000),
  ~ {
    z <- rexp(length(x$sd), 1 / mean(x$sd))
    ks.test(z, "pexp", 1 / mean(z))$p.value
  }
)

ggplot(tibble(x = simulated_d), aes(x = x)) +
  geom_histogram() +
  geom_vline(xintercept = actual_d)

mean(simulated_d <= actual_d)
```

### 2021-06-18 Simulation/Code/run.py

@@ -1,0 +1,193 @@
+#!/usr/bin/env python3
+"""Run simulation of honeybee foraging"""
+__appname__  = "run.py"
+__author__   = "Joseph Palmer <joseph.palmer18@imperial.ac.uk>"
+__version__  = "0.0.2"
+__date__     = "02-2020"
+
+import numpy as np
+import pandas as pd
+import scipy.stats
+import matplotlib.pyplot as plt
+import Agents
+import Methods
+import config
+
+def create_flowers(lwr, upr, res, quality_array):
+    # Simulation window parameters
+    r = upr-lwr
+    xx0 = 0
+    yy0 = 0
+
+    areaTotal = np.pi*r**2
+
+    # Point process parameters
+    lambda0 = res
+
+    # Simulate Poisson point process
+    numbPoints = scipy.stats.poisson(lambda0*areaTotal).rvs()
+    theta = 2*np.pi*scipy.stats.uniform.rvs(0,1,((numbPoints,1)))
+    rho = r*np.sqrt(scipy.stats.uniform.rvs(0,1,((numbPoints,1))))
+
+    # Convert from polar to Cartesian coordinates
+    xx = rho * np.cos(theta)
+    yy = rho * np.sin(theta)
+
+    # Shift centre of disk to (xx0,yy0)
+    xx=xx+xx0
+    yy=yy+yy0
+
+    # Set locations as flowers
+    locations = np.column_stack((xx, yy))
+    flowers = np.apply_along_axis(Agents.Flower, 1, locations, quality_array)
+    fdict = {
+        i : flowers[i] for i in range(res)
+    }
+    return fdict
+
+def initialize(lwr, upr, res, nbees, nscouts, quality_array):
+    # create flowers in environment and store globaly.
+    config.FLOWERDICT = create_flowers(lwr, upr, res, quality_array)
+
+    # create honeybee population
+    nrecruits = nbees-nscouts
+    scouts = np.array([Agents.Scout() for i in range(nscouts)])
+    recruits = np.array([Agents.Recruit() for i in range(nrecruits)])
+    return (scouts, recruits)
+
+def flower_turnover(n, lwr, upr, k, quality_array):
+    # select n flowers and remove
+    assert (n <= len(config.FLOWERDICT.keys())), "N > flowers in environment!"
+    rmv_idx = np.random.choice(list(config.FLOWERDICT.keys()), n, replace=False)
[…]

### 2021-06-29 analysis/agent_based_sim.Rmd
updated simulation plots to fit exp and minexp

@@
 
+# model functions
+exp_loglike <- function(x) {
+  rate <- 1 / mean(x)
+  exp_model <- rate * exp(-rate * x)
+  return(sum(log(exp_model)))
+}
+
+expmin_loglike <- function(x) {
+  rate <- 1 / (pi * (mean(x^2) - min(x)^2))
+  expmin_model <- 2 * pi * rate * x * exp(-pi * rate * x^2)
+  return(sum(log(expmin_model)))
+}
+
+# ccdf functions
+get_exp_prediction <- function(x, n = 100) {
+  rate <- 1 / mean(x)
+  print(rate)
+  pred_x <- seq(min(x), max(x), length.out = n)
+  exp_prediction <- 1 - exp(-rate * pred_x)
+  pred <- tibble(x = pred_x, y = 1 - exp_prediction)
+  return(pred)
+}
+
+get_expmin_prediction <- function(x, n = 100) {
+  rate <- 1 / (pi * (mean(x^2) - min(x)^2))
+  print(rate)
+  pred_x <- seq(min(x), max(x), length.out = n)
+  expmin_prediction <- 1 - exp(-pi * rate * pred_x^2)
+  pred <- tibble(x = pred_x, y = 1 - expmin_prediction)
+  return(pred)
+}
+
+# load data
 recruit <- read.csv(
@@
 
-### Comparison of foraging distance patterns between foraging strategies
+### Comparing model fits to scout and recruit data
 
-For the foraging distances simulated under our different scenarios we evaluate the hypothesis that these two distributions are different and attempt to describe the observed patterns.
-
-To start with describing the obseved pattern we first fit an exponential to the data. The MLE of an exponential distribution is derived as the reciprocal of the data mean.
+We will fit an exponential and minimum of an exponential to each model.
+If our theory holds water the exponential will fit to the scouts best and the minimum of an exponential will fit to the recruits best.
+To be fit with the exponential function we first need to 'normalise' the data by subtracting the minimum foraging distance for all foraging distances.
+This is to ensure the exonential starts at 0 rather than stepping out to the minimum value and skewing the model fit.
+In plotting we take the values for the exponential ad add on the minimum value to each point to get the model fit data back in line with the actual data so they can be compared graphically.
+The data does not need to be transformed for the minimum of the exponential in the 2d plane as this minimum value is incorporated into the model.
 
 ```{r}
-
-get_exp_prediction <- function(x, n = 100) {
-  rate <- 1 / mean(x)
-  pred_x <- seq(0, max(x), length.out = n)
-  exp_prediction <- purrr::map_dbl(
-    pred_x,
-    ~ {
-      1 - exp(-rate * .x)
-    }
-  )
-  pred <- tibble(x = pred_x, y = 1 - exp_prediction)
-  return(pred)
-}
-
-scout_cdf <- inverse_ccdf(scout$V1 - min(scout$V1))
-pred <- get_exp_prediction(scout_cdf$sd)
+scout_cdf <- inverse_ccdf(scout$V1)
+pred_exp <- get_exp_prediction(scout$V1 - min(scout$V1))
+pred_expmin <- get_expmin_prediction(scout$V1)
 
 # log plot
-ggplot(scout_cdf, aes(x = sd, y = log(prob))) +
-  geom_point() +
-  geom_line(aes(x = x, y = log(y)), pred, colour = "blue") +
+scout_fit <- ggplot(scout_cdf, aes(x = sd, y = log(prob))) +
+  geom_point(size = 0.2) +
+  geom_line(
+    aes(x = x + min(scout$V1), y = log(y)), pred_exp,
+    colour = "#00bfc4"
+  ) +
+  geom_line(aes(x = x, y = log(y)), pred_expmin, colour = "#f8766d") +
   labs(
-    x = "Foraging distance (km)",
-    y = "count"
+    x = "Simulated foraging distance (km)",
+    y = "ln Inverse cummulative frequency"
   )
 
-# nonlog plot
-ggplot(scout_cdf, aes(x = sd, y = prob)) +
-  geom_point() +
-  geom_line(aes(x = x, y = y), pred, colour = "blue") +
-  labs(
-    x = "Foraging distance (km)",
-    y = "count"
+ggsave(
+  plot = scout_fit,
+  filename = "Simulation/Results/scout_fit.png",
+  width = 70.1,
+  height = 81.79,
+  units = "mm",
+  dpi = 400
 )
 
-# assessing fit of exponential using sampled KS test
-actual_p <- ks.test(scout_cdf$sd, "pexp", 1 / mean(scout_cdf$sd))$p.value
-simulated_p <- purrr::map_dbl(
-  seq_len(10000),
-  ~ {
-    z <- rexp(length(scout_cdf$sd), 1 / mean(scout_cdf$sd))
-    ks.test(z, "pexp", 1 / mean(z))$p.value
+ll_exp <- exp_loglike(scout$V1 - min(scout$V1))
+ll_expmin <- expmin_loglike(scout$V1)
+s_aic <- calc_aic(1, ll_exp)
+r_aic <- calc_aic(1, ll_expmin)
+if (s_aic < r_aic) {
+  print("Scout model is the best")
+} else {
+  print("Recruit model is the best")
 }
-)
-
-ggplot(tibble(x = simulated_p), aes(x = x)) +
-  geom_histogram() +
-  geom_vline(xintercept = actual_p)
-
-mean(simulated_p <= actual_p)
+print(paste("Scout loglike:", ll_exp, "Recruit loglike:", ll_expmin))
+print(paste("Scout AIC:", s_aic, "Recruit AIC:", r_aic))
 ```
@@
 ```{r}
-# recruit
-recruit_cdf <- inverse_ccdf(recruit$V1 - min(recruit$V1))
-pred <- get_exp_prediction(recruit_cdf$sd)
+recruit_cdf <- inverse_ccdf(recruit$V1)
+pred_exp <- get_exp_prediction(recruit$V1 - min(recruit$V1))
+pred_expmin <- get_expmin_prediction(recruit$V1)
 
 # log plot
-ggplot(recruit_cdf, aes(x = sd, y = log(prob))) +
-  geom_point() +
-  geom_line(aes(x = x, y = log(y)), pred, colour = "blue") +
+recruit_fit <- ggplot(recruit_cdf, aes(x = sd, y = log(prob))) +
+  geom_point(size = 0.2) +
+  geom_line(
+    aes(x = x + min(recruit$V1), y = log(y)),
+    pred_exp,
+    colour = "#00bfc4"
+  ) +
+  geom_line(aes(x = x, y = log(y)), pred_expmin, colour = "#f8766d") +
   labs(
-    x = "Foraging distance (km)",
-    y = "count"
+    x = "Simulated foraging distance (km)",
+    y = "ln Inverse cummulative frequency"
   )
 
-# nonlog plot
-ggplot(recruit_cdf, aes(x = sd, y = prob)) +
-  geom_point() +
-  geom_line(aes(x = x, y = y), pred, colour = "blue") +
-  labs(
-    x = "Foraging distance (km)",
-    y = "count"
+ggsave(
+  plot = recruit_fit,
+  filename = "Simulation/Results/recruit_fit.png",
+  width = 70.1,
+  height = 81.79,
+  units = "mm",
+  dpi = 400
 )
 
-# assessing fit of exponential using sampled KS test
-actual_p <- ks.test(scout_cdf$sd, "pexp", 1 / mean(scout_cdf$sd))$p.value
-simulated_p <- purrr::map_dbl(
-  seq_len(1000),
-  ~ {
-    z <- rexp(length(scout_cdf$sd), 1 / mean(scout_cdf$sd))
-    ks.test(z, "pexp", 1 / mean(z))$p.value
+ll_exp <- exp_loglike(recruit$V1 - min(recruit$V1))
+ll_expmin <- expmin_loglike(recruit$V1)
+s_aic <- calc_aic(1, ll_exp)
+r_aic <- calc_aic(1, ll_expmin)
+if (s_aic < r_aic) {
+  print("Scout model is the best")
+} else {
+  print("Recruit model is the best")
 }
-)
-
-ggplot(tibble(x = simulated_p), aes(x = x)) +
-  geom_histogram() +
-  geom_vline(xintercept = actual_p)
-
-mean(simulated_p <= actual_p)
+print(paste("Scout loglike:", ll_exp, "Recruit loglike:", ll_expmin))
+print(paste("Scout AIC:", s_aic, "Recruit AIC:", r_aic))
 ```
 
-### Comparison of the two foraging distance distributions
-
-```{r}
-ks <- ks.boot(scout_cdf$sd, recruit_cdf$sd)
-print(ks)
-```
-
-The distributions are significantly different to each other (Bootstrapped two-sided Kolmogorov-Smirnov test: d = 0.1039, p < 0.001).
-Whilst the KS test can determine if two data sets stem from different distributions, the stat will also show a different if the data come from the same distribution with different parameter estimates (e.g.
-exponential with different rates).
-
-```{r}
-# testing
-
-x <- inverse_ccdf(rexp(10000, 1.8))
-y <- inverse_ccdf(rnorm(1000, mean = 5, sd = 1))
-pred <- get_exp_prediction(x$sd)
-
-ggplot(x, aes(x = sd, y = log(prob))) +
-  geom_point() +
-  geom_line(aes(x = x, y = log(y)), pred, colour = "blue") +
-  labs(
-    x = "Foraging distance (km)",
-    y = "count"
-  )
-
-ks.test(x$sd, "pexp", 1 / mean(x$sd))
-
-# assessing fit of exponential using KS test with simulation
-actual_d <- ks.test(x$sd, "pexp", 1 / mean(x$sd))$p.value
-simulated_d <- purrr::map_dbl(
-  seq_len(1000),
-  ~ {
-    z <- rexp(length(x$sd), 1 / mean(x$sd))
-    ks.test(z, "pexp", 1 / mean(z))$p.value
-  }
-)
-
-ggplot(tibble(x = simulated_d), aes(x = x)) +
-  geom_histogram() +
-  geom_vline(xintercept = actual_d)
-
-mean(simulated_d <= actual_d)
-```
+From this analysis we observe that the distribution of foraging distances reported by scouts is better described by an exponential than the minimum of an exponential, whereas the converse is true for foraging distances reported by reruits.
+These results support our hypothesis for how scout and recruit reported foraging distances occur through different processes and produce different patterns.

### 2021-06-29 Simulation/Code/run.py

@@ -181,12 +181,12 @@ def main():
             [total_sdist, np.random.choice(sdist, 60, replace = False)]
         )
 
-    sdist_path = "../Results/scout_distribution_v2.csv"
-    rdist_path = "../Results/recruit_distribution_v2.csv"
-    dists = [total_sdist, total_rdist]
-    paths = [sdist_path, rdist_path]
-    for i in range(2):
-        np.savetxt(paths[i], dists[i], delimiter = ",")
+    # sdist_path = "../Results/scout_distribution_v2.csv"
+    # rdist_path = "../Results/recruit_distribution_v2.csv"
+    # dists = [total_sdist, total_rdist]
+    # paths = [sdist_path, rdist_path]
+    # for i in range(2):
+    #     np.savetxt(paths[i], dists[i], delimiter = ",")
     return 0
 
 if __name__ == "__main__":

### 2021-08-20 analysis/proportions-of-scouts.Rmd
added some basic plots exploring the proportion of scouts between rural and urban landscapes. not seeing much of a significant difference

# Analysis of the proportion of scouts across environmental gradients
## Joseph Palmer
---

To start I am just having a play with the proportion of scouts from each model.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(family = "URWHelvetica", size = 8)
    )
)
library(cowplot)
library(dplyr)
library(tibble)
source("analysis/fit_data.R")
```

```{r}

# load analysis results
path <- "analysis/results_version1"
data <- get_data()
sites <- purrr::map_chr(
  list.files(path),
  ~ {
    return(gsub(".rds", "", .x))
  }
)

# read analysis results into a single dataframe
model_results <- purrr::map(
    sites,
    ~ {
      subdat <- data %>% dplyr::filter(.data$site == .x)
      df <- readRDS(paste0(path, "/", .x, ".rds"))[-3]
      results_tibble <- make_results_tibble(df)
      results_tibble$site = .x
      return(results_tibble)
    }
  ) %>%
  bind_rows()

# load raw data
df <- read.table('analysis/data/FullHBForagingData.csv', header = TRUE,  sep = ',',  stringsAsFactors = FALSE)

# filter for landscape type
df_site_info <- df %>% select(site, LU) %>% unique()

# join model results with landscape type
model_results_pscouts <- model_results %>%
  filter(data_name == "all") %>%
  inner_join(df_site_info, by = c("site" = "site"))

# point plot of proportion of scouts for each site, coloured by landscape type
model_results_pscouts %>%
  ggplot(aes(x = site, y = p)) +
  geom_point(aes(color = LU))

```

The preceding code block show the proportion of scouts in each urban and rural site.

Lets look at summary stats for each landscape tyepe.

```{r}

model_results_pscouts %>%
  ggplot(aes(x = LU, y = p)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center")

```

There is a much higher spread in the rural sites, but no real difference in the proportion of scouts between each groups.

Need to work more on this and write more on what we should do with this....

Also need to ask vincent what he thinks of this chapter idea. Send him an overview and some data to look over before our next meeting.

### 2021-09-02 analysis/analysis.Rmd
removed p from optimisation to focus on others, need to optimise p after

@@
             subplot_coords = c(0.5, 6, -6, -2.5)
+          )
+        )
+      )
+    } else if (.x == "CAD") {
+      ret <- list(
+        site_name = list(
+          results_tibble,
+          make_full_plot(
+            subdat$foraging_distance, df,
+            subplot_coords = c(0.8, 2.2, -5, -2)
           )

### 2021-09-02 analysis/fit_models.Rmd
removed p from optimisation to focus on others, need to optimise p after

# Fitting waggle dance models to waggle dance data
## Joseph Palmer
---

```{r, include = FALSE}
devtools::load_all()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(family = "URWHelvetica", size = 8)
    )
)
library(cowplot)
library(dplyr)
library(tibble)
source("analysis/fit_data.R")
```

### 2021-09-02 analysis/proportions-of-scouts.Rmd
removed p from optimisation to focus on others, need to optimise p after

@@
 
+# keep only best fitting model
+model_results <- model_results %>% group_by(site) %>% top_n(-1, AIC)
+
 # load raw data
@@
 Also need to ask vincent what he thinks of this chapter idea. Send him an overview and some data to look over before our next meeting.
+
+```{r}
+# bootstrapped mean / median
+boot.median <- function(data, n = 10000) {
+  resamples <- purrr::map_dbl(
+    seq_len(n),
+    ~ {
+      median(sample(data$p, replace = TRUE))
+    }
+  )
+  std.err <- sqrt(var(resamples))
+  return(list(std.err = std.err, medians = resamples))
+}
+
+urban.boot <- model_results_pscouts %>%
+  filter(LU == "Urban") %>%
+  select(p) %>%
+  boot.median()
+
+hist(urban.boot$medians)
+
+rural.boot <- model_results_pscouts %>%
+  filter(LU == "Rural") %>%
+  select(p) %>%
+  boot.median()
+
+hist(rural.boot$medians)
+
+```
+
+```{r}
+
+# beta regression
+library(betareg)
+
+model_results_pscouts
+
+betamod <- betareg(p ~ LU, data = model_results_pscouts)
+summary(betamod)
+model_results_pscouts$predbeta = predict(betamod, interval = "confidence")
+
+ggplot(data = model_results_pscouts, aes(x = LU, y = p)) +
+  geom_point() +
+  geom_point(aes(y = predbeta, color = LU))
+
+```
+
+The above plot (code to make the beta plot) fits a beta regression.
+The beta regression is to be used when the predictor variables are proprtions.
+Whilst there is no difference between the proportion of scouts in urban and rural landscapes, there is a much tighter spread in the urban colonies than the rural colonies.

### 2021-09-02 R/RcppExports.R

@@ -86,13 +86,12 @@ model_ccdf <- function(x, params, model = 0L) {
 #'
 #' @param x double Foraging distance
 #' @param m double Minimum foraging distance
-#' @param p double Proportion of scouts (0<=p<=1)
 #' @param ls double Scout rate
 #' @param qn double Quality
 #' @param a double alpha value
 #' @export
-scout_dist <- function(x, m, p, ls, qn, a) {
-    .Call('_wagglefit_scout_dist', PACKAGE = 'wagglefit', x, m, p, ls, qn, a)
+scout_dist <- function(x, m, ls, qn, a) {
+    .Call('_wagglefit_scout_dist', PACKAGE = 'wagglefit', x, m, ls, qn, a)
 }
 
 #' Model function for recruits
@@ -100,13 +99,14 @@ scout_dist <- function(x, m, p, ls, qn, a) {
 #' @param ln double Recruit rate
 #' @inheritParams scout_dist
 #' @export
-recruit_dist <- function(x, m, p, ln, qn, a) {
-    .Call('_wagglefit_recruit_dist', PACKAGE = 'wagglefit', x, m, p, ln, qn, a)
+recruit_dist <- function(x, m, ln, qn, a) {
+    .Call('_wagglefit_recruit_dist', PACKAGE = 'wagglefit', x, m, ln, qn, a)
 }
 
 #' Log-likelihood function for scout and recruit superposition
 #'
 #' @param x double* Pointer to array of foraging distances
+#' @param p double Proportion of scouts (0<=p<=1)
 #' @inheritParams scout_dist
 #' @inheritParams recruit_dist
 #' @export

### 2021-09-02 src/RcppExports.cpp

@@ -107,34 +107,32 @@ BEGIN_RCPP
 END_RCPP
 }
 // scout_dist
-double scout_dist(double x, double m, double p, double ls, double qn, double a);
-RcppExport SEXP _wagglefit_scout_dist(SEXP xSEXP, SEXP mSEXP, SEXP pSEXP, SEXP lsSEXP, SEXP qnSEXP, SEXP aSEXP) {
+double scout_dist(double x, double m, double ls, double qn, double a);
+RcppExport SEXP _wagglefit_scout_dist(SEXP xSEXP, SEXP mSEXP, SEXP lsSEXP, SEXP qnSEXP, SEXP aSEXP) {
 BEGIN_RCPP
     Rcpp::RObject rcpp_result_gen;
     Rcpp::RNGScope rcpp_rngScope_gen;
     Rcpp::traits::input_parameter< double >::type x(xSEXP);
     Rcpp::traits::input_parameter< double >::type m(mSEXP);
-    Rcpp::traits::input_parameter< double >::type p(pSEXP);
     Rcpp::traits::input_parameter< double >::type ls(lsSEXP);
     Rcpp::traits::input_parameter< double >::type qn(qnSEXP);
     Rcpp::traits::input_parameter< double >::type a(aSEXP);
-    rcpp_result_gen = Rcpp::wrap(scout_dist(x, m, p, ls, qn, a));
+    rcpp_result_gen = Rcpp::wrap(scout_dist(x, m, ls, qn, a));
     return rcpp_result_gen;
 END_RCPP
 }
 // recruit_dist
-double recruit_dist(double x, double m, double p, double ln, double qn, double a);
-RcppExport SEXP _wagglefit_recruit_dist(SEXP xSEXP, SEXP mSEXP, SEXP pSEXP, SEXP lnSEXP, SEXP qnSEXP, SEXP aSEXP) {
+double recruit_dist(double x, double m, double ln, double qn, double a);
+RcppExport SEXP _wagglefit_recruit_dist(SEXP xSEXP, SEXP mSEXP, SEXP lnSEXP, SEXP qnSEXP, SEXP aSEXP) {
 BEGIN_RCPP
     Rcpp::RObject rcpp_result_gen;
     Rcpp::RNGScope rcpp_rngScope_gen;
     Rcpp::traits::input_parameter< double >::type x(xSEXP);
     Rcpp::traits::input_parameter< double >::type m(mSEXP);
[…]

### 2021-09-02 src/models.cpp

@@ -20,6 +20,7 @@ typedef struct {
     bool verbose;
     int fcount;
     double best_est;
+    double p;
 } data_struct;
 
 /*
@@ -59,15 +60,13 @@ void check_optimise_result(int optimise_result, bool verbose = true) {
 //'
 //' @param x double Foraging distance
 //' @param m double Minimum foraging distance
-//' @param p double Proportion of scouts (0<=p<=1)
 //' @param ls double Scout rate
 //' @param qn double Quality
 //' @param a double alpha value
 //' @export
 // [[Rcpp::export]]
 double scout_dist(double x, double m,
-                  double p, double ls,
-                  double qn, double a)
+                  double ls, double qn, double a)
 {
     double maxpart = (-1+qn-x*a)/a;
     if (maxpart < 0)
@@ -75,7 +74,7 @@ double scout_dist(double x, double m,
         maxpart = 0;
     }
     double result =
-    p * (
+        (
         (exp(((-1 + qn + m * a - x * a) * ls) / a) * a * pow(ls, 2) * maxpart) /
         (exp(m * ls) * a - exp(((-1 + qn) * ls) / a) *
             (a + ls - qn * ls + m * a * ls))
@@ -90,8 +89,7 @@ double scout_dist(double x, double m,
 //' @export
 // [[Rcpp::export]]
 double recruit_dist(double x, double m,
-                    double p, double ln,
-                    double qn, double a)
+                    double ln, double qn, double a)
 {
     double maxpart = -x+((-1+qn)/(qn*a));
     if (maxpart < 0)
@@ -99,7 +97,7 @@ double recruit_dist(double x, double m,
         maxpart = 0;
     }
     double result =
-        (1 - p) * (
+        (
             (2 * exp(-M_PI * pow(x, 2) * ln) * M_PI * x * ln * maxpart) /
             (((exp(-pow(m, 2) * M_PI * ln) * (-1 + qn - m * qn * a)) /
                 (qn * a)) + ((-1 + erf(m * sqrt(M_PI) * sqrt(ln)) +
@@ -116,6 +114,7 @@ double recruit_dist(double x, double m,
 //' Log-likelihood function for scout and recruit superposition
 //'
 //' @param x double* Pointer to array of foraging distances
+//' @param p double Proportion of scouts (0<=p<=1)
 //' @inheritParams scout_dist
 //' @inheritParams recruit_dist
 //' @export
@@ -129,8 +128,8 @@ double loglike_model_all(NumericVector x,
[…]

### 2021-09-10 analysis/fit_models.Rmd
Added a reparameterised model which returns good likelihoods but can't reverse engineer for a fit....code/package quite messy but going to just keep backed up and fix later once we have this working. ATM it is changing to often to make spending loads of time cleaning it up worth while. Its a mess, but its logical to follow and can be cleaned up once its sorted and we no longer need to mess with the guts

@@
 source("analysis/fit_data.R")
+
 ```
+
+## Trial run of a single site using new routine
+
+Using the new routine in this branch (using Nelder-Mead simplex rather than COBYLA) and seperatly optimising p.
+
+```{r}
+# Run model analysis lots and lots of times.
+
+bel <- get_data() %>%
+  filter(site == "BEL")
+
+result_data <- purrr::map_df(
+  seq_len(10),
+  ~ {
+    print(paste("Itteration", .x))
+    result <- run_analysis(bel, n = 10, upper = 0.1, verbose_r = T, verbose = F)
+    return(
+      data.frame(
+        ll = result$fmax,
+        p = result$est[1],
+        ls = result$est[2],
+        ln = result$est[3],
+        qn = result$est[4],
+        a = result$est[5]
+      )
+    )
+  }
+)
+
+best_mod <- result_data %>% filter(ll == max(ll))
+params <- best_mod %>% select(p, ls, ln, qn, a)
+solution <- list(
+  fmax = best_mod$ll,
+  data_name = "all",
+  est = params %>% as.double()
+)
+solution
+
+result_data %>%
+  gather("measurement", "value", -ll) %>%
+  ggplot(aes(x = value, y = ll)) +
+  geom_point() +
+  facet_wrap(~measurement, scale = "free")
+
+```
+
+```{r}
+# convert new parameters to the old ones and show the fit
+# ls = bs*as
+# ln = br*ar^2
+# qn = ar / as
+# alpha = ar - as
+formated_solution <- solution
+formated_solution$est[2] <- solution$est[2] * solution$est[4]
+formated_solution$est[3] <- solution$est[3] * solution$est[5]^2
+formated_solution$est[4] <- solution$est[5] / solution$est[4]
+formated_solution$est[5] <- solution$est[4] - solution$est[5]
+
+# check likelihood
+loglike_model_all(
+  bel$foraging_distance,
+  formated_solution$est[1], formated_solution$est[2],
+  formated_solution$est[3], formated_solution$est[4],
+  formated_solution$est[5]
+)
+
+# show fit of the best model to the data
+make_full_plot(
+  bel$foraging_distance,
+  list("collective" = formated_solution),
+  subplot_coords = c(0.5, 8, -4.5, -2.5)
+)
+
+```
+
+```{r}
+# examine likelihood space of best model
+map_likelihood_space(bel$foraging_distance, params, n = 10000, upper = 0.1)
+
+# sanity check
+loglike_model_all_new(
+  bel$foraging_distance,
+  params[[1]], params[[2]],
+  params[[3]], params[[4]],
+  params[[5]]
+)
+
+loglike_model_all(
+  bel$foraging_distance,
+  params[[1]], params[[2]],
+  params[[3]], 5,
+  params[[5]]
+)
+
+```

### 2021-09-10 R/RcppExports.R

@@ -103,6 +103,39 @@ recruit_dist <- function(x, m, ln, qn, a) {
     .Call('_wagglefit_recruit_dist', PACKAGE = 'wagglefit', x, m, ln, qn, a)
 }
 
+#' New model function for scouts
+#'
+#' @param x double Foraging distance
+#' @param m double Minimum foraging distance
+#' @param bs double scout rate
+#' @param as double scout alpha
+#' @export
+scout_dist_new <- function(x, m, bs, as) {
+    .Call('_wagglefit_scout_dist_new', PACKAGE = 'wagglefit', x, m, bs, as)
+}
+
+#' New model function for recruits
+#'
+#' @param br double Recruit rate
+#' @param ar double Recruit alpha
+#' @inheritParams scout_dist_new
+#' @export
+recruit_dist_new <- function(x, m, br, ar) {
+    .Call('_wagglefit_recruit_dist_new', PACKAGE = 'wagglefit', x, m, br, ar)
+}
+
+#' Log-likelihood function for scout and recruit superposition using
+#' re-parameterised model
+#'
+#' @param x double* Pointer to array of foraging distances
+#' @param p double Proportion of scouts (0<=p<=1)
+#' @inheritParams scout_dist_new
+#' @inheritParams recruit_dist_new
+#' @export
+loglike_model_all_new <- function(x, p, bs, br, as, ar) {
+    .Call('_wagglefit_loglike_model_all_new', PACKAGE = 'wagglefit', x, p, bs, br, as, ar)
+}
+
 #' Log-likelihood function for scout and recruit superposition
 #'
 #' @param x double* Pointer to array of foraging distances
@@ -122,7 +155,7 @@ loglike_model_scout <- function(x, ls, qn, a) {
     .Call('_wagglefit_loglike_model_scout', PACKAGE = 'wagglefit', x, ls, qn, a)
 }
 
-#' Optimise function for fitting a model using NLOPT
+#' Optimise function for fitting the new collective model using NLOPT
 #'
 #' @param x NumericVector Foraging distance
 #' @param params NumericVector parameter estimates to run the model with
@@ -136,5 +169,21 @@ loglike_model_scout <- function(x, ls, qn, a) {
 #' @export
 optimise_model <- function(x, params, lb, ub, verbose = FALSE, xtol = 0, model = 0L) {
[…]

### 2021-09-10 R/fitting.R

@@ -33,10 +33,10 @@ check_upper_bound <- function(upper) {
 #'
 generate_bounds_all <- function(upper) {
   p_bnds <- c(0, 1.0)
-  ls_bnds <- c(1.0e-6, upper)
-  ln_bnds <- c(0, upper)
-  q_bnds <- c(1, upper)
-  a_bnds <- c(0, upper)
+  ls_bnds <- c(1.0e-6, 100)
+  ln_bnds <- c(1.0e-6, 100)
+  q_bnds <- c(1.0e-6, upper)
+  a_bnds <- c(1.0e-6, upper)
   bounds <- rbind(
     p_bnds, ls_bnds,
     ln_bnds, q_bnds,
@@ -93,7 +93,8 @@ generate_bounds_scout <- function(upper) {
 generate_starting_ests_all <- function(distance, bounds, verbose = FALSE) {
   message_verbose(verbose, "  -Generating viable starting estimates")
   res <- Inf
-  while (is.infinite(res) | is.nan(res)) {
+  startest <- c(0, 0, 0, 1, 0)
+  while (is.infinite(res) | is.nan(res) | startest[4] > startest[5]) {
     non_p_bounds <- bounds[2:nrow(bounds), ]
     p_startest <- trunc_normal(
       mean = 0.15, sd = 0.05,
@@ -107,7 +108,7 @@ generate_starting_ests_all <- function(distance, bounds, verbose = FALSE) {
       }
     )
     startest <- c(p_startest, rest_startest)
-    res <- loglike_model_all(
+    res <- loglike_model_all_new(
       distance,
       startest[1],
       startest[2],

### 2021-09-10 R/plotting.R

@@ -104,7 +104,11 @@ make_full_plot <- function(x, model_result_list,
   cdf_data <- map(
     model_result_list,
     ~ {
-      make_ccdf_plot_data(x, .x$est, model = .x$data_name)
+      result <- make_ccdf_plot_data(x, .x$est, model = .x$data_name)
+      if(any((is.nan(result$cumul_ccdf)))) {
+        message("The combination of parameters produces nan values in the ccdf")
+      }
+      return(result)
     }
   )
   df <- map_df(
@@ -112,7 +116,6 @@ make_full_plot <- function(x, model_result_list,
     I,
     .id = "Model"
   )
-
   plt <- make_base_plot(x, TRUE) +
     geom_line(
       data = df,

### 2021-09-10 R/utils.R

@@ -143,3 +143,103 @@ calc_ks_boot <- function(x, param_est, model_type, pvalue = TRUE) {
   ks <- ks.boot(ccdf_data$prob, ccdf_model)
   return(ks$ks.boot.pvalue)
 }
+
+#' Calculate likelihoods for a specified parameter
+#'
+#' @description Calculate likelihoods for a specified parameter to show the
+#' likelihood space for that parameter. This uses the bounds defined for that
+#' parameter and gets the specified number of values to show the likelihood
+#' space.
+#' @param data doubleArray The data to check the parameters against. This should
+#' be the data the parameters were fit to.
+#' @param var characterArray The name of the variable you want to examine the
+#' likelihood space for.
+#' @param p double The proportion of scouts.
+#' @param ls double The scout rate.
+#' @param ln double The recruit rate.
+#' @param q double The quality.
+#' @param a double The alpha value.
+#' @param n integer The number of points to sample, defaults to 1000.
+#' @param upper integer The upper bound
+#' @return tibble The varied parameter and associated likelihood and
+#' log-likelihood scores.
+#' @importFrom tibble tibble
+#' @importFrom purrr map_dbl
+#' @concept utility
+#'
+calc_var_likelihood <- function(
+  data, var, p, ls, ln, q, a, n = 1000, upper = 5) {
+  bnds <- generate_bounds_all(upper)
+  var_bnds <- paste0(var, "_bnds")
+  vals <- seq(bnds[var_bnds, 1], bnds[var_bnds, 2], length.out = n)
+  result <- map_dbl(
+    vals,
+    ~ {
+      if (var == "p") {
+        loglike_model_all_new(data, .x, ls, ln, q, a)
+      } else if (var == "ls") {
+        loglike_model_all_new(data, p, .x, ln, q, a)
+      } else if (var == "ln") {
+        loglike_model_all_new(data, p, ls, .x, q, a)
+      } else if (var == "q") {
+        loglike_model_all_new(data, p, ls, ln, .x, a)
+      } else {
+        loglike_model_all_new(data, p, ls, ln, q, .x)
+      }
+    }
+  )
+  return(tibble( var = vals, loglike = result, likelihood = exp(result)))
+}
+
+#' Calculate likelihoods for all parameters for a given dataset
+#'
+#' @description Calculate likelihoods for all parameters to show their
+#' likelihood space for a given set of data and optimised model parameters. The
[…]

### 2021-09-10 analysis/fit_data.R

@@ -31,12 +31,16 @@ get_data <- function() {
 }
 
 #' Run my analysis
-run_analysis <- function(data, model = "all", n = 5, xtol = 0) {
+run_analysis <- function(
+  data, model = "all", n = 5, upper = 5, xtol = 0,
+  verbose_r = FALSE, verbose = FALSE) {
   best_result <- fit_mutliple(
     data$foraging_distance,
     model = model,
     n = n,
-    verbose_r = FALSE,
+    upper = upper,
+    verbose_r = verbose_r,
+    verbose = verbose,
     xtol = xtol
   )
   best_result$data_name <- model

### 2021-09-10 src/RcppExports.cpp

@@ -136,6 +136,50 @@ BEGIN_RCPP
     return rcpp_result_gen;
 END_RCPP
 }
+// scout_dist_new
+double scout_dist_new(double x, double m, double bs, double as);
+RcppExport SEXP _wagglefit_scout_dist_new(SEXP xSEXP, SEXP mSEXP, SEXP bsSEXP, SEXP asSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< double >::type x(xSEXP);
+    Rcpp::traits::input_parameter< double >::type m(mSEXP);
+    Rcpp::traits::input_parameter< double >::type bs(bsSEXP);
+    Rcpp::traits::input_parameter< double >::type as(asSEXP);
+    rcpp_result_gen = Rcpp::wrap(scout_dist_new(x, m, bs, as));
+    return rcpp_result_gen;
+END_RCPP
+}
+// recruit_dist_new
+double recruit_dist_new(double x, double m, double br, double ar);
+RcppExport SEXP _wagglefit_recruit_dist_new(SEXP xSEXP, SEXP mSEXP, SEXP brSEXP, SEXP arSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< double >::type x(xSEXP);
+    Rcpp::traits::input_parameter< double >::type m(mSEXP);
+    Rcpp::traits::input_parameter< double >::type br(brSEXP);
+    Rcpp::traits::input_parameter< double >::type ar(arSEXP);
+    rcpp_result_gen = Rcpp::wrap(recruit_dist_new(x, m, br, ar));
+    return rcpp_result_gen;
+END_RCPP
+}
+// loglike_model_all_new
+double loglike_model_all_new(NumericVector x, double p, double bs, double br, double as, double ar);
+RcppExport SEXP _wagglefit_loglike_model_all_new(SEXP xSEXP, SEXP pSEXP, SEXP bsSEXP, SEXP brSEXP, SEXP asSEXP, SEXP arSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
[…]

### 2021-09-10 src/models.cpp

@@ -23,6 +23,24 @@ typedef struct {
     double p;
 } data_struct;
 
+typedef struct {
+    NumericVector x;
+    bool verbose;
+    int fcount;
+    double best_est;
+} data_struct_new;
+
+typedef struct {
+    NumericVector x;
+    bool verbose;
+    int fcount;
+    double best_est;
+    double ls;
+    double ln;
+    double qn;
+    double a;
+} data_struct_p;
+
 /*
 ------------------------------- helper functions -------------------------------
 */
@@ -107,9 +125,87 @@ double recruit_dist(double x, double m,
     return result;
 }
 
+//' New model function for scouts
+//'
+//' @param x double Foraging distance
+//' @param m double Minimum foraging distance
+//' @param bs double scout rate
+//' @param as double scout alpha
+//' @export
+// [[Rcpp::export]]
+double scout_dist_new(double x, double m, double bs, double as)
+{
+    double maxpart = 1-as*x;
+    if (maxpart < 0)
+    {
+        maxpart = 0;
+    }
+    double result =
+        as * (
+            (bs * exp(-bs*as*(x-m))*maxpart) /
+            ((1-as*m) - pow(bs, -1)*(1-exp(-bs*(1-as*m))))
+        );
+    return result;
+}
+
+//' New model function for recruits
+//'
+//' @param br double Recruit rate
+//' @param ar double Recruit alpha
+//' @inheritParams scout_dist_new
+//' @export
+// [[Rcpp::export]]
+double recruit_dist_new(double x, double m, double br, double ar)
+{
+    double maxpart = 1-ar*x;
+    if (maxpart < 0)
+    {
+        maxpart = 0;
+    }
+    double result =
+    ar * (
+        (maxpart*2*M_PI*ar*br*x*exp(-M_PI*br*pow(ar*x, 2))) /
+        ((1-ar*m)*exp(-M_PI*br*pow(ar*m, 2)) +
+        ((erf(ar*sqrt(M_PI*br*m)) - erf(sqrt(M_PI*br))) /(2*sqrt(br))))
+    );
+    return result;
+}
+
 /*
 ------------------------------- likelihood functions ---------------------------
 */
+
+//' Log-likelihood function for scout and recruit superposition using
+//' re-parameterised model
+//'
+//' @param x double* Pointer to array of foraging distances
+//' @param p double Proportion of scouts (0<=p<=1)
+//' @inheritParams scout_dist_new
+//' @inheritParams recruit_dist_new
[…]

### 2021-09-21 analysis/fit_models.Rmd
trouble shooting issues

--- 2021-09-02 analysis/fit_models.Rmd
@@
 source("analysis/fit_data.R")
+
 ```
+
+## Trial run of a single site using new routine
+
+Using the new routine in this branch (using Nelder-Mead simplex rather than COBYLA) and seperatly optimising p.
+
+```{r}
+# Run model analysis lots and lots of times.
+use_upper <- 50
+
+bel <- get_data() %>%
+  filter(site == "BEL")
+
+result_data <- purrr::map_df(
+  seq_len(10),
+  ~ {
+    print(paste("Itteration", .x))
+    result <- run_analysis(
+      bel, n = 10, upper = use_upper, verbose_r = T, verbose = F
+    )
+    return(
+      data.frame(
+        ll = result$fmax,
+        p = result$est[1],
+        bs = result$est[2],
+        br = result$est[3],
+        as = result$est[4],
+        ar = result$est[5]
+      )
+    )
+  }
+)
+
+best_mod <- result_data %>% filter(ll == max(ll))
+if (length(best_mod[, 1] > 1)) {
+  best_mod <- best_mod[1, ]
+}
+params <- best_mod %>% select(p, bs, br, as, ar)
+solution <- list(
+  fmax = best_mod$ll,
+  data_name = "all",
+  est = params %>% as.double()
+)
+
+map_likelihood_space(
+  bel$foraging_distance, params, n = 10000, upper = use_upper
+)
+
+solution
+
+```
+
+```{r}
+# show the fit
+
+make_full_plot(
+  bel$foraging_distance,
+  list("collective" = solution),
+  subplot_coords = c(0.5, 8, -4.5, -2.5)
+)
+
+```

### 2021-09-21 R/RcppExports.R

@@ -6,7 +6,7 @@
 #' @param x NumericArray
 #' @export
 alter_in_place <- function(x) {
-    invisible(.Call('_wagglefit_alter_in_place', PACKAGE = 'wagglefit', x))
+  invisible(.Call("_wagglefit_alter_in_place", PACKAGE = "wagglefit", x))
 }
 
 #' Heaviside equation. Replace negatives with 0 and positives with 1.
@@ -14,7 +14,18 @@ alter_in_place <- function(x) {
 #' @param x double the parameter to convert with heaviside
 #' @export
 heaviside <- function(x) {
-    .Call('_wagglefit_heaviside', PACKAGE = 'wagglefit', x)
+  .Call("_wagglefit_heaviside", PACKAGE = "wagglefit", x)
+}
+
+#'
+#' @param x double Foraging distance
+#' @param m double Minimum foraging distance
+#' @param bs double Scout rate
+#' @param as double scout alpha
+#' @return result double the ccdf for that foraging distance (x)
+#' @export
+scout_ccdf_new <- function(x, m, bs, as) {
+  .Call("_wagglefit_scout_ccdf_new", PACKAGE = "wagglefit", x, m, bs, as)
 }
 
 #' Model CCDF function for scouts
@@ -28,7 +39,19 @@ heaviside <- function(x) {
 #' @return result double the ccdf for that foraging distance (x)
 #' @export
 scout_ccdf <- function(x, p, ls, qn, a, m) {
-    .Call('_wagglefit_scout_ccdf', PACKAGE = 'wagglefit', x, p, ls, qn, a, m)
+  .Call("_wagglefit_scout_ccdf", PACKAGE = "wagglefit", x, p, ls, qn, a, m)
+}
+
+#' Model CCDF function for new recruit function
+#'
+#' @param x double Foraging distance
+#' @param m double Minimum foraging distance
+#' @param br double Recruit rate
+#' @param ar double recruit alpha
+#' @return result double the ccdf for that foraging distance (x)
+#' @export
+recruit_ccdf_new <- function(x, m, br, ar) {
+  .Call("_wagglefit_recruit_ccdf_new", PACKAGE = "wagglefit", x, m, br, ar)
 }
 
 #' Model CCDF function for recruits
@@ -43,7 +66,20 @@ scout_ccdf <- function(x, p, ls, qn, a, m) {
[…]

### 2021-09-21 R/fitting.R

@@ -33,8 +33,8 @@ check_upper_bound <- function(upper) {
 #'
 generate_bounds_all <- function(upper) {
   p_bnds <- c(0, 1.0)
-  ls_bnds <- c(1.0e-6, 100)
-  ln_bnds <- c(1.0e-6, 100)
+  ls_bnds <- c(1.0e-6, 10)
+  ln_bnds <- c(1.0e-6, 10)
   q_bnds <- c(1.0e-6, upper)
   a_bnds <- c(1.0e-6, upper)
   bounds <- rbind(

### 2021-09-21 R/utils.R

@@ -167,8 +167,7 @@ calc_ks_boot <- function(x, param_est, model_type, pvalue = TRUE) {
 #' @importFrom purrr map_dbl
 #' @concept utility
 #'
-calc_var_likelihood <- function(
-  data, var, p, ls, ln, q, a, n = 1000, upper = 5) {
+calc_var_likelihood <- function(data, var, p, ls, ln, q, a, n = 1000, upper = 5) {
   bnds <- generate_bounds_all(upper)
   var_bnds <- paste0(var, "_bnds")
   vals <- seq(bnds[var_bnds, 1], bnds[var_bnds, 2], length.out = n)
@@ -188,7 +187,7 @@ calc_var_likelihood <- function(
       }
     }
   )
-  return(tibble( var = vals, loglike = result, likelihood = exp(result)))
+  return(tibble(var = vals, loglike = result, likelihood = exp(result)))
 }
 
 #' Calculate likelihoods for all parameters for a given dataset
@@ -218,7 +217,7 @@ map_likelihood_space <- function(data, params, n = 1000, upper = 5) {
         n = n,
         upper = upper
       ) %>%
-      gather("measurement", "value", -var)
+        gather("measurement", "value", -var)
     }
   )
   names(result) <- names(params)
@@ -226,10 +225,11 @@ map_likelihood_space <- function(data, params, n = 1000, upper = 5) {
     names(result),
     ~ {
       result[[.x]] %>%
+        filter(value > -1e99) %>%
         ggplot(aes(x = var, y = value)) +
-          geom_line() +
-          facet_wrap(~measurement, scale = "free_y") +
-          labs(x = .x)
+        geom_line() +
+        facet_wrap(~measurement, scale = "free_y") +
+        labs(x = .x)
     }
   )
   master_plot <- plot_grid(

### 2021-09-21 src/RcppExports.cpp

@@ -5,11 +5,6 @@
 
 using namespace Rcpp;
 
-#ifdef RCPP_USE_GLOBAL_ROSTREAM
-Rcpp::Rostream<true>&  Rcpp::Rcout = Rcpp::Rcpp_cout_get();
-Rcpp::Rostream<false>& Rcpp::Rcerr = Rcpp::Rcpp_cerr_get();
-#endif
-
 // alter_in_place
 void alter_in_place(NumericVector x);
 RcppExport SEXP _wagglefit_alter_in_place(SEXP xSEXP) {
@@ -31,6 +26,20 @@ BEGIN_RCPP
     return rcpp_result_gen;
 END_RCPP
 }
+// scout_ccdf_new
+double scout_ccdf_new(double x, double m, double bs, double as);
+RcppExport SEXP _wagglefit_scout_ccdf_new(SEXP xSEXP, SEXP mSEXP, SEXP bsSEXP, SEXP asSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< double >::type x(xSEXP);
+    Rcpp::traits::input_parameter< double >::type m(mSEXP);
+    Rcpp::traits::input_parameter< double >::type bs(bsSEXP);
+    Rcpp::traits::input_parameter< double >::type as(asSEXP);
+    rcpp_result_gen = Rcpp::wrap(scout_ccdf_new(x, m, bs, as));
+    return rcpp_result_gen;
+END_RCPP
+}
 // scout_ccdf
 double scout_ccdf(double x, double p, double ls, double qn, double a, double m);
 RcppExport SEXP _wagglefit_scout_ccdf(SEXP xSEXP, SEXP pSEXP, SEXP lsSEXP, SEXP qnSEXP, SEXP aSEXP, SEXP mSEXP) {
@@ -44,6 +53,20 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< double >::type a(aSEXP);
     Rcpp::traits::input_parameter< double >::type m(mSEXP);
     rcpp_result_gen = Rcpp::wrap(scout_ccdf(x, p, ls, qn, a, m));
+    return rcpp_result_gen;
+END_RCPP
+}
+// recruit_ccdf_new
+double recruit_ccdf_new(double x, double m, double br, double ar);
[…]

### 2021-09-21 src/model_ccdfs.cpp

@@ -16,6 +16,25 @@ double heaviside(double x)
   else return 1;
 }
 
+ //' Model CCDF function for new scout function
+//'
+//' @param x double Foraging distance
+//' @param m double Minimum foraging distance
+//' @param bs double Scout rate
+//' @param as double scout alpha
+//' @return result double the ccdf for that foraging distance (x)
+//' @export
+// [[Rcpp::export]]
+double scout_ccdf_new(double x, double m,
+                  double bs, double as)
+{
+  double result;
+  result = ((1-as*x)-pow(bs, -1)*(exp(-as*bs*(x-m)) - exp(-bs*(1-as*m)))) /
+    ((1-as*m)-pow(bs, -1)*(1-exp(-bs*(1-as*m))));
+  return result;
+}
+
+
 //' Model CCDF function for scouts
 //'
 //' @param x double Foraging distance
@@ -39,6 +58,26 @@ double scout_ccdf(double x, double p,
     ) / (
       exp(m*ls)*a-exp(((-1+qn)*ls)/a)*(a+ls-qn*ls+m*a*ls)
     )) + heaviside(m-x);
+  return result;
+}
+
+//' Model CCDF function for new recruit function
+//'
+//' @param x double Foraging distance
+//' @param m double Minimum foraging distance
+//' @param br double Recruit rate
+//' @param ar double recruit alpha
+//' @return result double the ccdf for that foraging distance (x)
+//' @export
+// [[Rcpp::export]]
+double recruit_ccdf_new(double x, double m,
+                    double br, double ar)
+{
+  double result;
+  result = ((1-ar*x)*exp(-M_PI*br*pow(ar*x, 2)) +
+    ((erf(ar*sqrt(M_PI*br)*x)-erf(sqrt(M_PI*br)))/(2*sqrt(br)))) /
+    ((1-ar*m)*exp(-M_PI*br*pow(ar*m, 2)) +
+    ((erf(ar*sqrt(M_PI*br)*m)-erf(sqrt(M_PI*br)))/(2*sqrt(br))));
   return result;
 }
 
@@ -72,6 +111,27 @@ double recruit_ccdf(double x, double p,
     ))
     ) * heaviside((((-1+qn)/(qn*a))-x)*(x-m));
   return result;
+}
+
+//' Model ccdf function for new scout and recruit superposition. Stores results in
+//' given array (y)
+//'
+//' @param x NumericVector foraging distances
+//' @param y NumericVector storage array for the results
[…]

### 2021-09-21 src/models.cpp

@@ -162,12 +162,9 @@ double recruit_dist_new(double x, double m, double br, double ar)
     {
         maxpart = 0;
     }
-    double result =
-    ar * (
-        (maxpart*2*M_PI*ar*br*x*exp(-M_PI*br*pow(ar*x, 2))) /
+    double result = (maxpart*2*M_PI*pow(ar,2)*br*x*exp(-M_PI*br*pow(ar*x, 2))) /
         ((1-ar*m)*exp(-M_PI*br*pow(ar*m, 2)) +
-        ((erf(ar*sqrt(M_PI*br*m)) - erf(sqrt(M_PI*br))) /(2*sqrt(br))))
-    );
+            ((erf(ar*sqrt(M_PI*br)*m)-erf(sqrt(M_PI*br)))/(2*sqrt(br))));
     return result;
 }
 
@@ -196,9 +193,7 @@ double loglike_model_all_new(NumericVector x,
     {
         tmp_ll = (p * scout_dist_new(x[i], m, bs, as) +
             (1 - p) * recruit_dist_new(x[i], m, br, ar));
-        if (as > ar) {
-            ll += -1e99+ar/(as+1);
-        } else if (tmp_ll < 1e-99) {
+        if (tmp_ll < 1e-99) {
             ll += -1e99-ar;
         } else {
             ll += log(tmp_ll);

### 2021-09-22 analysis/fit_models.Rmd
new routine coded up and old removed, each site evaliuated. slight hiccups with the ccdf plotting and need to check with vincent. need to make plots on to the histograms as well as checking out with fits with vincent. unsure on CI's as for any estimate these will be very large, but may be able to make the case of these in the aggregate and get a single CI for many different sites within a given environment.

--- 2021-09-02 analysis/fit_models.Rmd
@@
 source("analysis/fit_data.R")
+
 ```
+
+## Optimising each site
+
+In this document I have fit the collective and individual model to each site.
+So al the sites have reasonable likelihood values, however in some cases the fits are simply wrong (i.e.
+they show the probability of sampling a value greater than on equal to $x$ as increasing in the tail).
+I need to trouble shoot these with Vincent.
+I also need to get some advice on how to plot the model fits to the histogram of data.
+
+Would be a good idea to knit this as a single document / run in R studio on the laptop.
+
+### BEL
+
+Provides a good fit on the data. All parameters look central and nicely covered.
+
+```{r}
+
+target_site <- "BEL"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 0.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 10, -5.5, -2)
+)
+
+# bring results together
+bel_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+bel_model_fits
+
+```
+
+### BFI
+
+All parameters look central in the likelihood space and a nice fit is returned.
+
+```{r}
+
+target_site <- "BFI"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 0.6)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 4.5, -8, -2.5)
+)
+
+# bring results together
+bfi_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+bfi_model_fits
+
+```
+
+### BLO
+
+CCDF plot is not right. Cant go back on its self which indicates there is probably some round off error somewhere throwing things off.
+
+The individual fit isnt great and the Bs paramater is increasing up to the boundry, indicating it can only really take on a straight line / exponential fit.
+The parameters for the collective model are fairly central which raises questions as to why the fit is misbehaved in the tail.
+
+```{r}
+
+target_site <- "BLO"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-10, 5)
+br_bnds <- c(1.0e-10, 5)
+as_bnds <- c(1.0e-5, 5)
+ar_bnds <- c(1.0e-10, 1.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 105)
+as_bnds <- c(1.0e-12, 0.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 12, -8., -2.5)
+)
+
+# bring results together
+blo_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+blo_model_fits
+
+```
+
+### BUR
+
+Bs and Br go in oposite directions. E.g. Bs approaches 0 whilst Br approaches an every higher number.
+
+Fit looks good.
+
+```{r}
+
+target_site <- "BUR"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 500)
+br_bnds <- c(1.0e-6, 500)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 0.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 5.5, -6.5, -2.5)
+)
+
+# bring results together
+bur_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+bur_model_fits
+
+```
+
+### CAD
+
+The only graph I can get to make sense is one where the collective reduces to the individual. The others show the ccdf going off.
+
+```{r}
+
+target_site <- "CAD"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 100)
+br_bnds <- c(1.0e-6, 100)
+as_bnds <- c(1.0e-12, 10)
+ar_bnds <- c(1.0e-12, 10)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 2)
+as_bnds <- c(1.0e-12, 0.8)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 2.5, -5, -1.5)
+)
+
+# bring results together
+cad_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+cad_model_fits
+
+```
+
+### GIL
+
+Again, the likelihood scores look good but the model fits are doing crazy things at the tail!
+
+```{r}
+
+target_site <- "GIL"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 0.8)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 2.7, -8., -2)
+)
+
+# bring results together
+gil_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+gil_model_fits
+
+```
+
+### HER
+
+Also has a weird CCDF plot
+
+```{r}
+
+target_site <- "HER"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 100)
+br_bnds <- c(1.0e-6, 100)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 2)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 5, -6.5, -1.5)
+)
+
+# bring results together
+her_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+her_model_fits
+
+```
+
+### HHS
+
+Shit ccdf.
+
+```{r}
+
+target_site <- "HHS"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 50)
+br_bnds <- c(1.0e-6, 50)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 6, -6.5, -2.5)
+)
+
+# bring results together
+hhs_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+hhs_model_fits
+
+```
+
+### HOR
+
+Agin weird behaviour in the tail.
+Seems to be something to do with the Bs parameter approaching the lower bound.
+If the lower bound is made closer to zero (just 1e-7) the plot goes mad and the log-likelihood score becomes very large, indicating a round off error or something stupid.
+
+```{r}
+
+target_site <- "HOR"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 4, -9, -2.5)
+)
+
+# bring results together
+hor_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+hor_model_fits
+
+```
+
+### MAK
+
+Same problem as with the others.
+
+```{r}
+
+target_site <- "MAK"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 5, -6.5, -2.5)
+)
+
+# bring results together
+mak_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+mak_model_fits
+
+```
+
+### MEL
+
+Slight misbehaviour in the tail creaping up.
+
+```{r}
+
+target_site <- "MEL"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 3.5, -7, -2.5)
+)
+
+# bring results together
+mel_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+mel_model_fits
+
+```
+
+### MPA
+
+Same problem with the tail.
+
+```{r}
+
+target_site <- "MPA"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 4.5, -7.2, -2.5)
+)
+
+# bring results together
+mpa_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+mpa_model_fits
+
+```
+
+### ROT
+
+Same problem with the tail.
+
+```{r}
+
+target_site <- "ROT"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 3., -5.5, -2.5)
+)
+
+# bring results together
+rot_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+rot_model_fits
+
+```
+
+### SAU
+
+No problems here!
+
+```{r}
+
+target_site <- "SAU"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 3.5, -6.5, -2.5)
+)
+
+# bring results together
+sau_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+sau_model_fits
+
+```
+
+### SOM
+
+Same problem with the tail.
+
+```{r}
+
+target_site <- "SOM"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site) %>%
+  filter(foraging_distance < 9) # remove outlier distance
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 3., -5.5, -2.5)
+)
+
+# bring results together
+som_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+som_model_fits
+
+```
+
+### SRA
+
+Same problem with the tail.
+
+```{r}
+
+target_site <- "SRA"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 4., -7.5, -2.5)
+)
+
+# bring results together
+sra_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+sra_model_fits
+
+```
+
+### STU
+
+Same problem with the tail.
+Increasing the bounds for as and ar seems to coerce it towards the better visual fit, despite it not changing what they optimise at.
+Weird thing is the model is getting close in the region of p but not actually to it.
+
+```{r}
+
+target_site <- "STU"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12,10)
+ar_bnds <- c(1.0e-12, 10)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 5.5, -6.5, -1.5)
+)
+
+# bring results together
+stu_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+stu_model_fits
+
+```
+
+### SWP
+
+Same problem with the tail. Weird P behaviour when we move some of the other parameters around. Looks like it can get a fit that looks very similar with p = 0
+
+```{r}
+
+target_site <- "SWP"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 15)
+br_bnds <- c(1.0e-6, 15)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 2.5, -6, -2.)
+)
+
+# bring results together
+swp_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+swp_model_fits
+
+```
+
+### YAL
+
+Same problem with the tail.
+
+```{r}
+
+target_site <- "YAL"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 5., -8, -2.5)
+)
+
+# bring results together
+yal_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+yal_model_fits
+
+```
+
+### ZSL
+
+Same problem with the tail.
+
+```{r}
+
+target_site <- "ZSL"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 50)
+as_bnds <- c(1.0e-12, 1.3)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 2.4, -8, -2)
+)
+
+# bring results together
+zsl_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  )
+)
+
+zsl_model_fits
+
+```

### 2021-09-22 R/RcppExports.R

@@ -6,17 +6,10 @@
 #' @param x NumericArray
 #' @export
 alter_in_place <- function(x) {
-  invisible(.Call("_wagglefit_alter_in_place", PACKAGE = "wagglefit", x))
+    invisible(.Call('_wagglefit_alter_in_place', PACKAGE = 'wagglefit', x))
 }
 
-#' Heaviside equation. Replace negatives with 0 and positives with 1.
-#'
-#' @param x double the parameter to convert with heaviside
-#' @export
-heaviside <- function(x) {
-  .Call("_wagglefit_heaviside", PACKAGE = "wagglefit", x)
-}
-
+#' Model CCDF for scout distribution
 #'
 #' @param x double Foraging distance
 #' @param m double Minimum foraging distance
@@ -24,25 +17,11 @@ heaviside <- function(x) {
 #' @param as double scout alpha
 #' @return result double the ccdf for that foraging distance (x)
 #' @export
-scout_ccdf_new <- function(x, m, bs, as) {
-  .Call("_wagglefit_scout_ccdf_new", PACKAGE = "wagglefit", x, m, bs, as)
+scout_ccdf <- function(x, m, bs, as) {
+    .Call('_wagglefit_scout_ccdf', PACKAGE = 'wagglefit', x, m, bs, as)
 }
 
-#' Model CCDF function for scouts
-#'
-#' @param x double Foraging distance
-#' @param p double Proportion of scouts (0<=p<=1)
-#' @param ls double Scout rate
-#' @param qn double Quality
-#' @param a double alpha value
-#' @param m double Minimum foraging distance
-#' @return result double the ccdf for that foraging distance (x)
-#' @export
-scout_ccdf <- function(x, p, ls, qn, a, m) {
-  .Call("_wagglefit_scout_ccdf", PACKAGE = "wagglefit", x, p, ls, qn, a, m)
-}
-
-#' Model CCDF function for new recruit function
+#' Model CCDF for recruit distribution
 #'
 #' @param x double Foraging distance
 #' @param m double Minimum foraging distance
@@ -50,36 +29,8 @@ scout_ccdf <- function(x, p, ls, qn, a, m) {
 #' @param ar double recruit alpha
 #' @return result double the ccdf for that foraging distance (x)
 #' @export
-recruit_ccdf_new <- function(x, m, br, ar) {
-  .Call("_wagglefit_recruit_ccdf_new", PACKAGE = "wagglefit", x, m, br, ar)
-}
-
-#' Model CCDF function for recruits
[…]

### 2021-09-22 R/fitting.R

@@ -33,14 +33,14 @@ check_upper_bound <- function(upper) {
 #'
 generate_bounds_all <- function(upper) {
   p_bnds <- c(0, 1.0)
-  ls_bnds <- c(1.0e-6, 10)
-  ln_bnds <- c(1.0e-6, 10)
-  q_bnds <- c(1.0e-6, upper)
-  a_bnds <- c(1.0e-6, upper)
+  bs_bnds <- c(1.0e-6, 10)
+  br_bnds <- c(1.0e-6, 10)
+  as_bnds <- c(1.0e-12, upper)
+  ar_bnds <- c(1.0e-12, upper)
   bounds <- rbind(
-    p_bnds, ls_bnds,
-    ln_bnds, q_bnds,
-    a_bnds
+    p_bnds, bs_bnds,
+    br_bnds, as_bnds,
+    ar_bnds
   )
   return(bounds)
 }
@@ -58,13 +58,11 @@ generate_bounds_all <- function(upper) {
 #'
 generate_bounds_scout <- function(upper) {
   upper <- as.double(upper)
-  ls_bnds <- c(1.0e-6, upper)
-  q_bnds <- c(1, upper)
-  a_bnds <- c(0, upper)
+  bs_bnds <- c(1.0e-6, upper)
+  as_bnds <- c(1.0e-6, upper)
   bounds <- rbind(
-    ls_bnds,
-    q_bnds,
-    a_bnds
+    bs_bnds,
+    as_bnds
   )
   return(bounds)
 }
@@ -108,7 +106,7 @@ generate_starting_ests_all <- function(distance, bounds, verbose = FALSE) {
       }
     )
     startest <- c(p_startest, rest_startest)
-    res <- loglike_model_all_new(
+    res <- loglike_model_collective(
       distance,
       startest[1],
       startest[2],
@@ -161,11 +159,10 @@ generate_starting_ests_scout <- function(distance, bounds, verbose = FALSE) {
         trunc_normal(mean = (.x + .y) / 2, sd = 1, lwr = .x, upr = .y)
       }
     )
-    res <- loglike_model_scout(
+    res <- loglike_model_individual(
       distance,
       startest[1],
-      startest[2],
-      startest[3]
+      startest[2]
     )
   }
   message_verbose(
@@ -188,8 +185,10 @@ generate_starting_ests_scout <- function(distance, bounds, verbose = FALSE) {
 #'
 #' @description Fits specified model to the data given using MLE.
 #' @param distance doubleArray The distance decoded from the waggle dance.
[…]

### 2021-09-22 R/plotting.R

@@ -46,6 +46,27 @@ make_ccdf_plot_data <- function(x, param_est, model, npoints = 100) {
   return(cumul)
 }
 
+#' Creates data for histogram plot model fit
+#'
+#' @description Using the optimised paramater estimates, creates the data for
+#' the pdf/histogram plots
+#' @param x doubleArray The foraging distance to plot
+#' @param param_est doubleArray The paramater estimates of the optimisation
+#' @param model characterArray The name of the model used
+#' @param npoints integer The number of points to plot the predicted data for.
+#' Defaults to 100.
+#' @return tibble of model data and cumulative probability
+#' @importFrom tibble tibble
+#' @export
+#'
+make_pdf_plot_data <- function(x, param_est, model, npoints = 100) {
+  model_n <- model_number_from_model(model)
+  x_seq <- seq(min(x), max(x), length.out = npoints)
+  pdf <- model_pdf(x_seq, param_est, model_n)
+  cumul <- tibble(x_seq, pdf)
+  return(cumul)
+}
+
 #' Creates ccdf plot of data
 #'
 #' @description Creates a ccdf plot for the data provided
@@ -111,8 +132,23 @@ make_full_plot <- function(x, model_result_list,
       return(result)
     }
   )
+  pdf_data <- map(
+    model_result_list,
+    ~ {
+      result <- make_pdf_plot_data(x, .x$est, model = .x$data_name)
+      if(any((is.nan(result$cumul_ccdf)))) {
+        message("The combination of parameters produces nan values in the ccdf")
+      }
+      return(result)
+    }
+  )
   df <- map_df(
     cdf_data,
+    I,
+    .id = "Model"
+  )
+  df_histo <- map_df(
+    pdf_data,
     I,
     .id = "Model"
   )
@@ -132,6 +168,14 @@ make_full_plot <- function(x, model_result_list,
       bins = 100,
       binwidth = (max(x) - min(x)) / 20,
       col = "white"
+    ) +
+    geom_line(
+      data = df_histo,
+      aes(
+        x = .data$x_seq,
+        y = .data$pdf,
+        colour = .data$Model
+      )
     ) +
     theme(
       axis.title.x = element_blank(),

### 2021-09-22 R/utils.R

@@ -6,8 +6,8 @@
 #' @concept utility
 model_number_from_model <- function(model) {
   model_list <- list(
-    "all" = 0,
-    "scout" = 1
+    "collective" = 0,
+    "individual" = 1
   )
   if (!(model %in% names(model_list))) {
     stop("Model name not found")
@@ -144,7 +144,7 @@ calc_ks_boot <- function(x, param_est, model_type, pvalue = TRUE) {
   return(ks$ks.boot.pvalue)
 }
 
-#' Calculate likelihoods for a specified parameter
+#' Calculate likelihoods for a specified parameter for the collective model
 #'
 #' @description Calculate likelihoods for a specified parameter to show the
 #' likelihood space for that parameter. This uses the bounds defined for that
@@ -155,10 +155,10 @@ calc_ks_boot <- function(x, param_est, model_type, pvalue = TRUE) {
 #' @param var characterArray The name of the variable you want to examine the
 #' likelihood space for.
 #' @param p double The proportion of scouts.
-#' @param ls double The scout rate.
-#' @param ln double The recruit rate.
-#' @param q double The quality.
-#' @param a double The alpha value.
+#' @param bs double The scout rate.
+#' @param as double The scout alpha.
+#' @param br double The recruit rate.
+#' @param ar double The recruit alpha.
 #' @param n integer The number of points to sample, defaults to 1000.
 #' @param upper integer The upper bound
 #' @return tibble The varied parameter and associated likelihood and
@@ -167,24 +167,79 @@ calc_ks_boot <- function(x, param_est, model_type, pvalue = TRUE) {
 #' @importFrom purrr map_dbl
 #' @concept utility
 #'
-calc_var_likelihood <- function(data, var, p, ls, ln, q, a, n = 1000, upper = 5) {
-  bnds <- generate_bounds_all(upper)
+calc_var_likelihood_collective <- function(
+  data, var, params, n = 1000, upper = 5, bounds = NULL
+  ) {
+  p <- params[[1]]
+  bs <- params[[2]]
+  br <- params[[3]]
+  as <- params[[4]]
+  ar <- params[[5]]
+
+  if (is.null(bounds)) {
+    bnds <- generate_bounds_all(upper)
+  } else {
+    bnds <- bounds
+  }
   var_bnds <- paste0(var, "_bnds")
   vals <- seq(bnds[var_bnds, 1], bnds[var_bnds, 2], length.out = n)
   result <- map_dbl(
     vals,
     ~ {
-      if (var == "p") {
[…]

### 2021-09-22 analysis/fit_data.R

@@ -32,13 +32,14 @@ get_data <- function() {
 
 #' Run my analysis
 run_analysis <- function(
-  data, model = "all", n = 5, upper = 5, xtol = 0,
+  data, model = "all", n = 5, upper = 5, bounds = NULL, xtol = 0,
   verbose_r = FALSE, verbose = FALSE) {
   best_result <- fit_mutliple(
     data$foraging_distance,
     model = model,
     n = n,
     upper = upper,
+    bounds = bounds,
     verbose_r = verbose_r,
     verbose = verbose,
     xtol = xtol
@@ -47,27 +48,93 @@ run_analysis <- function(
   return(best_result)
 }
 
-## run it
-run_all_models <- function() {
-  n <- 5
-  data <- get_data()
-  for (i in unique(data$site)) {
-    print(paste("Working on", i))
-    subdat <- data %>% dplyr::filter(site == i)
-    result_recruit <- run_analysis(
-      subdat,
-      n = n, model = "recruit", xtol = 0
-    )
-    result_all <- run_analysis(subdat, n = n, model = "all", xtol = 0)
-    result_scout <- run_analysis(subdat, n = n, model = "scout", xtol = 0)
-    result_list <- list(
-      "all" = result_all,
-      "scout" = result_scout,
-      "recruit" = result_recruit
-    )
-    saveRDS(object = result_list, file = paste0("analysis/results/", i, ".rds"))
-    print(paste(i, "Completed"))
+#' Run on specific site k times and output likelihood space and plots
+fit_collective_model_to_data <- function(data, bounds, k=10, n = 10,
+                                        verbose_r = FALSE, verbose = FALSE) {
+  # run optimisation k times and return best one
+  result_data <- purrr::map_df(
+    seq_len(k),
+    ~ {
+      print(paste("Itteration", .x))
+      result <- run_analysis(
+        data, model = "collective", n = n, upper = 10, bounds = bounds,
+        verbose_r = verbose_r, verbose = verbose
+      )
+      return(
+        data.frame(
+          ll = result$fmax,
+          p = result$est[1],
+          bs = result$est[2],
+          br = result$est[3],
+          as = result$est[4],
+          ar = result$est[5]
+        )
+      )
+    }
+  )
+
+  # get the best model params
+  best_mod <- result_data %>% filter(ll == max(ll))
+  if (length(best_mod[, 1] > 1)) {
+    best_mod <- best_mod[1, ]
   }
+  params <- best_mod %>% select(p, bs, br, as, ar)
[…]

### 2021-09-22 src/RcppExports.cpp

@@ -15,20 +15,9 @@ BEGIN_RCPP
     return R_NilValue;
 END_RCPP
 }
-// heaviside
-double heaviside(double x);
-RcppExport SEXP _wagglefit_heaviside(SEXP xSEXP) {
-BEGIN_RCPP
-    Rcpp::RObject rcpp_result_gen;
-    Rcpp::RNGScope rcpp_rngScope_gen;
-    Rcpp::traits::input_parameter< double >::type x(xSEXP);
-    rcpp_result_gen = Rcpp::wrap(heaviside(x));
-    return rcpp_result_gen;
-END_RCPP
-}
-// scout_ccdf_new
-double scout_ccdf_new(double x, double m, double bs, double as);
-RcppExport SEXP _wagglefit_scout_ccdf_new(SEXP xSEXP, SEXP mSEXP, SEXP bsSEXP, SEXP asSEXP) {
+// scout_ccdf
+double scout_ccdf(double x, double m, double bs, double as);
+RcppExport SEXP _wagglefit_scout_ccdf(SEXP xSEXP, SEXP mSEXP, SEXP bsSEXP, SEXP asSEXP) {
 BEGIN_RCPP
     Rcpp::RObject rcpp_result_gen;
     Rcpp::RNGScope rcpp_rngScope_gen;
@@ -36,59 +25,27 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< double >::type m(mSEXP);
     Rcpp::traits::input_parameter< double >::type bs(bsSEXP);
     Rcpp::traits::input_parameter< double >::type as(asSEXP);
-    rcpp_result_gen = Rcpp::wrap(scout_ccdf_new(x, m, bs, as));
-    return rcpp_result_gen;
-END_RCPP
-}
-// scout_ccdf
-double scout_ccdf(double x, double p, double ls, double qn, double a, double m);
-RcppExport SEXP _wagglefit_scout_ccdf(SEXP xSEXP, SEXP pSEXP, SEXP lsSEXP, SEXP qnSEXP, SEXP aSEXP, SEXP mSEXP) {
-BEGIN_RCPP
-    Rcpp::RObject rcpp_result_gen;
-    Rcpp::RNGScope rcpp_rngScope_gen;
-    Rcpp::traits::input_parameter< double >::type x(xSEXP);
-    Rcpp::traits::input_parameter< double >::type p(pSEXP);
[…]

### 2021-09-22 src/model_ccdfs.cpp

@@ -3,20 +3,7 @@
 #include <Rcpp.h>
 using namespace Rcpp;
 
-
-
-//' Heaviside equation. Replace negatives with 0 and positives with 1.
-//'
-//' @param x double the parameter to convert with heaviside
-//' @export
-// [[Rcpp::export]]
-double heaviside(double x)
-{
-  if (x < 0) return 0;
-  else return 1;
-}
-
- //' Model CCDF function for new scout function
+//' Model CCDF for scout distribution
 //'
 //' @param x double Foraging distance
 //' @param m double Minimum foraging distance
@@ -25,43 +12,16 @@ double heaviside(double x)
 //' @return result double the ccdf for that foraging distance (x)
 //' @export
 // [[Rcpp::export]]
-double scout_ccdf_new(double x, double m,
+double scout_ccdf(double x, double m,
                   double bs, double as)
 {
   double result;
-  result = ((1-as*x)-pow(bs, -1)*(exp(-as*bs*(x-m)) - exp(-bs*(1-as*m)))) /
-    ((1-as*m)-pow(bs, -1)*(1-exp(-bs*(1-as*m))));
+  result = ((1-as*x)*exp(-as*bs*x)-pow(bs, -1)*(exp(-as*bs*x)-exp(-bs))) /
+    ((1-as*m)*exp(-as*bs*m)-pow(bs, -1)*(exp(-as*bs*m)-exp(-bs)));
   return result;
 }
 
-
-//' Model CCDF function for scouts
-//'
-//' @param x double Foraging distance
-//' @param p double Proportion of scouts (0<=p<=1)
-//' @param ls double Scout rate
-//' @param qn double Quality
-//' @param a double alpha value
-//' @param m double Minimum foraging distance
-//' @return result double the ccdf for that foraging distance (x)
-//' @export
-// [[Rcpp::export]]
-double scout_ccdf(double x, double p,
-                  double ls, double qn,
-                  double a, double m)
-{
-  double result;
-  result = p*(
-    (exp((m-x-(1/a))*ls)*(
-      exp((x+(1/a))*ls)*a-exp((qn*ls)/a)*(a+ls-qn*ls+x*a*ls)
-    )*heaviside((-m+x)*(-x+((-1+qn)/a)))
-    ) / (
-      exp(m*ls)*a-exp(((-1+qn)*ls)/a)*(a+ls-qn*ls+m*a*ls)
-    )) + heaviside(m-x);
-  return result;
-}
-
[…]

### 2021-09-22 src/models.cpp

@@ -20,26 +20,8 @@ typedef struct {
     bool verbose;
     int fcount;
     double best_est;
-    double p;
 } data_struct;
 
-typedef struct {
-    NumericVector x;
-    bool verbose;
-    int fcount;
-    double best_est;
-} data_struct_new;
-
-typedef struct {
-    NumericVector x;
-    bool verbose;
-    int fcount;
-    double best_est;
-    double ls;
-    double ln;
-    double qn;
-    double a;
-} data_struct_p;
 
 /*
 ------------------------------- helper functions -------------------------------
@@ -78,62 +60,11 @@ void check_optimise_result(int optimise_result, bool verbose = true) {
 //'
 //' @param x double Foraging distance
 //' @param m double Minimum foraging distance
-//' @param ls double Scout rate
-//' @param qn double Quality
-//' @param a double alpha value
-//' @export
-// [[Rcpp::export]]
-double scout_dist(double x, double m,
-                  double ls, double qn, double a)
-{
-    double maxpart = (-1+qn-x*a)/a;
-    if (maxpart < 0)
-    {
-        maxpart = 0;
-    }
-    double result =
-        (
-        (exp(((-1 + qn + m * a - x * a) * ls) / a) * a * pow(ls, 2) * maxpart) /
-        (exp(m * ls) * a - exp(((-1 + qn) * ls) / a) *
-            (a + ls - qn * ls + m * a * ls))
-    );
-    return result;
-}
-
-//' Model function for recruits
-//'
-//' @param ln double Recruit rate
-//' @inheritParams scout_dist
-//' @export
-// [[Rcpp::export]]
-double recruit_dist(double x, double m,
-                    double ln, double qn, double a)
-{
-    double maxpart = -x+((-1+qn)/(qn*a));
-    if (maxpart < 0)
-    {
-        maxpart = 0;
-    }
-    double result =
-        (
-            (2 * exp(-M_PI * pow(x, 2) * ln) * M_PI * x * ln * maxpart) /
-            (((exp(-pow(m, 2) * M_PI * ln) * (-1 + qn - m * qn * a)) /
-                (qn * a)) + ((-1 + erf(m * sqrt(M_PI) * sqrt(ln)) +
-                erfc((sqrt(M_PI) * (-1 + qn) * sqrt(ln)) /
-                (qn * a))) / (2 * sqrt(ln))))
-    );
-    return result;
-}
-
-//' New model function for scouts
-//'
-//' @param x double Foraging distance
[…]

### 2021-09-24 R/plotting.R
fixed plotting problem in meeting with Vincent

@@ -62,7 +62,7 @@ make_ccdf_plot_data <- function(x, param_est, model, npoints = 100) {
 make_pdf_plot_data <- function(x, param_est, model, npoints = 100) {
   model_n <- model_number_from_model(model)
   x_seq <- seq(min(x), max(x), length.out = npoints)
-  pdf <- model_pdf(x_seq, param_est, model_n)
+  pdf <- model_pdf(x_seq, param_est, model_n) * length(x) * ((max(x) - min(x)) / 20)
   cumul <- tibble(x_seq, pdf)
   return(cumul)
 }

### 2021-09-24 src/model_ccdfs.cpp

@@ -16,8 +16,12 @@ double scout_ccdf(double x, double m,
                   double bs, double as)
 {
   double result;
-  result = ((1-as*x)*exp(-as*bs*x)-pow(bs, -1)*(exp(-as*bs*x)-exp(-bs))) /
+  if (x > 1/as) {
+    result = 0;
+  } else {
+    result = ((1-as*x)*exp(-as*bs*x)-pow(bs, -1)*(exp(-as*bs*x)-exp(-bs))) /
     ((1-as*m)*exp(-as*bs*m)-pow(bs, -1)*(exp(-as*bs*m)-exp(-bs)));
+  }
   return result;
 }
 
@@ -34,10 +38,14 @@ double recruit_ccdf(double x, double m,
                     double br, double ar)
 {
   double result;
-  result = ((1-ar*x)*exp(-M_PI*br*pow(ar*x, 2)) +
-    ((erf(ar*sqrt(M_PI*br)*x)-erf(sqrt(M_PI*br)))/(2*sqrt(br)))) /
-    ((1-ar*m)*exp(-M_PI*br*pow(ar*m, 2)) +
-    ((erf(ar*sqrt(M_PI*br)*m)-erf(sqrt(M_PI*br)))/(2*sqrt(br))));
+  if (x > 1/ar) {
+    result = 0;
+  } else {
+    result = ((1-ar*x)*exp(-M_PI*br*pow(ar*x, 2)) +
+      ((erf(ar*sqrt(M_PI*br)*x)-erf(sqrt(M_PI*br)))/(2*sqrt(br)))) /
+      ((1-ar*m)*exp(-M_PI*br*pow(ar*m, 2)) +
+      ((erf(ar*sqrt(M_PI*br)*m)-erf(sqrt(M_PI*br)))/(2*sqrt(br))));
+  }
   return result;
 }
 

### 2021-09-30 analysis/fit_models.Rmd
fixed plotting problem in meeting with Vincent
updated figures with working routine. nothing eally changed results wise but we have more confidence in the proportions of scouts coming back (i think)

--- 2021-09-21 analysis/fit_models.Rmd
@@
 
-```
-
-## Trial run of a single site using new routine
-
-Using the new routine in this branch (using Nelder-Mead simplex rather than COBYLA) and seperatly optimising p.
-
-```{r}
-# Run model analysis lots and lots of times.
-use_upper <- 50
-
-bel <- get_data() %>%
-  filter(site == "BEL")
-
-result_data <- purrr::map_df(
-  seq_len(10),
-  ~ {
-    print(paste("Itteration", .x))
-    result <- run_analysis(
-      bel, n = 10, upper = use_upper, verbose_r = T, verbose = F
-    )
-    return(
-      data.frame(
-        ll = result$fmax,
-        p = result$est[1],
-        bs = result$est[2],
-        br = result$est[3],
-        as = result$est[4],
-        ar = result$est[5]
-      )
-    )
-  }
-)
-
-best_mod <- result_data %>% filter(ll == max(ll))
-if (length(best_mod[, 1] > 1)) {
-  best_mod <- best_mod[1, ]
-}
-params <- best_mod %>% select(p, bs, br, as, ar)
-solution <- list(
-  fmax = best_mod$ll,
-  data_name = "all",
-  est = params %>% as.double()
-)
-
-map_likelihood_space(
-  bel$foraging_distance, params, n = 10000, upper = use_upper
-)
-
-solution
+all_data <- as.list(rep(0, 20))
 
@@
 
+## Optimising each site
+
+In this document I have fit the collective and individual model to each site.
+So al the sites have reasonable likelihood values, however in some cases the fits are simply wrong (i.e.
+they show the probability of sampling a value greater than on equal to $x$ as increasing in the tail).
+I need to trouble shoot these with Vincent.
+I also need to get some advice on how to plot the model fits to the histogram of data.
+
+Would be a good idea to knit this as a single document / run in R studio on the laptop.
+
+### BEL
+
+Provides a good fit on the data. All parameters look central and nicely covered.
+
 ```{r}
-# show the fit
 
+target_site <- "BEL"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 0.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
 make_full_plot(
-  bel$foraging_distance,
-  list("collective" = solution),
-  subplot_coords = c(0.5, 8, -4.5, -2.5)
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 10, -5.5, -2)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+bel_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[1]] <- bel_model_fits
+
+bel_model_fits
+
+```
+
+### BFI
+
+All parameters look central in the likelihood space and a nice fit is returned.
+
+```{r}
+
+target_site <- "BFI"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 0.6)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 4.5, -8, -2.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+bfi_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[2]] <- bfi_model_fits
+
+bfi_model_fits
+
+```
+
+### BLO
+
+The individual fit isnt great and the Bs paramater is increasing up to the boundry, indicating it can only really take on a straight line / exponential fit.
+The parameters for the collective model are fairly central in the likelihood space, the fit looks very good.
+
+```{r}
+
+target_site <- "BLO"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-10, 5)
+br_bnds <- c(1.0e-10, 5)
+as_bnds <- c(1.0e-5, 5)
+ar_bnds <- c(1.0e-10, 1.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 5505)
+as_bnds <- c(1.0e-12, 0.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 12, -8., -2.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+blo_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[3]] <- blo_model_fits
+
+blo_model_fits
+
+```
+
+### BUR
+
+Bs and Br go in oposite directions. E.g. Bs approaches 0 whilst Br approaches an every higher number.
+
+Fit looks good.
+
+```{r}
+
+target_site <- "BUR"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 500)
+br_bnds <- c(1.0e-6, 500)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 0.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 5.5, -6.5, -2.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+bur_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[4]] <- bur_model_fits
+
+bur_model_fits
+
+```
+
+### CAD
+
+The collective model roughly follows the individual model but is able to acheive a slightly improved fit to the shoulder, hence it provides a higher likelihood score.
+The AIC indicates this is overfitting, suggesting the individual model provides a more parsimonious explanation.
+
+$p$ does not approach 1 as one might expect, indicating there is overfitting and so a comparison with an individual model is required.
+
+```{r}
+
+target_site <- "CAD"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 10)
+ar_bnds <- c(1.0e-12, 10)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 2)
+as_bnds <- c(1.0e-12, 0.8)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 2.5, -5, -1.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+cad_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[5]] <- cad_model_fits
+
+cad_model_fits
+
+```
+
+### GIL
+
+The collective model provides the most parsimonious and best fit.
+
+```{r}
+
+target_site <- "GIL"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 0.8)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 2.7, -8., -2)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+gil_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[6]] <- gil_model_fits
+
+gil_model_fits
+
+```
+
+### HER
+
+The collective model provides the best fit to the data, but the proportion of scouts is high.
+
+```{r}
+
+target_site <- "HER"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 100)
+br_bnds <- c(1.0e-6, 100)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 2)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 5, -6.5, -1.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+her_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[7]] <- her_model_fits
+
+her_model_fits
+
+```
+
+### HHS
+
+Collective model provides the best fit but falls under the tail and shoulder. The individual model strugles to find a good fit, probably due to the tail.
+
+```{r}
+
+target_site <- "HHS"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 50)
+br_bnds <- c(1.0e-6, 50)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 6, -6.5, -2.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+hhs_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[8]] <- hhs_model_fits
+
+hhs_model_fits
+
+```
+
+### HOR
+
+Collective model provides a very good fit, whilst the individual model fails to find much traction.
+The proportion of scouts goes very low (~3%) suggesting the majority of the colony are following a small number of scouting individuals.
+
+```{r}
+
+target_site <- "HOR"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 4, -9, -2.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+hor_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[9]] <- hor_model_fits
+
+hor_model_fits
+
+```
+
+### MAK
+
+Again the collective model provides a good fit to the data, however, the individual model fits poorly, reducing to an exponential.
+
+```{r}
+
+target_site <- "MAK"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 20)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 5, -6.5, -2.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+mak_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[10]] <- mak_model_fits
+
+mak_model_fits
+
+```
+
+### MEL
+
+Collective model provides the best fit, however it misses a large section of the shoulder for the tail.
+The $bs$ parameter wants to go to zero, however when let go bellow 1e-6 the behaviour becomes very erratic and the fit deteriorates.
+
+```{r}
+
+target_site <- "MEL"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 3.5, -7, -2.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+mel_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[11]] <- mel_model_fits
+
+mel_model_fits
+
+```
+
+### MPA
+
+Collective provides the best fit.
+
+```{r}
+
+target_site <- "MPA"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 4.5, -7.2, -2.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+mpa_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[12]] <- mpa_model_fits
+
+mpa_model_fits
+
+```
+
+### ROT
+
+Colletive provides the best fit.
+
+```{r}
+
+target_site <- "ROT"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 3., -5.5, -2.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+rot_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[13]] <- rot_model_fits
+
+rot_model_fits
+
+```
+
+### SAU
+
+Collective provides the best fit.
+
+```{r}
+
+target_site <- "SAU"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 3.5, -6.5, -2.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+sau_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[14]] <- sau_model_fits
+
+sau_model_fits
+
+```
+
+### SOM
+
+Collective provides the best fit but misses a large section of the shoulder.
+
+```{r}
+
+target_site <- "SOM"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site) %>%
+  filter(foraging_distance < 9) # remove outlier distance
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 3., -5.5, -2.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+som_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[15]] <- som_model_fits
+
+som_model_fits
+
+```
+
+### SRA
+
+Collective provides the best fit.
+
+```{r}
+
+target_site <- "SRA"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 4., -7.5, -2.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+sra_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[16]] <- sra_model_fits
+
+sra_model_fits
+
+```
+
+### STU
+
+Individual provides the best fit. Again, the proportion of scouts does not move towards 1. I.e. the collective model fails to reduce to the individual in the MLE fit.
+
+```{r}
+
+target_site <- "STU"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12,10)
+ar_bnds <- c(1.0e-12, 10)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 5.5, -6.5, -1.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+stu_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[17]] <- stu_model_fits
+
+stu_model_fits
+
+```
+
+### SWP
+
+Collective provides the best fit.
+
+```{r}
+
+target_site <- "SWP"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 15)
+br_bnds <- c(1.0e-6, 15)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 2.5, -6, -2.)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+swp_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[18]] <- swp_model_fits
+
+swp_model_fits
+
+```
+
+### YAL
+
+Collective provides the best fit.
+
+```{r}
+
+target_site <- "YAL"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 5., -8, -2.5)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+yal_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[19]] <- yal_model_fits
+
+yal_model_fits
+
+```
+
+### ZSL
+
+Collective provides the best fit.
+
+```{r}
+
+target_site <- "ZSL"
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# fit collective model
+colletive_result <- fit_collective_model_to_data(data, bounds)
+
+# view collective model likelihood space to check bounds look ok
+colletive_result$llspace
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 50)
+as_bnds <- c(1.0e-12, 1.3)
+bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# run individual model
+individual_result <- fit_individual_model_to_data(data, bounds)
+
+# view individual model likelihood space to check bounds look ok
+individual_result$llspace
+
+# make plot of model fits
+make_full_plot(
+  data$foraging_distance,
+  list(
+    "collective" = colletive_result$solution,
+    "individual" = individual_result$solution
+  ),
+  subplot_coords = c(0.5, 2.4, -8, -2)
+)
+
+# calculate ks statistics
+ks_test_result_collective <- calc_ks_boot(
+  data$foraging_distance, colletive_result$solution$est, "collective"
+)
+ks_test_result_individual <- calc_ks_boot(
+  data$foraging_distance, individual_result$solution$est, "individual"
+)
+
+# bring results together
+zsl_model_fits <- tibble(
+  site = c(target_site, target_site),
+  model = c("collective", "individual"),
+  loglikelihood = c(
+    colletive_result$solution$fmax, individual_result$solution$fmax
+  ),
+  p = c(colletive_result$solution$est[1], 1),
+  k = c(
+    length(colletive_result$solution$est),
+    length(individual_result$solution$est)
+  ),
+  AIC = c(
+    calc_aic(
+      length(colletive_result$solution$est), colletive_result$solution$fmax
+    ),
+    calc_aic(
+      length(individual_result$solution$est), individual_result$solution$fmax
+    )
+  ),
+  ks_statistic = c(
+    ks_test_result_collective$ks$statistic[[1]],
+    ks_test_result_individual$ks$statistic[[1]]
+  ),
+  ks_pvalue = c(
+    ks_test_result_collective$ks.boot.pvalue,
+    ks_test_result_individual$ks.boot.pvalue
+  )
+)
+
+all_data[[20]] <- zsl_model_fits
+
+zsl_model_fits
+
+```
+
+## Overall findings
+
+```{r}
+
+# group all site results together
+df <- all_data %>%
+  bind_rows()
+
+# AIC plot
+aic_plot <- df %>%
+  group_by(site) %>%
+  slice(which.min(AIC)) %>%
+  select(model) %>%
+  group_by(model) %>%
+  summarise(lowest_AIC = n()) %>%
+  ggplot(aes(x = model, y = lowest_AIC)) +
+  geom_bar(stat = "identity") +
+  labs(x = "Model", y = "Count") +
+  scale_y_continuous(breaks = seq(0, 20, by = 1))
+
+ggsave(
+  plot = aic_plot,
+  filename = "analysis/figures/AIC_plot.png",
+  width = 86,
+  height = 90,
+  units = "mm",
+  dpi = 400
+)
+
+# ks plot
+ks_plot_dist <- df %>%
+  ggplot(aes(x = ks_pvalue)) +
+  geom_histogram(bins = 10, binwidth = 0.1, col = "white") +
+  geom_vline(xintercept = 0.05, color = "red", linetype = "dashed") +
+  scale_y_continuous(breaks = seq(0, 12, by = 2)) +
+  labs(x = "KS P value") +
+  facet_wrap(~model, nrow = 3)
+
+ggsave(
+  plot = ks_plot_dist,
+  filename = "analysis/figures/sites_ks.png",
+  width = 86,
+  height = 180,
+  units = "mm",
+  dpi = 400
 )
@@
 ```
+
+### Map plots
+
+```{r}
+
+library(ggplot2)
+library(cowplot)
+library(ggsn)
+library(sf)
+library(rworldmap)
+
+full_data_path <- "analysis/data/FullHBForagingData.csv"
+data_raw <- tibble(read.csv(full_data_path))
+
+map_data <- data_raw %>%
+  select(
+    site, lat, lon
+  ) %>%
+  distinct()
+
+map_data <- df %>%
+  group_by(site) %>%
+  slice(which.min(AIC)) %>%
+  select(site, model) %>%
+  left_join(map_data, on = "site")
+
+locations <- st_as_sf(
+  map_data,
+  coords = c("lon", "lat"), crs = 4326
+)
+
+zoomed_area <- st_read("analysis/shapefiles/english_region_region.shp")
+
+# cropped
+croped_data <- st_crop(
+  zoomed_area,
+  xmin = 470000,
+  xmax = 610000,
+  ymin = 130000,
+  ymax = 225000
+)
+
+# location of sites box
+selected_sites <- filter(locations, site %in% c("STU", "ZSL"))
+sites_sf_t <- st_transform(
+  selected_sites
+)
+
+points_area <- st_bbox(locations)
+
+zoom_plt <- ggplot() +
+  geom_sf(data = croped_data, fill = "#ededed") +
+  geom_sf(data = locations, aes(shape = model), size = 2) +
+  geom_sf_text(
+    data = selected_sites, aes(label = site),
+    nudge_x = c(0.5, 105), nudge_y = c(-0.5, -105)
+  ) +
+  scale_shape_manual(values = c(2, 17)) +
+  theme_nothing() +
+  scale_x_continuous(expand=c(0,0)) +
+  scale_y_continuous(expand=c(0,0)) +
+  labs(x = NULL, y = NULL) +
+  theme(
+    panel.border = element_rect(color = "black", fill = NA, size = 1),
+    text = element_text(family = "URWHelvetica", size = 8)
+  )
+
+zoom_plt
+
+# make full plot
+# get underlaying map
+worldmap <- getMap(resolution = "high")
+
+base <- ggplot() +
+  geom_polygon(data = worldmap,
+               aes(x = long, y = lat,
+                   group = group),
+               fill = '#c4cfc8',
+               color = '#4b4949d0') +
+  coord_fixed(ratio = 1.8,
+              xlim = c(-11, 13),
+              ylim = c(50, 59)) +
+  geom_point(data = map_data, aes(x = lon, y = lat), size = 0.2) +
+  geom_rect(
+    aes(
+      xmin = points_area$xmin[[1]] - 0.2, xmax = points_area$xmax[[1]] + 0.2,
+      ymin = points_area$ymin[[1]] - 0.2, ymax = points_area$ymax[[1]] + 0.2
+    ),
+    fill = NA,
+    colour = "black",
+    size = 0.2
+  ) +
+  theme_nothing() +
+  scale_x_continuous(expand=c(0,0)) +
+  scale_y_continuous(expand=c(0,0)) +
+  labs(x = NULL, y = NULL) +
+  theme(
+    panel.border = element_rect(color = "black", fill = NA, size = 1)
+  )
+
+base
+
+merge_plot <- ggdraw() +
+  draw_plot(ggplotGrob(base)) +
+  draw_plot(
+    ggplotGrob(zoom_plt), x = 0.47, y = 0.35, width = 0.5, height = 0.5
+  )
+
+merge_plot
+
+ggsave(
+  plot = merge_plot,
+  file = "analysis/figures/map_aic.png",
+  width = 86,
+  height = 90,
+  units = "mm",
+  dpi = 400
+)
+
+ggsave(
+  plot = plots[[20]][[1]],
+  file = "analysis/figures/ZSL.png",
+  width = 86,
+  height = 90,
+  units = "mm",
+  dpi = 400
+)
+
+ggsave(
+  plot = plots[[17]][[1]],
+  file = "analysis/figures/STU.png",
+  width = 86,
+  height = 90,
+  units = "mm",
+  dpi = 400
+)
+```

### 2021-09-30 R/plotting.R

@@ -46,27 +46,6 @@ make_ccdf_plot_data <- function(x, param_est, model, npoints = 100) {
   return(cumul)
 }
 
-#' Creates data for histogram plot model fit
-#'
-#' @description Using the optimised paramater estimates, creates the data for
-#' the pdf/histogram plots
-#' @param x doubleArray The foraging distance to plot
-#' @param param_est doubleArray The paramater estimates of the optimisation
-#' @param model characterArray The name of the model used
-#' @param npoints integer The number of points to plot the predicted data for.
-#' Defaults to 100.
-#' @return tibble of model data and cumulative probability
-#' @importFrom tibble tibble
-#' @export
-#'
-make_pdf_plot_data <- function(x, param_est, model, npoints = 100) {
-  model_n <- model_number_from_model(model)
-  x_seq <- seq(min(x), max(x), length.out = npoints)
-  pdf <- model_pdf(x_seq, param_est, model_n) * length(x) * ((max(x) - min(x)) / 20)
-  cumul <- tibble(x_seq, pdf)
-  return(cumul)
-}
-
 #' Creates ccdf plot of data
 #'
 #' @description Creates a ccdf plot for the data provided
@@ -132,23 +111,8 @@ make_full_plot <- function(x, model_result_list,
       return(result)
     }
   )
-  pdf_data <- map(
-    model_result_list,
-    ~ {
-      result <- make_pdf_plot_data(x, .x$est, model = .x$data_name)
-      if(any((is.nan(result$cumul_ccdf)))) {
-        message("The combination of parameters produces nan values in the ccdf")
-      }
-      return(result)
-    }
-  )
   df <- map_df(
     cdf_data,
-    I,
-    .id = "Model"
-  )
-  df_histo <- map_df(
-    pdf_data,
     I,
     .id = "Model"
   )
@@ -168,14 +132,6 @@ make_full_plot <- function(x, model_result_list,
       bins = 100,
       binwidth = (max(x) - min(x)) / 20,
       col = "white"
-    ) +
-    geom_line(
-      data = df_histo,
-      aes(
-        x = .data$x_seq,
-        y = .data$pdf,
-        colour = .data$Model
-      )
     ) +
     theme(
       axis.title.x = element_blank(),

### 2021-09-30 R/utils.R

@@ -87,61 +87,25 @@ calc_aic <- function(k, ll) {
   return(2 * k - 2 * ll)
 }
 
-#' Run a KS test on the data
-#'
-#' @description Calulctate the ks test stat and p value for a given model
-#' @param x doubleArray Foraging distance data
-#' @param param_est doubleArray The parameter estimates for a model
-#' @param model_type characterArray the name of the model ran
-#' @param pvalue Bool To return the p value (TRUE) or D statistic (FALSE).
-#' Defaults to TRUE.
-#' @importFrom stats ks.test
-#' @importFrom rlang .data
-#' @concept utility
-calc_ks <- function(x, param_est, model_type, pvalue = TRUE) {
-  if (model_type != "all") {
-    param_est <- param_est[-1]
-  }
-  param_est <- param_est[!is.na(param_est)]
-  ccdf_data <- inverse_ccdf(x)
-  ccdf_model <- model_ccdf(x, param_est, model_number_from_model(model_type))
-  if (all(is.nan(ccdf_model))) {
-    print(paste(model_type, "has bad results"))
-    return(NA)
-  }
-  ks <- ks.test(ccdf_data$prob, ccdf_model)
-  if (pvalue) {
-    return(ks$p.value)
-  } else {
-    return(ks$statistic)
-  }
-}
-
 #' Run a bootstapped KS test on the data
 #'
 #' @description Calulctate the ks test stat and p value for a given model
 #' @param x doubleArray Foraging distance data
 #' @param param_est doubleArray The parameter estimates for a model
 #' @param model_type characterArray the name of the model ran
-#' @param pvalue Bool To return the p value (TRUE) or D statistic (FALSE).
-#' Defaults to TRUE.
 #' @importFrom stats ks.test
 #' @importFrom rlang .data
 #' @importFrom Matching ks.boot
 #' @concept utility
-calc_ks_boot <- function(x, param_est, model_type, pvalue = TRUE) {
-  if (model_type != "all") {
-    param_est <- param_est[-1]
-  }
-  param_est <- param_est[!is.na(param_est)]
+calc_ks_boot <- function(x, param_est, model_type) {
   ccdf_data <- inverse_ccdf(x)
   ccdf_model <- model_ccdf(x, param_est, model_number_from_model(model_type))
   if (all(is.nan(ccdf_model))) {
-    print(paste(model_type, "has bad results"))
+    print(paste(model_type, "has NA results"))
[…]

### 2021-09-30 analysis/fit_models.Rmd
finished for the day, continue tomorrow

@@
 # make plot of model fits
-make_full_plot(
+bel_plot <- make_full_plot(
   data$foraging_distance,
@@
 )
+
+bel_plot
 
@@
 zoom_plt <- ggplot() +
-  geom_sf(data = croped_data, fill = "#ededed") +
-  geom_sf(data = locations, aes(shape = model), size = 2) +
-  geom_sf_text(
-    data = selected_sites, aes(label = site),
-    nudge_x = c(0.5, 105), nudge_y = c(-0.5, -105)
-  ) +
-  scale_shape_manual(values = c(2, 17)) +
+  geom_sf(data = croped_data, fill = "#ededed", size = .1) +
+  geom_sf(data = locations, aes(shape = model), size = 1.5) +
+  geom_sf(data = selected_sites, aes(shape = site, size = 1)) +
+  scale_shape_manual(values = c(1, 2, 17, 16)) +
   theme_nothing() +
@@
 
+zsl_model_fit
+
 ggsave(

### 2021-10-01 analysis/fit_models.Rmd
made individual plots. need combining for chapter 1 in gimp.

--- 2021-09-22 analysis/fit_models.Rmd
@@
     theme(
-      text = element_text(family = "URWHelvetica", size = 8)
+      text = element_text(family = "DejaVuSerif", size = 8)
     )
@@
 
-```
-
-## Optimising each site
-
-In this document I have fit the collective and individual model to each site.
-So al the sites have reasonable likelihood values, however in some cases the fits are simply wrong (i.e.
-they show the probability of sampling a value greater than on equal to $x$ as increasing in the tail).
-I need to trouble shoot these with Vincent.
-I also need to get some advice on how to plot the model fits to the histogram of data.
-
-Would be a good idea to knit this as a single document / run in R studio on the laptop.
-
-### BEL
-
-Provides a good fit on the data. All parameters look central and nicely covered.
-
-```{r}
-
-target_site <- "BEL"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 10)
-br_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 0.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 0.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
+library(showtext)
+showtext_auto()
+
+run_wagglefit_analysis <- function(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+  ) {
+
+  # run collective model
+  colletive_result <- fit_collective_model_to_data(data, collective_bounds)
 
   # run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
+  individual_result <- fit_individual_model_to_data(data, individual_bounds)
 
   # make plot of model fits
-make_full_plot(
+  full_plot <- make_full_plot(
     data$foraging_distance,
@@
     ),
-  subplot_coords = c(0.5, 10, -5.5, -2)
+    subplot_coords = subplot_coords
+  )
+
+  # calculate ks statistics
+  ks_test_result_collective <- calc_ks_boot(
+    data$foraging_distance, colletive_result$solution$est, "collective"
+  )
+  ks_test_result_individual <- calc_ks_boot(
+    data$foraging_distance, individual_result$solution$est, "individual"
   )
@@
   # bring results together
-bel_model_fits <- tibble(
+  model_fits <- tibble(
     site = c(target_site, target_site),
@@
     p = c(colletive_result$solution$est[1], 1),
+    bs = c(colletive_result$solution$est[2], individual_result$solution$est[1]),
+    br = c(colletive_result$solution$est[3], NA),
+    as = c(colletive_result$solution$est[4], individual_result$solution$est[2]),
+    ar = c(colletive_result$solution$est[5], NA),
     k = c(
@@
       )
-  )
-)
-
-bel_model_fits
+    ),
+    ks_statistic = c(
+      ks_test_result_collective$ks$statistic[[1]],
+      ks_test_result_individual$ks$statistic[[1]]
+    ),
+    ks_pvalue = c(
+      ks_test_result_collective$ks.boot.pvalue,
+      ks_test_result_individual$ks.boot.pvalue
+    )
+  )
+
+  return(
+    list(
+      fit_result = model_fits, fit = full_plot,
+      individual_llspace = individual_result$llspace,
+      collective_llspace = colletive_result$llspace
+    )
+  )
+}
+
+all_sites <- as.list(rep(0, 20))
+names(all_sites) <- get_data() %>%
+    select(site) %>%
+    unique() %>%
+    pull()
+
+```
+
+## Optimising each site
+
+In this document I have fit the collective and individual model to each site.
+So al the sites have reasonable likelihood values, however in some cases the fits are simply wrong (i.e.
+they show the probability of sampling a value greater than on equal to $x$ as increasing in the tail).
+I need to trouble shoot these with Vincent.
+I also need to get some advice on how to plot the model fits to the histogram of data.
+
+Would be a good idea to knit this as a single document / run in R studio on the laptop.
+
+### BEL
+
+Provides a good fit on the data. All parameters look central and nicely covered.
+
+```{r}
+
+target_site <- "BEL"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 0.1)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 10, -5.5, -2)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
 
@@
 target_site <- "BFI"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
 
@@
 
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
 # set up bounds for the individual model
@@
 
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 4.5, -8, -2.5)
-)
-
-# bring results together
-bfi_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-bfi_model_fits
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 4.5, -6.5, -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
 
@@
 
-CCDF plot is not right. Cant go back on its self which indicates there is probably some round off error somewhere throwing things off.
-
 The individual fit isnt great and the Bs paramater is increasing up to the boundry, indicating it can only really take on a straight line / exponential fit.
-The parameters for the collective model are fairly central which raises questions as to why the fit is misbehaved in the tail.
+The parameters for the collective model are fairly central in the likelihood space, the fit looks very good.
 
@@
 target_site <- "BLO"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
 
@@
 ar_bnds <- c(1.0e-10, 1.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 105)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 5505)
 as_bnds <- c(1.0e-12, 0.1)
-bounds <- rbind(
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 12, -8., -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### BUR
+
+Bs and Br go in oposite directions. E.g. Bs approaches 0 whilst Br approaches an every higher number.
+
+Fit looks good.
+
+```{r}
+
+target_site <- "BUR"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 500)
+br_bnds <- c(1.0e-6, 500)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 0.1)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 5.5, -6.5, -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### CAD
+
+The collective model roughly follows the individual model but is able to acheive a slightly improved fit to the shoulder, hence it provides a higher likelihood score.
+The AIC indicates this is overfitting, suggesting the individual model provides a more parsimonious explanation.
+
+$p$ does not approach 1 as one might expect, indicating there is overfitting and so a comparison with an individual model is required.
+
+```{r}
+
+target_site <- "CAD"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 10)
+ar_bnds <- c(1.0e-12, 10)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 2)
+as_bnds <- c(1.0e-12, 0.8)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 2.5, -5, -1.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### GIL
+
+The collective model provides the most parsimonious and best fit.
+
+```{r}
+
+target_site <- "GIL"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 0.8)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 2.7, -8., -2)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### HER
+
+The collective model provides the best fit to the data, but the proportion of scouts is high.
+
+```{r}
+
+target_site <- "HER"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 100)
+br_bnds <- c(1.0e-6, 100)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 2)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 5, -6.5, -1.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### HHS
+
+Collective model provides the best fit but falls under the tail and shoulder. The individual model strugles to find a good fit, probably due to the tail.
+
+```{r}
+
+target_site <- "HHS"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 50)
+br_bnds <- c(1.0e-6, 50)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 6, -6.5, -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### HOR
+
+Collective model provides a very good fit, whilst the individual model fails to find much traction.
+The proportion of scouts goes very low (~3%) suggesting the majority of the colony are following a small number of scouting individuals.
+
+```{r}
+
+target_site <- "HOR"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 4, -9, -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### MAK
+
+Again the collective model provides a good fit to the data, however, the individual model fits poorly, reducing to an exponential.
+
+```{r}
+
+target_site <- "MAK"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 20)
+as_bnds <- c(1.0e-12, 1.1)
+individual_bounds <- rbind(
   bs_bnds, as_bnds
@@
 
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 12, -8., -2.5)
-)
-
-# bring results together
-blo_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-blo_model_fits
-
-```
-
-### BUR
-
-Bs and Br go in oposite directions. E.g. Bs approaches 0 whilst Br approaches an every higher number.
-
-Fit looks good.
-
-```{r}
-
-target_site <- "BUR"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 500)
-br_bnds <- c(1.0e-6, 500)
-as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 1.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 0.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 5.5, -6.5, -2.5)
-)
-
-# bring results together
-bur_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-bur_model_fits
-
-```
-
-### CAD
-
-The only graph I can get to make sense is one where the collective reduces to the individual. The others show the ccdf going off.
-
-```{r}
-
-target_site <- "CAD"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 100)
-br_bnds <- c(1.0e-6, 100)
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 5, -6.5, -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### MEL
+
+Collective model provides the best fit, however it misses a large section of the shoulder for the tail.
+The $bs$ parameter wants to go to zero, however when let go bellow 1e-6 the behaviour becomes very erratic and the fit deteriorates.
+
+```{r}
+
+target_site <- "MEL"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 3.5, -7, -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### MPA
+
+Collective provides the best fit.
+
+```{r}
+
+target_site <- "MPA"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 4.5, -7.2, -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### ROT
+
+Colletive provides the best fit.
+
+```{r}
+
+target_site <- "ROT"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 3., -5.5, -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### SAU
+
+Collective provides the best fit.
+
+```{r}
+
+target_site <- "SAU"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# set coordinates for histogram subplot
+subplot_coords <- c(0.5, 3.5, -6.5, -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### SOM
+
+Collective provides the best fit but misses a large section of the shoulder.
+
+```{r}
+
+target_site <- "SOM"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site) %>%
+  filter(foraging_distance < 9) # remove outlier distance
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+ # set coordinates for histogram subplot
+subplot_coords <- c(0.5, 3., -5.5, -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### SRA
+
+Collective provides the best fit.
+
+```{r}
+
+target_site <- "SRA"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.1)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+ # set coordinates for histogram subplot
+subplot_coords <- c(0.5, 4., -7.5, -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+### STU
+
+Individual provides the best fit. Again, the proportion of scouts does not move towards 1. I.e. the collective model fails to reduce to the individual in the MLE fit.
+
+```{r}
+
+target_site <- "STU"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
 as_bnds <- c(1.0e-12, 10)
 ar_bnds <- c(1.0e-12, 10)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 2)
-as_bnds <- c(1.0e-12, 0.8)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 2.5, -5, -1.5)
-)
-
-# bring results together
-cad_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-cad_model_fits
-
-```
-
-### GIL
-
-Again, the likelihood scores look good but the model fits are doing crazy things at the tail!
-
-```{r}
-
-target_site <- "GIL"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 10)
-br_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 1.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 0.8)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 2.7, -8., -2)
-)
-
-# bring results together
-gil_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-gil_model_fits
-
-```
-
-### HER
-
-Also has a weird CCDF plot
-
-```{r}
-
-target_site <- "HER"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 100)
-br_bnds <- c(1.0e-6, 100)
-as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 1.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 2)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 5, -6.5, -1.5)
-)
-
-# bring results together
-her_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-her_model_fits
-
-```
-
-### HHS
-
-Shit ccdf.
-
-```{r}
-
-target_site <- "HHS"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 50)
-br_bnds <- c(1.0e-6, 50)
-as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 1.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
 
@@
 as_bnds <- c(1.0e-12, 1.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 6, -6.5, -2.5)
-)
-
-# bring results together
-hhs_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-hhs_model_fits
-
-```
-
-### HOR
-
-Agin weird behaviour in the tail.
-Seems to be something to do with the Bs parameter approaching the lower bound.
-If the lower bound is made closer to zero (just 1e-7) the plot goes mad and the log-likelihood score becomes very large, indicating a round off error or something stupid.
-
-```{r}
-
-target_site <- "HOR"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 10)
-br_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 1.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 4, -9, -2.5)
-)
-
-# bring results together
-hor_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-hor_model_fits
-
-```
-
-### MAK
-
-Same problem as with the others.
-
-```{r}
-
-target_site <- "MAK"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 10)
-br_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 0.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 5, -6.5, -2.5)
-)
-
-# bring results together
-mak_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-mak_model_fits
-
-```
-
-### MEL
-
-Slight misbehaviour in the tail creaping up.
-
-```{r}
-
-target_site <- "MEL"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 10)
-br_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 0.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 3.5, -7, -2.5)
-)
-
-# bring results together
-mel_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-mel_model_fits
-
-```
-
-### MPA
-
-Same problem with the tail.
-
-```{r}
-
-target_site <- "MPA"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 10)
-br_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 0.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 4.5, -7.2, -2.5)
-)
-
-# bring results together
-mpa_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-mpa_model_fits
-
-```
-
-### ROT
-
-Same problem with the tail.
-
-```{r}
-
-target_site <- "ROT"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 10)
-br_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 0.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 3., -5.5, -2.5)
-)
-
-# bring results together
-rot_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-rot_model_fits
-
-```
-
-### SAU
-
-No problems here!
-
-```{r}
-
-target_site <- "SAU"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 10)
-br_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 0.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 3.5, -6.5, -2.5)
-)
-
-# bring results together
-sau_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-sau_model_fits
-
-```
-
-### SOM
-
-Same problem with the tail.
-
-```{r}
-
-target_site <- "SOM"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 10)
-br_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 0.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site) %>%
-  filter(foraging_distance < 9) # remove outlier distance
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 3., -5.5, -2.5)
-)
-
-# bring results together
-som_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-som_model_fits
-
-```
-
-### SRA
-
-Same problem with the tail.
-
-```{r}
-
-target_site <- "SRA"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 10)
-br_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 0.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 4., -7.5, -2.5)
-)
-
-# bring results together
-sra_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-sra_model_fits
-
-```
-
-### STU
-
-Same problem with the tail.
-Increasing the bounds for as and ar seems to coerce it towards the better visual fit, despite it not changing what they optimise at.
-Weird thing is the model is getting close in the region of p but not actually to it.
-
-```{r}
-
-target_site <- "STU"
-
-# set up bounds for the collective model
-p_bnds <- c(0, 1.0)
-bs_bnds <- c(1.0e-6, 10)
-br_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12,10)
-ar_bnds <- c(1.0e-12, 10)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
-
-# set up bounds for the individual model
-bs_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 1.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 5.5, -6.5, -1.5)
-)
-
-# bring results together
-stu_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-stu_model_fits
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+ # set coordinates for histogram subplot
+subplot_coords <- c(0.5, 5.5, -6.5, -1.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
 
@@
 
-Same problem with the tail. Weird P behaviour when we move some of the other parameters around. Looks like it can get a fit that looks very similar with p = 0
+Collective provides the best fit.
 
@@
 target_site <- "SWP"
+
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
 
@@
 ar_bnds <- c(1.0e-12, 0.5)
-bounds <- rbind(
-  p_bnds, bs_bnds,
-  br_bnds, as_bnds,
-  ar_bnds
-)
-
-# subset data for target site
-data <- get_data() %>%
-  filter(site == target_site)
-
-# fit collective model
-colletive_result <- fit_collective_model_to_data(data, bounds)
-
-# view collective model likelihood space to check bounds look ok
-colletive_result$llspace
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
 
@@
 as_bnds <- c(1.0e-12, 1.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 2.5, -6, -2.)
-)
-
-# bring results together
-swp_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-swp_model_fits
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+ # set coordinates for histogram subplot
+subplot_coords <- c(0.5, 2.5, -6, -2.)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
 
@@
 
-Same problem with the tail.
+Collective provides the best fit.
 
@@
 
+# subset data for target site
+data <- get_data() %>%
+  filter(site == target_site)
+
 # set up bounds for the collective model
@@
 as_bnds <- c(1.0e-12, 1.1)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
-  ),
-  subplot_coords = c(0.5, 5., -8, -2.5)
-)
-
-# bring results together
-yal_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-yal_model_fits
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+ # set coordinates for histogram subplot
+subplot_coords <- c(0.5, 5., -8, -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
 
@@
 as_bnds <- c(1.0e-12, 1.3)
-bounds <- rbind(
-  bs_bnds, as_bnds
-)
-
-# run individual model
-individual_result <- fit_individual_model_to_data(data, bounds)
-
-# view individual model likelihood space to check bounds look ok
-individual_result$llspace
-
-# make plot of model fits
-make_full_plot(
-  data$foraging_distance,
-  list(
-    "collective" = colletive_result$solution,
-    "individual" = individual_result$solution
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+ # set coordinates for histogram subplot
+subplot_coords <- c(0.5, 2.4, -8, -2)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[[target_site]]
+
+all_sites[[target_site]]$fit
+
+```
+
+## Overall findings
+
+```{r}
+
+# group all site results together
+df <- map(all_sites, 1) %>%
+  bind_rows()
+
+# AIC plot
+aic_plot <- df %>%
+  group_by(site) %>%
+  slice(which.min(AIC)) %>%
+  select(model) %>%
+  group_by(model) %>%
+  summarise(lowest_AIC = n()) %>%
+  ggplot(aes(x = model, y = lowest_AIC)) +
+  geom_bar(stat = "identity") +
+  labs(x = "Model", y = "Count") +
+  scale_y_continuous(breaks = seq(0, 20, by = 1)) +
+  theme(
+    text = element_text(size = 12)
+  )
+
+ggsave(
+  plot = aic_plot,
+  filename = "analysis/figures/AIC_plot.png",
+  width = 90,
+  height = 110,
+  units = "mm",
+  dpi = 400
+)
+
+# ks plot
+ks_plot_dist <- df %>%
+  ggplot(aes(x = ks_pvalue)) +
+  geom_histogram(bins = 10, binwidth = 0.1, col = "white") +
+  geom_vline(xintercept = 0.05, color = "red", linetype = "dashed") +
+  scale_y_continuous(breaks = seq(0, 12, by = 2)) +
+  labs(x = "KS P value") +
+  facet_wrap(~model, nrow = 3) +
+  theme(
+    text = element_text(size = 12)
+  )
+
+ggsave(
+  plot = ks_plot_dist,
+  filename = "analysis/figures/sites_ks.png",
+  width = 86,
+  height = 180,
+  units = "mm",
+  dpi = 400
+)
+
+ggsave(
+  plot = all_sites$STU$fit,
+  filename = "analysis/figures/STU.png",
+  width = 90,
+  height = 110,
+  units = "mm",
+  dpi = 400
+)
+
+ggsave(
+  plot = all_sites$ZSL$fit,
+  filename = "analysis/figures/ZSL.png",
+  width = 90,
+  height = 110,
+  units = "mm",
+  dpi = 400
+)
+
+```
+
+### Map plots
+
+Code to make the individual mal plots. These figures are created standalone but are combined in to facets manually in an image processor.
+
+```{r}
+
+library(ggplot2)
+library(ggrepel)
+library(gridExtra)
+library(ggsn)
+library(sf)
+library(rworldmap)
+library(ggspatial)
+library(rnaturalearth)
+library(rnaturalearthdata)
+
+full_data_path <- "analysis/data/FullHBForagingData.csv"
+data_raw <- tibble(read.csv(full_data_path))
+
+map_data <- data_raw %>%
+  select(
+    site, lat, lon
+  ) %>%
+  distinct()
+
+map_data <- df %>%
+  group_by(site) %>%
+  slice(which.min(AIC)) %>%
+  select(site, model) %>%
+  left_join(map_data, on = "site") %>%
+  mutate(col = ifelse(site %in% c("STU", "ZSL"), "1", "0"))
+
+locations <- st_as_sf(
+  map_data,
+  coords = c("lon", "lat"), crs = 4326
+)
+
+# Extract selected sites for figure
+selected_sites <- filter(map_data, site %in% c("STU", "ZSL"))
+
+points_area <- st_bbox(locations)
+
+worldmap <- ne_countries(scale = "large", returnclass = "sf")
+
+# london area
+london <- st_read(
+  "analysis/shapefiles/statistical-gis-boundaries-london/ESRI/London_Ward.shp"
+)
+
+london <- st_transform(
+  london,
+  CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
+)
+
+inset <- ggplot() +
+  geom_sf(
+    data = worldmap,
+    fill = 'grey90',
+    color = '#4b4949d0'
+  ) +
+  geom_sf(
+    data = london,
+    fill = '#ced0cffc',
+    lwd = 0
+  ) +
+  coord_sf(
+    xlim = c(points_area[[1]] - 0.1, points_area[[3]] + 0.1),
+    ylim = c(points_area[[2]] - 0.1, points_area[[4]] + 0.1)
+  ) +
+  geom_point(
+    data = map_data,
+    aes(x = lon, y = lat, shape = model, colour = col), size = 1
+  ) +
+  geom_text(
+    data = selected_sites, aes(x = lon, y = lat, label = site),
+    nudge_x = c(0, -0.05), nudge_y = c(0.05, 0.05)
+  ) +
+  annotation_north_arrow(
+    location = "tr", which_north = "true",
+    style = north_arrow_fancy_orienteering
+  ) +
+  annotation_scale(location = "br") +
+  scale_shape_manual(values = c(1, 2)) +
+  scale_colour_manual(values = c("black", "red")) +
+  theme_nothing() +
+  scale_x_continuous(expand = c(0,0)) +
+  scale_y_continuous(expand = c(0,0)) +
+  labs(x = NULL, y = NULL) +
+  theme(
+    panel.border = element_rect(color = "black", fill = NA, size = .5),
+    text = element_text(size = 12)
+  )
+
+inset
+
+# make full plot
+base <- ggplot(data = worldmap) +
+  geom_sf(
+    fill = '#c4cfc8',
+    color = '#4b4949d0'
+  ) +
+  coord_sf(
+    xlim = c(-11, 14),
+    ylim = c(47.5, 60)
+  ) +
+  geom_point(data = map_data, aes(x = lon, y = lat), size = 0.2) +
+  geom_rect(
+    aes(
+      xmin = points_area$xmin[[1]] - 0.1, xmax = points_area$xmax[[1]] + 0.1,
+      ymin = points_area$ymin[[1]] - 0.1, ymax = points_area$ymax[[1]] + 0.1
     ),
-  subplot_coords = c(0.5, 2.4, -8, -2)
-)
-
-# bring results together
-zsl_model_fits <- tibble(
-  site = c(target_site, target_site),
-  model = c("collective", "individual"),
-  loglikelihood = c(
-    colletive_result$solution$fmax, individual_result$solution$fmax
-  ),
-  p = c(colletive_result$solution$est[1], 1),
-  k = c(
-    length(colletive_result$solution$est),
-    length(individual_result$solution$est)
-  ),
-  AIC = c(
-    calc_aic(
-      length(colletive_result$solution$est), colletive_result$solution$fmax
-    ),
-    calc_aic(
-      length(individual_result$solution$est), individual_result$solution$fmax
-    )
-  )
-)
-
-zsl_model_fits
-
-```
+    fill = NA,
+    colour = "black",
+    size = .02
+  ) +
+  theme_nothing() +
+  scale_x_continuous(expand = c(0,0)) +
+  scale_y_continuous(expand = c(0,0)) +
+  labs(x = NULL, y = NULL) +
+  theme(
+    panel.border = element_rect(color = "black", fill = NA, size = .5)
+  )
+
+base
+
+merge_plot <- base +
+  annotation_custom(
+    ggplotGrob(inset),
+    xmin = 1,
+    xmax = 13,
+    ymin = 52.5,
+    ymax = 60
+  )
+
+merge_plot
+
+ggsave(
+  plot = merge_plot,
+  filename = "analysis/figures/site_map.png",
+  width = 90,
+  height = 110,
+  units = "mm",
+  dpi = 400
+)
+
+```

### 2021-10-01 R/plotting.R

@@ -124,6 +124,9 @@ make_full_plot <- function(x, model_result_list,
         y = log(.data$cumul_ccdf),
         colour = .data$Model
       )
+    ) +
+    theme(
+      legend.position = "none"
     )
 
   histoplot <- as_tibble(x) %>%

### 2021-10-04 analysis/fit_models.Rmd
got plotting to work in a grid, different but looks ok

@@
   labs(x = "Model", y = "Count") +
-  scale_y_continuous(breaks = seq(0, 20, by = 1)) +
+  scale_y_continuous(breaks = seq(0, 20, by = 2)) +
   theme(
@@
 # Extract selected sites for figure
-selected_sites <- filter(map_data, site %in% c("STU", "ZSL"))
+selected_sites <- filter(map_data, site %in% c("STU", "ZSL")) %>%
+  mutate(
+    label = ifelse(site == "STU", "C (STU)", "D (ZSL)")
+  )
 
@@
     data = map_data,
-    aes(x = lon, y = lat, shape = model, colour = col), size = 1
+    aes(x = lon, y = lat, shape = model, colour = col), size = 1.5
   ) +
   geom_text(
-    data = selected_sites, aes(x = lon, y = lat, label = site),
-    nudge_x = c(0, -0.05), nudge_y = c(0.05, 0.05)
+    data = selected_sites, aes(x = lon, y = lat, label = label),
+    nudge_x = c(0, -0.07), nudge_y = c(0.05, 0.05)
   ) +
@@
     location = "tr", which_north = "true",
-    style = north_arrow_fancy_orienteering
+    style = north_arrow_fancy_orienteering,
+    height = unit(10, "mm"),
+    width = unit(10, "mm")
   ) +
@@
     fill = '#c4cfc8',
-    color = '#4b4949d0'
+    color = '#4b4949d0',
+    lwd = 0.2
   ) +
   coord_sf(
-    xlim = c(-11, 14),
-    ylim = c(47.5, 60)
+    xlim = c(-11, 3),
+    ylim = c(49.5, 60)
   ) +
-  geom_point(data = map_data, aes(x = lon, y = lat), size = 0.2) +
+  # geom_point(data = map_data, aes(x = lon, y = lat), size = 0.2) +
   geom_rect(
@@
 
-```
+merge_plot_2 <- inset +
+  annotation_custom(
+    ggplotGrob(base),
+    xmin = -3.35,
+    xmax = 2,
+    ymin = 51.05,
+    ymax = 51.35
+  )
+
+merge_plot_2
+
+sites_model_plot <- plot_grid(
+  merge_plot_2, aic_plot, all_sites$STU$fit, all_sites$ZSL$fit,
+  labels = c("A", "B", "C", "D")
+)
+
+ggsave(
+  plot = sites_model_plot,
+  filename = "analysis/figures/sites_model_plot.png",
+  width = 90,
+  height = 110,
+  units = "mm",
+  dpi = 300,
+  scaling = 0.5
+)
+
+```

### 2021-10-04 R/plotting.R

@@ -63,7 +63,7 @@ make_base_plot <- function(x, logit = TRUE) {
   theme_set(
     theme_classic() +
       theme(
-        text = element_text(family = "URWHelvetica", size = 8)
+        text = element_text(family = "URWHelvetica", size = 12)
       )
   )
   if (logit) {

### 2021-10-04 analysis/fit_models.Rmd
Updated figures so text is correct size for chapter 1 latex file

@@
 as_bnds <- c(1.0e-12, 1.5)
-ar_bnds <- c(1.0e-12, 0.5)
+ar_bnds <- c(1.0e-12, 1.5)
 collective_bounds <- rbind(
@@
   theme(
-    text = element_text(size = 12)
+    text = element_text(size = 42)
   )
@@
   units = "mm",
-  dpi = 400
+  dpi = 300
 )
@@
   theme(
-    text = element_text(size = 12)
+    text = element_text(size = 42),
+    strip.background = element_blank()
   )
@@
     panel.border = element_rect(color = "black", fill = NA, size = .5),
-    text = element_text(size = 12)
+    text = element_text(size = 82)
   )
@@
   filename = "analysis/figures/sites_model_plot.png",
-  width = 90,
-  height = 110,
+  width = 183,
+  height = 190,
   units = "mm",
-  dpi = 300,
-  scaling = 0.5
-)
-
-```
+  dpi = 300
+)
+
+```
+
+```{r}
+# other map plots
+inset <- ggplot() +
+  geom_sf(
+    data = worldmap,
+    fill = 'grey90',
+    color = '#4b4949d0'
+  ) +
+  geom_sf(
+    data = london,
+    fill = '#ced0cffc',
+    lwd = 0
+  ) +
+  coord_sf(
+    xlim = c(points_area[[1]] - 0.1, points_area[[3]] + 0.1),
+    ylim = c(points_area[[2]] - 0.1, points_area[[4]] + 0.1)
+  ) +
+  geom_point(
+    data = map_data,
+    aes(x = lon, y = lat, shape = model, colour = model), size = 1.5
+  ) +
+  annotation_north_arrow(
+    location = "tr", which_north = "true",
+    style = north_arrow_fancy_orienteering,
+    height = unit(10, "mm"),
+    width = unit(10, "mm")
+  ) +
+  annotation_scale(
+    location = "br",
+    text_cex = 3
+  ) +
+  scale_shape_manual(values = c(1, 2)) +
+  scale_colour_manual(values = c("black", "red")) +
+  theme_nothing() +
+  scale_x_continuous(expand = c(0,0)) +
+  scale_y_continuous(expand = c(0,0)) +
+  labs(x = NULL, y = NULL) +
+  theme(
+    panel.border = element_rect(color = "black", fill = NA, size = .5),
+    text = element_text(size = 42)
+  )
+
+inset
+
+# make full plot
+base <- ggplot(data = worldmap) +
+  geom_sf(
+    fill = '#c4cfc8',
+    color = '#4b4949d0',
+    lwd = 0.2
+  ) +
+  coord_sf(
+    xlim = c(-11, 15),
+    ylim = c(49.5, 60)
+  ) +
+  geom_point(data = map_data, aes(x = lon, y = lat), size = 0.2) +
+  geom_rect(
+    aes(
+      xmin = points_area$xmin[[1]] - 0.1, xmax = points_area$xmax[[1]] + 0.1,
+      ymin = points_area$ymin[[1]] - 0.1, ymax = points_area$ymax[[1]] + 0.1
+    ),
+    fill = NA,
+    colour = "black",
+    size = .02
+  ) +
+  theme_nothing() +
+  scale_x_continuous(expand = c(0,0)) +
+  scale_y_continuous(expand = c(0,0)) +
+  labs(x = NULL, y = NULL) +
+  theme(
+    panel.border = element_rect(color = "black", fill = NA, size = .5)
+  )
+
+base
+
+merge_plot <- base +
+  annotation_custom(
+    ggplotGrob(inset),
+    xmin = 1,
+    xmax = 13,
+    ymin = 52.5,
+    ymax = 60
+  )
+
+merge_plot
+
+ggsave(
+  plot = merge_plot,
+  filename = "analysis/figures/site_map.png",
+  width = 183,
+  height = 190,
+  units = "mm",
+  dpi = 300
+)
+
+sites_model_plot <- plot_grid(
+  all_sites$STU$fit, all_sites$ZSL$fit, aic_plot, ks_plot_dist,
+  labels = c("A", "B", "C", "D"), label_size = 42
+)
+
+sites_model_plot
+
+ggsave(
+  plot = sites_model_plot,
+  filename = "analysis/figures/results_model_plot.png",
+  width = 183,
+  height = 190,
+  units = "mm",
+  dpi = 300
+)
+
+```

### 2021-10-04 R/plotting.R

@@ -63,7 +63,7 @@ make_base_plot <- function(x, logit = TRUE) {
   theme_set(
     theme_classic() +
       theme(
-        text = element_text(family = "URWHelvetica", size = 12)
+        text = element_text(family = "URWHelvetica", size = 42)
       )
   )
   if (logit) {

### 2021-10-05 R/RcppExports.R
updated and fixed tests

@@ -33,7 +33,7 @@ recruit_ccdf <- function(x, m, br, ar) {
     .Call('_wagglefit_recruit_ccdf', PACKAGE = 'wagglefit', x, m, br, ar)
 }
 
-#' Model ccdf function for scout and recruit superposition. Stores results in
+#' Model ccdf function for collective model. Stores results in
 #' given array (y)
 #'
 #' @param x NumericVector foraging distances
@@ -42,18 +42,18 @@ recruit_ccdf <- function(x, m, br, ar) {
 #' @inheritParams scout_ccdf
 #' @inheritParams recruit_ccdf
 #' @export
-ccdf_model_all <- function(x, y, p, bs, br, as, ar) {
-    invisible(.Call('_wagglefit_ccdf_model_all', PACKAGE = 'wagglefit', x, y, p, bs, br, as, ar))
+ccdf_model_collective <- function(x, y, p, bs, br, as, ar) {
+    invisible(.Call('_wagglefit_ccdf_model_collective', PACKAGE = 'wagglefit', x, y, p, bs, br, as, ar))
 }
 
-#' Model ccdf function for scout model. Stores results in given array (y)
+#' Model ccdf function for individual model. Stores results in given array (y)
 #'
 #' @param x NumericVector foraging distances
 #' @param y NumericVector storage array for the results
 #' @inheritParams scout_ccdf
 #' @export
-ccdf_model_scout <- function(x, y, bs, as) {
-    invisible(.Call('_wagglefit_ccdf_model_scout', PACKAGE = 'wagglefit', x, y, bs, as))
+ccdf_model_individual <- function(x, y, bs, as) {
+    invisible(.Call('_wagglefit_ccdf_model_individual', PACKAGE = 'wagglefit', x, y, bs, as))
 }
 
 #' Get model ccdf for a given model

### 2021-10-05 src/RcppExports.cpp

@@ -43,9 +43,9 @@ BEGIN_RCPP
     return rcpp_result_gen;
 END_RCPP
 }
-// ccdf_model_all
-void ccdf_model_all(NumericVector x, NumericVector y, double p, double bs, double br, double as, double ar);
-RcppExport SEXP _wagglefit_ccdf_model_all(SEXP xSEXP, SEXP ySEXP, SEXP pSEXP, SEXP bsSEXP, SEXP brSEXP, SEXP asSEXP, SEXP arSEXP) {
+// ccdf_model_collective
+void ccdf_model_collective(NumericVector x, NumericVector y, double p, double bs, double br, double as, double ar);
+RcppExport SEXP _wagglefit_ccdf_model_collective(SEXP xSEXP, SEXP ySEXP, SEXP pSEXP, SEXP bsSEXP, SEXP brSEXP, SEXP asSEXP, SEXP arSEXP) {
 BEGIN_RCPP
     Rcpp::RNGScope rcpp_rngScope_gen;
     Rcpp::traits::input_parameter< NumericVector >::type x(xSEXP);
@@ -55,20 +55,20 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< double >::type br(brSEXP);
     Rcpp::traits::input_parameter< double >::type as(asSEXP);
     Rcpp::traits::input_parameter< double >::type ar(arSEXP);
-    ccdf_model_all(x, y, p, bs, br, as, ar);
+    ccdf_model_collective(x, y, p, bs, br, as, ar);
     return R_NilValue;
 END_RCPP
 }
-// ccdf_model_scout
-void ccdf_model_scout(NumericVector x, NumericVector y, double bs, double as);
-RcppExport SEXP _wagglefit_ccdf_model_scout(SEXP xSEXP, SEXP ySEXP, SEXP bsSEXP, SEXP asSEXP) {
+// ccdf_model_individual
+void ccdf_model_individual(NumericVector x, NumericVector y, double bs, double as);
+RcppExport SEXP _wagglefit_ccdf_model_individual(SEXP xSEXP, SEXP ySEXP, SEXP bsSEXP, SEXP asSEXP) {
 BEGIN_RCPP
     Rcpp::RNGScope rcpp_rngScope_gen;
     Rcpp::traits::input_parameter< NumericVector >::type x(xSEXP);
[…]

### 2021-10-05 tests/testthat/test-cpp_ccdfs.R

@@ -1,27 +1,15 @@
-test_heaviside <- function() {
-  tests <- c(-10, -1, -0.4, -0.001, 0, 0.6, 1, 100)
-  expected <- c(0, 0, 0, 0, 1, 1, 1, 1)
-  actual <- purrr::map_dbl(
-    tests,
-    heaviside
-  )
-  test_that("heaviside works as expected", {
-    expect_identical(expected, actual)
-  })
-}
-
-test_scout_ccdf <- function(x, p, ls, qn, a) {
+test_scout_ccdf <- function(x, m, bs, as) {
   expected <- c(
-    0.41257543, 0.2264957,
-    0.33921855, 0.18375567,
-    0.11885913, 0.2264957,
-    0.09461152, 0.18375567,
-    1.5, 0.23602505
+    0.4204785, 0.0000000,
+    0.1269553, 0.0000000,
+    0.0000000, 0.0000000,
+    0.0000000, 0.0000000,
+    1.0000000, 0.0000000
   )
   actual <- purrr::map_dbl(
     x,
     ~ {
-      scout_ccdf(.x, p, ls, qn, a, min(x))
+      scout_ccdf(.x, m, bs, as)
     }
   )
   test_that("scout ccdf returns expected results", {
@@ -29,18 +17,18 @@ test_scout_ccdf <- function(x, p, ls, qn, a) {
   })
 }
 
-test_recruit_ccdf <- function(x, p, ln, qn, a) {
+test_recruit_ccdf <- function(x, m, br, ar) {
   expected <- c(
-    0.41534078, 0.11970665,
-    0.30781376, 0.06187754,
-    0.0100956, 0.11970665,
-    0.00273772, 0.06187754,
-    0.5, 0.13444239
+    0.9502439, 0.6768375,
+    0.8748425, 0.5690817,
+    0.3669045, 0.6768375,
+    0.2807545, 0.5690817,
+    1.0000000, 0.6981969
   )
   actual <- purrr::map_dbl(
     x,
     ~ {
-      recruit_ccdf(.x, p, ln, qn, a, min(x))
+      recruit_ccdf(.x, m, br, ar)
     }
   )
   test_that("scout ccdf returns expected results", {
@@ -48,32 +36,32 @@ test_recruit_ccdf <- function(x, p, ln, qn, a) {
   })
 }
 
-test_ccdf_model_all <- function(x, p, ls, ln, qn, a) {
+test_ccdf_model_collective <- function(x, p, bs, br, as, ar) {
[…]

### 2021-10-05 tests/testthat/test-cpp_models.R

@@ -1,32 +1,32 @@
-test_scout_dist <- function(x, m, p, ls, qn, a) {
-  truth <- 0.8000732527680952
-  ans <- scout_dist(x, m, p, ls, qn, a)
+test_scout_dist <- function(x, m, bs, as) {
+  truth <- 4.15163
+  ans <- scout_dist(x, m, bs, as)
   test_that("scout_dist gives expected result", {
-    expect_equal(truth, ans)
+    expect_equal(truth, round(ans, 5))
   })
 }
 
-test_recruit_dist <- function(x, m, p, ln, qn, a) {
-  truth <- 1.0095546946431222
-  ans <- recruit_dist(x, m, p, ln, qn, a)
+test_recruit_dist <- function(x, m, br, ar) {
+  truth <- 0.63748
+  ans <- recruit_dist(x, m, br, ar)
   test_that("recruit_dist gives expected results", {
     expect_identical(round(truth, 5), round(ans, 5))
   })
 }
 
-test_loglike_model_all <- function(x, p, ls, ln, qn, a) {
-  truth <- -0.48062578001916395
-  ans <- loglike_model_all(x, p, ls, ln, qn, a)
-  test_that("loglike_model_all gives expected results", {
-    expect_equal(truth, ans)
+test_loglike_model_collective <- function(x, p, bs, br, as, ar) {
+  truth <- -2.233698
+  ans <- loglike_model_collective(x, p, bs, br, as, ar)
+  test_that("loglike_model_collective gives expected results", {
+    expect_equal(truth, round(ans, 6))
   })
 }
 
-test_loglike_model_scout <- function(x, ls, qn, a) {
-  truth <- -0.7072428399318523
-  ans <- loglike_model_scout(x, ls, qn, a)
-  test_that("loglike_model_scout gives expected results", {
-    expect_equal(truth, ans)
+test_loglike_model_individual <- function(x, br, ar) {
+  truth <- -0.83449
+  ans <- loglike_model_individual(x, br, ar)
+  test_that("loglike_model_individual gives expected results", {
+    expect_equal(truth, round(ans, 5))
   })
 }
 
@@ -62,20 +62,20 @@ model_tests <- function() {
     0.1, 0.48
   )
   p <- 0.5
-  ls <- 1.3
-  ln <- 1.3
-  qn <- 2.2
-  a <- 0.5
+  bs <- 1.3
+  br <- 1.3
+  as <- 2.2
+  ar <- 0.5
 
   # run through model tests
[…]

### 2021-10-05 tests/testthat/test-fitting.R

@@ -1,4 +1,4 @@
-# generate_bounds
+# get_bounds
 test_check_upper_bound <- function() {
   test_that("check_upper_bound gives error if upper < 0", {
     expect_error(check_upper_bound(-1), "upper cannot be < 0")
@@ -11,40 +11,40 @@ test_check_upper_bound <- function() {
   })
 }
 
-test_generate_bounds_all <- function() {
+test_get_bounds_collective <- function() {
   upper <- 5
   p_bnds <- c(0, 1.0)
-  ls_bnds <- c(1.0e-6, upper)
-  ln_bnds <- c(0, upper)
-  q_bnds <- c(1, upper)
-  a_bnds <- c(0, upper)
+  bs_bnds <- c(1.0e-6, 10)
+  br_bnds <- c(1.0e-6, 10)
+  as_bnds <- c(1.0e-12, upper)
+  ar_bnds <- c(1.0e-12, upper)
   actual_bounds <- rbind(
-    p_bnds, ls_bnds,
-    ln_bnds, q_bnds,
-    a_bnds
+    p_bnds, bs_bnds,
+    br_bnds, as_bnds,
+    ar_bnds
   )
-  test_that("generate_bounds_all generates the correct bounds", {
-    expect_identical(actual_bounds, generate_bounds_all(upper))
+  test_that("get_bounds_collective gets the correct bounds", {
+    expect_identical(actual_bounds, get_bounds_collective(upper))
   })
 }
 
-test_generate_bounds_scout <- function() {
+test_get_bounds_individual <- function() {
   upper <- 5
-  ls_bnds <- c(1.0e-6, upper)
-  q_bnds <- c(1, upper)
-  a_bnds <- c(0, upper)
+  bs_bnds <- c(1.0e-6, upper)
+  as_bnds <- c(1.0e-6, upper)
   actual_bounds <- rbind(
-    ls_bnds, q_bnds, a_bnds
+    bs_bnds,
+    as_bnds
   )
-  test_that("generate_bounds_all generates the correct bounds", {
-    expect_identical(actual_bounds, generate_bounds_scout(upper))
+  test_that("get_bounds_collective gets the correct bounds", {
+    expect_identical(actual_bounds, get_bounds_individual(upper))
   })
 }
 
-test_generate_starting_est_all <- function(x) {
-  bnds <- generate_bounds_all(5)
-  result <- generate_starting_ests_all(x, bounds = bnds)
[…]

### 2021-10-05 tests/testthat/test-plotting.R

@@ -20,7 +20,7 @@ test_make_ccdf_plot_data <- function(x, param_est, model, npoints) {
       0.100, 0.189, 0.278, 0.367, 0.456, 0.544, 0.633, 0.722, 0.811, 0.900
     ),
     cumul_ccdf = c(
-      1.000, 0.848, 0.687, 0.534, 0.402, 0.297, 0.219, 0.164, 0.125, 0.097
+      1.000, 0.713, 0.534, 0.425, 0.362, 0.314, 0.267, 0.221, 0.178, 0.140
     )
   )
   actual <- round(make_ccdf_plot_data(x, param_est, model, npoints), 3)
@@ -54,13 +54,13 @@ test_make_full_plot <- function(x, model_result_list) {
 test_make_results_tibble <- function(result) {
   expected <- tibble::tibble(
     fmax = c(-256, -256),
-    data_name = c("all", "scout"),
+    data_name = c("collective", "individual"),
     p = c(0.5, 1),
-    ls = c(1.3, 1.3),
-    ln = c(1.3, NA),
-    qn = c(2.2, 2.2),
-    a = c(0.5, 0.5),
-    AIC = c(522, 518)
+    bs = c(1.3, 1.3),
+    br = c(1.3, NA),
+    as = c(2.2, 2.2),
+    ar = c(0.5, NA),
+    AIC = c(522, 516)
   )
   actual <- make_results_tibble(result)
   test_that("make_results_tibble returns the expected result", {
@@ -77,26 +77,26 @@ plotting_tests <- function() {
     0.1, 0.48
   )
   p <- 0.5
-  ls <- 1.3
-  ln <- 1.3
-  qn <- 2.2
-  a <- 0.5
-  param_est <- c(p, ls, ln, qn, a)
+  bs <- 1.3
+  br <- 1.3
+  as <- 2.2
+  ar <- 0.5
+  param_est <- c(p, bs, br, as, ar)
   example_result <- list(
-    "all" = list(
+    "collective" = list(
       est = param_est,
       fmax = -256,
-      data_name = "all"
+      data_name = "collective"
     ),
-    "scout" = list(
-      est = c(ls, qn, a),
+    "individual" = list(
+      est = c(bs, as),
       fmax = -256,
-      data_name = "scout"
+      data_name = "individual"
     )
   )
 
   test_inverse_cdf(x)
-  test_make_ccdf_plot_data(x, param_est, "all", 10)
[…]

### 2021-10-05 tests/testthat/test-utils.R

@@ -38,7 +38,7 @@ test_that("message_verbose does not return a message when asked", {
 test_that("model_number_from_model returns correct models", {
   expected <- c(0, 1)
   results <- purrr::map_dbl(
-    c("all", "scout"),
+    c("collective", "individual"),
     model_number_from_model
   )
   expect_identical(expected, results)
@@ -57,7 +57,31 @@ test_that("calc_aic works as expected", {
   expect_equal(result, expected)
 })
 
-test_that("calc_ks works as expected", {
-  # add code to test the calc_ks function. Write in once bootstrapped.
-  expect_equal(1, 1)
+expect_collective_dstat <- 0.4
+expect_individual_dstat <- 0.7
+x <- c(
+  0.2, 0.5,
+  0.3, 0.6,
+  0.8, 0.5,
+  0.9, 0.6,
+  0.1, 0.48
+)
+p <- 0.5
+bs <- 1.3
+br <- 1.3
+as <- 2.2
+ar <- 0.5
+param_est_collective <- c(p, bs, br, as, ar)
+param_est_individual <- c(bs, as)
+collective_result <- calc_ks_boot(x, param_est_collective, "collective")
+individual_result <- calc_ks_boot(x, param_est_individual, "individual")
+test_that("calc_ks_boot collective gives correct D stat", {
+  expect_equal(
+    as.numeric(collective_result$ks$statistic), expect_collective_dstat
+  )
+})
+test_that("calc_ks_boot individual gives correct D stat", {
+  expect_equal(
+    as.numeric(individual_result$ks$statistic), expect_individual_dstat
+  )
 })

### 2021-10-05 analysis/fit_models.Rmd
removed shapefiles not used

@@
 ---
+
+This document contains code to fit the waggle dance model to each of our 20 different sites.
+The code here uses the wagglefit package, as well as some additional code stored in `fit_data.R` to help simplify things here.
 
@@
 london <- st_read(
-  "analysis/shapefiles/statistical-gis-boundaries-london/ESRI/London_Ward.shp"
+  "analysis/shapefiles/London_Ward.shp"
 )

### 2021-10-05 analysis/fit_data.R
removed analysis to keep in seperate branch, this allows dev to be just package code and free of specific analysis code.

@@ -1,249 +1,0 @@
-suppressMessages(library(dplyr))
-library(magrittr)
-library(tibble)
-suppressMessages(library(lubridate))
-library(wagglefit)
-suppressMessages(library(rlang))
-library(ggplot2)
-suppressMessages(library(tidyr))
-devtools::load_all()
-
-#' Get waggle dance data from file and calculate foraging distance
-#' @descriptionFor For specific data in the 'data' directory read it in,
-#' calculate foraging distance and get just the required columns
-#' @importFrom rlang .data
-#' @export
-#'
-get_data <- function() {
-  full_data_path <- "analysis/data/FullHBForagingData.csv"
-  data <- tibble(read.csv(full_data_path))
-  data <- data %>%
-    mutate(
-      date = ymd(.data$date),
-      foraging_distance = calc_dist(.data$duration.seconds)
-    ) %>%
-    select(
-      date,
-      site,
-      foraging_distance
-    )
-  return(data)
-}
-
-#' Run my analysis
-run_analysis <- function(
-  data, model = "all", n = 5, upper = 5, bounds = NULL, xtol = 0,
-  verbose_r = FALSE, verbose = FALSE) {
-  best_result <- fit_mutliple(
-    data$foraging_distance,
-    model = model,
-    n = n,
-    upper = upper,
-    bounds = bounds,
-    verbose_r = verbose_r,
-    verbose = verbose,
-    xtol = xtol
-  )
-  best_result$data_name <- model
-  return(best_result)
-}
-
-#' Run on specific site k times and output likelihood space and plots
-fit_collective_model_to_data <- function(data, bounds, k=10, n = 10,
-                                        verbose_r = FALSE, verbose = FALSE) {
-  # run optimisation k times and return best one
-  result_data <- purrr::map_df(
-    seq_len(k),
-    ~ {
-      print(paste("Itteration", .x))
-      result <- run_analysis(
-        data, model = "collective", n = n, upper = 10, bounds = bounds,
-        verbose_r = verbose_r, verbose = verbose
-      )
-      return(
-        data.frame(
-          ll = result$fmax,
-          p = result$est[1],
-          bs = result$est[2],
-          br = result$est[3],
-          as = result$est[4],
-          ar = result$est[5]
-        )
-      )
-    }
-  )
-
[…]

### 2021-10-05 R/utils.R
ran devtools check and all ok

@@ -118,22 +118,18 @@ calc_ks_boot <- function(x, param_est, model_type) {
 #' be the data the parameters were fit to.
 #' @param var characterArray The name of the variable you want to examine the
 #' likelihood space for.
-#' @param p double The proportion of scouts.
-#' @param bs double The scout rate.
-#' @param as double The scout alpha.
-#' @param br double The recruit rate.
-#' @param ar double The recruit alpha.
+#' @param params The parameters of the model to be checked in isolation.
 #' @param n integer The number of points to sample, defaults to 1000.
 #' @param upper integer The upper bound
+#' @param bounds list User supplied list of upper and lower bounds.
 #' @return tibble The varied parameter and associated likelihood and
 #' log-likelihood scores.
 #' @importFrom tibble tibble
 #' @importFrom purrr map_dbl
 #' @concept utility
 #'
-calc_var_likelihood_collective <- function(
-  data, var, params, n = 1000, upper = 5, bounds = NULL
-  ) {
+calc_var_likelihood_collective <- function(data, var, params, n = 1000,
+                                           upper = 5, bounds = NULL) {
   p <- params[[1]]
   bs <- params[[2]]
   br <- params[[3]]
@@ -141,7 +137,7 @@ calc_var_likelihood_collective <- function(
   ar <- params[[5]]
 
   if (is.null(bounds)) {
-    bnds <- get_bounds_all(upper)
+    bnds <- get_bounds_collective(upper)
   } else {
     bnds <- bounds
   }
@@ -176,13 +172,12 @@ calc_var_likelihood_collective <- function(
 #' @importFrom purrr map_dbl
 #' @concept utility
 #'
-calc_var_likelihood_individual <- function(
-  data, var, params, n = 1000, upper = 5, bounds = NULL
-  ) {
+calc_var_likelihood_individual <- function(data, var, params, n = 1000,
+                                           upper = 5, bounds = NULL) {
   bs <- params[[1]]
   as <- params[[2]]
   if (is.null(bounds)) {
-    bnds <- get_bounds_all(upper)
+    bnds <- get_bounds_collective(upper)
   } else {
     bnds <- bounds
   }
@@ -219,7 +214,7 @@ calc_var_likelihood_individual <- function(
 #' @concept utility
 #'
 map_likelihood_space <- function(data, params, model_name = "collective",
-                                n = 1000, upper = 5, bounds = NULL) {
[…]

### 2021-10-05 R/RcppExports.R
Prep for dev (#21)
* added ccdf cpp functions

* added tests for cpp ccdf functions

* added basic analysis overview for manuscript results and figures

* added general plotting functions for ccdf and stats functions to utils, as well as other helpers

* added some tests, more to come

* finished plotting and utils tests

* removed all recruit model code

* added back analysis files removed from merge

* created Rmd document for making manuscript figures

* added new figures, markdown and more plots

* added simulation python code and linked with R

* updated simulation plots to fit exp and mi

@@ -6,58 +6,112 @@
 #' @param x NumericArray
 #' @export
 alter_in_place <- function(x) {
-  invisible(.Call("_wagglefit_alter_in_place", PACKAGE = "wagglefit", x))
+    invisible(.Call('_wagglefit_alter_in_place', PACKAGE = 'wagglefit', x))
+}
+
+#' Model CCDF for scout distribution
+#'
+#' @param x double Foraging distance
+#' @param m double Minimum foraging distance
+#' @param bs double Scout rate
+#' @param as double scout alpha
+#' @return result double the ccdf for that foraging distance (x)
+#' @export
+scout_ccdf <- function(x, m, bs, as) {
+    .Call('_wagglefit_scout_ccdf', PACKAGE = 'wagglefit', x, m, bs, as)
+}
+
+#' Model CCDF for recruit distribution
+#'
+#' @param x double Foraging distance
+#' @param m double Minimum foraging distance
+#' @param br double Recruit rate
+#' @param ar double recruit alpha
+#' @return result double the ccdf for that foraging distance (x)
+#' @export
+recruit_ccdf <- function(x, m, br, ar) {
+    .Call('_wagglefit_recruit_ccdf', PACKAGE = 'wagglefit', x, m, br, ar)
+}
+
+#' Model ccdf function for collective model. Stores results in
+#' given array (y)
+#'
+#' @param x NumericVector foraging distances
+#' @param y NumericVector storage array for the results
+#' @param p double Proportion of scouts (0<=p<=1)
+#' @inheritParams scout_ccdf
+#' @inheritParams recruit_ccdf
+#' @export
+ccdf_model_collective <- function(x, y, p, bs, br, as, ar) {
+    invisible(.Call('_wagglefit_ccdf_model_collective', PACKAGE = 'wagglefit', x, y, p, bs, br, as, ar))
+}
+
+#' Model ccdf function for individual model. Stores results in given array (y)
+#'
+#' @param x NumericVector foraging distances
+#' @param y NumericVector storage array for the results
+#' @inheritParams scout_ccdf
+#' @export
+ccdf_model_individual <- function(x, y, bs, as) {
+    invisible(.Call('_wagglefit_ccdf_model_individual', PACKAGE = 'wagglefit', x, y, bs, as))
+}
+
+#' Get model ccdf for a given model
+#'
+#' @param x NumericVector foraging distances
[…]

### 2021-10-05 R/fitting.R

@@ -20,7 +20,7 @@ check_upper_bound <- function(upper) {
   }
 }
 
-#' Generates bounds for all parameters of the scout recruit superposition model
+#' Generates bounds for the collective model
 #'
 #' @description Generates upper and lower bounds for all parameters
 #' @param upper double The upper bound for the params
@@ -31,21 +31,21 @@ check_upper_bound <- function(upper) {
 #' generate_bounds_all()
 #' }
 #'
-generate_bounds_all <- function(upper) {
+get_bounds_collective <- function(upper) {
   p_bnds <- c(0, 1.0)
-  ls_bnds <- c(1.0e-6, upper)
-  ln_bnds <- c(0, upper)
-  q_bnds <- c(1, upper)
-  a_bnds <- c(0, upper)
+  bs_bnds <- c(1.0e-6, 10)
+  br_bnds <- c(1.0e-6, 10)
+  as_bnds <- c(1.0e-12, upper)
+  ar_bnds <- c(1.0e-12, upper)
   bounds <- rbind(
-    p_bnds, ls_bnds,
-    ln_bnds, q_bnds,
-    a_bnds
+    p_bnds, bs_bnds,
+    br_bnds, as_bnds,
+    ar_bnds
   )
   return(bounds)
 }
 
-#' Generates bounds for all parameters of the scout model
+#' Generates bounds for all parameters of the individual model
 #'
 #' @description Generates upper and lower bounds for all parameters
 #' @param upper double The upper bound for the params
@@ -56,22 +56,20 @@ generate_bounds_all <- function(upper) {
 #' generate_bounds_all()
 #' }
 #'
-generate_bounds_scout <- function(upper) {
+get_bounds_individual <- function(upper) {
   upper <- as.double(upper)
-  ls_bnds <- c(1.0e-6, upper)
-  q_bnds <- c(1, upper)
-  a_bnds <- c(0, upper)
+  bs_bnds <- c(1.0e-6, upper)
+  as_bnds <- c(1.0e-6, upper)
   bounds <- rbind(
-    ls_bnds,
-    q_bnds,
-    a_bnds
+    bs_bnds,
+    as_bnds
   )
   return(bounds)
 }
 
 
 #' Generates starting estimates for all numerical optimisation of all parameters
-#' of the scout recruit superposition model
+#' of the collective model
 #'
 #' @description Generates starting estimates for all parameters by taking a
 #' random value sampled from a normal distribution truncated at the upper and
[…]

### 2021-10-05 R/plotting.R

@@ -1,0 +1,193 @@
+
+#' Calculates the ccdf for given data
+#'
+#' @description Orders the data smallest to largest and calculates the
+#' probability of sampling a value greater than or equal to a given value.
+#' @param x The data, e.g. foraging distance
+#' @return tibble of the sorted foraging distances and their cumulative
+#' probability
+#' @importFrom purrr map_dbl
+#' @importFrom tibble tibble
+#' @export
+#'
+inverse_ccdf <- function(x) {
+  if (!is.double(x) | length(x) < 2) {
+    stop("x must be a double array")
+  }
+  sd <- -sort(-x)
+  prob <- map_dbl(
+    seq_len(length(x)),
+    ~ {
+      .x / length(x)
+    }
+  )
+  return(tibble(sd, prob))
+}
+
+#' Creates data for plots
+#'
+#' @description Using the optimised paramater estimates, creates the data for
+#' the ccdf plots
+#' @param x doubleArray The foraging distance to plot
+#' @param param_est doubleArray The paramater estimates of the optimisation
+#' @param model characterArray The name of the model used
+#' @param npoints integer The number of points to plot the predicted data for.
+#' Defaults to 100.
+#' @return tibble of model data and cumulative probability
+#' @importFrom tibble tibble
+#' @export
+#'
+make_ccdf_plot_data <- function(x, param_est, model, npoints = 100) {
+  model_n <- model_number_from_model(model)
+  x_seq <- seq(min(x), max(x), length.out = npoints)
+  cumul_ccdf <- model_ccdf(x_seq, param_est, model_n)
+  cumul_ccdf[[1]] <- 1
+  cumul <- tibble(x_seq, cumul_ccdf)
+  return(cumul)
+}
+
+#' Creates ccdf plot of data
+#'
+#' @description Creates a ccdf plot for the data provided
+#' @param x doubleArray The foraging distance to plot
+#' @param logit Bool to log the probabilities or not.
+#' @return ggplot plot of cumulative probability for foraging distances in Km
+#' @importFrom ggplot2 ggplot aes_string geom_point geom_line theme_set theme
+#' theme_classic element_text labs
+#' @importFrom tibble tibble
+#' @importFrom magrittr %>%
+#' @importFrom rlang .data
+#' @export
+#'
+make_base_plot <- function(x, logit = TRUE) {
+  theme_set(
+    theme_classic() +
+      theme(
[…]

### 2021-10-05 R/utils.R

@@ -1,3 +1,20 @@
+#' Get model number from model name
+#'
+#' @description Converts the model name to its number for cpp. E.g. 'all' is 0
+#' @param model characterArray The model to get the number for
+#' @export
+#' @concept utility
+model_number_from_model <- function(model) {
+  model_list <- list(
+    "collective" = 0,
+    "individual" = 1
+  )
+  if (!(model %in% names(model_list))) {
+    stop("Model name not found")
+  }
+  return(model_list[[model]])
+}
+
 #' Wrapper for message
 #'
 #' @description A wrapper for `message` that only prints output when
@@ -25,7 +42,7 @@ message_verbose <- function(verbose = TRUE, ...) {
 #' \dontrun{
 #' calc_dist(1.8)
 #' }
-#'
+#' @concept utility
 calc_dist <- function(duration) {
   intercept <- 0.17
   slope <- 1.38
@@ -47,7 +64,7 @@ calc_dist <- function(duration) {
 #' \dontrun{
 #' trunc_normal(0, 0.5, -1, 1)
 #' }
-#'
+#' @concept utility
 trunc_normal <- function(mean, sd, lwr, upr) {
   continue <- TRUE
   while (continue) {
@@ -57,4 +74,184 @@ trunc_normal <- function(mean, sd, lwr, upr) {
     }
   }
   return(x)
+}
+
+#' Calculate AIC values for a model
+#'
+#' @description Calulcates the AIC values for a given model
+#' @param k integer The number of parameters
+#' @param ll double The maximum log-likelihood (MLE) of the fitted model
+#' @export
+#' @concept utility
+calc_aic <- function(k, ll) {
+  return(2 * k - 2 * ll)
+}
+
+#' Run a bootstapped KS test on the data
+#'
+#' @description Calulctate the ks test stat and p value for a given model
+#' @param x doubleArray Foraging distance data
+#' @param param_est doubleArray The parameter estimates for a model
+#' @param model_type characterArray the name of the model ran
+#' @importFrom stats ks.test
+#' @importFrom rlang .data
+#' @importFrom Matching ks.boot
+#' @concept utility
+calc_ks_boot <- function(x, param_est, model_type) {
+  ccdf_data <- inverse_ccdf(x)
+  ccdf_model <- model_ccdf(x, param_est, model_number_from_model(model_type))
+  if (all(is.nan(ccdf_model))) {
[…]

### 2021-10-05 analysis/fit_data.R

@@ -1,40 +1,0 @@
-library(dplyr)
-library(magrittr)
-library(tibble)
-library(lubridate)
-library(wagglefit)
-
-#' Get waggle dance data from file and calculate foraging distance
-#' @descriptionFor For specific data in the 'data' directory read it in,
-#' calculate foraging distance and get just the required columns
-#' @importFrom rlang .data
-#' @export
-#'
-get_data <- function() {
-  full_data_path <- "analysis/data/FullHBForagingData.csv"
-  data <- tibble(read.csv(full_data_path))
-  data <- data %>%
-    mutate(
-      date = ymd(.data$date),
-      foraging_distance = calc_dist(.data$duration.seconds)
-    ) %>%
-    select(
-      date,
-      site,
-      foraging_distance
-    )
-  return(data)
-}
-
-#' Run my analysis
-run_analysis <- function(model = "all", n = 5, xtol = 0) {
-  data <- get_data()
-  best_result <- run_mutliple(
-    data$foraging_distance,
-    model = model,
-    n = n,
-    verbose_r = TRUE,
-    xtol = xtol
-  )
-  return(best_result)
-}

### 2021-10-05 src/RcppExports.cpp

@@ -15,85 +15,136 @@ BEGIN_RCPP
     return R_NilValue;
 END_RCPP
 }
-// scout_dist
-double scout_dist(double x, double m, double p, double ls, double qn, double a);
-RcppExport SEXP _wagglefit_scout_dist(SEXP xSEXP, SEXP mSEXP, SEXP pSEXP, SEXP lsSEXP, SEXP qnSEXP, SEXP aSEXP) {
+// scout_ccdf
+double scout_ccdf(double x, double m, double bs, double as);
+RcppExport SEXP _wagglefit_scout_ccdf(SEXP xSEXP, SEXP mSEXP, SEXP bsSEXP, SEXP asSEXP) {
 BEGIN_RCPP
     Rcpp::RObject rcpp_result_gen;
     Rcpp::RNGScope rcpp_rngScope_gen;
     Rcpp::traits::input_parameter< double >::type x(xSEXP);
     Rcpp::traits::input_parameter< double >::type m(mSEXP);
+    Rcpp::traits::input_parameter< double >::type bs(bsSEXP);
+    Rcpp::traits::input_parameter< double >::type as(asSEXP);
+    rcpp_result_gen = Rcpp::wrap(scout_ccdf(x, m, bs, as));
+    return rcpp_result_gen;
+END_RCPP
+}
+// recruit_ccdf
+double recruit_ccdf(double x, double m, double br, double ar);
+RcppExport SEXP _wagglefit_recruit_ccdf(SEXP xSEXP, SEXP mSEXP, SEXP brSEXP, SEXP arSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< double >::type x(xSEXP);
+    Rcpp::traits::input_parameter< double >::type m(mSEXP);
+    Rcpp::traits::input_parameter< double >::type br(brSEXP);
+    Rcpp::traits::input_parameter< double >::type ar(arSEXP);
+    rcpp_result_gen = Rcpp::wrap(recruit_ccdf(x, m, br, ar));
+    return rcpp_result_gen;
+END_RCPP
+}
+// ccdf_model_collective
+void ccdf_model_collective(NumericVector x, NumericVector y, double p, double bs, double br, double as, double ar);
[…]

### 2021-10-05 src/models.cpp

@@ -22,6 +22,7 @@ typedef struct {
     double best_est;
 } data_struct;
 
+
 /*
 ------------------------------- helper functions -------------------------------
 */
@@ -38,12 +39,15 @@ void check_user_input(int fcount) {
 // check optimise exit status and display important info about them to user
 //
 // @param optimise_result Integer The exit code from the nlopt optimisation
-void check_optimise_result(int optimise_result) {
-    if (optimise_result < 0) {
-        Rcout << "NLOPT FAIL! Error code: " << optimise_result << std::endl;
-        if (optimise_result == -4) {
-            Rcout << "Fail is due to round off errors"
-                "- result may still be usefull" << std::endl;
+// @param verbose Bool To display messages or not (defaults to TRUE)
+void check_optimise_result(int optimise_result, bool verbose = true) {
+    if (verbose) {
+        if (optimise_result < 0) {
+            Rcout << "NLOPT FAIL! Error code: " << optimise_result << std::endl;
+            if (optimise_result == -4) {
+                Rcout << "Fail is due to round off errors"
+                    "- result may still be usefull" << std::endl;
+            }
         }
     }
 }
@@ -56,53 +60,42 @@ void check_optimise_result(int optimise_result) {
 //'
 //' @param x double Foraging distance
 //' @param m double Minimum foraging distance
-//' @param p double Proportion of scouts (0<=p<=1)
-//' @param ls double Scout rate
-//' @param qn double Quality
-//' @param a double alpha value
+//' @param bs double scout rate
+//' @param as double scout alpha
 //' @export
 // [[Rcpp::export]]
-double scout_dist(double x, double m,
-                  double p, double ls,
-                  double qn, double a)
+double scout_dist(double x, double m, double bs, double as)
 {
-    double maxpart = (-1+qn-x*a)/a;
+    double maxpart = 1-as*x;
     if (maxpart < 0)
     {
         maxpart = 0;
     }
     double result =
-    p * (
-        (exp(((-1 + qn + m * a - x * a) * ls) / a) * a * pow(ls, 2) * maxpart) /
-        (exp(m * ls) * a - exp(((-1 + qn) * ls) / a) *
-            (a + ls - qn * ls + m * a * ls))
-    );
+        as * (
+            (bs * exp(-bs*as*(x-m))*maxpart) /
+            ((1-as*m) - pow(bs, -1)*(1-exp(-bs*(1-as*m))))
+        );
     return result;
 }
 
 //' Model function for recruits
 //'
[…]

### 2021-10-05 tests/testthat/test-cpp_ccdfs.R

@@ -1,0 +1,91 @@
+test_scout_ccdf <- function(x, m, bs, as) {
+  expected <- c(
+    0.4204785, 0.0000000,
+    0.1269553, 0.0000000,
+    0.0000000, 0.0000000,
+    0.0000000, 0.0000000,
+    1.0000000, 0.0000000
+  )
+  actual <- purrr::map_dbl(
+    x,
+    ~ {
+      scout_ccdf(.x, m, bs, as)
+    }
+  )
+  test_that("scout ccdf returns expected results", {
+    expect_identical(round(expected, 5), round(actual, 5))
+  })
+}
+
+test_recruit_ccdf <- function(x, m, br, ar) {
+  expected <- c(
+    0.9502439, 0.6768375,
+    0.8748425, 0.5690817,
+    0.3669045, 0.6768375,
+    0.2807545, 0.5690817,
+    1.0000000, 0.6981969
+  )
+  actual <- purrr::map_dbl(
+    x,
+    ~ {
+      recruit_ccdf(.x, m, br, ar)
+    }
+  )
+  test_that("scout ccdf returns expected results", {
+    expect_identical(round(expected, 5), round(actual, 5))
+  })
+}
+
+test_ccdf_model_collective <- function(x, p, bs, br, as, ar) {
+  expected <- c(
+    0.6853612, 0.3384187,
+    0.5008989, 0.2845408,
+    0.1834523, 0.3384187,
+    0.1403772, 0.2845408,
+    1.0000000, 0.3490985
+  )
+  result <- rep(0.0, length(x))
+  ccdf_model_collective(x, result, p, bs, br, as, ar)
+  test_that("ccdf_model_collective returns the expected result", {
+    expect_identical(round(expected, 5), round(result, 5))
+  })
+}
+
+test_ccdf_model_individual <- function(x, bs, as) {
+  expected <- c(
+    0.4204785, 0.0000000,
+    0.1269553, 0.0000000,
+    0.0000000, 0.0000000,
+    0.0000000, 0.0000000,
+    1.0000000, 0.0000000
+  )
+  result <- rep(0.0, length(x))
+  ccdf_model_individual(x, result, bs, as)
+  test_that("ccdf_model_individual returns the expected result", {
+    expect_identical(round(expected, 5), round(result, 5))
+  })
+}
+
[…]

### 2021-10-05 tests/testthat/test-cpp_models.R

@@ -1,40 +1,32 @@
-test_scout_dist <- function(x, m, p, ls, qn, a) {
-  truth <- 0.8000732527680952
-  ans <- scout_dist(x, m, p, ls, qn, a)
+test_scout_dist <- function(x, m, bs, as) {
+  truth <- 4.15163
+  ans <- scout_dist(x, m, bs, as)
   test_that("scout_dist gives expected result", {
-    expect_equal(truth, ans)
+    expect_equal(truth, round(ans, 5))
   })
 }
 
-test_recruit_dist <- function(x, m, p, ln, qn, a) {
-  truth <- 1.0095546946431222
-  ans <- recruit_dist(x, m, p, ln, qn, a)
+test_recruit_dist <- function(x, m, br, ar) {
+  truth <- 0.63748
+  ans <- recruit_dist(x, m, br, ar)
   test_that("recruit_dist gives expected results", {
     expect_identical(round(truth, 5), round(ans, 5))
   })
 }
 
-test_loglike_model_all <- function(x, p, ls, ln, qn, a) {
-  truth <- -0.48062578001916395
-  ans <- loglike_model_all(x, p, ls, ln, qn, a)
-  test_that("loglike_model_all gives expected results", {
-    expect_equal(truth, ans)
+test_loglike_model_collective <- function(x, p, bs, br, as, ar) {
+  truth <- -2.233698
+  ans <- loglike_model_collective(x, p, bs, br, as, ar)
+  test_that("loglike_model_collective gives expected results", {
+    expect_equal(truth, round(ans, 6))
   })
 }
 
-test_loglike_model_scout <- function(x, ls, qn, a) {
-  truth <- -0.7072428399318523
-  ans <- loglike_model_scout(x, ls, qn, a)
-  test_that("loglike_model_all gives expected results", {
-    expect_equal(truth, ans)
-  })
-}
-
-test_loglike_model_recruit <- function(x, ln, qn, a) {
-  truth <- -1.3118270596288413
-  ans <- loglike_model_recruit(x, ln, qn, a)
-  test_that("loglike_model_all gives expected results", {
-    expect_equal(truth, ans)
+test_loglike_model_individual <- function(x, br, ar) {
+  truth <- -0.83449
+  ans <- loglike_model_individual(x, br, ar)
+  test_that("loglike_model_individual gives expected results", {
[…]

### 2021-10-05 tests/testthat/test-fitting.R

@@ -1,4 +1,4 @@
-# generate_bounds
+# get_bounds
 test_check_upper_bound <- function() {
   test_that("check_upper_bound gives error if upper < 0", {
     expect_error(check_upper_bound(-1), "upper cannot be < 0")
@@ -11,40 +11,40 @@ test_check_upper_bound <- function() {
   })
 }
 
-test_generate_bounds_all <- function() {
+test_get_bounds_collective <- function() {
   upper <- 5
   p_bnds <- c(0, 1.0)
-  ls_bnds <- c(1.0e-6, upper)
-  ln_bnds <- c(0, upper)
-  q_bnds <- c(1, upper)
-  a_bnds <- c(0, upper)
+  bs_bnds <- c(1.0e-6, 10)
+  br_bnds <- c(1.0e-6, 10)
+  as_bnds <- c(1.0e-12, upper)
+  ar_bnds <- c(1.0e-12, upper)
   actual_bounds <- rbind(
-    p_bnds, ls_bnds,
-    ln_bnds, q_bnds,
-    a_bnds
+    p_bnds, bs_bnds,
+    br_bnds, as_bnds,
+    ar_bnds
   )
-  test_that("generate_bounds_all generates the correct bounds", {
-    expect_identical(actual_bounds, generate_bounds_all(upper))
+  test_that("get_bounds_collective gets the correct bounds", {
+    expect_identical(actual_bounds, get_bounds_collective(upper))
   })
 }
 
-test_generate_bounds_scout <- function() {
+test_get_bounds_individual <- function() {
   upper <- 5
-  ls_bnds <- c(1.0e-6, upper)
-  q_bnds <- c(1, upper)
-  a_bnds <- c(0, upper)
+  bs_bnds <- c(1.0e-6, upper)
+  as_bnds <- c(1.0e-6, upper)
   actual_bounds <- rbind(
-    ls_bnds, q_bnds, a_bnds
+    bs_bnds,
+    as_bnds
   )
-  test_that("generate_bounds_all generates the correct bounds", {
-    expect_identical(actual_bounds, generate_bounds_scout(upper))
+  test_that("get_bounds_collective gets the correct bounds", {
+    expect_identical(actual_bounds, get_bounds_individual(upper))
   })
 }
 
-test_generate_starting_est_all <- function(x) {
-  bnds <- generate_bounds_all(5)
-  result <- generate_starting_ests_all(x, bounds = bnds)
[…]

### 2021-10-05 tests/testthat/test-plotting.R

@@ -1,0 +1,107 @@
+test_inverse_cdf <- function(x) {
+  test_that("inverse_cdf works as expected", {
+    expected <- tibble::tibble(
+      sd = c(0.90, 0.80, 0.60, 0.60, 0.50, 0.50, 0.48, 0.30, 0.20, 0.10),
+      prob = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0)
+    )
+    expect_identical(round(inverse_ccdf(x), 3), expected)
+  })
+
+  test_that("inverse_cdf fails if input is not a double array", {
+    x <- tibble::tibble(a = c(0, 4, 2, 6, 8, 3), b = c(0, 5, 1, 8, 3, 9))
+    expect_error(inverse_ccdf(x), "x must be a double array")
+    expect_error(inverse_ccdf(0.5), "x must be a double array")
+  })
+}
+
+test_make_ccdf_plot_data <- function(x, param_est, model, npoints) {
+  expected <- tibble::tibble(
+    x_seq = c(
+      0.100, 0.189, 0.278, 0.367, 0.456, 0.544, 0.633, 0.722, 0.811, 0.900
+    ),
+    cumul_ccdf = c(
+      1.000, 0.713, 0.534, 0.425, 0.362, 0.314, 0.267, 0.221, 0.178, 0.140
+    )
+  )
+  actual <- round(make_ccdf_plot_data(x, param_est, model, npoints), 3)
+  test_that("make_ccdf_plot_data returns expected results", {
+    expect_identical(expected, actual)
+  })
+}
+
+test_make_base_plot <- function(x) {
+  plt <- make_base_plot(x)
+  expected_xlab <- "Foraging distance (Km)"
+  expected_ylab <- "Ln cumulative probability"
+  test_that("make_base_plot returns a ggplot object with the correct labels", {
+    expect_true(ggplot2::is.ggplot(plt))
+    expect_equal(expected_xlab, plt$labels$x)
+    expect_equal(expected_ylab, plt$labels$y)
+  })
+}
+
+test_make_full_plot <- function(x, model_result_list) {
+  plt <- make_full_plot(x, model_result_list)
+  dev.off()
+  expected_xlab <- "Foraging distance (Km)"
+  expected_ylab <- "Ln cumulative probability"
+  test_that("make_full_plot returns a ggplot object with the correct labels", {
[…]

### 2021-10-05 tests/testthat/test-utils.R

@@ -34,3 +34,54 @@ test_that("message_verbose does not return a message when asked", {
   )
   expect_equal(l, character(0))
 })
+
+test_that("model_number_from_model returns correct models", {
+  expected <- c(0, 1)
+  results <- purrr::map_dbl(
+    c("collective", "individual"),
+    model_number_from_model
+  )
+  expect_identical(expected, results)
+})
+
+test_that("model_number_from_model returns error if model name is not known", {
+  expect_error(
+    model_number_from_model("amadeupmodel"),
+    "Model name not found"
+  )
+})
+
+test_that("calc_aic works as expected", {
+  expected <- 314
+  result <- calc_aic(2, -155)
+  expect_equal(result, expected)
+})
+
+expect_collective_dstat <- 0.4
+expect_individual_dstat <- 0.7
+x <- c(
+  0.2, 0.5,
+  0.3, 0.6,
+  0.8, 0.5,
+  0.9, 0.6,
+  0.1, 0.48
+)
+p <- 0.5
+bs <- 1.3
+br <- 1.3
+as <- 2.2
+ar <- 0.5
+param_est_collective <- c(p, bs, br, as, ar)
+param_est_individual <- c(bs, as)
+collective_result <- calc_ks_boot(x, param_est_collective, "collective")
+individual_result <- calc_ks_boot(x, param_est_individual, "individual")
+test_that("calc_ks_boot collective gives correct D stat", {
+  expect_equal(
+    as.numeric(collective_result$ks$statistic), expect_collective_dstat
+  )
+})
+test_that("calc_ks_boot individual gives correct D stat", {
+  expect_equal(
+    as.numeric(individual_result$ks$statistic), expect_individual_dstat
+  )
+})

### 2021-10-05 analysis/fit_models.Rmd
moved analysis files from their working branches into this one

@@
 
-run_wagglefit_analysis <- function(
-  target_site, data, collective_bounds, individual_bounds, subplot_coords
-  ) {
+run_wagglefit_analysis <- function(target_site, data, collective_bounds, individual_bounds, subplot_coords) {
 
@@
   pull()
-
 ```
@@
 all_sites[[target_site]]$fit
-
 ```
@@
 )
-
 ```
@@
     data = worldmap,
-    fill = 'grey90',
-    color = '#4b4949d0'
+    fill = "grey90",
+    color = "#4b4949d0"
   ) +
@@
     data = london,
-    fill = '#ced0cffc',
+    fill = "#ced0cffc",
     lwd = 0
@@
   theme_nothing() +
-  scale_x_continuous(expand = c(0,0)) +
-  scale_y_continuous(expand = c(0,0)) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
   labs(x = NULL, y = NULL) +
@@
   geom_sf(
-    fill = '#c4cfc8',
-    color = '#4b4949d0',
+    fill = "#c4cfc8",
+    color = "#4b4949d0",
     lwd = 0.2
@@
 )
-
-```
+```

### 2021-10-05 analysis/Simulation/Code/Agents.py

@@ -1,0 +1,245 @@
+#!/usr/bin/env python3
+"""Classes of agents (workers, scouts, flowers)"""
+__appname__  = "Agents.py"
+__author__   = "Joseph Palmer <joseph.palmer18@imperial.ac.uk>"
+__version__  = "0.0.1"
+__date__     = "12-2019"
+
+import numpy as np
+import matplotlib.pyplot as plt
+import Methods
+import config
+import copy
+from shapely.geometry import Point, Polygon
+import shapely
+
+class Flower:
+    # Attributes #
+    location = (None, None)
+    location_point = None
+    distance = None
+    resource_level = None
+    exists = 0
+
+    # Methods #
+    def __init__(self, location, quality_array):
+        self.location = location
+        self.location_point = Point(*location)
+        self.distance = Methods.getdist(*location)
+        self.resource_level = 2000 #np.random.randint(10, 5000)
+        self.quality_array = quality_array[np.random.randint(0, len(quality_array))]
+        self.exists = 1
+        return None
+
+    def remove_flower(self):
+        self.location = (None, None)
+        self.distance = None
+        self.resource_level = None
+        self.exists = 0
+        return None
+
+    def visit(self):
+        if self.resource_level > 1:
+            self.resource_level -= 1
+        else:
+            self.remove_flower()
+        return None
+
+class Honeybee:
+    flower = None
+    distance = None
+    mode = None # -1 is scout, 1 is employed recruit, 0 is unemployed recruit
+
+class Scout(Honeybee):
+    def __init__(self):
+        self.mode = -1
+        return None
+
+    def _drawpath(self, s1, s2, flowers, infa=[], i=0, limmit=10, plot=True):
+        angle = np.random.uniform(0, 360)
+        distance = np.random.uniform(0., .25)
+        base_fa = [
+            tuple(s1),
+            (s1[0], distance+s1[1]),
+            (s2[0], distance+s2[1]),
+            tuple(s2)
+        ]
+        fa = Polygon(base_fa)
+        fa = shapely.affinity.rotate(fa, angle, origin=s1)
+        if plot:
+            x,y = fa.exterior.xy
+            plt.plot(x,y, color="red")
+        a, b = list(fa.exterior.coords)[1:3]
+        infa_idx = [fa.contains(j.location_point) for j in flowers]
+        [infa.append(ii) for ii in np.array([j.distance for j in flowers])[infa_idx]]
+        if i < limmit:
[…]

### 2021-10-05 analysis/Simulation/Code/run.py

@@ -1,0 +1,193 @@
+#!/usr/bin/env python3
+"""Run simulation of honeybee foraging"""
+__appname__  = "run.py"
+__author__   = "Joseph Palmer <joseph.palmer18@imperial.ac.uk>"
+__version__  = "0.0.2"
+__date__     = "02-2020"
+
+import numpy as np
+import pandas as pd
+import scipy.stats
+import matplotlib.pyplot as plt
+import Agents
+import Methods
+import config
+
+def create_flowers(lwr, upr, res, quality_array):
+    # Simulation window parameters
+    r = upr-lwr
+    xx0 = 0
+    yy0 = 0
+
+    areaTotal = np.pi*r**2
+
+    # Point process parameters
+    lambda0 = res
+
+    # Simulate Poisson point process
+    numbPoints = scipy.stats.poisson(lambda0*areaTotal).rvs()
+    theta = 2*np.pi*scipy.stats.uniform.rvs(0,1,((numbPoints,1)))
+    rho = r*np.sqrt(scipy.stats.uniform.rvs(0,1,((numbPoints,1))))
+
+    # Convert from polar to Cartesian coordinates
+    xx = rho * np.cos(theta)
+    yy = rho * np.sin(theta)
+
+    # Shift centre of disk to (xx0,yy0)
+    xx=xx+xx0
+    yy=yy+yy0
+
+    # Set locations as flowers
+    locations = np.column_stack((xx, yy))
+    flowers = np.apply_along_axis(Agents.Flower, 1, locations, quality_array)
+    fdict = {
+        i : flowers[i] for i in range(res)
+    }
+    return fdict
+
+def initialize(lwr, upr, res, nbees, nscouts, quality_array):
+    # create flowers in environment and store globaly.
+    config.FLOWERDICT = create_flowers(lwr, upr, res, quality_array)
+
+    # create honeybee population
+    nrecruits = nbees-nscouts
+    scouts = np.array([Agents.Scout() for i in range(nscouts)])
+    recruits = np.array([Agents.Recruit() for i in range(nrecruits)])
+    return (scouts, recruits)
+
+def flower_turnover(n, lwr, upr, k, quality_array):
+    # select n flowers and remove
+    assert (n <= len(config.FLOWERDICT.keys())), "N > flowers in environment!"
+    rmv_idx = np.random.choice(list(config.FLOWERDICT.keys()), n, replace=False)
[…]

### 2021-10-05 analysis/fit_data.R

@@ -1,0 +1,252 @@
+suppressMessages(library(dplyr))
+library(magrittr)
+library(tibble)
+suppressMessages(library(lubridate))
+library(wagglefit)
+suppressMessages(library(rlang))
+library(ggplot2)
+suppressMessages(library(tidyr))
+devtools::load_all()
+
+#' Get waggle dance data from file and calculate foraging distance
+#' @descriptionFor For specific data in the 'data' directory read it in,
+#' calculate foraging distance and get just the required columns
+#' @importFrom rlang .data
+#' @export
+#'
+get_data <- function() {
+  full_data_path <- "analysis/data/FullHBForagingData.csv"
+  data <- tibble(read.csv(full_data_path))
+  data <- data %>%
+    mutate(
+      date = ymd(.data$date),
+      foraging_distance = calc_dist(.data$duration.seconds)
+    ) %>%
+    select(
+      date,
+      site,
+      foraging_distance
+    )
+  return(data)
+}
+
+#' Run my analysis
+run_analysis <- function(data, model = "all", n = 5, upper = 5, bounds = NULL,
+                         xtol = 0, verbose_r = FALSE, verbose = FALSE) {
+  best_result <- fit_mutliple(
+    data$foraging_distance,
+    model = model,
+    n = n,
+    upper = upper,
+    bounds = bounds,
+    verbose_r = verbose_r,
+    verbose = verbose,
+    xtol = xtol
+  )
+  best_result$data_name <- model
+  return(best_result)
+}
+
+#' Run on specific site k times and output likelihood space and plots
+fit_collective_model_to_data <- function(data, bounds, k = 10, n = 10,
+                                         verbose_r = FALSE, verbose = FALSE) {
+  # run optimisation k times and return best one
+  result_data <- purrr::map_df(
+    seq_len(k),
+    ~ {
+      print(paste("Itteration", .x))
+      result <- run_analysis(
+        data,
+        model = "collective", n = n, upper = 10, bounds = bounds,
+        verbose_r = verbose_r, verbose = verbose
+      )
+      return(
+        data.frame(
+          ll = result$fmax,
+          p = result$est[1],
+          bs = result$est[2],
+          br = result$est[3],
+          as = result$est[4],
+          ar = result$est[5]
+        )
+      )
+    }
+  )
+
[…]

### 2021-10-05 analysis/analyse_proportion_of_scouts.Rmd
reorganised the code structures and stated exploring how the proportion of scouts varies accross landscapes.

# Analysing the proportion of scouts estimated in different environments
## Joseph Palmer
---

```{r, include = FALSE}
devtools::load_all()
library(showtext)
showtext_auto()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(size = 8)
    )
)
```

First we load in the data and make the tables required. I.e. site land type information along with the estimates proportion of scouts from the model fits in `fit_models.Rmd`.

```{r}
# load results of model fit (code to generate found in fit_models.Rmd)
sitedata <- readRDS("analysis/results/site_fit_results.Rda")

# filter for best model
sitedata <- sitedata %>%
  group_by(site) %>%
  slice(which.min(AIC))

# Load land use data taken (taken from Samuelson et al. 2021)
urban_lu <- read.csv(file = "analysis/data/urban-landuse.csv")
rural_lu <- read.csv(file = "analysis/data/agri-rural-landuse.csv")

# put proportion of scouts into landuse data
urban_lu_long <- sitedata %>%
  select(site, p) %>%
  inner_join(urban_lu) %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")

rural_lu_long <- sitedata %>%
  select(site, p) %>%
  inner_join(rural_lu) %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
```

Now we have everything in, lests do a basic plotting to see what the relationships between each landuse type and the proportion of scouts looks like.

For rural we can see that for most catagories there isn't much of a relationship, however, from a basic linear model the proportion of scouts seems to increase with arable and decrease with more built up areas.

```{r}

rural_lu_long %>%
  ggplot(aes(x = p, y = values)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~landuse, ncol = 2) +
  theme(
    text = element_text(size = 8)
  )
```

for urban we don't really see much at all. There is a slight increase the proportion of scouts in a continous central landscape, and a decrese in sparse residential areas. Everything else seems rather flat.

```{r}

urban_lu_long %>%
  ggplot(aes(x = p, y = values)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~landuse, ncol = 2) +
  theme(
    text = element_text(size = 8)
  )
```

### 2021-10-05 analysis/fit_models.Rmd
reorganised the code structures and stated exploring how the proportion of scouts varies accross landscapes.

@@
 
+# save results
+saveRDS(df, file = "analysis/results/site_fit_results.Rda")
+
 # AIC plot
@@
   plot = aic_plot,
-  filename = "analysis/figures/AIC_plot.png",
+  filename = "analysis/results/figures/AIC_plot.png",
   width = 90,
@@
   plot = ks_plot_dist,
-  filename = "analysis/figures/sites_ks.png",
+  filename = "analysis/results/figures/sites_ks.png",
   width = 86,
@@
   plot = all_sites$STU$fit,
-  filename = "analysis/figures/STU.png",
+  filename = "analysis/results/figures/STU.png",
   width = 90,
@@
   plot = all_sites$ZSL$fit,
-  filename = "analysis/figures/ZSL.png",
+  filename = "analysis/results/figures/ZSL.png",
   width = 90,
@@
   plot = merge_plot,
-  filename = "analysis/figures/site_map.png",
+  filename = "analysis/results/figures/site_map.png",
   width = 90,
@@
   plot = sites_model_plot,
-  filename = "analysis/figures/sites_model_plot.png",
+  filename = "analysis/results/figures/sites_model_plot.png",
   width = 183,
@@
   plot = merge_plot,
-  filename = "analysis/figures/site_map.png",
+  filename = "analysis/results/figures/site_map.png",
   width = 183,
@@
   plot = sites_model_plot,
-  filename = "analysis/figures/results_model_plot.png",
+  filename = "analysis/results/figures/results_model_plot.png",
   width = 183,

### 2021-10-05 analysis/analyse_proportion_of_scouts.Rmd
started with the PCA analysis, see if it makes sense

@@
 ---
+
+## Set up data and environment
 
@@
 # put proportion of scouts into landuse data
-urban_lu_long <- sitedata %>%
+urban_lu <- sitedata %>%
   select(site, p) %>%
-  inner_join(urban_lu) %>%
+  inner_join(urban_lu)
+
+urban_lu_long <- urban_lu %>%
   pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
 
-rural_lu_long <- sitedata %>%
+rural_lu <- sitedata %>%
   select(site, p) %>%
-  inner_join(rural_lu) %>%
+  inner_join(rural_lu)
+rural_lu_long <- rural_lu %>%
   pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
@@
 Now we have everything in, lests do a basic plotting to see what the relationships between each landuse type and the proportion of scouts looks like.
+
+## Basic analysis
 
@@
   )
+
 ```
+
+## Princple Component Analysis
+
+```{r}
+
+prcomp(urban_lu[, 3:length(names(urban_lu))])
+```

### 2021-10-06 analysis/analyse_proportion_of_scouts.Rmd
changed axis around for lm exploratory plots

@@
 rural_lu_long %>%
-  ggplot(aes(x = p, y = values)) +
+  ggplot(aes(x = values, y = p)) +
   geom_point() +
   geom_smooth(method = "lm") +
-  facet_wrap(~landuse, ncol = 2) +
+  facet_wrap(~landuse, ncol = 2, scales="free") +
   theme(
@@
 urban_lu_long %>%
-  ggplot(aes(x = p, y = values)) +
+  ggplot(aes(x = values, y = p)) +
   geom_point() +
   geom_smooth(method = "lm") +
-  facet_wrap(~landuse, ncol = 2) +
+  facet_wrap(~landuse, ncol = 2, scales="free") +
   theme(

### 2021-10-06 src/RcppExports.cpp

@@ -5,6 +5,11 @@
 
 using namespace Rcpp;
 
+#ifdef RCPP_USE_GLOBAL_ROSTREAM
+Rcpp::Rostream<true>&  Rcpp::Rcout = Rcpp::Rcpp_cout_get();
+Rcpp::Rostream<false>& Rcpp::Rcerr = Rcpp::Rcpp_cerr_get();
+#endif
+
 // alter_in_place
 void alter_in_place(NumericVector x);
 RcppExport SEXP _wagglefit_alter_in_place(SEXP xSEXP) {

### 2021-10-14 analysis/analyse_proportion_of_scouts.Rmd
Added rmd file for the analysis of the proportion of scouts. WIP

--- 2021-10-05 analysis/analyse_proportion_of_scouts.Rmd
@@
 )
+
+library(tidyr)
+library(dplyr)
+library(ggbiplot)
+library(energy)
+library(betareg)
+
 ```
@@
 rural_lu_long %>%
-  ggplot(aes(x = p, y = values)) +
+  ggplot(aes(x = values, y = p)) +
   geom_point() +
   geom_smooth(method = "lm") +
-  facet_wrap(~landuse, ncol = 2) +
+  facet_wrap(~landuse, ncol = 2, scales = "free") +
   theme(
@@
 urban_lu_long %>%
-  ggplot(aes(x = p, y = values)) +
+  ggplot(aes(x = values, y = p)) +
   geom_point() +
   geom_smooth(method = "lm") +
-  facet_wrap(~landuse, ncol = 2) +
+  facet_wrap(~landuse, ncol = 2, scales = "free") +
   theme(
@@
 
+These previous graphs shows us that some variables seem to correlate with our estimated proportions of scouts, however, there are many different landscape types which likely co-occur together in a given site.
+It is therefore necessary to reduce the dimentionality of the dataset by conducting a principle Component analysis (PCA).
+This allows us to see which variables in the two environments most explain the variance.
+We can then take the major components and examine their relationship with our estimated proportion of scouts.
+
 ```{r}
 
-prcomp(urban_lu[, 3:length(names(urban_lu))])
+u_pca_data <- urban_lu[, 3:length(names(urban_lu))]
+
+# First, check the data follows a multivariative normal, p>0.05
+unormres <- u_pca_data %>%
+  mvnorm.etest(R = 100)
+
+# run PCA
+upca <- prcomp(
+  u_pca_data, center = TRUE, scale = TRUE
+)
+
+# extract the sd, loadings and scores of the data
+sd <- upca$sdev
+loadings <- upca$rotation
+rownames(loadings) <- colnames(u_pca_data)
+scores <- upca$x
+
+# calculate the variance and percentage of overall variances directly
+var <- sd^2
+varpercent <- var / sum(var) * 100
+
+# make a scree plot
+dev.new()
+barplot(
+  varpercent, xlab = "PC",
+  ylab = "Percent Variance",
+  names.arg = 1:length(varpercent), las = 1,
+  ylim = c(0, max(varpercent)),
+  col = "gray")
+
+abline(h = 1 / ncol(u_pca_data) * 100, col = "red")
+
 ```
+
+The largest drop in the scree plot occurs between PC1 and PC2 (46 - 22% of the variance).
+The red line indicates the expected variance if each PC contributed equally to the variance.
+Here, PCs 1, 2 and 3 all contribute more than their expected and the remaining pcs contribute increasingly less than expected.
+The results indicate we should include the first three Principle Components in our interpretation.
+These three PCs combine to explain 83% of the variance.
+
+We now need to know how our variables (landuse types) contribute to each of the PCs.
+This is given by the loadings, with a positive loading suggesting the variable is positively correlated with the PC and a negative loading indicating a negative correlation with the PC.
+
+A large (negative or positive) loading indicates a large effect on the PC.
+The job now is to determine the lagest loadings.
+One way to do this is to see which ones are larger than the average if all variables contributed equally to a given PC.
+
+```{r}
+
+ggbiplot(upca)
+
+loadings[, 1:3]
+sqrt(1 / ncol(u_pca_data)) # cutoff for top loadings
+
+```
+
+The table shows that continous central, dense and sparse residential, amenity grassland and water most contribute to PC1.
+For PC2 the main contributors are railway, woodland and water.
+For PC3, the main variables are sparse residential, parks, allotments and cemeteries and woodland.
+
+Now we have identified the principle components which explain the majority of the data, along with the varibales which contribute to these PCs, we now need to see how these PCs predict the proportion of scouts we estimated from our model fitting.
+To do this we use Princple Component Regression (PCR).
+In each PCA column we have the coefficients of a linear combination of the land use types which make up the PC (and we derived the most important of these above).
+In a PCR, these values are used as independent variables in a regression on to some other variable.
+Here, we will use this to compare against the proportion of scouts accross our sites.
+
+```{r}
+
+n <- length(urban_lu$p)
+y <- (urban_lu$p * (n - 1) + 0.5) / n # transform for zero or 1 inflated data
+mydata <- cbind(y, data.frame(upca$x))
+colnames(mydata)[1] <- "tranformed.prop.scouts"
+cor(mydata)[, 1]
+
+mydata.mod <- betareg(tranformed.prop.scouts ~ ., data = mydata)
+summary(mydata.mod)
+
+asoriginal <- upca$rotation %*% mydata.mod$coefficients$mean[-1]
+asoriginal
+
+```
+
+```{r}
+n <- length(urban_lu$p)
+y <- (urban_lu$p * (n - 1) + 0.5) / n # transform for zero or 1 inflated data
+
+df <- cbind(y, upca$x) %>%
+  as.data.frame()
+
+# run beta regression with each PC
+mod <- betareg(
+  y ~ PC1 + PC2 + PC3,
+  data = df
+)
+
+# display results
+summary(mod)
+
+fitted(mod)
+
+plot(upca$x[, 1], y)
+
+library(pls)
+
+data <- urban_lu
+data$p_transform <- y
+
+pcr.model <- plsr(p_transform ~ ., data = data[, 3:11], scale = TRUE, validation = "CV")
+
+summary(pcr.model)
+
+library(plsRbeta)
+
+model <- plsRbeta(p_transform ~ ., data = data[, 3:11], nt = 7, model = "pls-beta")
+
+model
+
+model$pp
+
+```
+
+```{r}
+
+# First, check the data follows a multivariative normal, p>0.05
+rnormres <- rural_lu[, 3:10] %>%
+  mvnorm.etest(R = 100)
+
+rpca <- prcomp(
+  rural_lu[, 3:length(names(rural_lu))], center = TRUE, scale = TRUE
+)
+
+print(summary(rpca))
+ggbiplot(rpca)
+
+rpca_var_explained <- tibble(
+  landscape = names(rural_lu[, 3:length(names(rural_lu))]),
+  prop_variance_explained = rpca$sdev^2 / sum(rpca$sdev^2)
+)
+
+rural_lu_long %>%
+  filter(landuse %in% rpca_var_explained[1:3,1][[1]]) %>%
+  ggplot(aes(x = values, y = p)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~landuse, ncol = 2, scales = "free") +
+  theme(
+    text = element_text(size = 8)
+  )
+
+```
+
+In the agri-rural environment the results show that the first PCA (woodland) explains 29% of the variation, the second PCA (non-agricultural unmanaged green space) explains 23% of the variance and the third PCA (non-agricultural managed green space) explains 19% of the variance.
+Combined, the first 2 PCs explain 52% of the variance, whilst the top three explain 71% of the variance.

### 2021-10-14 analysis/scouts_landuse_pca.Rmd
Added rmd file for the analysis of the proportion of scouts. WIP

---
title: Analysing the proportion of scouts estimated in different environments
subtitle: A work in progress
author: Joseph Palmer
output:
  html_notebook:
    theme: yeti
    toc: true
    toc_float: true
---
---

```{r, include=FALSE}
devtools::load_all()
library(showtext)
showtext_auto()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(size = 8)
    )
)

library(tidyr)
library(dplyr)
library(ggfortify)
library(energy)
library(betareg)
library(plsRbeta)

```

# Background and outline

Now we have used our model to estimate the proportion of scouts used by hives in different locations, it would be interesting to evaluate what effect (if any) the different land-use types have on the proportion scouts. The simplest way to do this would be to perform a regression between the proportion of scouts and the proportion of land-use type across the different sites. However, many land-use types are inherently correlated, for example dense and sparse residential terrain. Consequently, this analysis requires the use of a multivariate statistic. There are many to choose from but the two most common methods are the Principle Component Analysis (PCA) (followed by a regression) and Partial Least Squares (PLS).

A PCA regression is done between the main axis of variation in land-use and a response variable. First a PCA is carried out on different factors to produce a linear combination of coefficients that explain the most variation in the data. These factors can then be linearly combined in a regression on a response variable. For our study we are interested in how land-use influences proportion of scouts. We would therefore remove the proportion of scouts from the data to start with and perform a PCA on different land-use types to see which combinations best explain the variation. We would then use these combinations of factors as the coefficients for a regression of the proportion of scouts. Alternatively, a PLS is a single step process as unlike the PCA the response variable, in our case proportion of scouts, is not removed from the data. Instead the PLS finds the factors which most contribute to the variation within the response variable, rather than the variation within the land-use types.

In essence, a PCA allows us to test if the main axis of land-use variation between sites is related to the proportion of scouts, where as a PLS is suited to questions about what land-use types most influence the variation in the proportion of scouts (Scott & Crone, 2021).

It appears that our question is geared more towards a PLS rather than PCA as we are most interested in what factors most relate to the proportion of scouts, however, both approaches are suitable. A key consideration, however, has to be that our response variable is a proportion and thus lies on the interval $[0, 1]$. This makes it unsuitable for a standard linear regression and instead we need to use a beta regression. For PLS there is a dedicated R package for conducting PLS on proportional data (plsRbeta), however, for a PCA I am not aware of an equivalent package. This could be as simple as performing a PCA and passing the corresponding PCAs to a beta regression, however, within a PCA, the eigenvalues derived describe a linear combination of factors and so I am not sure is this is directly relatable.

Here, I will work through trying to evaluate the effect of land-use type on the proportion of scouts using both a PCA and PLS method. I begin with the urban data set which I will format with the proportion of scout data.

# Principle Component Analysis

```{r}
# load results of model fit (code to generate found in fit_models.Rmd)
sitedata <- readRDS("results/site_fit_results.Rda")

# filter for best model
sitedata <- sitedata %>%
  group_by(site) %>%
  slice(which.min(AIC))

# Load land use data taken (taken from Samuelson et al. 2021)
urban_lu <- read.csv(file = "data/urban-landuse.csv")
rural_lu <- read.csv(file = "data/agri-rural-landuse.csv")

# put proportion of scouts into landuse data
urban_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(urban_lu)
urban_lu_long <- urban_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")

rural_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(rural_lu)
rural_lu_long <- rural_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
```

Before doing a PCA it is important to check the data follows a multivariate normal distribution:

```{r}
urban.pcadata <- urban_lu[, 3:length(names(urban_lu))]
# First, check the data follows a multivariative normal, p>0.05
unormres <- urban.pcadata %>%
  mvnorm.etest(R = 100)
print(unormres)
```

As $p > 0.05$ we fail to reject the null hypothesis that the data follows a multivariate normal distribution.

Lets first start with a PCA and then come to the PLS later for comparison.

```{r}
# run PCA
upca <- prcomp(
  urban.pcadata,
  center = TRUE, scale. = TRUE
)
print(upca)
print(summary(upca))

autoplot(upca, data = as.data.frame(urban_lu), colour = "p") +
  scale_color_gradient(low = "green", high = "red")
```

The plot above shows PC1 against PC2 coloured by the proportion of scouts for that data row (green being 0 to red being 1). This suggests there isn't much of a relationship with the first two PCs and the proportion of scouts (however, I am not sure I am ploting this correctly as their are 8 PCs but 10 rows of data...)

The PCA results contain a number of results, notably the standard deviations and loadings (noted by this package as rotation) of each PC. The standard deviations can be squared to get the variance which can then be used to obtain the percentage variance each PC explains. These can then be used to make a scree plot to show the percentage variance each PC explains. In addition, each PC can then be compared to the amount of variance they would explain if each PC contributed equally to the variance. This can help in choosing which PCs most contribute to the data.

```{r}
# extract the sd, loadings and scores of the data
sd <- upca$sdev
loadings <- upca$rotation
rownames(loadings) <- colnames(urban.pcadata)
scores <- upca$x

# calculate the variance and percentage of overall variances directly
var <- sd^2
varpercent <- var / sum(var) * 100

# make a scree plot
dev.new()
barplot(
  varpercent,
  xlab = "PC",
  ylab = "Percent Variance",
  names.arg = 1:length(varpercent), las = 1,
  ylim = c(0, max(varpercent)),
  col = "gray"
)

abline(h = 1 / ncol(urban.pcadata) * 100, col = "red")
```

The Scree plot here indicates PCs 1, 2 and 3 most contribute to the variance in the data. The loadings can now be evaluated for each of these PCs to identify how each land-use type contributes to the variance in each PC. Again, we can use the expected variance from each land-use if they all equally contributed to find a cut-off.

```{r}

print(loadings[, 1:3])
sqrt(1 / ncol(urban.pcadata)) # cutoff for top loadings
```

In each column, values which are larger than 0.35 or smaller than -0.35 (the expected variance under a hypothesis of equal variance contributions) can be said to be the main contributors to that PC.

The table shows that continuous central, dense residential and water correlates negatively with PC1, whilst sparse residential and amenity grassland correlates positively with PC1. For PC2 the main contributors are railway, woodland and water, all of which have negative relationships. For PC3, the main variables are sparse residential and railway having a negative correlation with PC3, whilst parks, allotments and cemeteries and woodland all have positive correlations.

## Principle Component Regression

Now the PCA is complete we need to relate these major axis of variation (PC1, 2 and 3) to the proportion of scouts. As stated, we cant use a linear model here as the response variable is a proportion, instead we must use a beta regression on the proportion of scouts. These values currently sit on the interval $(0, 1])$ and so must be transformed to lie on the interval $(0, 1)$ via the equation:

$$
y = \frac{x (n-1) + s}{n}
$$

where $n$ represents the number of data points in $x$ and $s$ represents a constant between 0 and 1. Here, $s$ acts like a Bayesian prior and so we choose $0.5$ to represent an uninformed prior (Smithson & Verkuilen, 2006).

Here I will fit the beta regression to the first 3 PCs which collectively explain 83% of the data.

```{r}

n <- length(urban_lu$p)
y <- (urban_lu$p * (n - 1) + 0.5) / n # transform 1 inflated data
mydata <- cbind(y, data.frame(upca$x))
colnames(mydata)[1] <- "transformed.prop.scouts"
cor(mydata)[, 1]

mydata.mod <- betareg(
  transformed.prop.scouts ~ .,
  data = mydata
)
summary(mydata.mod)
```

The results of the regression shows the intercept is significantly different to 0, and all the PCs are significantly different to the intercept. The $R^2$ suggests 97% of the variation in the proportion of scouts is explained by our full model.

But, we worked out before that ideal number of principle components is 3 so lets see what the first three look like:

```{r}

mydata.mod.3 <- betareg(
  transformed.prop.scouts ~ PC1+PC2+PC3,
  data = mydata
)
summary(mydata.mod.3)

```

Here, the intercept is significantly different to 0, but only PC3 is significantly different to the intercept, with the combined coefficients explaining 47% of the variation in the response variable. If we examine PC3 on its own we see that it accounts for 38% of the variation in the proportion of scouts.

```{r}

mydata.mod.p3 <- betareg(
  transformed.prop.scouts ~ PC3,
  data = mydata
)
summary(mydata.mod.p3)

```

What this analyis seems to suggest is that PC3, whilst explaining 15% of the variation in landuse types, explains 38% of the variation in the proportion of scouts. Whilst PCs 1 and 2 explain higher variances in the landuse data, they explain much less the variation in the response variable.

** Some thoughts**

As I understand it, PCs 1 and 2 collectively explain 68% of the variation in land-use types between sites (45 % PC1 + 22% PC2) whilst PC3 explains only 15% of the variation in land-use types. However, when using these PCs to describe the variation in the proportion of scouts, PC3 explains 38% of the variation, whilst all the other PCs describe much less. I would expect the $R^2$ value to increase as more PCs are added to the model, not for it to spike in the middle. I am Unsure why this happens but one explanation could be that the PCs explain the variation in the data not the proportion of scouts and so only when relating it back to the proportion of scouts do we see that the main axis of variation does not correlate well with the proportion of scouts, however, the variation described by PC3 does correlate with it. With a beta regression the coefficients need to be transformed to get them in terms of the original values. This is done through the following function (https://stats.stackexchange.com/questions/120472/how-to-interpret-coefficients-of-a-beta-regression-model-with-logit-link):

$$
x = \frac{exp(coefficient)}{1+exp(coefficient)}
$$

using this function to retrieve the proportional increase, or decrease, in $y$ for every unit change in $x$ (as the model is log-linked), every 1 unit change in PC3 corresponds to an 65% increase in the proportion of scouts (I think, this is tricky to understand fully). A unit change in PC3 is difficult to understand as this change will not be linear in the parameters. I.e. all the contributing factors could change or only one could change.

I struggle to understand what this means in the real world. Sparse residential and railway have a negative correlation with PC3, suggesting that when these factors are lower the proportion of scouts is higher. Conversely, parks, allotments and cemeteries and woodland all have a positive correlation with PC3, suggesting when these attributes are higher, there are a higher number of scouts.

These results potentially highlight a problem with using the PCA approach as the thing we are interested in, how these factors influence proportion of scouts, are kept separate during the PCA. The results from this analysis, if done correctly and I haven't made any mistakes ( big if!), indicates that the main axis of variation in these landscapes does not correlate with the proportion of scouts. However, one of the principle components identified does correlate with a proportion of scouts. The first 3 Principle components together explain 83% of the variation in the data (the land-use coverage across the different sites) and explain 47% of the variation in the proportion of scouts.

Perhaps a PLS will make more sense as this more aligns with the question of what land-use type best explain the variation in the proportion of scouts.

## PCR using GML instead of beta regression

Instead of using a beta distribution I will try using a GML with a logit linker and quasibinomial family. This, however, doesn't feel right because this family is for count valued proportions (e.g. male to female ratio). Whilst the proportion of scouts is technically an integer proportion (scouts vs non-scouts) it is a continuous estimate derived from a model and so may not be suitable for this kind of analysis.

```{r}

mydata <- cbind(urban_lu$p, data.frame(upca$x))
colnames(mydata)[1] <- "prop.scouts"
cor(mydata)[, 1]

glm.mod <- glm(
  prop.scouts ~ .,
  data = mydata,
  family = quasibinomial(logit)
)
summary(glm.mod)

```

This suggests the intercept is significantly different to 0 and that PC2 provides a significant fit over the other PCs (I think).

# Partial Least Squares

```{r}

urban.pcadata$p_transform <- y

model <- plsRbeta(
  p_transform ~ .,
  data = urban.pcadata,
  nt = ncol(urban.pcadata) - 2,
  model = "pls-beta"
)

print(model)
```

Instead of showing the variation each PC explains in the land-use types, the PLS shows the variation in the proportion of scouts explained by each PC. The $R^2$ values therefore shows us the cumulative amount of variation in the proportion of scouts explained by adding more and more PCs. Here, rather than by defining an expected variance each would contribute too as we did for the PCA, we can look for where the $R^2$ slows down in increasing. The AIC and BIC also indicates where the cut of of how many PCs should be considered is. In this case, it looks like the cut off should be at PC3. Unlike the PCA regression, PLS doesn't produce p values. Instead a goodness of fit measure is given by the $R^2$ values and parsimony is assessed through the AIC criteria.

So from this the first three principle components explain 88% of the variation in the proportion of scouts. These principle components can be characterised as follows:

```{r}
model$pp[, 1:3]
sqrt(1 / (ncol(urban.pcadata) - 1))

```

A larger proportion of scouts is most described by a combination of continuous central with dense residential areas, a lack of amenity grassland and a large amount of water. This combination of factors explains ~61% of the variance in the proportion of scouts. Having sparse residential areas with plenty of parks, allotments and cemeteries, amenity grasslands, woodland and not much water increases some of the variation in the proportion of scouts. The third PCA combines with a high degree of continuous central land, not dense but not sparse residential land (does that even make sense?), very little in the way of amenity grassland and parks, lots of railways and water to describe some more variation in the proportion of scouts.

The other thing to note i that Cross validation is often needed to determine which PCs to keep hold of.

```{r}

bbb <- PLS_beta_kfoldcv_formula(
  p_transform ~., data = urban.pcadata, nt = ncol(urban.pcadata) - 2,
  model = "pls-beta", verbose = FALSE
)

kfolds2CVinfos_beta(bbb)

```

The results (The $Q^2$ ended ones) indicate Comp 3 is the best cut off.

## Standard PLSR

Here I will run a standard PLS regression with the response variable transformed according to the arcsine transformation.

Here, we want the Root mean squared error (RMSE) to be the lowest.

```{r}
library(pls)

usedat <- urban_lu
usedat$p_transform <- asin(sqrt(usedat$p))

qqnorm(usedat$p_transform)
qqline(usedat$p_transform, col = "steelblue", lwd = 2)

pls.model <- plsr(
  p_transform ~ .,
  data = usedat[, 3:ncol(usedat)],
  scale = TRUE,
  validation = "CV"
)

pls.model

summary(pls.model)

selectNcomp(
  pls.model, method = "randomization", plot = TRUE
)

```

First thing to note is that the Q-Q plots don't look fully normal.

Following this we can see that the RMSE increases with additional PCs and that adding PCs makes the fit even worse. Combinded these results suggests the model provides a worse fit than a horizontal straight line. E.g. the best model is a null model. So there is no effect of landuse type of the proportion of scouts in the urban environment.

I can't get my head around this. How can it have a worse fit than a null model to the proportion of scouts, yet simultaneously explain 87% of the variation in the proportion of scouts?

# Conclusion

The PLS with beta regression gives the most sensical answers in that it finds PCs which expain the largest variance in the proportion of scouts. The issue is I can't tell if these are significant or really interperate them as well as the others. The documentation for the package is very obtuse and whilst I can grasp how the model works in theory, I can't understand how their results relate those you would get with a simple linear response variable.

The PCA followed by the beta regression sort of makes sense in that I can desceribe it as the third main axis of variation in landuse type correlates with the the proportion of scouts / explains the most variation in the proportion of scouts. This would suggests it is a few factors witin a landscape driving the proportion of scouts without driving the overall landuse classification. This could be that their is not correlation and that their is no axis of variation in landuse type which explains the proportion of scouts more than a null model.

The idea that there is no significant association betwen landuse type and proportion of scouts is supported by the fitting of the quasibinomial glm and the PLS on the transformed proportion of scouts data. However, both of theese methods may not be appropriate: A. because the glm is geared more towards inetger proportions and B. Because transformed proportions still don't look all that normal (see qqplots above)

# Refs

Scott, E.R., Crone, E.E., 2021. Using the right tool for the job: the difference between unsupervised and supervised analyses of multivariate ecological data. Oecologia 196, 13–25.. doi:10.1007/s00442-020-04848-w

Smithson, M. & Verkuilen, J. A better lemon squeezer? Maximum-likelihood regression with beta-distributed dependent variables. Psychol. Methods 11, 54–71 (2006). DOI: 10.1037/1082-989X.11.1.54

### 2021-10-19 analysis/landuse_effect_prop_scouts.Rmd
Finished PLS and PCA analysis. sent to Elli and Vincent for feedback

---
title: Analysing the effect of land-use type on the proportion of scouts
author: Joseph Palmer
output:
  html_notebook:
    theme: yeti
    toc: true
    toc_float: true
---

---

```{r preamble, include=FALSE}
devtools::load_all()
library(showtext)
showtext_auto()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(size = 18)
    )
)

library(tidyr)
library(dplyr)
library(ggfortify)
library(energy)
library(betareg)
library(plsRbeta)
library(lmtest)
```

```{r get-data, include=FALSE}
# load results of model fit (code to generate found in fit_models.Rmd)
sitedata <- readRDS("results/site_fit_results.Rda")

# filter for best model
sitedata <- sitedata %>%
  group_by(site) %>%
  slice(which.min(AIC))

# Load land use data taken (taken from Samuelson et al. 2021)
urban_lu <- read.csv(file = "data/urban-landuse.csv")
rural_lu <- read.csv(file = "data/agri-rural-landuse.csv")

# put proportion of scouts into landuse data
urban_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(urban_lu)
urban_lu_long <- urban_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")

rural_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(rural_lu)
rural_lu_long <- rural_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
```

# Objective

The objective of this study to see if / how land-use influences the proportion of scouts estimated by our model. To do this I carried out both a Principle Component Analysis (PCA) with regression and a Partial Least Squares (PLS) of the land-use data across our sites. I ran both the PLS and PCA on the urban and rural sites separately due to differing land-use categories. This document contains the rational, methods and code for this analysis. At the end of the document I have attempted to write a discussion of these points. This requires more thought to put this into a theoretical framework and is the area I feel most unsure of.

# Method rational

In chapter 1 (outlined in fit_models.Rmd) we fit our models to waggle dance data and estimated the proportion of scouts used by hives in different environments. For chapter 2 I will now try and evaluate what, if any, effect different land-use types have on the proportion scouts estimated by our model. The simplest way to do this would be to perform a regression between the proportion of scouts and the proportion of land-use type across the different sites. However, many land-use types are inherently correlated, for example dense and sparse residential terrain. This analysis therefore requires the use of a multivariate statistic. There are many to choose from but the two most common methods are the Principle Component Analysis (PCA) followed by a regression and Partial Least Squares (PLS).

A PCA regression is done between the main axis of variation in land-use and a response variable. First a PCA is carried out on different factors to produce a linear combination of coefficients that explain the most variation in land-use types. These factors can then be linearly combined in a regression on a response variable. For our study we are interested in how land-use influences proportion of scouts. We would therefore remove the proportion of scouts from the data to start with and perform a PCA on different land-use types to see which combinations best explain the variation. We would then use these combinations of factors as the coefficients for a regression with the proportion of scouts as the response variable.

Alternatively, a PLS is a single step process as unlike the PCA the response variable, in our case proportion of scouts, is not removed from the data. Instead the PLS finds the factors which most contribute to the variation within the response variable, rather than the variation within the land-use types.

In essence, a PCA allows us to test if the main axis of land-use variation between sites is related to the proportion of scouts, where as a PLS is suited to questions about what land-use types most influence the variation in the proportion of scouts (Scott & Crone, 2021). Our question is geared more towards a PLS rather than PCA as we are most interested in what factors most relate to the proportion of scouts, however, both approaches are suitable and will be carried out here.

A key consideration is that our response variable is a proportion and thus lies on the interval $[0, 1]$. This makes it unsuitable for a standard linear regression in the PLS and the PCA regression. One option would be to use a generalized linear regression with a logit link and quasibinomial family. This method is most suited for when the response variable is an integer proportion, where the response variable is a division of two integers, such as the proportion of males within a population. Our measure is in principle an integer proportion as we measure the proportion of scouts, however, this value is an estimate derived from our model and so is continuous on the interval $[0, 1]$. An alternative method would be to transform the proportions on to a continuos scale using the $arcsin \sqrt{x}$ transformation. However, the most appropriate method would be to use a beta regression which is designed specifically for proportional response variables. For the PLS there is a dedicated R package for conducting PLS on proportional data (`plsRbeta`), however, for a PCA we just need to pass the PCA coefficients into a beta regression using the R package `betareg`.

In this document I will conduct a PLS and PCA + regression to investigate how land-use type influences the proportion of scouts in both urban and agri-rural environments.

# Urban

Here I use both a partial least squares and principle component analysis to investigate how different land-use types in the urban environment most influence the proportion of scouts estimates from our model.

First lets look at the data...

```{r check_urban}

# check the land-use type values sum to 100%
testthat::test_that(
  "land-use type values sum to 100",
  {
    expected <- rep(100, 10)
    ans <- rowSums(urban_lu[, 3:ncol(urban_lu)])
    testthat::expect_identical(expected, ans)
  }
)
```

A sanity check on the data shows that the sum of the land-use types for each site doesn't always sum to 100. This was copied over from the data presented in Samuelson *et al*. (2021).

```{r, peek_urban}
# look at data
print(urban_lu)
```

Only one site has 100% scouts (CAD) the rest are distributed between 0 and 50%

## PCA

```{r, urban_pca}
# extract just the lans-use columns
urban.pcadata <- urban_lu[, 3:length(names(urban_lu))]

# run PCA
upca <- prcomp(
  urban.pcadata,
  center = TRUE, scale. = TRUE
)
print(upca)
print(summary(upca))
```

The PCA results contain a number of results, notably the standard deviations and loadings (noted by this package as rotation) of each PC. The standard deviations can be squared to get the variance which can then be used to obtain the percentage variance each PC explains. These are shown in the summary of the PCA model, but deriving them this way allows you to easily make a scree plot to show the percentage variance each PC explains. In addition, each PC can then be compared to the amount of variance they would explain if each PC contributed equally to the variance. This can help in choosing which PCs most contribute to the data.

```{r, urban_pca_scree}
# extract the sd, loadings and scores of the data
sd <- upca$sdev
loadings <- upca$rotation
rownames(loadings) <- colnames(urban.pcadata)
scores <- upca$x

# calculate the variance and percentage of overall variances directly
var <- sd^2
varpercent <- var / sum(var) * 100

# make a scree plot
ggplot(
  data.frame(PC = seq(1, 8), varpercent = varpercent),
  aes(x = PC, y = varpercent)
) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 1 / ncol(urban.pcadata) * 100, col = "red") +
  ylab("% Variance")
```

The red line on the plot indicates the expected variance each PC should contribute if they all contributed equally. PCs 1, 2 and 3 explain ~83% of the variance in the data. The loadings can now be evaluated for each of these PCs to identify how each land-use type contributes to the variance in each PC. Again, we can use the expected variance from each land-use if they all equally contributed to find a cut-off.

```{r, urban_pca_loadings}
print(loadings[, 1:3])
sqrt(1 / ncol(urban.pcadata)) # cut off for top loadings
```

The expected variance each should contribute is around 0.35 and so any values smaller than -0.35 and larger than 0.35 can be said to be contributing more than expected to the variance than under a null model and are thus the main contributors to that PC.

The table shows that continuous central, dense residential and water correlates negatively with PC1, whilst sparse residential and amenity grassland correlates positively with PC1. For PC2 the main contributors are railway, woodland and water, all of which have negative relationships. For PC3, the main variables are sparse residential and railway having a negative correlation with PC3, whilst parks, allotments and cemeteries and woodland all have positive correlations.

These PCs all show the factors that most explain the variance within land-use types across our study sites. To see how this influences the proportion of scouts I will now do a beta regression of these Principle components.

## Principle Component Regression

Now the PCA is complete we need to relate these major axis of variation (PC1, 2 and 3) to the proportion of scouts. As stated, we cant use a linear model here as the response variable is a proportion, instead we must use a beta regression on the proportion of scouts. These values currently sit on the interval $(0, 1])$ and so must be transformed to lie on the interval $(0, 1)$ via the equation:

$$
y = \frac{x (n-1) + s}{n}
$$

where $n$ represents the number of data points in $x$ and $s$ represents a constant between 0 and 1. Here, $s$ acts like a Bayesian prior and so we choose $0.5$ to represent an uninformed prior (Smithson & Verkuilen, 2006).

The X attribute of the data (`upca$x`) is the value of the rotated data (centred and scaled) multiplied by the rotation matrix. This is what we use to run the model on as it is the PCs multiplied by the land-use values.

```{r, urban_p_transform}
n <- length(urban_lu$p)
y <- (urban_lu$p * (n - 1) + 0.5) / n # transform 1 inflated data
mydata <- cbind(y, data.frame(upca$x))
colnames(mydata)[1] <- "transformed.prop.scouts"
```

Here I will fit the beta regression to the first 3 PCs which collectively explain 83% of the data.

```{r, urban_pca_regression_3p}
mydata.mod.3 <- betareg(
  transformed.prop.scouts ~ PC1 + PC2 + PC3,
  data = mydata
)
summary(mydata.mod.3)

newdat <- data_frame(
  PC1 = seq(-5, 4, 0.1),
  PC2 = seq(-5, 4, 0.1),
  PC3 = seq(-5, 4, 0.1),
)

newdat$fit <- predict(mydata.mod.3, newdat, type = "response")

plotdata <- mydata
plotdata$p <- urban_lu$p

ggplot(data = plotdata, aes(x = PC1 + PC2 + PC3, y = p)) +
  geom_point() +
  geom_line(data = newdat, aes(x = PC1 + PC2 + PC3, y = fit)) +
  xlim(min(-6), max(4))
```

Here, the intercept is significantly different to 0, but only PC3 is significantly different to the intercept, with the combined coefficients explaining 47% of the variation in the response variable. If we examine PC3 on its own we see that it accounts for 38% of the variation in the proportion of scouts:

```{r, urban_pca_regression_p3}
mydata.mod.p3 <- betareg(
  transformed.prop.scouts ~ PC3,
  data = mydata
)
summary(mydata.mod.p3)

newdat <- data_frame(
  PC1 = seq(-5, 4, 0.1),
  PC2 = seq(-5, 4, 0.1),
  PC3 = seq(-5, 4, 0.1),
)

newdat$fit <- predict(mydata.mod.p3, newdat, type = "response")

plotdata <- mydata
plotdata$p <- urban_lu$p

ggplot(data = plotdata, aes(x = PC3, y = p)) +
  geom_point() +
  geom_line(data = newdat, aes(x = PC3, y = fit)) +
  xlim(min(-2), max(3))
```

What this analysis seems to suggest is that PC3, whilst explaining 15% of the variation in landuse types, explains 38% of the variation in the proportion of scouts. Whilst PCs 1 and 2 explain higher variances in the landuse data, they explain much less of the variation in the proportion of scouts.

Which combination of PCs to use, either PC3 on its own or PC1 + PC2 + PC3, or even PC2 + PC3, can be established through a likelihood ratio test.

```{r, urban_pca_regression_lrt}
mydata.mod.p1_p2 <- betareg(
  transformed.prop.scouts ~ PC2 + PC3,
  data = mydata
)

lrtest(mydata.mod.3, mydata.mod.p3, mydata.mod.p1_p2)
```

The LR test confirms that just using PC3 is the best model.

## PLS

Instead of working out the major axis of variation in land-use across our sites and evaluating how these influence the proportion of scouts, a PLS calculates the principle components which explains the most variation in the proportion of scouts directly.

```{r, urban_pls}
n <- length(urban_lu$p)
y <- (urban_lu$p * (n - 1) + 0.5) / n
urban.pcadata$p_transform <- y

model <- plsRbeta(
  p_transform ~ .,
  data = urban.pcadata,
  nt = ncol(urban.pcadata) - 2,
  model = "pls-beta"
)
print(model)
```

The $R^2$ values shows us the cumulative amount of variation in the proportion of scouts explained by adding more and more PCs. Here, rather than by defining an expected variance each would contribute to as we did for the PCA, we can look at factors like the where the $R^2$ slows down in increasing or the AIC and BIC for those models. In this case, it looks like the cut off should be at PC3. A goodness of fit measure is given by the $R^2$ values and parsimony is assessed through the AIC criteria. The PC eigenvalues can be fed into a beta regression model, just as we did with the PCA regression, to see how they look as a fit and obtain significance values.

Before running the variables through the beta regression lets first look at the principle components. The first three principle components explain 88% of the variation in the proportion of scouts. These principle components can be characterised as follows (again using an expected variance under a null model to work out the major contributors):

```{r, urban_pls_loadings}
print(model$pp[, 1:3])
sqrt(1 / (ncol(urban.pcadata) - 1))

# when multiple PCs have the best fit, summing them gives the overall effect of
# the PCs, just as we do in the model formula.
rowSums(model$pp[, 1:3])
```

A larger proportion of scouts is most described by a combination of continuous central with dense residential areas, a lack of amenity grassland and a large amount of water. This combination of factors explains ~61% of the variance in the proportion of scouts. Having sparse residential areas with plenty of parks, allotments and cemeteries, amenity grasslands, woodland and not much water increases some of the variation in the proportion of scouts (PC2). The third PC combines with a high degree of continuous central land, not dense but not sparse residential land (does that even make sense?), very little in the way of amenity grassland and parks, lots of railways and water to describe some more variation in the proportion of scouts.

We can extract the regression statistics out of the model just like we did for the PCA analysis. First I will do this just for PC1 (61% variance explained) and then for PC1 + PC2 + PC3 which explains 88% of the variation in the proportion of scouts.

```{r, urban_pls_p1}

plsdata <- data.frame(
  prop.scouts = model$dataY,
  model$tt
)

plsbetamodel <- betareg(
  prop.scouts ~ Comp_.1,
  plsdata
)

summary(plsbetamodel)

newdat <- data_frame(
  Comp_.1 = seq(-50, 140, 0.1),
)

newdat$fit <- predict(plsbetamodel, newdat, type = "response")

plotdata <- plsdata
plotdata$p <- urban_lu$p

ggplot(data = plotdata, aes(x = Comp_.1, y = p)) +
  geom_point() +
  geom_line(data = newdat, aes(x = Comp_.1, y = fit)) +
  xlim(min(-2.5), max(3.5))
```

Relatively nice fit. Seems to be the outlier (site where p = 1 ) pulling the proportions up.

```{r, urban_pls_3p}

plsdata <- data.frame(
  prop.scouts = model$dataY,
  model$tt
)

plsbetamodel.3 <- betareg(
  prop.scouts ~ Comp_.1 + Comp_.2 + Comp_.3,
  plsdata
)

summary(plsbetamodel.3)

newdat <- data_frame(
  Comp_.1 = seq(-50, 140, 0.1),
  Comp_.2 = seq(-50, 140, 0.1),
  Comp_.3 = seq(-50, 140, 0.1)
)

newdat$fit <- predict(plsbetamodel.3, newdat, type = "response")

plotdata <- plsdata
plotdata$p <- urban_lu$p

ggplot(data = plotdata, aes(x = Comp_.1 + Comp_.2 + Comp_.3, y = p)) +
  geom_point() +
  geom_line(data = newdat, aes(x = Comp_.1 + Comp_.2 + Comp_.3, y = fit)) +
  xlim(min(-2.5), max(4.5))
```

Provides an even nicer looking fit, but is of course a more complex model. Again, we turn to the LR test to decide what to keep.

```{r, urban_pls_lrt}
plsbetamodel.2 <- betareg(
  prop.scouts ~ Comp_.1 + Comp_.2,
  plsdata
)

lrtest(plsbetamodel, plsbetamodel.2, plsbetamodel.3)
```

Although PC1 explains the largest share of the data, including the first three PCs provides an improvement, despite the extra complexity.

# Agri-rural

Lets first examine the data...

```{r, check_rural}

# check the land-use type values sum to 100%
testthat::test_that(
  "land-use type values sum to 100",
  {
    expected <- rep(100, 10)
    ans <- rowSums(rural_lu[, 3:ncol(rural_lu)])
    testthat::expect_identical(expected, ans)
  }
)
```

A sanity check on the data shows that the sum of the land-use types for each site doesn't always sum to 100. This was copied over from the data presented in Samuelson *et al*. (2021).

```{r, peek_rural}
print(rural_lu)
```

Much wider spread than in the urban data. Here, two sites have values at 1 (BEl and STU), whilst the rest distribute between 0 and .65.

## PCA

```{r, rural_pca_scree}
# extract just the lans-use columns
rural.pcadata <- rural_lu[, 3:length(names(rural_lu))]

# run PCA
rpca <- prcomp(
  rural.pcadata,
  center = TRUE, scale. = TRUE
)
print(rpca)
print(summary(rpca))

# extract the sd, loadings and scores of the data
sd <- rpca$sdev
loadings <- rpca$rotation
rownames(loadings) <- colnames(rural.pcadata)
scores <- rpca$x

# calculate the variance and percentage of overall variances directly
var <- sd^2
varpercent <- var / sum(var) * 100

# make a scree plot
ggplot(
  data.frame(PC = seq(1, 10), varpercent = varpercent),
  aes(x = PC, y = varpercent)
) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 1 / ncol(urban.pcadata) * 100, col = "red") +
  ylab("% Variance") +
  scale_x_continuous(breaks = seq(1, 10))
```

The red line on the plot indicates the expected variance each PC should contribute if they all contributed equally. PCs 1, 2, 3 and 4 explain ~83% of the variance in the data. The loadings can now be evaluated for each of these PCs to identify how each land-use type contributes to the variance in each PC. Again, we can use the expected variance from each land-use if they all equally contributed to find a cut-off.

```{r, rural_pca_loadings}
print(loadings[, 1:4])
sqrt(1 / ncol(rural.pcadata)) # cut off for top loadings
```

The expected variance each should contribute is around 0.32 and so any values smaller than -0.32 and larger than 0.32 can be said to be contributing more than expected to the variance than under a null model and are thus the main contributors to that PC.

The table shows that PC1 correlates positively with arable land and negatively with woodland, Non agricultural managed green space, other agricultural, fruit and water. PC2 correlates positively with Non-agricultural managed green space, other agricultural and OSR, whilst correlating negatively with pasture. PC3 correlates negatively with woodland and OSR, whilst correlating positively with Non agricultural unmanaged and managed green spaces and built up areas. Finally, PC4 correlates negatively with arable land, fruit and water, whilst correlating positively with built up areas.

These PCs all show the factors that most explain the variance within land-use types across our study sites. To see how this influences the proportion of scouts I will now do a beta regression of these Principle components.

## Principle Component Regression

This regression is carried out with the same rational as the urban sites.

```{r, rural_transform_p}
n <- length(rural_lu$p)
y <- (rural_lu$p * (n - 1) + 0.5) / n # transform 1 inflated data
mydata <- cbind(y, data.frame(rpca$x))
colnames(mydata)[1] <- "transformed.prop.scouts"
```

Here I will fit the beta regression to the first 4 PCs which collectively explain 83% of the data.

```{r, rural_pca_4p}
mydata.mod.4 <- betareg(
  transformed.prop.scouts ~ PC1 + PC2 + PC3 + PC4,
  data = mydata
)
summary(mydata.mod.4)

newdat <- data_frame(
  PC1 = seq(-5, 4, 0.1),
  PC2 = seq(-5, 4, 0.1),
  PC3 = seq(-5, 4, 0.1),
  PC4 = seq(-5, 4, 0.1)
)

newdat$fit <- predict(mydata.mod.4, newdat, type = "response")

plotdata <- mydata
plotdata$p <- rural_lu$p

ggplot(data = plotdata, aes(x = PC1 + PC2 + PC3 + PC4, y = p)) +
  geom_point() +
  geom_line(data = newdat, aes(x = PC1 + PC2 + PC3 + PC4, y = fit)) +
  xlim(min(-5), max(6))
```

Here, the intercept is not significantly different to 0, but PC1 and PC3 are both significantly different, with the combined coefficients of PCs 1:4 explaining 60% of the variation in the proportion of scouts in a negative correlation. If we examine PC1 on its own se see that is accounts for ~29% of the variation in the proportion of scouts in a positive correlation:

```{r, rural_pca_p1}

mydata.mod.p1 <- betareg(
  transformed.prop.scouts ~ PC1,
  data = mydata
)
summary(mydata.mod.p1)

newdat <- data_frame(
  PC1 = seq(-5, 4, 0.1),
  PC2 = seq(-5, 4, 0.1),
  PC3 = seq(-5, 4, 0.1),
)

newdat$fit <- predict(mydata.mod.p1, newdat, type = "response")

plotdata <- mydata
plotdata$p <- rural_lu$p

ggplot(data = plotdata, aes(x = PC1, y = p)) +
  geom_point() +
  geom_line(data = newdat, aes(x = PC1, y = fit)) +
  xlim(min(-3), max(4))
```

Examining PC3 on its own suggests it accounts for ~27% of the variation in the proportion of scouts, showing a singificant negative relationship.

```{r, rural_pca_p3}
mydata.mod.p3 <- betareg(
  transformed.prop.scouts ~ PC3,
  data = mydata
)
summary(mydata.mod.p3)

newdat <- data_frame(
  PC1 = seq(-5, 4, 0.1),
  PC2 = seq(-5, 4, 0.1),
  PC3 = seq(-5, 4, 0.1),
)

newdat$fit <- predict(mydata.mod.p3, newdat, type = "response")

plotdata <- mydata
plotdata$p <- rural_lu$p

ggplot(data = plotdata, aes(x = PC3, y = p)) +
  geom_point() +
  geom_line(data = newdat, aes(x = PC3, y = fit)) +
  xlim(min(-3), max(3))
```

This analysis suggests that PCs 1 and 3, which collectively explain ~51% of the variation in the proportion of scouts, explain both a significant postive (PC1) and negative (PC2) relationship with the proportion of scouts, accoutning for ~29% and ~27% of the variance respectively.

```{r, rural_pca_lrt}
lrtest(mydata.mod.p1, mydata.mod.4)
```

PC1 seems to be the best model compared to the others according to the LR test, e.g. the more complex model does not explain the data any better.

## PLS

The rational for the PLS follows that used on the urban data.

```{r, rural_pls}
n <- length(rural_lu$p)
y <- (rural_lu$p * (n - 1) + 0.5) / n
rural.pcadata$p_transform <- y

model <- plsRbeta(
  p_transform ~ .,
  data = rural.pcadata,
  nt = ncol(rural.pcadata) - 1,
  model = "pls-beta"
)
print(model)
```

The partial least squares method was only able to extract 8 Principle components. nevertheless, depending on what criteria you use you can place the cutoff in multiple places. Combined, all 8 PCs explain ~83% of the variation in the proportion of scouts. The AIC and BIC scores suggests keeping only the first PC, which explains ~64% of the variation in the proportion of scouts. If looking for the slow down in the rate of increae in $R^2$ you would choose up to PC3, which explains ~77% of the variaton in the proportion of scouts. Lets looks at these first three PCs.

```{r, rural_pls_loadings}
model$pp[, 1:3]
sqrt(1 / (ncol(rural.pcadata) - 1))
```

PC1 (accoutning for ~64% of the variance) correlates positively with arable land and negatively with Non agricultural unmanaged green space, built up areas and water. PC2 (acoutning for ~7% of the variation) correlates positively with other agriculutural, fruit and OSG, whilst correlating negatively with Non agricultural managed green space and built up areas. PC3 (also acounting for ~7% of the variation) correlates positively with other agricultural and built up areas, and negatively with fruit and water.

We can extract the regression statistics out of the model just like we did for the PCA analysis. First I will do this just for PC1 (~64% variance explained) and then for PC1 + PC2 + PC3 which explains ~77% of the variation in the proportion of scouts.

```{r, rural_pls_p1}

plsdata <- data.frame(
  prop.scouts = model$dataY,
  model$tt
)

plsbetamodel <- betareg(
  prop.scouts ~ Comp_.1,
  plsdata
)

summary(plsbetamodel)

newdat <- data_frame(
  Comp_.1 = seq(-50, 140, 0.1),
)

newdat$fit <- predict(plsbetamodel, newdat, type = "response")

plotdata <- plsdata
plotdata$p <- rural_lu$p

ggplot(data = plotdata, aes(x = Comp_.1, y = p)) +
  geom_point() +
  geom_line(data = newdat, aes(x = Comp_.1, y = fit)) +
  xlim(min(-2.5), max(3.))
```

Even with a lack of data, this looks like a pretty good relationship!

```{r, rural_pls_4p}

plsdata <- data.frame(
  prop.scouts = model$dataY,
  model$tt
)

plsbetamodel.3 <- betareg(
  prop.scouts ~ Comp_.1 + Comp_.2 + Comp_.3,
  plsdata
)

summary(plsbetamodel.3)

newdat <- data_frame(
  Comp_.1 = seq(-50, 140, 0.1),
  Comp_.2 = seq(-50, 140, 0.1),
  Comp_.3 = seq(-50, 140, 0.1)
)

newdat$fit <- predict(plsbetamodel.3, newdat, type = "response")

plotdata <- plsdata
plotdata$p <- rural_lu$p

ggplot(data = plotdata, aes(x = Comp_.1 + Comp_.2 + Comp_.3, y = p)) +
  geom_point() +
  geom_line(data = newdat, aes(x = Comp_.1 + Comp_.2 + Comp_.3, y = fit)) +
  xlim(min(-3.5), max(5.5))
```

Good looking relationship, but what does the LRT say?

```{r, rural_pls_lrt}
plsbetamodel.2 <- betareg(
  prop.scouts ~ Comp_.1 + Comp_.2,
  plsdata
)

lrtest(plsbetamodel, plsbetamodel.2, plsbetamodel.3)
```

The added complexity is not needed. PC1 on its own explains the data well enough.

# Discussion / Conclusions

Here I have walked through a PCA + beta regression and PLS analysis on both the urban and agri-rural datasets. Both of these methods reveal interesting relationships between the proportion of scouts and land-use types in urban and agri-rural environments. In both environments we are able to identify principle components which show a singificant relationship with our estimated proportion of scouts. The tricky part is placing these in a theoretical framework about how the environment influences foraging behaviours by a colony.

In the urban environment, the PCA + regression identifies three main axis of variation, yet it is only the third PC which has a singificant relationship with the proportion of scouts. This PC increases with parks, allotments, cemeteries and woodland whilst decreasing with sparse residential and railway. In essence, this suggests a landscape with higher residential areas as increasing the proportion of scouts, whilst more green spaces (parks, allotments and cemeteries and woodlands) decreases the proportion of scouts, suggesting more reliable foraging grounds. However, the relationship here is rather weak ($R^2$ = 0.38). This is likely due to the nature of the PCA. The PCA finds the priniciple axis of variation in the data rather than our response variable. Consequently, we end up only evaluating the effect of the main axis of variation in land-use type on the proportion of scouts. The PLS allows us to directly evaluate the land-use types most contributing to the proportion of scouts.

Just like the PCA, the PLS identifies 3 major axis of variation which together explain 88% of the variation in the proportion of scouts. Both in isolation and combined, these three PCs have a singificant positive correlation with the proportion of scouts. The best fit comes from a combination of these three PCs. Collectively, these first three PCs correlate positively with continous central land, parks allotments and cemeteries,railway  and water, and correlate negatively with dense residential areas and amenity grassland. Unlike with the PCA, this relationship is fairly strong ($R^2 = 0.88$) So, in environments where there are more green spaces (parks, allotments, woodland etc.), more continous central land and water, along with less residential and grasslands, honeybee colonies use a scout dominant foraging strategy.

These results suggests in urban environments, when forage is easy to find (open spaces?) colonies send out more scouts rather than optimising information by using recruits. When resources are harder to find a colony uses more recruits. This suggests, in urban environments at least, that the waggle dance is more benefician when resources are harder to find. However, it should be noted that this relationship may be biased by the single site where the proportion of scouts sat at 100%.

In the agri-rural environment the PCA + regression identifies 4 main axis of variation explaining 83% of the variation in the data, yet it is PCs 1 and 3 which show a singificant relationship with the proportion of scouts. PC1 describes the most variation in the data (~29%) and is a singificantly more parsimonious model than just PC3 or all four Pcs combined. PC1 correlates positively with arable land and negatively with woodland, Non agricultural managed green space, other agricultural, fruit and water. PC3, which correlates negatively with the proportion of scouts, correlates negatively with woodland and OSR, whilst correlating positively with Non agricultural unmanaged and managed green spaces and built up areas. Combined, these results sort of follow the results for the urban data, that when resources are easier to find a colony uses more recruits, where as when resources are harder to find they use more scouts. Again though, only ~30% of the variation is accounted for and so the association is quite weak.

The PLS identifies three PCs which combined explain 77% of the variation in the proportion of scouts, however, 64% of this is driven by a single PC. The regression suggests this single PC explains the data well enough and that adding more complexity through additional PCs is not needed. This PC (which correlates positively with the proportion of scouts) correlates positively with arable land and negatively with Non agricultural unmanaged green space, built up areas and water.

**Stats notes: things to add or improve in the stats**

* One thing it would be nice to do, but which I have not figured out how to do, is to see how much od the data the PLS PCs explain. With the PCA we obtain the variance a PC explains in the data: i.e. the variance in land-use types within our data ($R^2_x$). A regression after then tells us how much of the variance in the proportion of scouts that particular PC explains ($R^2_y$). The PLS optimise a PC to describe as much of the variation in the proportion of scouts as possible ($R^2_y$) but it would be also nice to know how much of the variation in land-use type ($R^2_x$) that PC explains.

* It would be nice to add confidence intervals to the model fit plots. Unsure how to do this with a beta regression and not all that sure it is that important?

### 2021-10-22 analysis/foraging_distance_analysis.Rmd
added PCA and PLS analysis of foraging distances

---
title: Analysing the effect of land-use type on honeybee foraging distances
author: Joseph Palmer
output:
  html_notebook:
    theme: yeti
    toc: true
    toc_float: true
---

---

```{r preamble, include=FALSE}
devtools::load_all()
library(showtext)
showtext_auto()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(size = 18)
    )
)

library(tidyr)
library(tidyverse)
library(dplyr)
library(ggfortify)
library(lmtest)
library(tibble)
library(plsRglm)
```

# TLDR

Land-use type appears to have a significant interaction with the average foraging distance of a colony. This relationship is only apparent when we subset the data to a season level and shows landscapes which are typically nutrient rich require travelling short distances than those which are typically nutrient poor.

# Overview

From Samuelson *et al*. (2021) we know that honeybees in agri-rural environments travel further when foraging than their urban counterparts. This is also evident in the data when comparing the foraging distances. As we have land-use data for these different sites it would be interesting to know what land-use types most correlate / influence the average foraging distances for a colony. To do this, I perform a regression between the average foraging distance in a given site and the principle axis of variation in land-use type across the different sites in both the urban and agri-rural environment. Please note that the term average here refers to a measure of central tendency and not exclusively the mean. In addition to evaluating to what extent the principle axis of land-use variation influences the average foraging distance of the site, I also identify the principle components which best describe the variation in average foraging distance and evaluate their relationship in both the urban and agri-rural environment. In addition to performing this on the urban and agri-rural sites, I also evaluate the relationship between land-use type and average foraging distance with the data subset by year and season.

As a start, lets first look at the foraging distances within each site.

```{r, foraging_distance_dist, message = FALSE, warning = FALSE}
# get the data
# Load land use data taken (taken from Samuelson et al. 2021)
urban_lu <- read.csv(file = "data/urban-landuse.csv")
rural_lu <- read.csv(file = "data/agri-rural-landuse.csv")

# foraging distance data
df <- read.csv("data/FullHBForagingData.csv")

ggplot(df, aes(x = ForagingDistance_km), fill = LU) +
  geom_histogram(binwidth = 1, colour = "black") +
  facet_wrap(~site)
```

In all cases the distributions are right skewed suggesting the median should be used instead of the mean as a measure of central tendency of the data.

```{r, median_foraging_distance_comparison, message = FALSE, warning = FALSE}
# calculate median foraging distances for each site
df.median.fd <- df %>%
  group_by(LU, site) %>%
  summarise(Avg.foraging.distance = median(ForagingDistance_km))

# add land-use types to the foraging data
rural.df <- df.median.fd %>%
  filter(LU == "Rural") %>%
  inner_join(rural_lu)

urban.df <- df.median.fd %>%
  filter(LU == "Urban") %>%
  inner_join(urban_lu)

# plot median foraging distances for each environment
df.median.fd %>%
  ggplot(aes(x = LU, y = Avg.foraging.distance)) +
  geom_boxplot() +
  geom_point() +
  ylab("Median foraging distance (km)") +
  xlab("Environment")
```

From a rough eyeball the hives in the agri-rural sites do appear to be foraging on average much further than those in the urban sites. I will now evaluate the land-use factors (if any) which may influence foraging distances in the urban and agri-rural sites, starting with urban.

# Urban

## PCA

```{r, urban_pca, message = FALSE, warning = FALSE}
urban.pcadata <- urban.df[, 4:ncol(urban.df)]

# run PCA
upca <- prcomp(
  urban.pcadata,
  center = TRUE, scale. = TRUE
)
print(upca)
print(summary(upca))

# extract the sd, loadings and scores of the data
sd <- upca$sdev
loadings <- upca$rotation
rownames(loadings) <- colnames(urban.pcadata)
scores <- upca$x

# calculate the variance and percentage of overall variances directly
var <- sd^2
varpercent <- var / sum(var) * 100

# make a scree plot
ggplot(
  data.frame(PC = seq(1, 8), varpercent = varpercent),
  aes(x = PC, y = varpercent)
) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 1 / ncol(urban.pcadata) * 100, col = "red") +
  ylab("% Variance")
```

The red line on the plot indicates the expected variance each PC should contribute if they all contributed equally. PCs 1, 2 and 3 explain ~83% of the variance in the data. The loadings can now be evaluated for each of these PCs to identify how each land-use type contributes to the variance in each PC. Again, we can use the expected variance from each land-use if they all equally contributed to find a cut-off.

```{r, urban_pca_loadings, message = FALSE, warning = FALSE}
print(loadings[, 1:3])
sqrt(1 / ncol(urban.pcadata)) # cut off for top loadings
```

The expected variance each should contribute is around 0.35 and so any values smaller than -0.35 and larger than 0.35 can be said to be contributing more than expected to the variance than under a null model and are thus the main contributors to that PC.

The table shows that continuous central, dense residential and water correlates negatively with PC1, whilst sparse residential and amenity grassland correlates positively with PC1. For PC2 the main contributors are railway, woodland and water, all of which have negative relationships. For PC3, the main variables are sparse residential and railway having a negative correlation with PC3, whilst parks, allotments and cemeteries and woodland all have positive correlations.

These PCs all show the factors that most explain the variance within land-use types across our study sites. To see how this influences the average foraging distance I will now perform a regression of these Principle components.

## PCA + regression

```{r, urban_pca_regression, warning = FALSE, message = FALSE}

mydata <- cbind(
  Avg.foraging.distance = urban.df$Avg.foraging.distance, data.frame(upca$x)
)

mydata.mod.3 <- lm(
  Avg.foraging.distance ~ PC1 + PC2 + PC3,
  data = mydata
)
summary(mydata.mod.3)

newdat <- data_frame(
  PC1 = seq(-5, 4, 0.1),
  PC2 = seq(-5, 4, 0.1),
  PC3 = seq(-5, 4, 0.1),
)

newdat$fit <- predict(mydata.mod.3, newdat, type = "response")

plotdata <- mydata

ggplot(data = plotdata, aes(x = PC1 + PC2 + PC3, y = Avg.foraging.distance)) +
  geom_point() +
  geom_line(data = newdat, aes(x = PC1 + PC2 + PC3, y = fit)) +
  xlim(-6, 4)

```

Whilst the intercept is significantly different to 0 their is no significant relationship between the average foraging distance and the main axis of variation in land-use type.

## PLS

Using a PLS we can evaluate what land-use types most explain the variation in the average foraging distances across the urban sites.

```{r urban_pls, warning = FALSE, message = FALSE}

urban.plsdata <- urban.df[, 3:ncol(urban.df)]

model <- plsRglm(
  Avg.foraging.distance ~ .,
  data = urban.plsdata,
  nt = ncol(urban.plsdata) - 1,
  verbose = FALSE
)

model
```

The PLS results indicate the first 4 PCs explain 54% of the variation in the average foraging distance. The four PCs are characterised as follows:

```{r, urban_pls_loadings, message = FALSE, warning = FALSE}

print(model$pp[, 1:4])
sqrt(1 / (ncol(urban.plsdata) - 1))
```

Values greater than 0.35 or less -0.35 can be said to be contributing more towards the variance than would be expected under a null model. PC1 shows a positive correlation with continuous central, railway and water, and a negative correlation with sparse residential and amenity grassland. PC2 shows a positive correlation sparse residential and amenity grassland and a negative correlation with water and dense residential. PC3 shows a positive correlation with continuous central, parks allotments and cemeteries, and a negative correlation with railway. PC4 describes a positive correlation with amenity grassland, railway and woodland, and a negative relationship with sparse residential.

We can pull out the coefficients to see what this relationship looks like graphically.

```{r, urban_pls_fit4, message = FALSE, warning = FALSE}
plsdata <- data.frame(
  Avg.foraging.distance = model$dataY,
  model$tt
)

plsmodel <- lm(
  Avg.foraging.distance ~ Comp_1 + Comp_2 + Comp_3 + Comp_4,
  plsdata
)

summary(plsmodel)

newdat <- data_frame(
  Comp_1 = seq(-50, 140, 0.1),
  Comp_2 = seq(-50, 140, 0.1),
  Comp_3 = seq(-50, 140, 0.1),
  Comp_4 = seq(-50, 140, 0.1)
)

newdat$fit <- predict(plsmodel, newdat, type = "response")

plotdata <- plsdata

ggplot(
  data = plotdata,
  aes(x = Comp_1 + Comp_2 + Comp_3 + Comp_4, y = Avg.foraging.distance)
) +
  geom_point() +
  geom_line(
    data = newdat, aes(x = Comp_1 + Comp_2 + Comp_3 + Comp_4, y = fit)
  ) +
  xlim(min(-4.5), max(3)) +
  ylim(0, 3.5)
```

Although the fit is much better than with the PCA and the intercept is significantly different to 0, there is no significant relationship between the PCs and the median foraging distances reported across the different sites. This suggests the land-use types we have recorded do not explain foraging distances in the urban environment.

Could do some kind of GLM maybe instead?

# Agri-rural

## PCA

```{r, agri-rural_pca, warning = FALSE, message = FALSE}
rural.pcadata <- rural.df[, 4:ncol(rural.df)]

# run PCA
rpca <- prcomp(
  rural.pcadata,
  center = TRUE, scale. = TRUE
)
print(rpca)
print(summary(rpca))

# extract the sd, loadings and scores of the data
sd <- rpca$sdev
loadings <- rpca$rotation
rownames(loadings) <- colnames(rural.pcadata)
scores <- rpca$x

# calculate the variance and percentage of overall variances directly
var <- sd^2
varpercent <- var / sum(var) * 100

# make a scree plot
ggplot(
  data.frame(PC = seq(1, 10), varpercent = varpercent),
  aes(x = PC, y = varpercent)
) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 1 / ncol(rural.pcadata) * 100, col = "red") +
  ylab("% Variance")
```

The red line on the plot indicates the expected variance each PC should contribute if they all contributed equally. PCs 1, 2, 3 and 4 explain ~83% of the variance in the data. The loadings can now be evaluated for each of these PCs to identify how each land-use type contributes to the variance in each PC. Again, we can use the expected variance from each land-use if they all equally contributed to find a cut-off.

```{r, agri-rural_pca_loadings, message = FALSE, warning = FALSE}
print(loadings[, 1:3])
sqrt(1 / ncol(rural.pcadata)) # cut off for top loadings
```

The expected variance each should contribute is around 0.32 and so any values smaller than -0.32 and larger than 0.32 can be said to be contributing more than expected to the variance than under a null model and are thus the main contributors to that PC.

The table shows that PC1 correlates positively with arable land and negatively with woodland, Non agricultural managed green space, other agricultural, fruit and water. PC2 correlates positively with Non-agricultural managed green space, other agricultural and OSR, whilst correlating negatively with pasture. PC3 correlates negatively with woodland and OSR, whilst correlating positively with Non agricultural unmanaged and managed green spaces and built up areas. Finally, PC4 correlates negatively with arable land, fruit and water, whilst correlating positively with built up areas.

These PCs all show the factors that most explain the variance within land-use types across our study sites. To see how this influences the proportion of scouts I will now do a beta regression of these Principle components.

## PCA + regression

```{r, agri-rural_pca_regression, warning = FALSE, message = FALSE}

mydata <- cbind(
  Avg.foraging.distance = rural.df$Avg.foraging.distance, data.frame(rpca$x)
)

mydata.mod.3 <- lm(
  Avg.foraging.distance ~ PC1 + PC2 + PC3 + PC4,
  data = mydata
)
summary(mydata.mod.3)

newdat <- data_frame(
  PC1 = seq(-5, 4, 0.1),
  PC2 = seq(-5, 4, 0.1),
  PC3 = seq(-5, 4, 0.1),
  PC4 = seq(-5, 4, 0.1),
)

newdat$fit <- predict(mydata.mod.3, newdat, type = "response")

plotdata <- mydata

ggplot(
  data = plotdata,
  aes(x = PC1 + PC2 + PC3 + PC4, y = Avg.foraging.distance)
) +
  geom_point() +
  geom_line(data = newdat, aes(x = PC1 + PC2 + PC3 + PC4, y = fit)) +
  xlim(min(-6), max(6))
```

Whilst the intercept is significantly different to 0 their is no significant relationship between the average foraging distance and the main axis of variation in land-use type.

## PLS

Using a PLS we can evaluate what land-use types most explain the variation in the average foraging distances across the urban sites.

```{r, agri-rural_pls, warning = FALSE, message = FALSE}

rural.plsdata <- rural.df[, 3:ncol(rural.df)]

model <- plsRglm(
  Avg.foraging.distance ~ .,
  data = rural.plsdata,
  nt = ncol(rural.plsdata) - 1,
  verbose = FALSE
)

model
```

The PLS results indicate the first 3 PCs explain 86% of the variation in the average foraging distance. The three PCs are characterised as follows:

```{r, agri-rural_pls_loadings, message = FALSE, warning = FALSE}
print(model$pp[, 1:3])
sqrt(1 / (ncol(rural.plsdata) - 1))
```

Values greater than 0.32 or less -0.32 can be said to be contributing more towards the variance than would be expected under a null model. PC1 correlates positively with arable and OSR, and negatively with pasture and built up areas. PC2 correlates positively with pasture and water, but negatively with woodland, non-agricultural managed green space, other agricultural land, OSR and built up areas. PC3 correlates positively with Non agricultural unmanaged green space, other agricultural and built up areas, and negatively with arable land and fruit.

We can pull out the coefficients to see what this relationship looks like graphically.

```{r, agri-rural_pls_pc_fits, message = FALSE, warning = FALSE}
plsdata <- data.frame(
  Avg.foraging.distance = model$dataY,
  model$tt
)

plsmodel <- lm(
  Avg.foraging.distance ~ Comp_1 + Comp_2 + Comp_3,
  plsdata
)

summary(plsmodel)

newdat <- data_frame(
  Comp_1 = seq(-50, 140, 0.1),
  Comp_2 = seq(-50, 140, 0.1),
  Comp_3 = seq(-50, 140, 0.1),
)

newdat$fit <- predict(plsmodel, newdat, type = "response")

plotdata <- plsdata

ggplot(
  data = plotdata,
  aes(x = Comp_1 + Comp_2 + Comp_3, y = Avg.foraging.distance)
) +
  geom_point() +
  geom_line(
    data = newdat, aes(x = Comp_1 + Comp_2 + Comp_3, y = fit)
  ) +
  xlim(min(-2.5), max(4)) +
  ylim(0, 4.5)
```

In addition to the intercept being significantly different to 0, PC1 is significantly different to the intercept, with every one unit increase in PC1 corresponding to an increase of 410 meters. The other PCs do not significantly influence this association.

Plotting PC1 on its own against the average foraging distance maintains a significant relationship.

```{r, agri-rural_pls_pc1_fits, message = FALSE, warning = FALSE}
plsdata <- data.frame(
  Avg.foraging.distance = model$dataY,
  model$tt
)

plsmodel.pc1 <- lm(
  Avg.foraging.distance ~ Comp_1,
  plsdata
)

summary(plsmodel.pc1)

newdat <- data_frame(
  Comp_1 = seq(-50, 140, 0.1),
)

newdat$fit <- predict(plsmodel.pc1, newdat, type = "response")

plotdata <- plsdata

ggplot(
  data = plotdata,
  aes(x = Comp_1, y = Avg.foraging.distance)
) +
  geom_point() +
  geom_line(
    data = newdat, aes(x = Comp_1, y = fit)
  ) +
  xlim(min(-2), max(3)) +
  ylim(0, 4.5)
```

Just to verify this we can perform a LRT.

```{r, agri-rural_pls_lrt, message = FALSE, warning = FALSE}
plsmodel.pc1_pc2 <- lm(
  Avg.foraging.distance ~ Comp_1 + Comp_2,
  plsdata
)

lrtest(plsmodel.pc1, plsmodel.pc1_pc2, plsmodel)
```

PC1+PC2+PC3 explains the data better than just PC1 on its own, despite these additional PCs not showing a significant relationship (although the p is not far from 0.05). The model suggests a combination of these PCs explains ~86% of the variation in average foraging distance.

# Discussion

For the urban sites there does not seem to be any effect of land-use on the median foraging distances reported by honey-bees. Whilst using a PLS regression provides a better fit than the PCA + regression, a significant effect is not identified. In the Agri-rural sites the PCA + regression did not identify any significant relationships between the main axis of land-use variation and the median foraging distance. However, the PLS regression identified a three PCs, accounting for 86% of the variation in average foraging distance, had a significant positive relationship with the average foraging distance. Of these only the first PC (explaining ~64% of the variance) was significantly different to the intercept, with every one unit increase in PC1 accounting for an increase in foraging distance of ~410 meters. This PC correlates positively with arable and OSR lands, and negatively correlated with pasture and built up areas.

One potential issue with this analysis could be that land-use types change in floral availability depending on the season. For example, gardens and parks may have more planted flowers in Spring than Summer. It is therefore prudent to investigate if season might influence foraging distance and so may influence any effects of land-use type on average foraging distance.

# Analysis over time

Unlike in the analysis of the proportion of scouts, where the model estimates are limited by the need for a significant number of data points, we can calculate an average foraging distance at practically any time scale. To break down the data, I will now repeat the PLS fitting for rural and urban sites for each years season to identify if any effect of land-use can be identified at this scale.

```{r, foraging-distance-over-time, message = FALSE, warning = FALSE}
# add season column
df <- df %>%
  mutate(
    season = ifelse(
      Month %in% seq(1, 3),
      "Winter",
      ifelse(
        Month %in% seq(4, 6),
        "Spring",
        ifelse(
          Month %in% seq(7, 9),
          "Summer",
          "Autumn"
        )
      )
    )
  )

# calculate mean foraging distances for each site
df.median.season.fd <- df %>%
  group_by(LU, site, Year, season) %>%
  summarise(Avg.foraging.distance = median(ForagingDistance_km))

# add land-use types to the foraging data
rural.df <- df.median.season.fd %>%
  filter(LU == "Rural") %>%
  inner_join(rural_lu, on = site)

urban.df <- df.median.season.fd %>%
  filter(LU == "Urban") %>%
  inner_join(urban_lu, on = site)

# plot mean foraging distances for each environment
df.median.season.fd %>%
  ggplot(aes(x = LU, y = Avg.foraging.distance)) +
  geom_boxplot() +
  geom_point() +
  ylab("Median foraging distance (km)") +
  xlab("Environment") +
  facet_wrap(~ Year + season)
```

The trend of further foraging in agri-rural environments is consistent at the season level, however, between landscape type there doesn't appear to be much of a difference. Agrirural foraging distances in spring have a similar spread as to those in summer and the same appears to be the case for urban also.
```{r, foraging-diatance-site-season, message = FALSE, warning = FALSE}
df.median.season.fd %>%
  ggplot(aes(x = site, y = Avg.foraging.distance, fill = season)) +
  geom_bar(stat = "identity", position = "dodge") +
  ylab("Median foraging distance (km)") +
  xlab("Site") +
  facet_wrap(~LU, ncol = 1, scale = "free")

```

When we break this down by site we see some differences, but their doesn't seem to be that much in it (however, this is missing confidence intervals).

Lets do a PLS on each one to see if any landscape factors emerge once we control for seasons. There only a few data points for 2016 so I will not include this and instead focus on 2017.

## Urban by season

### Spring 2017
```{r urban_spring_pls, message = FALSE, warning = FALSE}

# subset data by season and year and prepare for the model
urban.df$year_season <- paste(urban.df$season, urban.df$Year, sep = "_")
seasonal.foraging.df <- purrr::map(
  unique(urban.df$year_season),
  ~ {
    urban.df %>%
      filter(year_season == .x) %>%
      ungroup() %>%
      select(-LU, -site, -Year, -season, -year_season)
  }
)
names(seasonal.foraging.df) <- unique(urban.df$year_season)

# run model on spring 2017
model.spring.2017 <- plsRglm(
  Avg.foraging.distance ~ .,
  data = seasonal.foraging.df$Spring_2017,
  nt = ncol(seasonal.foraging.df$Spring_2017) - 1,
  verbose = FALSE
)

print(model.spring.2017)
```

In spring 2017 PCs 1 to 3 collectively explain 80% of the variation in foraging distances across the sites.

```{r, urban_spring_p3, message = FALSE, warning = FALSE}
plsdata <- data.frame(
  Avg.foraging.distance = model.spring.2017$dataY,
  model.spring.2017$tt
)

plsmodel.pc1_3 <- lm(
  Avg.foraging.distance ~ Comp_1 + Comp_2 + Comp_3,
  plsdata
)

summary(plsmodel.pc1_3)

newdat <- data_frame(
  Comp_1 = seq(-50, 140, 0.1),
  Comp_2 = seq(-50, 140, 0.1),
  Comp_3 = seq(-50, 140, 0.1),
)

newdat$fit <- predict(plsmodel.pc1_3, newdat, type = "response")

plotdata <- plsdata

ggplot(
  data = plotdata,
  aes(x = Comp_1 + Comp_2 + Comp_3, y = Avg.foraging.distance)
) +
  geom_point() +
  geom_line(
    data = newdat, aes(x = Comp_1 + Comp_2 + Comp_3, y = fit)
  ) +
  xlim(min(-4), max(4)) +
  ylim(0, 4)
```

PC1 looks like the only one significant so I will isolate this now.

```{r, urban_spring_p1, message = FALSE, warning = FALSE}
plsdata <- data.frame(
  Avg.foraging.distance = model.spring.2017$dataY,
  model.spring.2017$tt
)

plsmodel.pc1 <- lm(
  Avg.foraging.distance ~ Comp_1,
  plsdata
)

summary(plsmodel.pc1)

newdat <- data_frame(
  Comp_1 = seq(-50, 140, 0.1),
)

newdat$fit <- predict(plsmodel.pc1, newdat, type = "response")

plotdata <- plsdata

ggplot(
  data = plotdata,
  aes(x = Comp_1, y = Avg.foraging.distance)
) +
  geom_point() +
  geom_line(
    data = newdat, aes(x = Comp_1, y = fit)
  ) +
  xlim(min(-2), max(2)) +
  ylim(0, 4)

plsmodel.pc1_2 <- lm(
  Avg.foraging.distance ~ Comp_1 + Comp_2,
  plsdata
)

lrtest(plsmodel.pc1, plsmodel.pc1_2, plsmodel.pc1_3)
```

Its close but PC1 on its own appears to be the best model from the LRT.

Here we get a significant positive relationship between PC1 and the foraging distance of urban honeybees. For every 1 unit increase in PC1, foraging distance increase by ~300m. This PC correlates positively with continuos central land and railway whilst correlating negatively with parks allotments and cemeteries and woodland.

```{r, urban_spring_loadings, message = FALSE, warning = FALSE}
print(model.spring.2017$pp[, 1])
sqrt(1 / (ncol(urban.plsdata) - 1))
```

### Summer 2017

```{r, urban_summer_pls, message = FALSE, warning = FALSE}
# run model on summer 2017
model.summer.2017 <- plsRglm(
  Avg.foraging.distance ~ .,
  data = seasonal.foraging.df$Summer_2017,
  nt = ncol(seasonal.foraging.df$Summer_2017) - 1,
  verbose = FALSE
)

print(model.summer.2017)
```

PC1 accounts for ~41% of the variance in average foraging distance. Adding more PCs doesn't do a lot to this, will all eight PCs explaining only 50% of the variance.

```{r, urban_summer_pc1, message = FALSE, warning = FALSE}
plsdata <- data.frame(
  Avg.foraging.distance = model.summer.2017$dataY,
  model.summer.2017$tt
)

plsmodel.pc1 <- lm(
  Avg.foraging.distance ~ Comp_1,
  plsdata
)

summary(plsmodel.pc1)

newdat <- data_frame(
  Comp_1 = seq(-50, 140, 0.1),
)

newdat$fit <- predict(plsmodel.pc1, newdat, type = "response")

plotdata <- plsdata

ggplot(
  data = plotdata,
  aes(x = Comp_1, y = Avg.foraging.distance)
) +
  geom_point() +
  geom_line(
    data = newdat, aes(x = Comp_1, y = fit)
  ) +
  xlim(min(-2), max(3)) +
  ylim(0, 4)
```

Despite only showing a weak relationship ($R^2 = 0.41$) this is non the less significant, with every one unit increase in PC1 corresponding to an increase of ~250 meters in foraging distance.

This PC correlates positively with continuos central land, parks allotments and cemeteries whilst correlating negatively with railway and amenity grassland.

```{r, urban_summer_loadings, message = FALSE, warning = FALSE}
print(model.summer.2017$pp[, 1])
sqrt(1 / (ncol(urban.plsdata) - 1))
```

## Agri-rural by season

### Spring 2017
```{r, ar_spring_pls, message = FALSE, warning = FALSE}

# subset data by season and year and prepare for the model
rural.df$year_season <- paste(rural.df$season, rural.df$Year, sep = "_")
seasonal.foraging.df <- purrr::map(
  unique(rural.df$year_season),
  ~ {
    rural.df %>%
      filter(year_season == .x) %>%
      ungroup() %>%
      select(-LU, -site, -Year, -season, -year_season)
  }
)
names(seasonal.foraging.df) <- unique(rural.df$year_season)

# run model on spring 2017
model.spring.2017 <- plsRglm(
  Avg.foraging.distance ~ .,
  data = seasonal.foraging.df$Spring_2017,
  nt = ncol(seasonal.foraging.df$Spring_2017) - 1,
  verbose = FALSE
)

print(model.spring.2017)
```

The first 2 PCs explain ~88% of the variation in the average foraging distance.

```{r, ar_spring_pc2, message = FALSE, warning = FALSE}
plsdata <- data.frame(
  Avg.foraging.distance = model.spring.2017$dataY,
  model.spring.2017$tt
)

plsmodel.pc1_2 <- lm(
  Avg.foraging.distance ~ Comp_1 + Comp_2,
  plsdata
)

summary(plsmodel.pc1_2)

newdat <- data_frame(
  Comp_1 = seq(-50, 140, 0.1),
  Comp_2 = seq(-50, 140, 0.1),
  Comp_3 = seq(-50, 140, 0.1),
)

newdat$fit <- predict(plsmodel.pc1_2, newdat, type = "response")

plotdata <- plsdata

ggplot(
  data = plotdata,
  aes(x = Comp_1 + Comp_2, y = Avg.foraging.distance)
) +
  geom_point() +
  geom_line(
    data = newdat, aes(x = Comp_1 + Comp_2, y = fit)
  ) +
  xlim(min(-3), max(4)) +
  ylim(0, 4.5)
```

The intercept is significantly different to 0 but only PC1 looks significant. Lets isolate it to have a look.

```{r, ar_spring_pc1, message = FALSE, warning = FALSE}
plsdata <- data.frame(
  Avg.foraging.distance = model.spring.2017$dataY,
  model.spring.2017$tt
)

plsmodel.pc1 <- lm(
  Avg.foraging.distance ~ Comp_1,
  plsdata
)

summary(plsmodel.pc1)

newdat <- data_frame(
  Comp_1 = seq(-50, 140, 0.1),
  Comp_2 = seq(-50, 140, 0.1),
  Comp_3 = seq(-50, 140, 0.1),
)

newdat$fit <- predict(plsmodel.pc1, newdat, type = "response")

plotdata <- plsdata

ggplot(
  data = plotdata,
  aes(x = Comp_1, y = Avg.foraging.distance)
) +
  geom_point() +
  geom_line(
    data = newdat, aes(x = Comp_1, y = fit)
  ) +
  xlim(min(-2), max(2)) +
  ylim(0, 4.5)

lrtest(plsmodel.pc1, plsmodel.pc1_2)
```

PC1 on its own provides the best fit and explains ~83% of the variation in average foraging distance. There is a significant positive relationship where by every one unit increase in PC1 increases the foraging distance by ~640 meters. This PC is characterised by a positive correlation with arable land and Non.agricultural unmanaged green space and a negative correlation with pasture and built up areas.

```{r, ar_spring_loadings, message = FALSE, warning = FALSE}
print(model.spring.2017$pp[, 1])
sqrt(1 / (ncol(rural.plsdata) - 1))
```

### Summer 2017

```{r, ar_summer_pls, message = FALSE, warning = FALSE}
# run model on summer 2017
model.summer.2017 <- plsRglm(
  Avg.foraging.distance ~ .,
  data = seasonal.foraging.df$Summer_2017,
  nt = ncol(seasonal.foraging.df$Summer_2017) - 1,
  verbose = FALSE
)

print(model.summer.2017)

```

The first three PCs collectively explain ~71% of the variation in average foraging distances.

```{r, ar_summer_pc3, message = FALSE, warning = FALSE}
plsdata <- data.frame(
  Avg.foraging.distance = model.summer.2017$dataY,
  model.summer.2017$tt
)

plsmodel.pc1_3 <- lm(
  Avg.foraging.distance ~ Comp_1 + Comp_2 + Comp_3,
  plsdata
)

summary(plsmodel.pc1_3)

newdat <- data_frame(
  Comp_1 = seq(-50, 140, 0.1),
  Comp_2 = seq(-50, 140, 0.1),
  Comp_3 = seq(-50, 140, 0.1),
)

newdat$fit <- predict(plsmodel.pc1_3, newdat, type = "response")

plotdata <- plsdata

ggplot(
  data = plotdata,
  aes(x = Comp_1 + Comp_2 + Comp_3, y = Avg.foraging.distance)
) +
  geom_point() +
  geom_line(
    data = newdat, aes(x = Comp_1 + Comp_2 + Comp_3, y = fit)
  ) +
  xlim(min(-5), max(5)) +
  ylim(0, 4.5)
```

Both the intercept and PC1 looks significant so I will isolate them.

```{r, ar_summer_pc1, message = FALSE, warning = FALSE}
plsdata <- data.frame(
  Avg.foraging.distance = model.summer.2017$dataY,
  model.summer.2017$tt
)

plsmodel.pc1 <- lm(
  Avg.foraging.distance ~ Comp_1,
  plsdata
)

summary(plsmodel.pc1)

newdat <- data_frame(
  Comp_1 = seq(-50, 140, 0.1),
  Comp_2 = seq(-50, 140, 0.1),
  Comp_3 = seq(-50, 140, 0.1),
)

newdat$fit <- predict(plsmodel.pc1, newdat, type = "response")

plotdata <- plsdata

ggplot(
  data = plotdata,
  aes(x = Comp_1, y = Avg.foraging.distance)
) +
  geom_point() +
  geom_line(
    data = newdat, aes(x = Comp_1, y = fit)
  ) +
  xlim(min(-2), max(3)) +
  ylim(0, 4.5)

plsmodel.pc1_2 <- lm(
  Avg.foraging.distance ~ Comp_1 + Comp_2,
  plsdata
)

lrtest(plsmodel.pc1, plsmodel.pc1_2, plsmodel.pc1_3)
```

PC1 on its own looks to be the best model, explaining ~48% of the variation with a positive relationship. Every one unit increase in PC1 increasing foraging distance by ~331 meters. This PC is characterised by a positive correlation with arable land and a negative correlation with woodland, non-agricultural unmanaged green space, built up areas and water.

```{r, ar_summer_loadings, message = FALSE, warning = FALSE}
print(model.summer.2017$pp[, 1])
sqrt(1 / (ncol(rural.plsdata) - 1))
```

# Discussion

Median foraging distance do not seem to differ much between summer and spring for urban and agri-rural hives. Even between sites the difference looks to be minimal. Nevertheless, at this scale for both the urban and agri-rural sites there is a singificant relationship between land-use type and the average foraging distance.

In the urban environment in spring, average foraging distance correlates positively with a single PC, which is described by a positive correlation with continuos central land and railway whilst correlating negatively with parks allotments and cemeteries and woodland. So in landscape with more continous central land and railways (nutrient poor environments) average foraging distance increases, presumably because foragers must travel further to find suitable forage. However, in sites with more parks, allotments and cemeteries and woodland (nutrient rich areas) foragers do not have to travel as far to find enough food. In the summer season, land-use appears to have less of an influence and explains a much lower portion of the variance. Nevertheless, a single PC explaining ~40% of the variance correrlates positively with average foraging distance, but with a shallower slope (0.25 vs 3). This PO correlates positively with continuos central land, parks allotments and cemeteries whilst correlating negatively with railway and amenity grassland. Interestingly, this is the opposite of the spring scenario. This is possibly due to a gap in forage avaliability during this time in parks and allotments, forcing foragers to travel further.

In the agri-rural environment in spring, a single PC was found to explain 83% of the variation in the average foraging distance and had a singificant positive correlation with a steap slope (0.61). This PC correlated positively with arable land and Non.agricultural unmanaged green space and a negative correlation with pasture and built up areas. This makes sense as pasture and built up areas are typically nutrient poor, forcing foragers further to better grounds, whilst arable land and Non.agricultural unmanaged green space appear to provide enough quality forage. Just as with the urban sites, in summer land-use types explain less variation in average foraging distances. A single PC explains ~48% of the variation with a singificant if all beint smaller slope than in spring. This PC, as in the urban case, looks largely like a mirror of the dominant PC in spring. Again, this is possibly due to a gap in forage avaliability during this time in parks and allotments, forcing foragers to travel further.

### 2021-10-23 analysis/landuse_effect_prop_scouts.Rmd
added urban and rural joint PLS graph, needs more adding to it and sending to vincent and elli

@@
 library(ggfortify)
-library(energy)
 library(betareg)
@@
 * It would be nice to add confidence intervals to the model fit plots. Unsure how to do this with a beta regression and not all that sure it is that important?
+
+# Combining the data
+
+```{r}
+
+joined <- bind_rows(rural_lu, urban_lu)
+joined[is.na(joined)] <- 0
+
+joined.pcdata <- joined[, 3:length(joined)]
+
+n <- length(joined$p)
+y <- (joined$p * (n - 1) + 0.5) / n
+joined.pcdata$p_transform <- y
+
+model <- plsRbeta(
+  p_transform ~ .,
+  data = joined.pcdata,
+  nt = ncol(joined.pcdata) - 1,
+  model = "pls-beta"
+)
+
+print(model)
+
+plsdata <- data.frame(
+  prop.scouts = model$dataY,
+  model$tt
+)
+
+plsbetamodel <- betareg(
+  prop.scouts ~ Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4,
+    plsdata
+)
+
+summary(plsbetamodel)
+
+newdat <- data_frame(
+  Comp_.1 = seq(-50, 140, 0.1),
+  Comp_.2 = seq(-50, 140, 0.1),
+  Comp_.3 = seq(-50, 140, 0.1),
+  Comp_.4 = seq(-50, 140, 0.1)
+)
+
+newdat$fit <- predict(plsbetamodel, newdat, type = "response")
+
+plotdata <- plsdata
+plotdata$p <- joined$p
+
+plotdata$LU = ""
+plotdata$LU[1:10] = "Rural"
+plotdata$LU[11:20] = "Urban"
+
+ggplot(data = plotdata,
+  aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = p)) +
+  geom_point(aes(colour = LU)) +
+  geom_rug(aes(colour = LU)) +
+  scale_color_manual(values = c("Rural" = "black", "Urban" = "red")) +
+  geom_line(data = newdat, aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = fit)) +
+  xlim(min(-5.5), max(5.5))
+
+ggplot(data = plotdata,
+  aes(x = LU, y = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4)) +
+  geom_boxplot(aes(group = LU))
+
+```

### 2021-10-23 analysis/landuse_effect_prop_scouts.Rmd
started preentation of work in this analysis folder

@@
   xlim(min(-2.5), max(3.))
+
+ggsave("analysis/results/prop_scout_lu_analysis/rural_pls_p1.png")
 ```
@@
 
-plotdata$LU = ""
-plotdata$LU[1:10] = "Rural"
-plotdata$LU[11:20] = "Urban"
+plotdata$LU <- ""
+plotdata$LU[1:10] <- "Rural"
+plotdata$LU[11:20] <- "Urban"
 
-ggplot(data = plotdata,
-  aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = p)) +
+ggplot(
+  data = plotdata,
+  aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = p)
+) +
   geom_point(aes(colour = LU)) +
@@
 
-ggplot(data = plotdata,
-  aes(x = LU, y = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4)) +
+ggplot(
+  data = plotdata,
+  aes(x = LU, y = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4)
+) +
   geom_boxplot(aes(group = LU))
-
 ```

### 2021-10-23 analysis/presentation.Rmd
started preentation of work in this analysis folder

---
title: "Using the waggle dance"
author: "Joe Palmer"
date: "25/10/2021"
output: ioslides_presentation
widescreen: true
---

<style type="text/css">
.caption {
    font-size: small;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(showtext)
showtext_auto()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(size = 18)
    )
)

library(kableExtra)
library(tidyr)
library(dplyr)
```

## Constructing the waggle dance model | Profitability function

Seeley suggests dancing is regulated by efficiency: $\frac{gain-cost}{cost}$.

- Gain is determined by the nectar quality of the flower, $q$.
- Cost is determined by the distance to the flower, $x$.
- Energy gained is therefore $\propto$ $q$, cost increases linearly with $x$.
- Can therefore refactor efficiency as: $\frac{q}{1+\alpha x} - 1$, where $\alpha$ is the energy spend per unit of distance.
- Energy spend cannot be negative, so the function must be positive:

$$\sigma(q,x) = [\frac{q}{1+\alpha x}-1]_+$$

## Constructing the waggle dance model | Scout distribution

- Assume scouts search an environment without information and report back the first resource found.
- Floral resources should be randomly distributed around a hive and thus follow a Poisson distribution.
- The distance between a central hive and each point should therefore be exponentially distributed, with rate $\lambda_s$.
- This scout distribution is altered by the profitability function, which translates the distance of the resource to a number of dances reported on the dance floor.
- $\sigma(q, x) \lambda_s e^{-\lambda_s x}$

## Constructing the waggle dance model | Recruit distribution

- Recruits sample randomly from the dance on the dance floor (scouts + other recruits).
- Recruits over-represent the best resources, skewing the distribution of dances.
- The best resource is implicitly a function of distance and should bias towards the closest patches to the colony.
- The chance of discovering a food source within a radius of $x$ in a circular area is: $e^{-\lambda_r \pi x^2}$, where $\lambda_r$ is the scout discovery rate.
- The distribution of the nearest food sources that recruits would come for is: $2 \lambda_r \pi x e^{-\lambda_r \pi x^2}$

- $\sigma(q, x) 2 \lambda_r \pi x e^{-\lambda_r \pi x^2}$

## Constructing the waggle dance model | Putting it all together

<center>
$$\sigma(q, x) ( p \lambda_s e^{-\lambda_s x} + (1-p) 2 \lambda_r \pi x e^{-\lambda_r \pi x^2})$$
</center>

- The distribution of reported distances is a superposition of recruit and scout distance distributions.
- $p$ provides a proportional weighting of the number of dances coming from either a scout or recruit strategy.

---

<div style="float: left; width: 50%;">

![](results/figures/foraging-model.png){width=95%, height=100%}

</div>

<div style="float: right; width: 50%;">

<p class = caption>
Model of honeybee foraging. Flowers advertised by scouts are distributed exponentially (A), y axis is the log probability of sampling a value larger than $x$ (same for C and E). These dances are advertised on the dance floor (B) in relation to their profitability, meaning sampling recruits are biased to the more profitable (and closer) resources (C). Recruits also dance for these resources leading to higher recruitment (D) which overall skews the distribution of distances reported on the entire dance floor (E).
</p>

</div>

## Simulation

*Are different foraging strategies identifiable from the distribution of foraging distances reported on the dance floor?*

We created an Agent Based Simulation of a foraging honey bee colony.

1. Hive exists surrounded by resources which have a quality and distance from the hive. The resources are randomly removed and others added for simulate flower turnover.
2. A proportion of foragers act as scouts, leave the hive on a randomly determined path and report back the first patch discovered.
3. The rest act as recruits and sample randomly the pooled dances of scouts and other recruits. Once they have a location they visit that resource and also advertise it.
4. Scouts and recruits return to their flower but once that patch disappears they must return to searching or following dances respectively.

---

![A. Outline of scout foraging. Scouts leave the hive on a random flight path (white lines) and continue until they encounter a resource (flowers in the figure). B. Outline of foraging with recruitment. Scouts continue to identify resources but recruits' sample available dances provided by recruits and other scouts. C. Log inverse frequency distribution of foraging distances reported by scouts. D. Log inverse frequency distribution of foraging distances reported by recruits.](results/figures/simulation.png){width=100%, height=50%}

---

- **Different foraging strategies leave different patterns in the distribution of foraging distances reported on the dance floor**.

- Logic of the model holds water.

## Fit to actual data

*Do colonies in real world environments forage collectively, individually or through some other mechanism?*

- Waggle dance data was collected and decoded from 10 urban and 10 agri-rural sites over the 2017 season for a total of > 3000 dances (Credit: Ash Samuelson).
- Fit our full model to the distribution of foraging distances recorded for each site.
- Also fit a scouts only model ($p$ fixed to 1) to evaluate an individual only foraging strategy.
- Model selection used to determine which strategy was more likely.
- Model fit used to determine if our model offered a good explanation of the data.

---

<div style="float: left; width: 45%;">

![](results/figures/site_map.png){width=95%, height=95%}

<p class = caption>
Location of hive sites across southern England. Black circles indicates the collective model provided the best fit to the data in that site, whilst red triangles indicates the individual model provided a better explanation of the data. The greater London area is indicated by the dark grey shading (inset map).
</p>

</div>

<div style="float: right; width: 50%;">

![](results/figures/results_model_plot.png){width=86%, height=86%}

<p class = caption>
Model fits to foraging distance distributions decoded from waggle dances in sites ZSL (A) and STU (B). The Y axis in A and B is the inverse log-cumulative frequency, indicating the log probability of a foraging distance being greater than or equal to $x$. Inset plots show the frequency of dances advertising foraging distances. C. Number of sites where the collective and individual models provided the best explanation of the data. D. Distribution of KS statistic P values across the sites for the collective (top) and individual (bottom) models. A value greater than 0.05 indicates there is no significant difference between the empirical distribution of the data and the model applied.
</p>

</div>

## Conclusions

- We have successfully translated the dominant theoretical understanding of honeybee foraging into a quantifiable and thus measurable model.
- Evaluated this model against a large empirical dataset to show that in the majority of cases it provides an extremely good explanation of foraging patterns.
- In some cases, a model of individual behaviour is more suitable
- Provides new evidence that honeybees can regulate the use of the waggle dance and use personal rather than social information at a colony level.
- Provides a new, non-invasive  tool to evaluate what factors might influence how honeybees use the waggle dance.

## Latest work (Brand new stuff!)

- Environmental factors are thought to influence the use of the waggle dance and social information more broadly.
- Intensely studied in honeybees using theoretical and empirical studies.
- Empirical methods are open to problems.
- Our model provides a new way of doing things.
- Fitting our model to data gives us a proportion of scout dances $p$, providing an estimate of what the foragers are doing (working alone or following others) based on what they report.

---

```{r, prop-scouts-LU, echo = FALSE, message = FALSE, warning = FALSE}

# load results of model fit (code to generate found in fit_models.Rmd)
sitedata <- readRDS("results/site_fit_results.Rda")

# filter for best model
sitedata <- sitedata %>%
  group_by(site) %>%
  slice(which.min(AIC))

# foraging distance data
df <- read.csv("data/FullHBForagingData.csv")

# join LU on to sitedata
sitedata.lu <- df %>%
  select(site, LU) %>%
  unique() %>%
  inner_join(sitedata, on = "site")

ggplot(sitedata.lu, aes(y = p, x = LU)) +
  geom_boxplot(binwidth = 1, colour = "black")
```

- No real difference in the proportion of scouts between Agri-rural and Urban areas.

## Influence of land-use type on the proportion of scouts

- The foraging land around each site was classified by their different land-use types (Credit: Ash Samuelson)

- Urban: continuous central, dense residential, sparse residential, parks allotments cemeteries, amenity.grassland, Railway, woodland, and water.

- Agri-rural: woodland, Non agricultural unmanaged green space, Non, agricultural managed green space, arable, pasture, other agricultural, fruit, Oil Seed Rape, built up area, and water.

*Is there a relationship between land-use type and proportion of scouts?*

## Partial Least Squares analysis (PLS)

<div style="float: left; width: 50%;">

- Similar to a Principle Component Analysis (PCA).
- PCA finds the major axis of variation in the data and relates this to a response variable.
- PLS finds the major axis of variation in the response variable.
- Allows you to evaluate what land-use types most influence the variation in the proportion of scouts.

</div>

<div style="float: right; width: 50%;">

![Difference between PLS (supervised) and PCA (unsupervised) methods. Taken from Scott and Crone., 2021](https://media.springernature.com/lw685/springer-static/image/art%3A10.1007%2Fs00442-020-04848-w/MediaObjects/442_2020_4848_Fig1_HTML.png?as=webp){width=95%, height=95%}

</div>

## PLS - Agri-rural

<div style="float: left; width: 45%;">

- A single PC explained ~60% of the variation in the proportion of scouts.
- There is significant positive relationship between PC1 and the proportion of scouts (Beta regression: R$^2$ = 0.60, df = 3, p < 0.05).

</div>

<div style="float: right; width: 50%;">

![Effect of PC1 on the proportion of scouts in an agri-rural environment](man/figures/README-rural_pls_p1-1.png){width=95%, height=95%}

</div>

---

<div style="float: left; width: 45%;">

```{r, echo=FALSE,out.width="55%", out.height="10%",fig.cap=""}
land.types <- c(
  "https://stmaaprodfwsite.blob.core.windows.net/assets/sites/1/2015/04/farmland.jpg",
  "https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcSeyllkO-dj0vCIfqb7vSN18gpkiB-7fn74Mw&usqp=CAU",
  "https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcQVDhSwdnZ32zYO4gw39oKxY9QSygjB2gIXJg&usqp=CAU",
  "https://hips.hearstapps.com/hmg-prod.s3.amazonaws.com/images/haweswater-reservoir-cumbria-1531212664.jpg"
)
knitr::include_graphics(land.types)
```

</div>

<div style="float: right; width: 50%;">

- PC1 correlates:
- **positively** with **arable land**
- **negatively** with **Non agricultural unmanaged green space, built up areas and water**.

- Suggests more recruits in resource poor areas and more scouts in resource rich areas.

</div>

## PLS - Urban

<div style="float: left; width: 45%;">

- The first three PCs together explain ~88% of the variation in the proportion of scouts. There is significant positive relationship between these PCs and the proportion of scouts (Beta regression: R$^2$ = 0.88, df = 5, p < 0.05).

</div>

<div style="float: right; width: 50%;">

![Effect of PC1+PC2+PC3 on the proportion of scouts in an urban environment](man/figures/README-urban_pls_3p-1.png){width=95%, height=95%}

</div>

---

- PC1, explaining ~61% of the variance in the proportion of scouts, correlates:
- **Positively** with continuos central land, dense residential and water.
- **Negatively** with sparse residential and amenity grassland.

**insert pictures**

## Implications

- Land-use type has a significant impact of the proportion of scouts.
- Reinforces the idea that the waggle dance is used flexibly in certain environments.
- Suggests agricultural land pressures colonies to use more scout type strategies.

more implications...finish up.

### 2021-10-23 R/RcppExports.R

@@ -6,7 +6,7 @@
 #' @param x NumericArray
 #' @export
 alter_in_place <- function(x) {
-    invisible(.Call('_wagglefit_alter_in_place', PACKAGE = 'wagglefit', x))
+  invisible(.Call("_wagglefit_alter_in_place", PACKAGE = "wagglefit", x))
 }
 
 #' Model CCDF for scout distribution
@@ -18,7 +18,7 @@ alter_in_place <- function(x) {
 #' @return result double the ccdf for that foraging distance (x)
 #' @export
 scout_ccdf <- function(x, m, bs, as) {
-    .Call('_wagglefit_scout_ccdf', PACKAGE = 'wagglefit', x, m, bs, as)
+  .Call("_wagglefit_scout_ccdf", PACKAGE = "wagglefit", x, m, bs, as)
 }
 
 #' Model CCDF for recruit distribution
@@ -30,7 +30,7 @@ scout_ccdf <- function(x, m, bs, as) {
 #' @return result double the ccdf for that foraging distance (x)
 #' @export
 recruit_ccdf <- function(x, m, br, ar) {
-    .Call('_wagglefit_recruit_ccdf', PACKAGE = 'wagglefit', x, m, br, ar)
+  .Call("_wagglefit_recruit_ccdf", PACKAGE = "wagglefit", x, m, br, ar)
 }
 
 #' Model ccdf function for collective model. Stores results in
@@ -43,7 +43,7 @@ recruit_ccdf <- function(x, m, br, ar) {
 #' @inheritParams recruit_ccdf
 #' @export
 ccdf_model_collective <- function(x, y, p, bs, br, as, ar) {
-    invisible(.Call('_wagglefit_ccdf_model_collective', PACKAGE = 'wagglefit', x, y, p, bs, br, as, ar))
+  invisible(.Call("_wagglefit_ccdf_model_collective", PACKAGE = "wagglefit", x, y, p, bs, br, as, ar))
 }
 
 #' Model ccdf function for individual model. Stores results in given array (y)
@@ -53,7 +53,7 @@ ccdf_model_collective <- function(x, y, p, bs, br, as, ar) {
 #' @inheritParams scout_ccdf
 #' @export
 ccdf_model_individual <- function(x, y, bs, as) {
-    invisible(.Call('_wagglefit_ccdf_model_individual', PACKAGE = 'wagglefit', x, y, bs, as))
[…]

### 2021-10-23 src/RcppExports.cpp

@@ -5,11 +5,6 @@
 
 using namespace Rcpp;
 
-#ifdef RCPP_USE_GLOBAL_ROSTREAM
-Rcpp::Rostream<true>&  Rcpp::Rcout = Rcpp::Rcpp_cout_get();
-Rcpp::Rostream<false>& Rcpp::Rcerr = Rcpp::Rcpp_cerr_get();
-#endif
-
 // alter_in_place
 void alter_in_place(NumericVector x);
 RcppExport SEXP _wagglefit_alter_in_place(SEXP xSEXP) {

### 2021-10-24 analysis/presentation.Rmd
updated presentation

@@
 ---
-title: "Using the waggle dance"
+title: "Modelling the honeybee waggle dance"
 author: "Joe Palmer"
@@
     font-size: small;
+}
+</style>
+
+<style>
+body {
+    overflow: scroll;
 }
@@
 
+## How a collective behaviour works
+
+- Foraging is a complex task.
+- Social insects have evolved elaborate cooperative strategies to increase foraging success & efficiency.
+- Self organisation pools and weighs individual information to produce an informed collective response
+
+<div style="float: left; width: 49%;">
+![Flocking starlings](https://headguruteacher.files.wordpress.com/2019/11/wales-starling-murmuration.jpg){height=94%, width=94%}
+
+</div>
+
+<div style="float: right; width: 49%;">
+![*Lasius niger*](https://www.antwiki.org/wiki/images/thumb/5/54/Lasius_cf._niger%2C_Michal_Kukla_%282%29.JPG/450px-Lasius_cf._niger%2C_Michal_Kukla_%282%29.JPG)
+
+</div>
+
+## Honeybee foraging
+
+- Evolved a unique dance language to provide the distance and direction of a food source to nest-mates
+- Studied intensively to uncover the rules governing dancing
+
+1. Foragers dance in relation to foraging efficiency.
+2.
+Foragers use the time taken to find a worker to unload them of nectar when returning from foraging as an indicator of how desperate the colony is for nectar, and factored this into their dancing enthusiasm.
+3. Recruits don't compare dances.
+
+## Honeybee foraging
+
+- Combined, these rules are hypothesised to work like other collective behaviours.
+
+- Scouts explore the landscape for forage and perform dance runs based on foraging efficiency
+- Recruits, sampling randomly, are biased to the best resources
+- Recruits return to dance just like scouts
+- Individual assessments are pooled on the dance-floor
+- Collective finds the best resources
+- Individual parts have significant empirical support
+- Combine to produce a coherent theory widely accepted to be true
+
+**Overall hypothesis has not been empirically validated or tested!**
+
+## Why do we need to evaluate this theory?
+
+- Increasing attention on studying the waggle dance as a measure of social Vs collective information.
+- Key question in collective behaviour is what conditions lead to the evolution of sociality and cooperation?
+- Knowing how much the waggle dance is used provides an estimate of how much a colony values social vs personal information.
+- Multiple theoretical and empirical studies have reached conflicting conclusions about the importance of the waggle dance.
+- New methods are needed to better understand the waggle dance and hows its used.
+
+## Objective
+
+- Construct a model of honeybee foraging based off the generally accepted narrative
+- Contrast this with an individual based foraging model
+- Evaluate these models on real foraging data
+
 ## Constructing the waggle dance model | Profitability function
@@
 
-## Constructing the waggle dance model | Scout distribution
+### Scout distribution
 
@@
 - This scout distribution is altered by the profitability function, which translates the distance of the resource to a number of dances reported on the dance floor.
-- $\sigma(q, x) \lambda_s e^{-\lambda_s x}$
 
-## Constructing the waggle dance model | Recruit distribution
+$$\sigma(q, x) \lambda_s e^{-\lambda_s x}$$
+
+### Recruit distribution
 
@@
 
-- $\sigma(q, x) 2 \lambda_r \pi x e^{-\lambda_r \pi x^2}$
+$$\sigma(q, x) 2 \lambda_r \pi x e^{-\lambda_r \pi x^2}$$
 
-## Constructing the waggle dance model | Putting it all together
-
-<center>
-$$\sigma(q, x) ( p \lambda_s e^{-\lambda_s x} + (1-p) 2 \lambda_r \pi x e^{-\lambda_r \pi x^2})$$
-</center>
-
-- The distribution of reported distances is a superposition of recruit and scout distance distributions.
-- $p$ provides a proportional weighting of the number of dances coming from either a scout or recruit strategy.
-
----
+### Putting it all together
 
@@
 <div style="float: right; width: 50%;">
+
+<center>
+$$\sigma(q, x) ( p \lambda_s e^{-\lambda_s x} + (1-p) 2 \lambda_r \pi x e^{-\lambda_r \pi x^2})$$
+</center>
+
+- The distribution of reported distances is a superposition of recruit and scout distance distributions.
+- $p$ provides a proportional weighting of the number of dances coming from either a scout or recruit strategy.
 
@@
 
+- Waggle dance use tells us how social and personal information is valued
 - Environmental factors are thought to influence the use of the waggle dance and social information more broadly.
 - Intensely studied in honeybees using theoretical and empirical studies.
-- Empirical methods are open to problems.
-- Our model provides a new way of doing things.
+- Empirical studies rely on hive disorientation
+
+- Tells us if a colony does well with or without a dance in a given environment
+- Can't tell us if a colony actually responds to the environment by altering their foraging strategies
+
+- Our model provides a new way of approaching this question.
 - Fitting our model to data gives us a proportion of scout dances $p$, providing an estimate of what the foragers are doing (working alone or following others) based on what they report.
+
+---
+
+- In addition to understanding what conditions gave rise to the evolution of social foraging, there is significant attention on understanding how anthropogenic activities are impacting social insects.
+- Urbanisation and expanding Agricultural practices represent two of the biggest land-use changes.
+- How are these landscapes influencing how the waggle dance is being used?
 
@@
 
-<div style="float: left; width: 45%;">
+<div style="float: left; width: 25%;">
 
-```{r, echo=FALSE,out.width="55%", out.height="10%",fig.cap=""}
+```{r, echo=FALSE,out.width="65%", out.height="10%",fig.cap=""}
 land.types <- c(
@@
 
-<div style="float: right; width: 50%;">
+<div style="float: right; width: 75%;">
 
@@
 
-- Suggests more recruits in resource poor areas and more scouts in resource rich areas.
+- Arable land are typically deserts for pollinators
+- Unmanaged green space has space for wild flowers
+- agri-rural built up areas typically have green space in the form of gardens
+- water ways typically have lots of flora on river and pond banks
+
+Suggests when resources are sparse and difficult to find, a colony responds by increasing the proportion of scouts to explore the landscape.
 
@@
 - The first three PCs together explain ~88% of the variation in the proportion of scouts.
-There is significant positive relationship between these PCs and the proportion of scouts (Beta regression: R$^2$ = 0.88, df = 5, p < 0.05).
+- There is significant positive relationship between these PCs and the proportion of scouts (Beta regression: R$^2$ = 0.88, df = 5, p < 0.05).
 
@@
 
-- PC1, explaining ~61% of the variance in the proportion of scouts, correlates:
+- PC1, explains ~61% of the variance in the proportion of scouts, correlates:
 - **Positively** with continuos central land, dense residential and water.
@@
 
-**insert pictures**
+- Continuos central land typically has very few resources for pollinators.
+- Dense residential areas have fewer green spaces such as gardens.
+- Urban waterways have fewer plants available.
+- Sparse residential has more area for gardens.
+- Amenity grassland typically managed, but has benefited from recent drives to increase green boarders and 'wild zones'.
+
+Suggests when resources are sparse and difficult to find, a colony responds by increasing the proportion of scouts to explore the landscape.
 
@@
 - Reinforces the idea that the waggle dance is used flexibly in certain environments.
-- Suggests agricultural land pressures colonies to use more scout type strategies.
+- In both Agri-rural and Urban landscapes, resource poor land-use types correlate with an increase in the proportion of scouts.
+- Both the urban and agri-rural analysis suggests anthropogenic activity selects colonies to ignore the waggle dance more.
 
-more implications...finish up.
+## Future directions
+
+- Collect more data!
+- Invest in new technologies such as automatic dance decoding software
+- Generate an automated pipeline to analyse honeybee waggle dance data from our methods.
+
+## Thank you for listening
+### Acknowledgments
+- Vincent Jansen
+- Elli Leadbeater
+- Ash Samuelson
+
+### References

### 2021-10-25 analysis/presentation.Rmd
finished final draft of presentation for Gill-lab

--- 2021-10-23 analysis/presentation.Rmd
@@
 ---
-title: "Using the waggle dance"
+title: "Modelling the honeybee waggle dance"
 author: "Joe Palmer"
 date: "25/10/2021"
-output: ioslides_presentation
+output:
+  ioslides_presentation:
     widescreen: true
+    incremental: true
+bibliography: reffs.bib
+
 ---
@@
 .caption {
+    font-size: small;
+}
+.refs {
     font-size: small;
@@
 
-## Constructing the waggle dance model | Profitability function
+## Collective behaviour through self-organisation
 
-Seeley suggests dancing is regulated by efficiency: $\frac{gain-cost}{cost}$.
+- Social insects often live in groups of hundreds or even thousands strong
+- Successfully coordinate themselves to work as a cohesive unit
+- Not a directed process, occurs through self-organisation
+- All individuals follow a given set of rules and a collective pattern **emerges**
+- Self organisation pools and weighs individual information to produce an informed collective response [@Sumpter2006,Myerscough2003,Franks2002].
 
-- Gain is determined by the nectar quality of the flower, $q$.
-- Cost is determined by the distance to the flower, $x$.
-- Energy gained is therefore $\propto$ $q$, cost increases linearly with $x$.
-- Can therefore refactor efficiency as: $\frac{q}{1+\alpha x} - 1$, where $\alpha$ is the energy spend per unit of distance.
-- Energy spend cannot be negative, so the function must be positive:
+## Honeybee foraging
 
-$$\sigma(q,x) = [\frac{q}{1+\alpha x}-1]_+$$
+>- Evolved a unique dance language to provide the distance and direction of a food source to nest-mates
 
-## Constructing the waggle dance model | Scout distribution
+>- Studied intensively to uncover the rules governing dancing [@Seeley1995]
 
-- Assume scouts search an environment without information and report back the first resource found.
-- Floral resources should be randomly distributed around a hive and thus follow a Poisson distribution.
-- The distance between a central hive and each point should therefore be exponentially distributed, with rate $\lambda_s$.
-- This scout distribution is altered by the profitability function, which translates the distance of the resource to a number of dances reported on the dance floor.
-- $\sigma(q, x) \lambda_s e^{-\lambda_s x}$
+<div style="float: left; width: 49%;">
 
-## Constructing the waggle dance model | Recruit distribution
-
-- Recruits sample randomly from the dance on the dance floor (scouts + other recruits).
-- Recruits over-represent the best resources, skewing the distribution of dances.
-- The best resource is implicitly a function of distance and should bias towards the closest patches to the colony.
-- The chance of discovering a food source within a radius of $x$ in a circular area is: $e^{-\lambda_r \pi x^2}$, where $\lambda_r$ is the scout discovery rate.
-- The distribution of the nearest food sources that recruits would come for is: $2 \lambda_r \pi x e^{-\lambda_r \pi x^2}$
-
-- $\sigma(q, x) 2 \lambda_r \pi x e^{-\lambda_r \pi x^2}$
-
-## Constructing the waggle dance model | Putting it all together
-
-<center>
-$$\sigma(q, x) ( p \lambda_s e^{-\lambda_s x} + (1-p) 2 \lambda_r \pi x e^{-\lambda_r \pi x^2})$$
-</center>
-
-- The distribution of reported distances is a superposition of recruit and scout distance distributions.
-- $p$ provides a proportional weighting of the number of dances coming from either a scout or recruit strategy.
-
----
-
-<div style="float: left; width: 50%;">
-
-![](results/figures/foraging-model.png){width=95%, height=100%}
+![waggle dance](https://thumbs.gfycat.com/WarySlowArizonaalligatorlizard-max-14mb.gif)
 
@@
 
-<div style="float: right; width: 50%;">
+<div style="float: right; width: 49%;">
 
-<p class = caption>
-Model of honeybee foraging.
-Flowers advertised by scouts are distributed exponentially (A), y axis is the log probability of sampling a value larger than $x$ (same for C and E).
-These dances are advertised on the dance floor (B) in relation to their profitability, meaning sampling recruits are biased to the more profitable (and closer) resources (C).
-Recruits also dance for these resources leading to higher recruitment (D) which overall skews the distribution of distances reported on the entire dance floor (E).
-</p>
+1. Foragers dance in relation to foraging efficiency [@Seeley1986,Seeley1994A].
+2. Foragers use the time taken to find a worker to unload them of nectar when returning from foraging as an indicator of how desperate the colony is for nectar [@Seeley1986].
+3. Recruits don't compare dances [@Seeley1992].
 
 </div>
+
+## Honeybee foraging
+
+- Combined, these rules are hypothesised to work as a self-organised process [@Seeley1995].
+
+- Scouts explore the landscape for forage and perform dance runs based on foraging efficiency
+- Recruits, sampling randomly, are biased to the best resources
+- Recruits return to dance just like scouts
+- Individual assessments are pooled on the dance-floor
+- Collective finds the best resources
+- Individual parts have significant empirical support
+- Combine to produce a coherent theory widely accepted to be true
+
+- **Overall hypothesis has not been empirically validated or tested!**
+
+## Why do we need to evaluate this theory?
+
+- Key question in collective behaviour is what conditions lead to the evolution of sociality and cooperation? [@King2007,Sumpter2006,Gordon2016]
+- Knowing how the waggle dance is used indicates how much a colony values social vs personal information [@IAnsonPrice2015].
+- Multiple theoretical [@Dornhaus2006,Beekman2008] and empirical studies [@Gruter2009,Gruter2011,Wray2012,Schurch2014,Price2019] have reached conflicting conclusions about the importance of the waggle dance.
+- New methods are needed to better understand the waggle dance and how it is used.
+
+## Objective
+
+- Construct a model of honeybee foraging based off the generally accepted narrative
+- Contrast this with an individual based foraging model
+- Evaluate these models on real foraging data
+
+## Where to start?
+
+- Impossible to track individual behaviour to model data against
+- Does a particular foraging behaviour leave an imprint on the dance-floor?
+- If scouts and recruits forage differently, will this difference reflect in the pattern of waggle dances each group performs?
 
@@
 We created an Agent Based Simulation of a foraging honey bee colony.
-
-1. Hive exists surrounded by resources which have a quality and distance from the hive. The resources are randomly removed and others added for simulate flower turnover.
-2. A proportion of foragers act as scouts, leave the hive on a randomly determined path and report back the first patch discovered.
-3. The rest act as recruits and sample randomly the pooled dances of scouts and other recruits. Once they have a location they visit that resource and also advertise it.
-4. Scouts and recruits return to their flower but once that patch disappears they must return to searching or following dances respectively.
 
@@
 
-- **Different foraging strategies leave different patterns in the distribution of foraging distances reported on the dance floor**.
+<center>
+**Different foraging strategies leave different patterns in the distribution of foraging distances reported on the dance floor**.
+</center>
 
-- Logic of the model holds water.
+## Constructing the waggle dance model
+
+- Assume scouts search an environment without information and report back the first resource found. **$\lambda_s e^{-\lambda_s x}$**
+- Recruits sample randomly from the dance floor (scouts + other recruits).
+- Recruits over-represent the best resources, skewing the distribution of dances. **$2 \lambda_r \pi x e^{-\lambda_r \pi x^2}$**
+- Seeley suggests dancing is regulated by efficiency: $\frac{gain-cost}{cost}$.
+- Energy gained is $\propto$ quality $q$, cost increases linearly with distance, $x$.
+- Can therefore refactor efficiency as: $\frac{q}{1+\alpha x} - 1$, where $\alpha$ is the energy spend per unit of distance.
+
+## Constructing the waggle dance model | Putting it all together
+
+<center>
+$$\sigma(q, x) ( p \lambda_s e^{-\lambda_s x} + (1-p) 2 \lambda_r \pi x e^{-\lambda_r \pi x^2})$$
+</center>
+
+<br>
+
+>- The distribution of reported distances is a superposition of recruit and scout distance distributions.
+>- $p$ provides a proportional weighting of the number of dances coming from either a scout or recruit strategy.
 
@@
 - Fit our full model to the distribution of foraging distances recorded for each site.
-- Also fit a scouts only model ($p$ fixed to 1) to evaluate an individual only foraging strategy.
+- Also fit an individual only model ($p$ fixed to 1) to evaluate an individual only foraging strategy.
 - Model selection used to determine which strategy was more likely.
@@
 
+- Waggle dance use tells us how social and personal information is valued
 - Environmental factors are thought to influence the use of the waggle dance and social information more broadly.
 - Intensely studied in honeybees using theoretical and empirical studies.
-- Empirical methods are open to problems.
-- Our model provides a new way of doing things.
+- Empirical studies rely on hive disorientation
+- Tells us if a colony does well with or without the dance in a given environment
+- Can't tell us if a colony actually **responds** to the environment by altering their foraging strategies
+
+- Our model provides a new way of evaluating how external factors might inlfunece the use of the waggle dance.
 - Fitting our model to data gives us a proportion of scout dances $p$, providing an estimate of what the foragers are doing (working alone or following others) based on what they report.
+
+---
+
+>- In addition to understanding what conditions gave rise to the evolution of social foraging, there is significant attention on understanding how anthropogenic activities are impacting social insects.
+>- Urbanisation and expanding Agricultural practices represent two of the biggest land-use changes imposed by humans.
+>- Understanding what effect this has on the use of social information is an important question.
+
+<div style="float: left; width: 50%;">
+
+![](https://www.pbctoday.co.uk/news/wp-content/uploads/2019/11/thumbnail_image010-300x200.jpg){width=95%, height=95%}
+
+<p class = caption>
+Urbanisation.
+</p>
+
+</div>
+
+<div style="float: right; width: 50%;">
+
+![](https://upload.wikimedia.org/wikipedia/commons/b/b8/Claas_Tucano_430.jpg){width=84%, height=84%}
+
+<p class = caption>
+Agricultural expansion
+</p>
+
+</div>
 
@@
 
+*Is there a relationship between land-use type and proportion of scouts?*
+
 - The foraging land around each site was classified by their different land-use types (Credit: Ash Samuelson)
-
 - Urban: continuous central, dense residential, sparse residential, parks allotments cemeteries, amenity.grassland, Railway, woodland, and water.
-
 - Agri-rural: woodland, Non agricultural unmanaged green space, Non, agricultural managed green space, arable, pasture, other agricultural, fruit, Oil Seed Rape, built up area, and water.
-
-*Is there a relationship between land-use type and proportion of scouts?*
 
@@
 
-- Similar to a Principle Component Analysis (PCA).
-- PCA finds the major axis of variation in the data and relates this to a response variable.
-- PLS finds the major axis of variation in the response variable.
-- Allows you to evaluate what land-use types most influence the variation in the proportion of scouts.
+>- Similar to a Principle Component Analysis (PCA).
+>- PCA finds the major axis of variation in the data and relates this to a response variable.
+>- PLS finds the major axis of variation in the response variable.
+>- Allows you to evaluate what land-use types most influence the variation in the proportion of scouts.
 
@@
 
-- A single PC explained ~60% of the variation in the proportion of scouts.
-- There is significant positive relationship between PC1 and the proportion of scouts (Beta regression: R$^2$ = 0.60, df = 3, p < 0.05).
+>- A single PC explained ~60% of the variation in the proportion of scouts.
+>- There is significant positive relationship between PC1 and the proportion of scouts (Beta regression: R$^2$ = 0.60, df = 3, p < 0.05).
 
@@
 
-<div style="float: right; width: 50%;">
+<div style="float: right; width: 75%;">
 
-- PC1 correlates:
-- **positively** with **arable land**
-- **negatively** with **Non agricultural unmanaged green space, built up areas and water**.
+>- PC1 correlates:
+>- **positively** with **arable land**
+>- **negatively** with **Non agricultural unmanaged green space, built up areas and water**.
 
-- Suggests more recruits in resource poor areas and more scouts in resource rich areas.
+- Arable land are typically deserts for pollinators
+- Unmanaged green space has space for wild flowers
+- agri-rural built up areas typically have green space in the form of gardens
+- water ways typically have lots of flora on river and pond banks
+- Suggests when resources are sparse and difficult to find, a colony responds by increasing the proportion of scouts to explore the landscape.
 
@@
 
-- The first three PCs together explain ~88% of the variation in the proportion of scouts.
-There is significant positive relationship between these PCs and the proportion of scouts (Beta regression: R$^2$ = 0.88, df = 5, p < 0.05).
+>- The first PC explains ~61% of the variation in the proportion of scouts.
+>- There is significant positive relationship between this PC and the proportion of scouts (Beta regression: R$^2$ = 0.61, df = 3, p < 0.05).
 
@@
 
-![Effect of PC1+PC2+PC3 on the proportion of scouts in an urban environment](man/figures/README-urban_pls_3p-1.png){width=95%, height=95%}
+![Effect of PC1 on the proportion of scouts in an urban environment](man/figures/README-urban_pls_p1-1.png){width=95%, height=95%}
 
@@
 
-- PC1, explaining ~61% of the variance in the proportion of scouts, correlates:
-- **Positively** with continuos central land, dense residential and water.
-- **Negatively** with sparse residential and amenity grassland.
+<div style="float: left; width: 25%;">
 
-**insert pictures**
+```{r, echo=FALSE,out.width="65%", out.height="10%",fig.cap=""}
+land.types <- c(
+  "https://olympia.london/sites/default/files/styles/main_full_width__mobile/public/local-area/High_street_kensington_olympia_Main.jpg?itok=7wzMLoSz",
+  "https://www.telegraph.co.uk/content/dam/property/2016/09/28/1507111_Views_From_The_Barbican_Estate_On_Its_40th_AnniversaryLONDON_ENGLAND_-_JULY_07__A_general_vi_trans_NvBQzQNjv4BqyILZCJqFaBEpuz3ivC4GStDyCNk0iIla_0j06bseVEE.jpg",
+  "https://architizer-prod.imgix.net/media/1478153409182project_london-river-park_07.jpg?fit=max&w=1680&q=60&auto=format&auto=compress&cs=strip",
+  "https://labmonline.co.uk/wp-content/uploads/2020/03/MHCLG-eviction-ban.jpg",
+  "https://i.guim.co.uk/img/media/913e25493e4d9b4cd1d2e7e535639a11765285a3/0_0_4032_3024/master/4032.jpg?width=445&quality=45&auto=format&fit=max&dpr=2&s=ff60726e82d18bb4c8af7e696e74cc7c"
+)
+knitr::include_graphics(land.types)
+```
+
+</div>
+
+<div style="float: right; width: 75%;">
+>- PC1, explains ~61% of the variance in the proportion of scouts, correlates:
+>- **Positively** with continuos central land, dense residential and water.
+>- **Negatively** with sparse residential and amenity grassland.
+
+- Continuos central typically has very few resources for pollinators.
+- Dense residential areas have fewer gardens.
+- Urban waterways have fewer flora.
+- Sparse residential has more area for gardens.
+- Amenity grassland have green boarders and space for flowers.
+
+- Suggests when resources are sparse and difficult to find, a colony responds by increasing the proportion of scouts to explore the landscape.
+
+</div>
 
@@
 - Land-use type has a significant impact of the proportion of scouts.
-- Reinforces the idea that the waggle dance is used flexibly in certain environments.
-- Suggests agricultural land pressures colonies to use more scout type strategies.
+- Reinforces the idea that the waggle dance is used flexibly.
+- In both Agri-rural and Urban landscapes, resource poor land-use types correlate with an increase in the proportion of scouts.
+- Both the urban and agri-rural analysis suggests anthropogenic activity selects colonies to ignore the waggle dance more.
+- These findings support previous theoretical studies [@Dornhaus2006]
+- Provides new insight into how human activity influences social information use
 
-more implications...finish up.
+## Future directions
+
+- Collect more data!
+- Invest in new technologies such as automatic dance decoding software
+- Generate an automated pipeline to analyse honeybee waggle dance data from our methods.
+
+## Thank you for listening, Questions?
+
+### Acknowledgments
+Vincent Jansen, Elli Leadbeater, Ash Samuelson
+
+### References
+<div id = "refs" class = "refs columns-2"></div>
+
+---
+
+<div style="float: right; width: 50%;">
+![Collective choice of a food source in ants: Taken from [@Detrain2006]](assets/ant-foraging-fig.jpg){width=100%, height=100%}
+</div>
+
+<div style="float: left; width: 50%;">
+![*Lasius niger*](https://www.antwiki.org/wiki/images/thumb/5/54/Lasius_cf._niger%2C_Michal_Kukla_%282%29.JPG/450px-Lasius_cf._niger%2C_Michal_Kukla_%282%29.JPG)
+
+- Self organisation pools and weighs individual information to produce an informed collective response [@Sumpter2006].
+- A similar principle is also observed in honeybees when searching for a new nest [@Myerscough2003,Franks2002]
+
+</div>
+
+---
+
+1. Hive exists surrounded by resources which have a quality and distance from the hive. The resources are randomly removed and others added for simulate flower turnover.
+2. A proportion of foragers act as scouts, leave the hive on a randomly determined path and report back the first patch discovered.
+3. The rest act as recruits and sample randomly the pooled dances of scouts and other recruits. Once they have a location they visit that resource and also advertise it.
+4. Scouts and recruits return to their flower but once that patch disappears they must return to searching or following dances respectively.
+
+---
+
+## Constructing the waggle dance model | Scout distribution {.build}
+
+- Assume scouts search an environment without information and report back the first resource found.
+- Floral resources should be randomly distributed around a hive and thus follow a Poisson distribution.
+- The distance between a central hive and each point should therefore be exponentially distributed, with rate $\lambda_s$.
+
+<br>
+$$\lambda_s e^{-\lambda_s x}$$
+
+## Constructing the waggle dance model | Recruit distribution {.build}
+
+- Recruits sample randomly from the dance on the dance floor (scouts + other recruits).
+- Recruits over-represent the best resources, skewing the distribution of dances.
+- The best resource is implicitly a function of distance and should bias towards the closest patches to the colony.
+- The chance of discovering a food source within a radius of $x$ in a circular area is: $e^{-\lambda_r \pi x^2}$, where $\lambda_r$ is the recruit discovery rate.
+- The distribution of the nearest food sources that recruits would come for is: $2 \lambda_r \pi x e^{-\lambda_r \pi x^2}$
+
+$$2 \lambda_r \pi x e^{-\lambda_r \pi x^2}$$
+
+## Constructing the waggle dance model | Profitability function {.build}
+
+Seeley suggests dancing is regulated by efficiency: $\frac{gain-cost}{cost}$.
+
+- Gain is determined by the nectar quality of the flower, $q$.
+- Cost is determined by the distance to the flower, $x$.
+- Energy gained is therefore $\propto$ $q$, cost increases linearly with $x$.
+- Can therefore refactor efficiency as: $\frac{q}{1+\alpha x} - 1$, where $\alpha$ is the energy spend per unit of distance.
+- Number of dances cannot be negative, so the function must be positive:
+
+$$\sigma(q,x) = [\frac{q}{1+\alpha x}-1]_+$$
+
+## Constructing the waggle dance model | Putting it all together
+
+<center>
+$$\sigma(q, x) ( p \lambda_s e^{-\lambda_s x} + (1-p) 2 \lambda_r \pi x e^{-\lambda_r \pi x^2})$$
+</center>
+
+<br>
+
+>- The distribution of reported distances is a superposition of recruit and scout distance distributions.
+>- $p$ provides a proportional weighting of the number of dances coming from either a scout or recruit strategy.

### 2021-10-28 analysis/install-analysis-packages.R
added script to search analysis for packages and install if not installed

@@ -1,0 +1,16 @@
+library(stringr)
+library_calls <- purrr::map(
+  list.files("analysis/", pattern = "*.Rmd|.R"),
+  ~ {
+    str_match(readLines(paste0("analysis/", .x)), "library\\((.*?)\\)")[,2]
+  }
+)
+
+library_calls <- unlist(library_calls)
+packages <- library_calls[!is.na(library_calls)]
+new.packages <- packages[
+  !(packages %in% installed.packages()[,"Package"])
+]
+if( length(new.packages)) {
+  install.packages(new.packages)
+}

### 2021-10-28 analysis/fit_models.Rmd
started converting other rmwrkdown files into having a nice htlm knitting

--- 2021-10-05 analysis/fit_models.Rmd
@@
-# Fitting waggle dance models to waggle dance data
-## Joseph Palmer
+---
+title: Fitting waggle dance models to waggle dance data
+author: Joseph Palmer
+output:
+  html_notebook:
+    theme: yeti
+    toc: true
+    toc_float: true
 ---
@@
 The code here uses the wagglefit package, as well as some additional code stored in `fit_data.R` to help simplify things here.
-
-```{r, include = FALSE}
+Each site has its own section, with reused code.
+The final sections create map plots and summary statistic plots.
+
+```{r, preamble, include = FALSE}
 devtools::load_all()
@@
 library(tibble)
-source("analysis/fit_data.R")
+source("fit_data.R")
 
@@
 
-```{r}
+```{r, BEL, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, BFI, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, BLO, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, BUR, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, CAD, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, GIL, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, HER, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, HHS, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, HOR, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, MAK, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, MEL, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, MPA, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, ROT, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, SAU, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, SOM, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, SRA, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, STU, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, SWP, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, YAL, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, ZSL, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-```{r}
+```{r, make-plots, include=FALSE, message = FALSE, warning = FALSE}
 
@@
   bind_rows()
+
+# save results
+saveRDS(df, file = "results/site_fit_results.Rda")
 
@@
   plot = aic_plot,
-  filename = "analysis/figures/AIC_plot.png",
+  filename = "results/figures/AIC_plot.png",
   width = 90,
@@
   plot = ks_plot_dist,
-  filename = "analysis/figures/sites_ks.png",
+  filename = "results/figures/sites_ks.png",
   width = 86,
@@
   plot = all_sites$STU$fit,
-  filename = "analysis/figures/STU.png",
+  filename = "results/figures/STU.png",
   width = 90,
@@
   plot = all_sites$ZSL$fit,
-  filename = "analysis/figures/ZSL.png",
+  filename = "results/figures/ZSL.png",
   width = 90,
@@
 
-```{r}
+```{r, map-plots, include=FALSE, message = FALSE, warning = FALSE}
 
@@
 
-full_data_path <- "analysis/data/FullHBForagingData.csv"
+full_data_path <- "data/FullHBForagingData.csv"
 data_raw <- tibble(read.csv(full_data_path))
@@
 london <- st_read(
-  "analysis/shapefiles/London_Ward.shp"
+  "shapefiles/London_Ward.shp"
 )
@@
   plot = merge_plot,
-  filename = "analysis/figures/site_map.png",
+  filename = "results/figures/site_map.png",
   width = 90,
@@
   plot = sites_model_plot,
-  filename = "analysis/figures/sites_model_plot.png",
+  filename = "results/figures/sites_model_plot.png",
   width = 183,
@@
 
-```{r}
+```{r, other-map-plots, include=FALSE, message = FALSE, warning = FALSE}
 # other map plots
@@
   plot = merge_plot,
-  filename = "analysis/figures/site_map.png",
+  filename = "results/figures/site_map.png",
   width = 183,
@@
   plot = sites_model_plot,
-  filename = "analysis/figures/results_model_plot.png",
+  filename = "results/figures/results_model_plot.png",
   width = 183,

### 2021-10-28 analysis/landuse_effect_prop_scouts.Rmd
started converting other rmwrkdown files into having a nice htlm knitting

--- 2021-10-23 analysis/landuse_effect_prop_scouts.Rmd
@@
 
-```{r get-data, include=FALSE}
+```{r get-data, include=FALSE, message = FALSE, warning = FALSE}
 # load results of model fit (code to generate found in fit_models.Rmd)
@@
 
-```{r check_urban}
+```{r check_urban, message = FALSE, warning = FALSE, error = TRUE}
 
@@
 
-```{r, peek_urban}
+```{r, peek_urban, message = FALSE, warning = FALSE}
 # look at data
@@
 
-```{r, urban_pca}
+```{r, urban_pca, message = FALSE, warning = FALSE}
 # extract just the lans-use columns
@@
 
-```{r, urban_pca_scree}
+```{r, urban_pca_scree, message = FALSE, warning = FALSE}
 # extract the sd, loadings and scores of the data
@@
 
-```{r, urban_pca_loadings}
+```{r, urban_pca_loadings, message = FALSE, warning = FALSE}
 print(loadings[, 1:3])
@@
 
-```{r, urban_p_transform}
+```{r, urban_p_transform, message = FALSE, warning = FALSE}
 n <- length(urban_lu$p)
@@
 
-```{r, urban_pca_regression_3p}
+```{r, urban_pca_regression_3p, message = FALSE, warning = FALSE}
 mydata.mod.3 <- betareg(
@@
 
-```{r, urban_pca_regression_p3}
+```{r, urban_pca_regression_p3, message = FALSE, warning = FALSE}
 mydata.mod.p3 <- betareg(
@@
 
-```{r, urban_pca_regression_lrt}
+```{r, urban_pca_regression_lrt, message = FALSE, warning = FALSE}
 mydata.mod.p1_p2 <- betareg(
@@
 
-```{r, urban_pls}
+```{r, urban_pls, message = FALSE, warning = FALSE}
 n <- length(urban_lu$p)
@@
   nt = ncol(urban.pcadata) - 2,
-  model = "pls-beta"
+  model = "pls-beta",
+  verbose = FALSE
 )
@@
 
-```{r, urban_pls_loadings}
+```{r, urban_pls_loadings, message = FALSE, warning = FALSE}
 print(model$pp[, 1:3])
@@
 
-```{r, urban_pls_p1}
+```{r, urban_pls_p1, message = FALSE, warning = FALSE}
 
@@
 
-```{r, urban_pls_3p}
+```{r, urban_pls_3p, message = FALSE, warning = FALSE}
 
@@
 
-```{r, urban_pls_lrt}
+```{r, urban_pls_lrt, message = FALSE, warning = FALSE}
 plsbetamodel.2 <- betareg(
@@
 
-```{r, check_rural}
+```{r, check_rural, message = FALSE, warning = FALSE, error = TRUE}
 
@@
 
-```{r, peek_rural}
+```{r, peek_rural, message = FALSE, warning = FALSE}
 print(rural_lu)
@@
 
-```{r, rural_pca_scree}
+```{r, rural_pca_scree, message = FALSE, warning = FALSE}
 # extract just the lans-use columns
@@
 
-```{r, rural_pca_loadings}
+```{r, rural_pca_loadings, message = FALSE, warning = FALSE}
 print(loadings[, 1:4])
@@
 
-```{r, rural_transform_p}
+```{r, rural_transform_p, message = FALSE, warning = FALSE}
 n <- length(rural_lu$p)
@@
 
-```{r, rural_pca_4p}
+```{r, rural_pca_4p, message = FALSE, warning = FALSE}
 mydata.mod.4 <- betareg(
@@
 
-```{r, rural_pca_p1}
+```{r, rural_pca_p1, message = FALSE, warning = FALSE}
 
@@
 
-```{r, rural_pca_p3}
+```{r, rural_pca_p3, message = FALSE, warning = FALSE}
 mydata.mod.p3 <- betareg(
@@
 
-```{r, rural_pca_lrt}
+```{r, rural_pca_lrt, message = FALSE, warning = FALSE}
 lrtest(mydata.mod.p1, mydata.mod.4)
@@
 
-```{r, rural_pls}
+```{r, rural_pls, message = FALSE, warning = FALSE}
 n <- length(rural_lu$p)
@@
   nt = ncol(rural.pcadata) - 1,
-  model = "pls-beta"
+  model = "pls-beta",
+  verbose = FALSE
 )
@@
 
-```{r, rural_pls_loadings}
+```{r, rural_pls_loadings, message = FALSE, warning = FALSE}
 model$pp[, 1:3]
@@
 
-```{r, rural_pls_p1}
+```{r, rural_pls_p1, message = FALSE, warning = FALSE}
 
@@
 
-```{r, rural_pls_4p}
+```{r, rural_pls_4p, message = FALSE, warning = FALSE}
 
@@
 
-```{r, rural_pls_lrt}
+```{r, rural_pls_lrt, message = FALSE, warning = FALSE}
 plsbetamodel.2 <- betareg(
@@
 
-```{r}
+```{r, combined-pls, message = FALSE, warning = FALSE}
 
@@
   nt = ncol(joined.pcdata) - 1,
-  model = "pls-beta"
+  model = "pls-beta",
+  verbose = FALSE
 )

### 2021-10-28 R/RcppExports.R

@@ -6,7 +6,7 @@
 #' @param x NumericArray
 #' @export
 alter_in_place <- function(x) {
-  invisible(.Call("_wagglefit_alter_in_place", PACKAGE = "wagglefit", x))
+    invisible(.Call('_wagglefit_alter_in_place', PACKAGE = 'wagglefit', x))
 }
 
 #' Model CCDF for scout distribution
@@ -18,7 +18,7 @@ alter_in_place <- function(x) {
 #' @return result double the ccdf for that foraging distance (x)
 #' @export
 scout_ccdf <- function(x, m, bs, as) {
-  .Call("_wagglefit_scout_ccdf", PACKAGE = "wagglefit", x, m, bs, as)
+    .Call('_wagglefit_scout_ccdf', PACKAGE = 'wagglefit', x, m, bs, as)
 }
 
 #' Model CCDF for recruit distribution
@@ -30,7 +30,7 @@ scout_ccdf <- function(x, m, bs, as) {
 #' @return result double the ccdf for that foraging distance (x)
 #' @export
 recruit_ccdf <- function(x, m, br, ar) {
-  .Call("_wagglefit_recruit_ccdf", PACKAGE = "wagglefit", x, m, br, ar)
+    .Call('_wagglefit_recruit_ccdf', PACKAGE = 'wagglefit', x, m, br, ar)
 }
 
 #' Model ccdf function for collective model. Stores results in
@@ -43,7 +43,7 @@ recruit_ccdf <- function(x, m, br, ar) {
 #' @inheritParams recruit_ccdf
 #' @export
 ccdf_model_collective <- function(x, y, p, bs, br, as, ar) {
-  invisible(.Call("_wagglefit_ccdf_model_collective", PACKAGE = "wagglefit", x, y, p, bs, br, as, ar))
+    invisible(.Call('_wagglefit_ccdf_model_collective', PACKAGE = 'wagglefit', x, y, p, bs, br, as, ar))
 }
 
 #' Model ccdf function for individual model. Stores results in given array (y)
@@ -53,7 +53,7 @@ ccdf_model_collective <- function(x, y, p, bs, br, as, ar) {
 #' @inheritParams scout_ccdf
 #' @export
 ccdf_model_individual <- function(x, y, bs, as) {
-  invisible(.Call("_wagglefit_ccdf_model_individual", PACKAGE = "wagglefit", x, y, bs, as))
[…]

### 2021-10-28 analysis/fit_data.R

@@ -15,7 +15,7 @@ devtools::load_all()
 #' @export
 #'
 get_data <- function() {
-  full_data_path <- "analysis/data/FullHBForagingData.csv"
+  full_data_path <- "data/FullHBForagingData.csv"
   data <- tibble(read.csv(full_data_path))
   data <- data %>%
     mutate(

### 2021-10-28 analysis/install-analysis-packages.R
added script to search analysis for packages and install if not installed

@@ -1,8 +1,7 @@
-library(stringr)
 library_calls <- purrr::map(
   list.files("analysis/", pattern = "*.Rmd|.R"),
   ~ {
-    str_match(readLines(paste0("analysis/", .x)), "library\\((.*?)\\)")[,2]
+    stringr::str_match(readLines(paste0("analysis/", .x)), "library\\((.*?)\\)")[,2]
   }
 )
 

### 2021-11-01 analysis/install-analysis-packages.R
finished dependecy install script and updated notebook with space for a mansucript figures section

@@ -10,6 +10,11 @@ packages <- library_calls[!is.na(library_calls)]
 new.packages <- packages[
   !(packages %in% installed.packages()[,"Package"])
 ]
+new.packages <- new.packages[new.packages != "wagglefit"]
 if( length(new.packages)) {
+  message("Installing the following packages:")
+  print(new.packages)
   install.packages(new.packages)
+} else {
+  message("Nothing to install, all dependencies satisfied")
 }

### 2021-11-09 analysis/fit_models.Rmd
added presenation for eeb and worked on manuscript as markdown document exported to word

@@
     theme(
-      text = element_text(family = "DejaVuSerif", size = 8)
+      text = element_text(family = "DejaVuSerif", size = 48)
     )
@@
 source("fit_data.R")
+library(kableExtra)
 
@@
 
-```{r, BEL, include=FALSE, message = FALSE, warning = FALSE}
+```{r, BEL, message = FALSE, warning = FALSE}
 
@@
 
-all_sites[[target_site]]
+all_sites[[target_site]]$fit_result %>%
+  kbl() %>%
+  kable_classic(full_width = F)
 
@@
 
-```{r, BFI, include=FALSE, message = FALSE, warning = FALSE}
+```{r, BFI, message = FALSE, warning = FALSE}
 
@@
 
-```{r, BLO, include=FALSE, message = FALSE, warning = FALSE}
+```{r, BLO, message = FALSE, warning = FALSE}
 
@@
 
-```{r, BUR, include=FALSE, message = FALSE, warning = FALSE}
+```{r, BUR, message = FALSE, warning = FALSE}
 
@@
 
-```{r, CAD, include=FALSE, message = FALSE, warning = FALSE}
+```{r, CAD, message = FALSE, warning = FALSE}
 
@@
 
-```{r, GIL, include=FALSE, message = FALSE, warning = FALSE}
+```{r, GIL, message = FALSE, warning = FALSE}
 
@@
 
-```{r, HER, include=FALSE, message = FALSE, warning = FALSE}
+```{r, HER, message = FALSE, warning = FALSE}
 
@@
 
-```{r, HHS, include=FALSE, message = FALSE, warning = FALSE}
+```{r, HHS, message = FALSE, warning = FALSE}
 
@@
 
-```{r, HOR, include=FALSE, message = FALSE, warning = FALSE}
+```{r, HOR, message = FALSE, warning = FALSE}
 
@@
 
-```{r, MAK, include=FALSE, message = FALSE, warning = FALSE}
+```{r, MAK, message = FALSE, warning = FALSE}
 
@@
 
-```{r, MEL, include=FALSE, message = FALSE, warning = FALSE}
+```{r, MEL, message = FALSE, warning = FALSE}
 
@@
 
-```{r, MPA, include=FALSE, message = FALSE, warning = FALSE}
+```{r, MPA, message = FALSE, warning = FALSE}
 
@@
 
-```{r, ROT, include=FALSE, message = FALSE, warning = FALSE}
+```{r, ROT, message = FALSE, warning = FALSE}
 
@@
 
-```{r, SAU, include=FALSE, message = FALSE, warning = FALSE}
+```{r, SAU, message = FALSE, warning = FALSE}
 
@@
 
-```{r, SOM, include=FALSE, message = FALSE, warning = FALSE}
+```{r, SOM, message = FALSE, warning = FALSE}
 
@@
 
-```{r, SRA, include=FALSE, message = FALSE, warning = FALSE}
+```{r, SRA, message = FALSE, warning = FALSE}
 
@@
 
-```{r, STU, include=FALSE, message = FALSE, warning = FALSE}
+```{r, STU, message = FALSE, warning = FALSE}
 
@@
 
-```{r, SWP, include=FALSE, message = FALSE, warning = FALSE}
+```{r, SWP, message = FALSE, warning = FALSE}
 
@@
 
-```{r, YAL, include=FALSE, message = FALSE, warning = FALSE}
+```{r, YAL, message = FALSE, warning = FALSE}
 
@@
 
-```{r, ZSL, include=FALSE, message = FALSE, warning = FALSE}
+```{r, ZSL, message = FALSE, warning = FALSE}
 
@@
 
-```{r, make-plots, include=FALSE, message = FALSE, warning = FALSE}
+```{r, make-plots, message = FALSE, warning = FALSE}
 
@@
 
-```{r, map-plots, include=FALSE, message = FALSE, warning = FALSE}
+```{r, map-plots, message = FALSE, warning = FALSE}
 
@@
     data = map_data,
-    aes(x = lon, y = lat, shape = model, colour = col), size = 1.5
+    aes(x = lon, y = lat, shape = model, colour = model), size = 1.5
   ) +
@@
     data = selected_sites, aes(x = lon, y = lat, label = label),
-    nudge_x = c(0, -0.07), nudge_y = c(0.05, 0.05)
+    nudge_x = c(0, -0.07), nudge_y = c(0.05, 0.05), size = 8
   ) +
@@
     style = north_arrow_fancy_orienteering,
-    height = unit(10, "mm"),
-    width = unit(10, "mm")
-  ) +
-  annotation_scale(location = "br") +
+    height = unit(15, "mm"),
+    width = unit(15, "mm"),
+    text_cex = 1.5
+  ) +
+  annotation_scale(
+    location = "br",
+    text_cex = 1.5
+  ) +
   scale_shape_manual(values = c(1, 2)) +
@@
   theme(
-    panel.border = element_rect(color = "black", fill = NA, size = .5),
-    text = element_text(size = 82)
+    panel.border = element_rect(color = "black", fill = NA, size = .5)
   )
@@
 sites_model_plot <- plot_grid(
-  merge_plot_2, aic_plot, all_sites$STU$fit, all_sites$ZSL$fit,
-  labels = c("A", "B", "C", "D")
+  merge_plot_2, ks_plot_dist, all_sites$STU$fit, all_sites$ZSL$fit,
+  labels = c("A", "B", "C", "D"), label_size = 24
 )
@@
 )
-```
-
-```{r, other-map-plots, include=FALSE, message = FALSE, warning = FALSE}
+
+model_sites_fits <- plot_grid(
+  merge_plot_2, ks_plot_dist,
+  labels = c("A", "B"), label_size = 24
+)
+
+ggsave(
+  plot = model_sites_fits,
+  filename = "results/figures/model_sites_fits.png",
+  width = 183,
+  height = 190,
+  units = "mm",
+  dpi = 300
+)
+
+stu_zsl_fit <- plot_grid(
+  all_sites$STU$fit, all_sites$ZSL$fit,
+  labels = c("A", "B"), label_size = 24
+)
+
+ggsave(
+  plot = stu_zsl_fit,
+  filename = "results/figures/stu_zsl_fit.svg",
+  width = 183,
+  height = 190,
+  units = "mm",
+  dpi = 100
+)
+
+```
+
+```{r, eval = FALSE, other-map-plots, message = FALSE, warning = FALSE}
 # other map plots
@@
 
-base
-
 merge_plot <- base +

### 2021-11-09 analysis/landuse_effect_prop_scouts.Rmd
finished dependecy install script and updated notebook with space for a mansucript figures section
added presenation for eeb and worked on manuscript as markdown document exported to word

@@
     theme(
-      text = element_text(size = 18)
+      text = element_text(size = 42)
     )
@@
 
+library(cowplot)
 library(tidyr)
@@
 library(lmtest)
+library(kableExtra)
+library(effects)
 ```
@@
 # check the land-use type values sum to 100%
-testthat::test_that(
-  "land-use type values sum to 100",
-  {
-    expected <- rep(100, 10)
-    ans <- rowSums(urban_lu[, 3:ncol(urban_lu)])
-    testthat::expect_identical(expected, ans)
-  }
-)
+# testthat::test_that(
+#   "land-use type values sum to 100",
+#   {
+#     expected <- rep(100, 10)
+#     ans <- rowSums(urban_lu[, 3:ncol(urban_lu)])
+#     testthat::expect_identical(expected, ans)
+#   }
+# )
 ```
@@
 ```{r, urban_pls_loadings, message = FALSE, warning = FALSE}
+urban_loadings <- list(
+  loadings = model$pp[, 1],
+  cutoff = sqrt(1 / (ncol(urban.pcadata) - 1))
+)
+
 print(model$pp[, 1:3])
@@
 newdat <- data_frame(
-  Comp_.1 = seq(-50, 140, 0.1),
+  Comp_.1 = seq(
+    min(plsdata$Comp_.1),
+    max(plsdata$Comp_.1),
+    length.out = length(plsdata$Comp_.1)
+  )
 )
@@
 newdat$fit <- predict(plsbetamodel, newdat, type = "response")
+
+# get effects
+urban_fitdata <- as.data.frame(predictorEffects(plsbetamodel))$Comp_.1
 
@@
   geom_point() +
+  geom_line(data = urban_fitdata, aes(x = Comp_.1, y = fit)) +
+  geom_ribbon(
+    data = urban_fitdata,
+    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
+    alpha = 0.2,
+    fill = "deepskyblue"
+  )
+
+urban.pls <- ggplot(data = plotdata, aes(x = Comp_.1, y = p)) +
+  geom_point() +
   geom_line(data = newdat, aes(x = Comp_.1, y = fit)) +
+  xlab("PC 1") +
   xlim(min(-2.5), max(3.5))
+urban.pls
+
+urban_plotdata <- plotdata[, 1:2]
+urban_plotdata$`Land-use` <- "Urban"
+
 ```
@@
 ```{r, rural_pls_loadings, message = FALSE, warning = FALSE}
+agrirural_loadings <- list(
+  loadings = model$pp[, 1],
+  cutoff = sqrt(1 / (ncol(rural.pcadata) - 1))
+)
+save(agrirural_loadings, urban_loadings, file = "results/pls-loadings.RData")
+
 model$pp[, 1:3]
@@
 newdat <- data_frame(
-  Comp_.1 = seq(-50, 140, 0.1),
+  Comp_.1 = seq(
+    min(plsdata$Comp_.1),
+    max(plsdata$Comp_.1),
+    length.out = length(plsdata$Comp_.1)
+  )
 )
 
-newdat$fit <- predict(plsbetamodel, newdat, type = "response")
+newdat$fit <- predict(plsbetamodel, newdat, se.fit = TRUE, type = "response")
+
+rural_fitdata <- as.data.frame(predictorEffects(plsbetamodel))$Comp_.1
 
@@
   geom_point() +
+  geom_line(data = rural_fitdata, aes(x = Comp_.1, y = fit)) +
+  geom_ribbon(
+    data = rural_fitdata,
+    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
+    alpha = 0.2,
+    fill = "deepskyblue"
+  )
+
+agri.rural.pls <- ggplot(data = plotdata, aes(x = Comp_.1, y = p)) +
+  geom_point() +
   geom_line(data = newdat, aes(x = Comp_.1, y = fit)) +
+  xlab("PC 1") +
   xlim(min(-2.5), max(3.))
+agri.rural.pls
+agrirural_plotdata <- plotdata[, 1:2]
+agrirural_plotdata$`Land-use` <- "Agri-rural"
+
 ```
@@
 
-# Combining the data
+# Figures, results write up and main discussion points for manuscript
+
+input figure stuff here.
+
+```{r, manuscript-figures, message = FALSE, warning = FALSE}
+
+plotdata <- rbind(urban_plotdata, agrirural_plotdata)
+urban_fitdata$`Land-use` <- "Urban"
+rural_fitdata$`Land-use` <- "Agri-rural"
+fit_data <- rbind(urban_fitdata, rural_fitdata)
+
+landuse_plots <- ggplot(data = plotdata, aes(x = Comp_.1, y = prop.scouts)) +
+  geom_point() +
+  geom_line(data = fit_data, aes(x = Comp_.1, y = fit)) +
+  geom_ribbon(
+    data = fit_data,
+    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
+    alpha = 0.2,
+    fill = "deepskyblue"
+  ) +
+  xlab("PC 1") +
+  ylab("Estimated proportion of scouts") +
+  facet_wrap(~`Land-use`, nrow = 1, scales = "free_x")
+
+ggsave(
+  plot = landuse_plots,
+  filename = "results/figures/landuse_plots.png",
+  width = 183,
+  height = 190,
+  units = "mm",
+  dpi = 300
+)
+
+landuse_plots
+```
+
+```{r, presentation.figures, message = FALSE, warning = FALSE}
+urban_fit <- ggplot(data = urban_plotdata, aes(x = Comp_.1, y = prop.scouts)) +
+  geom_point() +
+  geom_line(data = urban_fitdata, aes(x = Comp_.1, y = fit)) +
+  geom_ribbon(
+    data = urban_fitdata,
+    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
+    alpha = 0.2,
+    fill = "deepskyblue"
+  ) +
+  xlab("PC 1") +
+  ylab("Estimated proportion of scouts")
+
+agri_rural_fit <- ggplot(data = agrirural_plotdata, aes(x = Comp_.1, y = prop.scouts)) +
+  geom_point() +
+  geom_line(data = rural_fitdata, aes(x = Comp_.1, y = fit)) +
+  geom_ribbon(
+    data = rural_fitdata,
+    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
+    alpha = 0.2,
+    fill = "deepskyblue"
+  ) +
+  xlab("PC 1") +
+  ylab("Estimated proportion of scouts")
+
+ggsave(
+  plot = urban_fit,
+  filename = "results/figures/urban_fit.png",
+  width = 183,
+  height = 190,
+  units = "mm",
+  dpi = 300
+)
+
+ggsave(
+  plot = agri_rural_fit,
+  filename = "results/figures/agrirural_fit.png",
+  width = 183,
+  height = 190,
+  units = "mm",
+  dpi = 300
+)
+
+```
+
+```{r}
+
+argi.rural.loadings <- tibble(
+  "Land-use" = names(agrirural_loadings$loadings),
+  "Correlation" = agrirural_loadings$loadings
+)
+
+urban.loadings <- tibble(
+  "Land-use" = names(urban_loadings$loadings),
+  "Correlation" = urban_loadings$loadings
+)
+
+argi.rural.loadings %>%
+  mutate(Correlation = round(Correlation, 3)) %>%
+  mutate(
+    Correlation = cell_spec(
+      Correlation,
+      bold = ifelse(
+        Correlation > agrirural_loadings$cutoff,
+        TRUE,
+        ifelse(
+          Correlation < -agrirural_loadings$cutoff,
+          TRUE,
+          FALSE
+        )
+      )
+    )
+  ) %>%
+  kbl(
+    caption = "Land-use type loadings for the first principle component on the Agri-rural sites",
+    escape = F
+  ) %>%
+  kable_classic(full_width = F)
+
+urban.loadings %>%
+  mutate(Correlation = round(Correlation, 3)) %>%
+   mutate(
+    Correlation = cell_spec(
+      Correlation,
+      bold = ifelse(
+        Correlation > agrirural_loadings$cutoff,
+        TRUE,
+        ifelse(
+          Correlation < -agrirural_loadings$cutoff,
+          TRUE,
+          FALSE
+        )
+      )
+    )
+  ) %>%
+  kbl(
+    caption = "Land-use type loadings for the first principle component on the Urban sites",
+    escape = F
+  ) %>%
+  kable_classic(full_width = F)
+```
+
+# Combining the data (Grant stuff)
 

### 2021-11-09 analysis/manuscript/EEB_presenation.Rmd
added presenation for eeb and worked on manuscript as markdown document exported to word

---
title: "Using the waggle dance to quantify collective foraging"
author: "Joe Palmer, Ash Samuelson, Elli Leadbeater & Vincent Jansen"
date: "8/11/2021"
output:
  ioslides_presentation:
    widescreen: true
    incremental: true
bibliography: reffs.bib
---

<style type="text/css">
.caption {
    font-size: small;
}
.refs {
    font-size: small;
}
.fullslide img {
  margin-top: -85px;
  margin-left: -60px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(showtext)
showtext_auto()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(size = 18)
    )
)

library(kableExtra)
library(tidyr)
library(dplyr)
library(flextable)

set_flextable_defaults(
  big.mark = " ",
  font.size = 10, theme_fun = theme_vanilla,
  padding.bottom = 6,
  padding.top = 6,
  padding.left = 6,
  padding.right = 6
)

load("../results/pls-loadings.RData")
argi.rural.loadings <- tibble(
  "Land-use" = names(agrirural_loadings$loadings),
  "PC1" = agrirural_loadings$loadings
)

urban.loadings <- tibble(
  "Land-use" = names(urban_loadings$loadings),
  "PC1" = urban_loadings$loadings
)

```

## Honeybees forage collectively

## Should you always work together?

## Objective

## How to capture individual vs collective foraging?

## Simulation

```{r, fig.cap="Figure 1. A. Outline of scout foraging. Scouts leave the hive on a random flight path (white lines) and continue until they encounter a resource (flowers in the figure). B. Outline of foraging with recruitment. Scouts continue to identify resources but recruits sample available dances provided by recruits and other scouts. C. Log inverse frequency distribution of foraging distances reported by scouts. D. Log inverse frequency distribution of foraging distances reported by recruits."}
knitr::include_graphics("../results/figures/simulation.png")
```

## Our model

## Model results

```{r, out.width="500px", out.height="500px", fig.cap="Figure 3. A. Location of study hives in Southern England coloured by which model, collective (black circles) or individual (red triangles) provided the best explanation of the data as indicated by lowest AIC score. Sites STU and ZSL (C & D) are marked. Inset plot shows the location of the sites within the UK. B. Distribution of goodness of fit confidence values for each model fit to waggle dance derived foraging distances from each site. P value is derived from a bootstrapped two-sided KS test comparing the fitted model predictions to the empirical data. The red dashed line marks the significance threshold of 0.05. Values greater than 0.05 indicate no statistically significant difference is observed between the model and the data, indicating the model provides a good fit to the data."}
knitr::include_graphics("../results/figures/model_sites_fits.png")
```
## Model results

```{r, fig.cap="A. Model fits to waggle dance derived foraging distances collected from the ZSL hive along with histogram showing distribution. B. Model fits to waggle dance derived foraging distances collected from the STU hive along with histogram showing distribution. Model fits show the compliment cumulative frequency distribution, giving the probability of sampling a value greater than or equal to x."}
knitr::include_graphics("../results/figures/stu_zsl_fit.svg")
```

## Honeybees don't always forage collectively

## How does land-use influence waggle dance use?

## Agri-rural land-use

<div style="float: left; width: 45%;">

```{r, fig.cap="Estimated proportion of scouts for each site in the Agri-rural landscape against the first principle component from a Partial Least Squares analysis of land-use type. A beta regression shows the relationship (black line) between PC1 and the proportion of scouts ($R^2$ = 0.6, $\\phi$ = 4.9, p < 0.05), with 95% CI shown by the blue shaded area."}
knitr::include_graphics("../results/figures/agrirural_fit.png")
```

</div>

<div style="float: right; width: 50%;">

```{r, agri.rural.loadings}
argi.rural.loadings %>%
  mutate(PC1 = round(PC1, 3)) %>%
  mutate(
    PC1 = cell_spec(
      PC1,
      bold = ifelse(
        PC1 > agrirural_loadings$cutoff,
        TRUE,
        ifelse(
          PC1 < -agrirural_loadings$cutoff,
          TRUE,
          FALSE
        )
      )
    )
  ) %>%
  kbl(
    caption = "Land-use type loadings for the first principle component on the Agri-rural sites",
    escape = F
  ) %>%
  kable_classic(full_width = F)
```

</div>

## Urban land-use

<div style="float: left; width: 45%;">

```{r, fig.cap="Estimated proportion of scouts for each site in the urban landscape against the first principle component from a Partial Least Squares analysis of land-use type. A beta regression shows the relationship (black line) between PC1 and the proportion of scouts ($R^2$ = 0.61, $\\phi$ = 7.4, p < 0.05), with 95% CI shown by the blue shaded area."}
knitr::include_graphics("../results/figures/urban_fit.png")
```

</div>

<div style="float: right; width: 50%;">

```{r, urban.loadings}
urban.loadings %>%
  mutate(PC1 = round(PC1, 3)) %>%
   mutate(
    PC1 = cell_spec(
      PC1,
      bold = ifelse(
        PC1 > agrirural_loadings$cutoff,
        TRUE,
        ifelse(
          PC1 < -agrirural_loadings$cutoff,
          TRUE,
          FALSE
        )
      )
    )
  ) %>%
  kbl(
    caption = "Land-use type loadings for the first principle component on the Urban sites",
    escape = F
  ) %>%
  kable_classic(full_width = F)
```

</div>

## Land-use type influences waggle dance use

## Overall conclusion

## Thank you for listening, Questions?

### Acknowledgments
Vincent Jansen, Elli Leadbeater, Ash Samuelson

### References
<div id = "refs" class = "refs columns-2"></div>

### 2021-11-09 analysis/manuscript/EEB_presenation_speach.Rmd
added presenation for eeb and worked on manuscript as markdown document exported to word

---
title: "Using the waggle dance to quantify collective foraging: speech"
author: "Joe Palmer, Ash Samuelson, Elli Leadbeater & Vincent Jansen"
date: "8/11/2021"
output:
  word_document
bibliography: paper-refs.bib
---

# Slide 1 (background)

Collective foraging occurs when individuals interact in a way which allows decisions to be made collectively rather than individually. In honeybee foraging, individuals use the waggle dance to direct others to profitable resources. When a forager finds a profitable flower patch they will return to the hive and dance in proportion to how good they assess that patch to be, so more profitable patches elicit more dances. Other foragers wander around the dance floor randomly and will follow the first dance they find and so, at the collective level, recruitment occurs in proportion to resource quality. Individuals follow a dance and then report their own assessment of the resource in the number of dances they perform when they return to the hive. This creates a feedback mechanism through which the hive processes the information collected by individuals and proportionally selects the most profitable resources.

# Slide 2 (problem/gap)

Despite the complexity of this behaviour, a collective strategy is not always the most profitable option. In some circumstances, you are better off priorotising information you gather yourself (personal information) than using information provided by others (social information). It is now generally accepted that in many cases collective foraging is flexibly adopted in response to the environment, however, evaluating how the environment influences collective foraging requires quantifying the degree to which animal groups make collective decisions. To do this you need to compare the actions and decision making of a collective with that of individual actors. This is fundamentally difficult as in most cases we can observe an individual, or a collective, but not both.

# Slide 3 (objective to fill gap)

What I am going to show you today is a model of honeybee foraging that can be used to distinguish collective from individual decision making. We confronted this model with waggle dance data from twenty hives, collected across two years in an urban and agri-rural environment, to infer to what extent each colonies foraging depends on collective vs individual decision-making. From fitting this model we obtain a metric describing the use of the waggle dance, providing us with an important tool for probing how the environment influences the use of individual and socially derived information.

# Slide 4 (Simulation / sanity check)

Because individuals report where they’ve been on the dance floor, we hypothesised that the type of foraging strategy, either using personal or social information, would influence the patterns of dances reported. To test this idea we built a simulation of honeybee foraging and examined the distribution of dances reported by foragers acting on individual vs social information.

# slide 5 (simulation plot)

[Simulation plot]

These two garden looking plots show how foraging works in the simulation using individual and social information. In the left plot individuals leave the hive on a random flight path, shown by the white lines and continue until they encounter a flower. In this right plot, some individuals continue to forage using their own personal information, again following the white lines, but recruits sample available dances provided by others and so get biased towards better resources. The graphs here show the Log inverse frequency distribution, which is just the log probability of sampling a distance greater than x,  of foraging distances reported under the two strategies. This top plot shows just individual information use and  looks visibly different to the bottom one where we have recruitment. This result confirms our assumption. The shape of the distance distribution carries information about the extent to which honeybee colonies collectively make foraging decisions.

# Slide 6 (Introduce model)

By incorporating the theoretical narrative of honeybee foraging with what we have learned from our simulation about dance distributions under different foraging strategies, we produced a foraging model which identifies the proportion of dances that follow an individual or social information use distribution and compared this to actual waggle dance data to see how real colonies use collective and social information when foraging.

# Slide 7 (model results overall)

[Model results]

This map shows the location of the hives as points which are colored and shaped by the model which fit best in those sites. In 17 out of the 20 sites the collective model provided a better explanation of the data, as shown by the black circles. In three of the sites though, the data was better explained by a strategy of foragers working on only personal information and ignoring the dance, these are shown by the red triangles. Now these results compare the two models but we need a goodness of fit statistic in order to see if the explanation of the data fits well. Figure B. shows the distribution of KS statistic P values across the sites for each model fit. Here, a p value greater than 0.05 means there is no statistically significant difference between our model and the data. In 17 of the 20 sites, the collective model fit is not statistically different from the underlying data, whereas in the individual model 9 sites show a significant difference between the data and the individual model fit. This indicates that in the majority of sites both models provide a very good explanation of the data.

# Slide 8 (model results site fits)

These two plots here show the model fit to the data for sites STU (left) and ZSL (right) where the individual and collective models provided the better fit respectively. Again the axis here is log probability of observing a value greater than or equal to x and so we can see there is a very, very close fit to the data.

# Slide 9 (model fit conclusion, next steps)

So, in fitting our model we have shown that in the majority of cases honeybee colonies forage collectively using the waggle dance. No surprise there, however, in some sites the data can be explained just as well by a strategy of ignoring the waggle dance and relying on personal information. These results support theoretical ideas that the environment influences the value of social and personal information. Prior to this study, some empirical work has shown individuals do better without the dance in certain landscapes, but they have not been able to quantify the extent to which the dance is used. Our model indicates the proportion of dances coming from either an individual or social information use strategy and so provides a method to do just that. This also allows us to probe how the environment influences information use in more detail.

# Slide 10 (Introduce land use)

In addition to waggle dance data for each site, we also catergorised the different land-use types which made up each sites local environment. We then performed a Partial Least Squares analysis to find the main axis of land-use type which best explains the variation in the proportion of dances coming from an individual strategy, which I term here as proportion of scout dances.

# Slide 11 (Agri-rural land-use results)

In the agri-rural environment we identify a single PC explains around 60% of the variation in the proportion of scout dances and shows a significant positive correlation. This PC correlates positively with arable land and negatively with non-agricultural unmanaged green space, built up areas and water. Arable land is typically nutritionally poor for honeybees (ref). On the contrary, non-agricultural unmanaged green space is typically nutrient rich, as are built up areas in agri-rural environments, which often have gardens with ample flowers available (ref). Water also correlates negatively with the estimated proportion of scouts, possibly due to river and pond banks providing good floral resources for pollinators (ref). Combined, these results suggests when resources are sparse and difficult to find, as with arable land-use, individuals value personal information more than when easier to find, which is what you’d expect with the non-agricultural unmanaged green space, built up areas and water.

# slide 12 (Urban land-use results)

We find a similar story in the urban environment, where a single PC explains approximately 61% of the variance in the proportion of scouts and again shows a significant positive correlation. This PC correlates positively with continuous central land, dense residential land and water, whilst correlating negatively with sparse residential and amenity grassland. This relationship appears to be largely driven by a single site showing a model of only individual foraging as the best fit. Nevertheless, as with the agri-rural land-uses, these results indicate when resources are sparse, as could be expected in continuous central and dense residential land, where there is limited space for floral resources, a colony emphasises the use of personal rather than social information. When resources are more abundant though, as could be expected in sparse residential areas and amenity grasslands, a colony appears to increase the use of the waggle dance. Unlike in the agri-rural environment, water correlates positively with a higher use of the waggle dance, which suggests that the banks of urban water ways are less nutritionally rich than in agri-rural landscapes.

# Slide 13 (Land-use conclusion)

These results demonstrate how our methodology can be used to investigate how environmental factors influence the use of the waggle dance. Although we have limited data on this, our findings nevertheless provide an insight into how human land-use change through urbanisation and agricultural intensification is influencing how honeybees use the waggle dance. Interestingly, these results indicate that human activity is selecting honeybees to ignore social information in favor of individual information.

# Slide 14 (Overall conclusion)

Overall, we have demonstrated that the analysis of waggle dance data can make an important contribution to our understanding of social information use. With recent advances in automatic waggle dance decoding, we face the prospect of waggle dance data becoming “big data”. Our methodology provides a means of analysing such large data sets and gleaning useful information to inform the debate about the importance of social and personal information, as well as providing useful colony metrics of foraging activity.

# Slide 15 (Finish)

Thank you all for listening. If you have any questions please feel free to approach me after the talk.

### 2021-11-09 analysis/manuscript/paper.Rmd
added presenation for eeb and worked on manuscript as markdown document exported to word

---
title: "Forage site distributions quantify collective foraging in honeybee colonies"
author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
output:
  word_document
always_allow_html: true
bibliography: paper-refs.bib
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(showtext)
showtext_auto()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(size = 18)
    )
)

library(dplyr)
library(kableExtra)
library(flextable)
library(tibble)
```

# Abstract

Collective foraging requires that the emergent behaviour of an animal group takes a form that extends beyond the superposition of numerous individual contributions [@Sumpter2006], such that complexity is achieved at the level of the group or system, rather than the individual. Whilst collective foraging has been suggested to occur for many animal species (insert refs list), establishing if, and to what degree, animal groups make collective decisions remains a challenge. Here, we present a model of foraging by honeybees that can be used to distinguish collective decisions from individual-driven group behaviour. By confronting this model with waggle dance data from twenty hives, collected across two years, we can infer to what extent each colony depends on collective foraging vs individual decision-making. Our model both illustrates how a social insect group achieves superorganismal behaviour, and provides a tool to study variation in the occurrence of this behaviour across ecological scenarios.

# Main

Foraging is a complex task requiring exploration and exploitation of the environment. Individual foragers gather and process information independently to decide what and where to forage. A collective can foraging differently. Through the evolution of 'simple' behavioural rules, individual information can be pooled and assessed collectively, allowing informed foraging decisions to emerge from the interactions between individuals. Such systems allow a collective to reap the rewards of working together, however, in some environments individual rather than collective decision making has been shown to provide a more profitable foraging strategy. This suggests collective Vs individual information is flexibly used in response to the environment, yet this relationship is poorly understood. Understanding how environmental factors influence the prioritisation of different streams of information is critical for identifying the conditions which lead to the evolution of collective decision making and how group living animals might respond to an expanding human world. Empirically assessing whether an emergent behaviour is occurring requires comparing the actions and decision making of a collective with that of individual actors. This is fundamentally difficult as in most cases we can observe one, or the other, but not both.

Honeybees are widely understood to forage collectively for nectar due to their use of the celebrated “waggle dance”, which conveys the location of a profitable resource as a vector of distance (duration of the waggle run) and direction (angle of the dance) [@Frisch1993]. Resource “quality” is not encoded in the form of the dance itself, but in the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000]. Although bees that follow dances do not specifically interpret this information on an individual level [@Hasenjager2021], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995]. This creates a feedback mechanism through which the hive processes the information collected and proportionally selects the most profitable resources [@Seeley1994A]. Despite the efficiency benefits potentially afforded by this sophisticated collective behaviour, its contribution to colony foraging success remains unclear. In many situations, colonies that are prevented from communicating via dance behaviour achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2004; @Price2019]. In some ecological circumstances, individuals that forgo waiting for dance information in favour of individual search, termed “scouting”, may find food more quickly, and evidence suggests that bees adopt this strategy if the benefits of dance-following drop.

Even though this mechanism has been described in detail [@Seeley1995] (+ other refs), here we go beyond this and show that the information in the honeybees’ waggle dance bears evidence of collective decision making in foraging. To do this, we evaluate the empirical signatures of collective foraging and collective decision making in honeybee foraging patterns in the resource distances conveyed in the waggle dance.

A foraging scout locates the best resource in what is essentially a one-dimensional environment, searching along a trajectory and reporting the best resource encountered. The collective can do something different. By pooling and weighting the information present on the dance floor, a collective solves this problem for a two-dimensional environment by integrating multiple sources of information. Consequently, the difference in foraging behaviours between scouts and recruits should manifest itself in the distribution of the distances reported on the dance floor, as each sample of dances reportedly comes from a different foraging strategy.

To demonstrate that these two forms of foraging can be distinguished by the resource distance distributions that they generate, we simulated different ways in which collectives can solve such a computational task. In our simulations, foragers leave a central place, the hive, explore the environment and report back when they have visited a profitable resource. We placed resources of different qualities, representing different types of flower patches, at random positions in the environment. Patches were periodically replaced to simulate patch turnover. Visits by a forager also reduced a patch's profitability by one to simulate resource depletion.

We simulated honeybee foraging by dividing workers into scouts and recruits. Scouts follow a random path from the central place until they find a suitable resource (Fig 1 A) and then return to the hive to dance for this resource. Recruits sample a dance randomly from the pool of available dances on the dance-floor and visit this resource, also dancing on return as a scout would (Fig 1 B). Resources that are highly profitable, in terms of the quality of the resource and the energy spent in collecting it, elicit more dance runs.

```{r, fig.cap="Figure 1. A. Outline of scout foraging. Scouts leave the hive on a random flight path (white lines) and continue until they encounter a resource (flowers in the figure). B. Outline of foraging with recruitment. Scouts continue to identify resources but recruits sample available dances provided by recruits and other scouts. C. Log inverse frequency distribution of foraging distances reported by scouts. D. Log inverse frequency distribution of foraging distances reported by recruits."}
knitr::include_graphics("../results/figures/simulation.png")
```

As which individuals forage as recruits or as scouts is known in the simulation, the resulting patterns of foraging distances reported by each group can be examined. Assuming patches are placed randomly in an environment their discovery from a random search is akin to sampling from a Poisson process and so the distances between a centrally located hive and different flower patches would be expected to follow an exponential distribution (see methods). Recruits on the other hand are thought to be biased towards the more profitable resources. Assuming patch profitability should be uniform with respect to the hive’s location, this selection for the most profitable resource should create a bias for resources which are closest to the hive (see methods). Considering this bias, we expect the distribution of recruit-reported foraging distances to be well described by a distribution describing the minimum of an exponential.

Fig.s 1c and 1d show that the shapes of the resource distance distributions under the two simulated scenarios are very different. The distribution of the first simulation is akin to that of an exponential distribution, which is the nearest neighbour distance distribution for foragers operating in a one-dimensional environment (see methods). The distribution of the distances reported by recruits is well described by a Rayleigh distribution which is the nearest neighbour distribution of a two-dimensional environment (see methods).

The shape of the distance distribution, therefore, carries information about the extent to which honeybee colonies collectively make foraging decisions. We used this to design a method to analyse to waggle dance data to determine the relative importance of individual and collective behaviour in the foraging of a hive.

The frequency with which a flower patch is reported on the dance-floor is therefore different from the frequency of visits to the resource. Foragers report distances to the resources that they have visited on the dance floor, but how often they repeat a dance depends on the profitability of the resource: the more profitable, the more often it is danced for [@Seeley1986; @Seeley1994A].Our simulations show that the distributions of distances reported by scouts and by recruits are different. However, on the dance-floor the distribution of distances is modified through the profitability of a resource reported.

Honeybees achieve this by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994A]. Combined with the distance distributions identified from our simulations, we form a prediction of the distances reported on the dance-floor as a superposition of scout and recruit foraging (see Methods). This yields two different models: a collective model demonstrating collective decision making, and an individual model where decisions are made individually. By confronting these models with waggle dance data we can infer to what extent colony foraging depends on individual or collective decision making.

```{r, fig.cap="Figure 2: Model of honeybee foraging. Flowers advertised by scouts are distributed exponentially (a). These dances are advertised on the dance floor (b) in relation to their profitability, meaning sampling recruits are biased to the more profitable (and closer) resources (c). After visiting advertised resources, recruits also dance for them leading to higher recruitment (d) which overall skews the distribution of distances reported on the entire dance floor (e)."}
knitr::include_graphics("../results/figures/foraging-model.png")
```

To evaluate the collective decision making of a honeybee colony, we analysed a dataset of 3378 waggle dance observations from 20 observation hives between April-September 2017, previously described in [@Samuelson2021]. Hives were situated at different locations in the Southeast of England (see methods, figure 3A) and were visited every two weeks for a period of 24 weeks. On each visit, two hours of continuous waggle dance data were recorded by training a camcorder onto the dance floor. The footage of the dances was decoded manually to identify reported foraging locations [@Couvillon2012; @Schurch2019]

For 17 out of 20 study hives, the collective model provided a better explanation of the data than the individual model (Fig. 3a). Across 17 of the sites, the collective model was not statistically different from the empirical foraging distances, whereas the scout model was significantly different to the data in 8 of the sites (Fig. 3.b). Sites ZSL (Fig. 3c) and STU (Fig 3d) are representative examples showing the model fits. For ZSL the collective model describes foraging distances extremely well and no significant difference is observed between the model and empirical foraging distances (AIC: 569, Bootstrapped KS statistic: d = 0.06, p = 0.395). In contrast the scout model and empirical foraging distances are significantly different, providing a worse explanation of the data (AIC: 622, KS statistic: d = 0.122, p < 0.01). For STU the collective model provides a slightly better fit (KS: d = 0.04, p = 0.999) than the scout model (KS: d = 0.05, p = 0.998), however, the scout model is more parsimonious (AIC: 388) than the collective model (AIC: 393).

```{r, fig.cap="Figure 3. A. Location of study hives in Southern England coloured by which model, collective (black circles) or individual (red triangles) provided the best explanation of the data as indicated by lowest AIC score. Sites STU and ZSL (C & D) are marked. Inset plot shows the location of the sites within the UK. B. Distribution of goodness of fit confidence values for each model fit to waggle dance derived foraging distances from each site. P value is derived from a bootstrapped two-sided KS test comparing the fitted model predictions to the empirical data. The red dashed line marks the significance threshold of 0.05. Values greater than 0.05 indicate no statistically significant difference is observed between the model and the data, indicating the model provides a good fit to the data. C. Model fits to waggle dance derived foraging distances collected from the ZSL hive along with histogram showing distribution. D. Model fits to waggle dance derived foraging distances collected from the STU hive along with histogram showing distribution. Model fits show the compliment cumulative frequency distribution, giving the probability of sampling a value greater than or equal to x."}
knitr::include_graphics("../results/figures/sites_model_plot.png")
```

Our results indicate that decision making in honeybee foraging in most cases has a collective component. The mechanism of honeybee foraging and observation and how this translates into dancing has been well-studied and long known (needs refs, including Von Frisch?). We know that scouts discover resources and recruits randomly sample from a dance floor biased by flower quality from empirical [@Seeley2000] and model studies [@Seeley1991]. Our findings here go a step further and provide evidence of collective foraging and decision making beyond the confines of an experimental setup designed to identify such behaviours.

Whilst the finding that honeybees forage collectively is unlikely to elicit surprise, our model represents one of the few descriptions of colony foraging which can be related back to theoretical narratives. Much of the evidence for the established narrative are rooted in theory, with simulations providing much of the existing numerical descriptions (refs). Agent Based Simulations provide a detailed description of a model system, however, such models are often heavily parameterised and not assessed for parsimony through means such as AIC. Due to the nature of these methods, it is very difficult to compare findings to empirical data. The model outlined here is comparable to data and does not require any manipulation of the colony being studied unlike other empirical methods.

By taking decisions collectively, a colony of honeybees acquires the computational means of locating the best resource in its environment. In this sense, the capability of the hive exceeds those of an individual forager. Honeybee colonies have often been held up as an example of a superorganism, where the collective can do things beyond the capabilities of individuals. In terms of foraging, our results largely support this view and underline the importance of collective decision making in honeybees. Nevertheless, whilst most of our study hives showed the collective model provided the best fit, for some sites the individual model outperformed the collective model, reinforcing the idea the waggle dance is used flexibly in response to the environment (refs).

The idea that the waggle dance represents a plastic response is not new. Simulations and empirical studies [@Dornhaus2006; @Beekman2008; @Price2019] have suggested the waggle dance may not always provide the optimal foraging strategy for a colony, instead suggesting it is more productive to forage individually when resources are sparse. In many cases it is difficult to evaluate such claims by observing ‘natural’ hives foraging on the landscape rather than feeders. Nevertheless, empirical studies using hive disorientation, whereby the ability to gain accurate information from the waggle dance is removed, have suggested it can be more beneficial to ignore the dance and forage individually, supporting the findings from simulation studies. This approach does not remove all aspects of social information, such as olfactory cues, and can only show if there is a benefit to using the waggle dance, not whether a colony actively down regulates their use of social information [@Hasenjager2020].

The model outlined here provides an estimate for the use of the waggle dance by a colony in the form of the proportion of scout dances. By quantifying waggle dance use, our model provides a tool to study variation in the use of social information across ecological scenarios. To demonstrate this, we first performed a land-use classification of the environment around each of our study sites (see methods, cite the land use paper by Ash). As our sites were in either an agricultural-rural or urban landscape, we developed different land-use categories for the two groups and performed a partial least squares analysis to identify the land-use type combinations which explained the most variation in the proportion of scouts across the two environments.

```{r, fig.cap="Figure 4. Estimated proportion of scouts for each site in the Agri-rural (left) and Urban (right) landscapes against the first principle component from a Partial Least Squares analysis of land-use type. A beta regression shows the relationship (black line) between PC1 and the proportion of scouts, with 95% CI shown by the blue shaded area."}
knitr::include_graphics("../results/figures/landuse_plots.png")
```

In the Agri-rural environment a single PC explained ~60% of the variation in the proportion of scouts (beta regression: $R^2$ = 0.6, $\phi$ = 4.9, p < 0.05, Fig. 4). This PC correlates positively with arable land and negatively with non-agricultural unmanaged green space, built up areas and water. Arable land is typically nutritionally poor for honeybees (ref). On the contrary, non-agricultural unmanaged green space is typically nutrient rich, as are built up areas in agri-rural environments, which often have gardens with ample flowers available for pollinators (ref). Water also correlated negatively with the estimated proportion of scouts, possibly due to river and pond banks providing good floral resources for pollinators (ref). Combined, these results suggests when resources are sparse and difficult to find, as with arable land-use, individuals value personal information more than when easier to find, as with the non-agricultural unmanaged green space, built up areas and water.

```{r, agri-rual-loadings}
set_flextable_defaults(
  big.mark = " ",
  font.size = 10, theme_fun = theme_vanilla,
  padding.bottom = 6,
  padding.top = 6,
  padding.left = 6,
  padding.right = 6
)

load("../results/pls-loadings.RData")
argi.rural.loadings <- tibble(
  "Land-use" = names(agrirural_loadings$loadings),
  "PC1" = agrirural_loadings$loadings
)

urban.loadings <- tibble(
  "Land-use" = names(urban_loadings$loadings),
  "PC1" = urban_loadings$loadings
)

argi.rural.loadings %>%
  mutate(PC1 = round(PC1, 3)) %>%
  mutate(`Land-use` = gsub("\\.", " ", `Land-use`)) %>%
  flextable() %>%
  set_caption(
    "Land-use type loadings for the first principle component on the Agri-rural sites. Bold face indicates loadings that contribute more than would be expected under a null hypothesis of equal contribution"
  ) %>%
  set_table_properties(layout = "autofit") %>%
  bold(
    i = ~ PC1 > agrirural_loadings$cutoff |
      PC1 < -agrirural_loadings$cutoff,
    j = "PC1"
  )

```

A similar effect is also observed in the urban environment, where a single PC explained ~61% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.61, $\phi$ = 7.4, p < 0.05, Fig. 4). This PC correlates positively with continuos central land, dense residential land and water, whilst correlating negatively with sparse residential and amenity grassland. This relationship appears to be largely driven by a single site showing a model of only individual foraging as the best fit. Nevertheless, as with the agri-rural land-uses, these results indicate when resources are sparse, as could be expected in continuos central and dense residential land, where there is limited space for floral resources, a colony emphasises the use of personal rather than social information. However, when resources are more abundant, as could be expected in sparse residential areas and amenity grasslands, a colony appears to increase the use of information contained in the waddle dance. Unlike in the agri-rural environment, water correlates positively with a higher use of the waggle dance. This likely indicates that the banks of urban water ways are less nutritionally rich than in agri-rural landscapes.

```{r, urban.loadings}
urban.loadings %>%
  mutate(PC1 = round(PC1, 3)) %>%
  mutate(`Land-use` = gsub("\\.", " ", `Land-use`)) %>%
  flextable() %>%
  set_caption(
    "Land-use type loadings for the first principle component on the urban sites. Bold face indicates loadings that contribute more than would be expected under a null hypothesis of equal contribution"
  ) %>%
  set_table_properties(layout = "autofit") %>%
  bold(
    i = ~ PC1 > urban_loadings$cutoff |
      PC1 < -urban_loadings$cutoff,
    j = "PC1"
  )

```

This analysis identifies associations between land-use types and the use of the waggle dance which follow what is predicted under theoretical models [@Dornhaus2006; @Beekman2008]. In addition, these findings provide an insight into how human land-use change through urbanisation and agricultural intensification influence how honeybees use the waggle dance. These results indicate that human activity is selecting honeybees to ignore social information in favour of individual information. However, in the urban environment, the land-use type which would be expected to be the most florally rich for pollinators (parks, allotments and cemeteries) showed no significant correlation with the main axis of variation in the proportion of scouts, indicating the relationship is more nuanced than previously thought.

To what extent social information is used when making foraging decisions, as well as the value of personal information, has been a matter of considerable debate (refs). Much of that debate is informed by simulation studies (ref), and empirical work in which a relatively small number of individuals are observed (ref). Here, we demonstrate that the analysis of waggle dance data can make an important contribution to this debate. With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”. Our methodology provides a means of analysing such large data sets and gleaning useful information on the importance of such data to inform the debate about the importance of social information, as well as providing useful colony metrics of foraging activity.

Overall, our results strongly suggest honeybee foraging occurs through collective decision making. In developing this model and applying it to real waggle dance data, our results provide evidence supporting collective foraging in honeybees and supports existing theoretical models describing collective foraging in honeybees. The methods described herein provide a less invasive and empirically reinforced means of approaching questions such as the importance of the waggle dance in foraging, and with this, deeper questions about the value of social vs individual information for honeybees in different environments.

# References

### 2021-11-11 analysis/manuscript/paper.Rmd
edits to manuscript

@@
 ---
-title: "Forage site distributions quantify collective foraging in honeybee colonies"
+title: "Forage site distributions quantify collective foraging in honey bee colonies"
 author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
@@
 Whilst collective foraging has been suggested to occur for many animal species (insert refs list), establishing if, and to what degree, animal groups make collective decisions remains a challenge.
-Here, we present a model of foraging by honeybees that can be used to distinguish collective decisions from individual-driven group behaviour.
+Here, we present a model of foraging by honey bees that can be used to distinguish collective decisions from individual-driven group behaviour.
 By confronting this model with waggle dance data from twenty hives, collected across two years, we can infer to what extent each colony depends on collective foraging vs individual decision-making.
@@
 
-Foraging is a complex task requiring exploration and exploitation of the environment.
-Individual foragers gather and process information independently to decide what and where to forage.
-A collective can foraging differently.
-Through the evolution of 'simple' behavioural rules, individual information can be pooled and assessed collectively, allowing informed foraging decisions to emerge from the interactions between individuals.
-Such systems allow a collective to reap the rewards of working together, however, in some environments individual rather than collective decision making has been shown to provide a more profitable foraging strategy.
-This suggests collective Vs individual information is flexibly used in response to the environment, yet this relationship is poorly understood.
-Understanding how environmental factors influence the prioritisation of different streams of information is critical for identifying the conditions which lead to the evolution of collective decision making and how group living animals might respond to an expanding human world.
-Empirically assessing whether an emergent behaviour is occurring requires comparing the actions and decision making of a collective with that of individual actors.
-This is fundamentally difficult as in most cases we can observe one, or the other, but not both.
+Collective decision making provides animals groups with the ability to pool individual information and make informed decisions.
+Despite the enhanced ability to process information, not every circumstance requires a collective response.
+In many cases, animal groups do better when individuals prioritise their own personal information over social information provided by conspecifics [@King2007; @Rieucau2011; @Gruter2013].
+Identifying the conditions which favour different information use strategies remains a key challenge, both for determining what conditions likely lead to the evolution of collective decision making and for understanding how animal groups might interact in an increasingly anthropomorphic world.
+Nevertheless, assessing the importance of different information use strategies can be extremely difficult, requiring the actions and decision making of a collective to be compared with that of individual actors.
+This is fundamentally difficult as making such comparisons demands observations of individual and collective behaviours.
+A side from limiting the group size that can be studied, in many cases we can observe one, or the other, but not both.
 
-Honeybees are widely understood to forage collectively for nectar due to their use of the celebrated “waggle dance”, which conveys the location of a profitable resource as a vector of distance (duration of the waggle run) and direction (angle of the dance) [@Frisch1993].
+In honey bees, foraging is collectively coordinated via the famous "waggle dance", which conveys the location of a profitable resource as a vector of distance (duration of the waggle run) and direction (angle of the dance) [@Frisch1993].
 Resource “quality” is not encoded in the form of the dance itself, but in the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000].
@@
 This creates a feedback mechanism through which the hive processes the information collected and proportionally selects the most profitable resources [@Seeley1994A].
-Despite the efficiency benefits potentially afforded by this sophisticated collective behaviour, its contribution to colony foraging success remains unclear.
+
+Despite the efficiency benefits the waggle dance reportedly provides, its contribution to colony foraging success remains unclear.
 In many situations, colonies that are prevented from communicating via dance behaviour achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2004; @Price2019].
-In some ecological circumstances, individuals that forgo waiting for dance information in favour of individual search, termed “scouting”, may find food more quickly, and evidence suggests that bees adopt this strategy if the benefits of dance-following drop.
+In some ecological circumstances, individuals that forgo waiting for dance information in favour of individual search, termed “scouting”, may find food more quickly, and evidence suggests that bees adopt this strategy if the benefits of dance-following drop [@Gruter2008].
+Overall, these studies suggest that honey bees use the waggle dance flexibly in response to environmental conditions [@Guter2011; @Gruter2014].
+However, it is difficult to empirically probe this relationship further without new methods which can tease apart the extent to which foraging occurs with or without use of the dance.
 
-Even though this mechanism has been described in detail [@Seeley1995] (+ other refs), here we go beyond this and show that the information in the honeybees’ waggle dance bears evidence of collective decision making in foraging.
-To do this, we evaluate the empirical signatures of collective foraging and collective decision making in honeybee foraging patterns in the resource distances conveyed in the waggle dance.
+Here, we describe a model to quantify waggle dance use in honey bee foraging and fit it to a large empirical waggle dance dataset.
+After estimating the extent to which colony foraging occurs individually or collectively, we investigate the influence different land-use types have on waggle dance use.
+Our methods provide novel evidence for collective honey bee foraging and a framework to evaluate how environmental conditions influence waggle dance use in honey bees.
 
-A foraging scout locates the best resource in what is essentially a one-dimensional environment, searching along a trajectory and reporting the best resource encountered.
-The collective can do something different.
-By pooling and weighting the information present on the dance floor, a collective solves this problem for a two-dimensional environment by integrating multiple sources of information.
-Consequently, the difference in foraging behaviours between scouts and recruits should manifest itself in the distribution of the distances reported on the dance floor, as each sample of dances reportedly comes from a different foraging strategy.
-
-To demonstrate that these two forms of foraging can be distinguished by the resource distance distributions that they generate, we simulated different ways in which collectives can solve such a computational task.
-In our simulations, foragers leave a central place, the hive, explore the environment and report back when they have visited a profitable resource.
-We placed resources of different qualities, representing different types of flower patches, at random positions in the environment.
-Patches were periodically replaced to simulate patch turnover.
-Visits by a forager also reduced a patch's profitability by one to simulate resource depletion.
-
-We simulated honeybee foraging by dividing workers into scouts and recruits.
-Scouts follow a random path from the central place until they find a suitable resource (Fig 1 A) and then return to the hive to dance for this resource.
-Recruits sample a dance randomly from the pool of available dances on the dance-floor and visit this resource, also dancing on return as a scout would (Fig 1 B).
-Resources that are highly profitable, in terms of the quality of the resource and the energy spent in collecting it, elicit more dance runs.
+Whether an individual forager follows a waggle dance from the dance floor or finds a resource for themselves, all foragers report resource information on the dance-floor.
+Consequently, the difference in foraging behaviours between individuals using the dance and individuals relying on personal information should manifest itself in the pattern of dances reported on the dance-floor.
+To evaluate this idea, we simulated honey bees foraging in a landscape using both personal and waggle dance information (Fig.
+1).
+As which individuals foraged under what strategy is known in the simulation, we can compare the distributions of foraging distances reported on the dance floor.
+Figs 1c and 1d show that the shapes of the resource distance distributions under the two simulated scenarios are very different.
+The distribution of the first simulation is akin to that of an exponential distribution (Fig.
+1 C), which is the nearest neighbour distance distribution for foragers operating in a one-dimensional environment (see methods).
+The distribution of the distances reported by recruits (Fig.
+1 D) is appears to resemble a Rayleigh distribution which is the nearest neighbour distribution of a two-dimensional environment (see methods).
 
@@
 A.
-Outline of scout foraging.
-Scouts leave the hive on a random flight path (white lines) and continue until they encounter a resource (flowers in the figure).
+Outline of individual foraging.
+Foragers leave the hive on a random flight path (white lines) and continue until they encounter a resource (flowers in the figure).
 B.
 Outline of foraging with recruitment.
-Scouts continue to identify resources but recruits sample available dances provided by recruits and other scouts.
+Some foragers (scouts) continue to identify resources but others (recruits) sample available dances provided by scouts and other recruits.
 C.
@@
 
-As which individuals forage as recruits or as scouts is known in the simulation, the resulting patterns of foraging distances reported by each group can be examined.
-Assuming patches are placed randomly in an environment their discovery from a random search is akin to sampling from a Poisson process and so the distances between a centrally located hive and different flower patches would be expected to follow an exponential distribution (see methods).
-Recruits on the other hand are thought to be biased towards the more profitable resources.
-Assuming patch profitability should be uniform with respect to the hive’s location, this selection for the most profitable resource should create a bias for resources which are closest to the hive (see methods).
-Considering this bias, we expect the distribution of recruit-reported foraging distances to be well described by a distribution describing the minimum of an exponential.
-
-Fig.s 1c and 1d show that the shapes of the resource distance distributions under the two simulated scenarios are very different.
-The distribution of the first simulation is akin to that of an exponential distribution, which is the nearest neighbour distance distribution for foragers operating in a one-dimensional environment (see methods).
-The distribution of the distances reported by recruits is well described by a Rayleigh distribution which is the nearest neighbour distribution of a two-dimensional environment (see methods).
-
-The shape of the distance distribution, therefore, carries information about the extent to which honeybee colonies collectively make foraging decisions.
-We used this to design a method to analyse to waggle dance data to determine the relative importance of individual and collective behaviour in the foraging of a hive.
-
-The frequency with which a flower patch is reported on the dance-floor is therefore different from the frequency of visits to the resource.
-Foragers report distances to the resources that they have visited on the dance floor, but how often they repeat a dance depends on the profitability of the resource: the more profitable, the more often it is danced for [@Seeley1986; @Seeley1994A].Our simulations show that the distributions of distances reported by scouts and by recruits are different.
-However, on the dance-floor the distribution of distances is modified through the profitability of a resource reported.
-
-Honeybees achieve this by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994A].
-Combined with the distance distributions identified from our simulations, we form a prediction of the distances reported on the dance-floor as a superposition of scout and recruit foraging (see Methods).
-This yields two different models: a collective model demonstrating collective decision making, and an individual model where decisions are made individually.
+Simulating honey bee foraging under a strategy of independent foraging and waggle dance following shows the shape of the foraging distance distributions reported on the dance-floor carries information about the extent to which honey bee colonies collectively make foraging decisions.
+On a real honey bee dance-floor, the distribution of distances is modified through the profitability of a resource reported.
+Honey bees achieve this by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994A].
+Combined with the distance distributions identified from our simulations, we form a prediction of the distances reported on the dance-floor as a superposition of independent and waggle dance foraging (Fig 2, see Methods).
+This yields two different models: a collective model demonstrating collective decision making, comprised of the proportion of dances following an individual foraging strategy (scout) along with the proportion of dances following a waggle dance use strategy (recruit), and an individual model where decisions are made entirely individually (scout only) (Fig 2).
 By confronting these models with waggle dance data we can infer to what extent colony foraging depends on individual or collective decision making.
 
-```{r, fig.cap="Figure 2: Model of honeybee foraging.
+```{r, fig.cap="Figure 2: Model of honey bee foraging.
 Flowers advertised by scouts are distributed exponentially (a).
@@
 
-To evaluate the collective decision making of a honeybee colony, we analysed a dataset of 3378 waggle dance observations from 20 observation hives between April-September 2017, previously described in [@Samuelson2021].
+To evaluate collective decision making in honey bee colonies foraging on 'natural' landscapes we analysed a dataset of 3378 waggle dance observations from 20 observation hives between April-September 2017, previously described in [@Samuelson2021].
 Hives were situated at different locations in the Southeast of England (see methods, figure 3A) and were visited every two weeks for a period of 24 weeks.
 On each visit, two hours of continuous waggle dance data were recorded by training a camcorder onto the dance floor.
-The footage of the dances was decoded manually to identify reported foraging locations [@Couvillon2012; @Schurch2019]
+The footage of the dances was decoded manually to identify reported foraging locations [@Couvillon2012; @Schurch2019].
+For each site we fit both the collective and individual models.
+We used model selection to determine which model provided the better explanation of the data and model fit to determine if the model provided a good explanation of the data.
 
-For 17 out of 20 study hives, the collective model provided a better explanation of the data than the individual model (Fig.
-3a).
+For 17 out of 20 study hives, a model of collective foraging provided a better explanation of the data than one of individual foraging (Fig.
+3a), however, in 3 sites the individual model provided the best explanation of foraging patterns.
 Across 17 of the sites, the collective model was not statistically different from the empirical foraging distances, whereas the scout model was significantly different to the data in 8 of the sites (Fig.
@@
 
-Our results indicate that decision making in honeybee foraging in most cases has a collective component.
-The mechanism of honeybee foraging and observation and how this translates into dancing has been well-studied and long known (needs refs, including Von Frisch?).
-We know that scouts discover resources and recruits randomly sample from a dance floor biased by flower quality from empirical [@Seeley2000] and model studies [@Seeley1991].
-Our findings here go a step further and provide evidence of collective foraging and decision making beyond the confines of an experimental setup designed to identify such behaviours.
+Our results indicate that, whilst foraging decisions are primarily reached collectively through the use of the waggle dance, in some circumstances, colony foraging is best explained by a model of individual foraging.
+This suggests that in some environments, honey bees actively ignore the waggle dance and instead prioritise the use of personal information.
+Previous empirical studies have shown that individual foragers flexibly alter their use of waggle dance information in response to unrewarding feeders [@Gruter2011], suggesting experience drives information use at the individual level.
+Evaluating how these individual decisions influence the collective, however, has always been a challenge due to fact that it is effectively impossible to track an individuals foraging behaviour over a landscape.
 
-Whilst the finding that honeybees forage collectively is unlikely to elicit surprise, our model represents one of the few descriptions of colony foraging which can be related back to theoretical narratives.
-Much of the evidence for the established narrative are rooted in theory, with simulations providing much of the existing numerical descriptions (refs).
-Agent Based Simulations provide a detailed description of a model system, however, such models are often heavily parameterised and not assessed for parsimony through means such as AIC.
-Due to the nature of these methods, it is very difficult to compare findings to empirical data.
-The model outlined here is comparable to data and does not require any manipulation of the colony being studied unlike other empirical methods.
+As well as providing direct evidence that honey bees forage collectively according to the widely accepted theory (refs), our results demonstrate that waggle dance use depends on environmental context.
+Although this has been suspected for some time (refs), empirical evidence that colonies foraging on the landscape might regulate the use of the waggle dance have relied on the method of dance disorientation (refs for dance disorientation study), whereby a colony is disoriented to remove the spatial component of the waggle dance.
+Foraging success, measured as quantities of nectar or pollen retrieved, is then compared to an undisturbed colony in the same environment to determine what dance use strategy is the most effective.
+Whilst this method allows for an evaluation of whether or not colonies forage more effectively with or without the dance in a given environment, whether a colony actively regulates waggle dance use in response to the environment is not directly shown.
+In showing that foraging patterns in three of our sites are best explained by individual foraging, our results demonstrate honey bee colonies tune waggle dance use in relation to environmental factors.
 
-By taking decisions collectively, a colony of honeybees acquires the computational means of locating the best resource in its environment.
-In this sense, the capability of the hive exceeds those of an individual forager.
-Honeybee colonies have often been held up as an example of a superorganism, where the collective can do things beyond the capabilities of individuals.
-In terms of foraging, our results largely support this view and underline the importance of collective decision making in honeybees.
-Nevertheless, whilst most of our study hives showed the collective model provided the best fit, for some sites the individual model outperformed the collective model, reinforcing the idea the waggle dance is used flexibly in response to the environment (refs).
+As our model predicts the proportion of dances coming from an individual or collective foraging strategy, its fit to waggle dance data provides a tool to estimate a colonies use of the waggle dance.
 
-The idea that the waggle dance represents a plastic response is not new.
-Simulations and empirical studies [@Dornhaus2006; @Beekman2008; @Price2019] have suggested the waggle dance may not always provide the optimal foraging strategy for a colony, instead suggesting it is more productive to forage individually when resources are sparse.
-In many cases it is difficult to evaluate such claims by observing ‘natural’ hives foraging on the landscape rather than feeders.
-Nevertheless, empirical studies using hive disorientation, whereby the ability to gain accurate information from the waggle dance is removed, have suggested it can be more beneficial to ignore the dance and forage individually, supporting the findings from simulation studies.
-This approach does not remove all aspects of social information, such as olfactory cues, and can only show if there is a benefit to using the waggle dance, not whether a colony actively down regulates their use of social information [@Hasenjager2020].
-
-The model outlined here provides an estimate for the use of the waggle dance by a colony in the form of the proportion of scout dances.
-By quantifying waggle dance use, our model provides a tool to study variation in the use of social information across ecological scenarios.
-To demonstrate this, we first performed a land-use classification of the environment around each of our study sites (see methods, cite the land use paper by Ash).
-As our sites were in either an agricultural-rural or urban landscape, we developed different land-use categories for the two groups and performed a partial least squares analysis to identify the land-use type combinations which explained the most variation in the proportion of scouts across the two environments.
+To evaluate how waggle dance use is being influenced by the environment, we compared the different land-use types for each site to the estimated proportion of dances stemming from an individual foraging strategy (proportion of scout dances).
+Each site used in this study has a unique landscape comprised of different land-use types.
+We first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments.
+As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis to determine the combination of land-use types which explained the most variation in the proportion of scout dances.
 
@@
 This PC correlates positively with arable land and negatively with non-agricultural unmanaged green space, built up areas and water.
-Arable land is typically nutritionally poor for honeybees (ref).
+Arable land is typically nutritionally poor for honey bees (ref).
 On the contrary, non-agricultural unmanaged green space is typically nutrient rich, as are built up areas in agri-rural environments, which often have gardens with ample flowers available for pollinators (ref).
@@
   )
-
 ```
@@
 This relationship appears to be largely driven by a single site showing a model of only individual foraging as the best fit.
-Nevertheless, as with the agri-rural land-uses, these results indicate when resources are sparse, as could be expected in continuos central and dense residential land, where there is limited space for floral resources, a colony emphasises the use of personal rather than social information.
-However, when resources are more abundant, as could be expected in sparse residential areas and amenity grasslands, a colony appears to increase the use of information contained in the waddle dance.
+As with the agri-rural land-uses, these results indicate when resources are sparse, as could be expected in continuos central and dense residential land, where there is limited space for floral resources, a colony emphasises the use of personal rather than social information.
+However, when resources are more abundant, as could be expected in sparse residential areas and amenity grasslands, a colony appears to increase the use of information contained in the waggle dance.
 Unlike in the agri-rural environment, water correlates positively with a higher use of the waggle dance.
@@
   )
-
 ```
 
-This analysis identifies associations between land-use types and the use of the waggle dance which follow what is predicted under theoretical models [@Dornhaus2006; @Beekman2008].
-In addition, these findings provide an insight into how human land-use change through urbanisation and agricultural intensification influence how honeybees use the waggle dance.
-These results indicate that human activity is selecting honeybees to ignore social information in favour of individual information.
-However, in the urban environment, the land-use type which would be expected to be the most florally rich for pollinators (parks, allotments and cemeteries) showed no significant correlation with the main axis of variation in the proportion of scouts, indicating the relationship is more nuanced than previously thought.
+In both the agri-rural and Urban environments, land-use types which associate with resource scarcity correlate positively with a higher proportion of scouts, whilst those which correlate with resource abundance correlate with a lower proportion of scouts, indicating a greater use of the waggle dance.
+These results appear to follow what is predicted under theoretical models: in landscapes where resources are likely to more difficult to find, a colony tends to have a higher number of scouts than in landscapes where resources should be easier to locate [@Dornhaus2006; @Beekman2008].
+However, it is important to note that other combinations which didn't show as having a significant association with the first principle component could have just as well fit into this argument, arguably better than others.
+In both the agri-rural and urban environments water seems to correlate in contrasting ways with the proportion of scouts.
+This could be due to differences in the floral richness of urban and agri-rural water ways (reff), however, if land-use type is to be taken as a proxy for floral availability, other variables such as parks, allotments and cemeteries in the urban environment and fruit or Oil Seed Rape in the agri-rural environment might be expected to be included in the first PC over water.
+An alternative explanation could be that water acts as barriers between land-use types and so gets bundled into the association (more thought on this needed i think).
 
@@
 Much of that debate is informed by simulation studies (ref), and empirical work in which a relatively small number of individuals are observed (ref).
-Here, we demonstrate that the analysis of waggle dance data can make an important contribution to this debate.
+Here, we demonstrate that the analysis of waggle dance data can make an important contribution to this debate and provide a methodology to further evaluate how honey bees use this unique behaviour.
 With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”.
@@
 
-Overall, our results strongly suggest honeybee foraging occurs through collective decision making.
-In developing this model and applying it to real waggle dance data, our results provide evidence supporting collective foraging in honeybees and supports existing theoretical models describing collective foraging in honeybees.
-The methods described herein provide a less invasive and empirically reinforced means of approaching questions such as the importance of the waggle dance in foraging, and with this, deeper questions about the value of social vs individual information for honeybees in different environments.
-
 # References

### 2021-11-12 analysis/manuscript/paper.Rmd
latest update to manuscript

@@
 
+Notes from Elli.
+
+The big thing is collective vs individual information use and what caused the evolution of sociality.
+With honeybees we have an advanced dance language.
+Wider context question is of how the waggle dance has evolved / what are the conditions that lead to the evolution of the waggle dance, as it has only happened once.
+This is sort of what Dornhaus, Herman and Visscher are hinting at.
+
+Gruter's stuff shows individuals can vary their use of the dance depending on foraging experience.
+The hive disorientation stuff shows that a collective can do better without the dance in some environments.
+Suggests honeybees flexibly use the dance in colony foraging.
+However, despite this we have no way of quantifying the extent to which honeybees use the dance.
+We designed a method to quantify its use and fit it against a large empirical data set.
+
+Results are that honeybees in the majority of cases forage collectively but in some sites use personal information. We can then apply the results to evaluate how this varies across environments.
+
+$$
+\pi r^2
+$$
+
 # Abstract
@@
 
-Collective decision making provides animals groups with the ability to pool individual information and make informed decisions.
-Despite the enhanced ability to process information, not every circumstance requires a collective response.
+Collective decision making provides animals groups with the ability to pool individual information and make more informed decisions.
+Despite an enhanced processing capacity, not every circumstance requires a collective response.
 In many cases, animal groups do better when individuals prioritise their own personal information over social information provided by conspecifics [@King2007; @Rieucau2011; @Gruter2013].
 Identifying the conditions which favour different information use strategies remains a key challenge, both for determining what conditions likely lead to the evolution of collective decision making and for understanding how animal groups might interact in an increasingly anthropomorphic world.
-Nevertheless, assessing the importance of different information use strategies can be extremely difficult, requiring the actions and decision making of a collective to be compared with that of individual actors.
-This is fundamentally difficult as making such comparisons demands observations of individual and collective behaviours.
+Assessing the importance of different information use strategies is not trivial, requiring the actions and decision making of a collective to be compared with that of individual actors.
+Making such comparisons demands observations of individual and collective behaviours.
 A side from limiting the group size that can be studied, in many cases we can observe one, or the other, but not both.
@@
 In some ecological circumstances, individuals that forgo waiting for dance information in favour of individual search, termed “scouting”, may find food more quickly, and evidence suggests that bees adopt this strategy if the benefits of dance-following drop [@Gruter2008].
-Overall, these studies suggest that honey bees use the waggle dance flexibly in response to environmental conditions [@Guter2011; @Gruter2014].
-However, it is difficult to empirically probe this relationship further without new methods which can tease apart the extent to which foraging occurs with or without use of the dance.
+Whilst direct experiments can show how individuals respond to different scenarios [@Guter2011; @Gruter2014], empirically evaluating a collective foraging strategy in response to different environmental conditions remains a challenge.
 
@@
 To what extent social information is used when making foraging decisions, as well as the value of personal information, has been a matter of considerable debate (refs).
-Much of that debate is informed by simulation studies (ref), and empirical work in which a relatively small number of individuals are observed (ref).
 Here, we demonstrate that the analysis of waggle dance data can make an important contribution to this debate and provide a methodology to further evaluate how honey bees use this unique behaviour.
+Recent advances in computational methods have allowed individuals to be tracked within colonies in ways which would be almost impossible to do by hand (Langraff paper).
+When combined with social network analyses, the use of such technologies have shed new light on many in colony behaviours.
+By piecing together the behavioural response of individuals into a mathematical description of colony foraging, the model outlined here extends our ability to predict individual behaviour beyond the colony walls and in to the environment.
 With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”.

### 2021-11-16 analysis/manuscript/paper.Rmd
updated manuscript and sent to Vincent and Elli

--- 2021-11-11 analysis/manuscript/paper.Rmd
@@
 
-# Main
+# Main (Word count excluding figures and legend = 2357)
 
-Collective decision making provides animals groups with the ability to pool individual information and make informed decisions.
-Despite the enhanced ability to process information, not every circumstance requires a collective response.
-In many cases, animal groups do better when individuals prioritise their own personal information over social information provided by conspecifics [@King2007; @Rieucau2011; @Gruter2013].
-Identifying the conditions which favour different information use strategies remains a key challenge, both for determining what conditions likely lead to the evolution of collective decision making and for understanding how animal groups might interact in an increasingly anthropomorphic world.
-Nevertheless, assessing the importance of different information use strategies can be extremely difficult, requiring the actions and decision making of a collective to be compared with that of individual actors.
-This is fundamentally difficult as making such comparisons demands observations of individual and collective behaviours.
-A side from limiting the group size that can be studied, in many cases we can observe one, or the other, but not both.
+Collective decision making provides animals groups with the ability to pool individual experiences and make more informed decisions.
+Despite an enhanced capacity to process information, not every circumstance requires a collective response.
+In many cases, animal groups have been shown to do better when individuals prioritise information they obtain through their own experiences to that provided by conspecifics [@King2007; @Rieucau2011; @Gruter2013].
+Identifying the conditions which favour different information use strategies is therefore important, both for determining what conditions might have lead to the evolution of sociality and for understanding how animal groups might interact in an increasingly anthropomorphic world.
+However, assessing the importance of different information use strategies is not trivial and requires comparing the actions and behaviours of individuals with that of a collective.
+Often, information from different sources gets pooled together and so determining the importance of social Vs individual information use requires tools which can tease apart individual and collective behaviours.
 
-In honey bees, foraging is collectively coordinated via the famous "waggle dance", which conveys the location of a profitable resource as a vector of distance (duration of the waggle run) and direction (angle of the dance) [@Frisch1993].
+Honeybees provide an ideal study system in which to evaluate social information use thanks to the famous "waggle dance", which conveys the location of a profitable resource as a vector of distance (duration of the waggle run) and direction (angle of the dance) [@Frisch1993].
 Resource “quality” is not encoded in the form of the dance itself, but in the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000].
@@
 In some ecological circumstances, individuals that forgo waiting for dance information in favour of individual search, termed “scouting”, may find food more quickly, and evidence suggests that bees adopt this strategy if the benefits of dance-following drop [@Gruter2008].
-Overall, these studies suggest that honey bees use the waggle dance flexibly in response to environmental conditions [@Guter2011; @Gruter2014].
-However, it is difficult to empirically probe this relationship further without new methods which can tease apart the extent to which foraging occurs with or without use of the dance.
+Whilst individuals can be observed under experimental conditions to determine how they use different sources of information, translating this up to a colony level is logistically impossible due to honey bees small size, extensive foraging ranges and shear number of foragers.
+New methods which can quantify waggle dance use at a colony level are therefore needed if we are to assess the conditions where the dance provides the most benefit and derive the conditions user which this unique behaviour evolved.
 
-Here, we describe a model to quantify waggle dance use in honey bee foraging and fit it to a large empirical waggle dance dataset.
-After estimating the extent to which colony foraging occurs individually or collectively, we investigate the influence different land-use types have on waggle dance use.
-Our methods provide novel evidence for collective honey bee foraging and a framework to evaluate how environmental conditions influence waggle dance use in honey bees.
-
-Whether an individual forager follows a waggle dance from the dance floor or finds a resource for themselves, all foragers report resource information on the dance-floor.
-Consequently, the difference in foraging behaviours between individuals using the dance and individuals relying on personal information should manifest itself in the pattern of dances reported on the dance-floor.
+Whether an individual forager follows a waggle dance from the dance floor or finds a resource for themselves, all foragers report resource information on the dance-floor [@Seeley1995].
+Consequently, the difference in foraging behaviours between individuals following the dance and individuals relying on personal information should manifest itself in the pattern of dances reported on the dance-floor.
 To evaluate this idea, we simulated honey bees foraging in a landscape using both personal and waggle dance information (Fig.
 1).
-As which individuals foraged under what strategy is known in the simulation, we can compare the distributions of foraging distances reported on the dance floor.
+As which individuals forage under what strategy is known in the simulation, we can compare the distributions of foraging distances reported on the dance floor.
 Figs 1c and 1d show that the shapes of the resource distance distributions under the two simulated scenarios are very different.
@@
 
-As well as providing direct evidence that honey bees forage collectively according to the widely accepted theory (refs), our results demonstrate that waggle dance use depends on environmental context.
-Although this has been suspected for some time (refs), empirical evidence that colonies foraging on the landscape might regulate the use of the waggle dance have relied on the method of dance disorientation (refs for dance disorientation study), whereby a colony is disoriented to remove the spatial component of the waggle dance.
-Foraging success, measured as quantities of nectar or pollen retrieved, is then compared to an undisturbed colony in the same environment to determine what dance use strategy is the most effective.
-Whilst this method allows for an evaluation of whether or not colonies forage more effectively with or without the dance in a given environment, whether a colony actively regulates waggle dance use in response to the environment is not directly shown.
-In showing that foraging patterns in three of our sites are best explained by individual foraging, our results demonstrate honey bee colonies tune waggle dance use in relation to environmental factors.
-
-As our model predicts the proportion of dances coming from an individual or collective foraging strategy, its fit to waggle dance data provides a tool to estimate a colonies use of the waggle dance.
+As well as providing direct evidence that honey bees forage collectively according to a widely accepted theory [@Seeley1994; @Sherman2002], in showing that foraging patterns in three of our sites are best explained by individual foraging, our results demonstrate honey bee colonies tune waggle dance use in relation to environmental factors.
+These results support previous efforts to quantify collective foraging [@Sherman2002; @Dornhaus2006; @Gruter2013; @Price2019].
+Furthermore, as our model predicts the proportion of dances coming from an individual or social foraging strategy, its fit to waggle dance data provides a tool to estimate a colonies use of the waggle dance.
 
@@
 Water also correlated negatively with the estimated proportion of scouts, possibly due to river and pond banks providing good floral resources for pollinators (ref).
-Combined, these results suggests when resources are sparse and difficult to find, as with arable land-use, individuals value personal information more than when easier to find, as with the non-agricultural unmanaged green space, built up areas and water.
+Combined, these results suggest when resources are sparse and difficult to find, as with arable land-use, individuals value personal information more than when easier to find, as with the non-agricultural unmanaged green space, built up areas and water.
 
@@
 4).
-This PC correlates positively with continuos central land, dense residential land and water, whilst correlating negatively with sparse residential and amenity grassland.
+This PC correlates positively with continuous central land, dense residential land and water, whilst correlating negatively with sparse residential and amenity grassland.
 This relationship appears to be largely driven by a single site showing a model of only individual foraging as the best fit.
-As with the agri-rural land-uses, these results indicate when resources are sparse, as could be expected in continuos central and dense residential land, where there is limited space for floral resources, a colony emphasises the use of personal rather than social information.
+As with the agri-rural land-uses, these results indicate when resources are sparse, as could be expected in continuous central and dense residential land, where there is limited space for floral resources, a colony emphasises the use of personal rather than social information.
 However, when resources are more abundant, as could be expected in sparse residential areas and amenity grasslands, a colony appears to increase the use of information contained in the waggle dance.
@@
 In both the agri-rural and Urban environments, land-use types which associate with resource scarcity correlate positively with a higher proportion of scouts, whilst those which correlate with resource abundance correlate with a lower proportion of scouts, indicating a greater use of the waggle dance.
-These results appear to follow what is predicted under theoretical models: in landscapes where resources are likely to more difficult to find, a colony tends to have a higher number of scouts than in landscapes where resources should be easier to locate [@Dornhaus2006; @Beekman2008].
+These results appear to follow what is predicted under theoretical models: in landscapes where resources are likely to more difficult to find, a colony tends to have a higher number of scouts than in landscapes where resources should be easier to locate [@Dornhaus2006; @Price2019].
 However, it is important to note that other combinations which didn't show as having a significant association with the first principle component could have just as well fit into this argument, arguably better than others.
@@
 This could be due to differences in the floral richness of urban and agri-rural water ways (reff), however, if land-use type is to be taken as a proxy for floral availability, other variables such as parks, allotments and cemeteries in the urban environment and fruit or Oil Seed Rape in the agri-rural environment might be expected to be included in the first PC over water.
-An alternative explanation could be that water acts as barriers between land-use types and so gets bundled into the association (more thought on this needed i think).
+An alternative explanation could be that water acts as barriers between land-use types and so gets bundled into the association (more thought on this needed I think).
 
-To what extent social information is used when making foraging decisions, as well as the value of personal information, has been a matter of considerable debate (refs).
-Much of that debate is informed by simulation studies (ref), and empirical work in which a relatively small number of individuals are observed (ref).
-Here, we demonstrate that the analysis of waggle dance data can make an important contribution to this debate and provide a methodology to further evaluate how honey bees use this unique behaviour.
+Here, we demonstrate that the analysis of waggle dance data can make an important contribution to our understanding of social information use and provide a methodology to further evaluate how honey bees use this unique behaviour.
+Recently, technological advances have emerged which enable colony metrics to be collected faster, more accurately and over greater time spans than could be gathered by hand .
+
+Recent advances in computational methods have allowed individuals to be tracked within colonies in ways which would be almost impossible to do by hand, allowing theories of individual behaviour to be evaluated in more depth than could have been done previously [@Wild2021].
+We are not yet in a position to be able to track behaviours outside the colony in the same detail, nevertheless, by piecing together the behavioural response of individuals into a mathematical description of colony foraging, the model outlined here extends our ability to predict individual behaviour beyond the colony walls and in to the environment.
 With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”.
-Our methodology provides a means of analysing such large data sets and gleaning useful information on the importance of such data to inform the debate about the importance of social information, as well as providing useful colony metrics of foraging activity.
+Our methodology provides a means of analysing such large data sets to inform the debate about the importance of social information, as well as providing useful colony metrics of foraging activity.
 

### 2021-11-16 analysis/landuse_effect_prop_scouts.Rmd
started work on updating figure

@@
 urban_plotdata$`Land-use` <- "Urban"
-
 ```
@@
 agrirural_plotdata$`Land-use` <- "Agri-rural"
-
 ```
@@
 )
-
+```
+
+## Land-use maps
+
+```{r, agrirural_landuse_maps, message = FALSE, warning = FALSE}
+library(sf)
+library(sp)
+library(ggspatial)
+
+# make fit map with agrirural landuses
+agrirural_plotdata$site <- rural_lu$site
+agrirural_plotdata$prop_scouts <- rural_lu$p
+ar_sortedpropscouts <- arrange(agrirural_plotdata, prop_scouts)
+ar_marked <- ar_sortedpropscouts[
+  c(1, nrow(ar_sortedpropscouts) / 2, nrow(ar_sortedpropscouts)),
+]
+
+# make model fit graph of pc1 against prop scouts and mark the sites with the
+# lowest, middle and highest proportion of scouts
+ar_fit <- ggplot(
+  data = ar_sortedpropscouts, aes(x = Comp_.1, y = prop_scouts)
+) +
+  geom_point() +
+  geom_point(
+    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, colour = "red")
+  ) +
+  geom_text(
+    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, label = site),
+    nudge_x = c(0, 0), nudge_y = c(-0.05, -0.05), size = 8
+  ) +
+  geom_line(data = rural_fitdata, aes(x = Comp_.1, y = fit)) +
+  geom_ribbon(
+    data = rural_fitdata,
+    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
+    alpha = 0.2,
+    fill = "deepskyblue"
+  ) +
+  theme(legend.position = "none") +
+  xlab("PC 1") +
+  ylab("Estimated proportion of scouts")
+
+# make the maps for each of the sites selected in the previous step
+site_shape_data <- list()
+for (i in seq_len(3)) {
+  path <- paste0(
+    "shapefiles/agri-rural/",
+    ar_marked[i, "site"], "/", ar_marked[i, "site"],
+    ".shp"
+  )
+  df <- st_read(path) %>%
+    st_transform(
+      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
+    )
+  site_shape_data[[ar_marked[i, "site"]]] <- df
+}
+
+site_shape_data$STU$descriptio
+
+ggplot() +
+  geom_sf(
+    data = site_shape_data$STU
+  )
 ```
@@
 
-plotdata$LU = ""
-plotdata$LU[1:10] = "Rural"
-plotdata$LU[11:20] = "Urban"
-
-ggplot(data = plotdata,
-  aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = p)) +
+plotdata$LU <- ""
+plotdata$LU[1:10] <- "Rural"
+plotdata$LU[11:20] <- "Urban"
+
+ggplot(
+  data = plotdata,
+  aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = p)
+) +
   geom_point(aes(colour = LU)) +
@@
 
-ggplot(data = plotdata,
-  aes(x = LU, y = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4)) +
+ggplot(
+  data = plotdata,
+  aes(x = LU, y = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4)
+) +
   geom_boxplot(aes(group = LU))
-
-```
+```

### 2021-11-17 analysis/landuse_effect_prop_scouts.Rmd
created new landuse plots for the PLS analysis. Awaiting a land-use conversion document from Elli, once I have this I can colour the map with the significant land-use catagories and thus show visually what changes in the environment as the proportion of scouts increases

@@
 library(ggplot2)
-theme_set(
-  theme_classic() +
-    theme(
-      text = element_text(size = 42)
-    )
-)
+# theme_set(
+#   theme_classic() +
+#     theme(
+#       text = element_text(size = rel(42))
+#     )
+# )
 
@@
 
+Here I create figures which show the relationship between PC1 and the proportion of scouts as well as land-use maps of sites long the gradient which are coloured by the land-use types identified as most singificant in the PLS (PC1).
+
+### Agri-rural
 ```{r, agrirural_landuse_maps, message = FALSE, warning = FALSE}
@@
 ar_sortedpropscouts <- arrange(agrirural_plotdata, prop_scouts)
-ar_marked <- ar_sortedpropscouts[
-  c(1, nrow(ar_sortedpropscouts) / 2, nrow(ar_sortedpropscouts)),
-]
-
-# make model fit graph of pc1 against prop scouts and mark the sites with the
-# lowest, middle and highest proportion of scouts
+
+# select sites to draw attention to, those sitting on the trend
+ar_marked <- ar_sortedpropscouts %>%
+  filter(site %in% c("MEL", "SOM", "BEL"))
+
+# make model fit graph of pc1 against prop scouts and mark the sites
 ar_fit <- ggplot(
@@
   geom_point() +
-  geom_point(
-    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, colour = "red")
-  ) +
-  geom_text(
-    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, label = site),
-    nudge_x = c(0, 0), nudge_y = c(-0.05, -0.05), size = 8
-  ) +
   geom_line(data = rural_fitdata, aes(x = Comp_.1, y = fit)) +
@@
   ) +
+  geom_point(
+    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, colour = "red")
+  ) +
+  geom_text(
+    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, label = site),
+    nudge_x = c(0, 0), nudge_y = c(-0.05, 0.05), size = rel(3)
+  ) +
+  theme_classic() +
   theme(legend.position = "none") +
@@
 site_shape_data <- list()
+site_plots <- list()
 for (i in seq_len(3)) {
@@
   site_shape_data[[ar_marked[i, "site"]]] <- df
+  site_plots[[ar_marked[i, "site"]]] <- ggplot() +
+    geom_sf(
+      data = df,
+      aes(fill = descriptio)
+    ) +
+    theme_void()
 }
 
-site_shape_data$STU$descriptio
-
-ggplot() +
+# make loadings plot
+argi.rural.loadings <- tibble(
+  "Landuse" = names(agrirural_loadings$loadings),
+  "Correlation" = agrirural_loadings$loadings
+) %>%
+  mutate(
+    significant = ifelse(
+      Correlation > agrirural_loadings$cutoff,
+      Landuse,
+      ifelse(
+        Correlation < -agrirural_loadings$cutoff,
+        Landuse,
+        "Other"
+      )
+    )
+  ) %>%
+  mutate(
+    Landuse = ifelse(
+      Landuse == "Non.agricultural.managed.green.space",
+      "NAMGS",
+      ifelse(
+        Landuse == "Non.agricultural.unmanaged.green.space",
+        "NAUMGS",
+        Landuse
+      )
+    )
+  )
+
+loading_plot <- ggplot(argi.rural.loadings, aes(x = Landuse, y = Correlation)) +
+  geom_point(aes(colour = significant), size = rel(3)) +
+  geom_segment(aes(
+    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
+  )) +
+  geom_text(
+    aes(x = Landuse, y = Correlation, label = Landuse),
+    hjust = 0.25,
+    nudge_y = c(0, 0), nudge_x = c(0.5, 0.5), size = rel(3)
+  ) +
+  geom_hline(
+    yintercept = 0, color = "black"
+  ) +
+  geom_rect(
+    aes(
+      ymin = -agrirural_loadings$cutoff,
+      ymax = agrirural_loadings$cutoff,
+      xmin = 0, xmax = Inf
+    ),
+    alpha = 0.01
+  ) +
+  coord_flip() +
+  theme_classic() +
+  theme(
+    axis.line.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "None"
+  ) +
+  scale_colour_manual(
+    values = c(
+      "arable" = "red3", "built.up.area" = "orange3",
+      "Non.agricultural.unmanaged.green.space" = "seagreen3",
+      "water" = "cornflowerblue", other = "grey"
+    )
+  ) +
+  ylim(-0.6, 0.6)
+
+loading_plot
+
+# combine plots into facet
+
+# first the maps in a column of three
+sitelumaps <- plot_grid(
+  site_plots[[3]] + theme(legend.position = "none"),
+  site_plots[[2]] + theme(legend.position = "none"),
+  site_plots[[1]] + theme(legend.position = "none"),
+  ncol = 1,
+  align = "v",
+  labels = c(
+    names(site_plots)[3], names(site_plots)[2], names(site_plots)[1]
+  )
+)
+
+# now the model fit and pls results in a plot of two
+mfitpls <- plot_grid(
+  ar_fit + theme(plot.margin = unit(c(0., 0., 1., 0.), "cm")),
+  loading_plot,
+  ncol = 1,
+  align = "v",
+  labels = c(
+    "A", "B"
+  )
+)
+
+# all together
+
+comb <- plot_grid(
+  mfitpls, sitelumaps,
+  nrow = 1, scale = 0.9, align = "h"
+)
+
+comb
+```
+
+### Urban
+
+```{r, agrirural_landuse_maps, message = FALSE, warning = FALSE}
+
+# make fit map with agrirural landuses
+urban_plotdata$site <- urban_lu$site
+urban_plotdata$prop_scouts <- urban_lu$p
+u_sortedpropscouts <- arrange(urban_plotdata, prop_scouts)
+u_marked <- u_sortedpropscouts %>%
+  filter(site %in% c("HOR", "SWP", "CAD"))
+
+# make model fit graph of pc1 against prop scouts and mark the sites with the
+# lowest, middle and highest proportion of scouts
+u_fit <- ggplot(
+  data = u_sortedpropscouts, aes(x = Comp_.1, y = prop_scouts)
+) +
+  geom_point() +
+  geom_point(
+    data = u_marked, aes(x = Comp_.1, y = prop_scouts, colour = "red")
+  ) +
+  geom_text(
+    data = u_marked, aes(x = Comp_.1, y = prop_scouts, label = site),
+    nudge_x = c(0, 0), nudge_y = c(-0.05, 0.05), size = rel(3)
+  ) +
+  geom_line(data = urban_fitdata, aes(x = Comp_.1, y = fit)) +
+  geom_ribbon(
+    data = urban_fitdata,
+    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
+    alpha = 0.2,
+    fill = "deepskyblue"
+  ) +
+  theme_classic() +
+  theme(legend.position = "none") +
+  xlab("PC 1") +
+  ylab("Estimated proportion of scouts")
+
+# make the maps for each of the sites selected in the previous step
+site_shape_data <- list()
+site_plots <- list()
+for (i in seq_len(3)) {
+  path <- paste0(
+    "shapefiles/urban/",
+    u_marked[i, "site"], "/", u_marked[i, "site"],
+    ".shp"
+  )
+  df <- st_read(path) %>%
+    st_transform(
+      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
+    )
+  site_shape_data[[u_marked[i, "site"]]] <- df
+  site_plots[[u_marked[i, "site"]]] <- ggplot() +
     geom_sf(
-    data = site_shape_data$STU
-  )
-```
-
+      data = df,
+      aes(fill = descriptio)
+    ) +
+    theme_void()
+}
+
+# make loadings plot
+urban.loadings <- tibble(
+  "Landuse" = names(urban_loadings$loadings),
+  "Correlation" = urban_loadings$loadings
+) %>%
+  mutate(
+    significant = ifelse(
+      Correlation > urban_loadings$cutoff,
+      Landuse,
+      ifelse(
+        Correlation < -urban_loadings$cutoff,
+        Landuse,
+        "Other"
+      )
+    )
+  )
+
+loading_plot <- ggplot(urban.loadings, aes(x = Landuse, y = Correlation)) +
+  geom_point(aes(colour = significant), size = rel(3)) +
+  geom_segment(aes(
+    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
+  )) +
+  geom_text(
+    aes(x = Landuse, y = Correlation, label = Landuse),
+    hjust = 0.25,
+    nudge_y = c(-0, -0), nudge_x = c(0.5, 0.5), size = rel(3)
+  ) +
+  geom_rect(
+    aes(
+      ymin = -agrirural_loadings$cutoff,
+      ymax = agrirural_loadings$cutoff,
+      xmin = 0, xmax = Inf
+    ),
+    alpha = 0.01
+  ) +
+  coord_flip() +
+  theme_classic() +
+  theme(
+    axis.line.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "None"
+  ) +
+  scale_colour_manual(
+    values = c(
+      "continuous..central" = "red3", "dense.residential" = "orange3",
+      "sparse.residential" = "purple3", "amenity.grassland" = "seagreen3",
+      "water" = "cornflowerblue", other = "grey"
+    )
+  ) +
+  scale_y_continuous(breaks = seq(-.9, .9, by = 0.4), limits = c(-1, 1))
+
+loading_plot
+
+# combine plots into facet
+
+# first the maps in a column of three
+sitelumaps <- plot_grid(
+  site_plots[[3]] + theme(legend.position = "none"),
+  site_plots[[2]] + theme(legend.position = "none"),
+  site_plots[[1]] + theme(legend.position = "none"),
+  ncol = 1,
+  align = "v",
+  labels = c(
+    names(site_plots)[3], names(site_plots)[2], names(site_plots)[1]
+  )
+)
+
+# now the model fit and pls results in a plot of two
+mfitpls <- plot_grid(
+  u_fit + theme(plot.margin = unit(c(0., 0., 1., 0.), "cm")),
+  loading_plot,
+  ncol = 1,
+  align = "v",
+  labels = c(
+    "A", "B"
+  )
+)
+
+# all together
+
+comb <- plot_grid(
+  mfitpls, sitelumaps,
+  nrow = 1, scale = 0.9, align = "h"
+)
+
+comb
+```
+
+# Loadings
 ```{r}
@@
   kbl(
-    caption = "Land-use type loadings for the first principle component on the Agri-rural sites",
+    caption = "Land-use type loadings for the first principle component on the
+    Agri-rural sites",
     escape = F
@@
   kbl(
-    caption = "Land-use type loadings for the first principle component on the Urban sites",
+    caption = "Land-use type loadings for the first principle component on the
+     Urban sites",
     escape = F

### 2021-11-26 analysis/landuse_effect_prop_scouts.Rmd
updated LU map plots

@@
 library(ggplot2)
-# theme_set(
-#   theme_classic() +
-#     theme(
-#       text = element_text(size = rel(42))
-#     )
-# )
+theme_set(
+  theme_classic()
+)
 
@@
 
-# check the land-use type values sum to 100%
-testthat::test_that(
-  "land-use type values sum to 100",
-  {
-    expected <- rep(100, 10)
-    ans <- rowSums(rural_lu[, 3:ncol(rural_lu)])
-    testthat::expect_identical(expected, ans)
-  }
-)
+# # check the land-use type values sum to 100%
+# testthat::test_that(
+#   "land-use type values sum to 100",
+#   {
+#     expected <- rep(100, 10)
+#     ans <- rowSums(rural_lu[, 3:ncol(rural_lu)])
+#     testthat::expect_identical(expected, ans)
+#   }
+# )
 ```
@@
 
+# load land-use conversion table
+luconversions <- read.csv("data/LUCategoryConversion.csv")
+agrirural_luconversions <- luconversions %>%
+  filter(LU == "agri-rural")
+
 # make fit map with agrirural landuses
@@
 # make the maps for each of the sites selected in the previous step
-site_shape_data <- list()
+site_shape_data_ar <- list()
 site_plots <- list()
@@
     )
-  site_shape_data[[ar_marked[i, "site"]]] <- df
+  # add major LU categories
+  df_merged <- df %>%
+    mutate(descriptio = toupper(descriptio)) %>%
+    full_join(agrirural_luconversions, by = c("descriptio" = "code"))
+  site_shape_data_ar[[ar_marked[i, "site"]]] <- df_merged
   site_plots[[ar_marked[i, "site"]]] <- ggplot() +
     geom_sf(
-      data = df,
-      aes(fill = descriptio)
-    ) +
-    theme_void()
+      data = df_merged,
+      aes(fill = category)
+    ) +
+    theme_void() +
+    scale_fill_manual(
+      values = c(
+        "arable" = "#EE6677", "built-up area" = "#CCBB44",
+        "Non-agricultural unmanaged green space" = "#228833",
+        "water" = "#4477AA", "pasture" = "#BBBBBB", "woodland" = "#BBBBBB",
+        "Non-agricultural managed green space" = "#BBBBBB", "fruit" = "#BBBBBB",
+        "OSR" = "#BBBBBB", "other agricultural" = "#BBBBBB"
+      )
+    )
 }
@@
     values = c(
-      "arable" = "red3", "built.up.area" = "orange3",
-      "Non.agricultural.unmanaged.green.space" = "seagreen3",
-      "water" = "cornflowerblue", other = "grey"
+      "arable" = "#EE6677", "built.up.area" = "#CCBB44",
+      "Non.agricultural.unmanaged.green.space" = "#228833",
+      "water" = "#4477AA", other = "#BBBBBB"
     )
@@
   ylim(-0.6, 0.6)
-
-loading_plot
 
@@
 mfitpls <- plot_grid(
-  ar_fit + theme(plot.margin = unit(c(0., 0., 1., 0.), "cm")),
+  ar_fit + theme(plot.margin = unit(c(0., 0., 0.2, 0.), "cm")),
   loading_plot,
@@
 # all together
-
 comb <- plot_grid(
   mfitpls, sitelumaps,
-  nrow = 1, scale = 0.9, align = "h"
+  nrow = 1, scale = 0.9, align = "h",
+  rel_widths = c(1, .5)
+)
+
+# MAPS BELLOW PLOT
+# first the maps in a column of three
+sitelumaps <- plot_grid(
+  site_plots[[1]] + theme(
+    legend.position = "none",
+    plot.margin = unit(c(0.1, 0.2, -1.2, 0.2), "cm")
+  ),
+  site_plots[[2]] + theme(
+    legend.position = "none",
+    plot.margin = unit(c(0.1, 0.2, -1.2, 0.2), "cm")
+  ),
+  site_plots[[3]] + theme(
+    legend.position = "none",
+    plot.margin = unit(c(0.1, 0.2, -1.2, 0.2), "cm")
+  ),
+  nrow = 1,
+  align = "h",
+  labels = c(
+    names(site_plots)[1], names(site_plots)[2], names(site_plots)[3]
+  )
+)
+
+mfitpls <- plot_grid(
+  ar_fit + theme(plot.margin = unit(c(0.2, 0.2, 1.2, 0.2), "cm")),
+  sitelumaps,
+  ncol = 1,
+  align = "v",
+  rel_heights = c(1, .5),
+  labels = c(
+    "A", ""
+  )
+)
+
+mfitpls_top <- plot_grid(
+  sitelumaps + theme(plot.margin = unit(c(0.2, 0.2, 1.2, 0.2), "cm")),
+  ar_fit + theme(plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")),
+  ncol = 1,
+  align = "v",
+  rel_heights = c(.5, 1),
+  labels = c(
+    "", "A"
+  )
+)
+
+# all together
+comb <- plot_grid(
+  mfitpls_top, loading_plot,
+  nrow = 1, align = "b",
+  rel_widths = c(1, .4),
+  labels = c("", "B", "")
 )
@@
 comb
+
 ```
@@
 ```{r, agrirural_landuse_maps, message = FALSE, warning = FALSE}
+
+# load land-use conversion table
+luconversions <- read.csv("data/LUCategoryConversion.csv")
+urban_luconversions <- luconversions %>%
+  filter(LU == "urban")
 
@@
 # make the maps for each of the sites selected in the previous step
-site_shape_data <- list()
+site_shape_data_u <- list()
 site_plots <- list()
@@
     )
-  site_shape_data[[u_marked[i, "site"]]] <- df
+  # add major LU categories
+  df_merged <- df %>%
+    mutate(descriptio = toupper(descriptio)) %>%
+    full_join(urban_luconversions, by = c("descriptio" = "code"))
+  site_shape_data_u[[u_marked[i, "site"]]] <- df_merged
   site_plots[[u_marked[i, "site"]]] <- ggplot() +
     geom_sf(
-      data = df,
-      aes(fill = descriptio)
-    ) +
-    theme_void()
+      data = df_merged,
+      aes(fill = category)
+    ) +
+    theme_void() +
+    scale_fill_manual(
+      values = c(
+        "continuous central" = "#EE6677", "dense residential" = "#AA3377",
+        "sparse residential" = "#CCBB44", "amenity grassland" = "#228833",
+        "water" = "#4477AA", "parks allotments cemeteries" = "#BBBBBB",
+        "Railway" = "#BBBBBB", "woodland" = "#BBBBBB"
+      )
+    )
 }
@@
   ) +
+  geom_hline(
+    yintercept = 0, color = "black"
+  ) +
   geom_rect(
@@
     values = c(
-      "continuous..central" = "red3", "dense.residential" = "orange3",
-      "sparse.residential" = "purple3", "amenity.grassland" = "seagreen3",
-      "water" = "cornflowerblue", other = "grey"
+      "continuous.central" = "#EE6677", "dense.residential" = "#AA3377",
+      "sparse.residential" = "#CCBB44", "amenity.grassland" = "#228833",
+      "water" = "#4477AA", other = "#BBBBBB"
     )
@@
   scale_y_continuous(breaks = seq(-.9, .9, by = 0.4), limits = c(-1, 1))
-
-loading_plot
 
@@
 
+# MAPS BELLOW PLOT
+# first the maps in a column of three
+sitelumaps <- plot_grid(
+  site_plots[[1]] + theme(
+    legend.position = "none",
+    plot.margin = unit(c(0.1, 0.2, -1.2, 0.2), "cm")
+  ),
+  site_plots[[2]] + theme(
+    legend.position = "none",
+    plot.margin = unit(c(0.1, 0.2, -1.2, 0.2), "cm")
+  ),
+  site_plots[[3]] + theme(
+    legend.position = "none",
+    plot.margin = unit(c(0.1, 0.2, -1.2, 0.2), "cm")
+  ),
+  nrow = 1,
+  align = "h",
+  labels = c(
+    names(site_plots)[1], names(site_plots)[2], names(site_plots)[3]
+  )
+)
+
+mfitpls <- plot_grid(
+  u_fit + theme(plot.margin = unit(c(0.2, 0.2, 1.2, 0.2), "cm")),
+  sitelumaps,
+  ncol = 1,
+  align = "v",
+  rel_heights = c(1, .4),
+  labels = c(
+    "A", ""
+  )
+)
+
+# all together
+comb <- plot_grid(
+  mfitpls, loading_plot,
+  nrow = 1, align = "h",
+  rel_widths = c(1, .8),
+  labels = c("", "B", "")
+)
+
 comb
+
+```
+
+```{r}
+
+vals <- c()
+for (l in c(site_shape_data_ar, site_shape_data_u)) {
+   vals <- c(vals, unique(as.vector(l$descriptio)))
+}
+
+unique(vals)
+
 ```

### 2021-11-29 analysis/landuse_effect_prop_scouts.Rmd
updated new LU plots to include in the manuscript. Needs typing up properly in the markdown section with vincents section and an abstract including

--- 2021-11-16 analysis/landuse_effect_prop_scouts.Rmd
@@
 theme_set(
-  theme_classic() +
-    theme(
-      text = element_text(size = 42)
-    )
+  theme_classic()
 )
@@
 
+textsize <- 8
+
+# load land-use conversion table
+luconversions <- read.csv("data/LUCategoryConversion.csv")
+agrirural_luconversions <- luconversions %>%
+  filter(LU == "agri-rural")
+
 # make fit map with agrirural landuses
@@
   ) +
-  theme(legend.position = "none") +
+  geom_point(
+    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, colour = "red")
+  ) +
+  geom_text(
+    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, label = site),
+    nudge_x = c(0, 0), nudge_y = c(-0.05, 0.05), size = rel(textsize)
+  ) +
+  theme_classic() +
+  theme(legend.position = "none", text = element_text(size=rel(textsize-2))) +
   xlab("PC 1") +
@@
 # make the maps for each of the sites selected in the previous step
-site_shape_data <- list()
+site_shape_data_ar <- list()
+site_plots <- list()
 for (i in seq_len(3)) {
@@
     )
-  site_shape_data[[ar_marked[i, "site"]]] <- df
+  # add major LU categories
+  df_merged <- df %>%
+    mutate(descriptio = toupper(descriptio)) %>%
+    full_join(agrirural_luconversions, by = c("descriptio" = "code"))
+  site_shape_data_ar[[ar_marked[i, "site"]]] <- df_merged
+  site_plots[[ar_marked[i, "site"]]] <- ggplot() +
+    geom_sf(
+      data = df_merged,
+      aes(fill = category)
+    ) +
+    theme_void() +
+    scale_fill_manual(
+      values = c(
+        "arable" = "#CCBB44", "built-up area" = "#EE6677",
+        "Non-agricultural unmanaged green space" = "#228833",
+        "water" = "#4477AA", "pasture" = "#BBBBBB", "woodland" = "#BBBBBB",
+        "Non-agricultural managed green space" = "#BBBBBB", "fruit" = "#BBBBBB",
+        "OSR" = "#BBBBBB", "other agricultural" = "#BBBBBB"
+      )
+    )
 }
 
-site_shape_data$STU$descriptio
-
-ggplot() +
+# make loadings plot
+argi.rural.loadings <- tibble(
+  "Landuse" = names(agrirural_loadings$loadings),
+  "Correlation" = agrirural_loadings$loadings
+) %>%
+  mutate(
+    significant = ifelse(
+      Correlation > agrirural_loadings$cutoff,
+      Landuse,
+      ifelse(
+        Correlation < -agrirural_loadings$cutoff,
+        Landuse,
+        "Other"
+      )
+    )
+  ) %>%
+  mutate(
+    Landuse = ifelse(
+      Landuse == "Non.agricultural.managed.green.space",
+      "NAMGS",
+      ifelse(
+        Landuse == "Non.agricultural.unmanaged.green.space",
+        "NAUMGS",
+        Landuse
+      )
+    )
+  )
+
+loading_plot <- ggplot(argi.rural.loadings, aes(x = Landuse, y = Correlation)) +
+  geom_point(aes(colour = significant), size = rel(textsize)) +
+  geom_segment(aes(
+    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
+  )) +
+  geom_text(
+    aes(x = Landuse, y = Correlation, label = Landuse),
+    hjust = 0.25,
+    nudge_y = c(0, 0), nudge_x = c(0.5, 0.5), size = rel(textsize)
+  ) +
+  geom_hline(
+    yintercept = 0, color = "black"
+  ) +
+  geom_rect(
+    aes(
+      ymin = -agrirural_loadings$cutoff,
+      ymax = agrirural_loadings$cutoff,
+      xmin = 0, xmax = Inf
+    ),
+    alpha = 0.01
+  ) +
+  coord_flip() +
+  theme_classic() +
+  theme(
+    axis.line.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "None",
+    text = element_text(size=rel(textsize-2))
+  ) +
+  scale_colour_manual(
+    values = c(
+      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
+      "Non.agricultural.unmanaged.green.space" = "#228833",
+      "water" = "#4477AA", other = "#BBBBBB"
+    )
+  ) +
+  ylim(-0.6, 0.6)
+
+# combine plots into facet
+
+# MAPS BELLOW PLOT
+# first the maps in a column of three
+sitelumaps <- plot_grid(
+  site_plots[[1]] + theme(
+    legend.position = "none",
+    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
+  ),
+  site_plots[[2]] + theme(
+    legend.position = "none",
+    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
+  ),
+  site_plots[[3]] + theme(
+    legend.position = "none",
+    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
+  ),
+  nrow = 1,
+  align = "h",
+  labels = c(
+    names(site_plots)[1], names(site_plots)[2], names(site_plots)[3]
+  ),
+  label_size = rel(textsize*5)
+)
+
+mfitpls <- plot_grid(
+  ar_fit + theme(plot.margin = unit(c(0.2, 0.2, 1.2, 0.2), "cm")),
+  sitelumaps,
+  ncol = 1,
+  align = "l",
+  rel_heights = c(1, .5),
+  labels = c(
+    "A", ""
+  ),
+  label_size = rel(textsize*5)
+)
+
+# all together
+comb <- plot_grid(
+  mfitpls, loading_plot,
+  nrow = 1, align = "h",
+  rel_widths = c(1, .4),
+  labels = c("", "B", ""),
+  label_size = rel(textsize*5)
+)
+
+comb
+
+ggsave(
+  plot = comb,
+  filename = "results/figures/lu_prop_scouts_agrirural.png",
+  width = 183,
+  height = 183,
+  units = "mm",
+  dpi = 300
+)
+
+```
+
+### Urban
+
+```{r, agrirural_landuse_maps, message = FALSE, warning = FALSE}
+
+# load land-use conversion table
+luconversions <- read.csv("data/LUCategoryConversion.csv")
+urban_luconversions <- luconversions %>%
+  filter(LU == "urban")
+
+# make fit map with agrirural landuses
+urban_plotdata$site <- urban_lu$site
+urban_plotdata$prop_scouts <- urban_lu$p
+u_sortedpropscouts <- arrange(urban_plotdata, prop_scouts)
+u_marked <- u_sortedpropscouts %>%
+  filter(site %in% c("HOR", "SWP", "CAD"))
+
+# make model fit graph of pc1 against prop scouts and mark the sites with the
+# lowest, middle and highest proportion of scouts
+u_fit <- ggplot(
+  data = u_sortedpropscouts, aes(x = Comp_.1, y = prop_scouts)
+) +
+  geom_point() +
+  geom_point(
+    data = u_marked, aes(x = Comp_.1, y = prop_scouts, colour = "red")
+  ) +
+  geom_text(
+    data = u_marked, aes(x = Comp_.1, y = prop_scouts, label = site),
+    nudge_x = c(0, 0), nudge_y = c(-0.05, 0.05), size = rel(textsize)
+  ) +
+  geom_line(data = urban_fitdata, aes(x = Comp_.1, y = fit)) +
+  geom_ribbon(
+    data = urban_fitdata,
+    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
+    alpha = 0.2,
+    fill = "deepskyblue"
+  ) +
+  theme_classic() +
+  theme(legend.position = "none", text = element_text(size=rel(textsize-2))) +
+  xlab("PC 1") +
+  ylab("Estimated proportion of scouts")
+
+# make the maps for each of the sites selected in the previous step
+site_shape_data_u <- list()
+site_plots <- list()
+for (i in seq_len(3)) {
+  path <- paste0(
+    "shapefiles/urban/",
+    u_marked[i, "site"], "/", u_marked[i, "site"],
+    ".shp"
+  )
+  df <- st_read(path) %>%
+    st_transform(
+      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
+    )
+  # add major LU categories
+  df_merged <- df %>%
+    mutate(descriptio = toupper(descriptio)) %>%
+    full_join(urban_luconversions, by = c("descriptio" = "code"))
+  site_shape_data_u[[u_marked[i, "site"]]] <- df_merged
+  site_plots[[u_marked[i, "site"]]] <- ggplot() +
     geom_sf(
-    data = site_shape_data$STU
-  )
-```
-
+      data = df_merged,
+      aes(fill = category)
+    ) +
+    theme_void() +
+    scale_fill_manual(
+      values = c(
+        "continuous central" = "#EE6677", "dense residential" = "#AA3377",
+        "sparse residential" = "#CCBB44", "amenity grassland" = "#228833",
+        "water" = "#4477AA", "parks allotments cemeteries" = "#BBBBBB",
+        "Railway" = "#BBBBBB", "woodland" = "#BBBBBB"
+      )
+    )
+}
+
+# make loadings plot
+urban.loadings <- tibble(
+  "Landuse" = names(urban_loadings$loadings),
+  "Correlation" = urban_loadings$loadings
+) %>%
+  mutate(
+    significant = ifelse(
+      Correlation > urban_loadings$cutoff,
+      Landuse,
+      ifelse(
+        Correlation < -urban_loadings$cutoff,
+        Landuse,
+        "Other"
+      )
+    )
+  )
+
+loading_plot <- ggplot(urban.loadings, aes(x = Landuse, y = Correlation)) +
+  geom_point(aes(colour = significant), size = rel(textsize)) +
+  geom_segment(aes(
+    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
+  )) +
+  geom_text(
+    aes(x = Landuse, y = Correlation, label = Landuse),
+    hjust = 0.55,
+    nudge_y = c(-0, -0), nudge_x = c(0.5, 0.5), size = rel(textsize)
+  ) +
+  geom_hline(
+    yintercept = 0, color = "black"
+  ) +
+  geom_rect(
+    aes(
+      ymin = -agrirural_loadings$cutoff,
+      ymax = agrirural_loadings$cutoff,
+      xmin = 0, xmax = Inf
+    ),
+    alpha = 0.01
+  ) +
+  coord_flip() +
+  theme_classic() +
+  theme(
+    axis.line.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "None",
+    text = element_text(size=rel(textsize-2))
+  ) +
+  scale_colour_manual(
+    values = c(
+      "continuous.central" = "#EE6677", "dense.residential" = "#AA3377",
+      "sparse.residential" = "#CCBB44", "amenity.grassland" = "#228833",
+      "water" = "#4477AA", other = "#BBBBBB"
+    )
+  ) +
+  scale_y_continuous(breaks = seq(-1, 1, by = 0.4), limits = c(-1, 1))
+
+# combine plots into facet
+
+# MAPS BELLOW PLOT
+# first the maps in a column of three
+sitelumaps <- plot_grid(
+  site_plots[[1]] + theme(
+    legend.position = "none",
+    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
+  ),
+  site_plots[[2]] + theme(
+    legend.position = "none",
+    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
+  ),
+  site_plots[[3]] + theme(
+    legend.position = "none",
+    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
+  ),
+  nrow = 1,
+  align = "h",
+  labels = c(
+    names(site_plots)[1], names(site_plots)[2], names(site_plots)[3]
+  ),
+  label_size = rel(textsize*5)
+)
+
+mfitpls <- plot_grid(
+  u_fit + theme(plot.margin = unit(c(0.2, 0.2, 1.2, 0.2), "cm")),
+  sitelumaps,
+  ncol = 1,
+  align = "l",
+  rel_heights = c(1, .5),
+  labels = c(
+    "A", ""
+  ),
+  label_size = rel(textsize*5)
+)
+
+# all together
+comb <- plot_grid(
+  mfitpls, loading_plot,
+  nrow = 1, align = "h",
+  rel_widths = c(1, .45),
+  labels = c("", "B", ""),
+  label_size = rel(textsize*5)
+)
+
+comb
+
+ggsave(
+  plot = comb,
+  filename = "results/figures/lu_prop_scouts_urban.png",
+  width = 183,
+  height = 183,
+  units = "mm",
+  dpi = 300
+)
+
+```
+
+```{r}
+
+vals <- c()
+for (l in c(site_shape_data_ar, site_shape_data_u)) {
+   vals <- c(vals, unique(as.vector(l$descriptio)))
+}
+
+unique(vals)
+
+```
+
+# Loadings
 ```{r}

### 2021-11-29 analysis/manuscript/paper.Rmd
updated manuscript with new figures and vincent changes. Stil WIP before I can send it to Elli and Vincent

@@
 ---
-title: "Forage site distributions quantify collective foraging in honey bee colonies"
+title: "Forage site distributions show social and private information use in honeybees "
 author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
@@
 
-Collective foraging requires that the emergent behaviour of an animal group takes a form that extends beyond the superposition of numerous individual contributions [@Sumpter2006], such that complexity is achieved at the level of the group or system, rather than the individual.
-Whilst collective foraging has been suggested to occur for many animal species (insert refs list), establishing if, and to what degree, animal groups make collective decisions remains a challenge.
-Here, we present a model of foraging by honey bees that can be used to distinguish collective decisions from individual-driven group behaviour.
-By confronting this model with waggle dance data from twenty hives, collected across two years, we can infer to what extent each colony depends on collective foraging vs individual decision-making.
-Our model both illustrates how a social insect group achieves superorganismal behaviour, and provides a tool to study variation in the occurrence of this behaviour across ecological scenarios.
+Environmental conditions are widely understood to influence the value of social and private information when foraging.
+Exploring this relationship in detail requires quantifying the extent to which foraging decisions are made using social or private information.
+In honeybees, foragers use the waggle dance to organise collective foraging, yet under some conditions it’s use is less effective than prioritising private information.
+Identifying how information is used at a collective level is challenging due to the difficulty in evaluating individual behaviours on the colony scale.
+Here, we outline a model to quantify the extent to which honeybees use the waggle dance by evaluating the foraging distance distributions reported on the dance floor.
+In confronting this model with waggle dance data from twenty hives, we demonstrate that waggle dance use depends on the local environment of a hive.
+Further, by combining our estimates of waggle dance use with local landscape characteristics, our methods demonstrate a means of visualising the relationship between major landscape type and waggle dance use.
+These methods offer a novel approach to evaluate what factors likely led to the evolution of this unique behaviour, as well as helping determine how honeybees might adapt to an increasingly anthropogenic world.
 
-# Main (Word count excluding figures and legend = 2357)
+# Main
 
-Collective decision making provides animals groups with the ability to pool individual experiences and make more informed decisions.
-Despite an enhanced capacity to process information, not every circumstance requires a collective response.
-In many cases, animal groups have been shown to do better when individuals prioritise information they obtain through their own experiences to that provided by conspecifics [@King2007; @Rieucau2011; @Gruter2013].
-Identifying the conditions which favour different information use strategies is therefore important, both for determining what conditions might have lead to the evolution of sociality and for understanding how animal groups might interact in an increasingly anthropomorphic world.
-However, assessing the importance of different information use strategies is not trivial and requires comparing the actions and behaviours of individuals with that of a collective.
-Often, information from different sources gets pooled together and so determining the importance of social Vs individual information use requires tools which can tease apart individual and collective behaviours.
+Decision making by social insects typically requires the use of both social and private information (ref).
+However, to what degree social or private information is needed depends on circumstance (ref), and the mixture of private and social information varies both spatially and temporally.
+Although collective decision making using social information is an adaption to foraging in challenging environments (ref), in some cases, groups have been shown to do better when individuals prioritise information obtained through their own experiences to that provided by conspecifics [@King2007; @Rieucau2011; @Gruter2013].
+Identifying the conditions which favour the use of these two types of information is important to understand how social insects make collective decisions, and determining what conditions might have led to the evolution of sociality as well as how social insects might respond to anthropogenic changes.
 
@@
 In some ecological circumstances, individuals that forgo waiting for dance information in favour of individual search, termed “scouting”, may find food more quickly, and evidence suggests that bees adopt this strategy if the benefits of dance-following drop [@Gruter2008].
-Whilst individuals can be observed under experimental conditions to determine how they use different sources of information, translating this up to a colony level is logistically impossible due to honey bees small size, extensive foraging ranges and shear number of foragers.
+Whilst individuals can be observed under experimental conditions to determine how they use different sources of information, translating this up to a colony level is challenging due to honeybees small size, extensive foraging ranges and shear number of foragers.
 New methods which can quantify waggle dance use at a colony level are therefore needed if we are to assess the conditions where the dance provides the most benefit and derive the conditions user which this unique behaviour evolved.
 
-Whether an individual forager follows a waggle dance from the dance floor or finds a resource for themselves, all foragers report resource information on the dance-floor [@Seeley1995].
-Consequently, the difference in foraging behaviours between individuals following the dance and individuals relying on personal information should manifest itself in the pattern of dances reported on the dance-floor.
-To evaluate this idea, we simulated honey bees foraging in a landscape using both personal and waggle dance information (Fig.
+Whether an individual forager uses social information and follows a waggle dance from the dance floor or acquires private information through scouting, all foragers report resource information on the dance-floor [@Seeley1995].
+Consequently, the difference in foraging behaviours between recruited individuals following dances and scouting individuals relying on private information should manifest itself in the pattern of dances reported on the dance-floor.
+Here, we will demonstrate that in honeybees it is possible quantify the use of social and private information through the observation of honeybees’ waggle dances.
+
+To reveal what foraging patterns might look like under a social or personal information use strategy, we simulated honeybees foraging in a landscape using both personal and waggle dance information (Fig.
 1).
-As which individuals forage under what strategy is known in the simulation, we can compare the distributions of foraging distances reported on the dance floor.
+As it is known in the simulation which individuals forage under what strategy, we can compare the distributions of foraging distances reported on the dance floor.
 Figs 1c and 1d show that the shapes of the resource distance distributions under the two simulated scenarios are very different.
@@
 The distribution of the distances reported by recruits (Fig.
-1 D) is appears to resemble a Rayleigh distribution which is the nearest neighbour distribution of a two-dimensional environment (see methods).
+1 D) appears to be a Rayleigh distribution which is the nearest neighbour distribution of a two-dimensional environment (see methods).
 
@@
 
-Simulating honey bee foraging under a strategy of independent foraging and waggle dance following shows the shape of the foraging distance distributions reported on the dance-floor carries information about the extent to which honey bee colonies collectively make foraging decisions.
-On a real honey bee dance-floor, the distribution of distances is modified through the profitability of a resource reported.
+Scouts and recruits report different foraging distance distributions on the dance-floor.
+This means the distances reported on the dance floor can be used to quantify the relative use of private and social information.
+Foragers that use private information (scouts) report a distribution of distances that reflects their probability of encountering a resource in the environment (see methods), whilst foragers that use social information (recruits) report distances that reflect the distribution of the most profitable resource available (see methods).
+The distance distribution reported on the dance floor therefore carries information about the extent to which private and social information is used.
+
+On the dance-floor, the distribution of distances is modified through the profitability of a resource reported.
 Honey bees achieve this by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994A].
-Combined with the distance distributions identified from our simulations, we form a prediction of the distances reported on the dance-floor as a superposition of independent and waggle dance foraging (Fig 2, see Methods).
-This yields two different models: a collective model demonstrating collective decision making, comprised of the proportion of dances following an individual foraging strategy (scout) along with the proportion of dances following a waggle dance use strategy (recruit), and an individual model where decisions are made entirely individually (scout only) (Fig 2).
-By confronting these models with waggle dance data we can infer to what extent colony foraging depends on individual or collective decision making.
+Combined with the distance distributions identified in our simulations, we can describe the distances reported on the dance-floor as a superposition of private and social information use (Fig 2, see Methods).
+We assume that scouts report distances as they encounter resources on their scouting forays.
+If resources are distributed randomly this is approximated by an exponential distribution.
+Recruits report the most profitable resource among the resources advertised on the dance floor and follow a Rayleigh distribution (see Methods).
+
+This yields two models for foraging: one model based on collective decision making which uses a mixture of private and social information, comprised of the proportion of dances following an individual foraging strategy (scouts), denoted by the parameter $p$, along with the proportion of dances, $1-p$, following a waggle dance use strategy (recruit), and a model based on the sole use of private information (scouts only, $p = 1$) (Fig 2).
+By using maximum likelihood methods and model selection, we inferred if and, by estimating the parameter $p$, to what extent colonies use social information.
 
@@
 To evaluate collective decision making in honey bee colonies foraging on 'natural' landscapes we analysed a dataset of 3378 waggle dance observations from 20 observation hives between April-September 2017, previously described in [@Samuelson2021].
-Hives were situated at different locations in the Southeast of England (see methods, figure 3A) and were visited every two weeks for a period of 24 weeks.
+Hives were situated at different locations in South East England (see methods, figure 3A) and were visited every two weeks for a period of 24 weeks.
 On each visit, two hours of continuous waggle dance data were recorded by training a camcorder onto the dance floor.
 The footage of the dances was decoded manually to identify reported foraging locations [@Couvillon2012; @Schurch2019].
-For each site we fit both the collective and individual models.
-We used model selection to determine which model provided the better explanation of the data and model fit to determine if the model provided a good explanation of the data.
+For each site we fit both the collective and individual foraging models.
+We used model selection to determine which model provided the better explanation of the data, and quantify the relative use of social information through estimating the parameter $p$.
+We calculated goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data.
 
 For 17 out of 20 study hives, a model of collective foraging provided a better explanation of the data than one of individual foraging (Fig.
-3a), however, in 3 sites the individual model provided the best explanation of foraging patterns.
-Across 17 of the sites, the collective model was not statistically different from the empirical foraging distances, whereas the scout model was significantly different to the data in 8 of the sites (Fig.
+3a), in 3 sites the individual model provided the best explanation of foraging patterns.
+Across 17 of the sites, the collective model had a good fit to the empirical foraging distances, whereas the scout model was significantly different to the data in 8 of the sites, and therefore not deemed a plausible model (Fig.
 3.b).
-Sites ZSL (Fig.
-3c) and STU (Fig 3d) are representative examples showing the model fits.
-For ZSL the collective model describes foraging distances extremely well and no significant difference is observed between the model and empirical foraging distances (AIC: 569, Bootstrapped KS statistic: d = 0.06, p = 0.395).
+The sites shown in Figs  3c-d are representative examples showing the model fits.
+For the data in Fig 3c the collective model describes foraging distances extremely well and no significant difference is observed between the model and empirical foraging distances (AIC: 569, Bootstrapped KS statistic: d = 0.06, p = 0.395).
 In contrast the scout model and empirical foraging distances are significantly different, providing a worse explanation of the data (AIC: 622, KS statistic: d = 0.122, p < 0.01).
-For STU the collective model provides a slightly better fit (KS: d = 0.04, p = 0.999) than the scout model (KS: d = 0.05, p = 0.998), however, the scout model is more parsimonious (AIC: 388) than the collective model (AIC: 393).
+For Fig 3d the collective model provides a slightly better fit (KS: d = 0.04, p = 0.999) than the scout model (KS: d = 0.05, p = 0.998), however, the scout model is more parsimonious (AIC: 388) than the collective model (AIC: 393).
 
@@
 
-Our results indicate that, whilst foraging decisions are primarily reached collectively through the use of the waggle dance, in some circumstances, colony foraging is best explained by a model of individual foraging.
-This suggests that in some environments, honey bees actively ignore the waggle dance and instead prioritise the use of personal information.
+Our results show that, whilst foraging decisions are primarily reached using social and private information, in some circumstances, colony foraging is best explained by a model of individual foraging and private information use.
+This suggests that in some environments, honeybees actively ignore the waggle dance and instead prioritise the use of private information.
 Previous empirical studies have shown that individual foragers flexibly alter their use of waggle dance information in response to unrewarding feeders [@Gruter2011], suggesting experience drives information use at the individual level.
-Evaluating how these individual decisions influence the collective, however, has always been a challenge due to fact that it is effectively impossible to track an individuals foraging behaviour over a landscape.
+Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individuals foraging behaviour over a landscape.
 
 As well as providing direct evidence that honey bees forage collectively according to a widely accepted theory [@Seeley1994; @Sherman2002], in showing that foraging patterns in three of our sites are best explained by individual foraging, our results demonstrate honey bee colonies tune waggle dance use in relation to environmental factors.
-These results support previous efforts to quantify collective foraging [@Sherman2002; @Dornhaus2006; @Gruter2013; @Price2019].
-Furthermore, as our model predicts the proportion of dances coming from an individual or social foraging strategy, its fit to waggle dance data provides a tool to estimate a colonies use of the waggle dance.
 
-To evaluate how waggle dance use is being influenced by the environment, we compared the different land-use types for each site to the estimated proportion of dances stemming from an individual foraging strategy (proportion of scout dances).
-Each site used in this study has a unique landscape comprised of different land-use types.
-We first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments.
+Using the $p$ parameter estimated from our model fits, we evaluated how the use of social information varies across different environments.
+Each site used in this study has a unique landscape comprised of different proportions of common land-use types.
+We first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments separately.
 As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis to determine the combination of land-use types which explained the most variation in the proportion of scout dances.
 
+In the Agri-rural environment a single PC explained ~60% of the variation in the proportion of scouts (beta regression: $R^2$ = 0.6, $\phi$ = 4.9, p < 0.05, Fig.
+4.
+A).
+This PC correlates positively with arable land and negatively with non-agricultural unmanaged green space (NAUMGS), built-up areas and water (Fig.
+4.
+B).
+This analysis allows us to visualise what aspects of a colonies local environment most influence waggle dance use.
+As arable land increases whilst NAUMGS and built-up areas decrease, the proportion of individuals prioritising private information increases.
+
 ```{r, fig.cap="Figure 4.
-Estimated proportion of scouts for each site in the Agri-rural (left) and Urban (right) landscapes against the first principle component from a Partial Least Squares analysis of land-use type.
-A beta regression shows the relationship (black line) between PC1 and the proportion of scouts, with 95% CI shown by the blue shaded area."}
-knitr::include_graphics("../results/figures/landuse_plots.png")
-```
-
-In the Agri-rural environment a single PC explained ~60% of the variation in the proportion of scouts (beta regression: $R^2$ = 0.6, $\phi$ = 4.9, p < 0.05, Fig.
-4).
-This PC correlates positively with arable land and negatively with non-agricultural unmanaged green space, built up areas and water.
-Arable land is typically nutritionally poor for honey bees (ref).
-On the contrary, non-agricultural unmanaged green space is typically nutrient rich, as are built up areas in agri-rural environments, which often have gardens with ample flowers available for pollinators (ref).
-Water also correlated negatively with the estimated proportion of scouts, possibly due to river and pond banks providing good floral resources for pollinators (ref).
-Combined, these results suggest when resources are sparse and difficult to find, as with arable land-use, individuals value personal information more than when easier to find, as with the non-agricultural unmanaged green space, built up areas and water.
-
-```{r, agri-rual-loadings}
-set_flextable_defaults(
-  big.mark = " ",
-  font.size = 10, theme_fun = theme_vanilla,
-  padding.bottom = 6,
-  padding.top = 6,
-  padding.left = 6,
-  padding.right = 6
-)
-
-load("../results/pls-loadings.RData")
-argi.rural.loadings <- tibble(
-  "Land-use" = names(agrirural_loadings$loadings),
-  "PC1" = agrirural_loadings$loadings
-)
-
-urban.loadings <- tibble(
-  "Land-use" = names(urban_loadings$loadings),
-  "PC1" = urban_loadings$loadings
-)
-
-argi.rural.loadings %>%
-  mutate(PC1 = round(PC1, 3)) %>%
-  mutate(`Land-use` = gsub("\\.", " ", `Land-use`)) %>%
-  flextable() %>%
-  set_caption(
-    "Land-use type loadings for the first principle component on the Agri-rural sites.
-Bold face indicates loadings that contribute more than would be expected under a null hypothesis of equal contribution"
-  ) %>%
-  set_table_properties(layout = "autofit") %>%
-  bold(
-    i = ~ PC1 > agrirural_loadings$cutoff |
-      PC1 < -agrirural_loadings$cutoff,
-    j = "PC1"
-  )
+A.
+Estimated proportion of scouts for each site in the Agri-rural landscape against the first principle component derived from a Partial Least Squares analysis of land-use type.
+A beta regression shows the relationship (black line) between PC1 and the proportion of scouts, with 95% CI shown by the blue shaded area.
+B.
+Loadings for PC1 showing the correlations between PC1 and each land-use type (note NAUMGS stands for Non-agricultural unmanaged green space).
+All points outside the shaded area significantly contribute to PC1.
+The colours of these land-use types are shown in the local environment of sites HOR, SWP and CAD (maps in bottom left) chosen as representative sites along the model gradient."}
+knitr::include_graphics("../results/figures/lu_prop_scouts_agrirural.png")
 ```
@@
 A similar effect is also observed in the urban environment, where a single PC explained ~61% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.61, $\phi$ = 7.4, p < 0.05, Fig.
-4).
-This PC correlates positively with continuous central land, dense residential land and water, whilst correlating negatively with sparse residential and amenity grassland.
+5.
+A).
+This PC correlates positively with continuous central land, dense residential land and water, whilst correlating negatively with sparse residential and amenity grassland (Fig.
+5.
+B).
 This relationship appears to be largely driven by a single site showing a model of only individual foraging as the best fit.
-As with the agri-rural land-uses, these results indicate when resources are sparse, as could be expected in continuous central and dense residential land, where there is limited space for floral resources, a colony emphasises the use of personal rather than social information.
-However, when resources are more abundant, as could be expected in sparse residential areas and amenity grasslands, a colony appears to increase the use of information contained in the waggle dance.
-Unlike in the agri-rural environment, water correlates positively with a higher use of the waggle dance.
-This likely indicates that the banks of urban water ways are less nutritionally rich than in agri-rural landscapes.
+Visualising how waggle dance use is influenced by changes in these land-use types shows that when the landscape is dominated by sparse residential land, as in site HOR, waggle dance use dominates the colonies foraging strategy.
+However, this reduces as sparse residential land is replaced by continuous central and dense residential lands (Fig.
+5)
 
-```{r, urban.loadings}
-urban.loadings %>%
-  mutate(PC1 = round(PC1, 3)) %>%
-  mutate(`Land-use` = gsub("\\.", " ", `Land-use`)) %>%
-  flextable() %>%
-  set_caption(
-    "Land-use type loadings for the first principle component on the urban sites.
-Bold face indicates loadings that contribute more than would be expected under a null hypothesis of equal contribution"
-  ) %>%
-  set_table_properties(layout = "autofit") %>%
-  bold(
-    i = ~ PC1 > urban_loadings$cutoff |
-      PC1 < -urban_loadings$cutoff,
-    j = "PC1"
-  )
+```{r, fig.cap="Figure 4.
+A.
+Estimated proportion of scouts for each site in the urban landscape against the first principle component derived from a Partial Least Squares analysis of land-use type.
+A beta regression shows the relationship (black line) between PC1 and the proportion of scouts, with 95% CI shown by the blue shaded area.
+B.
+Loadings for PC1 showing the correlations between PC1 and each land-use type.
+All points outside the shaded area significantly contribute to PC1.
+The colours of these land-use types are shown in the local environment of sites MEL, SOM and BEL (maps in bottom left) chosen as representative sites along the model gradient."}
+knitr::include_graphics("../results/figures/lu_prop_scouts_urban.png")
 ```
 
+The fitting of our model to waggle dance observations provides an innovative new tool to quantify the relative importance of social and private information in honeybees.
+Further, by identifying the land-use combinations which most influence variation in the proportion of scouts, our analysis allows us to map the environment along the major axis of honeybee information use and visualise how land-use influences the use of the waggle dance.
+Despite being limited to 10 sites in an urban and agri-rural environment, our results appear to agree with previous findings that the environment influences waggle dance use [@Dornhaus2006; @Price2019].
+
 In both the agri-rural and Urban environments, land-use types which associate with resource scarcity correlate positively with a higher proportion of scouts, whilst those which correlate with resource abundance correlate with a lower proportion of scouts, indicating a greater use of the waggle dance.
-These results appear to follow what is predicted under theoretical models: in landscapes where resources are likely to more difficult to find, a colony tends to have a higher number of scouts than in landscapes where resources should be easier to locate [@Dornhaus2006; @Price2019].
-However, it is important to note that other combinations which didn't show as having a significant association with the first principle component could have just as well fit into this argument, arguably better than others.
-In both the agri-rural and urban environments water seems to correlate in contrasting ways with the proportion of scouts.
-This could be due to differences in the floral richness of urban and agri-rural water ways (reff), however, if land-use type is to be taken as a proxy for floral availability, other variables such as parks, allotments and cemeteries in the urban environment and fruit or Oil Seed Rape in the agri-rural environment might be expected to be included in the first PC over water.
-An alternative explanation could be that water acts as barriers between land-use types and so gets bundled into the association (more thought on this needed I think).
+Arable land is typically nutritionally poor for honey bees (ref).
+On the contrary, non-agricultural unmanaged green space is typically nutrient rich, as are built up areas in agri-rural environments, which often have gardens with ample flowers available for pollinators (ref).
+The same can also be said of the urban land-use types, where continuos central and dense residential areas are also nutritionally poor (ref) compared to sparse residential and amenity grasslands.
+In both landscapes, water shows as significant contributor but in opposite directions, increasing waggle dance use in the agri-rural environment whilst decreasing it in the urban.
+This might indicate that agri-rural water ways provide good floral resources for honeybees, whereas urban water ways do not.
+However, the map visualisations suggests that these land-use types offer little contribution to the overall environment.
+More data is therefore required to evaluate this relationship further.
 
-Here, we demonstrate that the analysis of waggle dance data can make an important contribution to our understanding of social information use and provide a methodology to further evaluate how honey bees use this unique behaviour.
-Recently, technological advances have emerged which enable colony metrics to be collected faster, more accurately and over greater time spans than could be gathered by hand .
-
-Recent advances in computational methods have allowed individuals to be tracked within colonies in ways which would be almost impossible to do by hand, allowing theories of individual behaviour to be evaluated in more depth than could have been done previously [@Wild2021].
-We are not yet in a position to be able to track behaviours outside the colony in the same detail, nevertheless, by piecing together the behavioural response of individuals into a mathematical description of colony foraging, the model outlined here extends our ability to predict individual behaviour beyond the colony walls and in to the environment.
+Overall, we have demonstrated here that the analysis of waggle dance data can make an important contribution to our understanding of social information use and provides a methodology to further evaluate how honey bees use the unique waggle dance.
+Recently, technological advances have emerged which enable colony metrics to be collected faster, more accurately and over greater time spans than could be gathered by hand, allowing individuals to be tracked within colonies and theories of individual behaviour to be evaluated in more depth than could have been done previously [@Wild2021].
+We are not yet in a position to be able to track behaviours outside the colony in the same detail, nevertheless, by piecing together the behavioural response of individuals into a mathematical description of colony foraging, the methodology outlined here extends our ability to predict individual behaviour beyond the colony walls and in to the environment.
 With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”.
-Our methodology provides a means of analysing such large data sets to inform the debate about the importance of social information, as well as providing useful colony metrics of foraging activity.
+Our methodology thus provides a means of analysing such large data sets to inform the debate about the importance of social information, as well as providing useful colony metrics of foraging activity.
 

### 2022-01-27 analysis/fit_models_durations.Rmd
changed analysis to fit models to waggle run duration

--- 2021-11-09 analysis/fit_models.Rmd
@@
 
+REMEBER: Once we know how we are doing the conversion from duration to distance, or what we are doing with it, reset this branch to the remote one and check everthing is as was.
+Then make the edits in the dev branch, merge the relevent bits in properly and then redo the analysis, remaking the figures for the code.
+You MUST reset this to remote to prevent overwiting graphs and what not.
+
 This document contains code to fit the waggle dance model to each of our 20 different sites.
-The code here uses the wagglefit package, as well as some additional code stored in `fit_data.R` to help simplify things here.
+The code here uses the wagglefit package, as well as some additional code stored in `fit_data.R` to help simplify things.
 Each site has its own section, with reused code.
@@
 
+```{r}
+
+calc_dist_actual <- function(duration) {
+  m = 1.38
+  b = 0. #0.17
+  distance = (duration-b) / m
+  # distance = (1/m)*duration-(b/m)
+  return(distance)
+}
+
+data <- read.csv("data/FullHBForagingData.csv")
+
+head(data)
+
+data$newforaging_distance <- calc_dist_actual(data$duration.seconds)
+
+head(data)
+
+par(mfrow = c(2, 2))
+hist(data$duration.seconds)
+hist(data$ForagingDistance_km)
+hist(data$newforaging_distance)
+
+min(data$newforaging_distance)
+min(data$duration.seconds)
+
+alldata <- data %>%
+  select(date, site, duration.seconds) %>%
+  rename(foraging_distance = duration.seconds)
+
+head(alldata)
+
+```
+
 ### BEL
@@
 # subset data for target site
-data <- get_data() %>%
+data <- alldata %>%
   filter(site == target_site)
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 10, -5.5, -2)
+subplot_coords <- c(0.1, 8, -5.5, -1.8)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 4.5, -6.5, -2.5)
+subplot_coords <- c(0.2, 3.5, -6.5, -2.)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 12, -8., -2.5)
+subplot_coords <- c(0.2, 9.5, -8., -2.2)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 5.5, -6.5, -2.5)
+subplot_coords <- c(0.1, 4, -6.5, -2.2)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 2.5, -5, -1.5)
+subplot_coords <- c(0.1, 1.8, -6, -1.2)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 2.7, -8., -2)
+subplot_coords <- c(0.1, 2.1, -8., -2)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 5, -6.5, -1.5)
+subplot_coords <- c(0.1, 4, -6.5, -1.5)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 6, -6.5, -2.5)
+subplot_coords <- c(0.1, 4, -6.5, -2.2)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 4, -9, -2.5)
+subplot_coords <- c(0.1, 3, -9, -2.5)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 5, -6.5, -2.5)
+subplot_coords <- c(0.1, 4.2, -6.5, -2.5)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 3.5, -7, -2.5)
+subplot_coords <- c(0.1, 3., -7, -2.1)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 4.5, -7.2, -2.5)
+subplot_coords <- c(0.1, 4., -7.2, -2.1)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 3., -5.5, -2.5)
+subplot_coords <- c(0.1, 2., -5.5, -2.1)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 3.5, -6.5, -2.5)
+subplot_coords <- c(0.1, 2.5, -6.5, -2.1)
 
@@
 # subset data for target site
-data <- get_data() %>%
+data <- alldata %>%
   filter(site == target_site) %>%
-  filter(foraging_distance < 9) # remove outlier distance
+  filter(foraging_distance < 6) # remove outlier distance
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 4., -7.5, -2.5)
+subplot_coords <- c(0.1, 3., -7.5, -2.1)
 
@@
 br_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 10)
-ar_bnds <- c(1.0e-12, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
 collective_bounds <- rbind(
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 5.5, -6.5, -1.5)
+subplot_coords <- c(0.1, 4, -6.5, -1.5)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 2.5, -6, -2.)
+subplot_coords <- c(0.1, 2., -6, -1.5)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 5., -8, -2.5)
+subplot_coords <- c(0.1, 4., -8, -2.)
 
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.5, 2.4, -8, -2)
+subplot_coords <- c(0.1, 1.8, -8, -2)
 

### 2022-01-27 analysis/fit_models_newdist.Rmd
changed analysis to fit models to waggle run duration

--- 2021-11-09 analysis/fit_models.Rmd
@@
 
+```{r}
+
+calc_dist_actual <- function(duration) {
+  m = 1.38
+  b = 0. #0.17
+  distance = (duration-b) / m
+  # distance = (1/m)*duration-(b/m)
+  return(distance)
+}
+
+data <- read.csv("data/FullHBForagingData.csv")
+
+head(data)
+
+data$newforaging_distance <- calc_dist_actual(data$duration.seconds)
+
+head(data)
+
+par(mfrow = c(2, 2))
+hist(data$duration.seconds)
+hist(data$ForagingDistance_km)
+hist(data$newforaging_distance)
+
+min(data$newforaging_distance)
+min(data$duration.seconds)
+
+alldata <- data %>%
+  select(date, site, newforaging_distance) %>%
+  rename(foraging_distance = newforaging_distance)
+
+```
+
 ### BEL
@@
 # subset data for target site
-data <- get_data() %>%
+data <- alldata %>%
   filter(site == target_site) %>%

### 2022-01-27 analysis/landuse_effect_prop_scouts.Rmd
changed analysis to fit models to waggle run duration

@@
 
+print(ar_sortedpropscouts)
+
 # select sites to draw attention to, those sitting on the trend
 ar_marked <- ar_sortedpropscouts %>%
-  filter(site %in% c("MEL", "SOM", "BEL"))
+  filter(site %in% c("MEL", "ROT", "BEL"))
 
@@
     hjust = 0.25,
-    nudge_y = c(0, 0), nudge_x = c(0.5, 0.5), size = rel(textsize)
+    nudge_y = c(0, -0.05), nudge_x = c(0.5, 0.5), size = rel(textsize)
   ) +
@@
 
+ggsave(
+  plot = comb,
+  filename = "results/figures/lu_prop_scouts_agrirural.svg",
+  width = 183,
+  height = 183,
+  units = "mm",
+  dpi = 300
+)
+
 ```
@@
 
-```{r, agrirural_landuse_maps, message = FALSE, warning = FALSE}
+```{r, urban_landuse_maps, message = FALSE, warning = FALSE}
 
@@
 u_sortedpropscouts <- arrange(urban_plotdata, prop_scouts)
+
+print(u_sortedpropscouts)
+
 u_marked <- u_sortedpropscouts %>%
-  filter(site %in% c("HOR", "SWP", "CAD"))
+  filter(site %in% c("HHS", "MPA", "CAD"))
 
@@
       values = c(
-        "continuous central" = "#EE6677", "dense residential" = "#AA3377",
-        "sparse residential" = "#CCBB44", "amenity grassland" = "#228833",
-        "water" = "#4477AA", "parks allotments cemeteries" = "#BBBBBB",
-        "Railway" = "#BBBBBB", "woodland" = "#BBBBBB"
+        "continuous central" = "#BBBBBB", "dense residential" = "#BBBBBB",
+        "sparse residential" = "#EE6677", "amenity grassland" = "#BBBBBB",
+        "water" = "#BBBBBB", "parks allotments cemeteries" = "#228833",
+        "Railway" = "#CCBB44", "woodland" = "#BBBBBB"
       )
@@
     aes(x = Landuse, y = Correlation, label = Landuse),
-    hjust = 0.55,
-    nudge_y = c(-0, -0), nudge_x = c(0.5, 0.5), size = rel(textsize)
-  ) +
+    hjust = 0.5,
+    vjust = 0.5,
+    nudge_y = c(-0., -0.3), nudge_x = c(0.5, 0.5), size = rel(textsize)
+  ) +
+  theme(text = element_text(size = 10)) +
   geom_hline(
@@
     aes(
-      ymin = -agrirural_loadings$cutoff,
-      ymax = agrirural_loadings$cutoff,
+      ymin = -urban_loadings$cutoff,
+      ymax = urban_loadings$cutoff,
       xmin = 0, xmax = Inf
@@
     values = c(
-      "continuous.central" = "#EE6677", "dense.residential" = "#AA3377",
-      "sparse.residential" = "#CCBB44", "amenity.grassland" = "#228833",
-      "water" = "#4477AA", other = "#BBBBBB"
+      "continuous.central" = "#BBBBBB", "dense.residential" = "#BBBBBB",
+      "sparse.residential" = "#EE6677", "railway" = "#CCBB44",
+      "parks.allotments.cemeteries" = "#228833",
+      "water" = "#BBBBBB", other = "#BBBBBB"
     )

### 2022-01-27 analysis/manuscript/paper.Rmd
changed analysis to fit models to waggle run duration

@@
 ---
-title: "Forage site distributions show social and private information use in honeybees "
+title: "Forage site distributions quantify collective foraging in honey bee colonies"
 author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
@@
 
+Alternative title:
+
+- Waggle run durations quantify collective foraging in honey bee colonies
+- Waggle run duration distributions quantify collective foraging in honey bee colonies
+- Distributions of waggle dance durations quantifies collective foraging in honey bee colonies
+
+Or something like this.
+
 # Abstract
 
-Environmental conditions are widely understood to influence the value of social and private information when foraging.
-Exploring this relationship in detail requires quantifying the extent to which foraging decisions are made using social or private information.
-In honeybees, foragers use the waggle dance to organise collective foraging, yet under some conditions it’s use is less effective than prioritising private information.
-Identifying how information is used at a collective level is challenging due to the difficulty in evaluating individual behaviours on the colony scale.
-Here, we outline a model to quantify the extent to which honeybees use the waggle dance by evaluating the foraging distance distributions reported on the dance floor.
-In confronting this model with waggle dance data from twenty hives, we demonstrate that waggle dance use depends on the local environment of a hive.
-Further, by combining our estimates of waggle dance use with local landscape characteristics, our methods demonstrate a means of visualising the relationship between major landscape type and waggle dance use.
-These methods offer a novel approach to evaluate what factors likely led to the evolution of this unique behaviour, as well as helping determine how honeybees might adapt to an increasingly anthropogenic world.
+Honeybee foraging is a collective behaviour that is directed through the extraordinary waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops.
+Despite having been described as the most astounding example of non-primate communication, recent evidence suggests that the benefits of using dance information vary across environments and are sometimes even negligible.
+Yet, because we cannot easily quantify collective foraging, pinpointing which environments promote collective foraging is challenging.
+Here, we present a mathematical model which allows inferring the proportion of 'scout' and 'recruit' dances on a hive's dance floor and quantify collective foraging.
+By applying this methodology to waggle dance data from twenty hives, we show that dance information may be used less in environments where resources are sparse, potentially because colonies are required to send out more scouts to find food in these circumstances.
+Our methodology provides a tool to quantify waggle dance use in undisturbed hives at a much larger scale than hitherto has been possible and to identify those ecological conditions that may have driven the evolution of this remarkable collective behaviour.
+
+Needs referencing.
 
@@
 
-Decision making by social insects typically requires the use of both social and private information (ref).
-However, to what degree social or private information is needed depends on circumstance (ref), and the mixture of private and social information varies both spatially and temporally.
-Although collective decision making using social information is an adaption to foraging in challenging environments (ref), in some cases, groups have been shown to do better when individuals prioritise information obtained through their own experiences to that provided by conspecifics [@King2007; @Rieucau2011; @Gruter2013].
-Identifying the conditions which favour the use of these two types of information is important to understand how social insects make collective decisions, and determining what conditions might have led to the evolution of sociality as well as how social insects might respond to anthropogenic changes.
+Within honeybee colonies, a series of simple rules that determine when and how much bees perform the celebrated waggle dance (Box 1) mean that choices between feeding sites occur at the level of the group rather than the individual.
+For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the sweeter of two equidistant sources (ref), or the closer of two equally sweet sources (ref), without requiring any individual bee to compare options.
+This extraordinary system is a key example of how social insect colony behaviour can take on a form extending beyond that of the individual units [@Sumpter2006].
 
-Honeybees provide an ideal study system in which to evaluate social information use thanks to the famous "waggle dance", which conveys the location of a profitable resource as a vector of distance (duration of the waggle run) and direction (angle of the dance) [@Frisch1993].
-Resource “quality” is not encoded in the form of the dance itself, but in the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000].
-Although bees that follow dances do not specifically interpret this information on an individual level [@Hasenjager2021], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995].
-This creates a feedback mechanism through which the hive processes the information collected and proportionally selects the most profitable resources [@Seeley1994A].
+Despite such impressive complexity, in some situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2006; @Gruter2013; @Price2019].
+This indicates that bees frequently ignore the spatial information in dances.
 
-Despite the efficiency benefits the waggle dance reportedly provides, its contribution to colony foraging success remains unclear.
-In many situations, colonies that are prevented from communicating via dance behaviour achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2004; @Price2019].
-In some ecological circumstances, individuals that forgo waiting for dance information in favour of individual search, termed “scouting”, may find food more quickly, and evidence suggests that bees adopt this strategy if the benefits of dance-following drop [@Gruter2008].
-Whilst individuals can be observed under experimental conditions to determine how they use different sources of information, translating this up to a colony level is challenging due to honeybees small size, extensive foraging ranges and shear number of foragers.
-New methods which can quantify waggle dance use at a colony level are therefore needed if we are to assess the conditions where the dance provides the most benefit and derive the conditions user which this unique behaviour evolved.
+The prevalent explanation for this apparent redundancy portrays dance-based collective foraging as an evolutionary anachronism that was useful in ancient tropical environments but offers little benefit in modern temperate environments (ref).
+However, fossil evidence has questioned the long-held assumption that the genus Apis evolved in tropical climates (Kotthoff et al 2013).
+Moreover, given that dancing comes at a significant time cost (ref), this raises the issue of why all temperate *Apis mellifera* colonies continue to dance.
+The dual questions of what foraging selection pressures drove the evolution of this uniquely complex collective behaviour, and why it continues to exist, are thus yet to be satisfactorily answered.
+Empirical studies have shown that honeybee colonies use dance information flexibly in response to their environment, suggesting that bees cease to rely on dance information if it provides little benefit [@Price2019].
+Thus, identifying those environments whereby modern honeybee colonies forage collectively rather than through individual search is a key step towards understanding dance evolution (ref).
 
-Whether an individual forager uses social information and follows a waggle dance from the dance floor or acquires private information through scouting, all foragers report resource information on the dance-floor [@Seeley1995].
-Consequently, the difference in foraging behaviours between recruited individuals following dances and scouting individuals relying on private information should manifest itself in the pattern of dances reported on the dance-floor.
-Here, we will demonstrate that in honeybees it is possible quantify the use of social and private information through the observation of honeybees’ waggle dances.
+---
+**Box 1**
 
-To reveal what foraging patterns might look like under a social or personal information use strategy, we simulated honeybees foraging in a landscape using both personal and waggle dance information (Fig.
+[figure of waggle dance, maybe have text in like a box or something? Or just the text]
+
+Honeybees provide an ideal study system in which to evaluate collective foraging thanks to the famous “waggle dance”, which conveys information about location of a resource as a vector of distance (duration of the waggle run) and direction (angle of the dance) (von Frisch 1993).
+Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions.
+Resource “quality” is conveyed through the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000]).
+Although bees that follow dances do not specifically interpret this information on an individual level [@Hasenjager2021], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995], and provides the colony with a mechanism to select the most profitable resources in their environment [@Seeley1994A].
+---
+
+Overall, evidence that colony-level collective foraging offers any benefit over individual search has proved surprisingly hard to come by, mainly because this requires to distinguish scouts from recruits.
+Scouts are foragers that have located a resource through individual search behaviour, whilst recruits are foragers that have found a forage site by using the information provided in the dance of another bee.
+Scouting and recruiting are not fixed behavioural categories, because individual bees can engage in both over the course of their foraging lifetime, and foragers can dance on return from any successful trip irrespective of whether they were recruited to the forage site or found it individually.
+So far, scouts and recruits are distinguished by means of the foragers' individual search history, which requires following individual bees and is time and labour intensive.
+
+Here, we will use a different approach and develop a model that describes the proportion of foraging trips in which bees act as scouts, and as recruits, within a colony, based on the distribution of foraging distances reported by waggle dancing, encoded as the duration of the run phase of the dance [@Frisch1993].
+We will use this model to infer the proportion of scouts and recruits on the dance floor from dance data, and to demonstrate the principle, apply this methodology to dance data from a collection of hives in different environments.
+
+To reveal what the distribution of waggle run durations reported on the dance floor look like, and to demonstrate that dances from recruits and scouts lead to different distributions, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
+Foragers could locate these under two different strategies: either acting as a scout and locating resources themselves, or following a recruit strategy and locating resources by following a random dance from the dance floor (Fig.
 1).
 As it is known in the simulation which individuals forage under what strategy, we can compare the distributions of foraging distances reported on the dance floor.
-Figs 1c and 1d show that the shapes of the resource distance distributions under the two simulated scenarios are very different.
-The distribution of the first simulation is akin to that of an exponential distribution (Fig.
+Figs 1c and 1d show that the shapes of the resource distance distributions for the two types of foraging trips are very different.
+The distance distribution of the scouts is akin to that of an exponential distribution (Fig.
 1 C), which is the nearest neighbour distance distribution for foragers operating in a one-dimensional environment (see methods).
 The distribution of the distances reported by recruits (Fig.
-1 D) appears to be a Rayleigh distribution which is the nearest neighbour distribution of a two-dimensional environment (see methods).
+1 D) is a Rayleigh distribution which is the nearest-neighbour distribution of a two-dimensional environment [@Pyke1978] (Pyke 1978) (see Methods).
 
@@
 
-Scouts and recruits report different foraging distance distributions on the dance-floor.
-This means the distances reported on the dance floor can be used to quantify the relative use of private and social information.
-Foragers that use private information (scouts) report a distribution of distances that reflects their probability of encountering a resource in the environment (see methods), whilst foragers that use social information (recruits) report distances that reflect the distribution of the most profitable resource available (see methods).
-The distance distribution reported on the dance floor therefore carries information about the extent to which private and social information is used.
+Scouts and recruits create different foraging distance distributions on the dance-floor.
+These different distance distributions can be used to quantify the relative importance of collective foraging.
+Foragers that scout report a distribution of distances that reflects their probability of encountering a resource in the environment (see methods), whilst foragers that recruits report distances that reflect the distribution of the most profitable resource available (see methods).
+On the dance-floor, the distribution of distances is modified through the number of runs performed for a resource, which depends on the profitability of a resource, which honeybees achieve by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994A].
+By combining this profitability with the distributions identified in our simulations, we develop a method to describe the distribution of waggle runs reported on the dance-floors of real honeybee colonies as a superposition of scout and recruit distributions (Fig 2, see Methods).
 
-On the dance-floor, the distribution of distances is modified through the profitability of a resource reported.
-Honey bees achieve this by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994A].
-Combined with the distance distributions identified in our simulations, we can describe the distances reported on the dance-floor as a superposition of private and social information use (Fig 2, see Methods).
-We assume that scouts report distances as they encounter resources on their scouting forays.
-If resources are distributed randomly this is approximated by an exponential distribution.
-Recruits report the most profitable resource among the resources advertised on the dance floor and follow a Rayleigh distribution (see Methods).
-
-This yields two models for foraging: one model based on collective decision making which uses a mixture of private and social information, comprised of the proportion of dances following an individual foraging strategy (scouts), denoted by the parameter $p$, along with the proportion of dances, $1-p$, following a waggle dance use strategy (recruit), and a model based on the sole use of private information (scouts only, $p = 1$) (Fig 2).
-By using maximum likelihood methods and model selection, we inferred if and, by estimating the parameter $p$, to what extent colonies use social information.
+This method intrinsically captures honeybee foraging as collective decision making as the foraging sites represented on the dance floor derive from a mixture of individual search and waggle dance information, captured in the proportions p and 1-p, respectively.
+Consequently, by setting p to 1, we can also derive a model based on the sole use of individual search (scouts only) (Fig 2).
+By fitting these two models to waggle run duration distributions from real honeybee colonies we are thus able, using model selection [@BurnhamAnderson?](Burnham and Anderson), to infer if and, by estimating the parameter p, to what extent honeybee colonies use social information when foraging.
 
@@
 
-To evaluate collective decision making in honey bee colonies foraging on 'natural' landscapes we analysed a dataset of 3378 waggle dance observations from 20 observation hives between April-September 2017, previously described in [@Samuelson2021].
+To evaluate the use of waggle dance information and individual search in honeybee colonies foraging in 'natural' landscapes, we analysed an existing dataset of 3378 waggle dance observations from 20 observation hives, recorded between April-September 2017, (previously described in [@Samuelson2021]).
 Hives were situated at different locations in South East England (see methods, figure 3A) and were visited every two weeks for a period of 24 weeks.
-On each visit, two hours of continuous waggle dance data were recorded by training a camcorder onto the dance floor.
-The footage of the dances was decoded manually to identify reported foraging locations [@Couvillon2012; @Schurch2019].
-For each site we fit both the collective and individual foraging models.
-We used model selection to determine which model provided the better explanation of the data, and quantify the relative use of social information through estimating the parameter $p$.
-We calculated goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data.
+On each visit, two hours of continuous waggle dance data was recorded by training a camcorder onto the dance floor.
+The footage of the dances was decoded manually ([@Couvillon2012; @Schurch2019] to extract waggle run durations.
+For each site we fit both the collective and individual models and used model selection to determine which provided the better explanation of the data, and (if the collective model provided a better fit) to quantify the relative use of social information through estimating the parameter p.
+We calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data.
 
 For 17 out of 20 study hives, a model of collective foraging provided a better explanation of the data than one of individual foraging (Fig.
-3a), in 3 sites the individual model provided the best explanation of foraging patterns.
-Across 17 of the sites, the collective model had a good fit to the empirical foraging distances, whereas the scout model was significantly different to the data in 8 of the sites, and therefore not deemed a plausible model (Fig.
-3.b).
-The sites shown in Figs  3c-d are representative examples showing the model fits.
-For the data in Fig 3c the collective model describes foraging distances extremely well and no significant difference is observed between the model and empirical foraging distances (AIC: 569, Bootstrapped KS statistic: d = 0.06, p = 0.395).
-In contrast the scout model and empirical foraging distances are significantly different, providing a worse explanation of the data (AIC: 622, KS statistic: d = 0.122, p < 0.01).
-For Fig 3d the collective model provides a slightly better fit (KS: d = 0.04, p = 0.999) than the scout model (KS: d = 0.05, p = 0.998), however, the scout model is more parsimonious (AIC: 388) than the collective model (AIC: 393).
+3a).
+In each of these cases, the collective model had a good fit (using a Kolmogorov-Smirnov statistic, see methods) to the empirical waggle run durations, and for 9 of these sites, the scout model was significantly different to the data (Fig.
+3b).
+The sites shown in Figs 3c-d are representative examples showing the model fits, showing the overall quality of the model fit to the data.
 
@@
 B.
-Distribution of goodness of fit confidence values for each model fit to waggle dance derived foraging distances from each site.
+Distribution of goodness of fit confidence values for each model fit to waggle run durations from each site.
 P value is derived from a bootstrapped two-sided KS test comparing the fitted model predictions to the empirical data.
@@
 C.
-Model fits to waggle dance derived foraging distances collected from the ZSL hive along with histogram showing distribution.
+Model fits to waggle run durations collected from the ZSL hive along with histogram showing distribution.
 D.
-Model fits to waggle dance derived foraging distances collected from the STU hive along with histogram showing distribution.
+Model fits to waggle run durations collected from the STU hive along with histogram showing distribution.
 Model fits show the compliment cumulative frequency distribution, giving the probability of sampling a value greater than or equal to x."}
@@
 
-Our results show that, whilst foraging decisions are primarily reached using social and private information, in some circumstances, colony foraging is best explained by a model of individual foraging and private information use.
-This suggests that in some environments, honeybees actively ignore the waggle dance and instead prioritise the use of private information.
-Previous empirical studies have shown that individual foragers flexibly alter their use of waggle dance information in response to unrewarding feeders [@Gruter2011], suggesting experience drives information use at the individual level.
+Our results show that, whilst colony-level foraging mostly use a mixture of scouts and recruits, in some circumstances, colony foraging is best explained by a model of individual foraging.
+This suggests that in some environments, honey bees actively ignore the waggle dance and instead prioritise the use of private information.
+Previous empirical studies have shown that individual foragers flexibly alter their use of waggle dance information in response to unrewarding resources [@Gruter2011; @Price2019].
 Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individuals foraging behaviour over a landscape.
+As our results show that these individual behaviours accurately describe colony foraging in three different sites, these findings support the idea that these individual behaviours can dominate colony foraging.
 
-As well as providing direct evidence that honey bees forage collectively according to a widely accepted theory [@Seeley1994; @Sherman2002], in showing that foraging patterns in three of our sites are best explained by individual foraging, our results demonstrate honey bee colonies tune waggle dance use in relation to environmental factors.
+Our finding that foraging in some landscapes is more accurately described by ignoring the dance and focusing on private information supports the hypothesis that the dance is used flexibly in response to environmental factors.
+As the collective model quantifies waggle dance use, the fit of our model provides a tool to evaluate how environmental factors influence waggle dance use, and thus how a colony values social and private information in different environmental circumstances.
+This allows for comparing waggle dance use across environments.
 
-Using the $p$ parameter estimated from our model fits, we evaluated how the use of social information varies across different environments.
-Each site used in this study has a unique landscape comprised of different proportions of common land-use types.
-We first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments separately.
-As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis to determine the combination of land-use types which explained the most variation in the proportion of scout dances.
+We compared the estimates of the proportion of scout and recruits for colonies in different environments.
+Each site used in this study is situated in a landscape comprised of different proportions of land-use types.
+To evaluate how these different land-use types influenced waggle dance use, we first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments separately (see methods).
+As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances.
 
-In the Agri-rural environment a single PC explained ~60% of the variation in the proportion of scouts (beta regression: $R^2$ = 0.6, $\phi$ = 4.9, p < 0.05, Fig.
+In the agri-rural environment one combination of land use types explained ~52% of the variation in the proportion of scouts (beta regression: $R^2 = 0.52$, $\phi$ = 3.7, p < 0.05, Fig.
 4.
 A).
-This PC correlates positively with arable land and negatively with non-agricultural unmanaged green space (NAUMGS), built-up areas and water (Fig.
+This principal component correlates positively with arable land and negatively with non-agricultural unmanaged green space, built-up areas and water (Fig.
 4.
 B).
-This analysis allows us to visualise what aspects of a colonies local environment most influence waggle dance use.
-As arable land increases whilst NAUMGS and built-up areas decrease, the proportion of individuals prioritising private information increases.
+From the land-use maps coloured by each landscape dominating PC1 (Fig 4) we can see that arable land and built-up areas show the most coverage.
+As arable land increases whilst unmanged green space and built-up areas decrease, the proportion of individuals prioritising private information increases.
 
@@
 B.
-Loadings for PC1 showing the correlations between PC1 and each land-use type (note NAUMGS stands for Non-agricultural unmanaged green space).
+Loadings for PC1 showing the correlations between PC1 and each land-use type (NAUMGS stands for Non-agricultural unmanaged green space).
 All points outside the shaded area significantly contribute to PC1.
-The colours of these land-use types are shown in the local environment of sites HOR, SWP and CAD (maps in bottom left) chosen as representative sites along the model gradient."}
+The colours of these land-use types are shown in the local environment of sites MEL, ROT and BEL (maps in bottom left) chosen as representative sites along the model gradient."}
 knitr::include_graphics("../results/figures/lu_prop_scouts_agrirural.png")
@@
 
-A similar effect is also observed in the urban environment, where a single PC explained ~61% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.61, $\phi$ = 7.4, p < 0.05, Fig.
+A similar effect is also observed in the urban environment, where a single PC explained ~59% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.59, $\phi$ = 6.2, p < 0.05, Fig.
 5.
 A).
-This PC correlates positively with continuous central land, dense residential land and water, whilst correlating negatively with sparse residential and amenity grassland (Fig.
+This PC correlates positively with parks, allotments and cemeteries, whilst correlating negatively with sparse residential lands and railways (Fig.
 5.
 B).
-This relationship appears to be largely driven by a single site showing a model of only individual foraging as the best fit.
-Visualising how waggle dance use is influenced by changes in these land-use types shows that when the landscape is dominated by sparse residential land, as in site HOR, waggle dance use dominates the colonies foraging strategy.
-However, this reduces as sparse residential land is replaced by continuous central and dense residential lands (Fig.
-5)
+Visualising how waggle dance use is influenced by changes in these land-use types shows that when the landscape is dominated by sparse residential land, as in site HHS, waggle dance use dominates the colonies foraging strategy.
+However, this reduces as sparse residential land is replaced by parks, allotments, and cemeteries (Fig.
+5).
 
-```{r, fig.cap="Figure 4.
+```{r, fig.cap="Figure 5.
 A.
@@
 All points outside the shaded area significantly contribute to PC1.
-The colours of these land-use types are shown in the local environment of sites MEL, SOM and BEL (maps in bottom left) chosen as representative sites along the model gradient."}
+The colours of these land-use types are shown in the local environment of sites HHS, MPA and CAD (maps in bottom left) chosen as representative sites along the model gradient."}
 knitr::include_graphics("../results/figures/lu_prop_scouts_urban.png")
@@
 
-The fitting of our model to waggle dance observations provides an innovative new tool to quantify the relative importance of social and private information in honeybees.
-Further, by identifying the land-use combinations which most influence variation in the proportion of scouts, our analysis allows us to map the environment along the major axis of honeybee information use and visualise how land-use influences the use of the waggle dance.
-Despite being limited to 10 sites in an urban and agri-rural environment, our results appear to agree with previous findings that the environment influences waggle dance use [@Dornhaus2006; @Price2019].
-
-In both the agri-rural and Urban environments, land-use types which associate with resource scarcity correlate positively with a higher proportion of scouts, whilst those which correlate with resource abundance correlate with a lower proportion of scouts, indicating a greater use of the waggle dance.
+In agri-rural environments, land-use types which typically associate with resource scarcity correlate positively with a higher proportion of scouts, whilst those which correlate with a higher resource abundance correlate with a lower proportion of scouts.
 Arable land is typically nutritionally poor for honey bees (ref).
 On the contrary, non-agricultural unmanaged green space is typically nutrient rich, as are built up areas in agri-rural environments, which often have gardens with ample flowers available for pollinators (ref).
-The same can also be said of the urban land-use types, where continuos central and dense residential areas are also nutritionally poor (ref) compared to sparse residential and amenity grasslands.
-In both landscapes, water shows as significant contributor but in opposite directions, increasing waggle dance use in the agri-rural environment whilst decreasing it in the urban.
-This might indicate that agri-rural water ways provide good floral resources for honeybees, whereas urban water ways do not.
-However, the map visualisations suggests that these land-use types offer little contribution to the overall environment.
-More data is therefore required to evaluate this relationship further.
+These findings indicate that in agri-rural environments, when resources are more abundant and thus easier to find, the collective foraging is used more, presumably to identify the best resources in the environment (ref).
+Conversely, where resources are more difficult to find, a decreased use of the waggle dance suggests the colony focuses more on individual exploration rather than in trying to locate on the best resources.
+A similar result is observed for urban environments, where the dependence appears to be driven by the presence of flower rich gardens in residential areas.
+However, an increase in the amount of parks, allotments and cemeteries decreases the use of the dance.
 
-Overall, we have demonstrated here that the analysis of waggle dance data can make an important contribution to our understanding of social information use and provides a methodology to further evaluate how honey bees use the unique waggle dance.
+The fitting of our model to waggle dance observations provides a novel tool to quantify the relative importance of social and private information in honeybees.
+Further, by identifying the land-use combinations which most influence variation in the proportion of scouts, our analysis allows us to map the environment along the major axis of honeybee information use and visualise how land-use influences the use of the waggle dance.
+Our results within the agri-rural sites are in broad agreement with other studies which evaluated foraging performance with and without the dance [@Dornhaus2006; Price2019].
+These results suggest that, even though collective foraging is not always beneficial, that in environments where high quality resources are present, but relatively scarce, collective foraging is being carried out by colonies, suggesting that under these conditions exchanging social information through the waggle dance confers benefits.
+These benefits of dancing appear to outweigh the costs associated with dancing and could explain why this remarkable behaviour is present in modern-day bees.
+However, the results for our urban sites do not produce a clear biological rational and present an all together more varied picture.
+More waggle dance data from urban colonies is therefore required to further probe how honeybees use the waggle dance in this heavily anthropogenic environment.
+
+The analysis of waggle dance data can make an important contribution to our understanding of social information use and provides a methodology to further evaluate how honey bees use their unique dance language.
 Recently, technological advances have emerged which enable colony metrics to be collected faster, more accurately and over greater time spans than could be gathered by hand, allowing individuals to be tracked within colonies and theories of individual behaviour to be evaluated in more depth than could have been done previously [@Wild2021].
-We are not yet in a position to be able to track behaviours outside the colony in the same detail, nevertheless, by piecing together the behavioural response of individuals into a mathematical description of colony foraging, the methodology outlined here extends our ability to predict individual behaviour beyond the colony walls and in to the environment.
+By piecing together the behavioural response of individuals and combining these with landscape analyses, we have found a mathematical description of colony foraging which extends our ability to predict individual behaviour beyond the colony walls and in to the environment.
 With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”.

### 2022-01-27 R/plotting.R

@@ -70,12 +70,12 @@ make_base_plot <- function(x, logit = TRUE) {
     plt <- inverse_ccdf(x) %>%
       ggplot(aes(x = .data$sd, y = log(.data$prob))) +
       geom_point() +
-      labs(x = "Foraging distance (Km)", y = "Ln cumulative probability")
+      labs(x = "Waggle run duration (seconds)", y = "Ln cumulative probability")
   } else {
     plt <- inverse_ccdf(x) %>%
       ggplot(aes(x = .data$sd, y = .data$prob)) +
       geom_point() +
-      labs(x = "Foraging distance (Km)", y = "Ln cumulative probability")
+      labs(x = "Waggle run duration (seconds)", y = "Ln cumulative probability")
   }
   return(plt)
 }

### 2022-01-27 analysis/manuscript/methods.Rmd
added refs to MS and updated a methods section

---
title: "Forage site distributions quantify collective foraging in honey bee colonies: Methods"
author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
output:
  word_document
always_allow_html: true
bibliography: paper-refs.bib
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(showtext)
showtext_auto()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(size = 18)
    )
)

library(dplyr)
library(kableExtra)
library(flextable)
library(tibble)
```

# Acknowledgments

# Conflict of interest

The authors declare no conflict of interest.

# Code and data availability

All code for analysis is available on GitHub at (insert link to finalised public github repo or archive). All data is available via the Dryad Digital Repository https://doi.org/10.5061/dryad.c2fqz618f [@Samuelson2021].

# Data collection

Methods for how the data used for this study was collected can be found in full in Materials and Methods sections 2.1, 2.2 and 2.3 of [@Samuelson2021].

## Land-use preference analysis

Methods for how land-use types were classified can be found in Materials and Methods sections 2.6 of [@Samuelson2021].

## Waggle dance decoding

Methods for waggle dance decoding are fully described in Materials and Methods section 2.4 in [@Samuelson2021].

# Simulation

A circular environment is first created with radius $r = 2.5$. The number of flowers in the environment is generated as a random Poisson variable with rate equal to 5000 multiplied by the area of the environment. These are placed on polar coordinates with a uniformly selected angle, $\theta$, between 0 and 2$\pi$ and a radial value, $\rho$, between 0 and $r$, determined from the square root of the uniform position values multiplied by $r$. These polar coordinates are converted to Cartesian coordinates using the formulas $x = \rho\cos(\theta)$ and $y = \rho\sin(\theta$). Each location is then assigned to an instance of a flower object along with a random quality of between 0 and 10. This quality is combined with the distance of the flower to the centrally located hive to form a measure of how profitable the flower is (see model, equation?).

One-hundred honeybee objects are created, 20 of which 20 are scouts and the rest recruits. Scouts leave the environment following a random path through the environment generated by sampling a uniform random step length and angle. The number of paths the scout draws when searching is also determined as a uniform random number. Each straight line in the random path is converted to a rectangle with length equal to the path section length and a constant width of ~0.01 to represent an area the scout searches along that path. Of all the flowers contained in the boxes drawn from the scout's path, the one closest to the colony is selected as the flower the scout will report and will communicate its location if the quality of the flower is $ge$ 0. Flower communication is simulated by pooling together all the flower objects found. If no flowers are contained in the scout's path, they will not add any flowers to the scout pool and draw a new path in the next foraging iteration.

Recruits represent honeybee objects which do not perform the searches the scouts do. Instead, they sample from the pool of flower objects reported by scouts. This sampling is done by selecting flowers with a probability which is skewed towards the profitability of the flower, meaning more profitable flowers have a greater chance of being selected by recruits. Recruits will then visit these flowers and in the next iteration will add their flower to the pool of scout dances. Consequently, the pool of dances represents flowers discovered by scouts and flowers exploited by recruits. When a flower is depleted,

the flower is removed from the environment and so any foragers that were foraging on it would select a different flower from the dance floor.

The simulation was run 100 times and every 5 time steps all distances reported by scouts and recruits were recorded and combined. This was done to reflect the way foraging data is collected in real honeybee studies. We fit an exponential and minimum of an exponential distribution to both the distribution of foraging distances reported by the scout and recruit objects. Fitting was done by deriving the maximum likelihood estimate for each model fit on each data source through their analytical solutions: $\hat{\lambda} = \frac{1}/{\bar{x}}$, minimum of the exponential with a minimum foraging distance: $\hat{\lambda} = \frac{1}{\pi \bar{x^{2}}}$. As the exponential assumes distributions start from 0 the data was transformed to start from 0 by subtracting the minimum foraging distance from all foraging distances ($x = x - min(x)$) before fitting.

All simulation code was written in Python version 3.9 and uses the Pandas [@Mckinney2011] and Scipy [@Jones2001] packages.

# Model

We make the initial assumption that there are $n$ different resources distributed within an environment according to a spatial Poisson point process with rate $\lambda_{I}$ for resource $I$. As a scout travels a small distance $dx$ it will cover an area of $c_sdx$, with $c_s$ depending upon the width of the area scanned, the detection rate and degree to which a scout's path covers new ground. The probability density of the distance travelled after which the first resource is discovered $x$ follows an exponential distribution with a rate derived from the intensity of the flowers within the environment: $\lambda_s e^{-\lambda_s x}$, where $\lambda_s = c_s \sum_{I = 1}^{n} \lambda_I$.

The number of repeats of a dance on the dance floor is dictated by profitability and is defined as the net energy efficiency per amount of energy spent: $\frac{Gain-Cost}{Cost}$. The gain is determined by the sucrose concentration of nectar from a flower, $q$, whilst the cost is determined by the distance, $x$. The energy gained, $G$, is therefore proportional to the quality $q$, whilst the cost increases linearly with distance. We can therefore refactor energy efficiency as $\frac{q}{1+\alpha x}-1$, where $\alpha$ represents the energy spend per distance relative to a fixed cost of travel. As bees should not be expected to dance for a negative profitability, we can derive a function to translate quality and distance into a measure of the number of dances on a dance floor: $\sigma(q,x) = [\frac{q}{1+\alpha x}-1]_+$, where the notation of $[x]_+ = max(x, 0)$. This function has the property that for a fixed $x$ there is a minimum quality $q = 1+\alpha x$ below which bees don't dance. The equivalent quality of two resources at distance $x1$ and $x2$ is $\frac{1+\alpha x2}{1+\alpha x1} = \frac{q2}{q1}$. The critical distance $\epsilon_{ij} = [\frac{1}{\alpha} \frac{qj - qi}{qi}+\frac{qj}{qi} x]_+$.

For a resource of type $I$ located at a distance $x$, the expected number of dances would be $c_s \lambda_I \sigma(qi, x) e^{-\lambda_s x}$. It follows that the expected total number of scout waggle dances on the dance floor is given by: $M_s = \sum_{I=1}^{n} \int_{0}^{\inf} \frac{c_s \lambda_I}{\lambda_s} \sigma(qi, x) \lambda_s e^{-\lambda_s x} dx$. The distribution of scout dances is: $\frac{\lambda_s e^{-\lambda_s x} \sum_{I=1}^{n} \frac{c_s \lambda_I}{\lambda_s} \sigma(qi, x)}{M_s}$.

Using the same logic we derive what the distribution of recruits will look like. If two locations differ in resource quality and distance then a hive can prefer a further location if it is sufficiently better and a lesser quality resource if it is sufficiently nearer. For a site with quality $q_I$ and at distance $x$ from the hive, for a second location with quality $q_j$ this threshold distance $\epsilon_{ij}$ satisfies $\sigma(q_i, x) = \sigma(qj, \epsilon_{ij}$. Consequently, the critical distance is an implicit function of distance and it follows from implicit differentiation that $\frac{d\epsilon_{ij}}{dq_j} = - \frac{d\sigma{q_j,x}/dq{j}}{d\sigma{q,x}/dx} > 0$ and $\frac{d\epsilon_{ij}}{dq_I} = \frac{d\sigma{q_i,x}/dq{i}}{d\sigma{q,x}/dx} < 0$. If $qi = qj$ then $\epsilon_{ij}(x) = x$ and it follows that if $q_I > q_j$ then $\epsilon_{ij} < x$ and if $q_I < q_j$ then $\epsilon_{ij} > x$.  As recruits get their information from scouts (as well as recruits) how well the scouts cover depends upon the number of scouts and how well their collective paths cover an area and the chance of discovery of each resource. To capture this we let the fraction of resources that can be discovered by a collection of scouts be $c_r$ and assume $0 \leq c_r \leq 1$. The probability that there is a not a food source discovered worth reporting on within a radius of $x$ $e^{-c_r \lambda_r \pi \epsilon_{ij}(x)^2}$, where $\lambda_r = \sum_I={1}^{n} \lambda_i$. This function also described the probability of having the nearest food source that the recruits will come out for discovered at least a distance of $x$ away.

# Statistical analysis

All analysis code is written in R [@RCore2020].

## Model fitting

All models are fit using Maximum likelihood estimation [@Burnham2002] by summation of the log of the model function outlined in the methods section: model. The numerical optimisation routine is written in c++ and uses the Nelder-Mead simplex algorithm [@Nelder1965] implemented in the 'NLopt' library [@Johnson2020] and interfaced to R [@RCore2020] using 'Rcpp' [@Eddelbuettel2011].

Model parsimony is assessed using Akaike information criterion (AIC) [@Aho2014] and the model with the lowest AIC score is deemed to be the most parsimonious.

Goodness of fit is assessed using the Kolmorgorov-Smirnov (KS) test [@Goldstein2004] and implemented in R using the ks.boot function of the package Matching in R [@Sekhon2011].

## Partial Least Squares analysis

A key consideration is that our response variable is a proportion and thus lies on the interval $[0, 1]$. This makes it unsuitable for a standard linear regression. One option would be to use a generalized linear regression with a logit link and quasibinomial family. This method is most suited for when the response variable is an integer proportion, where the response variable is a division of two integers, such as the proportion of males within a population. Our measure is in principle an integer proportion as we measure the proportion of scouts, however, this value is an estimate derived from our model and so is continuous on the interval $[0, 1]$. An alternative method would be to transform the proportions on to a continuos scale using the $arcsin \sqrt{x}$ transformation. The most appropriate method would be to use a beta regression which is designed specifically for proportional response variables.

We used the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010].

# References

### 2022-01-27 analysis/manuscript/paper.Rmd
added refs to MS and updated a methods section

@@
 
-Honeybee foraging is a collective behaviour that is directed through the extraordinary waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops.
-Despite having been described as the most astounding example of non-primate communication, recent evidence suggests that the benefits of using dance information vary across environments and are sometimes even negligible.
+Honeybee foraging is a collective behaviour that is directed through the extraordinary waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops [@Frisch1993; @Seeley1995].
+Despite having been described as the most astounding example of non-primate communication [@Frisch1993], recent evidence suggests that the benefits of using dance information vary across environments and are sometimes even negligible [@Dornhaus2006; @Price2019].
 Yet, because we cannot easily quantify collective foraging, pinpointing which environments promote collective foraging is challenging.
@@
 
-Needs referencing.
-
 # Main
@@
 Within honeybee colonies, a series of simple rules that determine when and how much bees perform the celebrated waggle dance (Box 1) mean that choices between feeding sites occur at the level of the group rather than the individual.
-For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the sweeter of two equidistant sources (ref), or the closer of two equally sweet sources (ref), without requiring any individual bee to compare options.
+For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the sweeter of two equidistant sources [@Seeley1991], or the closer of two equally sweet sources [@Hasenjager2022], without requiring any individual bee to compare options.
 This extraordinary system is a key example of how social insect colony behaviour can take on a form extending beyond that of the individual units [@Sumpter2006].
@@
 
-The prevalent explanation for this apparent redundancy portrays dance-based collective foraging as an evolutionary anachronism that was useful in ancient tropical environments but offers little benefit in modern temperate environments (ref).
-However, fossil evidence has questioned the long-held assumption that the genus Apis evolved in tropical climates (Kotthoff et al 2013).
-Moreover, given that dancing comes at a significant time cost (ref), this raises the issue of why all temperate *Apis mellifera* colonies continue to dance.
+The prevalent explanation for this apparent redundancy portrays dance-based collective foraging as an evolutionary anachronism that was useful in ancient tropical environments but offers little benefit in modern temperate environments [@Dornhaus2004; @Donaldson-Matasci2012; @Price2015].
+However, fossil evidence has questioned the long-held assumption that the genus Apis evolved in tropical climates [@Kotthoff2013].
+Moreover, given that dancing comes at a significant time cost [@Seeley1988], this raises the issue of why all temperate *Apis mellifera* colonies continue to dance.
 The dual questions of what foraging selection pressures drove the evolution of this uniquely complex collective behaviour, and why it continues to exist, are thus yet to be satisfactorily answered.
-Empirical studies have shown that honeybee colonies use dance information flexibly in response to their environment, suggesting that bees cease to rely on dance information if it provides little benefit [@Price2019].
-Thus, identifying those environments whereby modern honeybee colonies forage collectively rather than through individual search is a key step towards understanding dance evolution (ref).
+Empirical studies have shown that honeybee colonies use dance information flexibly in response to their environment [@Sherman2002; @Dornhaus2004; @Dornhaus2006], suggesting that bees cease to rely on dance information if it provides little benefit [@Price2019].
+Thus, identifying those environments whereby modern honeybee colonies forage collectively rather than through individual search is a key step towards understanding dance evolution [@Price2015].
 
@@
 
-Honeybees provide an ideal study system in which to evaluate collective foraging thanks to the famous “waggle dance”, which conveys information about location of a resource as a vector of distance (duration of the waggle run) and direction (angle of the dance) (von Frisch 1993).
+Honeybees provide an ideal study system in which to evaluate collective foraging thanks to the famous “waggle dance”, which conveys information about location of a resource as a vector of distance (duration of the waggle run) and direction (angle of the dance) [@Frisch1993].
 Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions.
@@
 Although bees that follow dances do not specifically interpret this information on an individual level [@Hasenjager2021], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995], and provides the colony with a mechanism to select the most profitable resources in their environment [@Seeley1994A].
+
 ---
@@
 Overall, evidence that colony-level collective foraging offers any benefit over individual search has proved surprisingly hard to come by, mainly because this requires to distinguish scouts from recruits.
-Scouts are foragers that have located a resource through individual search behaviour, whilst recruits are foragers that have found a forage site by using the information provided in the dance of another bee.
-Scouting and recruiting are not fixed behavioural categories, because individual bees can engage in both over the course of their foraging lifetime, and foragers can dance on return from any successful trip irrespective of whether they were recruited to the forage site or found it individually.
-So far, scouts and recruits are distinguished by means of the foragers' individual search history, which requires following individual bees and is time and labour intensive.
+Scouts are foragers that have located a resource through individual search behaviour, whilst recruits are foragers that have found a forage site by using the information provided in the dance of another bee [@Seeley1995].
+Scouting and recruiting are not fixed behavioural categories, because individual bees can engage in both over the course of their foraging lifetime, and foragers can dance on return from any successful trip irrespective of whether they were recruited to the forage site or found it individually [@Beekman2007].
+So far, scouts and recruits are distinguished by means of the foragers' individual search history, which requires following individual bees and is time and labour intensive [@Seeley1995].
 
@@
 To reveal what the distribution of waggle run durations reported on the dance floor look like, and to demonstrate that dances from recruits and scouts lead to different distributions, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
-Foragers could locate these under two different strategies: either acting as a scout and locating resources themselves, or following a recruit strategy and locating resources by following a random dance from the dance floor (Fig.
+Foragers could locate these under two different strategies: either acting as a scout and locating resources themselves, or following a recruit strategy and locating resources by following a random dance from the dance floor [@Seeley1992] (Fig.
 1).
@@
 The distribution of the distances reported by recruits (Fig.
-1 D) is a Rayleigh distribution which is the nearest-neighbour distribution of a two-dimensional environment [@Pyke1978] (Pyke 1978) (see Methods).
+1 D) is a Rayleigh distribution which is the nearest-neighbour distribution of a two-dimensional environment [@Pyke1978] (see Methods).
 
@@
 These different distance distributions can be used to quantify the relative importance of collective foraging.
-Foragers that scout report a distribution of distances that reflects their probability of encountering a resource in the environment (see methods), whilst foragers that recruits report distances that reflect the distribution of the most profitable resource available (see methods).
+Scouting foragers report a distribution of distances that reflects their probability of encountering a resource in the environment (see methods), whilst foragers that recruits report distances that reflect the distribution of the most profitable resource available (see methods).
 On the dance-floor, the distribution of distances is modified through the number of runs performed for a resource, which depends on the profitability of a resource, which honeybees achieve by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994A].
-By combining this profitability with the distributions identified in our simulations, we develop a method to describe the distribution of waggle runs reported on the dance-floors of real honeybee colonies as a superposition of scout and recruit distributions (Fig 2, see Methods).
+By combining this profitability with the distributions identified in our simulations, we develop a novel method to describe the distribution of waggle runs reported on the dance-floors of real honeybee colonies as a superposition of scout and recruit distributions (Fig 2, see Methods).
 
-This method intrinsically captures honeybee foraging as collective decision making as the foraging sites represented on the dance floor derive from a mixture of individual search and waggle dance information, captured in the proportions p and 1-p, respectively.
-Consequently, by setting p to 1, we can also derive a model based on the sole use of individual search (scouts only) (Fig 2).
-By fitting these two models to waggle run duration distributions from real honeybee colonies we are thus able, using model selection [@BurnhamAnderson?](Burnham and Anderson), to infer if and, by estimating the parameter p, to what extent honeybee colonies use social information when foraging.
+This method intrinsically captures honeybee foraging as collective decision making, as the foraging sites represented on the dance floor derive from a mixture of individual search and waggle dance information, captured in the proportions $p$ and $1-p$, respectively.
+Consequently, by setting $p = 1$, we can also derive a model based on the sole use of individual search (scouts only) (Fig 2).
+By fitting these two models to distribution of waggle run durations decoded from real honeybee colonies we are thus able, using model selection [@Burnham2002], to infer if and, by estimating the parameter $p$, to what extent honeybee colonies use waggle dance information when foraging.
 
@@
 For each site we fit both the collective and individual models and used model selection to determine which provided the better explanation of the data, and (if the collective model provided a better fit) to quantify the relative use of social information through estimating the parameter p.
-We calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data.
+We calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004].
 
@@
 
-Our results show that, whilst colony-level foraging mostly use a mixture of scouts and recruits, in some circumstances, colony foraging is best explained by a model of individual foraging.
+Our results show that, whilst colony-level foraging mostly use a mixture of scouts and recruits, in some circumstances, colony foraging is better described by a model of individual foraging.
 This suggests that in some environments, honey bees actively ignore the waggle dance and instead prioritise the use of private information.
@@
 Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individuals foraging behaviour over a landscape.
-As our results show that these individual behaviours accurately describe colony foraging in three different sites, these findings support the idea that these individual behaviours can dominate colony foraging.
+As our results show that individual foraging accurately describes colony foraging in three different sites, these findings support the idea that individual decision making can dominate colony foraging and demonstrates further evidence for flexible waggle dance use by honey bee colonies.
 
 Our finding that foraging in some landscapes is more accurately described by ignoring the dance and focusing on private information supports the hypothesis that the dance is used flexibly in response to environmental factors.
-As the collective model quantifies waggle dance use, the fit of our model provides a tool to evaluate how environmental factors influence waggle dance use, and thus how a colony values social and private information in different environmental circumstances.
-This allows for comparing waggle dance use across environments.
+As the collective model quantifies waggle dance use, the fit of our model provides a new tool to evaluate how environmental factors influence waggle dance use, and thus how a colony values waggle dance information under different environmental circumstances.
 
@@
 To evaluate how these different land-use types influenced waggle dance use, we first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments separately (see methods).
-As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances.
+As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances.
 
@@
 In agri-rural environments, land-use types which typically associate with resource scarcity correlate positively with a higher proportion of scouts, whilst those which correlate with a higher resource abundance correlate with a lower proportion of scouts.
-Arable land is typically nutritionally poor for honey bees (ref).
-On the contrary, non-agricultural unmanaged green space is typically nutrient rich, as are built up areas in agri-rural environments, which often have gardens with ample flowers available for pollinators (ref).
-These findings indicate that in agri-rural environments, when resources are more abundant and thus easier to find, the collective foraging is used more, presumably to identify the best resources in the environment (ref).
-Conversely, where resources are more difficult to find, a decreased use of the waggle dance suggests the colony focuses more on individual exploration rather than in trying to locate on the best resources.
+Arable land is typically nutritionally poor for honey bees [@Samuelson2021].
+On the contrary, non-agricultural unmanaged green space is typically nutrient rich, as are built up areas, which often have gardens with ample flowers available for pollinators [@Baldock2019].
+These findings indicate that in agri-rural environments, when resources are more abundant and thus easier to find, the collective foraging is used more, presumably to identify the best resources in the environment.
+Conversely, where resources are sparse and more difficult to find, a decreased use of the waggle dance suggests the colony focuses more on individual exploration rather than in trying to locate on the best resources.
 A similar result is observed for urban environments, where the dependence appears to be driven by the presence of flower rich gardens in residential areas.
-However, an increase in the amount of parks, allotments and cemeteries decreases the use of the dance.
+However, an increase in the amount of parks, allotments and cemeteries appears to decrease the use of the dance.
+
+( **NOTE:** I feel like we need more of an explanation here.
+For the agri-rural environment we can assemble the results into a coherent story which has some logical basis and suggests more data would investigate this further.
+The results for the urban sites are contradictory.
+Parks, allotments and cemeteries increases the proportion of scouts so following our logic these areas should be barron, but the evidence suggests they are not and along side gardens should be full of flora.
+This indicates there is something else in play here, such as the spatial arrangement of the locations?
+Or that we just don't have enough data to comment on what processes are going on in urban sites.
+However, we do get a 'significant' result in the PLS.
+Maybe we need more fine grained data to evaluate how season influences, it could be that these results are muddled by an interaction with season or temperature or something.
+Also, railway is a random thing to pop up as significant.
+I am not sure how best to describe these results without it looking like blatant data mining...)
 

### 2022-02-01 analysis/landuse_effect_jackknife.Rmd
started jackknife on the pls analysis. Results remain significant, procedure is mostly there but need to double check the logic all adds up and the stats are ok

---
title: Analysing the effect of land-use type on the proportion of scouts - a *jackknife approach*
author: Joseph Palmer
output:
  html_notebook:
    theme: yeti
    toc: true
    toc_float: true
---

---

```{r preamble, include=FALSE}
devtools::load_all()
library(showtext)
showtext_auto()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic()
)

library(cowplot)
library(tidyr)
library(dplyr)
library(ggfortify)
library(betareg)
library(plsRbeta)
library(lmtest)
library(kableExtra)
library(effects)
```

```{r get-data, include=FALSE, message = FALSE, warning = FALSE}
# load results of model fit (code to generate found in fit_models.Rmd)
sitedata <- readRDS("results/site_fit_results.Rda")

# filter for best model
sitedata <- sitedata %>%
  group_by(site) %>%
  slice(which.min(AIC))

# Load land use data taken (taken from Samuelson et al. 2021)
urban_lu <- read.csv(file = "data/urban-landuse.csv")
rural_lu <- read.csv(file = "data/agri-rural-landuse.csv")

# put proportion of scouts into landuse data
urban_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(urban_lu)
urban_lu_long <- urban_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")

rural_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(rural_lu)
rural_lu_long <- rural_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
```

Here, I will run a jackknifed PLS regression to identify how land-use type influences the proportion of scouts estimated from our model. The purpose of the jackknife is to assess how sensitive the results of the PLS are to a given sample and allow for a more robust assessment of any relationship that is minimally skewed by particular sites.

The jackknife will proceed as follows:

A PLS will be be first ran on the full dataset, and then subsets of the data with a single site, $n_i$, removed. Running the PLS on this data yields loadings and the $x$ values for a beta regression to explore the relationship between the primary axis of land-use variation and the proportion of scouts.

Here the land-use categories stay the same, but a single row (site and associated proportion of scouts) is removed. The PLS is then carried out on the remaining data. This results in a loading value for each land-use type and a beta regressor for each site. Once the replicates for $n-1$ data points have been collected (that is, once we have loadings and regressors for the data with each site individually removed) we the plot the distribution of the loading values for each land-use type as a box plot to see how they vary in relation to the included sites. We also then take all the regressor values and calculate the median of these values for each site. These medians are then compared against the proportion of scouts in a beta regression and the results Bonferroni corrected.

---

Alternatively, the beta regression slope for each iteration in the jackknife could be used to estimate an overall mean and standard error that could be used to determine significance.

Maybe a plot of the slope coefficients against $p$ might also show if the removal of a single site greatly influences the results of the analysis.

---

```{r, jackknife_functions, message=FALSE, warning=FALSE}
run_betareg <- function(model) {
  # requires global assignment for predictorEffects to find it
  plsdata <<- data.frame(
    prop.scouts = model$dataY,
    model$tt
  )

  plsbetamodel <- betareg(
    prop.scouts ~ Comp_.1,
    plsdata
  )
  print(summary(plsbetamodel))
  newdat <- data_frame(
    Comp_.1 = seq(
      min(plsdata$Comp_.1),
      max(plsdata$Comp_.1),
      length.out = length(plsdata$Comp_.1)
    )
  )
  newdat$fit <- predict(plsbetamodel, newdat, type = "response")

  # get effects
  fitdata <- as.data.frame(predictorEffects(plsbetamodel))$Comp_.1
  plotdata <- plsdata
  plotdata$p <- plsdata$prop.scouts

  # make plot
  plt <- ggplot(data = plotdata, aes(x = Comp_.1, y = p)) +
    geom_point() +
    geom_line(data = fitdata, aes(x = Comp_.1, y = fit)) +
    geom_ribbon(
      data = fitdata,
      aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
      alpha = 0.2,
      fill = "deepskyblue"
    )
  print(plt)
  readline(prompt = "Continue?")
}

run_pls <- function(data) {
  model_pls <- plsRbeta(
    p_transform ~ .,
    data = data,
    nt = ncol(data) - 1,
    model = "pls-beta",
    verbose = FALSE
  )
  # run_betareg(model_pls)
  plsdata <- data.frame(
    prop.scouts = model_pls$dataY,
    pc1 = model_pls$tt[, 1]
  )
  plsbetamodel <- betareg(
    prop.scouts ~ pc1,
    plsdata
  )
  pls_loadings <- model_pls$pp[, 1]
  # get beta regression data
  plsdata <- data.frame(
    prop.scouts = model_pls$dataY,
    model_pls$tt
  )
  returnlist <- list(
    "loadings" = pls_loadings, "plsdata" = plsdata[, 1:2],
    "coef" = coef(plsbetamodel)[[2]]
  )
  return(returnlist)
}
```

## Urban

```{r, urban_pls, message = FALSE, warning = FALSE}

# set up data for pls
urban.pcadata <- urban_lu[, 3:length(names(urban_lu))]
sites <- urban_lu$site
urban_cuttoff <- sqrt(1 / (ncol(urban.pcadata) - 1))
n <- length(urban_lu$p)
y <- (urban_lu$p * (n - 1) + 0.5) / n
urban.pcadata$p_transform <- y

# create empty lists to store jackknifed results in
jkloadings <- list()
jkplsdata <- list()
jkcoefs <- list()

# jackknife the pls analysis
for (i in c(0, seq_len(nrow(urban.pcadata)))) {
  data <- urban.pcadata
  removed <- 0
  usesites <- "None"
  if (i > 0) {
    data <- data[-i, ]
    removed <- urban.pcadata[i, ncol(urban.pcadata)][[1]]
    usesites <- sites[-i]
  }
  result <- run_pls(data)
  loadings <- result[["loadings"]]
  plsdata <- result[["plsdata"]]
  coef <- result[["coef"]]
  landuse_type <- names(loadings)

  resloadings <- tibble(landuse_type, loadings)
  resloadings$p_removed <- removed
  jkloadings[[paste(i)]] <- resloadings

  resplsdata <- plsdata
  resplsdata$p_removed <- removed
  resplsdata$site <- usesites
  jkplsdata[[paste(i)]] <- resplsdata
  coefres <- tibble(coef, removed, usesites)
  jkcoefs[[paste(i)]] <- coefres
}

# process loadings
urban_jk_loadings <- do.call(bind_rows, jkloadings)

# plot loadings
urban_loading_text <- tibble(landuse_type = unique(landuse_type), result = 0)
ggplot(data = urban_jk_loadings, aes(
  x = landuse_type, y = loadings, fill = landuse_type
)) +
  geom_boxplot() +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = urban_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -urban_cuttoff,
    ymax = urban_cuttoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "continuous.central" = "#BBBBBB", "dense.residential" = "#BBBBBB",
      "sparse.residential" = "#EE6677", "railway" = "#CCBB44",
      "parks.allotments.cemeteries" = "#228833",
      "water" = "#BBBBBB", other = "#BBBBBB"
    )
  )

# process plsdata for model fit on the median or mean response from JK PLS
urban_jk_pls <- do.call(bind_rows, jkplsdata)
n <- 10
urban_jk_pls$p <- (2 * n * urban_jk_pls$prop.scouts - 1) / (2 * (n - 1)) # convert to p

forjoin <- urban_lu %>%
  select(site, p)

urban_median_pls_data <- urban_jk_pls %>%
  group_by(site) %>%
  summarise(pc1 = median(Comp_.1, na.rm = TRUE)) %>%
  right_join(forjoin, by = "site") %>%
  mutate(prop.scouts = (p * (n - 1) + 0.5) / n)

plsbetamodel <- betareg(
  prop.scouts ~ pc1,
  urban_median_pls_data
)

# show pls summary
print(summary(plsbetamodel))

# set up plot of pls average fit
newdat <- data_frame(
  pc1 = seq(
    min(urban_median_pls_data$pc1),
    max(urban_median_pls_data$pc1),
    length.out = length(urban_median_pls_data$pc1)
  )
)
newdat$fit <- predict(plsbetamodel, newdat, type = "response")

# get effects
fitdata <- as.data.frame(predictorEffects(plsbetamodel))$pc1
urban_median_pls_data$p <- (2 * n * urban_median_pls_data$prop.scouts - 1) / (2 * (n - 1))

# make plot of average pls fit
plt <- ggplot(data = urban_median_pls_data, aes(x = pc1, y = p)) +
  geom_point() +
  geom_line(data = fitdata, aes(x = pc1, y = fit)) +
  geom_ribbon(
    data = fitdata,
    aes(x = pc1, y = fit, ymin = lower, ymax = upper),
    alpha = 0.2,
    fill = "deepskyblue"
  )
print(plt)

# Show distribution of all response from the JK PLS
ggplot(data = urban_jk_pls, aes(
  x = Comp_.1, y = as.factor(p)
)) +
  geom_violin(outlier.shape = NA) +
  stat_summary(fun = median, geom = "point", colour = "red")

# show coef calculated with peticular prop scouts removed from analysis
urban_jk_coef <- do.call(bind_rows, jkcoefs)
ggplot(urban_jk_coef[2:nrow(urban_jk_coef), ], aes(x = removed, y = coef)) +
  geom_point() +
  geom_hline(
    yintercept = mean(urban_jk_coef$coef), color = "red"
  ) +
  geom_hline(
    yintercept = urban_jk_coef[1, 1][[1]], color = "blue"
  ) +
  geom_hline(
    yintercept = median(urban_jk_coef$coef), color = "green"
  )
```

## Agri-rural PLS

```{r, rural_pls, message = FALSE, warning = FALSE}

# set up data for pls
rural.pcadata <- rural_lu[, 3:length(names(rural_lu))]
sites <- rural_lu$site
rural_cuttoff <- sqrt(1 / (ncol(rural.pcadata) - 1))
n <- length(rural_lu$p)
y <- (rural_lu$p * (n - 1) + 0.5) / n
rural.pcadata$p_transform <- y

# create empty lists to store jackknifed results in
jkloadings <- list()
jkplsdata <- list()
jkcoefs <- list()

# jackknife the pls analysis
for (i in c(0, seq_len(nrow(rural.pcadata)))) {
  data <- rural.pcadata
  removed <- 0
  usesites <- "None"
  if (i > 0) {
    data <- data[-i, ]
    removed <- rural.pcadata[i, ncol(rural.pcadata)][[1]]
    usesites <- sites[-i]
  }
  result <- run_pls(data)
  loadings <- result[["loadings"]]
  plsdata <- result[["plsdata"]]
  coef <- result[["coef"]]
  landuse_type <- names(loadings)

  resloadings <- tibble(landuse_type, loadings)
  resloadings$p_removed <- removed
  jkloadings[[paste(i)]] <- resloadings

  resplsdata <- plsdata
  resplsdata$p_removed <- removed
  resplsdata$site <- usesites
  jkplsdata[[paste(i)]] <- resplsdata
  coefres <- tibble(coef, removed, usesites)
  jkcoefs[[paste(i)]] <- coefres
}

# process loadings
rural_jk_loadings <- do.call(bind_rows, jkloadings)

# plot loadings
rural_loading_text <- tibble(landuse_type = unique(landuse_type), result = 0)
ggplot(data = rural_jk_loadings, aes(
  x = landuse_type, y = loadings, fill = landuse_type
)) +
  geom_boxplot() +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = rural_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -rural_cuttoff,
    ymax = rural_cuttoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
      "Non.agricultural.unmanaged.green.space" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  )

# process plsdata for model fit on the median or mean response from JK PLS
rural_jk_pls <- do.call(bind_rows, jkplsdata)
n <- 10
rural_jk_pls$p <- (2 * n * rural_jk_pls$prop.scouts - 1) / (2 * (n - 1)) # convert to p

forjoin <- rural_lu %>%
  select(site, p)

rural_median_pls_data <- rural_jk_pls %>%
  group_by(site) %>%
  summarise(pc1 = median(Comp_.1, na.rm = TRUE)) %>%
  right_join(forjoin, by = "site") %>%
  mutate(prop.scouts = (p * (n - 1) + 0.5) / n)

plsbetamodel <- betareg(
  prop.scouts ~ pc1,
  rural_median_pls_data
)

# show pls summary
print(summary(plsbetamodel))

# set up plot of pls average fit
newdat <- data_frame(
  pc1 = seq(
    min(rural_median_pls_data$pc1),
    max(rural_median_pls_data$pc1),
    length.out = length(rural_median_pls_data$pc1)
  )
)
newdat$fit <- predict(plsbetamodel, newdat, type = "response")

# get effects
fitdata <- as.data.frame(predictorEffects(plsbetamodel))$pc1
rural_median_pls_data$p <- (2 * n * rural_median_pls_data$prop.scouts - 1) / (2 * (n - 1))

# make plot of average pls fit
plt <- ggplot(data = rural_median_pls_data, aes(x = pc1, y = p)) +
  geom_point() +
  geom_line(data = fitdata, aes(x = pc1, y = fit)) +
  geom_ribbon(
    data = fitdata,
    aes(x = pc1, y = fit, ymin = lower, ymax = upper),
    alpha = 0.2,
    fill = "deepskyblue"
  )
print(plt)

# Show distribution of all response from the JK PLS
ggplot(data = rural_jk_pls, aes(
  x = Comp_.1, y = as.factor(p)
)) +
  geom_violin(outlier.shape = NA) +
  stat_summary(fun = median, geom = "point", colour = "red")

# show coef calculated with peticular prop scouts removed from analysis
rural_jk_coef <- do.call(bind_rows, jkcoefs)
ggplot(rural_jk_coef[2:nrow(rural_jk_coef), ], aes(x = removed, y = coef)) +
  geom_point() +
  geom_hline(
    yintercept = mean(rural_jk_coef$coef), color = "red"
  ) +
  geom_hline(
    yintercept = rural_jk_coef[1, 1][[1]], color = "blue"
  ) +
  geom_hline(
    yintercept = median(rural_jk_coef$coef), color = "green"
  )
```

### 2022-02-02 analysis/landuse_effect_jackknife.Rmd
finished jackknife analysis. awaiting on meeting with vincent before proceding

@@
 
-```{r get-data, include=FALSE, message = FALSE, warning = FALSE}
+```{r get-data, include=FALSE, message = FALSE, error = FALSE, warning = FALSE}
 # load results of model fit (code to generate found in fit_models.Rmd)
@@
 
-Here, I will run a jackknifed PLS regression to identify how land-use type influences the proportion of scouts estimated from our model.
-The purpose of the jackknife is to assess how sensitive the results of the PLS are to a given sample and allow for a more robust assessment of any relationship that is minimally skewed by particular sites.
+# TLDR:
+
+A jackknife was carried out on the PLS beta regression whereby each site was excluded from the analysis and the fits to the others established.
+This was done to assess if a single site biased the model and should be treated as an outlier.
+
+In the urban dataset, site CAD, with a proportion of scouts value of 1 bears the hallmarks of an outlier, the removal of which influences the results of the study more than the removal of any other point.
+No outliers were identified in the agri-rural dataset.
+
+In addition to showing fits with and without outliers for the urban data, for both environments the jackknifed analysis shows how the loadings vary with each site removed and how the model fits look on the average of the PLS beta values from across the jackknifed analyses.
+These results suggest the agri-rural data is fairly robust to removing individual sites, however, the urban data is less robust, exemplifying the bias introduced by site CAD.
+The urban results are overall more volatile than the agri-rural.
+
+By doing this analysis I find more coherent results for the urban sites and demonstrate that more data is needed to draw more firm conclusions.
+The agri-rural data, although only having ten sites, shows a consistent pattern.
+
+In addition, I also plot the loadings to show the amount of area a particular landuse type corresponds to in the total land of all the sites in that environment as a percentage.
+This allows us to see how much land each land-use type covers.
+A linear regression between the loadings of PC1 and the amount of land identified shows no significant relationship between the two variables, indicating the amount of land is not influencing the PLS results.
+
+# Jackknife analysis to investigate PLS sensitivity to outliers
+
+Looking over the results of the PLS it is apparent that in the urban PLS the beta regression analysis appears skewed by a single site, CAD, which has a proportion of scouts much higher than the others.
+It is important to investigate how outliers are influencing our results, particularly in the urban dataset but also with the agri-rural sites.
+To do this I present an analysis to see if there are any outliers and what effect these have on the results of the PLS and beta regression.
+
+Although site CAD appears to be an outlier, it is prudent to formally establish outliers in both datasets to ensure no other outliers are missed and ensure statistical rigour.
+To do this I will perform a jackknifed PLS regression to identify how land-use type influences the proportion of scouts estimated from our model.
+The purpose of the jackknife is to assess how sensitive the results of the PLS are to a given sample and allow for a more robust assessment of any relationship that is skewed by particular sites.
 
@@
 This results in a loading value for each land-use type and a beta regressor for each site.
-Once the replicates for $n-1$ data points have been collected (that is, once we have loadings and regressors for the data with each site individually removed) we the plot the distribution of the loading values for each land-use type as a box plot to see how they vary in relation to the included sites.
+Once the replicates for $n-1$ data points have been collected (that is, once we have loadings and regressors for the data with each site individually removed) I will examine the distribution of slope values identified from the beta regression.
+I will also the plot the distribution of the loading values for each land-use type as a box plot to see how they vary in relation to the included sites.
 We also then take all the regressor values and calculate the median of these values for each site.
-These medians are then compared against the proportion of scouts in a beta regression and the results Bonferroni corrected.
+These medians are then compared against the proportion of scouts in a beta regression.
 
----
+```{r, jackknife_functions, message=FALSE, warning=FALSE, echo = FALSE}
 
-Alternatively, the beta regression slope for each iteration in the jackknife could be used to estimate an overall mean and standard error that could be used to determine significance.
-
-Maybe a plot of the slope coefficients against $p$ might also show if the removal of a single site greatly influences the results of the analysis.
-
----
-
-```{r, jackknife_functions, message=FALSE, warning=FALSE}
-run_betareg <- function(model) {
-  # requires global assignment for predictorEffects to find it
-  plsdata <<- data.frame(
-    prop.scouts = model$dataY,
-    model$tt
-  )
-
-  plsbetamodel <- betareg(
-    prop.scouts ~ Comp_.1,
-    plsdata
-  )
-  print(summary(plsbetamodel))
-  newdat <- data_frame(
-    Comp_.1 = seq(
-      min(plsdata$Comp_.1),
-      max(plsdata$Comp_.1),
-      length.out = length(plsdata$Comp_.1)
+ggplotRegression <- function(fit) {
+  plt <- ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) +
+    geom_point() +
+    stat_smooth(method = "lm", col = "red") +
+    labs(
+      title = paste(
+        "Adj R2 = ", signif(summary(fit)$adj.r.squared, 5),
+        "Intercept =", signif(fit$coef[[1]], 5),
+        " Slope =", signif(fit$coef[[2]], 5),
+        " P =", signif(summary(fit)$coef[2, 4], 5)
       )
     )
-  newdat$fit <- predict(plsbetamodel, newdat, type = "response")
-
-  # get effects
-  fitdata <- as.data.frame(predictorEffects(plsbetamodel))$Comp_.1
-  plotdata <- plsdata
-  plotdata$p <- plsdata$prop.scouts
-
-  # make plot
-  plt <- ggplot(data = plotdata, aes(x = Comp_.1, y = p)) +
-    geom_point() +
-    geom_line(data = fitdata, aes(x = Comp_.1, y = fit)) +
-    geom_ribbon(
-      data = fitdata,
-      aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
-      alpha = 0.2,
-      fill = "deepskyblue"
-    )
-  print(plt)
-  readline(prompt = "Continue?")
+  return(plt)
 }
@@
   )
-  # run_betareg(model_pls)
   plsdata <- data.frame(
@@
 
-## Urban
+## Urban jackknife
 
-```{r, urban_pls, message = FALSE, warning = FALSE}
+Starting with the urban dataset, the data is first jackknifed and the metrics reported and discussed in turn.
+
+```{r, urban_pls, message = FALSE, error = FALSE, warning = FALSE, include = FALSE}
 
@@
   usesites <- "None"
+  removed_site <- "None"
   if (i > 0) {
@@
     usesites <- sites[-i]
+    removed_site <- sites[i]
   }
@@
   resloadings$p_removed <- removed
+  resloadings$site_removed <- removed_site
   jkloadings[[paste(i)]] <- resloadings
@@
 
-# process loadings
+# process results of jackknife for further analysis
 urban_jk_loadings <- do.call(bind_rows, jkloadings)
+urban_jk_pls <- do.call(bind_rows, jkplsdata)
+urban_jk_coef <- do.call(bind_rows, jkcoefs)
+```
+
+### Identifying outlier sites from the slopes of the jackknifed beta regression
+
+After running the beta regression of the jackknifed data we can formally assess outliers by examining the distribution of slopes identified with each site removed.
+
+```{r, urban_jk_slope_dist, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
+# show coef calculated with peticular prop scouts removed from analysis
+urban_jk_coef <- do.call(bind_rows, jkcoefs)
+hist(urban_jk_coef$coef, col = "pink")
+```
+
+With only 10 points to investigate the quantiles are not particuarly informative, however, from looking at the graph we can see that the data is left skewed and that there is an influencial point in the left hand bar, as including this point casues the slope to be above 0.6.
+The majority of points sit around 0.8 to 1.1, however the other bars standing alone (0.6 and 1.2) are in range of the body of weight.
+Excluding just the 0.4 value is enough to center the data, suggesting these points lie as extremes rather than outliers.
+With such a small dataset, it is therefore not beyond the realm of reason that the gaps between them would fill in.
+There is fundimentally a judgment call to be made here, and for the reasons I have outlined I think 0.4 should be determined as the only outlier.
+
+Now lets see what this slope references to.
+
+```{r, urban_jk_slope_dist_value, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
+urban_jk_coef %>% filter(coef < 0.4)
+```
+
+The site missing from this dataset is CAD.
+This is what I would have expected, but it is good to establish this formally.
+The next task is to now look at how the loadings and betaregression fit looks with and without the inclusion of this site.
+
+```{r, urban_jk_outlier_removed, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
+
+# full data fit
+model_pls <- plsRbeta(
+  p_transform ~ .,
+  data = urban.pcadata,
+  nt = ncol(urban.pcadata) - 1,
+  model = "pls-beta",
+  verbose = FALSE
+)
+plsdata <- data.frame(
+  prop.scouts = model_pls$dataY,
+  pc1 = model_pls$tt[, 1]
+)
+plsbetamodel <- betareg(
+  prop.scouts ~ pc1,
+  plsdata
+)
+
+print("Full data beta regression")
+print(summary(plsbetamodel))
+
+# set up plot of pls fit
+newdat <- data_frame(
+  pc1 = seq(
+    min(plsdata$pc1),
+    max(plsdata$pc1),
+    length.out = length(plsdata$pc1)
+  )
+)
+newdat$fit <- predict(plsbetamodel, newdat, type = "response")
+
+# get effects
+fitdata <- as.data.frame(predictorEffects(plsbetamodel))$pc1
+plsdata$p <- (2 * n * plsdata$prop.scouts - 1) / (2 * (n - 1))
+
+# make plot of pls fit
+plt_full <- ggplot(data = plsdata, aes(x = pc1, y = p)) +
+  geom_point() +
+  geom_line(data = fitdata, aes(x = pc1, y = fit)) +
+  geom_ribbon(
+    data = fitdata,
+    aes(x = pc1, y = fit, ymin = lower, ymax = upper),
+    alpha = 0.2,
+    fill = "deepskyblue"
+  )
+
+# plot loadings
+full_loadings <- as_tibble(model_pls$pp[, 1])
+full_loadings$landuse_type <- names(model_pls$pp[, 1])
+
+urban_loading_text <- tibble(landuse_type = unique(landuse_type), result = 0)
+full_loading_plt <- ggplot(data = full_loadings, aes(
+  x = landuse_type, y = value, fill = landuse_type
+)) +
+  geom_point(aes(colour = landuse_type)) +
+  geom_segment(aes(
+    x = landuse_type, xend = landuse_type,
+    y = 0, yend = value, colour = landuse_type
+  )) +
+  geom_hline(
+    yintercept = 0, color = "black"
+  ) +
+  geom_text(
+    data = urban_loading_text,
+    aes(x = landuse_type, y = result, label = landuse_type),
+    hjust = 0.5,
+    vjust = 0.5,
+    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
+  ) +
+  annotate(
+    "rect",
+    ymin = -urban_cuttoff,
+    ymax = urban_cuttoff,
+    xmin = 0, xmax = Inf,
+    alpha = 0.2
+  ) +
+  coord_flip() +
+  theme_classic() +
+  theme(
+    axis.line.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "None",
+    text = element_text()
+  ) +
+  scale_colour_manual(
+    values = c(
+      "continuous.central" = "#BBBBBB", "dense.residential" = "#BBBBBB",
+      "sparse.residential" = "#EE6677", "railway" = "#CCBB44",
+      "parks.allotments.cemeteries" = "#228833",
+      "water" = "#BBBBBB", other = "#BBBBBB"
+    )
+  )
+
+# outlier removed fit
+removed_data <- urban.pcadata %>% filter(p_transform < 0.95)
+
+model_pls <- plsRbeta(
+  p_transform ~ .,
+  data = removed_data,
+  nt = ncol(removed_data) - 1,
+  model = "pls-beta",
+  verbose = FALSE
+)
+plsdata <- data.frame(
+  prop.scouts = model_pls$dataY,
+  pc1 = model_pls$tt[, 1]
+)
+plsbetamodel <- betareg(
+  prop.scouts ~ pc1,
+  plsdata
+)
+
+print("Outlier removed beta regression")
+print(summary(plsbetamodel))
+
+# set up plot of pls fit
+newdat <- data_frame(
+  pc1 = seq(
+    min(plsdata$pc1),
+    max(plsdata$pc1),
+    length.out = length(plsdata$pc1)
+  )
+)
+newdat$fit <- predict(plsbetamodel, newdat, type = "response")
+
+# get effects
+fitdata <- as.data.frame(predictorEffects(plsbetamodel))$pc1
+plsdata$p <- (2 * n * plsdata$prop.scouts - 1) / (2 * (n - 1))
+
+# make plot of pls fit
+plt_outlier_removed <- ggplot(data = plsdata, aes(x = pc1, y = p)) +
+  geom_point() +
+  geom_line(data = fitdata, aes(x = pc1, y = fit)) +
+  geom_ribbon(
+    data = fitdata,
+    aes(x = pc1, y = fit, ymin = lower, ymax = upper),
+    alpha = 0.2,
+    fill = "deepskyblue"
+  ) +
+  scale_y_continuous(limits = c(0, 1))
+
+# plot loadings
+outlier_rmv_loadings <- as_tibble(model_pls$pp[, 1])
+outlier_rmv_loadings$landuse_type <- names(model_pls$pp[, 1])
+
+outlier_rmv_loading_plt <- ggplot(data = outlier_rmv_loadings, aes(
+  x = landuse_type, y = value, fill = landuse_type
+)) +
+  geom_point(aes(colour = landuse_type)) +
+  geom_segment(aes(
+    x = landuse_type, xend = landuse_type,
+    y = 0, yend = value, colour = landuse_type
+  )) +
+  geom_hline(
+    yintercept = 0, color = "black"
+  ) +
+  geom_text(
+    data = urban_loading_text,
+    aes(x = landuse_type, y = result, label = landuse_type),
+    hjust = 0.5,
+    vjust = 0.5,
+    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
+  ) +
+  annotate(
+    "rect",
+    ymin = -urban_cuttoff,
+    ymax = urban_cuttoff,
+    xmin = 0, xmax = Inf,
+    alpha = 0.2
+  ) +
+  coord_flip() +
+  theme_classic() +
+  theme(
+    axis.line.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "None",
+    text = element_text()
+  ) +
+  scale_colour_manual(
+    values = c(
+      "continuous.central" = "#BBBBBB", "dense.residential" = "#BBBBBB",
+      "sparse.residential" = "#EE6677", "railway" = "#CCBB44",
+      "parks.allotments.cemeteries" = "#228833",
+      "water" = "#4477AA", other = "#BBBBBB"
+    )
+  )
+
+# show all plots
+plot_grid(
+  plt_full, plt_outlier_removed,
+  full_loading_plt, outlier_rmv_loading_plt,
+  ncol = 2, labels = "AUTO"
+)
+```
+
+plots A and C show the PLS beta regression of PC1 to the proportion of scouts and the loadings for PC1 respectively for the full data.
+Plots B and D show the same thing but to the data with site CAD removed.
+The first thing to note is the change in loadings when CAD is removed, with water showing a significant negative correlation, whilst sparse residential flips to become a singificant positive correlation.
+The fit beta regression fit with CAD removed remains singificant, but is less significant than with the outlier included: p < 0.05 vs p < 0.01 with CAD indluced.
+Whats more, the amount of variation explained, Psudo-$R^2$ is much lower at ~0.41.
+Overall, this suggests that landuse type explains some, but not a majority of variation in the proportion of scouts identified in our model, indicating other factors are likely contributing, or we simply dont have enough data.
+
+Overall, this result shows the urban dataset is vulnerable to the site CAD, which seems to be acting as an outlier.
+
+### Examining how the loadings and fit changes through the jackknifed analysis
+
+The jacknife computes a PLS beta regression for the sample with each site removed.
+It is thus possible to observe how the loadings change with each site removed and also to look at the regression fit of the average of the regressors outlined from the beta regressions run over the data.
+
+```{r, urban_jk_averaged_fits, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
 
@@
 urban_loading_text <- tibble(landuse_type = unique(landuse_type), result = 0)
-ggplot(data = urban_jk_loadings, aes(
+urban_average_loadings_plt <- ggplot(data = urban_jk_loadings, aes(
   x = landuse_type, y = loadings, fill = landuse_type
@@
 # make plot of average pls fit
-plt <- ggplot(data = urban_median_pls_data, aes(x = pc1, y = p)) +
+urban_average_fit_plt <- ggplot(data = urban_median_pls_data, aes(x = pc1, y = p)) +
   geom_point() +
@@
   )
-print(plt)
 
-# Show distribution of all response from the JK PLS
-ggplot(data = urban_jk_pls, aes(
-  x = Comp_.1, y = as.factor(p)
-)) +
-  geom_violin(outlier.shape = NA) +
-  stat_summary(fun = median, geom = "point", colour = "red")
-
-# show coef calculated with peticular prop scouts removed from analysis
-urban_jk_coef <- do.call(bind_rows, jkcoefs)
-ggplot(urban_jk_coef[2:nrow(urban_jk_coef), ], aes(x = removed, y = coef)) +
-  geom_point() +
-  geom_hline(
-    yintercept = mean(urban_jk_coef$coef), color = "red"
-  ) +
-  geom_hline(
-    yintercept = urban_jk_coef[1, 1][[1]], color = "blue"
-  ) +
-  geom_hline(
-    yintercept = median(urban_jk_coef$coef), color = "green"
+# show all plots
+plot_grid(
+  urban_average_fit_plt, urban_average_loadings_plt,
+  ncol = 1, labels = "AUTO"
 )
@@
 
+Plot A shows the fit of the beta regression to the median PC1 values (I.e.
+the median of the pc1 values identified at each itteration in the jackknife).
+Plot B shows the distribution of the loadings identified for each itteration in the jackknife.
+What is clear is that, on average, parks allotments and cemeteries remains sigificantly positive whilst sparse residential is singificantly negative and railways boarders on significant negative.
+The points (black dots) show the outliers for all of these and for sparse residential the outlier is in the position identified from conducting the PLS without CAD.
+
+The fit of the model is also better than for the full data, with the psudo $R^2$ value increasing to ~0.63 vs 0.59, suggesting averaging over jackknife itterations helps reduce variability in the beta regression.
+
+In many cases there are multiple outliers identifies in the distributions, such as woodland and water.
+However, drilling down further into these is unlikely to be that informative, as we are dealing with high dimensional spaces with only 9 data points.
+
+## Further analysis of sensitivity to site removal and landuse proportions: digging into the the distributions of PLS loadings
+
+Here I show the loadings for each jackknife itteration, that is each PLS ran with a specific site removed. The following plot shows abreak down of the boxplots shown previously.
+
+```{r, urban_jk_loadings_all, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
+
+ggplot(data = urban_jk_loadings, aes(
+  x = landuse_type, y = loadings, fill = landuse_type
+)) +
+  geom_point(aes(colour = landuse_type)) +
+  geom_segment(aes(
+    x = landuse_type, xend = landuse_type,
+    y = 0, yend = loadings, colour = landuse_type
+  )) +
+  geom_hline(
+    yintercept = 0, color = "black"
+  ) +
+  geom_text(
+    data = urban_loading_text,
+    aes(x = landuse_type, y = result, label = landuse_type),
+    hjust = 0.5,
+    vjust = 0.5,
+    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
+  ) +
+  annotate(
+    "rect",
+    ymin = -urban_cuttoff,
+    ymax = urban_cuttoff,
+    xmin = 0, xmax = Inf,
+    alpha = 0.2
+  ) +
+  coord_flip() +
+  theme_classic() +
+  theme(
+    axis.line.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "None",
+    text = element_text()
+  ) +
+  scale_fill_manual(
+    values = c(
+      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
+      "Non.agricultural.unmanaged.green.space" = "#228833",
+      "water" = "#4477AA", other = "#BBBBBB"
+    )
+  ) +
+  facet_wrap(~site_removed)
+```
+
+There is a fair amount of variation in these plots, more than is shown with the boxplots for sure.
+Many landuse types move around in an out of being significant depending on what site is removed, but parks allotments and cemeteries and sparse residential remain relatively consistent in their position, except for when CAD is removed.
+
+## How does this compare with the proportion of land-coverage for each site?
+
+I now recreate the main plot but using the percentage of total area that covers as the argument to the size parameter. for the plot.
+
+```{r, urban_jk_loadings_landuse_area, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
+
+area_sites_used <- urban_lu %>% filter(site != "CAD")
+
+total_area <- area_sites_used[, 3:ncol(area_sites_used)] %>%
+  colSums()
+total_area <- tibble(landuse_type = names(total_area), area = total_area)
+total_area$area_pct <- (total_area$area / sum(total_area$area)) * 100
+
+urban_lu_loadings_with_area <- urban_jk_loadings %>%
+  left_join(total_area, by = "landuse_type") %>%
+  filter(site_removed == "CAD") # <---- remove outlier site
+
+ggplot(data = urban_lu_loadings_with_area, aes(
+  x = landuse_type, y = loadings, fill = landuse_type
+)) +
+  geom_point(aes(colour = landuse_type, size = area_pct)) +
+  geom_segment(aes(
+    x = landuse_type, xend = landuse_type,
+    y = 0, yend = loadings, colour = landuse_type
+  )) +
+  geom_hline(
+    yintercept = 0, color = "black"
+  ) +
+  geom_text(
+    data = urban_loading_text,
+    aes(x = landuse_type, y = result, label = landuse_type),
+    hjust = 0.5,
+    vjust = 0.5,
+    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
+  ) +
+  annotate(
+    "rect",
+    ymin = -urban_cuttoff,
+    ymax = urban_cuttoff,
+    xmin = 0, xmax = Inf,
+    alpha = 0.2
+  ) +
+  coord_flip() +
+  theme_classic() +
+  theme(
+    axis.line.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "None",
+    text = element_text()
+  ) +
+  scale_fill_manual(
+    values = c(
+      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
+      "Non.agricultural.unmanaged.green.space" = "#228833",
+      "water" = "#4477AA", other = "#BBBBBB"
+    )
+  )
+
+urban_lu_loadings_with_area %>%
+  arrange(-area_pct) %>%
+  print()
+```
+
+As we can see from the graph and the table, sparse residential coveres the most land, around 39%.
+Parks allotments and cemeteries describes around 8% of the land, whilst water explains around 3%.
+Railway, a singificant contributor, accounts just 1% of the total land area.
+
+It is also important to check if there is a correlation between the loadings and the area for each landuse type.
+This was done using a linear model between the loadings for PC1 and the area associated with that landuse type.
+
+```{r, urban_jk_loadings_landuse_area_correlation, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
+
+fit1 <- lm(area ~ loadings, data = urban_lu_loadings_with_area)
+print(summary(fit1))
+
+print(
+  ggplotRegression(fit1)
+)
+```
+
+The results show no significant correlation, indicating the amount of land in each landuse type catagory is not influencing the model results.
+
+### Conclusions
+
+Site CAD seems to be having a large effect on the results of the PLS.
+The jackknife analysis shows this site skews the slopes more than the others.
+Plotting the averages of the PLS provides a means of seeing how the PLS results change in relation to removing sites.
+They seem fairly robust to site removals, but this could also mean site CAD is heavily influencing the relationship.
+
+I am not sure if plotting the averages adds anything and it could be making some questionable assumptions if we really dig into this.
+The fact that the PLS slope changes greatly with CAD removed, whereas the others dont as much, suggests that this site should be excluded from the PLS analysis and the result without it used.
+
 ## Agri-rural PLS
 
-```{r, rural_pls, message = FALSE, warning = FALSE}
+Although the agri-rural data doesnt show any ovbious outliers it is prudent to check it with a jackknife to be sure.
+The first thing to do is look at a histogram of the slopes to see if any one site greatly influences it.
+
+```{r, rural_pls, message = FALSE, error = FALSE, warning = FALSE, include = FALSE}
 
@@
 
-# process loadings
+# process results of jackknife for further analysis
 rural_jk_loadings <- do.call(bind_rows, jkloadings)
+rural_jk_pls <- do.call(bind_rows, jkplsdata)
+rural_jk_coef <- do.call(bind_rows, jkcoefs)
+```
+
+```{r, rural_jk_slope_dist, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
+# show coef calculated with peticular prop scouts removed from analysis
+rural_jk_coef <- do.call(bind_rows, jkcoefs)
+jkci <- quantile(rural_jk_coef$coef, probs = c(0.025, .975))
+hist(rural_jk_coef$coef, col = "pink")
+```
+
+The distribution is pretty well centered.
+There is a gap between 0.85 and 0.8, however this is likely due to having a small sample size rather than outliers.
+Given these results it does not seem nesesery to remove any sites from the agri-rural PLS annalysis.
+
+### Examining how the loadings and fit changes through the jackknifed analysis
+
+Although we did not see any outliers from the jackknife, it would still be usefull to evaluate the distribution of loadings and median model fit from the jackknifed analysis, as we did for the urban dataset.
+This will show us if the results change much accross the jackknifes.
+
+```{r, rural_jk_averaged_fits, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
 
@@
 rural_loading_text <- tibble(landuse_type = unique(landuse_type), result = 0)
-ggplot(data = rural_jk_loadings, aes(
+rural_average_loadings_plt <- ggplot(data = rural_jk_loadings, aes(
   x = landuse_type, y = loadings, fill = landuse_type
@@
 n <- 10
-rural_jk_pls$p <- (2 * n * rural_jk_pls$prop.scouts - 1) / (2 * (n - 1)) # convert to p
+# convert back to p
+rural_jk_pls$p <- (2 * n * rural_jk_pls$prop.scouts - 1) / (2 * (n - 1))
 
@@
 # make plot of average pls fit
-plt <- ggplot(data = rural_median_pls_data, aes(x = pc1, y = p)) +
+rural_average_fit_plt <- ggplot(data = rural_median_pls_data, aes(x = pc1, y = p)) +
   geom_point() +
@@
   )
-print(plt)
 
-# Show distribution of all response from the JK PLS
-ggplot(data = rural_jk_pls, aes(
-  x = Comp_.1, y = as.factor(p)
-)) +
-  geom_violin(outlier.shape = NA) +
-  stat_summary(fun = median, geom = "point", colour = "red")
-
-# show coef calculated with peticular prop scouts removed from analysis
-rural_jk_coef <- do.call(bind_rows, jkcoefs)
-ggplot(rural_jk_coef[2:nrow(rural_jk_coef), ], aes(x = removed, y = coef)) +
-  geom_point() +
-  geom_hline(
-    yintercept = mean(rural_jk_coef$coef), color = "red"
-  ) +
-  geom_hline(
-    yintercept = rural_jk_coef[1, 1][[1]], color = "blue"
-  ) +
-  geom_hline(
-    yintercept = median(rural_jk_coef$coef), color = "green"
+# show all plots
+plot_grid(
+  rural_average_fit_plt, rural_average_loadings_plt,
+  ncol = 1, labels = "AUTO"
 )
 ```
+
+Compared to the same plots for the urban site, the distributions are much tighter, and although there are still individual outliers, it seems that the jackknife shows our results don't change much when sites are excluded.
+
+## Further analysis of sensitivity to site removal and landuse proportions: digging into the the distributions of PLS loadings
+
+Here I show the loadings for each jackknife itteration, that is each PLS ran with a specific site removed. The following plot shows abreak down of the boxplots shown previously.
+
+```{r, rural_jk_loadings_all, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
+
+ggplot(data = rural_jk_loadings, aes(
+  x = landuse_type, y = loadings, fill = landuse_type
+)) +
+  geom_point(aes(colour = landuse_type)) +
+  geom_segment(aes(
+    x = landuse_type, xend = landuse_type,
+    y = 0, yend = loadings, colour = landuse_type
+  )) +
+  geom_hline(
+    yintercept = 0, color = "black"
+  ) +
+  geom_text(
+    data = rural_loading_text,
+    aes(x = landuse_type, y = result, label = landuse_type),
+    hjust = 0.5,
+    vjust = 0.5,
+    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
+  ) +
+  annotate(
+    "rect",
+    ymin = -rural_cuttoff,
+    ymax = rural_cuttoff,
+    xmin = 0, xmax = Inf,
+    alpha = 0.2
+  ) +
+  coord_flip() +
+  theme_classic() +
+  theme(
+    axis.line.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "None",
+    text = element_text()
+  ) +
+  scale_fill_manual(
+    values = c(
+      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
+      "Non.agricultural.unmanaged.green.space" = "#228833",
+      "water" = "#4477AA", other = "#BBBBBB"
+    )
+  ) +
+  facet_wrap(~site_removed)
+```
+
+There is a fair amount of variation in these plots, more than is shown with the boxplots for sure.
+Many landuse types move around in an out of being significant depending on what site is removed, but many, such as arable and built up areas, remain consistent in their position.
+
+## How does this compare with the proportion of land-coverage for each site?
+
+I now recreate the main plot but using the percentage of total area that covers as the argument to the size parameter. for the plot.
+
+```{r, rural_jk_loadings_landuse_area, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
+
+total_area <- rural_lu[, 3:ncol(rural_lu)] %>%
+  colSums()
+total_area <- tibble(landuse_type = names(total_area), area = total_area)
+total_area$area_pct <- (total_area$area / sum(total_area$area)) * 100
+
+rural_lu_loadings_with_area <- rural_jk_loadings %>%
+  left_join(total_area, by = "landuse_type") %>%
+  filter(site_removed == "None")
+
+ggplot(data = rural_lu_loadings_with_area, aes(
+  x = landuse_type, y = loadings, fill = landuse_type
+)) +
+  geom_point(aes(colour = landuse_type, size = area_pct)) +
+  geom_segment(aes(
+    x = landuse_type, xend = landuse_type,
+    y = 0, yend = loadings, colour = landuse_type
+  )) +
+  geom_hline(
+    yintercept = 0, color = "black"
+  ) +
+  geom_text(
+    data = rural_loading_text,
+    aes(x = landuse_type, y = result, label = landuse_type),
+    hjust = 0.5,
+    vjust = 0.5,
+    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
+  ) +
+  annotate(
+    "rect",
+    ymin = -rural_cuttoff,
+    ymax = rural_cuttoff,
+    xmin = 0, xmax = Inf,
+    alpha = 0.2
+  ) +
+  coord_flip() +
+  theme_classic() +
+  theme(
+    axis.line.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "None",
+    text = element_text()
+  ) +
+  scale_colour_manual(
+    values = c(
+      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
+      "Non.agricultural.unmanaged.green.space" = "#228833",
+      "water" = "#4477AA", other = "#BBBBBB"
+    )
+  )
+
+rural_lu_loadings_with_area %>%
+  arrange(-area_pct) %>%
+  print()
+```
+
+As we can see from the graph and the table, arable cover the largest percentage of land included in thie study and has a singificant positive association with PC1.
+Built up area, which has a singificant negative association, makes up around 17% of the land, whilst Non-agricultural unmanaged green space accounts for ~3% of the land.
+Water accounts for less than 1% of the total land coverage and so is association is likely to be an artefact of the small sample size used in this analysis.
+
+It is also important to check if there is a correlation between the loadings and the area for each landuse type.
+This was done using a linear model between the loadings for PC1 and the area associated with that landuse type.
+
+```{r, rural_jk_loadings_landuse_area_correlation, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
+
+fit1 <- lm(area ~ loadings, data = rural_lu_loadings_with_area)
+print(summary(fit1))
+
+print(
+  ggplotRegression(fit1)
+)
+```
+
+The results show no significant correlation, indicating the amount of land in each landuse type catagory is not influencing the model results.

### 2022-02-08 analysis/fit_models_durations.Rmd
reran with export to SVG
updated main figures with loading boxplots

@@
 ---
-
-REMEBER: Once we know how we are doing the conversion from duration to distance, or what we are doing with it, reset this branch to the remote one and check everthing is as was.
-Then make the edits in the dev branch, merge the relevent bits in properly and then redo the analysis, remaking the figures for the code.
-You MUST reset this to remote to prevent overwiting graphs and what not.
 
@@
 
-In this document I have fit the collective and individual model to each site.
-So al the sites have reasonable likelihood values, however in some cases the fits are simply wrong (i.e.
-they show the probability of sampling a value greater than on equal to $x$ as increasing in the tail).
-I need to trouble shoot these with Vincent.
-I also need to get some advice on how to plot the model fits to the histogram of data.
-
-Would be a good idea to knit this as a single document / run in R studio on the laptop.
-
 ```{r}
-
-calc_dist_actual <- function(duration) {
-  m = 1.38
-  b = 0. #0.17
-  distance = (duration-b) / m
-  # distance = (1/m)*duration-(b/m)
-  return(distance)
-}
-
+# set waggle dance duration in seconds as foraging distance for analysis
 data <- read.csv("data/FullHBForagingData.csv")
-
-head(data)
-
-data$newforaging_distance <- calc_dist_actual(data$duration.seconds)
-
-head(data)
-
-par(mfrow = c(2, 2))
-hist(data$duration.seconds)
-hist(data$ForagingDistance_km)
-hist(data$newforaging_distance)
-
-min(data$newforaging_distance)
-min(data$duration.seconds)
 
@@
   rename(foraging_distance = duration.seconds)
-
-head(alldata)
-
 ```
@@
 
+ggsave(
+  plot = sites_model_plot,
+  filename = "results/figures/sites_model_plot.svg",
+  width = 183,
+  height = 190,
+  units = "mm",
+  dpi = 300
+)
+
 model_sites_fits <- plot_grid(

### 2022-02-08 analysis/landuse_effect_jackknife.Rmd
updated main figures with loading boxplots

@@
 urban_jk_loadings <- do.call(bind_rows, jkloadings)
+urban_jk_loadings$cutoff <- sqrt(
+  1 / (length(unique(urban_jk_loadings$landuse_type)) - 1)
+)
 urban_jk_pls <- do.call(bind_rows, jkplsdata)
@@
 rural_jk_loadings <- do.call(bind_rows, jkloadings)
+rural_jk_loadings$cutoff <- sqrt(
+  1 / (length(unique(rural_jk_loadings$landuse_type)) - 1)
+)
+save(urban_jk_loadings, rural_jk_loadings, file = "results/pls-loadings.RData")
 rural_jk_pls <- do.call(bind_rows, jkplsdata)

### 2022-02-08 analysis/landuse_effect_prop_scouts.Rmd
reran with export to SVG
updated main figures with loading boxplots

@@
 sqrt(1 / (ncol(urban.pcadata) - 1))
-
-# when multiple PCs have the best fit, summing them gives the overall effect of
-# the PCs, just as we do in the model formula.
-rowSums(model$pp[, 1:3])
 ```
@@
 )
-save(agrirural_loadings, urban_loadings, file = "results/pls-loadings.RData")
 
@@
 
-**Stats notes: things to add or improve in the stats**
-
-* One thing it would be nice to do, but which I have not figured out how to do, is to see how much od the data the PLS PCs explain.
-With the PCA we obtain the variance a PC explains in the data: i.e.
-the variance in land-use types within our data ($R^2_x$).
-A regression after then tells us how much of the variance in the proportion of scouts that particular PC explains ($R^2_y$).
-The PLS optimise a PC to describe as much of the variation in the proportion of scouts as possible ($R^2_y$) but it would be also nice to know how much of the variation in land-use type ($R^2_x$) that PC explains.
-
-* It would be nice to add confidence intervals to the model fit plots. Unsure how to do this with a beta regression and not all that sure it is that important?
-
-# Figures, results write up and main discussion points for manuscript
+# Figures, results write up and main discussion points
 
@@
   theme_classic() +
-  theme(legend.position = "none", text = element_text(size=rel(textsize-2))) +
+  theme(legend.position = "none", text = element_text(size = rel(textsize - 2))) +
   xlab("PC 1") +
@@
 # make loadings plot
-argi.rural.loadings <- tibble(
-  "Landuse" = names(agrirural_loadings$loadings),
-  "Correlation" = agrirural_loadings$loadings
+argi.rural.loadings <- rural_jk_loadings %>%
+  rename(
+    Landuse = landuse_type,
+    Correlation = loadings
   ) %>%
   mutate(
+    Landuse = tools::toTitleCase(
+      gsub("\\.", " ", rural_jk_loadings$Landuse)
+    ),
     significant = ifelse(
-      Correlation > agrirural_loadings$cutoff,
+      Correlation > rural_jk_loadings$cutoff,
       Landuse,
       ifelse(
-        Correlation < -agrirural_loadings$cutoff,
+        Correlation < -rural_jk_loadings$cutoff,
         Landuse,
@@
 
-loading_plot <- ggplot(argi.rural.loadings, aes(x = Landuse, y = Correlation)) +
-  geom_point(aes(colour = significant), size = rel(textsize)) +
-  geom_segment(aes(
-    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
+rural_loading_text <- tibble(
+  landuse_type = unique(argi.rural.loadings$Loadings), result = 0
+)
+loading_plot <- ggplot(data = argi.rural.loadings, aes(
+  x = Landuse, y = Correlation, fill = Landuse
 )) +
-  geom_text(
-    aes(x = Landuse, y = Correlation, label = Landuse),
-    hjust = 0.25,
-    nudge_y = c(0, -0.05), nudge_x = c(0.5, 0.5), size = rel(textsize)
-  ) +
+  geom_boxplot() +
   geom_hline(
@@
   ) +
-  geom_rect(
-    aes(
-      ymin = -agrirural_loadings$cutoff,
-      ymax = agrirural_loadings$cutoff,
-      xmin = 0, xmax = Inf
-    ),
-    alpha = 0.01
+  geom_text(
+    data = rural_loading_text,
+    aes(x = landuse_type, y = result, label = landuse_type),
+    hjust = 0.5,
+    vjust = 0.5,
+    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
+  ) +
+  annotate(
+    "rect",
+    ymin = -rural_cuttoff,
+    ymax = rural_cuttoff,
+    xmin = 0, xmax = Inf,
+    alpha = 0.2
   ) +
@@
     legend.position = "None",
-    text = element_text(size=rel(textsize-2))
+    text = element_text()
   ) +
@@
     )
-  ) +
-  ylim(-0.6, 0.6)
+  )
 
@@
   ),
-  label_size = rel(textsize*5)
+  label_size = rel(textsize * 5)
 )
@@
   labels = c("", "B", ""),
-  label_size = rel(textsize*5)
+  label_size = rel(textsize * 5)
 )
@@
 # make loadings plot
-urban.loadings <- tibble(
-  "Landuse" = names(urban_loadings$loadings),
-  "Correlation" = urban_loadings$loadings
+urban.loadings <- urban_jk_loadings %>%
+  rename(
+    Landuse = landuse_type,
+    Correlation = loadings
   ) %>%
   mutate(
+    Landuse = tools::toTitleCase(
+      gsub("\\.", " ", urban_jk_loadings$Landuse)
+    ),
     significant = ifelse(
-      Correlation > urban_loadings$cutoff,
+      Correlation > urban_jk_loadings$cutoff,
       Landuse,
       ifelse(
-        Correlation < -urban_loadings$cutoff,
+        Correlation < -urban_jk_loadings$cutoff,
         Landuse,
@@
 
-loading_plot <- ggplot(urban.loadings, aes(x = Landuse, y = Correlation)) +
-  geom_point(aes(colour = significant), size = rel(textsize)) +
-  geom_segment(aes(
-    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
+urban_loading_text <- tibble(
+  landuse_type = unique(urban.loadings$Landuse), result = 0
+)
+loading_plot <- ggplot(data = urban.loadings, aes(
+  x = landuse_type, y = loadings, fill = landuse_type
 )) +
+  geom_boxplot() +
+  geom_hline(
+    yintercept = 0, color = "black"
+  ) +
   geom_text(
-    aes(x = Landuse, y = Correlation, label = Landuse),
+    data = urban_loading_text,
+    aes(x = landuse_type, y = result, label = landuse_type),
     hjust = 0.5,
     vjust = 0.5,
-    nudge_y = c(-0., -0.3), nudge_x = c(0.5, 0.5), size = rel(textsize)
-  ) +
-  theme(text = element_text(size = 10)) +
-  geom_hline(
-    yintercept = 0, color = "black"
-  ) +
-  geom_rect(
-    aes(
-      ymin = -urban_loadings$cutoff,
-      ymax = urban_loadings$cutoff,
-      xmin = 0, xmax = Inf
-    ),
-    alpha = 0.01
+    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
+  ) +
+  annotate(
+    "rect",
+    ymin = -urban_cuttoff,
+    ymax = urban_cuttoff,
+    xmin = 0, xmax = Inf,
+    alpha = 0.2
   ) +
@@
     legend.position = "None",
-    text = element_text(size=rel(textsize-2))
-  ) +
-  scale_colour_manual(
+    text = element_text()
+  ) +
+  scale_fill_manual(
     values = c(
-      "continuous.central" = "#BBBBBB", "dense.residential" = "#BBBBBB",
-      "sparse.residential" = "#EE6677", "railway" = "#CCBB44",
-      "parks.allotments.cemeteries" = "#228833",
-      "water" = "#BBBBBB", other = "#BBBBBB"
-    )
-  ) +
-  scale_y_continuous(breaks = seq(-1, 1, by = 0.4), limits = c(-1, 1))
+      "Continuous Central" = "#BBBBBB", "Dense Residential" = "#BBBBBB",
+      "Sparse Residential" = "#EE6677", "Railway" = "#CCBB44",
+      "Parks Allotments Cemeteries" = "#228833",
+      "Water" = "#BBBBBB", Other = "#BBBBBB"
+    )
+  )
 
@@
 
+ggsave(
+  plot = comb,
+  filename = "results/figures/lu_prop_scouts_urban.svg",
+  width = 183,
+  height = 183,
+  units = "mm",
+  dpi = 300
+)
 ```
@@
 unique(vals)
-
 ```

### 2022-02-08 analysis/landuse_effect_prop_scouts.Rmd
fixed errors with figs and rendered them

@@
 
+# load loadings data from the jk analysis
+loading_path <- "results/pls-loadings.RData"
+if (file.exists(loading_path)) {
+  load(loading_path)
+} else {
+   stop(
+    paste(
+      "File", loading_path, "not found.",
+      "Please run the notebook `landuse_effect_jackknife.Rmd` to create it."
+    )
+  )
+}
+
 plotdata <- rbind(urban_plotdata, agrirural_plotdata)
@@
     Landuse = tools::toTitleCase(
-      gsub("\\.", " ", rural_jk_loadings$Landuse)
+      gsub("\\.", " ", rural_jk_loadings$landuse_type)
     ),
@@
       )
-    )
-  ) %>%
-  mutate(
+    ),
     Landuse = ifelse(
-      Landuse == "Non.agricultural.managed.green.space",
+      Landuse == "Non Agricultural Managed Green Space",
       "NAMGS",
       ifelse(
-        Landuse == "Non.agricultural.unmanaged.green.space",
+        Landuse == "Non Agricultural Unmanaged Green Space",
         "NAUMGS",
@@
 
+cutoff <- unique(argi.rural.loadings$cutoff)[[1]]
+
 rural_loading_text <- tibble(
-  landuse_type = unique(argi.rural.loadings$Loadings), result = 0
-)
-loading_plot <- ggplot(data = argi.rural.loadings, aes(
-  x = Landuse, y = Correlation, fill = Landuse
-)) +
-  geom_boxplot() +
+  landuse_type = unique(argi.rural.loadings$Landuse), result = 0
+)
+
+loading_plot <- ggplot(
+  data = argi.rural.loadings,
+  aes(
+    x = Landuse, y = Correlation
+  )
+) +
+  geom_boxplot(aes(fill = Landuse), lwd=0.1) +
   geom_hline(
@@
     "rect",
-    ymin = -rural_cuttoff,
-    ymax = rural_cuttoff,
+    ymin = -cutoff,
+    ymax = cutoff,
     xmin = 0, xmax = Inf,
@@
   ) +
-  scale_colour_manual(
+  scale_fill_manual(
     values = c(
-      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
-      "Non.agricultural.unmanaged.green.space" = "#228833",
-      "water" = "#4477AA", other = "#BBBBBB"
-    )
-  )
+      "Arable" = "#CCBB44", "Built Up Area" = "#EE6677",
+      "Non Agricultural Unmanaged Green Space" = "#228833",
+      "Water" = "#4477AA", Other = "#BBBBBB"
+    )
+  )
+
+loading_plot
 
@@
     Landuse = tools::toTitleCase(
-      gsub("\\.", " ", urban_jk_loadings$Landuse)
+      gsub("\\.", " ", urban_jk_loadings$landuse_type)
     ),
@@
 
+cutoff <- unique(urban.loadings$cutoff)[[1]]
+
 urban_loading_text <- tibble(
@@
 )
-loading_plot <- ggplot(data = urban.loadings, aes(
-  x = landuse_type, y = loadings, fill = landuse_type
-)) +
-  geom_boxplot() +
+
+loading_plot <- ggplot(
+  data = urban.loadings,
+  aes(
+    x = Landuse, y = Correlation
+  ),
+  fill = Landuse
+) +
+  geom_boxplot(aes(fill = Landuse), lwd=0.1) +
   geom_hline(
@@
     "rect",
-    ymin = -urban_cuttoff,
-    ymax = urban_cuttoff,
+    ymin = -cutoff,
+    ymax = cutoff,
     xmin = 0, xmax = Inf,

### 2022-02-08 analysis/landuse_effect_prop_scouts.Rmd
increased size of loading plot labels

@@
     vjust = 0.5,
-    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
+    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5), size = rel(textsize)
   ) +

### 2022-02-08 analysis/landuse_effect_prop_scouts.Rmd
made changes recomended by Elli

--- 2022-02-08 analysis/landuse_effect_prop_scouts.Rmd
@@
 In this document I will conduct a PLS and PCA + regression to investigate how land-use type influences the proportion of scouts in both urban and agri-rural environments.
+
+# Proportion of landuse in Urban and Agri-rural
+
+```{r land-use-coverage, message = FALSE, warning = FALSE, error = TRUE}
+rural_total_area <- rural_lu[, 3:ncol(rural_lu)] %>%
+  colSums()
+rural_total_area <- tibble(
+  landuse_type = names(rural_total_area), area = rural_total_area
+)
+rural_total_area$area_pct <- (
+  rural_total_area$area / sum(rural_total_area$area)
+) * 100
+
+rural_total_area <- rural_total_area %>%
+  select(landuse_type, area_pct) %>%
+  rename(
+    Landuse = landuse_type, "% coverage" = area_pct
+  ) %>%
+  mutate(
+    Landuse = tools::toTitleCase(
+      gsub("\\.", " ", rural_total_area$landuse_type)
+    ),
+    Landuse = ifelse(Landuse == "OSR", "Oilseed Rape", Landuse)
+  ) %>%
+  rename(
+    "Land-use" = Landuse
+  )
+
+rural_total_area
+
+urban_total_area <- urban_lu[, 3:ncol(urban_lu)] %>%
+  colSums()
+urban_total_area <- tibble(
+  landuse_type = names(urban_total_area), area = urban_total_area
+)
+urban_total_area$area_pct <- (
+  urban_total_area$area / sum(urban_total_area$area)
+) * 100
+
+urban_total_area <- urban_total_area %>%
+  select(landuse_type, area_pct) %>%
+  rename(
+    Landuse = landuse_type, "% coverage" = area_pct
+  ) %>%
+  mutate(
+    Landuse = tools::toTitleCase(
+      gsub("\\.", " ", urban_total_area$landuse_type)
+    )
+  ) %>%
+  rename(
+    "Land-use" = Landuse
+  )
+
+urban_total_area
+
+save(urban_total_area, rural_total_area, file = "results/area_coverage.RData")
+```
 
@@
 ) +
-  geom_boxplot(aes(fill = Landuse), lwd=0.1) +
+  geom_boxplot(aes(fill = Landuse), lwd = 0.1) +
   geom_hline(

### 2022-02-08 analysis/manuscript/paper.Rmd
fixed box not rendering
made changes recomended by Elli

@@
 Yet, because we cannot easily quantify collective foraging, pinpointing which environments promote collective foraging is challenging.
-Here, we present a mathematical model which allows inferring the proportion of 'scout' and 'recruit' dances on a hive's dance floor and quantify collective foraging.
+Here, we present a mathematical model which quantifies collective foraging by inferring the proportion of 'scout' and 'recruit' dances on a hive's dance floor.
 By applying this methodology to waggle dance data from twenty hives, we show that dance information may be used less in environments where resources are sparse, potentially because colonies are required to send out more scouts to find food in these circumstances.
@@
 
-Within honeybee colonies, a series of simple rules that determine when and how much bees perform the celebrated waggle dance (Box 1) mean that choices between feeding sites occur at the level of the group rather than the individual.
+Within honeybee colonies, a series of simple rules that determine when and how much bees perform the celebrated waggle dance (Figure of dance waggle dance) mean that choices between feeding sites occur at the level of the group rather than the individual.
 For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the sweeter of two equidistant sources [@Seeley1991], or the closer of two equally sweet sources [@Hasenjager2022], without requiring any individual bee to compare options.
@@
 Despite such impressive complexity, in some situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2006; @Gruter2013; @Price2019].
-This indicates that bees frequently ignore the spatial information in dances.
+This indicates that using the dance does not always provide a foraging advantage and suggests bees may ignore spatial information in dances in favour of independent foraging.
+The prevalent explanation for this apparent redundancy portrays dance-based collective foraging as an evolutionary anachronism that was useful in ancient tropical environments but offers little benefit in modern temperate environments [@Dornhaus2004; @Donaldson-Matasci2012; @Price2015].
 
-The prevalent explanation for this apparent redundancy portrays dance-based collective foraging as an evolutionary anachronism that was useful in ancient tropical environments but offers little benefit in modern temperate environments [@Dornhaus2004; @Donaldson-Matasci2012; @Price2015].
-However, fossil evidence has questioned the long-held assumption that the genus Apis evolved in tropical climates [@Kotthoff2013].
-Moreover, given that dancing comes at a significant time cost [@Seeley1988], this raises the issue of why all temperate *Apis mellifera* colonies continue to dance.
+Given that dancing comes at a significant time cost [@Seeley1988], this raises the issue of why all temperate *Apis mellifera* colonies continue to dance.
 The dual questions of what foraging selection pressures drove the evolution of this uniquely complex collective behaviour, and why it continues to exist, are thus yet to be satisfactorily answered.
 Empirical studies have shown that honeybee colonies use dance information flexibly in response to their environment [@Sherman2002; @Dornhaus2004; @Dornhaus2006], suggesting that bees cease to rely on dance information if it provides little benefit [@Price2019].
-Thus, identifying those environments whereby modern honeybee colonies forage collectively rather than through individual search is a key step towards understanding dance evolution [@Price2015].
+Identifying those environments whereby modern honeybee colonies forage collectively rather than through individual search is a key step towards understanding dance evolution [@Price2015].
 
----
 **Box 1**
@@
 
-Honeybees provide an ideal study system in which to evaluate collective foraging thanks to the famous “waggle dance”, which conveys information about location of a resource as a vector of distance (duration of the waggle run) and direction (angle of the dance) [@Frisch1993].
-Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions.
-Resource “quality” is conveyed through the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000]).
-Although bees that follow dances do not specifically interpret this information on an individual level [@Hasenjager2021], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995], and provides the colony with a mechanism to select the most profitable resources in their environment [@Seeley1994A].
+Honeybees provide an ideal study system in which to evaluate collective foraging thanks to the famous “waggle dance”, which conveys information about location of a resource as a vector of distance (duration of the waggle run) and direction (angle of the dance relative to the sun) [@Frisch1993].Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions.
+Resource “quality” is provided through the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000]).
+Although bees that follow dances do not specifically interpret this information on an individual level [@Seeley1991], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995], and provides the colony with a mechanism to select the most profitable resources in their environment [@Seeley1994A].
 
----
+**End Box 1**
 
-Overall, evidence that colony-level collective foraging offers any benefit over individual search has proved surprisingly hard to come by, mainly because this requires to distinguish scouts from recruits.
-Scouts are foragers that have located a resource through individual search behaviour, whilst recruits are foragers that have found a forage site by using the information provided in the dance of another bee [@Seeley1995].
+Overall, evidence that colony-level collective foraging offers any benefit over individual search has proved surprisingly hard to come by.
+Whilst it is entirely possible the dance no longer provides a benefit in the modern world, a key barrier to assessing waggle dance use is that this requires distinguishing between foraging trips made by
+scouting and recruited individuals.
+Scout foraging trips occur when a food source is located through individual search behaviour, whilst recruit foraging trips occur when a forage site is found by using the information provided in the dance of another bee [@Seeley1995].
 Scouting and recruiting are not fixed behavioural categories, because individual bees can engage in both over the course of their foraging lifetime, and foragers can dance on return from any successful trip irrespective of whether they were recruited to the forage site or found it individually [@Beekman2007].
@@
 
-To reveal what the distribution of waggle run durations reported on the dance floor look like, and to demonstrate that dances from recruits and scouts lead to different distributions, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
+To reveal how the patterns of waggle dance encoded foraging distances might differ under a scout and recruit foraging strategy, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
 Foragers could locate these under two different strategies: either acting as a scout and locating resources themselves, or following a recruit strategy and locating resources by following a random dance from the dance floor [@Seeley1992] (Fig.
@@
 3a).
-In each of these cases, the collective model had a good fit (using a Kolmogorov-Smirnov statistic, see methods) to the empirical waggle run durations, and for 9 of these sites, the scout model was significantly different to the data (Fig.
+In all but one of these cases, the collective model had a good fit (using a Kolmogorov-Smirnov statistic, see methods) to the empirical waggle run durations, whereas for 9 of these sites, the distribution predicted by the scout model was significantly different to the observed data (Fig.
 3b).
@@
 
-Our results show that, whilst colony-level foraging mostly use a mixture of scouts and recruits, in some circumstances, colony foraging is better described by a model of individual foraging.
-This suggests that in some environments, honey bees actively ignore the waggle dance and instead prioritise the use of private information.
+Our results show that, whilst colony-level foraging is mostly comprised of a mixture of scout and recruit foraging trips, in some circumstances, colony foraging can be better described by individual foraging alone.
+This suggests that in some environments honey bees engage in scouting to find new food sites rather than follow dances.
 Previous empirical studies have shown that individual foragers flexibly alter their use of waggle dance information in response to unrewarding resources [@Gruter2011; @Price2019].
@@
 
-Our finding that foraging in some landscapes is more accurately described by ignoring the dance and focusing on private information supports the hypothesis that the dance is used flexibly in response to environmental factors.
-As the collective model quantifies waggle dance use, the fit of our model provides a new tool to evaluate how environmental factors influence waggle dance use, and thus how a colony values waggle dance information under different environmental circumstances.
+Our finding that foraging in some landscapes is more accurately described by scouting supports the hypothesis that the dance is used flexibly in response to environmental factors.
+As the collective model quantifies the contribution of recruitment to colony foraging, the fit of our model provides a new tool to evaluate how environmental factors influence waggle dance use, and thus how a colony uses waggle dance information under different environmental circumstances.
 
-We compared the estimates of the proportion of scout and recruits for colonies in different environments.
+We estimated the proportion of scout and recruit foraging trips for each individual colony and investigated whether these estimates might correlate with land-use.
 Each site used in this study is situated in a landscape comprised of different proportions of land-use types.
-To evaluate how these different land-use types influenced waggle dance use, we first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments separately (see methods).
-As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances.
+To evaluate how these different land-use types influenced waggle dance use, we first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments separately as many land-use types present in urban areas do not occur in agri-rural environments and vice versa (see methods).
+As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances within agri-rural and urban environments.
+As we have a small sample size, this analysis was jackknifed to evaluate the robustness of our results to influential points (see methods, supplementary material section _._)
 
@@
 B).
-From the land-use maps coloured by each landscape dominating PC1 (Fig 4) we can see that arable land and built-up areas show the most coverage.
-As arable land increases whilst unmanged green space and built-up areas decrease, the proportion of individuals prioritising private information increases.
+These land-use types maintain a significant correlation with PC1 over the jackknifed PLS (Fig.
+4.
+B) indicating the results are robust.
+As arable land increases whilst unmanaged green space and built-up areas decrease, the proportion of individuals prioritising private information increases.
+Arable land dominates the agri-rural environment accounting for over 28% of land-coverage, whilst built-up areas represent around 17% of lands (Table 1).
+Non-agricultural unmanaged green space represent smaller amounts of land, accounting for less than 3% of the total land combined (Table 1).
+
+```{r, agri-rural-area-table}
+set_flextable_defaults(
+  big.mark = " ",
+  font.size = 10, theme_fun = theme_vanilla,
+  padding.bottom = 6,
+  padding.top = 6,
+  padding.left = 6,
+  padding.right = 6
+)
+
+load("../results/area_coverage.RData")
+
+rural_total_area %>%
+  arrange(desc(`% coverage`)) %>%
+  mutate(`% coverage` = signif(`% coverage`, 3)) %>%
+  flextable() %>%
+  set_caption(
+    "Percentage area covered for each land-use type in the agri-rural environment"
+  ) %>%
+  set_table_properties(layout = "autofit")
+```
 
@@
 B).
+However, the jackknifed PLS identifies CAD as a highly influential site, the removal of which causes sparse residential land to have a positive correlation with PC1 (Supplementary Material _._).
 Visualising how waggle dance use is influenced by changes in these land-use types shows that when the landscape is dominated by sparse residential land, as in site HHS, waggle dance use dominates the colonies foraging strategy.
@@
 5).
+Sparse residential land dominates the urban environment accounting for over 34% of land-coverage, whilst parks allotments and cemeteries and railway corresponds to ~8% and ~1 respectively (Table 2).
+
+```{r, urban-area-table}
+urban_total_area %>%
+  arrange(desc(`% coverage`)) %>%
+  mutate(`% coverage` = signif(`% coverage`, 3)) %>%
+  flextable() %>%
+  set_caption(
+    "Percentage area covered for each land-use type in the agri-rural environment"
+  ) %>%
+  set_table_properties(layout = "autofit")
+```
 
@@
 Arable land is typically nutritionally poor for honey bees [@Samuelson2021].
-On the contrary, non-agricultural unmanaged green space is typically nutrient rich, as are built up areas, which often have gardens with ample flowers available for pollinators [@Baldock2019].
-These findings indicate that in agri-rural environments, when resources are more abundant and thus easier to find, the collective foraging is used more, presumably to identify the best resources in the environment.
-Conversely, where resources are sparse and more difficult to find, a decreased use of the waggle dance suggests the colony focuses more on individual exploration rather than in trying to locate on the best resources.
-A similar result is observed for urban environments, where the dependence appears to be driven by the presence of flower rich gardens in residential areas.
-However, an increase in the amount of parks, allotments and cemeteries appears to decrease the use of the dance.
+On the contrary, built up areas often have gardens with ample flowers available for pollinators [@Baldock2019].
+Although non-agricultural unmanaged green space and water show as significantly correlating with reduced use of the dance, as they cover such a small amount of area, combined with a small sample size of only 10 sites, we cannot ignore the possibility that their significance is a random effect rather than evidence of a biological mechanism.
+Overall, these findings indicate that in agri-rural environments, when resources are more abundant and thus easier to find, the waggle dance is used more, presumably to identify the best resources of those on offer.
+Conversely, where resources are sparse and more difficult to find, a decreased use of the waggle dance suggests the colony focuses more on individual exploration rather than in trying to locate the best resources.
 
-( **NOTE:** I feel like we need more of an explanation here.
-For the agri-rural environment we can assemble the results into a coherent story which has some logical basis and suggests more data would investigate this further.
-The results for the urban sites are contradictory.
-Parks, allotments and cemeteries increases the proportion of scouts so following our logic these areas should be barron, but the evidence suggests they are not and along side gardens should be full of flora.
-This indicates there is something else in play here, such as the spatial arrangement of the locations?
-Or that we just don't have enough data to comment on what processes are going on in urban sites.
-However, we do get a 'significant' result in the PLS.
-Maybe we need more fine grained data to evaluate how season influences, it could be that these results are muddled by an interaction with season or temperature or something.
-Also, railway is a random thing to pop up as significant.
-I am not sure how best to describe these results without it looking like blatant data mining...)
+A similar result is observed in urban sites, where the presence of flower rich gardens in residential areas appears to promote increased use of the dance.
+Conversely, an increase in the amount of parks, allotments and cemeteries appears to decrease the use of the dance.
+Such land-use types could be expected to be quite resource rich for pollinators yet our findings suggests honey bee colonies decrease the use of the dance in such environments.
+It is possible that such environments vary seasonally in terms of their forage availability, allowing them to function both as floral deserts and oasis at different times of the year.
+As our waggle dance observations were collected over the course of an entire season, we may only be picking up the dominant signal over this time frame.
+However, the jackknife analysis shows our results are highly sensitive to influential points.
+Further data is therefore needed to investigate how the urban environment influences waggle dance use in honey bees.
 

### 2022-02-08 analysis/landuse_effect_prop_scouts.Rmd
fixed naumgs colours on plot

@@
       "Arable" = "#CCBB44", "Built Up Area" = "#EE6677",
-      "Non Agricultural Unmanaged Green Space" = "#228833",
+      "NAUMGS" = "#228833",
       "Water" = "#4477AA", Other = "#BBBBBB"

### 2022-03-14 analysis/manuscript/paper.Rmd
updated changes made by Elli. Going to read through fresh in the morning and update before sending to Vicent for final edits.

@@
 ---
-title: "Forage site distributions quantify collective foraging in honey bee colonies"
+title: "Waggle run distributions quantify collective foraging in honey bee colonies"
 author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
@@
 
-Alternative title:
-
-- Waggle run durations quantify collective foraging in honey bee colonies
-- Waggle run duration distributions quantify collective foraging in honey bee colonies
-- Distributions of waggle dance durations quantifies collective foraging in honey bee colonies
-
-Or something like this.
-
 # Abstract
 
-Honeybee foraging is a collective behaviour that is directed through the extraordinary waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops [@Frisch1993; @Seeley1995].
-Despite having been described as the most astounding example of non-primate communication [@Frisch1993], recent evidence suggests that the benefits of using dance information vary across environments and are sometimes even negligible [@Dornhaus2006; @Price2019].
-Yet, because we cannot easily quantify collective foraging, pinpointing which environments promote collective foraging is challenging.
-Here, we present a mathematical model which quantifies collective foraging by inferring the proportion of 'scout' and 'recruit' dances on a hive's dance floor.
-By applying this methodology to waggle dance data from twenty hives, we show that dance information may be used less in environments where resources are sparse, potentially because colonies are required to send out more scouts to find food in these circumstances.
-Our methodology provides a tool to quantify waggle dance use in undisturbed hives at a much larger scale than hitherto has been possible and to identify those ecological conditions that may have driven the evolution of this remarkable collective behaviour.
+Honeybee foraging is an extraordinary collective behaviour that is directed through the celebrated waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops [@Frisch1993; @Seeley1995].
+Here, we present a mathematical model that quantifies the extent to which honeybee colonies behave collectively in different ecological environments by inferring the proportion of dances performed by bees that have explored individually, and those that have been recruited to food sources.
+By applying this methodology to waggle dance data from twenty hives, we show that colonies are less reliant on dance information where resources are sparse, potentially because colonies are required to send out more scouts to find food in these circumstances.
+Our methodology provides a tool to identify those ecological conditions in which honeybee colonies rely on dance communication, and so to understand the selection pressures that may have driven the evolution of this remarkable collective behaviour.
 
@@
 
-Despite such impressive complexity, in some situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2006; @Gruter2013; @Price2019].
-This indicates that using the dance does not always provide a foraging advantage and suggests bees may ignore spatial information in dances in favour of independent foraging.
-The prevalent explanation for this apparent redundancy portrays dance-based collective foraging as an evolutionary anachronism that was useful in ancient tropical environments but offers little benefit in modern temperate environments [@Dornhaus2004; @Donaldson-Matasci2012; @Price2015].
+Despite such impressive complexity, a wealth of research now shows that foraging bees frequently do not use dance information to find forage sites [@Gruter2009].
+Accordingly, in many situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2006; @Gruter2013; @Price2019].
+This apparent variation in the use of collective behaviour is intriguing because it may provide clues as to those ecological selection pressures that were critical in the evolution of the dance communication system.
+Despite the extraordinary amount of research interest that has focused on the proximate mechanisms of collective decision-making via dance communication, the ultimate evolutionary drivers of its unique evolution within *Apis* -unparalleled in even a single other social insect- remain obscure [@Price2015].
+Identifying those circumstances in which bees use the dance is likely to be a key step towards understanding when it is likely to afford fitness benefits at the colony level, but attempts to do so are challenging because they require constant monitoring of real-world colony weight [@Kirchner1998; @Sherman2002; @Dornhaus2006; @Gruter2013; @Nurnberger2017; @Price2019] -a noisy proxy of foraging success that is influenced by many other abiotic and biotic factors, including colony health- over ecologically credible time periods [@Schurch2014].
+To do this at a scale that allows inter-colony variation in foraging environments at sufficient replication is a major logistical hurdle (but see [@Nurnberger2017]).
 
-Given that dancing comes at a significant time cost [@Seeley1988], this raises the issue of why all temperate *Apis mellifera* colonies continue to dance.
-The dual questions of what foraging selection pressures drove the evolution of this uniquely complex collective behaviour, and why it continues to exist, are thus yet to be satisfactorily answered.
-Empirical studies have shown that honeybee colonies use dance information flexibly in response to their environment [@Sherman2002; @Dornhaus2004; @Dornhaus2006], suggesting that bees cease to rely on dance information if it provides little benefit [@Price2019].
-Identifying those environments whereby modern honeybee colonies forage collectively rather than through individual search is a key step towards understanding dance evolution [@Price2015].
+**Figure of honeybee foraging**
 
-**Box 1**
-
-[figure of waggle dance, maybe have text in like a box or something? Or just the text]
+[figure of waggle dance]
 
@@
 
-Overall, evidence that colony-level collective foraging offers any benefit over individual search has proved surprisingly hard to come by.
-Whilst it is entirely possible the dance no longer provides a benefit in the modern world, a key barrier to assessing waggle dance use is that this requires distinguishing between foraging trips made by
-scouting and recruited individuals.
-Scout foraging trips occur when a food source is located through individual search behaviour, whilst recruit foraging trips occur when a forage site is found by using the information provided in the dance of another bee [@Seeley1995].
+Here, we present a a model that can identify the proportion of foraging trips that are made by bees recruited through waggle dances, and thus quantify collective behaviour, in unmanipulated hives with no requirement to manipulate hive orientations or assay foraging efficiency.
+Our model seeks to distinguish between “scout” foraging trips, in which food sources that have been located through individual search are reported by waggle dancing, and “recruit” trips in which a forage site is found by using the information provided in the dance of another bee [@Seeley1995].
 Scouting and recruiting are not fixed behavioural categories, because individual bees can engage in both over the course of their foraging lifetime, and foragers can dance on return from any successful trip irrespective of whether they were recruited to the forage site or found it individually [@Beekman2007].
-So far, scouts and recruits are distinguished by means of the foragers' individual search history, which requires following individual bees and is time and labour intensive [@Seeley1995].
-
-Here, we will use a different approach and develop a model that describes the proportion of foraging trips in which bees act as scouts, and as recruits, within a colony, based on the distribution of foraging distances reported by waggle dancing, encoded as the duration of the run phase of the dance [@Frisch1993].
-We will use this model to infer the proportion of scouts and recruits on the dance floor from dance data, and to demonstrate the principle, apply this methodology to dance data from a collection of hives in different environments.
+In a hypothetical colony that relied only on individual search to find foraging sites, all trips would be “scout” trips, but as recruitment becomes more important, the proportion of “recruit” trips will increase.
+Previously, scout and recruit trips have only been distinguishable by means of the foragers’ individual search history, which requires following individual bees and is time and labour intensive [@Seeley1995].
 
@@
 D.
-Log inverse frequency distribution of foraging distances reported by recruits."}
+Log inverse frequency distribution of foraging distances reported by recruits.
+Here, inverse frequency distribution refers to probability of sampling a value greater than $X$."}
 knitr::include_graphics("../results/figures/simulation.png")
@@
 
-Scouts and recruits create different foraging distance distributions on the dance-floor.
-These different distance distributions can be used to quantify the relative importance of collective foraging.
-Scouting foragers report a distribution of distances that reflects their probability of encountering a resource in the environment (see methods), whilst foragers that recruits report distances that reflect the distribution of the most profitable resource available (see methods).
+Scouts and recruits create different waggle run distributions on the dance-floor.
+These different distributions can be used to quantify the relative importance of collective foraging.
+Scouting foragers report a distribution of distances that reflects their probability of encountering a resource in the environment (see methods), whilst recruits report distances that reflect the distribution of the most profitable resources available (see methods).
 On the dance-floor, the distribution of distances is modified through the number of runs performed for a resource, which depends on the profitability of a resource, which honeybees achieve by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994A].
@@
 
-To evaluate the use of waggle dance information and individual search in honeybee colonies foraging in 'natural' landscapes, we analysed an existing dataset of 3378 waggle dance observations from 20 observation hives, recorded between April-September 2017, (previously described in [@Samuelson2021]).
-Hives were situated at different locations in South East England (see methods, figure 3A) and were visited every two weeks for a period of 24 weeks.
+To evaluate the use of waggle dance information and individual search in honeybee colonies foraging in 'natural' landscapes, we analysed a pre-existing dataset of 3378 waggle dance observations from 20 observation hives, recorded between April-September 2017, (previously described in [@Samuelson2021]).
+Hives that contributed to this dataset had been situated at different locations in South East England (see methods, figure 3A) and visited every two weeks for a period of 24 weeks.
 On each visit, two hours of continuous waggle dance data was recorded by training a camcorder onto the dance floor.
-The footage of the dances was decoded manually ([@Couvillon2012; @Schurch2019] to extract waggle run durations.
-For each site we fit both the collective and individual models and used model selection to determine which provided the better explanation of the data, and (if the collective model provided a better fit) to quantify the relative use of social information through estimating the parameter p.
+The footage of the dances was decoded manually [@Couvillon2012; @Schurch2019] to extract waggle run durations.
+Using this data, for each site we fit both the collective and individual models and used model selection to determine which provided the better explanation of the data, and (if the collective model provided a better fit) to quantify the relative use of social information through estimating the parameter p.
 We calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004].
@@
 Sites STU and ZSL (C & D) are marked.
-Inset plot shows the location of the sites within the UK.
+Inset plot shows the location of the sites within the UK; shaded area in the main plot indicates Greater London.
 B.
@@
 C.
-Model fits to waggle run durations collected from the ZSL hive along with histogram showing distribution.
+Model fits to waggle run durations collected from the ZSL hive, for which the model did not outperform the individual model, along with histogram showing distribution.
 D.
@@
 
-Our results show that, whilst colony-level foraging is mostly comprised of a mixture of scout and recruit foraging trips, in some circumstances, colony foraging can be better described by individual foraging alone.
-This suggests that in some environments honey bees engage in scouting to find new food sites rather than follow dances.
-Previous empirical studies have shown that individual foragers flexibly alter their use of waggle dance information in response to unrewarding resources [@Gruter2011; @Price2019].
-Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individuals foraging behaviour over a landscape.
+These results show that, whilst colony-level foraging is mostly comprised of a mixture of scout and recruit foraging trips, in some circumstances, colony foraging can be better described by individual foraging alone.
+This suggests that in some environments honey bees mainly engage in scouting to find new food sites rather than follow dances.
+Previous empirical studies have shown that individual foragers may flexibly alter their use of waggle dance information if its use is proving unproductive [@Gruter2011; @Price2019].
+Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individual's foraging behaviour over a landscape.
 As our results show that individual foraging accurately describes colony foraging in three different sites, these findings support the idea that individual decision making can dominate colony foraging and demonstrates further evidence for flexible waggle dance use by honey bee colonies.
 
-Our finding that foraging in some landscapes is more accurately described by scouting supports the hypothesis that the dance is used flexibly in response to environmental factors.
-As the collective model quantifies the contribution of recruitment to colony foraging, the fit of our model provides a new tool to evaluate how environmental factors influence waggle dance use, and thus how a colony uses waggle dance information under different environmental circumstances.
+Further quantification of the use of waggle-dance recruitment within all colonies, as a proportion of all foraging trips, can be achieved by extracting the estimated proportion of scout trips (p) for each site.
+Since our sites varied in land-use characteristics and potentially thus forage distributions (although in this case, not by design for this study), we investigated whether these estimates might correlate with land-use.
+We first classified the different land-use types of the area surrounding each site [@Samuelson2018] to obtain a standardised land-use profile for the urban and agri-rural environments separately as many land-use types present in urban areas do not occur in agri-rural environments and vice versa (see methods).
+We then performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances within agri-rural and urban environments.
+As we have a small sample size (20 sites; 10 urban and 10 agri-rural), this analysis was jackknifed to evaluate the robustness of our results to influential points (see methods, supplementary material section _._)
 
-We estimated the proportion of scout and recruit foraging trips for each individual colony and investigated whether these estimates might correlate with land-use.
-Each site used in this study is situated in a landscape comprised of different proportions of land-use types.
-To evaluate how these different land-use types influenced waggle dance use, we first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments separately as many land-use types present in urban areas do not occur in agri-rural environments and vice versa (see methods).
-As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances within agri-rural and urban environments.
-As we have a small sample size, this analysis was jackknifed to evaluate the robustness of our results to influential points (see methods, supplementary material section _._)
-
-In the agri-rural environment one combination of land use types explained ~52% of the variation in the proportion of scouts (beta regression: $R^2 = 0.52$, $\phi$ = 3.7, p < 0.05, Fig.
+In the agri-rural environments one combination of land use types explained ~52% of the variation in the proportion of scouts (beta regression: $R^2 = 0.52$, $\phi$ = 3.7, p < 0.05, Fig.
 4.
 A).
-This principal component correlates positively with arable land and negatively with non-agricultural unmanaged green space, built-up areas and water (Fig.
+This principal component correlates positively with arable land (29% of land coverage; Table 1) and negatively with built-up areas (17% of land coverage) ; it also correlates negatively with non-agricultrual unmanaged green space and water , but together these represent less than 3% of land-use (Fig.
 4.
@@
 B) indicating the results are robust.
-As arable land increases whilst unmanaged green space and built-up areas decrease, the proportion of individuals prioritising private information increases.
-Arable land dominates the agri-rural environment accounting for over 28% of land-coverage, whilst built-up areas represent around 17% of lands (Table 1).
-Non-agricultural unmanaged green space represent smaller amounts of land, accounting for less than 3% of the total land combined (Table 1).
+As arable land increases whilst built-up areas decrease, the proportion of trips that are driven by individual search increases.
 
@@
 
-A similar effect is also observed in the urban environment, where a single PC explained ~59% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.59, $\phi$ = 6.2, p < 0.05, Fig.
+A similar effect was observed in the urban environment, where a single PC explained ~59% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.59, $\phi$ = 6.2, p < 0.05, Fig.
 5.
 A).
-This PC correlates positively with parks, allotments and cemeteries, whilst correlating negatively with sparse residential lands and railways (Fig.
+This PC correlates positively with parks, allotments and cemeteries (8% of land cover; Table 2), whilst correlating negatively with sparse residential lands (35% of land cover) and railways but these represent only 1% of land cover.
+(Fig.
 5.
 B).
-However, the jackknifed PLS identifies CAD as a highly influential site, the removal of which causes sparse residential land to have a positive correlation with PC1 (Supplementary Material _._).
-Visualising how waggle dance use is influenced by changes in these land-use types shows that when the landscape is dominated by sparse residential land, as in site HHS, waggle dance use dominates the colonies foraging strategy.
-However, this reduces as sparse residential land is replaced by parks, allotments, and cemeteries (Fig.
-5).
-Sparse residential land dominates the urban environment accounting for over 34% of land-coverage, whilst parks allotments and cemeteries and railway corresponds to ~8% and ~1 respectively (Table 2).
+However, we have low confidence in this finding because the jackknifed PLS identifies CAD as a highly influential site, the removal of which causes sparse residential land to have a positive correlation with PC1 (Supplementary Material _._).
 
@@
 
-In agri-rural environments, land-use types which typically associate with resource scarcity correlate positively with a higher proportion of scouts, whilst those which correlate with a higher resource abundance correlate with a lower proportion of scouts.
-Arable land is typically nutritionally poor for honey bees [@Samuelson2021].
-On the contrary, built up areas often have gardens with ample flowers available for pollinators [@Baldock2019].
-Although non-agricultural unmanaged green space and water show as significantly correlating with reduced use of the dance, as they cover such a small amount of area, combined with a small sample size of only 10 sites, we cannot ignore the possibility that their significance is a random effect rather than evidence of a biological mechanism.
-Overall, these findings indicate that in agri-rural environments, when resources are more abundant and thus easier to find, the waggle dance is used more, presumably to identify the best resources of those on offer.
-Conversely, where resources are sparse and more difficult to find, a decreased use of the waggle dance suggests the colony focuses more on individual exploration rather than in trying to locate the best resources.
-
-A similar result is observed in urban sites, where the presence of flower rich gardens in residential areas appears to promote increased use of the dance.
-Conversely, an increase in the amount of parks, allotments and cemeteries appears to decrease the use of the dance.
-Such land-use types could be expected to be quite resource rich for pollinators yet our findings suggests honey bee colonies decrease the use of the dance in such environments.
-It is possible that such environments vary seasonally in terms of their forage availability, allowing them to function both as floral deserts and oasis at different times of the year.
-As our waggle dance observations were collected over the course of an entire season, we may only be picking up the dominant signal over this time frame.
-However, the jackknife analysis shows our results are highly sensitive to influential points.
-Further data is therefore needed to investigate how the urban environment influences waggle dance use in honey bees.
+Across both environments our results hint at a decrease in reliance on waggle dance recruitment as resources become harder to find and foraging trips become longer [@Samuelson2021].
+Arable land is likely to be nutritionally poor for pollinators (note that oilseed rape field and fruit crops were not included in this category), while residential areas -particularly gardens- are considered nutritional hotspots for bees [@Baldock2019].
+Similarly, in the urban environments sparse-residential coverage correlated negatively with the proportion of scout trips.
+However, we are cautious in this interpretation for two reasons.
+Firstly, because our study was designed as a proof-of-principle based on an existing dataset, we did not systematically assess the availability of forage within these sites.
+Secondly, although we saw a similar pattern in the urban data for sparse residential areas (35% of land-cover), which correlated negatively with the proportion of scout trips, this pattern reversed when a highly influential site (CAD; identified by the jacknifed PLS) was removed, and we also found parks, allotments and cemeteries (8% land cover), which could be expected to provide abundant forage for bees [@Baldock2019], to correlate positively with independent search.
 
 The fitting of our model to waggle dance observations provides a novel tool to quantify the relative importance of social and private information in honeybees.
+Whilst recruitment is well known to occur through other mechanisms besides the waggle dance, such as through olfactory cues, the exceptionally close fit of our model to waggle run durations underlines the importance of the waggle dance in honey bee foraging.
 Further, by identifying the land-use combinations which most influence variation in the proportion of scouts, our analysis allows us to map the environment along the major axis of honeybee information use and visualise how land-use influences the use of the waggle dance.
+
 Our results within the agri-rural sites are in broad agreement with other studies which evaluated foraging performance with and without the dance [@Dornhaus2006; Price2019].
@@
 However, the results for our urban sites do not produce a clear biological rational and present an all together more varied picture.
-More waggle dance data from urban colonies is therefore required to further probe how honeybees use the waggle dance in this heavily anthropogenic environment.
+An alternative hypothesis could be that the the dance is influenced by resource stability more than abundance and that longer term effects may contribute to the benefits of the dance [@Schurch2014].
+Whilst the limiting number of sites used in this study handicaps our ability to progress this debate here, as a proof of concept, the methods derived here provide a new toolkit for evaluating factors influencing waggle dance recruitment
 

### 2022-03-15 analysis/manuscript/paper.Rmd
updated ms with latest edits

@@
 
-**Figure of honeybee foraging**
-
 [figure of waggle dance]
 
-Honeybees provide an ideal study system in which to evaluate collective foraging thanks to the famous “waggle dance”, which conveys information about location of a resource as a vector of distance (duration of the waggle run) and direction (angle of the dance relative to the sun) [@Frisch1993].Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions.
-Resource “quality” is provided through the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000]).
-Although bees that follow dances do not specifically interpret this information on an individual level [@Seeley1991], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995], and provides the colony with a mechanism to select the most profitable resources in their environment [@Seeley1994A].
+*Honeybees provide an ideal study system in which to evaluate collective foraging thanks to the famous “waggle dance”, which conveys information about location of a resource as a vector of distance (duration of the waggle run) and direction (angle of the dance relative to the sun) [@Frisch1993].Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions.
+Resource “quality” is provided through the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000].
+Although bees that follow dances do not specifically interpret this information on an individual level [@Seeley1991], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995], and provides the colony with a mechanism to select the most profitable resources in their environment [@Seeley1994A].*
 
-**End Box 1**
+[end of figure of waggle dance]
 
@@
 3b).
-The sites shown in Figs 3c-d are representative examples showing the model fits, showing the overall quality of the model fit to the data.
+The sites STU and ZSL shown in Figs 3c-d are representative examples showing the model fits where the collective (STU) and individual (ZSL) models fit best, showing the overall quality of the model fit to the data.
 
@@
 C.
-Model fits to waggle run durations collected from the ZSL hive, for which the model did not outperform the individual model, along with histogram showing distribution.
+Model fits to waggle run durations collected from the STU hive, for which the collective model (red line) did not outperform the individual model (blue line), along with histogram showing distribution.
 D.
-Model fits to waggle run durations collected from the STU hive along with histogram showing distribution.
+Model fits to waggle run durations collected from the ZSL hive, where the collective model (red line) outperformed the individual model (blue line), along with histogram showing distribution.
 Model fits show the compliment cumulative frequency distribution, giving the probability of sampling a value greater than or equal to x."}
@@
 A).
-This principal component correlates positively with arable land (29% of land coverage; Table 1) and negatively with built-up areas (17% of land coverage) ; it also correlates negatively with non-agricultrual unmanaged green space and water , but together these represent less than 3% of land-use (Fig.
+This principal component correlates positively with arable land (29% of land coverage; Table 1) and negatively with built-up areas (17% of land coverage) ; it also correlates negatively with non-agricultrual unmanaged green space and water, but together these represent less than 3% of land-use (Fig.
 4.
@@
 A).
-This PC correlates positively with parks, allotments and cemeteries (8% of land cover; Table 2), whilst correlating negatively with sparse residential lands (35% of land cover) and railways but these represent only 1% of land cover.
+This PC correlates positively with parks, allotments and cemeteries (8% of land cover; Table 2), whilst correlating negatively with sparse residential lands (35% of land cover) and railways, but these represent only 1% of land cover.
 (Fig.
@@
 
-Our results within the agri-rural sites are in broad agreement with other studies which evaluated foraging performance with and without the dance [@Dornhaus2006; Price2019].
+Our results within the agri-rural sites are in broad agreement with other studies which evaluated foraging performance with and without the dance [@Dornhaus2006; @Price2019].
 These results suggest that, even though collective foraging is not always beneficial, that in environments where high quality resources are present, but relatively scarce, collective foraging is being carried out by colonies, suggesting that under these conditions exchanging social information through the waggle dance confers benefits.

### 2022-04-04 analysis/manuscript/paper.Rmd
Updated MS edits and new figures. final edits!

@@
 ---
-title: "Waggle run distributions quantify collective foraging in honey bee colonies"
+title: "Waggle dance distributions quantify collective foraging in honey bee colonies"
 author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
@@
 bibliography: paper-refs.bib
+# csl: nature.csl
 ---
@@
 Here, we present a mathematical model that quantifies the extent to which honeybee colonies behave collectively in different ecological environments by inferring the proportion of dances performed by bees that have explored individually, and those that have been recruited to food sources.
-By applying this methodology to waggle dance data from twenty hives, we show that colonies are less reliant on dance information where resources are sparse, potentially because colonies are required to send out more scouts to find food in these circumstances.
-Our methodology provides a tool to identify those ecological conditions in which honeybee colonies rely on dance communication, and so to understand the selection pressures that may have driven the evolution of this remarkable collective behaviour.
+By applying this methodology to waggle dance data from twenty hives, we show our model closely fits real-world honeybee foraging patterns and demonstrate that colonies vary their use of waggle dance information across different landscapes.
+Our methodology provides a novel tool to identify the ecological conditions in which honeybee colonies rely on dance communication, opening the door to large-scale experimental exploration of the selection pressures that may have driven the evolution of this remarkable collective behaviour.
 
-# Main
+# Introduction
 
-Within honeybee colonies, a series of simple rules that determine when and how much bees perform the celebrated waggle dance (Figure of dance waggle dance) mean that choices between feeding sites occur at the level of the group rather than the individual.
+Within honeybee colonies, a series of simple rules that determine when and how much bees perform the celebrated waggle dance (Fig.
+1) mean that choices between feeding sites occur at the level of the group rather than the individual.
 For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the sweeter of two equidistant sources [@Seeley1991], or the closer of two equally sweet sources [@Hasenjager2022], without requiring any individual bee to compare options.
-This extraordinary system is a key example of how social insect colony behaviour can take on a form extending beyond that of the individual units [@Sumpter2006].
+This extraordinary system is a key example of how social insect colony behaviour can take on a form extending beyond that of the individual units [@Sumpter2006], functioning as a collective that makes decisions as a single entity.
 
-Despite such impressive complexity, a wealth of research now shows that foraging bees frequently do not use dance information to find forage sites [@Gruter2009].
-Accordingly, in many situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2006; @Gruter2013; @Price2019].
+Despite its intricacy and precision, a wealth of research now shows that foraging bees frequently do not use dance information to find forage sites [@Gruter2009].
+Accordingly, in many situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2004 @Dornhaus2006; @Gruter2013; @Price2019].
 This apparent variation in the use of collective behaviour is intriguing because it may provide clues as to those ecological selection pressures that were critical in the evolution of the dance communication system.
-Despite the extraordinary amount of research interest that has focused on the proximate mechanisms of collective decision-making via dance communication, the ultimate evolutionary drivers of its unique evolution within *Apis* -unparalleled in even a single other social insect- remain obscure [@Price2015].
-Identifying those circumstances in which bees use the dance is likely to be a key step towards understanding when it is likely to afford fitness benefits at the colony level, but attempts to do so are challenging because they require constant monitoring of real-world colony weight [@Kirchner1998; @Sherman2002; @Dornhaus2006; @Gruter2013; @Nurnberger2017; @Price2019] -a noisy proxy of foraging success that is influenced by many other abiotic and biotic factors, including colony health- over ecologically credible time periods [@Schurch2014].
-To do this at a scale that allows inter-colony variation in foraging environments at sufficient replication is a major logistical hurdle (but see [@Nurnberger2017]).
+
+Identifying the circumstances in which bees use the dance is likely to be a key step towards understanding when it is likely to afford fitness benefits at the colony level.
+Despite the amount of research interest that has focused on the proximate mechanisms of collective decision-making via dance communication, the ultimate evolutionary drivers of its unique evolution within *Apis* -unparalleled in even a single other social insect- remain obscure [@Price2015].
+While it is likely that the spatiotemporal distribution of forage was key, some studies find the benefits of dance communication to be realized only in challenging environments where resources are clumped or ephemeral [@Sherman2002; @Dornhaus2004; @Dornhaus2006], while others find no effects of landscape heterogeneity [@Nurnberger2017], or that dancing comes at a cost in challenging environments [@Price2019] and is most beneficial when species richness is high [@Donaldson-Matasci2012].
+
+Attempts to identify the circumstances in which dance communication is important have been challenging because they require constant monitoring of real-world colony weight [@Kirchner1998; @Sherman2002; @Dornhaus2006; @Gruter2013; @Nurnberger2017; @Price2019] -a noisy proxy of foraging success that is influenced by many other abiotic and biotic factors, including colony health- over ecologically credible time periods [@Schurch2014].
+To do this at a scale that allows inter-colony variation in foraging environments at sufficient replication is a major logistical hurdle (but see [@Nurnberger2017]), particularly given that multiple landscape variables may interact to determine the utility of dance communication.
 
@@
 
-*Honeybees provide an ideal study system in which to evaluate collective foraging thanks to the famous “waggle dance”, which conveys information about location of a resource as a vector of distance (duration of the waggle run) and direction (angle of the dance relative to the sun) [@Frisch1993].Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions.
-Resource “quality” is provided through the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000].
+**The honeybee waggle dance carries information about the location of a resource** *as a vector of distance (duration of the waggle run) and direction (angle of the dance relative to the sun) [@Frisch1993].Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions.
+Overall resource “quality” -the net energetic gain of a foraging trip- is provided through the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000].
 Although bees that follow dances do not specifically interpret this information on an individual level [@Seeley1991], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995], and provides the colony with a mechanism to select the most profitable resources in their environment [@Seeley1994A].*
@@
 
-Here, we present a a model that can identify the proportion of foraging trips that are made by bees recruited through waggle dances, and thus quantify collective behaviour, in unmanipulated hives with no requirement to manipulate hive orientations or assay foraging efficiency.
-Our model seeks to distinguish between “scout” foraging trips, in which food sources that have been located through individual search are reported by waggle dancing, and “recruit” trips in which a forage site is found by using the information provided in the dance of another bee [@Seeley1995].
-Scouting and recruiting are not fixed behavioural categories, because individual bees can engage in both over the course of their foraging lifetime, and foragers can dance on return from any successful trip irrespective of whether they were recruited to the forage site or found it individually [@Beekman2007].
-In a hypothetical colony that relied only on individual search to find foraging sites, all trips would be “scout” trips, but as recruitment becomes more important, the proportion of “recruit” trips will increase.
-Previously, scout and recruit trips have only been distinguishable by means of the foragers’ individual search history, which requires following individual bees and is time and labour intensive [@Seeley1995].
+Here, we present a method to determine the extent of waggle dance use in colony foraging within specific landscapes by identifying the proportion of foraging trips that are made by bees recruited through waggle dances.
+Our method examines the distribution of waggle run durations reported on the dance-floor and infers from its shape the relative contribution of dances for resources identified by individual search, "scouting" versus dances for resources found by following dances, “recruitment” [@Seeley1995].
+Previously, such stips have only been distinguishable by documenting foragers' individual search history, which requires tracking individual bees and is time -and labour- intensive [@Seeley1995].
+Our methods provide an efficient means of quantifying collective behaviour with no requirement to manipulate hive orientations, assay foraging efficiency or document individuals' previous experience.
 
-To reveal how the patterns of waggle dance encoded foraging distances might differ under a scout and recruit foraging strategy, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
+# Results
+
+To reveal how the patterns of waggle dance-encoded foraging distances might differ under scenarios where colonies rely on individual search vs recruitment, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
 Foragers could locate these under two different strategies: either acting as a scout and locating resources themselves, or following a recruit strategy and locating resources by following a random dance from the dance floor [@Seeley1992] (Fig.
-1).
-As it is known in the simulation which individuals forage under what strategy, we can compare the distributions of foraging distances reported on the dance floor.
-Figs 1c and 1d show that the shapes of the resource distance distributions for the two types of foraging trips are very different.
-The distance distribution of the scouts is akin to that of an exponential distribution (Fig.
-1 C), which is the nearest neighbour distance distribution for foragers operating in a one-dimensional environment (see methods).
-The distribution of the distances reported by recruits (Fig.
-1 D) is a Rayleigh distribution which is the nearest-neighbour distribution of a two-dimensional environment [@Pyke1978] (see Methods).
+2, details in Methods1).
+As it is known in the simulation which individuals in the hive forage under what strategy, we can compare the distributions of foraging distances reported on the dance floor.
+Fig 2c shows that the shapes of the resource distance distributions for bees engaging in the two types of foraging trips are different.
+The distance distribution for the scout trips is akin close to that of an exponential distribution (Fig.
+2c), which is the nearest neighbour distance distribution for foragers operating in a one-dimensional environment (see methods).
+The distribution of the distances reported for recruit trips (Fig.
+2c) is a Rayleigh distribution which is the nearest-neighbour distribution in a two-dimensional environment [@Pyke1978] (see Methods).
 
-```{r, fig.cap="Figure 1.
-A.
-Outline of individual foraging.
-Foragers leave the hive on a random flight path (white lines) and continue until they encounter a resource (flowers in the figure).
-B.
-Outline of foraging with recruitment.
-Some foragers (scouts) continue to identify resources but others (recruits) sample available dances provided by scouts and other recruits.
-C.
-Log inverse frequency distribution of foraging distances reported by scouts.
-D.
-Log inverse frequency distribution of foraging distances reported by recruits.
-Here, inverse frequency distribution refers to probability of sampling a value greater than $X$."}
-knitr::include_graphics("../results/figures/simulation.png")
+```{r, fig.cap="Figure 2.
+**Simulating honey bee foraging**.
+In our simulation model with scouting only (a), foragers leave the hive on a search path (white lines) and continue until they encounter a resource (circles, colours indicate different resource quality).
+When foraging with recruitment (b) foragers continue to identify resources in scouting trips (white lines) and convey this information on the dance floor (brown disc) where foragers can sample dances reporting on scouting and recruiting trips and follow these directions (yellow lines).
+(c) Complementary cumulative frequencies of foraging distances reported from scouting and recruit trips.
+Note the difference in the shape of the distributions"}
+knitr::include_graphics("../results/figures/manuscript_figures/fig2.png")
 ```
 
-Scouts and recruits create different waggle run distributions on the dance-floor.
-These different distributions can be used to quantify the relative importance of collective foraging.
-Scouting foragers report a distribution of distances that reflects their probability of encountering a resource in the environment (see methods), whilst recruits report distances that reflect the distribution of the most profitable resources available (see methods).
-On the dance-floor, the distribution of distances is modified through the number of runs performed for a resource, which depends on the profitability of a resource, which honeybees achieve by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994A].
-By combining this profitability with the distributions identified in our simulations, we develop a novel method to describe the distribution of waggle runs reported on the dance-floors of real honeybee colonies as a superposition of scout and recruit distributions (Fig 2, see Methods).
+On the dance-floor, the number of waggle runs performed for a resource depends on its profitability [@Seeley1994].
+Honeybees achieve this by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994; @Seeley1994A].
+By combining this profitability bias with the distributions identified in our simulations, we can accurately describe the distribution of waggle runs reported on the dance-floors of real honeybee colonies as a superposition of scout and recruit distributions (Fig 3, see Methods for details)
 
-This method intrinsically captures honeybee foraging as collective decision making, as the foraging sites represented on the dance floor derive from a mixture of individual search and waggle dance information, captured in the proportions $p$ and $1-p$, respectively.
-Consequently, by setting $p = 1$, we can also derive a model based on the sole use of individual search (scouts only) (Fig 2).
+This description intrinsically captures honeybee foraging as collective decision making, whereby the foraging sites represented on the dance floor derive from a mixture of individual search and waggle dance information, modified by the profitability rule that biases recruitment towards closer or sweeter patches.
+The extent to which collective decision making is used is expressed in the proportion of scout dances, $p$, and the proportion of recruit dances, $1-p$.
+Scouting and recruiting are not fixed behavioural categories, because individual bees can engage in both over the course of their foraging lifetime, and foragers can dance on return from any successful trip irrespective of whether they were recruited to the forage site or found it individually [@Beekman2007].
+
+In a hypothetical colony that relied only on individual search to find foraging sites, all trips are "scout" trips, but as recruitment becomes more important, the proportion of “recruit” trips will increase.
+By setting $p = 1$, we obtain a model based on the sole use of individual search (scouts only) and by allowing the proportion of scout trips, $p$, to take on any value between 0 and 1, we can derive the extent to which foraging occurs collectively.
+(Fig 2).
 By fitting these two models to distribution of waggle run durations decoded from real honeybee colonies we are thus able, using model selection [@Burnham2002], to infer if and, by estimating the parameter $p$, to what extent honeybee colonies use waggle dance information when foraging.
 
-```{r, fig.cap="Figure 2: Model of honey bee foraging.
-Flowers advertised by scouts are distributed exponentially (a).
-These dances are advertised on the dance floor (b) in relation to their profitability, meaning sampling recruits are biased to the more profitable (and closer) resources (c).
-After visiting advertised resources, recruits also dance for them leading to higher recruitment (d) which overall skews the distribution of distances reported on the entire dance floor (e)."}
-knitr::include_graphics("../results/figures/foraging-model.png")
+```{r, fig.cap="Figure 3.
+**The rationale underlying the foraging model**.
+The distances of resources encountered by scouts are distributed exponentially (a).
+These dances are advertised on the dance floor (b).
+Dances for resources that are closer or higher in quality are repeated more often (c).
+As a consequence, dances for more profitable resource are over-represented and sampling foragers are biased to the more profitable resources (d).
+After successfully visiting advertised resources, recruits also dance for them leading to further amplification of this bias towards the most profitable resource in the vicinity of the hive (e).
+The distances of recruiting trips are than distributed through a Rayleigh distribution (f).
+The distances reported on the dance floor are a mixture of the scout and recruiting trips and can be calculated from the distance distributions of the scouting and recruiting trips, taking the reporting bias into account (see Methods for detail)."}
+knitr::include_graphics("../results/figures/manuscript_figures/fig3.png")
 ```
@@
 To evaluate the use of waggle dance information and individual search in honeybee colonies foraging in 'natural' landscapes, we analysed a pre-existing dataset of 3378 waggle dance observations from 20 observation hives, recorded between April-September 2017, (previously described in [@Samuelson2021]).
-Hives that contributed to this dataset had been situated at different locations in South East England (see methods, figure 3A) and visited every two weeks for a period of 24 weeks.
+Hives that contributed to this dataset had been situated at different locations in South East England (see Methods, figure 4A) and visited every two weeks for a period of 24 weeks.
 On each visit, two hours of continuous waggle dance data was recorded by training a camcorder onto the dance floor.
-The footage of the dances was decoded manually [@Couvillon2012; @Schurch2019] to extract waggle run durations.
-Using this data, for each site we fit both the collective and individual models and used model selection to determine which provided the better explanation of the data, and (if the collective model provided a better fit) to quantify the relative use of social information through estimating the parameter p.
-We calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004].
+The footage of the dances was decoded manually [@Couvillon2012; @Schurch2019; @Samuelson2021] to extract waggle run durations.
+Using this data, for each site we fit both the collective and individual models and used model selection to determine which provided the better explanation of the data, and (if the collective model provided a better fit) to quantify the relative use of social information through estimating the parameter $p$.
+In each case, We calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004].
 
@@
 3b).
-The sites STU and ZSL shown in Figs 3c-d are representative examples showing the model fits where the collective (STU) and individual (ZSL) models fit best, showing the overall quality of the model fit to the data.
+The sites STU and ZSL shown in Figs 3c-d are representative examples showing the model fits where the collective (STU) and individual (ZSL) models fit best.
+Note the closeness of the fit to the data, illustrating the overall quality of the model description.
 
-```{r, fig.cap="Figure 3.
-A.
-Location of study hives in Southern England coloured by which model, collective (black circles) or individual (red triangles) provided the best explanation of the data as indicated by lowest AIC score.
-Sites STU and ZSL (C & D) are marked.
-Inset plot shows the location of the sites within the UK; shaded area in the main plot indicates Greater London.
-B.
+[Bit on AIC, link to SI showing full tables and use of delta AIC]
+In all three sites where the individual model had a lower AIC than the collective model, the maximum likelihood estimate (MLE) was higher for the collective model.
+This indicates that the individual models do not explain the data any better than the collective model, and that the added points to the MLE comes from the increased complexity of the collective model.
+This is further demonstrated by the finding that the $p$ values estimated from the collective model do not settle close to 1, which would be expected if the individual model had a higher MLE, and thus suggests the collective model over fits in these sites (See supplementary material).
+
+```{r, fig.cap="Figure 4.
+**The honey bee foraging model fitted to data from 20 hives**.
+(a) Location of study hives in Southern England, shaded area in the main plot indicates Greater London.
+For 17 hives for the collective foraging model provided best explanation (black circles) or for 3 hives the individual search model provided the best explanation(red triangles) as indicated by lowest AIC score.
+(b).
 Distribution of goodness of fit confidence values for each model fit to waggle run durations from each site.
-P value is derived from a bootstrapped two-sided KS test comparing the fitted model predictions to the empirical data.
-The red dashed line marks the significance threshold of 0.05.
-Values greater than 0.05 indicate no statistically significant difference is observed between the model and the data, indicating the model provides a good fit to the data.
-C.
-Model fits to waggle run durations collected from the STU hive, for which the collective model (red line) did not outperform the individual model (blue line), along with histogram showing distribution.
-D.
-Model fits to waggle run durations collected from the ZSL hive, where the collective model (red line) outperformed the individual model (blue line), along with histogram showing distribution.
-Model fits show the compliment cumulative frequency distribution, giving the probability of sampling a value greater than or equal to x."}
-knitr::include_graphics("../results/figures/sites_model_plot.png")
+The p-value is derived from a bootstrapped two-sided KS test comparing the fitted model predictions to the empirical data, the red dashed line marks the significance threshold of 0.05.
+For values exceeding the threshold there is no statistically significant difference between the model and the data, indicating the model provides a good fit.
+(c) For the STU hive the individual model (blue line) provided a better fit than the collective foraging model (red line).
+(d).
+For the ZSL hive the collective foraging model (red line) outperformed the individual model (blue line).
+Panels show the compliment cumulative frequencies with binned frequency distributions as inset."}
+knitr::include_graphics("../results/figures/manuscript_figures/fig4.png")
 ```
 
-These results show that, whilst colony-level foraging is mostly comprised of a mixture of scout and recruit foraging trips, in some circumstances, colony foraging can be better described by individual foraging alone.
-This suggests that in some environments honey bees mainly engage in scouting to find new food sites rather than follow dances.
-Previous empirical studies have shown that individual foragers may flexibly alter their use of waggle dance information if its use is proving unproductive [@Gruter2011; @Price2019].
+Our results show that, whilst colony-level foraging is mostly comprised of a mixture of scout and recruit foraging trips, in some circumstances, colony foraging can be better described by individual foraging alone.
+Thus, in some environments, the majority of foraging trips involve scouting to find new food sites rather than recruitment through dances.
+Note that this does not imply that these bees do not engage in dance following, because bees regularly follow dances but choose not to visit the advertised site [@Gruter2011], although it has also been shown that bees may cease dance following if it is proving unproductive [@Gruter2011; @Price2019].
 Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individual's foraging behaviour over a landscape.
@@
 
-Further quantification of the use of waggle-dance recruitment within all colonies, as a proportion of all foraging trips, can be achieved by extracting the estimated proportion of scout trips (p) for each site.
+Further quantification of the use of waggle-dance recruitment within all colonies, as a proportion of all foraging trips, can be achieved by extracting the estimated proportion of scout trips, $p$, for each site.
 Since our sites varied in land-use characteristics and potentially thus forage distributions (although in this case, not by design for this study), we investigated whether these estimates might correlate with land-use.
@@
 We then performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances within agri-rural and urban environments.
-As we have a small sample size (20 sites; 10 urban and 10 agri-rural), this analysis was jackknifed to evaluate the robustness of our results to influential points (see methods, supplementary material section _._)
+As we have a small sample size (10 urban and 10 agri-rural sites), we used jackknife resampling to evaluate the robustness of our results to influential points (see methods, supplementary material)
 
-In the agri-rural environments one combination of land use types explained ~52% of the variation in the proportion of scouts (beta regression: $R^2 = 0.52$, $\phi$ = 3.7, p < 0.05, Fig.
+In the agri-rural environments the first principle component is a combination of land use types that explained ~52% of the variation in the proportion of scouts (beta regression: $R^2 = 0.52$, $\phi$ = 3.7, p < 0.05, Fig.
 4.
 A).
-This principal component correlates positively with arable land (29% of land coverage; Table 1) and negatively with built-up areas (17% of land coverage) ; it also correlates negatively with non-agricultrual unmanaged green space and water, but together these represent less than 3% of land-use (Fig.
-4.
-B).
-These land-use types maintain a significant correlation with PC1 over the jackknifed PLS (Fig.
+This principal component correlates positively with arable land (29% of land coverage; Table 1) and negatively with built-up areas (17% of land coverage); note that it also correlates negatively with non-agricultural unmanaged green space and water, but together these represent less than 3% of land-use (Fig.
+5b).
+These land-use types maintain a significant correlation with the first principle component over the jackknifed PLS (Fig.
 4.
@@
 As arable land increases whilst built-up areas decrease, the proportion of trips that are driven by individual search increases.
+Arable land in the UK is typically considered nutritionally poor for bees (note that oilseed rape fields were not included within this category), while there is evidence to suggest that the residential areas that were captured within the “built-up” category are forage-rich hotspots, typically supporting relatively high bee diversity and abundance within gardens [@Baldock2019].
 
@@
 
-rural_total_area %>%
-  arrange(desc(`% coverage`)) %>%
+urban_total_area$Environment <- "Urban"
+rural_total_area$Environment <- "Agri-rural"
+landuse_area <- rbind(urban_total_area, rural_total_area) %>%
+  select(Environment, `Land-use`, `% coverage`)
+
+landuse_area %>%
+  arrange(Environment, desc(`% coverage`)) %>%
   mutate(`% coverage` = signif(`% coverage`, 3)) %>%
@@
   set_caption(
-    "Percentage area covered for each land-use type in the agri-rural environment"
-  ) %>%
-  set_table_properties(layout = "autofit")
-```
-
-```{r, fig.cap="Figure 4.
-A.
-Estimated proportion of scouts for each site in the Agri-rural landscape against the first principle component derived from a Partial Least Squares analysis of land-use type.
-A beta regression shows the relationship (black line) between PC1 and the proportion of scouts, with 95% CI shown by the blue shaded area.
-B.
-Loadings for PC1 showing the correlations between PC1 and each land-use type (NAUMGS stands for Non-agricultural unmanaged green space).
-All points outside the shaded area significantly contribute to PC1.
-The colours of these land-use types are shown in the local environment of sites MEL, ROT and BEL (maps in bottom left) chosen as representative sites along the model gradient."}
-knitr::include_graphics("../results/figures/lu_prop_scouts_agrirural.png")
-```
-
-A similar effect was observed in the urban environment, where a single PC explained ~59% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.59, $\phi$ = 6.2, p < 0.05, Fig.
-5.
-A).
-This PC correlates positively with parks, allotments and cemeteries (8% of land cover; Table 2), whilst correlating negatively with sparse residential lands (35% of land cover) and railways, but these represent only 1% of land cover.
-(Fig.
-5.
-B).
-However, we have low confidence in this finding because the jackknifed PLS identifies CAD as a highly influential site, the removal of which causes sparse residential land to have a positive correlation with PC1 (Supplementary Material _._).
-
-```{r, urban-area-table}
-urban_total_area %>%
-  arrange(desc(`% coverage`)) %>%
-  mutate(`% coverage` = signif(`% coverage`, 3)) %>%
-  flextable() %>%
-  set_caption(
-    "Percentage area covered for each land-use type in the agri-rural environment"
+    "Percentage area covered for each land-use type in the agri-rural and urban environments in the sites studied."
   ) %>%
@@
 ```{r, fig.cap="Figure 5.
-A.
-Estimated proportion of scouts for each site in the urban landscape against the first principle component derived from a Partial Least Squares analysis of land-use type.
-A beta regression shows the relationship (black line) between PC1 and the proportion of scouts, with 95% CI shown by the blue shaded area.
-B.
-Loadings for PC1 showing the correlations between PC1 and each land-use type.
-All points outside the shaded area significantly contribute to PC1.
-The colours of these land-use types are shown in the local environment of sites HHS, MPA and CAD (maps in bottom left) chosen as representative sites along the model gradient."}
-knitr::include_graphics("../results/figures/lu_prop_scouts_urban.png")
+**Collective foraging correlates with land-use**.
+The proportion of scouts for each site against the first principal component derived from a Partial Least Squares analysis of land-use type.
+Beta regression shows the relationship (black line) between first principal component and the proportion of scouts, with 95% CI shown by the grey shaded area for agri-rural (a) and urban landscapes (c).
+The correlations between first principal component and each land-use type are shown for agri-rural (b) and urban landscapes (d).
+Correlations outside the shaded area significantly contribute to the first principal component.
+Colours correspond to the land use as shown in maps (circular insets) for selected sites.
+NAUMGS (resp.
+NAMGS) stands for non-agricultural unmanaged (resp.
+managed) green space."}
+knitr::include_graphics("../results/figures/manuscript_figures/fig5.png")
 ```
 
-Across both environments our results hint at a decrease in reliance on waggle dance recruitment as resources become harder to find and foraging trips become longer [@Samuelson2021].
-Arable land is likely to be nutritionally poor for pollinators (note that oilseed rape field and fruit crops were not included in this category), while residential areas -particularly gardens- are considered nutritional hotspots for bees [@Baldock2019].
-Similarly, in the urban environments sparse-residential coverage correlated negatively with the proportion of scout trips.
-However, we are cautious in this interpretation for two reasons.
-Firstly, because our study was designed as a proof-of-principle based on an existing dataset, we did not systematically assess the availability of forage within these sites.
-Secondly, although we saw a similar pattern in the urban data for sparse residential areas (35% of land-cover), which correlated negatively with the proportion of scout trips, this pattern reversed when a highly influential site (CAD; identified by the jacknifed PLS) was removed, and we also found parks, allotments and cemeteries (8% land cover), which could be expected to provide abundant forage for bees [@Baldock2019], to correlate positively with independent search.
+In in the urban environment, the first principal component explained ~59% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.59, $\phi$ = 6.2, p < 0.05, Fig.
+5c).
+Similarly to the agri-rural environments, this component correlates negatively with a significant residential category (sparse residential lands; 35% of land cover; Table 1), such that scout trips would appear to increase as forage availability decreases.
+However, we have limited confidence in this pattern because (a) this component correlates positively with parks, allotments and cemeteries (8% of land cover), which are typically also forage-rich [@Baldock2019] (b) the jackknifed PLS identified it being influenced strongly by a single site, the removal of which causes significant distribution to both of these relationships (Supplementary Material).
 
-The fitting of our model to waggle dance observations provides a novel tool to quantify the relative importance of social and private information in honeybees.
-Whilst recruitment is well known to occur through other mechanisms besides the waggle dance, such as through olfactory cues, the exceptionally close fit of our model to waggle run durations underlines the importance of the waggle dance in honey bee foraging.
-Further, by identifying the land-use combinations which most influence variation in the proportion of scouts, our analysis allows us to map the environment along the major axis of honeybee information use and visualise how land-use influences the use of the waggle dance.
+# Discussion
+
+Here, we have presented a model to quantify the use of waggle-dance communication in collective decision-making by honeybee colonies.
+Whilst recruitment is well known to occur through other mechanisms besides the waggle dance, such as through olfactory cues [@Arenas2007], the exceptionally close fit of our model to waggle run durations underlines the importance of the waggle dance in honey bee foraging.
+The fitting of our model to waggle dance observations provides a time and labour efficient methodology to quantify collective behaviour.
+Further, by identifying the land-use combinations which most influence variation in the proportion of scouts, our model provides a tool to map the environment along the major axis of honeybee information use and visualise how land-use influences the use of the waggle dance.
+
+Whilst our results hint at a decrease in reliance on waggle dance recruitment as resources become harder to find and foraging trips become longer [@Samuelson2021], they do not yet provide strong evidence for such a pattern.
+Firstly, within the urban sites, our findings did not consistently map to this pattern, and nor were they robust to sensitivity analysis.
+Secondly, because our study was designed as a proof-of-principle based on an existing dataset, we did not systematically choose sites based on forage availability, and instead inferred forage availability based on land-use at each site.
+The dance recordings in our dataset were collected over an extended period of five months, over which time forage availability in the landscape likely changed considerably and non-uniformly across sites.
+We thus cannot rule out that longer term effects not captured in the data used in this study, such as resource stability [@Schurch2014], may also have contributed to our estimates of waggle dance use.
 
 Our results within the agri-rural sites are in broad agreement with other studies which evaluated foraging performance with and without the dance [@Dornhaus2006; @Price2019].
-These results suggest that, even though collective foraging is not always beneficial, that in environments where high quality resources are present, but relatively scarce, collective foraging is being carried out by colonies, suggesting that under these conditions exchanging social information through the waggle dance confers benefits.
-These benefits of dancing appear to outweigh the costs associated with dancing and could explain why this remarkable behaviour is present in modern-day bees.
-However, the results for our urban sites do not produce a clear biological rational and present an all together more varied picture.
-An alternative hypothesis could be that the the dance is influenced by resource stability more than abundance and that longer term effects may contribute to the benefits of the dance [@Schurch2014].
-Whilst the limiting number of sites used in this study handicaps our ability to progress this debate here, as a proof of concept, the methods derived here provide a new toolkit for evaluating factors influencing waggle dance recruitment
+These results suggest that, even though collective foraging is not always beneficial, in environments where high quality resources are present, but relatively scarce, collective foraging is being carried out by colonies, suggesting that under these conditions exchanging social information through the waggle dance confers benefits.
+In both agri-rural and urban landscapes we have shown that the proportion of scouting trips change with land-use (illustrated in the change in land-use maps from left to right in Fig 5).
+We arrived at this conclusion through the analysis of waggle dance data, which provides a time and labour efficient methodology to quantify collective behaviour.
+Whilst the limiting number of sites used in this study handicaps our ability to progress this debate here, as a proof of concept, the methods derived here provide a new toolkit for evaluating factors influencing waggle dance recruitment.
 
@@
 Recently, technological advances have emerged which enable colony metrics to be collected faster, more accurately and over greater time spans than could be gathered by hand, allowing individuals to be tracked within colonies and theories of individual behaviour to be evaluated in more depth than could have been done previously [@Wild2021].
-By piecing together the behavioural response of individuals and combining these with landscape analyses, we have found a mathematical description of colony foraging which extends our ability to predict individual behaviour beyond the colony walls and in to the environment.
+By piecing together the behavioural response of individuals and combining these with landscape analyses, we have found an accurate mathematical description of colony foraging which extends our ability to quantify collective behaviour across environments.
 With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”.

### 2022-04-05 R/plotting.R
updated with laptop files, including presentation notes and similar stuff

@@ -69,12 +69,12 @@ make_base_plot <- function(x, logit = TRUE) {
   if (logit) {
     plt <- inverse_ccdf(x) %>%
       ggplot(aes(x = .data$sd, y = log(.data$prob))) +
-      geom_point() +
+      geom_point(shape = 1) +
       labs(x = "Waggle run duration (seconds)", y = "Ln cumulative probability")
   } else {
     plt <- inverse_ccdf(x) %>%
       ggplot(aes(x = .data$sd, y = .data$prob)) +
-      geom_point() +
+      geom_point(shape = 1) +
       labs(x = "Waggle run duration (seconds)", y = "Ln cumulative probability")
   }
   return(plt)
@@ -92,7 +92,7 @@ make_base_plot <- function(x, logit = TRUE) {
 #' along with the model fits to this data.
 #' @importFrom ggplot2 ggplot aes geom_histogram geom_line theme_set theme
 #' theme_classic element_text element_blank element_rect annotation_custom
-#' ggplotGrob
+#' ggplotGrob scale_colour_manual
 #' @importFrom tibble tibble
 #' @importFrom purrr map map_df
 #' @importFrom rlang .data
@@ -124,6 +124,9 @@ make_full_plot <- function(x, model_result_list,
         y = log(.data$cumul_ccdf),
         colour = .data$Model
       )
+    ) +
+    scale_colour_manual(
+      values = c("individual" = "#00AFBB", "collective" = "#FC4E07")
     ) +
     theme(
       legend.position = "none"

### 2022-04-05 analysis/agent_based_sim.Rmd
started converting other rmwrkdown files into having a nice htlm knitting
updated main figures with loading boxplots

@@
-# Honeybee foraging simulation
-## Joseph Palmer
+---
+title: Honeybee foraging simulation
+author: Joseph Palmer
+output:
+  html_notebook:
+    theme: yeti
+    toc: true
+    toc_float: true
 ---
@@
 Once you have ran the simulation you can run through this Rmd doc to generate the figures.
-To run the python code you will need python3 installed, along with the dependencies listed in **init.py?**.
+To run the python code you will need python 3 installed.
 
@@
 To be fit with the exponential function we first need to 'normalise' the data by subtracting the minimum foraging distance for all foraging distances.
-This is to ensure the exonential starts at 0 rather than stepping out to the minimum value and skewing the model fit.
+This is to ensure the exponential starts at 0 rather than stepping out to the minimum value and skewing the model fit.
 In plotting we take the values for the exponential ad add on the minimum value to each point to get the model fit data back in line with the actual data so they can be compared graphically.
@@
 
+```{r}
+# combined ccdfs
+combplot <- ggplot() +
+  geom_point(
+    data = scout_cdf, aes(x = sd, y = log(prob), color = "blue"), size = 0.2
+  ) +
+  geom_point(
+    data = recruit_cdf, aes(x = sd, y = log(prob), color = "red"), size = 0.2
+  ) +
+  labs(
+    x = "Simulated foraging distance (km)",
+    y = "ln Inverse cummulative frequency"
+  ) +
+  scale_color_manual(labels = c("Scout", "Recruit"), values = c("blue", "red")) +
+  theme(legend.position="bottom")
+
+ggsave(
+  plot = combplot,
+  filename = "Simulation/Results/combined_fit.png",
+  width = 70.1,
+  height = 81.79,
+  units = "mm",
+  dpi = 400
+)
+```
 From this analysis we observe that the distribution of foraging distances reported by scouts is better described by an exponential than the minimum of an exponential, whereas the converse is true for foraging distances reported by reruits.

### 2022-04-05 analysis/fit_models_durations.Rmd
started Suppmat, need to compute plots on laptop
updated with laptop files, including presentation notes and similar stuff

@@
 ```{r, make-plots, message = FALSE, warning = FALSE}
+# save the analysis results for all sites
+save(all_sites, file = "results/site_fit_results.Rdata")
 
@@
 
+ggsave(
+  plot = sites_model_plot,
+  filename = "results/figures/sites_model_plot.pdf",
+  width = 183,
+  height = 190,
+  units = "mm",
+  dpi = 300
+)
+
 model_sites_fits <- plot_grid(

### 2022-04-05 analysis/manuscript/supplementary_material.Rmd

---
title: "Waggle dance distributions quantify collective foraging in honey bee colonies: Supplementary material"
author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
output:
  word_document
always_allow_html: true
bibliography: paper-refs.bib
# csl: nature.csl
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(showtext)
showtext_auto()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(size = 18)
    )
)

library(dplyr)
library(kableExtra)
library(flextable)
library(tibble)
```

# Waggle dance model

# Waggle dance model fit

Fig 1. shows the models fit to all 20 sites.

Table 1. shows the results of the model fitting, including likelihood scores, parsimony measures -AIC, delta_AIC, rAIC (relative delta_AIC) and wAIC (Akaike weights)- and goodness of fit -Ks statistics (ks_statistic, ks_pvalue)- for all 20 sites.

```{r, model-fit-results}
set_flextable_defaults(
  big.mark = " ",
  font.size = 10, theme_fun = theme_vanilla,
  padding.bottom = 6,
  padding.top = 6,
  padding.left = 6,
  padding.right = 6
)

# load fit data
df <- readRDS(file = "../results/site_fit_results.Rda")

# calculate other AIC stats
df2 <- df %>%
  group_by(site) %>%
  mutate(
    delta_AIC = AIC - min(AIC),
    rAIC = exp(-0.5 * delta_AIC),
    wAIC = rAIC / sum(rAIC)
  ) %>%
  select(
    site, model, loglikelihood, AIC, delta_AIC, rAIC, wAIC,
    p, bs, br, as, ar, k, ks_statistic, ks_pvalue
  )

# format table
df2 %>%
  arrange(site) %>%
  mutate(across(where(is.numeric), round, 3)) %>%
  flextable() %>%
  set_caption(
    "Fit results for the fitting of the waggle dance model to each site."
  ) %>%
  set_table_properties(layout = "autofit")
```

### 2022-04-05 analysis/manuscript/supplementary_material.Rmd
ran code and updated results of fitting
reran to make plots with out histograms for easier viewing in the SM

@@
 library(ggplot2)
+library(cowplot)
 theme_set(
@@
 library(tibble)
+library(purrr)
+
+# load fitting results / stats
+load("results/site_fit_results.Rdata")
 ```
@@
 Fig 1. shows the models fit to all 20 sites.
+
+```{r, all-model-fits}
+
+plots <- map(all_sites, 2)
+plots <- plots %>%
+  map(
+    ~ {.x + theme(
+      axis.title.x = element_blank(),
+      axis.title.y = element_blank()
+    )}
+  )
+print(plots$BEL)
+
+plot_grid(plotlist = plots, nrow = 5)
+
+```
 
@@
 # load fit data
-df <- readRDS(file = "../results/site_fit_results.Rda")
+load("results/site_fit_results.Rdata")
+
+# group all site results together
+df <- map(all_sites, 1) %>%
+  bind_rows()
 

### 2022-04-05 analysis/manuscript/supplementary_material.Rmd
updated SI docs. Need to rerun as CAD looking odd

@@
 library(purrr)
+library(gridExtra)
 
@@
   map(
-    ~ {.x + theme(
+    ~ {
+      .x + theme(
         axis.title.x = element_blank(),
         axis.title.y = element_blank()
-    )}
       )
-print(plots$BEL)
+    }
+  )
 
-plot_grid(plotlist = plots, nrow = 5)
+allsites_plot <- plot_grid(
+  plotlist = plots, nrow = 5,
+  hjust = -0.1, label_x = 0.7,
+  labels = names(plots)
+)
+y.grob <- textGrob(
+  "cumulative frequency",
+  gp = gpar(),
+  rot = 90
+)
+x.grob <- textGrob(
+  "waggle run duration (s)",
+  gp = gpar()
+)
 
+# add to plot
+grid.arrange(arrangeGrob(allsites_plot, left = y.grob, bottom = x.grob))
 ```
@@
 )
-
-# load fit data
-load("results/site_fit_results.Rdata")
 

### 2022-04-05 analysis/fit_models_durations.Rmd
updated run through. Sometime the routine finds a weid pocket and the likelihood scores can stray out of bounds.

@@
 # save the analysis results for all sites
-save(all_sites, file = "results/site_fit_results.Rdata")
+# save(all_sites, file = "results/site_fit_results.Rdata")
 

### 2022-04-05 analysis/manuscript/paper.Rmd
updated SI and MS in text with AIC data.

@@
 (Fig 2).
-By fitting these two models to distribution of waggle run durations decoded from real honeybee colonies we are thus able, using model selection [@Burnham2002], to infer if and, by estimating the parameter $p$, to what extent honeybee colonies use waggle dance information when foraging.
+In fitting these two models to distribution of waggle run durations decoded from real honeybee colonies we are thus able, using model selection [@Burnham2002], to infer if and, by estimating the parameter $p$, to what extent honeybee colonies use waggle dance information when foraging.
 
@@
 3b).
-The sites STU and ZSL shown in Figs 3c-d are representative examples showing the model fits where the collective (STU) and individual (ZSL) models fit best.
+The sites STU and ZSL shown in Figs 3c-d are representative examples showing the model fits where the collective (ZSL) and individual (STU) models fit best.
 Note the closeness of the fit to the data, illustrating the overall quality of the model description.
 
-[Bit on AIC, link to SI showing full tables and use of delta AIC]
-In all three sites where the individual model had a lower AIC than the collective model, the maximum likelihood estimate (MLE) was higher for the collective model.
-This indicates that the individual models do not explain the data any better than the collective model, and that the added points to the MLE comes from the increased complexity of the collective model.
-This is further demonstrated by the finding that the $p$ values estimated from the collective model do not settle close to 1, which would be expected if the individual model had a higher MLE, and thus suggests the collective model over fits in these sites (See supplementary material).
+In all twenty sites the collective model provided a higher maximum likelihood estimate (MLE) than the individual model, however, this is anticipated as the individual model is a subset of the collective model.
+In 17 of the sites the collective model accounted for over 95% of Akaike weights compared to the individual models fit (supplementary table 1).
+In 3 of the sites the individual model explained a higher proportion of Akaike weight than the collective model (supplementary table 1) and so honeybee foraging behaviour in these colonies is better explained by the individual rather than collective foraging model.
 

### 2022-04-05 analysis/manuscript/supplementary_material.Rmd
updated SI and MS in text with AIC data.

--- 2022-04-05 analysis/manuscript/supplementary_material.Rmd
@@
 library(purrr)
+library(gridExtra)
+library(grid)
 
 # load fitting results / stats
-load("results/site_fit_results.Rdata")
+load("../results/site_fit_results.Rdata")
 ```
@@
 
-Fig 1. shows the models fit to all 20 sites.
-
-```{r, all-model-fits}
-
+```{r, all-model-fits, fig.width=10,fig.height=11, fig.cap="Supplementary Fig 1.
+log Compliment cumulative frequency distribution of waggle dance durations and fits of the collective (red line) and individual (blue line) models for all 20 sites."}
 plots <- map(all_sites, 2)
@@
   map(
-    ~ {.x + theme(
+    ~ {
+      .x + theme(
         axis.title.x = element_blank(),
         axis.title.y = element_blank()
-    )}
       )
-print(plots$BEL)
+    }
+  )
 
-plot_grid(plotlist = plots, nrow = 5)
+allsites_plot <- plot_grid(
+  plotlist = plots, nrow = 5,
+  hjust = -0.1, label_x = 0.7,
+  labels = names(plots)
+)
+y.grob <- textGrob(
+  "cumulative frequency",
+  gp = gpar(fontsize = 18),
+  rot = 90
+)
+x.grob <- textGrob(
+  "waggle run duration (s)",
+  gp = gpar(fontsize = 18)
+)
 
+# add to plot
+grid.arrange(arrangeGrob(allsites_plot, left = y.grob, bottom = x.grob))
 ```
@@
 
+# set up proper names
 # format table
@@
   mutate(across(where(is.numeric), round, 3)) %>%
+  rename(
+    MLE = loglikelihood,
+    "ΔAIC" = delta_AIC,
+    D = ks_statistic,
+    p_value = ks_pvalue
+  ) %>%
+  select(-k) %>%
   flextable() %>%
@@
   ) %>%
-  set_table_properties(layout = "autofit")
+  set_table_properties(layout = "autofit") %>%
+  fit_to_width(7.5)
 ```

### 2022-04-06 analysis/landuse_effect_jackknife.Rmd
set up SI for PLS additional info. Need to render plots.

@@
 
-ggplot(data = urban_jk_loadings, aes(
+plt <- ggplot(data = urban_jk_loadings, aes(
   x = landuse_type, y = loadings, fill = landuse_type
@@
   facet_wrap(~site_removed)
+ggsave("results/figures/SI_figures/urban_jk_individual_loadings.png")
+plt
 ```
@@
 
-ggplot(data = rural_jk_loadings, aes(
+plt <- ggplot(data = rural_jk_loadings, aes(
   x = landuse_type, y = loadings, fill = landuse_type
@@
   facet_wrap(~site_removed)
+ggsave("results/figures/SI_figures/agrirural_jk_individual_loadings.png")
+plt
 ```

### 2022-04-06 analysis/manuscript/supplementary_material.Rmd
set up SI for PLS additional info. Need to render plots.

@@
 ```
+
+# Jackknifed partial least squares analysis
+
+```{r, urban-pls-loadings, fig.width=10,fig.height=11, fig.cap="Loadings of PLS calculated for each site removed for the urban sites.
+Each plot shows the loadings of the first principle component with that site removed from the analysis, showing the individual points making up the overall box plot loadings in Fig 4d."}
+
+```

### 2022-04-06 analysis/landuse_effect_jackknife.Rmd
updated SI with plots. May need refining further

--- 2022-02-08 analysis/landuse_effect_jackknife.Rmd
@@
     legend.position = "None",
-    text = element_text()
+    text = element_text(size = 12),
+    strip.text.x = element_text(size = 18),
+    axis.title.x = element_text(size = 18)
   ) +
@@
   ) +
-  facet_wrap(~site_removed)
+  facet_wrap(~site_removed, ncol = 2)
+ggsave(
+  plot = plt,
+  filename = "results/figures/SI_figures/urban_jk_individual_loadings.png",
+  width = 120,
+  height = 200,
+  units = "mm",
+  dpi = 300
+)
+plt
 ```
@@
   geom_point(aes(colour = landuse_type)) +
+  geom_segment(aes(
+    x = landuse_type, xend = landuse_type,
+    y = 0, yend = loadings, colour = landuse_type
+  )) +
+  geom_hline(
+    yintercept = 0, color = "black"
+  ) +
+  geom_text(
+    data = rural_loading_text,
+    aes(x = landuse_type, y = result, label = landuse_type),
+    hjust = 0.5,
+    vjust = 0.5,
+    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
+  ) +
+  annotate(
+    "rect",
+    ymin = -rural_cuttoff,
+    ymax = rural_cuttoff,
+    xmin = 0, xmax = Inf,
+    alpha = 0.2
+  ) +
+  coord_flip() +
+  theme_classic() +
+  theme(
+    axis.line.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "None",
+    text = element_text(size = 15),
+    strip.text.x = element_text(size = 18),
+    axis.title.x = element_text(size = 18)
+  ) +
+  scale_fill_manual(
+    values = c(
+      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
+      "Non.agricultural.unmanaged.green.space" = "#228833",
+      "water" = "#4477AA", other = "#BBBBBB"
+    )
+  ) +
+  facet_wrap(~site_removed, ncol = 2)
+ggsave(
+  plot = plt,
+  filename = "results/figures/SI_figures/agrirural_jk_individual_loadings.png",
+  width = 120,
+  height = 220,
+  units = "mm",
+  dpi = 300
+)
+plt
+```
+
+There is a fair amount of variation in these plots, more than is shown with the boxplots for sure.
+Many landuse types move around in an out of being significant depending on what site is removed, but many, such as arable and built up areas, remain consistent in their position.
+
+## How does this compare with the proportion of land-coverage for each site?
+
+I now recreate the main plot but using the percentage of total area that covers as the argument to the size parameter. for the plot.
+
+```{r, rural_jk_loadings_landuse_area, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
+
+total_area <- rural_lu[, 3:ncol(rural_lu)] %>%
+  colSums()
+total_area <- tibble(landuse_type = names(total_area), area = total_area)
+total_area$area_pct <- (total_area$area / sum(total_area$area)) * 100
+
+rural_lu_loadings_with_area <- rural_jk_loadings %>%
+  left_join(total_area, by = "landuse_type") %>%
+  filter(site_removed == "None")
+
+ggplot(data = rural_lu_loadings_with_area, aes(
+  x = landuse_type, y = loadings, fill = landuse_type
+)) +
+  geom_point(aes(colour = landuse_type, size = area_pct)) +
   geom_segment(aes(
@@
   ) +
-  scale_fill_manual(
-    values = c(
-      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
-      "Non.agricultural.unmanaged.green.space" = "#228833",
-      "water" = "#4477AA", other = "#BBBBBB"
-    )
-  ) +
-  facet_wrap(~site_removed)
-```
-
-There is a fair amount of variation in these plots, more than is shown with the boxplots for sure.
-Many landuse types move around in an out of being significant depending on what site is removed, but many, such as arable and built up areas, remain consistent in their position.
-
-## How does this compare with the proportion of land-coverage for each site?
-
-I now recreate the main plot but using the percentage of total area that covers as the argument to the size parameter. for the plot.
-
-```{r, rural_jk_loadings_landuse_area, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
-
-total_area <- rural_lu[, 3:ncol(rural_lu)] %>%
-  colSums()
-total_area <- tibble(landuse_type = names(total_area), area = total_area)
-total_area$area_pct <- (total_area$area / sum(total_area$area)) * 100
-
-rural_lu_loadings_with_area <- rural_jk_loadings %>%
-  left_join(total_area, by = "landuse_type") %>%
-  filter(site_removed == "None")
-
-ggplot(data = rural_lu_loadings_with_area, aes(
-  x = landuse_type, y = loadings, fill = landuse_type
-)) +
-  geom_point(aes(colour = landuse_type, size = area_pct)) +
-  geom_segment(aes(
-    x = landuse_type, xend = landuse_type,
-    y = 0, yend = loadings, colour = landuse_type
-  )) +
-  geom_hline(
-    yintercept = 0, color = "black"
-  ) +
-  geom_text(
-    data = rural_loading_text,
-    aes(x = landuse_type, y = result, label = landuse_type),
-    hjust = 0.5,
-    vjust = 0.5,
-    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
-  ) +
-  annotate(
-    "rect",
-    ymin = -rural_cuttoff,
-    ymax = rural_cuttoff,
-    xmin = 0, xmax = Inf,
-    alpha = 0.2
-  ) +
-  coord_flip() +
-  theme_classic() +
-  theme(
-    axis.line.y = element_blank(),
-    axis.title.y = element_blank(),
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "None",
-    text = element_text()
-  ) +
   scale_colour_manual(

### 2022-04-06 analysis/manuscript/supplementary_material.Rmd
updated SI with plots. May need refining further

--- 2022-04-05 analysis/manuscript/supplementary_material.Rmd
@@
 ```
+
+# Jackknifed partial least squares analysis
+
+```{r, rural-pls-loadings, fig.width=10,fig.height=11, fig.cap="Loadings of PLS calculated for each site removed for the agri-rural sites.
+Each plot shows the loadings of the first principle component with that site removed from the analysis, showing the individual points making up the overall box plot loadings in Fig 4b."}
+knitr::include_graphics("../results/figures/SI_figures/agrirural_jk_individual_loadings.png")
+```
+
+```{r, urban-pls-loadings, fig.width=10,fig.height=11, fig.cap="Loadings of PLS calculated for each site removed for the urban sites.
+Each plot shows the loadings of the first principle component with that site removed from the analysis, showing the individual points making up the overall box plot loadings in Fig 4d."}
+knitr::include_graphics("../results/figures/SI_figures/urban_jk_individual_loadings.png")
+```

### 2022-04-08 analysis/manuscript/methods.Rmd
updated methods

@@
 
-We make the initial assumption that there are $n$ different resources distributed within an environment according to a spatial Poisson point process with rate $\lambda_{I}$ for resource $I$.
-As a scout travels a small distance $dx$ it will cover an area of $c_sdx$, with $c_s$ depending upon the width of the area scanned, the detection rate and degree to which a scout's path covers new ground.
-The probability density of the distance travelled after which the first resource is discovered $x$ follows an exponential distribution with a rate derived from the intensity of the flowers within the environment: $\lambda_s e^{-\lambda_s x}$, where $\lambda_s = c_s \sum_{I = 1}^{n} \lambda_I$.
+To describe the distribution of dance durations on the dance floor we formulated a generic model the duration of waggle dances.
+In the model resource patches are randomly placed in the environment.
+Foragers scout for these patches.
+The rationale of the model is illustrated in Fig.
+3: upon visiting a resource patch, foragers translate the profitability of a resource into the number of repeats of the dance.
+The number of repeats of the dance is a function of quality an distance.
+Recruits sample random dances and report the location of successful visits to resource patches on the dance floor.
+Through the feedback and over-representation of profitable resources on the dance floor recruits will converge to visiting the most profitable resource in vicinity of the hive.
+The distribution of dance durations is the superposition of scouting and recruiting trips.
 
-The number of repeats of a dance on the dance floor is dictated by profitability and is defined as the net energy efficiency per amount of energy spent: $\frac{Gain-Cost}{Cost}$.
-The gain is determined by the sucrose concentration of nectar from a flower, $q$, whilst the cost is determined by the distance, $x$.
-The energy gained, $G$, is therefore proportional to the quality $q$, whilst the cost increases linearly with distance.
-We can therefore refactor energy efficiency as $\frac{q}{1+\alpha x}-1$, where $\alpha$ represents the energy spend per distance relative to a fixed cost of travel.
-As bees should not be expected to dance for a negative profitability, we can derive a function to translate quality and distance into a measure of the number of dances on a dance floor: $\sigma(q,x) = [\frac{q}{1+\alpha x}-1]_+$, where the notation of $[x]_+ = max(x, 0)$.
-This function has the property that for a fixed $x$ there is a minimum quality $q = 1+\alpha x$ below which bees don't dance.
-The equivalent quality of two resources at distance $x1$ and $x2$ is $\frac{1+\alpha x2}{1+\alpha x1} = \frac{q2}{q1}$.
-The critical distance $\epsilon_{ij} = [\frac{1}{\alpha} \frac{qj - qi}{qi}+\frac{qj}{qi} x]_+$.
+As the resources patches are randomly placed in the environment, the distance after which the first resource is discovered approximately follows an exponential distribution (given by $\lambda e^{- \lambda x}$).
+Through the feedback mechanism that the dance floor provides, the colony can, collectively, locate the most profitable resource in its environment.
+For randomly placed resources in a two dimensional environment the distance to the nearest point is distributed according to a Rayleigh distribution (given by $2 \lambda \pi x e^{- \pi \lambda x^2}$) (Pyke 1978).
+Our simulation model shows that this describes the distances at which recruits visit resources well.
+Knowing the distance distributions of scout and recruit trips we then assume that the a proportion $p$ of all trips are scout trips.
+With this information we can specify the distributions of distances on the dance floor (see Supplementary Material for details).
 
-For a resource of type $I$ located at a distance $x$, the expected number of dances would be $c_s \lambda_I \sigma(qi, x) e^{-\lambda_s x}$.
-It follows that the expected total number of scout waggle dances on the dance floor is given by: $M_s = \sum_{I=1}^{n} \int_{0}^{\inf} \frac{c_s \lambda_I}{\lambda_s} \sigma(qi, x) \lambda_s e^{-\lambda_s x} dx$.
-The distribution of scout dances is: $\frac{\lambda_s e^{-\lambda_s x} \sum_{I=1}^{n} \frac{c_s \lambda_I}{\lambda_s} \sigma(qi, x)}{M_s}$.
+We implemented this in a full model which describes the distance distribution of the environment has $n$ different resource types (See Supplementary Material).
+However, in the full model the number of parameters increases with $2n$.
+Even if the number of resources is low, it turned out to be cumbersome to estimate the parameters and the model tends to overfit.
+To facilitate estimation of the parameter $p$ we therefore used a simplified model to estimate the fraction of scout trips.
 
-Using the same logic we derive what the distribution of recruits will look like.
-If two locations differ in resource quality and distance then a hive can prefer a further location if it is sufficiently better and a lesser quality resource if it is sufficiently nearer.
-For a site with quality $q_I$ and at distance $x$ from the hive, for a second location with quality $q_j$ this threshold distance $\epsilon_{ij}$ satisfies $\sigma(q_i, x) = \sigma(qj, \epsilon_{ij}$.
-Consequently, the critical distance is an implicit function of distance and it follows from implicit differentiation that $\frac{d\epsilon_{ij}}{dq_j} = - \frac{d\sigma{q_j,x}/dq{j}}{d\sigma{q,x}/dx} > 0$ and $\frac{d\epsilon_{ij}}{dq_I} = \frac{d\sigma{q_i,x}/dq{i}}{d\sigma{q,x}/dx} < 0$.
-If $qi = qj$ then $\epsilon_{ij}(x) = x$ and it follows that if $q_I > q_j$ then $\epsilon_{ij} < x$ and if $q_I < q_j$ then $\epsilon_{ij} > x$.
-As recruits get their information from scouts (as well as recruits) how well the scouts cover depends upon the number of scouts and how well their collective paths cover an area and the chance of discovery of each resource.
-To capture this we let the fraction of resources that can be discovered by a collection of scouts be $c_r$ and assume $0 \leq c_r \leq 1$.
-The probability that there is a not a food source discovered worth reporting on within a radius of $x$ $e^{-c_r \lambda_r \pi \epsilon_{ij}(x)^2}$, where $\lambda_r = \sum_I={1}^{n} \lambda_i$.
-This function also described the probability of having the nearest food source that the recruits will come out for discovered at least a distance of $x$ away.
+For the simplified the model we assumed that the the by assuming that the number of dances depends weakly on distance and there is a sizable quality differences between resources of a decent size and that there is a sizable intensity of the high quality resource.
+Foragers on scouting trips are more likely to report larger distances than foragers on recruiting trips.
+By linearising the function that translates the profitability into the number of waggle dance run for the largest distance and normalising, we arrive at simplified distribution for distances for scouting trips:
+
+$$ f_s(x)=M_s^{-1} a_s b_s e^{-b_s a_s x}  \left[ 1-a_s x\right]_+ ,$$
+
+where we used the shorthand ${x}_+=\max(x,0).$ The parameter $a_s^{-1}$ is the maximum reported distance by scouts, $a_s b_s$ is the intensity of resources found by scouts and $M_s=1-b_s^{-1} \left(1-e^{-b_s}\right)$ is factor that normalises the distribution.
+
+Recruit trips will be predominantly to high quality resource.
+Only if the nearest high quality patch is very far away will there be a more profitable patch of lesser quality available, and this happens only rarely if the intensity of the best quality resource is sizable.
+After linearising the function that translates the profitability into the number of waggle dance run for short distances and normalising the distribution of distances reported from recruit trips in the simplified model is:
+
+$$f_r(x)=M_r^{-1} 2 \pi a_r^2   b_r x e^{- \pi  b_r  (a_r x)^2}\left[1- a_r x\right]_+$$
+
+where $M_r={1-\frac{erf\left(  \sqrt{\pi b_r}\right)}{2 \sqrt{b_r }}}$ is the normalisation factor, the parameter $a_r$ is the rate with which dances repeats depends on distance for recruit trips and $a_r^2 b_r$ is the intensity of high quality resources reported by recruited foragers.
+
+The simplified distribution function is
+
+$$P(\underline x= x )=p f_s(x)+ (1-p) f_r(x), $$ which we used for parameter estimation and model fittings.
 
@@
 
-All models are fit using Maximum likelihood estimation [@Burnham2002] by summation of the log of the model function outlined in the methods section: model.
+All models are fit using Maximum likelihood estimation [@Burnham2002] by summation of the log of the simplified distribution function outlined in the methods section: model.
 The numerical optimisation routine is written in c++ and uses the Nelder-Mead simplex algorithm [@Nelder1965] implemented in the 'NLopt' library [@Johnson2020] and interfaced to R [@RCore2020] using 'Rcpp' [@Eddelbuettel2011].
 
-Model parsimony is assessed using Akaike information criterion (AIC) [@Aho2014] and the model with the lowest AIC score is deemed to be the most parsimonious.
+Model parsimony is assessed using Akaike information criterion (AIC) [@Aho2014] and Akaike weights. The model with the lowest AIC score is deemed to be the most parsimonious.
 
-Goodness of fit is assessed using the Kolmorgorov-Smirnov (KS) test [@Goldstein2004] and implemented in R using the ks.boot function of the package Matching in R [@Sekhon2011].
+Goodness of fit is assessed using the Kolmorgorov-Smirnov (KS) test [@Goldstein2004] and implemented in R using the ks.boot function of the package 'Matching' in R [@Sekhon2011].
 
@@
 
-A key consideration is that our response variable is a proportion and thus lies on the interval $[0, 1]$.
-This makes it unsuitable for a standard linear regression.
-One option would be to use a generalized linear regression with a logit link and quasibinomial family.
-This method is most suited for when the response variable is an integer proportion, where the response variable is a division of two integers, such as the proportion of males within a population.
-Our measure is in principle an integer proportion as we measure the proportion of scouts, however, this value is an estimate derived from our model and so is continuous on the interval $[0, 1]$.
-An alternative method would be to transform the proportions on to a continuos scale using the $arcsin \sqrt{x}$ transformation.
-The most appropriate method would be to use a beta regression which is designed specifically for proportional response variables.
+As our estimated proportion of scouts is continuos on the interval $[0, 1]$ we performed a beta regression using the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010].
+As the betareg package only works on the open interval $(0, 1)$ the data, $x$, was transformed using the following equation: $\frac{x(n-1)+0.5}{n}, as outlined in the betareg package documentation.
+After analysis the data was back transformed to the original values for the plots in Fig 5.
 
-We used the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010].
+For the jackknifed resampling we iterated through the each site and removed it from the pool of data and then ran the PLS as described above, recoding the loadings for each iteration (see Supplementary Material for loadings with each site removed).
+The PLS loadings for each land-use type are plotted as a box plot in Fig 5.
+to show the spread of these variable types.
 

### 2022-04-11 analysis/fit_models_durations.Rmd
updated to use 2017 data, updated ms and si and methods. Lots of stuff as we approach final edits for submission

@@
 alldata <- data %>%
+  filter(Year == 2017) %>% # remove pilot data conducted in 2016
   select(date, site, duration.seconds) %>%
@@
 br_bnds <- c(1.0e-6, 10)
-as_bnds <- c(1.0e-12, 10)
-ar_bnds <- c(1.0e-12, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 1.5)
 collective_bounds <- rbind(
@@
 # save results
-saveRDS(df, file = "results/site_fit_results.Rda")
+# saveRDS(df, file = "results/site_fit_results.Rda")
 

### 2022-04-11 analysis/landuse_effect_jackknife.Rmd
updated to use 2017 data, updated ms and si and methods. Lots of stuff as we approach final edits for submission

@@
   inner_join(urban_lu)
+
 urban_lu_long <- urban_lu %>%
@@
   select(site, p) %>%
-  inner_join(rural_lu)
+  inner_join(rural_lu) %>%
+  filter(site != "ROT") # remove ROT as fit is bad
+
 rural_lu_long <- rural_lu %>%
   pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
+
 ```
@@
     legend.position = "None",
-    text = element_text(size = 12),
+    text = element_text(size = 15),
     strip.text.x = element_text(size = 18),

### 2022-04-11 analysis/landuse_effect_prop_scouts.Rmd
updated to use 2017 data, updated ms and si and methods. Lots of stuff as we approach final edits for submission

@@
   select(site, p) %>%
-  inner_join(rural_lu)
+  inner_join(rural_lu)  %>%
+  filter(site != "ROT") # remove ROT as fit is bad
+
 rural_lu_long <- rural_lu %>%
@@
 ggplot(
-  data.frame(PC = seq(1, 10), varpercent = varpercent),
+  data.frame(PC = seq(1, 9), varpercent = varpercent),
   aes(x = PC, y = varpercent)
@@
 agrirural_plotdata$prop_scouts <- rural_lu$p
-ar_sortedpropscouts <- arrange(agrirural_plotdata, prop_scouts)
+ar_sortedpropscouts <- arrange(agrirural_plotdata, prop_scouts) %>%
+  filter(site != "ROT")
 
@@
 ar_marked <- ar_sortedpropscouts %>%
-  filter(site %in% c("MEL", "ROT", "BEL"))
+  filter(site %in% c("MEL", "SRA", "HER"))
 
@@
       values = c(
-        "arable" = "#CCBB44", "built-up area" = "#EE6677",
-        "Non-agricultural unmanaged green space" = "#228833",
-        "water" = "#4477AA", "pasture" = "#BBBBBB", "woodland" = "#BBBBBB",
+        "arable" = "#018B27", "built-up area" = "#ED6778",
+        "Non-agricultural unmanaged green space" = "#BEC330",
+        "water" = "#536CAB", "pasture" = "#BBBBBB", "woodland" = "#BBBBBB",
         "Non-agricultural managed green space" = "#BBBBBB", "fruit" = "#BBBBBB",
@@
     values = c(
-      "Arable" = "#CCBB44", "Built Up Area" = "#EE6677",
-      "NAUMGS" = "#228833",
-      "Water" = "#4477AA", Other = "#BBBBBB"
+      "Arable" = "#018B27", "Built Up Area" = "#ED6778",
+      "NAUMGS" = "#BEC330",
+      "Water" = "#536CAB", Other = "#BBBBBB"
     )
@@
         "continuous central" = "#BBBBBB", "dense residential" = "#BBBBBB",
-        "sparse residential" = "#EE6677", "amenity grassland" = "#BBBBBB",
-        "water" = "#BBBBBB", "parks allotments cemeteries" = "#228833",
-        "Railway" = "#CCBB44", "woodland" = "#BBBBBB"
+        "sparse residential" = "#ED6778", "amenity grassland" = "#BBBBBB",
+        "water" = "#BBBBBB", "parks allotments cemeteries" = "#018B27",
+        "Railway" = "#BEC330", "woodland" = "#BBBBBB"
       )
@@
       "Continuous Central" = "#BBBBBB", "Dense Residential" = "#BBBBBB",
-      "Sparse Residential" = "#EE6677", "Railway" = "#CCBB44",
-      "Parks Allotments Cemeteries" = "#228833",
+      "Sparse Residential" = "#ED6778", "Railway" = "#BEC330",
+      "Parks Allotments Cemeteries" = "#018B27",
       "Water" = "#BBBBBB", Other = "#BBBBBB"
@@
 
-joined <- bind_rows(rural_lu, urban_lu)
-joined[is.na(joined)] <- 0
-
-joined.pcdata <- joined[, 3:length(joined)]
-
-n <- length(joined$p)
-y <- (joined$p * (n - 1) + 0.5) / n
-joined.pcdata$p_transform <- y
-
-model <- plsRbeta(
-  p_transform ~ .,
-  data = joined.pcdata,
-  nt = ncol(joined.pcdata) - 1,
-  model = "pls-beta",
-  verbose = FALSE
-)
-
-print(model)
-
-plsdata <- data.frame(
-  prop.scouts = model$dataY,
-  model$tt
-)
-
-plsbetamodel <- betareg(
-  prop.scouts ~ Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4,
-  plsdata
-)
-
-summary(plsbetamodel)
-
-newdat <- data_frame(
-  Comp_.1 = seq(-50, 140, 0.1),
-  Comp_.2 = seq(-50, 140, 0.1),
-  Comp_.3 = seq(-50, 140, 0.1),
-  Comp_.4 = seq(-50, 140, 0.1)
-)
-
-newdat$fit <- predict(plsbetamodel, newdat, type = "response")
-
-plotdata <- plsdata
-plotdata$p <- joined$p
-
-plotdata$LU <- ""
-plotdata$LU[1:10] <- "Rural"
-plotdata$LU[11:20] <- "Urban"
-
-ggplot(
-  data = plotdata,
-  aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = p)
-) +
-  geom_point(aes(colour = LU)) +
-  geom_rug(aes(colour = LU)) +
-  scale_color_manual(values = c("Rural" = "black", "Urban" = "red")) +
-  geom_line(data = newdat, aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = fit)) +
-  xlim(min(-5.5), max(5.5))
-
-ggplot(
-  data = plotdata,
-  aes(x = LU, y = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4)
-) +
-  geom_boxplot(aes(group = LU))
-```
+# joined <- bind_rows(rural_lu, urban_lu)
+# joined[is.na(joined)] <- 0
+
+# joined.pcdata <- joined[, 3:length(joined)]
+
+# n <- length(joined$p)
+# y <- (joined$p * (n - 1) + 0.5) / n
+# joined.pcdata$p_transform <- y
+
+# model <- plsRbeta(
+#   p_transform ~ .,
+#   data = joined.pcdata,
+#   nt = ncol(joined.pcdata) - 1,
+#   model = "pls-beta",
+#   verbose = FALSE
+# )
+
+# print(model)
+
+# plsdata <- data.frame(
+#   prop.scouts = model$dataY,
+#   model$tt
+# )
+
+# plsbetamodel <- betareg(
+#   prop.scouts ~ Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4,
+#   plsdata
+# )
+
+# summary(plsbetamodel)
+
+# newdat <- data_frame(
+#   Comp_.1 = seq(-50, 140, 0.1),
+#   Comp_.2 = seq(-50, 140, 0.1),
+#   Comp_.3 = seq(-50, 140, 0.1),
+#   Comp_.4 = seq(-50, 140, 0.1)
+# )
+
+# newdat$fit <- predict(plsbetamodel, newdat, type = "response")
+
+# plotdata <- plsdata
+# plotdata$p <- joined$p
+
+# plotdata$LU <- ""
+# plotdata$LU[1:10] <- "Rural"
+# plotdata$LU[11:20] <- "Urban"
+
+# ggplot(
+#   data = plotdata,
+#   aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = p)
+# ) +
+#   geom_point(aes(colour = LU)) +
+#   geom_rug(aes(colour = LU)) +
+#   scale_color_manual(values = c("Rural" = "black", "Urban" = "red")) +
+#   geom_line(data = newdat, aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = fit)) +
+#   xlim(min(-5.5), max(5.5))
+
+# ggplot(
+#   data = plotdata,
+#   aes(x = LU, y = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4)
+# ) +
+#   geom_boxplot(aes(group = LU))
+```

### 2022-04-11 analysis/manuscript/methods.Rmd
updated to use 2017 data, updated ms and si and methods. Lots of stuff as we approach final edits for submission

@@
 These are placed on polar coordinates with a uniformly selected angle, $\theta$, between 0 and 2$\pi$ and a radial value, $\rho$, between 0 and $r$, determined from the square root of the uniform position values multiplied by $r$.
-These polar coordinates are converted to Cartesian coordinates using the formulas $x = \rho\cos(\theta)$ and $y = \rho\sin(\theta$).
+These polar coordinates are converted to Cartesian coordinates using the formulas $x = \rho\cos(\theta)$ and $y = \rho\sin(\theta)$.
 Each location is then assigned to an instance of a flower object along with a random quality of between 0 and 10.
@@
 We fit an exponential and minimum of an exponential distribution to both the distribution of foraging distances reported by the scout and recruit objects.
-Fitting was done by deriving the maximum likelihood estimate for each model fit on each data source through their analytical solutions: $\hat{\lambda} = \frac{1}/{\bar{x}}$, minimum of the exponential with a minimum foraging distance: $\hat{\lambda} = \frac{1}{\pi \bar{x^{2}}}$.
+Fitting was done by deriving the maximum likelihood estimate for each model fit on each data source through their analytical solutions: $\hat{\lambda} = \frac{1}{\bar{x}}$, minimum of the exponential with a minimum foraging distance: $\hat{\lambda} = \frac{1}{\pi \bar{x^{2}}}$.
 As the exponential assumes distributions start from 0 the data was transformed to start from 0 by subtracting the minimum foraging distance from all foraging distances ($x = x - min(x)$) before fitting.
@@
 Even if the number of resources is low, it turned out to be cumbersome to estimate the parameters and the model tends to overfit.
-To facilitate estimation of the parameter $p$ we therefore used a simplified model to estimate the fraction of scout trips.
+To facilitate estimation of the parameter $p$ we therefore used a simplified model to estimate the fraction of scout trips, where $m$ represents the lowest duration recorded, as there exists a minimum waggle run duration.
 
@@
 
-$$ f_s(x)=M_s^{-1} a_s b_s e^{-b_s a_s x}  \left[ 1-a_s x\right]_+ ,$$
+$$ f_s(x)=a_s \left[ M_s^{-1} b_s e^{-b_s a_s (x-m)}  \left[ 1-a_s x \right]_+ \right] $$
 
-where we used the shorthand ${x}_+=\max(x,0).$ The parameter $a_s^{-1}$ is the maximum reported distance by scouts, $a_s b_s$ is the intensity of resources found by scouts and $M_s=1-b_s^{-1} \left(1-e^{-b_s}\right)$ is factor that normalises the distribution.
+where we used the shorthand ${x}_+=\max(x,0)$ The parameter $a_s^{-1}$ is the maximum reported distance by scouts, $a_s b_s$ is the intensity of resources found by scouts and $M_s=(1-a_s m)-b_s^{-1} \left(1-e^{-b_s(1-a_s m)}\right)$ is factor that normalises the distribution.
 
@@
 
-$$f_r(x)=M_r^{-1} 2 \pi a_r^2   b_r x e^{- \pi  b_r  (a_r x)^2}\left[1- a_r x\right]_+$$
+$$f_r(x)=M_r^{-1} 2 \pi a_r^2   b_r x e^{- \pi  b_r  (a_r x)^2} \left[1- a_r x\right]_+$$
 
-where $M_r={1-\frac{erf\left(  \sqrt{\pi b_r}\right)}{2 \sqrt{b_r }}}$ is the normalisation factor, the parameter $a_r$ is the rate with which dances repeats depends on distance for recruit trips and $a_r^2 b_r$ is the intensity of high quality resources reported by recruited foragers.
+where
+
+$$M_r={(1-a_r m)e^{-\pi br(a_r m)^2} + \frac{erf(a_r \sqrt{\pi b_r m}) - erf(\sqrt{\pi b_r}))}{2 \sqrt{b_r}}}$$
+
+is the normalisation factor, the parameter $a_r$ is the rate with which dances repeats depends on distance for recruit trips and $a_r^2 b_r$ is the intensity of high quality resources reported by recruited foragers.
 
@@
 
-Goodness of fit is assessed using the Kolmorgorov-Smirnov (KS) test [@Goldstein2004] and implemented in R using the ks.boot function of the package 'Matching' in R [@Sekhon2011].
+Goodness of fit is assessed using the two-sample Kolmorgorov-Smirnov (KS) test [@Goldstein2004] and implemented in R using the ks.boot function of the package 'Matching' in R [@Sekhon2011].
 
@@
 
-As our estimated proportion of scouts is continuos on the interval $[0, 1]$ we performed a beta regression using the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010].
+Prior to conducing the PLS we removed any sites in which the models fit was significantly different to the waggle dance durations for that site.
+This resulted in one agri-rural site (ROT) and no urban sites being removed from the analysis.
+
+As our estimated proportion of scouts is continuos on the interval $[0, 1]$ we used the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010].
 As the betareg package only works on the open interval $(0, 1)$ the data, $x$, was transformed using the following equation: $\frac{x(n-1)+0.5}{n}, as outlined in the betareg package documentation.

### 2022-04-11 analysis/manuscript/paper.Rmd
updated to use 2017 data, updated ms and si and methods. Lots of stuff as we approach final edits for submission

@@
 
-For 17 out of 20 study hives, a model of collective foraging provided a better explanation of the data than one of individual foraging (Fig.
-3a).
-In all but one of these cases, the collective model had a good fit (using a Kolmogorov-Smirnov statistic, see methods) to the empirical waggle run durations, whereas for 9 of these sites, the distribution predicted by the scout model was significantly different to the observed data (Fig.
-3b).
-The sites STU and ZSL shown in Figs 3c-d are representative examples showing the model fits where the collective (ZSL) and individual (STU) models fit best.
+For 16 out of 20 study hives, a model of collective foraging provided a better explanation of the data than one of individual foraging ($wAIC > 0.95$, Fig.
+4a).
+In the other 4 sites, despite the collective model having the higher maximum likelihood estimate, the individual model had higher Akaike weights and so is judged to be more parsimonious (Supplementary Table 1).
+In all but one site, the collective model had a good fit (using a Kolmogorov-Smirnov statistic of $p > 0.05$, see methods) to the empirical waggle run durations (Fig.
+4b), whereas the scout model was significantly different to the observed data in 8 sites (Kolmogorov-Smirnov statistic $p < 0.05$, Fig.
+4b).
+The sites STU and ZSL shown in Figs 4c-d are representative examples showing the model fits where the collective (ZSL) and individual (STU) models fit best.
 Note the closeness of the fit to the data, illustrating the overall quality of the model description.
-
-In all twenty sites the collective model provided a higher maximum likelihood estimate (MLE) than the individual model, however, this is anticipated as the individual model is a subset of the collective model.
-In 17 of the sites the collective model accounted for over 95% of Akaike weights compared to the individual models fit (supplementary table 1).
-In 3 of the sites the individual model explained a higher proportion of Akaike weight than the collective model (supplementary table 1) and so honeybee foraging behaviour in these colonies is better explained by the individual rather than collective foraging model.
 
@@
 (a) Location of study hives in Southern England, shaded area in the main plot indicates Greater London.
-For 17 hives for the collective foraging model provided best explanation (black circles) or for 3 hives the individual search model provided the best explanation(red triangles) as indicated by lowest AIC score.
+For 16 hives for the collective foraging model provided best explanation (black circles) or for 4 hives the individual search model provided the best explanation(red triangles) as indicated by lowest AIC score.
 (b).
@@
 Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individual's foraging behaviour over a landscape.
-As our results show that individual foraging accurately describes colony foraging in three different sites, these findings support the idea that individual decision making can dominate colony foraging and demonstrates further evidence for flexible waggle dance use by honey bee colonies.
+As our results show that individual foraging accurately describes colony foraging in 11 different sites, and more parsimoniously than a model of collective foraging in 4 sites, these findings support the idea that individual decision making can dominate colony foraging and demonstrates further evidence for flexible waggle dance use by honey bee colonies.
 
@@
 We then performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances within agri-rural and urban environments.
-As we have a small sample size (10 urban and 10 agri-rural sites), we used jackknife resampling to evaluate the robustness of our results to influential points (see methods, supplementary material)
+As one of our sites (ROT) was not well fit by either of our models, this sites was removed from the PLS analysis.
+Due to our small sample size (10 urban and 9 agri-rural sites), we used jackknife resampling to evaluate the robustness of our results to influential points (see methods, supplementary material).
 
-In the agri-rural environments the first principle component is a combination of land use types that explained ~52% of the variation in the proportion of scouts (beta regression: $R^2 = 0.52$, $\phi$ = 3.7, p < 0.05, Fig.
-4.
-A).
-This principal component correlates positively with arable land (29% of land coverage; Table 1) and negatively with built-up areas (17% of land coverage); note that it also correlates negatively with non-agricultural unmanaged green space and water, but together these represent less than 3% of land-use (Fig.
+In the agri-rural environments the first principle component is a combination of land use types that explained ~73% of the variation in the proportion of scouts (beta regression: $R^2 = 0.73$, $\phi$ = 4.9, p < 0.05, Fig.
+5a).
+This principal component correlates positively with arable land (29% of land coverage; Table 1) and negatively with built-up areas (17% of land coverage); note that it also correlates negatively with non-agricultural unmanaged green space and water, but together these represent less than 3% of land-use (Table 1; Fig.
 5b).
 These land-use types maintain a significant correlation with the first principle component over the jackknifed PLS (Fig.
-4.
-B) indicating the results are robust.
+5b), with the exception of non-agricultural unmanaged green space which sits on the boarder, indicating the results are robust.
 As arable land increases whilst built-up areas decrease, the proportion of trips that are driven by individual search increases.
@@
 
-In in the urban environment, the first principal component explained ~59% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.59, $\phi$ = 6.2, p < 0.05, Fig.
+In in the urban environment, the first principal component explained ~73% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.73, $\phi$ = 10.4, p < 0.05, Fig.
 5c).
-Similarly to the agri-rural environments, this component correlates negatively with a significant residential category (sparse residential lands; 35% of land cover; Table 1), such that scout trips would appear to increase as forage availability decreases.
-However, we have limited confidence in this pattern because (a) this component correlates positively with parks, allotments and cemeteries (8% of land cover), which are typically also forage-rich [@Baldock2019] (b) the jackknifed PLS identified it being influenced strongly by a single site, the removal of which causes significant distribution to both of these relationships (Supplementary Material).
+This component correlates positively with parks, allotments and cemeteries (8% of land cover; Table 1) and negatively with railways, however this accounts for ~1% of land cover (Table 1).
+The dominant land by coverage , sparse residential (land coverage ~35%; Table 1), has a significant negative correlation with the first principle component for some sites removed over the jackknife sampling (Supplementary Fig.
+2.), however, it is overall not a significant contributor to the variance in the loadings (Fig.
+5c.).
+These results seem to be in contrast to what was observed in the urban landscapes as parks, allotments and cemeteries are typically forage-rich for honeybees [@Baldock2019].
+However, we have limited confidence in these results as (a) parks, allotments and cemeteries account for just 8% of land cover and (b) the jackknifed PLS identified substantial variation in the loadings identified for the first principle components (Supplementary Fig.
+3.).
 

### 2022-04-11 analysis/manuscript/supplementary_material.Rmd
updated to use 2017 data, updated ms and si and methods. Lots of stuff as we approach final edits for submission

@@
 ```{r, all-model-fits, fig.width=10,fig.height=11, fig.cap="Supplementary Fig 1.
-log Compliment cumulative frequency distribution of waggle dance durations and fits of the collective (red line) and individual (blue line) models for all 20 sites."}
+**Cumulative frequency distribution** of waggle dance durations and fits of the collective (red line) and individual (blue line) models for all 20 sites."}
 plots <- map(all_sites, 2)
@@
 
+write.csv(df2, "../results/modelfits_allsites.csv")
+
 # set up proper names
@@
 
-```{r, rural-pls-loadings, fig.width=10,fig.height=11, fig.cap="Loadings of PLS calculated for each site removed for the agri-rural sites.
+```{r, rural-pls-loadings, fig.width=10,fig.height=11, fig.cap="Supplementary Fig 2.
+**Loadings of PLS calculated for each site removed for the agri-rural sites**.
 Each plot shows the loadings of the first principle component with that site removed from the analysis, showing the individual points making up the overall box plot loadings in Fig 4b."}
@@
 
-```{r, urban-pls-loadings, fig.width=10,fig.height=11, fig.cap="Loadings of PLS calculated for each site removed for the urban sites.
+```{r, urban-pls-loadings, fig.width=10,fig.height=11, fig.cap="Supplementary Fig 3.
+**Loadings of PLS calculated for each site removed for the urban sites**.
 Each plot shows the loadings of the first principle component with that site removed from the analysis, showing the individual points making up the overall box plot loadings in Fig 4d."}

### 2022-04-11 R/plotting.R

@@ -70,12 +70,12 @@ make_base_plot <- function(x, logit = TRUE) {
     plt <- inverse_ccdf(x) %>%
       ggplot(aes(x = .data$sd, y = log(.data$prob))) +
       geom_point(shape = 1) +
-      labs(x = "Waggle run duration (seconds)", y = "Ln cumulative probability")
+      labs(x = "Waggle dance duration (s)", y = "cumulative frequency")
   } else {
     plt <- inverse_ccdf(x) %>%
       ggplot(aes(x = .data$sd, y = .data$prob)) +
       geom_point(shape = 1) +
-      labs(x = "Waggle run duration (seconds)", y = "Ln cumulative probability")
+      labs(x = "Waggle dance duration (s)", y = "cumulative frequency")
   }
   return(plt)
 }

### 2022-04-12 analysis/manuscript/methods.Rmd
made latest MS and methods updates. including adding in new figures and changes

@@
 
+This work was supported by the Biotechnology and Biological Sciences Research Council (BBSRC) through grant BB/M011178/1.
+
 # Conflict of interest
@@
 
+All simulation code was written in Python version 3.9 and uses the Pandas [@Mckinney2011] and Scipy [@Jones2001] packages.
+
 A circular environment is first created with radius $r = 2.5$.
-The number of flowers in the environment is generated as a random Poisson variable with rate equal to 5000 multiplied by the area of the environment.
+The number of resources in the environment is generated as a random Poisson variable with rate equal to 5000 multiplied by the area of the environment.
 These are placed on polar coordinates with a uniformly selected angle, $\theta$, between 0 and 2$\pi$ and a radial value, $\rho$, between 0 and $r$, determined from the square root of the uniform position values multiplied by $r$.
-These polar coordinates are converted to Cartesian coordinates using the formulas $x = \rho\cos(\theta)$ and $y = \rho\sin(\theta)$.
-Each location is then assigned to an instance of a flower object along with a random quality of between 0 and 10.
-This quality is combined with the distance of the flower to the centrally located hive to form a measure of how profitable the flower is (see model, equation?).
+These polar coordinates are converted to Cartesian coordinates.
+Each location is then assigned to an instance of a resource object along with a random quality of between 0 and 10.
+This quality is combined with the distance of the resource to the centrally located hive to form a measure of how profitable the resource is (see model, equation?).
 
-One-hundred honeybee objects are created, 20 of which 20 are scouts and the rest recruits.
+One-hundred honeybee objects are created, 20 of which 20 are on scouting trips and the rest recruited to follow scout dances.
 Scouts leave the environment following a random path through the environment generated by sampling a uniform random step length and angle.
@@
 Each straight line in the random path is converted to a rectangle with length equal to the path section length and a constant width of ~0.01 to represent an area the scout searches along that path.
-Of all the flowers contained in the boxes drawn from the scout's path, the one closest to the colony is selected as the flower the scout will report and will communicate its location if the quality of the flower is $ge$ 0.
-Flower communication is simulated by pooling together all the flower objects found.
-If no flowers are contained in the scout's path, they will not add any flowers to the scout pool and draw a new path in the next foraging iteration.
+Of all the resources contained in the boxes drawn from the scout's path, the one closest to the colony is selected as the resource patch that the scout will report and will communicate its location if the quality of the resource resource exceeds a minimum threshold.
+Communication is simulated by pooling together all the resource patches found.
+If no resources are contained in the scout's path, they will not add any resources to the scout pool and draw a new path in the next foraging iteration.
 
 Recruits represent honeybee objects which do not perform the searches the scouts do.
-Instead, they sample from the pool of flower objects reported by scouts.
-This sampling is done by selecting flowers with a probability which is skewed towards the profitability of the flower, meaning more profitable flowers have a greater chance of being selected by recruits.
-Recruits will then visit these flowers and in the next iteration will add their flower to the pool of scout dances.
-Consequently, the pool of dances represents flowers discovered by scouts and flowers exploited by recruits.
-When a flower is depleted,
-
-the flower is removed from the environment and so any foragers that were foraging on it would select a different flower from the dance floor.
+Instead, they sample from the pool of resource objects reported by scouts.
+This sampling is done by selecting resources with a probability which is skewed towards the profitability of the resource, meaning more profitable resources have a greater chance of being selected by recruits.
+Recruits will then visit these resources and in the next iteration will add their resource to the pool of scout dances.
+Consequently, the pool of dances represents resources discovered by scouts and resources exploited by recruits.
+When a resource is depleted, it is removed from the environment and so any foragers that were foraging on it would select a different resource from the dance floor.
 
@@
 
-All simulation code was written in Python version 3.9 and uses the Pandas [@Mckinney2011] and Scipy [@Jones2001] packages.
-
 # Model
@@
 To describe the distribution of dance durations on the dance floor we formulated a generic model the duration of waggle dances.
-In the model resource patches are randomly placed in the environment.
+In the model resource patches are assumed to be randomly placed in the environment.
 Foragers scout for these patches.
@@
 Even if the number of resources is low, it turned out to be cumbersome to estimate the parameters and the model tends to overfit.
-To facilitate estimation of the parameter $p$ we therefore used a simplified model to estimate the fraction of scout trips, where $m$ represents the lowest duration recorded, as there exists a minimum waggle run duration.
+To facilitate estimation of the parameter $p$ we therefore used a simplified model to estimate the fraction of scout trips, where $m$ represents the lowest duration considered, and here a minimum waggle run duration in the data set.
 
-For the simplified the model we assumed that the the by assuming that the number of dances depends weakly on distance and there is a sizable quality differences between resources of a decent size and that there is a sizable intensity of the high quality resource.
+For the simplified the model we assumed that the number of dances depends weakly on distance and there is a sizable quality differences between resources of a decent size and that there is a sizable intensity of the high quality resource.
 Foragers on scouting trips are more likely to report larger distances than foragers on recruiting trips.
-By linearising the function that translates the profitability into the number of waggle dance run for the largest distance and normalising, we arrive at simplified distribution for distances for scouting trips:
+By linearising the function that translates the profitability into the number of waggle dance run for the largest dance duration and normalising, we arrive at simplified distribution for dance durations for scouting trips:
 
@@
 
-where we used the shorthand ${x}_+=\max(x,0)$ The parameter $a_s^{-1}$ is the maximum reported distance by scouts, $a_s b_s$ is the intensity of resources found by scouts and $M_s=(1-a_s m)-b_s^{-1} \left(1-e^{-b_s(1-a_s m)}\right)$ is factor that normalises the distribution.
+where we used the shorthand ${x}_+=\max(x,0)$ The parameter $a_s^{-1}$ is the maximum dance duration by scouts, $a_s b_s$ is the intensity of resources found by scouts and $M_s=(1-a_s m)-b_s^{-1} \left(1-e^{-b_s(1-a_s m)}\right)$ is factor that normalises the distribution.
 
-Recruit trips will be predominantly to high quality resource.
+Recruit trips will be predominantly to high quality resources.
 Only if the nearest high quality patch is very far away will there be a more profitable patch of lesser quality available, and this happens only rarely if the intensity of the best quality resource is sizable.
-After linearising the function that translates the profitability into the number of waggle dance run for short distances and normalising the distribution of distances reported from recruit trips in the simplified model is:
+After linearising the function that translates the profitability into the number of waggle dance runs for short dance durations and normalising the distribution of dance durations reported from recruit trips in the simplified model is:
 
@@
 
-Model parsimony is assessed using Akaike information criterion (AIC) [@Aho2014] and Akaike weights. The model with the lowest AIC score is deemed to be the most parsimonious.
+The most parsimonious model is assessed using Akaike information criterion (AIC) [@Burnham2002; @Aho2014] and Akaike weights. The model with the lowest AIC score is deemed to be the most parsimonious.
 
@@
 to show the spread of these variable types.
+A loading was determined to be significantly correlating with the first principal component if contributed more than its expected variance.
 

### 2022-04-12 analysis/manuscript/paper.Rmd
made latest MS and methods updates. including adding in new figures and changes

@@
 title: "Waggle dance distributions quantify collective foraging in honey bee colonies"
-author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
+author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent A. A. Jansen"
 output:
@@
 
-Honeybee foraging is an extraordinary collective behaviour that is directed through the celebrated waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops [@Frisch1993; @Seeley1995].
+Honeybee foraging is an extraordinary collective behaviour that is directed through the waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops [@Frisch1993; @Seeley1995].
 Here, we present a mathematical model that quantifies the extent to which honeybee colonies behave collectively in different ecological environments by inferring the proportion of dances performed by bees that have explored individually, and those that have been recruited to food sources.
 By applying this methodology to waggle dance data from twenty hives, we show our model closely fits real-world honeybee foraging patterns and demonstrate that colonies vary their use of waggle dance information across different landscapes.
-Our methodology provides a novel tool to identify the ecological conditions in which honeybee colonies rely on dance communication, opening the door to large-scale experimental exploration of the selection pressures that may have driven the evolution of this remarkable collective behaviour.
+Our methodology provides a tool to identify the ecological conditions in which honeybee colonies rely on dance communication, opening the door to large-scale experimental exploration of the selection pressures that may have driven the evolution of this remarkable collective behaviour.
 
@@
 Accordingly, in many situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2004 @Dornhaus2006; @Gruter2013; @Price2019].
-This apparent variation in the use of collective behaviour is intriguing because it may provide clues as to those ecological selection pressures that were critical in the evolution of the dance communication system.
+This apparent variation in the use of collective behaviour is intriguing because it may provide clues as to the selection pressures that were critical in the evolution of the dance communication system.
 
@@
 
-[figure of waggle dance]
-
+```{r, fig.cap="Figure 1.
 **The honeybee waggle dance carries information about the location of a resource** *as a vector of distance (duration of the waggle run) and direction (angle of the dance relative to the sun) [@Frisch1993].Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions.
 Overall resource “quality” -the net energetic gain of a foraging trip- is provided through the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000].
-Although bees that follow dances do not specifically interpret this information on an individual level [@Seeley1991], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995], and provides the colony with a mechanism to select the most profitable resources in their environment [@Seeley1994A].*
-
-[end of figure of waggle dance]
+Although bees that follow dances do not specifically interpret this information on an individual level [@Seeley1991], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995], and provides the colony with a mechanism to select the most profitable resources in their environment [@Seeley1994A]."}
+knitr::include_graphics("../results/figures/manuscript_figures/fig1.png")
+```
 
@@
 Our method examines the distribution of waggle run durations reported on the dance-floor and infers from its shape the relative contribution of dances for resources identified by individual search, "scouting" versus dances for resources found by following dances, “recruitment” [@Seeley1995].
-Previously, such stips have only been distinguishable by documenting foragers' individual search history, which requires tracking individual bees and is time -and labour- intensive [@Seeley1995].
+Previously, the relative amount of scouting and recruitment trips have only been quantifiable by documenting foragers' individual search history, which requires tracking individual bees and is time -and labour- intensive [@Seeley1995].
 Our methods provide an efficient means of quantifying collective behaviour with no requirement to manipulate hive orientations, assay foraging efficiency or document individuals' previous experience.
@@
 
-To reveal how the patterns of waggle dance-encoded foraging distances might differ under scenarios where colonies rely on individual search vs recruitment, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
+To reveal how the patterns of waggle dance-encoded foraging distances might differ under scenarios where colonies rely on individual search versus recruitment, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
 Foragers could locate these under two different strategies: either acting as a scout and locating resources themselves, or following a recruit strategy and locating resources by following a random dance from the dance floor [@Seeley1992] (Fig.
-2, details in Methods1).
+2a,b, details in Methods).
 As it is known in the simulation which individuals in the hive forage under what strategy, we can compare the distributions of foraging distances reported on the dance floor.
@@
 
-In a hypothetical colony that relied only on individual search to find foraging sites, all trips are "scout" trips, but as recruitment becomes more important, the proportion of “recruit” trips will increase.
+In a hypothetical colony that relies only on individual search to find foraging sites, all trips are "scout" trips, but as recruitment becomes more important, the proportion of “recruit” trips will increase.
 By setting $p = 1$, we obtain a model based on the sole use of individual search (scouts only) and by allowing the proportion of scout trips, $p$, to take on any value between 0 and 1, we can derive the extent to which foraging occurs collectively.
@@
 ```{r, fig.cap="Figure 3.
-**The rationale underlying the foraging model**.
+**The rationale of the foraging model**.
 The distances of resources encountered by scouts are distributed exponentially (a).
@@
 
-To evaluate the use of waggle dance information and individual search in honeybee colonies foraging in 'natural' landscapes, we analysed a pre-existing dataset of 3378 waggle dance observations from 20 observation hives, recorded between April-September 2017, (previously described in [@Samuelson2021]).
+To evaluate the use of waggle dance information and individual search in honeybee colonies foraging in 'natural' landscapes, we analysed a pre-existing dataset of 2827 waggle dance observations from 20 observation hives, recorded between April-September 2017, (previously described in [@Samuelson2021]).
 Hives that contributed to this dataset had been situated at different locations in South East England (see Methods, figure 4A) and visited every two weeks for a period of 24 weeks.
@@
 Using this data, for each site we fit both the collective and individual models and used model selection to determine which provided the better explanation of the data, and (if the collective model provided a better fit) to quantify the relative use of social information through estimating the parameter $p$.
-In each case, We calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004].
+In each case, We calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004; @Clauset2009].
 
-For 16 out of 20 study hives, a model of collective foraging provided a better explanation of the data than one of individual foraging ($wAIC > 0.95$, Fig.
+For 16 out of 20 study hives, a model of collective foraging provided a better description of the data than one of individual foraging (Fig.
 4a).
-In the other 4 sites, despite the collective model having the higher maximum likelihood estimate, the individual model had higher Akaike weights and so is judged to be more parsimonious (Supplementary Table 1).
+In the other 4 sites, despite the collective model having the higher maximum likelihood, the individual model had a higher AIC value and so is more parsimonious (Supplementary Table 1).
 In all but one site, the collective model had a good fit (using a Kolmogorov-Smirnov statistic of $p > 0.05$, see methods) to the empirical waggle run durations (Fig.
@@
 4b).
-The sites STU and ZSL shown in Figs 4c-d are representative examples showing the model fits where the collective (ZSL) and individual (STU) models fit best.
+The sites shown in Figs 4c-d are representative examples showing the model fits where the individual (Fig.
+4c) and collective (Fig.
+4d) models fit best.
 Note the closeness of the fit to the data, illustrating the overall quality of the model description.
@@
 For values exceeding the threshold there is no statistically significant difference between the model and the data, indicating the model provides a good fit.
-(c) For the STU hive the individual model (blue line) provided a better fit than the collective foraging model (red line).
-(d).
-For the ZSL hive the collective foraging model (red line) outperformed the individual model (blue line).
+For the hive in (c) the individual model (blue line) provided a better fit than the collective foraging model (red line).
+For the hive in (d) the collective foraging model (red line) provided a better fit than the individual model (blue line).
 Panels show the compliment cumulative frequencies with binned frequency distributions as inset."}
@@
 We then performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances within agri-rural and urban environments.
-As one of our sites (ROT) was not well fit by either of our models, this sites was removed from the PLS analysis.
-Due to our small sample size (10 urban and 9 agri-rural sites), we used jackknife resampling to evaluate the robustness of our results to influential points (see methods, supplementary material).
+As for one of our sites neither model provided a plausible description, this site was removed from the PLS analysis.
+Due to our small sample size (10 urban and 9 agri-rural sites), we used jackknife resampling to evaluate the robustness of our results to influential points (see Methods, Supplementary Material).
 
-In the agri-rural environments the first principle component is a combination of land use types that explained ~73% of the variation in the proportion of scouts (beta regression: $R^2 = 0.73$, $\phi$ = 4.9, p < 0.05, Fig.
+In the agri-rural environments the first principal component is a combination of land use types that explained ~73% of the variation in the proportion of scouts (beta regression: $R^2 = 0.73$, $\phi$ = 4.9, p < 0.05, Fig.
 5a).
@@
 5b).
-These land-use types maintain a significant correlation with the first principle component over the jackknifed PLS (Fig.
+These land-use types maintain a significant correlation with the first principal component over the jackknifed PLS (Fig.
 5b), with the exception of non-agricultural unmanaged green space which sits on the boarder, indicating the results are robust.
@@
 
-In in the urban environment, the first principal component explained ~73% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.73, $\phi$ = 10.4, p < 0.05, Fig.
+In the urban environment, the first principal component explained ~73% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.73, $\phi$ = 10.4, p < 0.05, Fig.
 5c).
 This component correlates positively with parks, allotments and cemeteries (8% of land cover; Table 1) and negatively with railways, however this accounts for ~1% of land cover (Table 1).
-The dominant land by coverage , sparse residential (land coverage ~35%; Table 1), has a significant negative correlation with the first principle component for some sites removed over the jackknife sampling (Supplementary Fig.
+The dominant land-use type by coverage, sparse residential (land coverage ~35%; Table 1), has a significant negative correlation with the first principal component for some sites removed over the jackknife sampling (Supplementary Fig.
 2.), however, it is overall not a significant contributor to the variance in the loadings (Fig.
 5c.).
-These results seem to be in contrast to what was observed in the urban landscapes as parks, allotments and cemeteries are typically forage-rich for honeybees [@Baldock2019].
-However, we have limited confidence in these results as (a) parks, allotments and cemeteries account for just 8% of land cover and (b) the jackknifed PLS identified substantial variation in the loadings identified for the first principle components (Supplementary Fig.
+Parks, allotments and cemeteries are typically forage-rich for honeybees [@Baldock2019], a result which was not replicated here.
+However, note that the amount of land cover for this component is limited and the jackknifed PLS identified substantial variation in the loadings identified for the first principal components (Supplementary Fig.
 3.).
@@
 
-Whilst our results hint at a decrease in reliance on waggle dance recruitment as resources become harder to find and foraging trips become longer [@Samuelson2021], they do not yet provide strong evidence for such a pattern.
-Firstly, within the urban sites, our findings did not consistently map to this pattern, and nor were they robust to sensitivity analysis.
-Secondly, because our study was designed as a proof-of-principle based on an existing dataset, we did not systematically choose sites based on forage availability, and instead inferred forage availability based on land-use at each site.
+Whilst our results show a decrease in reliance on waggle dance recruitment as resources become harder to find and foraging trips become longer [@Samuelson2021], they do not yet provide conclusive evidence for such a pattern.
+Our study was designed as a proof-of-principle based on an existing dataset, we did not systematically choose sites based on forage availability, and instead inferred forage availability based on land-use at each site.
 The dance recordings in our dataset were collected over an extended period of five months, over which time forage availability in the landscape likely changed considerably and non-uniformly across sites.
@@
 
-Our results within the agri-rural sites are in broad agreement with other studies which evaluated foraging performance with and without the dance [@Dornhaus2006; @Price2019].
+Our results within the agri-rural sites are in agreement with other studies which evaluated foraging performance with and without the dance [@Dornhaus2006; @Price2019].
 These results suggest that, even though collective foraging is not always beneficial, in environments where high quality resources are present, but relatively scarce, collective foraging is being carried out by colonies, suggesting that under these conditions exchanging social information through the waggle dance confers benefits.
-In both agri-rural and urban landscapes we have shown that the proportion of scouting trips change with land-use (illustrated in the change in land-use maps from left to right in Fig 5).
-We arrived at this conclusion through the analysis of waggle dance data, which provides a time and labour efficient methodology to quantify collective behaviour.
-Whilst the limiting number of sites used in this study handicaps our ability to progress this debate here, as a proof of concept, the methods derived here provide a new toolkit for evaluating factors influencing waggle dance recruitment.
+In both agri-rural and urban landscapes we have shown that the proportion of scouting trips change with land-use (illustrated in the change in land-use shown in the maps from left to right in Fig 5).
+We arrived at this conclusion through the analysis of waggle dance data, using a mathematical model to fit the waggle dance durations against.
+This method provides a time and labour efficient methodology to quantify collective behaviour.
+We analysed an existing data set and thus provide a proof of concept for how this new toolkit can be used to evaluate factors influencing waggle dance recruitment.
 
@@
 With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”.
-Our methodology thus provides a means of analysing such large data sets to inform the debate about the importance of social information, as well as providing useful colony metrics of foraging activity.
+Our methodology thus provides a means of analysing such large data sets to inform the debate about the importance of collective decision making, as well as providing useful colony metrics of foraging activity.
 

### 2022-04-29 analysis/manuscript/methods.Rmd
updated the MS with Vincent changes for Rich. Need to merge in Rich + Elli's changes and update for final draft.

@@
 title: "Forage site distributions quantify collective foraging in honey bee colonies: Methods"
-author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
+# author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
 output:
@@
 bibliography: paper-refs.bib
+csl: nature.csl
 ---

### 2022-04-29 analysis/manuscript/paper.Rmd
updated the MS with Vincent changes for Rich. Need to merge in Rich + Elli's changes and update for final draft.

@@
 title: "Waggle dance distributions quantify collective foraging in honey bee colonies"
-author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent A. A. Jansen"
+# author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent A. A. Jansen"
 output:
@@
 bibliography: paper-refs.bib
-# csl: nature.csl
+csl: nature.csl
 ---
@@
 ```{r, fig.cap="Figure 1.
-**The honeybee waggle dance carries information about the location of a resource** *as a vector of distance (duration of the waggle run) and direction (angle of the dance relative to the sun) [@Frisch1993].Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions.
+**The honeybee waggle dance carries information about the location of a resource.** The duration of the waggle run indicates the distance to the resource and the angle of the dance relative to the vertical indicates the direction of the resource (circle in right panel), relative to the direction of the sun [@Frisch1993].
+Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions.
 Overall resource “quality” -the net energetic gain of a foraging trip- is provided through the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000].

### 2022-05-03 analysis/fit_models_durations.Rmd
updated plots and layouts per Rich comments

@@
 
-run_wagglefit_analysis <- function(target_site, data, collective_bounds, individual_bounds, subplot_coords) {
+run_wagglefit_analysis <- function(target_site, data, collective_bounds, individual_bounds, subplot_coords, xlims = NULL) {
 
@@
     ),
-    subplot_coords = subplot_coords
+    subplot_coords = subplot_coords,
+    xlims = xlims
   )
@@
 # set coordinates for histogram subplot
-subplot_coords <- c(0.1, 4, -6.5, -2.2)
-
-all_sites[[target_site]] <- run_wagglefit_analysis(
-  target_site, data, collective_bounds, individual_bounds, subplot_coords
+subplot_coords <- c(0.1, 4, -6.8, -2.5)
+
+all_sites[[target_site]] <- run_wagglefit_analysis(
+  target_site, data, collective_bounds, individual_bounds, subplot_coords,
+  xlims = c(0, 6)
 )
@@
 # saveRDS(df, file = "results/site_fit_results.Rda")
+
+# all_sites$BFI$fit <- all_sites$BFI$fit +
+#   ylim(0, 6) +
+#   scale_x_continuous(breaks = seq(0, 6, by = 2))
 
@@
 ggsave(
-  plot = all_sites$ZSL$fit,
-  filename = "results/figures/ZSL.png",
+  plot = all_sites$BFI$fit,
+  filename = "results/figures/BUR.png",
   width = 90,
@@
   left_join(map_data, on = "site") %>%
-  mutate(col = ifelse(site %in% c("STU", "ZSL"), "1", "0"))
+  mutate(col = ifelse(site %in% c("STU", "BUR"), "1", "0"))
 
@@
 # Extract selected sites for figure
-selected_sites <- filter(map_data, site %in% c("STU", "ZSL")) %>%
+selected_sites <- filter(map_data, site %in% c("STU", "BUR")) %>%
   mutate(
-    label = ifelse(site == "STU", "C (STU)", "D (ZSL)")
+    label = ifelse(site == "STU", "c", "d")
   )
@@
     nudge_x = c(0, -0.07), nudge_y = c(0.05, 0.05), size = 8
+  ) +
+  geom_point(
+    data = selected_sites, aes(x = lon, y = lat), size = 1.5, shape = 4
   ) +
@@
     ggplotGrob(base),
-    xmin = -3.35,
-    xmax = 2,
-    ymin = 51.05,
-    ymax = 51.35
+    xmin = -3.45,
+    xmax = 2.2,
+    ymin = 51.01,
+    ymax = 51.45
   )
@@
 sites_model_plot <- plot_grid(
-  merge_plot_2, ks_plot_dist, all_sites$STU$fit, all_sites$ZSL$fit,
+  merge_plot_2, ks_plot_dist, all_sites$STU$fit, all_sites$BUR$fit,
   labels = c("A", "B", "C", "D"), label_size = 24
@@
 stu_zsl_fit <- plot_grid(
-  all_sites$STU$fit, all_sites$ZSL$fit,
+  all_sites$STU$fit, all_sites$BUR$fit,
   labels = c("A", "B"), label_size = 24
@@
 sites_model_plot <- plot_grid(
-  all_sites$STU$fit, all_sites$ZSL$fit, aic_plot, ks_plot_dist,
+  all_sites$STU$fit, all_sites$BUR$fit, aic_plot, ks_plot_dist,
   labels = c("A", "B", "C", "D"), label_size = 42

### 2022-05-03 analysis/manuscript/paper.Rmd
updated plots and layouts per Rich comments

@@
   arrange(Environment, desc(`% coverage`)) %>%
-  mutate(`% coverage` = signif(`% coverage`, 3)) %>%
+  mutate(`% coverage` = round(`% coverage`, digits = 1)) %>%
   flextable() %>%

### 2022-05-03 R/plotting.R

@@ -88,6 +88,7 @@ make_base_plot <- function(x, logit = TRUE) {
 #' estimates ($est) and the data name ($data_name).
 #' @param logit Bool To log the data or not. Defaults to TRUE
 #' @param subplot_coords doubleArray Coordinates for location of inner histogram
+#' @param xlims array limits for x axis
 #' @return ggplot plot of cumulative probability of foraging distances in Km
 #' along with the model fits to this data.
 #' @importFrom ggplot2 ggplot aes geom_histogram geom_line theme_set theme
@@ -100,7 +101,8 @@ make_base_plot <- function(x, logit = TRUE) {
 #'
 make_full_plot <- function(x, model_result_list,
                            logit = TRUE,
-                           subplot_coords = c(0.5, 2.5, -7.8, -2.5)) {
+                           subplot_coords = c(0.5, 2.5, -7.8, -2.5),
+                           xlims = NULL) {
   cdf_data <- map(
     model_result_list,
     ~ {
@@ -145,6 +147,17 @@ make_full_plot <- function(x, model_result_list,
       panel.background = element_rect(fill = "transparent"),
       plot.background = element_rect(fill = "transparent", colour = NA)
     )
+
+  if(!is.null(xlims)) {
+    plt <- plt + scale_x_continuous(
+      breaks = seq(xlims[1], xlims[2], 2),
+      limits = xlims
+    )
+    histoplot <- histoplot + scale_x_continuous(
+      breaks = seq(xlims[1], xlims[2], 2),
+      limits = xlims
+    )
+  }
 
   full_plot <- plt + annotation_custom(
     ggplotGrob(histoplot),

### 2022-05-03 analysis/manuscript/methods.Rmd
updated final MS edits. Just last meeting to go

@@
 Through the feedback mechanism that the dance floor provides, the colony can, collectively, locate the most profitable resource in its environment.
-For randomly placed resources in a two dimensional environment the distance to the nearest point is distributed according to a Rayleigh distribution (given by $2 \lambda \pi x e^{- \pi \lambda x^2}$) (Pyke 1978).
+For randomly placed resources in a two dimensional environment the distance to the nearest point is distributed according to a Rayleigh distribution (given by $2 \lambda \pi x e^{- \pi \lambda x^2}$)) [@Pyke1978].
 Our simulation model shows that this describes the distances at which recruits visit resources well.
@@
 
-For the simplified the model we assumed that the number of dances depends weakly on distance and there is a sizable quality differences between resources of a decent size and that there is a sizable intensity of the high quality resource.
+For the simplified the model we assumed that the number of dances depends weakly on distance and there is a sizable quality differences between resources of a non-negligible size and that there is a sizable intensity of the high quality resource (See Supplementary Material for detailed derivation).
 Foragers on scouting trips are more likely to report larger distances than foragers on recruiting trips.

### 2022-05-03 analysis/manuscript/paper.Rmd
updated final MS edits. Just last meeting to go

@@
 
-Honeybee foraging is an extraordinary collective behaviour that is directed through the waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops [@Frisch1993; @Seeley1995].
-Here, we present a mathematical model that quantifies the extent to which honeybee colonies behave collectively in different ecological environments by inferring the proportion of dances performed by bees that have explored individually, and those that have been recruited to food sources.
-By applying this methodology to waggle dance data from twenty hives, we show our model closely fits real-world honeybee foraging patterns and demonstrate that colonies vary their use of waggle dance information across different landscapes.
+Honeybee foraging is an extraordinary collective behaviour that is directed through the waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops.
+Here, we present a mathematical framework to explain how the intrinsic process of the waggle dance can allow a colony to exploit different landscape types.
+From waggle dance observations, we infer the proportion of dances performed by bees that forage individually, and those that forage collectively.
+By applying this methodology to waggle dance data from twenty hives, we show our model closely fits real-world honeybee foraging patterns and we demonstrate that colonies vary their use of waggle dance information across different landscapes.
 Our methodology provides a tool to identify the ecological conditions in which honeybee colonies rely on dance communication, opening the door to large-scale experimental exploration of the selection pressures that may have driven the evolution of this remarkable collective behaviour.
@@
 
-Within honeybee colonies, a series of simple rules that determine when and how much bees perform the celebrated waggle dance (Fig.
-1) mean that choices between feeding sites occur at the level of the group rather than the individual.
-For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the sweeter of two equidistant sources [@Seeley1991], or the closer of two equally sweet sources [@Hasenjager2022], without requiring any individual bee to compare options.
+No communication system is more celebrated than the honeybee waggle dance, yet there remains aspects of the behaviour which have not yet been quantified.
+
+Within honeybee colonies, a series of simple rules determine when and how much bees perform the celebrated waggle dance (Fig.
+1) mean that choices between feeding sites occur at the level of the group rather than the individual [@Seeley1994A].
+For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the richer of two equidistant sources [@Seeley1991], or the closer of two equally rich sources [@Hasenjager2022], without requiring any individuals to go out and visit multiple resources to make a directly informed choice.
 This extraordinary system is a key example of how social insect colony behaviour can take on a form extending beyond that of the individual units [@Sumpter2006], functioning as a collective that makes decisions as a single entity.
@@
 Accordingly, in many situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2004 @Dornhaus2006; @Gruter2013; @Price2019].
-This apparent variation in the use of collective behaviour is intriguing because it may provide clues as to the selection pressures that were critical in the evolution of the dance communication system.
+This plasticity in the use of collective behaviour is intriguing because it may provide clues as to the selection pressures that were critical in the evolution of the dance communication system.
 
@@
 Our methods provide an efficient means of quantifying collective behaviour with no requirement to manipulate hive orientations, assay foraging efficiency or document individuals' previous experience.
+As such, they can identify when colonies adopt a collective foraging strategy- a critical step towards understanding when such a strategy outperforms individual-level behaviour.
 
@@
 
-To reveal how the patterns of waggle dance-encoded foraging distances might differ under scenarios where colonies rely on individual search versus recruitment, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
+The patterns of waggle dance-encoded foraging distances might differ under scenarios where colonies rely on individual search versus recruitment.
+To show this, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
 Foragers could locate these under two different strategies: either acting as a scout and locating resources themselves, or following a recruit strategy and locating resources by following a random dance from the dance floor [@Seeley1992] (Fig.
 2a,b, details in Methods).
-As it is known in the simulation which individuals in the hive forage under what strategy, we can compare the distributions of foraging distances reported on the dance floor.
+As it is known in the simulation which individuals in the hive forage under what strategy, we can compare the distributions of foraging distances reported on the dance floor by each type of forager.
 Fig 2c shows that the shapes of the resource distance distributions for bees engaging in the two types of foraging trips are different.
@@
 
-This description intrinsically captures honeybee foraging as collective decision making, whereby the foraging sites represented on the dance floor derive from a mixture of individual search and waggle dance information, modified by the profitability rule that biases recruitment towards closer or sweeter patches.
+This description intrinsically captures honeybee foraging as collective decision making, whereby the foraging sites represented on the dance floor derive from a mixture of individual search and waggle dance information, modified by the profitability rule that biases recruitment towards closer or richer patches.
 The extent to which collective decision making is used is expressed in the proportion of scout dances, $p$, and the proportion of recruit dances, $1-p$.
@@
 
-In a hypothetical colony that relies only on individual search to find foraging sites, all trips are "scout" trips, but as recruitment becomes more important, the proportion of “recruit” trips will increase.
-By setting $p = 1$, we obtain a model based on the sole use of individual search (scouts only) and by allowing the proportion of scout trips, $p$, to take on any value between 0 and 1, we can derive the extent to which foraging occurs collectively.
+We constructed two models of foraging: an individual model describing a hypothetical colony that relies only on individual search to find foraging sites, and a collective model which describes a colony using waggle dance communication to exploit resources.
+In the individual model, all trips are "scout" trips, but as recruitment becomes more important, the proportion of “recruit” trips will increase.
+By setting $p = 1$, we obtain a model based on the sole use of individual search (scouts only) and, alternatively, by allowing the proportion of scout trips, $p$, to take on any value between 0 and 1, we can model the extent to which foraging occurs collectively.
 (Fig 2).
@@
 
-For 16 out of 20 study hives, a model of collective foraging provided a better description of the data than one of individual foraging (Fig.
+For 16 out of 20 study hives, the collective model provided a better description of the data than the individual model (Fig.
 4a).
@@
 In all but one site, the collective model had a good fit (using a Kolmogorov-Smirnov statistic of $p > 0.05$, see methods) to the empirical waggle run durations (Fig.
-4b), whereas the scout model was significantly different to the observed data in 8 sites (Kolmogorov-Smirnov statistic $p < 0.05$, Fig.
+4b), whereas the individual model was significantly different to the observed data in 8 sites (Kolmogorov-Smirnov statistic $p < 0.05$, Fig.
 4b).
@@
 Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individual's foraging behaviour over a landscape.
-As our results show that individual foraging accurately describes colony foraging in 11 different sites, and more parsimoniously than a model of collective foraging in 4 sites, these findings support the idea that individual decision making can dominate colony foraging and demonstrates further evidence for flexible waggle dance use by honey bee colonies.
+As our results show that our model of individual foraging provides a good fit to colony foraging in 11 different sites, and provides a more parsimonious description of foraging than a model of collective foraging in 4 sites, these findings support the idea that individual decision making can dominate colony foraging and demonstrates further evidence for flexible waggle dance use by honey bee colonies.
 
@@
 5c.).
-Parks, allotments and cemeteries are typically forage-rich for honeybees [@Baldock2019], a result which was not replicated here.
+Parks, allotments and cemeteries are typically forage-rich for honeybees [@Baldock2019], yet our model indicates more scouting trips in these landscapes.
 However, note that the amount of land cover for this component is limited and the jackknifed PLS identified substantial variation in the loadings identified for the first principal components (Supplementary Fig.
@@
 
-Here, we have presented a model to quantify the use of waggle-dance communication in collective decision-making by honeybee colonies.
+Wwe have presented a model to quantify the use of waggle-dance communication in collective decision-making by honeybee colonies.
 Whilst recruitment is well known to occur through other mechanisms besides the waggle dance, such as through olfactory cues [@Arenas2007], the exceptionally close fit of our model to waggle run durations underlines the importance of the waggle dance in honey bee foraging.
 The fitting of our model to waggle dance observations provides a time and labour efficient methodology to quantify collective behaviour.
-Further, by identifying the land-use combinations which most influence variation in the proportion of scouts, our model provides a tool to map the environment along the major axis of honeybee information use and visualise how land-use influences the use of the waggle dance.
+Our model provides a tool to map the environment along the major axis of honeybee information use and visualise how land-use influences the use of the waggle dance and thus identifies the land-use combinations which most influence variation in the proportion of scouts.
 
 Whilst our results show a decrease in reliance on waggle dance recruitment as resources become harder to find and foraging trips become longer [@Samuelson2021], they do not yet provide conclusive evidence for such a pattern.
-Our study was designed as a proof-of-principle based on an existing dataset, we did not systematically choose sites based on forage availability, and instead inferred forage availability based on land-use at each site.
+Although the agri-rural results agree with this assessment, the urban results seem to go against this trend.
+As we used an existing dataset, we did not systematically choose sites based on forage availability, and instead inferred forage availability based on land-use at each site.
 The dance recordings in our dataset were collected over an extended period of five months, over which time forage availability in the landscape likely changed considerably and non-uniformly across sites.
 We thus cannot rule out that longer term effects not captured in the data used in this study, such as resource stability [@Schurch2014], may also have contributed to our estimates of waggle dance use.
+Our findings illustrate how our model can be used to estimate reliance on collective behaviour within different landscapes, for exploration within datasets where land-use is systematically chosen by design.
 
@@
 We analysed an existing data set and thus provide a proof of concept for how this new toolkit can be used to evaluate factors influencing waggle dance recruitment.
+The methods described here can also provide a framework to start exploring how collective decision making can help to buffer the effects of landscape change, and in this case how flexibility in the use of the waggle dance may help to exploit resources when experiencing novel landscape mosaics such as in the highly urbanised conurbations.
 

### 2022-05-03 analysis/manuscript/supplementary_material.Rmd
updated final MS edits. Just last meeting to go

@@
 title: "Waggle dance distributions quantify collective foraging in honey bee colonies: Supplementary material"
-author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
+# author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
 output:
@@
 grid.arrange(arrangeGrob(allsites_plot, left = y.grob, bottom = x.grob))
+
+ggsave(
+  "../results/figures/SI_figures/allsites_plot.png"
+)
+
 ```
@@
   fit_to_width(7.5)
+
+textable <- kbl(
+  df2,
+  caption = "Fit results for the fitting of the waggle dance model to each site.",
+  booktabs = TRUE
+  ) %>%
+  kable_styling(latex_options = "striped")
+
+print(textable)
+
+print(
+  xtable(
+    df2, type = "latex",
+    caption = "Fit results for the fitting of the waggle dance model to each site."
+  ),
+  scalebox='0.65',
+  floating = FALSE,
+  caption.placement = "top",
+  file = "../results/figures/SI_figures/fittable.tex"
+)
+
 ```

### 2022-06-28 analysis/manuscript/methods.Rmd
made some updates for submittion to sci-advances

@@
 As our estimated proportion of scouts is continuos on the interval $[0, 1]$ we used the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010].
-As the betareg package only works on the open interval $(0, 1)$ the data, $x$, was transformed using the following equation: $\frac{x(n-1)+0.5}{n}, as outlined in the betareg package documentation.
+As the betareg package only works on the open interval $(0, 1)$ the data, $x$, was transformed using the following equation: $\frac{x(n-1)+0.5}{n}$, as outlined in the betareg package documentation.
 After analysis the data was back transformed to the original values for the plots in Fig 5.

### 2022-06-28 analysis/manuscript/paper.Rmd
made some updates for submittion to sci-advances

@@
 bibliography: paper-refs.bib
-csl: nature.csl
+csl: science.csl
 ---
@@
 
-Honeybee foraging is an extraordinary collective behaviour that is directed through the waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops.
-Here, we present a mathematical framework to explain how the intrinsic process of the waggle dance can allow a colony to exploit different landscape types.
-From waggle dance observations, we infer the proportion of dances performed by bees that forage individually, and those that forage collectively.
-By applying this methodology to waggle dance data from twenty hives, we show our model closely fits real-world honeybee foraging patterns and we demonstrate that colonies vary their use of waggle dance information across different landscapes.
+Honeybee foraging is an extraordinary collective behaviour that is directed by the waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops.
+Here, we present a mathematical framework to quantify how use of the waggle dance shifts in different landscape types.
+By decoding dances inside hives, we infer the proportion that are performed by bees that forage individually, and those that forage collectively.
+By applying this methodology to data from twenty hives, we show our model closely fits real-world honeybee foraging patterns and we demonstrate that colonies vary their use of waggle dance information across different landscapes.
 Our methodology provides a tool to identify the ecological conditions in which honeybee colonies rely on dance communication, opening the door to large-scale experimental exploration of the selection pressures that may have driven the evolution of this remarkable collective behaviour.
@@
 
-No communication system is more celebrated than the honeybee waggle dance, yet there remains aspects of the behaviour which have not yet been quantified.
+The extraordinary waggle dance communication system of honeybees (Apis spp.) is a key example of how social behaviour can take on a form extending beyond that of the individual units [@Sumpter2006], such that groups function as collectives that make decisions as a single entity.
+Within colonies, a series of simple rules determine when and how much bees perform the waggle dance (Fig.
+1) mean that choices between feeding sites occur at the level of the group rather than the individual [@Seeley1994].
+For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the richer of two equidistant sources [@Seeley1991], or the closer of two equally rich sources [@Hasenjager2022].
+This is without requiring any individuals to visit multiple resources to make a direct comparison.
+The resulting colony-level decisions are a prime example of how self-organized collective behaviour could increase the foraging efficiency of groups.
 
-Within honeybee colonies, a series of simple rules determine when and how much bees perform the celebrated waggle dance (Fig.
-1) mean that choices between feeding sites occur at the level of the group rather than the individual [@Seeley1994A].
-For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the richer of two equidistant sources [@Seeley1991], or the closer of two equally rich sources [@Hasenjager2022], without requiring any individuals to go out and visit multiple resources to make a directly informed choice.
-This extraordinary system is a key example of how social insect colony behaviour can take on a form extending beyond that of the individual units [@Sumpter2006], functioning as a collective that makes decisions as a single entity.
-
-Despite its intricacy and precision, a wealth of research now shows that foraging bees frequently do not use dance information to find forage sites [@Gruter2009].
-Accordingly, in many situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2004 @Dornhaus2006; @Gruter2013; @Price2019].
+A wealth of research now shows that foraging bees frequently do not use dance information to find forage sites [@Gruter2009].
+Accordingly, in many situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002;@Dornhaus2004;@Gruter2013;@Price2019].
 This plasticity in the use of collective behaviour is intriguing because it may provide clues as to the selection pressures that were critical in the evolution of the dance communication system.
@@
 Identifying the circumstances in which bees use the dance is likely to be a key step towards understanding when it is likely to afford fitness benefits at the colony level.
-Despite the amount of research interest that has focused on the proximate mechanisms of collective decision-making via dance communication, the ultimate evolutionary drivers of its unique evolution within *Apis* -unparalleled in even a single other social insect- remain obscure [@Price2015].
-While it is likely that the spatiotemporal distribution of forage was key, some studies find the benefits of dance communication to be realized only in challenging environments where resources are clumped or ephemeral [@Sherman2002; @Dornhaus2004; @Dornhaus2006], while others find no effects of landscape heterogeneity [@Nurnberger2017], or that dancing comes at a cost in challenging environments [@Price2019] and is most beneficial when species richness is high [@Donaldson-Matasci2012].
+Despite the amount of research interest that has focused on the proximate mechanisms of collective decision-making via dance communication, the ultimate evolutionary drivers of its unique evolution within Apis -unparalleled in even a single other social insect- remain obscure [@Price2015].
+While it is likely that the spatiotemporal distribution of forage was key, some studies find the benefits of dance communication to be realized only in challenging environments where resources are clumped or ephemeral [@Sherman2002;@Dornhaus2004;@Dornhaus2006], while others find no effects of landscape heterogeneity [@Nurnberger2017].
+Others have found that dancing comes at a cost in challenging environments [@Price2019] and is most beneficial when species richness is high [@Donaldson-Matasci2012].
+However, identifying how variation in landscape structure is associated with waggle dance use is practically difficult.
 
-Attempts to identify the circumstances in which dance communication is important have been challenging because they require constant monitoring of real-world colony weight [@Kirchner1998; @Sherman2002; @Dornhaus2006; @Gruter2013; @Nurnberger2017; @Price2019] -a noisy proxy of foraging success that is influenced by many other abiotic and biotic factors, including colony health- over ecologically credible time periods [@Schurch2014].
+Attempts to identify the circumstances in which dance communication is important have been challenging because they require constant monitoring of real-world colony weight [@Sherman2002;@Dornhaus2004;Gruter20013;@Price2019;@Dornhaus2006;@Nurnberger2017;@Kirchner1998] -a noisy proxy of foraging success that is influenced by many other abiotic and biotic factors, including colony health- over ecologically credible time periods [@Schurch2014].
 To do this at a scale that allows inter-colony variation in foraging environments at sufficient replication is a major logistical hurdle (but see [@Nurnberger2017]), particularly given that multiple landscape variables may interact to determine the utility of dance communication.
@@
 Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions.
-Overall resource “quality” -the net energetic gain of a foraging trip- is provided through the number of waggle runs performed [@Frisch1993; @Boch1956; @Esch1961; @Seeley1994; @Seeley2000].
+Overall resource “quality” -the net energetic gain of a foraging trip- is provided through the number of waggle runs performed [@Frisch1993;@Boch1956;@Esch1961;@Seeley1994;@Seeley2000].
 Although bees that follow dances do not specifically interpret this information on an individual level [@Seeley1991], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995], and provides the colony with a mechanism to select the most profitable resources in their environment [@Seeley1994A]."}
@@
 Here, we present a method to determine the extent of waggle dance use in colony foraging within specific landscapes by identifying the proportion of foraging trips that are made by bees recruited through waggle dances.
-Our method examines the distribution of waggle run durations reported on the dance-floor and infers from its shape the relative contribution of dances for resources identified by individual search, "scouting" versus dances for resources found by following dances, “recruitment” [@Seeley1995].
+Our method examines the distribution of waggle run durations reported on the dance-floor and infers from its shape the relative contribution of dances for resources identified by individual search, “scouting” versus dances for resources found by following dances, “recruitment” [@Seeley1995].
 Previously, the relative amount of scouting and recruitment trips have only been quantifiable by documenting foragers' individual search history, which requires tracking individual bees and is time -and labour- intensive [@Seeley1995].
@@
 (c) Complementary cumulative frequencies of foraging distances reported from scouting and recruit trips.
-Note the difference in the shape of the distributions"}
+Note the difference in the shape of the distributions.
+The scout distribution is best fit by an exponential (grey fit line), the recruit distribution by a Rayleigh distribution (black line)."}
 knitr::include_graphics("../results/figures/manuscript_figures/fig2.png")
@@
 On the dance-floor, the number of waggle runs performed for a resource depends on its profitability [@Seeley1994].
-Honeybees achieve this by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994; @Seeley1994A].
+Honeybees achieve this by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994;@Seeley1994A].
 By combining this profitability bias with the distributions identified in our simulations, we can accurately describe the distribution of waggle runs reported on the dance-floors of real honeybee colonies as a superposition of scout and recruit distributions (Fig 3, see Methods for details)
@@
 On each visit, two hours of continuous waggle dance data was recorded by training a camcorder onto the dance floor.
-The footage of the dances was decoded manually [@Couvillon2012; @Schurch2019; @Samuelson2021] to extract waggle run durations.
+The footage of the dances was decoded manually [@Couvillon2012;@Schurch2019;@Samuelson2021] to extract waggle run durations.
 Using this data, for each site we fit both the collective and individual models and used model selection to determine which provided the better explanation of the data, and (if the collective model provided a better fit) to quantify the relative use of social information through estimating the parameter $p$.
-In each case, We calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004; @Clauset2009].
+In each case, We calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004;@Clauset2009].
 
@@
 
-In the urban environment, the first principal component explained ~73% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.73, $\phi$ = 10.4, p < 0.05, Fig.
+In the urban environment, our findings were more equivocal.
+The first principal component explained ~73% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.73, $\phi$ = 10.4, p < 0.05, Fig.
 5c).
-This component correlates positively with parks, allotments and cemeteries (8% of land cover; Table 1) and negatively with railways, however this accounts for ~1% of land cover (Table 1).
-The dominant land-use type by coverage, sparse residential (land coverage ~35%; Table 1), has a significant negative correlation with the first principal component for some sites removed over the jackknife sampling (Supplementary Fig.
-2.), however, it is overall not a significant contributor to the variance in the loadings (Fig.
+This correlated negatively with the dominant land-use type by coverage, sparse residential (land coverage ~35%; Table 1), as would be expected if the proportion of scouts decreases when rich forage is available, since such land includes a high proportion of gardens.
+However, this relationship is very sensitive to site removal through the jackknife sampling (Supplementary Fig.
+2.), and is not a significant contributor to the variance in the loadings (Fig.
 5c.).
-Parks, allotments and cemeteries are typically forage-rich for honeybees [@Baldock2019], yet our model indicates more scouting trips in these landscapes.
-However, note that the amount of land cover for this component is limited and the jackknifed PLS identified substantial variation in the loadings identified for the first principal components (Supplementary Fig.
+Furthermore, the first principal component also correlates positively with parks, allotments and cemeteries (8% of land cover; also railways but these constitute ~1% of land cover; Table 1), which would typically be considered forage-rich for honeybees [@Baldock2019] (although note that the amount of land cover for this component is limited and the jackknifed PLS identified substantial variation in the loadings identified for the first principal components; Supplementary Fig.
 3.).
@@
 
-Our results within the agri-rural sites are in agreement with other studies which evaluated foraging performance with and without the dance [@Dornhaus2006; @Price2019].
+Our results within the agri-rural sites are in agreement with other studies which evaluated foraging performance with and without the dance [@Dornhaus2006;@Price2019].
 These results suggest that, even though collective foraging is not always beneficial, in environments where high quality resources are present, but relatively scarce, collective foraging is being carried out by colonies, suggesting that under these conditions exchanging social information through the waggle dance confers benefits.
@@
 
+# Materials and Methods
+
+## Data collection
+
+Methods for how the data used for this study was collected can be found in full in Materials and Methods sections 2.1, 2.2 and 2.3 of [@Samuelson2021].
+
+### Land-use preference analysis
+
+Methods for how land-use types were classified can be found in Materials and Methods sections 2.6 of [@Samuelson2021].
+
+### Waggle dance decoding
+
+Methods for waggle dance decoding are fully described in Materials and Methods section 2.4 in [@Samuelson2021].
+
+## Simulation
+
+All simulation code was written in Python version 3.9 and uses the Pandas [@Mckinney2011] and Scipy [@Jones2001] packages.
+
+A circular environment is first created with radius $r = 2.5$.
+The number of resources in the environment is generated as a random Poisson variable with rate equal to 5000 multiplied by the area of the environment.
+These are placed on polar coordinates with a uniformly selected angle, $\theta$, between 0 and 2$\pi$ and a radial value, $\rho$, between 0 and $r$, determined from the square root of the uniform position values multiplied by $r$.
+These polar coordinates are converted to Cartesian coordinates.
+Each location is then assigned to an instance of a resource object along with a random quality of between 0 and 10.
+This quality is combined with the distance of the resource to the centrally located hive to form a measure of how profitable the resource is (see model, equation?).
+
+One-hundred honeybee objects are created, 20 of which 20 are on scouting trips and the rest recruited to follow scout dances.
+Scouts leave the environment following a random path through the environment generated by sampling a uniform random step length and angle.
+The number of paths the scout draws when searching is also determined as a uniform random number.
+Each straight line in the random path is converted to a rectangle with length equal to the path section length and a constant width of ~0.01 to represent an area the scout searches along that path.
+Of all the resources contained in the boxes drawn from the scout's path, the one closest to the colony is selected as the resource patch that the scout will report and will communicate its location if the quality of the resource resource exceeds a minimum threshold.
+Communication is simulated by pooling together all the resource patches found.
+If no resources are contained in the scout's path, they will not add any resources to the scout pool and draw a new path in the next foraging iteration.
+
+Recruits represent honeybee objects which do not perform the searches the scouts do.
+Instead, they sample from the pool of resource objects reported by scouts.
+This sampling is done by selecting resources with a probability which is skewed towards the profitability of the resource, meaning more profitable resources have a greater chance of being selected by recruits.
+Recruits will then visit these resources and in the next iteration will add their resource to the pool of scout dances.
+Consequently, the pool of dances represents resources discovered by scouts and resources exploited by recruits.
+When a resource is depleted, it is removed from the environment and so any foragers that were foraging on it would select a different resource from the dance floor.
+
+The simulation was run 100 times and every 5 time steps all distances reported by scouts and recruits were recorded and combined.
+This was done to reflect the way foraging data is collected in real honeybee studies.
+We fit an exponential and minimum of an exponential distribution to both the distribution of foraging distances reported by the scout and recruit objects.
+Fitting was done by deriving the maximum likelihood estimate for each model fit on each data source through their analytical solutions: $\hat{\lambda} = \frac{1}{\bar{x}}$, minimum of the exponential with a minimum foraging distance: $\hat{\lambda} = \frac{1}{\pi \bar{x^{2}}}$.
+As the exponential assumes distributions start from 0 the data was transformed to start from 0 by subtracting the minimum foraging distance from all foraging distances ($x = x - min(x)$) before fitting.
+
+## Model
+
+To describe the distribution of dance durations on the dance floor we formulated a generic model the duration of waggle dances.
+In the model resource patches are assumed to be randomly placed in the environment.
+Foragers scout for these patches.
+The rationale of the model is illustrated in Fig.
+3: upon visiting a resource patch, foragers translate the profitability of a resource into the number of repeats of the dance.
+The number of repeats of the dance is a function of quality an distance.
+Recruits sample random dances and report the location of successful visits to resource patches on the dance floor.
+Through the feedback and over-representation of profitable resources on the dance floor recruits will converge to visiting the most profitable resource in vicinity of the hive.
+The distribution of dance durations is the superposition of scouting and recruiting trips.
+
+As the resources patches are randomly placed in the environment, the distance after which the first resource is discovered approximately follows an exponential distribution (given by $\lambda e^{- \lambda x}$).
+Through the feedback mechanism that the dance floor provides, the colony can, collectively, locate the most profitable resource in its environment.
+For randomly placed resources in a two dimensional environment the distance to the nearest point is distributed according to a Rayleigh distribution (given by $2 \lambda \pi x e^{- \pi \lambda x^2}$)) [@Pyke1978].
+Our simulation model shows that this describes the distances at which recruits visit resources well.
+Knowing the distance distributions of scout and recruit trips we then assume that the a proportion $p$ of all trips are scout trips.
+With this information we can specify the distributions of distances on the dance floor (see Supplementary Material for details).
+
+We implemented this in a full model which describes the distance distribution of the environment has $n$ different resource types (See Supplementary Material).
+However, in the full model the number of parameters increases with $2n$.
+Even if the number of resources is low, it turned out to be cumbersome to estimate the parameters and the model tends to overfit.
+To facilitate estimation of the parameter $p$ we therefore used a simplified model to estimate the fraction of scout trips, where $m$ represents the lowest duration considered, and here a minimum waggle run duration in the data set.
+
+For the simplified the model we assumed that the number of dances depends weakly on distance and there is a sizable quality differences between resources of a non-negligible size and that there is a sizable intensity of the high quality resource (See Supplementary Material for detailed derivation).
+Foragers on scouting trips are more likely to report larger distances than foragers on recruiting trips.
+By linearising the function that translates the profitability into the number of waggle dance run for the largest dance duration and normalising, we arrive at simplified distribution for dance durations for scouting trips:
+
+$$ f_s(x)=a_s \left[ M_s^{-1} b_s e^{-b_s a_s (x-m)}  \left[ 1-a_s x \right]_+ \right] $$
+
+where we used the shorthand ${x}_+=\max(x,0)$ The parameter $a_s^{-1}$ is the maximum dance duration by scouts, $a_s b_s$ is the intensity of resources found by scouts and $M_s=(1-a_s m)-b_s^{-1} \left(1-e^{-b_s(1-a_s m)}\right)$ is factor that normalises the distribution.
+
+Recruit trips will be predominantly to high quality resources.
+Only if the nearest high quality patch is very far away will there be a more profitable patch of lesser quality available, and this happens only rarely if the intensity of the best quality resource is sizable.
+After linearising the function that translates the profitability into the number of waggle dance runs for short dance durations and normalising the distribution of dance durations reported from recruit trips in the simplified model is:
+
+$$f_r(x)=M_r^{-1} 2 \pi a_r^2   b_r x e^{- \pi  b_r  (a_r x)^2} \left[1- a_r x\right]_+$$
+
+where
+
+$$M_r={(1-a_r m)e^{-\pi br(a_r m)^2} + \frac{erf(a_r \sqrt{\pi b_r m}) - erf(\sqrt{\pi b_r}))}{2 \sqrt{b_r}}}$$
+
+is the normalisation factor, the parameter $a_r$ is the rate with which dances repeats depends on distance for recruit trips and $a_r^2 b_r$ is the intensity of high quality resources reported by recruited foragers.
+
+The simplified distribution function is
+
+$$P(\underline x= x )=p f_s(x)+ (1-p) f_r(x), $$ which we used for parameter estimation and model fittings.
+
+## Statistical analysis
+
+All analysis code is written in R [@RCore2020].
+
+### Model fitting
+
+All models are fit using Maximum likelihood estimation [@Burnham2002] by summation of the log of the simplified distribution function outlined in the methods section: model.
+The numerical optimisation routine is written in c++ and uses the Nelder-Mead simplex algorithm [@Nelder1965] implemented in the 'NLopt' library [@Johnson2020] and interfaced to R [@RCore2020] using 'Rcpp' [@Eddelbuettel2011].
+
+The most parsimonious model is assessed using Akaike information criterion (AIC) [@Burnham2002; @Aho2014] and Akaike weights. The model with the lowest AIC score is deemed to be the most parsimonious.
+
+Goodness of fit is assessed using the two-sample Kolmorgorov-Smirnov (KS) test [@Goldstein2004] and implemented in R using the ks.boot function of the package 'Matching' in R [@Sekhon2011].
+
+### Partial Least Squares analysis
+
+Prior to conducing the PLS we removed any sites in which the models fit was significantly different to the waggle dance durations for that site.
+This resulted in one agri-rural site (ROT) and no urban sites being removed from the analysis.
+
+As our estimated proportion of scouts is continuos on the interval $[0, 1]$ we used the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010].
+As the betareg package only works on the open interval $(0, 1)$ the data, $x$, was transformed using the following equation: $\frac{x(n-1)+0.5}{n}$, as outlined in the betareg package documentation.
+After analysis the data was back transformed to the original values for the plots in Fig 5.
+
+For the jackknifed resampling we iterated through the each site and removed it from the pool of data and then ran the PLS as described above, recoding the loadings for each iteration (see Supplementary Material for loadings with each site removed).
+The PLS loadings for each land-use type are plotted as a box plot in Fig 5.
+to show the spread of these variable types.
+A loading was determined to be significantly correlating with the first principal component if contributed more than its expected variance.
+
+# Acknowledgments
+
+**Funding**
+
+This work was supported by the Biotechnology and Biological Sciences Research Council (BBSRC) through grant BB/M011178/1.
+
+**Author Contributions**
+
+Conceptualization: JP, RG, EL, VJ.
+Methodology: JP, EL, VJ.
+Software: JP.
+Formal analysis: JP.
+Investigation: AS, EL.
+Writing - Original Draft: JP, VJ.
+Writing - review & editing: JP, RG, EL, VJ.
+Supervision: RG, EL, VJ.
+
+**Competing interests**
+
+Authors declare they have no competing interests.
+
+**Data and materials availability**
+
+All data and code is available on zenodo at....
+
 # References

### 2022-06-29 analysis/landuse_effect_jackknife.Rmd
updated the code to use proportion of recruits rather than proportion of scouts, this was we can more easily compare landscape type with a measure of waggle dance use

@@
   pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
+
+```
+
+```{r, convert_p}
+
+ms_mr <- read.csv(file = "data/site_ms_mr.csv")
+
+urban_lu <- urban_lu %>%
+  # inner_join(ms_mr, by = "site") %>%
+  # select(-ms, -mr, -p.y) %>%
+  # rename(p = p_trips) %>%
+  # mutate(p = ifelse(p.x == 1, 1, p)) %>%
+  mutate(p = 1-p)
+  # select(site, p, 3:10)
+
+rural_lu <- rural_lu %>%
+  # inner_join(ms_mr, by = "site") %>%
+  # select(-ms, -mr, -p.y) %>%
+  # rename(p = p_trips) %>%
+  # mutate(p = ifelse(p.x == 1, 1, p)) %>%
+  mutate(p = 1-p)
+  # select(site, p, woodland, 3:12)
 

### 2022-06-29 analysis/landuse_effect_prop_scouts.Rmd
updated the code to use proportion of recruits rather than proportion of scouts, this was we can more easily compare landscape type with a measure of waggle dance use

@@
   pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
+```
+
+```{r, convert_p}
+
+ms_mr <- read.csv(file = "data/site_ms_mr.csv")
+
+urban_lu <- urban_lu %>%
+  # inner_join(ms_mr, by = "site") %>%
+  # select(-ms, -mr, -p.y) %>%
+  # rename(p = p_trips) %>%
+  # mutate(p = ifelse(p.x == 1, 1, p)) %>%
+  mutate(p = 1-p)
+  # select(site, p, 3:10)
+
+rural_lu <- rural_lu %>%
+  # inner_join(ms_mr, by = "site") %>%
+  # select(-ms, -mr, -p.y) %>%
+  # rename(p = p_trips) %>%
+  # mutate(p = ifelse(p.x == 1, 1, p)) %>%
+  mutate(p = 1-p)
+  # select(site, p, woodland, 3:12)
+
 ```
@@
   xlab("PC 1") +
-  ylab("Estimated proportion of scouts")
+  ylab("Estimated waggle dance use")
 
@@
   xlab("PC 1") +
-  ylab("Estimated proportion of scouts")
+  ylab("Estimated waggel dance use")
 

### 2022-07-26 analysis/manuscript/paper.Rmd
updated MS to new verion for current biology

@@
 
-The extraordinary waggle dance communication system of honeybees (Apis spp.) is a key example of how social behaviour can take on a form extending beyond that of the individual units [@Sumpter2006], such that groups function as collectives that make decisions as a single entity.
-Within colonies, a series of simple rules determine when and how much bees perform the waggle dance (Fig.
-1) mean that choices between feeding sites occur at the level of the group rather than the individual [@Seeley1994].
-For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the richer of two equidistant sources [@Seeley1991], or the closer of two equally rich sources [@Hasenjager2022].
-This is without requiring any individuals to visit multiple resources to make a direct comparison.
-The resulting colony-level decisions are a prime example of how self-organized collective behaviour could increase the foraging efficiency of groups.
+In group living animals, many decisions are taken collectively by integrating information from individuals to produce a social behaviour which extends beyond that of the individual units [@Sumpter2006].
+Such systems are self-organised, relying on the constituents using socially provided information and following behavioural 'rules' to produce a collective outcome [@Couzin2003;@Bonabeau1997].
+However, in addition to information provided by others, individuals also have knowledge of their own (private information) and so can choose to work collectively or individually [@Dall2005;@Reiucau2011;@Gruter2014].
+The benefits of working collectively versus independently vary across environmental contexts and individuals often adjust their own use of information accordingly [@Dechaume-Moncharmont2005;@Beekman2008].
+The impact these individual variations in information use have at a group level, however, is less well understood and difficult to empirically evaluate [@Rieucau2011;@Giardina2008;@Delgado2018].
 
-A wealth of research now shows that foraging bees frequently do not use dance information to find forage sites [@Gruter2009].
-Accordingly, in many situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002;@Dornhaus2004;@Gruter2013;@Price2019].
-This plasticity in the use of collective behaviour is intriguing because it may provide clues as to the selection pressures that were critical in the evolution of the dance communication system.
+Understanding the collective ramifications of individual variation in information use is not only important for deciphering how social animals collectively regulate information use [@Gruter2014], relating this to the environmental context of the behaviour is key to illuminating the ultimate evolutionary drivers of collective processes [@Price2015].
+However, establishing the group level effects of individual variation in information use is not trivial and requires measuring information use across animal groups, an often laborious process which requires monitoring individuals' behaviours and interactions.
+This becomes logistically impossible in very large animal groups, such as in social insects, which display some of the most intricate group decision-making mechanisms and consist of often thousands of individuals, each using a mix of individual and collective information to make decisions [@Bonabeau1997].
 
-Identifying the circumstances in which bees use the dance is likely to be a key step towards understanding when it is likely to afford fitness benefits at the colony level.
-Despite the amount of research interest that has focused on the proximate mechanisms of collective decision-making via dance communication, the ultimate evolutionary drivers of its unique evolution within Apis -unparalleled in even a single other social insect- remain obscure [@Price2015].
-While it is likely that the spatiotemporal distribution of forage was key, some studies find the benefits of dance communication to be realized only in challenging environments where resources are clumped or ephemeral [@Sherman2002;@Dornhaus2004;@Dornhaus2006], while others find no effects of landscape heterogeneity [@Nurnberger2017].
-Others have found that dancing comes at a cost in challenging environments [@Price2019] and is most beneficial when species richness is high [@Donaldson-Matasci2012].
-However, identifying how variation in landscape structure is associated with waggle dance use is practically difficult.
-
-Attempts to identify the circumstances in which dance communication is important have been challenging because they require constant monitoring of real-world colony weight [@Sherman2002;@Dornhaus2004;Gruter20013;@Price2019;@Dornhaus2006;@Nurnberger2017;@Kirchner1998] -a noisy proxy of foraging success that is influenced by many other abiotic and biotic factors, including colony health- over ecologically credible time periods [@Schurch2014].
-To do this at a scale that allows inter-colony variation in foraging environments at sufficient replication is a major logistical hurdle (but see [@Nurnberger2017]), particularly given that multiple landscape variables may interact to determine the utility of dance communication.
+Social insects are defined by their highly coordinated collective responses and so provide ideal systems to investigate the impact of individual variation in information use on emergent collective processes.
+Nevertheless, the sheer quantity of individuals in many colonies is a barrier to further understanding these dynamics.
+Honey bees (Apis spp.) provide an opportunity to sidestep this problem of scale as they provide a detailed record of colony foraging activity in the form of their unique waggle dance -unparalleled in even a single other social insect [@Seeley1995].
+Using the dance, individuals share information about profitable resource locations.
+A series of simple rules which determine when and how much bees perform waggle dances (Fig.
+1) mean that choices between feeding sites occur at the level of the group rather than the individual [@Seeley1991;@Seeley2004;@Hasenjager2022].
+Despite this, a wealth of research shows that individual foragers frequently do not use dance information to find forage sites and instead rely on their own knowledge [@Gruter2009].
+Moreover, studies investigating the value of the dance have shown that when prevented from communicating via dances, honey bee colonies can achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002;@Dornhaus2004;@Gruter2013;@Price2019].
 
@@
 
-Here, we present a method to determine the extent of waggle dance use in colony foraging within specific landscapes by identifying the proportion of foraging trips that are made by bees recruited through waggle dances.
-Our method examines the distribution of waggle run durations reported on the dance-floor and infers from its shape the relative contribution of dances for resources identified by individual search, “scouting” versus dances for resources found by following dances, “recruitment” [@Seeley1995].
-Previously, the relative amount of scouting and recruitment trips have only been quantifiable by documenting foragers' individual search history, which requires tracking individual bees and is time -and labour- intensive [@Seeley1995].
-Our methods provide an efficient means of quantifying collective behaviour with no requirement to manipulate hive orientations, assay foraging efficiency or document individuals' previous experience.
-As such, they can identify when colonies adopt a collective foraging strategy- a critical step towards understanding when such a strategy outperforms individual-level behaviour.
+The waggle dance provides a model system to study the group effects of collective and individual information use.
+Although it is likely that individual variation in dance use translates into variations in collective decision making at the colony level [@Donaldson-Matasci2012;@Price2019], this dynamic has not been empirically established.
+Individual based models, which explore the collective effects of individuals following different information streams provide an insight into how individual behaviours scale to the level of the group [@Dornhaus2006;Beekman2008], however, these are difficult to empirically relate to real observations and so are limited to showing what a colony under certain conditions might do, rather than quantifying what they are doing.
+Empirically determining how dance information is used at a colony level requires new methods which can identify the behaviours of individual foragers and integrate them with the mechanisms by which the dance functions to make collective decisions.
+Previously, this has only been quantifiable by documenting foragers' individual search history, which requires tracking individual bees and is time -and labour- intensive [@Seeley1995].
+
+Here, we investigate how variation in individual information use influences the collective foraging behaviour of honey bees.
+Using a model based on well-established foraging behaviours [@Seeley1986;@Seeley1991;@Seeley1992;@Seeley1994;@Seeley1994A;@Seeley1995], we examine the distribution of waggle run durations and infer from its shape the relative contribution of dances for resources identified by individual search - “scouting”- versus dances for resources found by following dances - “recruitment”- [@Seeley1995].
+This allows us to quantify the extent to which colonies use dance communication and thus forage collectively.
+To investigate how information use varies across environment, we fit our model to waggle dance observations from 20 sites in an agri-rural and urban environment (10 in each) and evaluated the landscape characteristics which best explained the variance in dance use.
 
@@
 The patterns of waggle dance-encoded foraging distances might differ under scenarios where colonies rely on individual search versus recruitment.
-To show this, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
+To test this, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
 Foragers could locate these under two different strategies: either acting as a scout and locating resources themselves, or following a recruit strategy and locating resources by following a random dance from the dance floor [@Seeley1992] (Fig.
-2a,b, details in Methods).
+2a,b, details in Materials and Methods).
 As it is known in the simulation which individuals in the hive forage under what strategy, we can compare the distributions of foraging distances reported on the dance floor by each type of forager.
@@
 The distance distribution for the scout trips is akin close to that of an exponential distribution (Fig.
-2c), which is the nearest neighbour distance distribution for foragers operating in a one-dimensional environment (see methods).
+2c), which is the nearest neighbour distance distribution for foragers operating in a one-dimensional environment (see Materials and Methods).
 The distribution of the distances reported for recruit trips (Fig.
-2c) is a Rayleigh distribution which is the nearest-neighbour distribution in a two-dimensional environment [@Pyke1978] (see Methods).
+2c) is a Rayleigh distribution which is the nearest-neighbour distribution in a two-dimensional environment [@Pyke1978] (see Materials and Methods).
 
@@
 Honeybees achieve this by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994;@Seeley1994A].
-By combining this profitability bias with the distributions identified in our simulations, we can accurately describe the distribution of waggle runs reported on the dance-floors of real honeybee colonies as a superposition of scout and recruit distributions (Fig 3, see Methods for details)
+By combining this profitability bias with the distributions identified in our simulations, we can accurately describe the distribution of waggle runs reported on the dance-floors of real honeybee colonies as a superposition of scout and recruit distributions (Fig 3, see Materials and Methods for details)
 
@@
 The distances of recruiting trips are than distributed through a Rayleigh distribution (f).
-The distances reported on the dance floor are a mixture of the scout and recruiting trips and can be calculated from the distance distributions of the scouting and recruiting trips, taking the reporting bias into account (see Methods for detail)."}
+The distances reported on the dance floor are a mixture of the scout and recruiting trips and can be calculated from the distance distributions of the scouting and recruiting trips, taking the reporting bias into account (see Materials and Methods for detail)."}
 knitr::include_graphics("../results/figures/manuscript_figures/fig3.png")
@@
 To evaluate the use of waggle dance information and individual search in honeybee colonies foraging in 'natural' landscapes, we analysed a pre-existing dataset of 2827 waggle dance observations from 20 observation hives, recorded between April-September 2017, (previously described in [@Samuelson2021]).
-Hives that contributed to this dataset had been situated at different locations in South East England (see Methods, figure 4A) and visited every two weeks for a period of 24 weeks.
+Hives that contributed to this dataset had been situated at different locations in South East England (see Materials and Methods, figure 4A) and visited every two weeks for a period of 24 weeks.
 On each visit, two hours of continuous waggle dance data was recorded by training a camcorder onto the dance floor.
@@
 In the other 4 sites, despite the collective model having the higher maximum likelihood, the individual model had a higher AIC value and so is more parsimonious (Supplementary Table 1).
-In all but one site, the collective model had a good fit (using a Kolmogorov-Smirnov statistic of $p > 0.05$, see methods) to the empirical waggle run durations (Fig.
+In all but one site, the collective model had a good fit (using a Kolmogorov-Smirnov statistic of $p > 0.05$, see Materials and Methods) to the empirical waggle run durations (Fig.
 4b), whereas the individual model was significantly different to the observed data in 8 sites (Kolmogorov-Smirnov statistic $p < 0.05$, Fig.
@@
 Since our sites varied in land-use characteristics and potentially thus forage distributions (although in this case, not by design for this study), we investigated whether these estimates might correlate with land-use.
-We first classified the different land-use types of the area surrounding each site [@Samuelson2018] to obtain a standardised land-use profile for the urban and agri-rural environments separately as many land-use types present in urban areas do not occur in agri-rural environments and vice versa (see methods).
-We then performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances within agri-rural and urban environments.
+We first classified the different land-use types of the area surrounding each site [@Samuelson2018] to obtain a standardised land-use profile for the urban and agri-rural environments separately as many land-use types present in urban areas do not occur in agri-rural environments and vice versa (see Materials and Methods).
+We then performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see Materials and Methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances within agri-rural and urban environments.
 As for one of our sites neither model provided a plausible description, this site was removed from the PLS analysis.
-Due to our small sample size (10 urban and 9 agri-rural sites), we used jackknife resampling to evaluate the robustness of our results to influential points (see Methods, Supplementary Material).
+Due to our small sample size (10 urban and 9 agri-rural sites), we used jackknife resampling to evaluate the robustness of our results to influential points (see Materials and Methods, Supplementary Material).
 
@@
 
-# Discussion
+# Discussion New
 
-Wwe have presented a model to quantify the use of waggle-dance communication in collective decision-making by honeybee colonies.
+Understanding how individual variation in information use scales to produce flexible collective processes at the group level is a key challenge in behavioural ecology.
+Here, we have presented a method to quantify information use at a collective level in a model social insect.
+Using this method, we have shown that collective decision-making is used variably.
 Whilst recruitment is well known to occur through other mechanisms besides the waggle dance, such as through olfactory cues [@Arenas2007], the exceptionally close fit of our model to waggle run durations underlines the importance of the waggle dance in honey bee foraging.
-The fitting of our model to waggle dance observations provides a time and labour efficient methodology to quantify collective behaviour.
-Our model provides a tool to map the environment along the major axis of honeybee information use and visualise how land-use influences the use of the waggle dance and thus identifies the land-use combinations which most influence variation in the proportion of scouts.
+Further examining our estimates of dance use show that a colonies use of the waggle dance varies across environments and does not appear to be randomly determined.
 
-Whilst our results show a decrease in reliance on waggle dance recruitment as resources become harder to find and foraging trips become longer [@Samuelson2021], they do not yet provide conclusive evidence for such a pattern.
-Although the agri-rural results agree with this assessment, the urban results seem to go against this trend.
+Within both the agri-rural and urban environments, dance use had a significant interaction with land-use type, ranging from total dance use to complete reliance on individual information.
+The notion that individual variation in information use influences the collective effect has only previously been inferred theoretically.
+Our findings demonstrate this empirically.
+Individual-based models [@Dornhaus2006;Beekman2008;@Price2019] have outlined how variations at an individual level could drive group level dynamics, however, without verifying this upon empirical data the insights remain theoretical.
+Here, we have shown how an individual based model can be used to better understand a system in a way which allows us to build more informative models which can be empirically verified.
+Not only do our findings add support for our understanding of how the dance functions to make collective decisions [@Seeley1995], but they also provide a new dimension to our understanding of how variation at an individual level influences the collective level.
+
+Whilst the benefits of dance communication are known to vary across environments [@Donaldson-Matasci2012;@Gruter2011;@Schurch2014;@Price2019], our results are the first to show colonies vary it's use in response, affording the colony a flexible dance use strategy which they can tailor to different conditions to maximise foraging success.
+Identifying the circumstances in which dance communication is important have been hampered by the requirement to monitor real-world colony weight [@Sherman2002;@Dornhaus2004;Gruter2013;@Price2019;@Dornhaus2006;@Nurnberger2017;@Kirchner1998] -a noisy proxy of foraging success that is influenced by many other abiotic and biotic factors, including colony health- over ecologically credible time periods [@Schurch2014].
+Doing this at a scale that allows inter-colony variation in foraging environments at sufficient replication is a major logistical hurdle (but see [@Nurnberger2017]), particularly given that multiple landscape variables may interact to determine the utility of dance communication.
+Our method circumvents this issue by providing a time and labour efficient methodology to quantify this key collective behaviour and map the environment along the major axis of honeybee information use.
+
+The methods described here provide a framework to start exploring how plasticity in collective decision making can help to buffer the effects of landscape change, and in this case, how flexibility in the use of the waggle dance may help to exploit resources when experiencing novel landscape mosaics, such as in highly urbanized conurbations.
 As we used an existing dataset, we did not systematically choose sites based on forage availability, and instead inferred forage availability based on land-use at each site.
@@
 We thus cannot rule out that longer term effects not captured in the data used in this study, such as resource stability [@Schurch2014], may also have contributed to our estimates of waggle dance use.
-Our findings illustrate how our model can be used to estimate reliance on collective behaviour within different landscapes, for exploration within datasets where land-use is systematically chosen by design.
+The number of colonies used in our analysis, although comparable to studies in which the benefits of collective foraging have been evaluated [@Sherman2002;@Dornhaus2004;@Price2019;@Nurnberger2017;Kirchner1998], limits the statistical power of our analysis.
+Despite this, we gain useful insights into the drivers of this relationship by exploring how land-use type influences waggle dance use.
 
-Our results within the agri-rural sites are in agreement with other studies which evaluated foraging performance with and without the dance [@Dornhaus2006;@Price2019].
-These results suggest that, even though collective foraging is not always beneficial, in environments where high quality resources are present, but relatively scarce, collective foraging is being carried out by colonies, suggesting that under these conditions exchanging social information through the waggle dance confers benefits.
-In both agri-rural and urban landscapes we have shown that the proportion of scouting trips change with land-use (illustrated in the change in land-use shown in the maps from left to right in Fig 5).
-We arrived at this conclusion through the analysis of waggle dance data, using a mathematical model to fit the waggle dance durations against.
-This method provides a time and labour efficient methodology to quantify collective behaviour.
-We analysed an existing data set and thus provide a proof of concept for how this new toolkit can be used to evaluate factors influencing waggle dance recruitment.
-The methods described here can also provide a framework to start exploring how collective decision making can help to buffer the effects of landscape change, and in this case how flexibility in the use of the waggle dance may help to exploit resources when experiencing novel landscape mosaics such as in the highly urbanised conurbations.
+Our agri-rural results suggest a potential decrease in reliance on waggle dance recruitment as resources become harder to find and foraging trips become longer [@Samuelson2021].
+Our results within the agri-rural sites broadly agree with the findings from other theoretical [@Dornhaus2006] (11) and empirical [@Price2019] studies which examined foraging performance with and without the dance.
+These results suggest that, even though collective foraging is not always beneficial, in environments where high quality resources are present but relatively scarce, collective foraging is being carried out by colonies and under these conditions exchanging social information through the waggle dance confers benefits.
+In both agri-rural and urban landscapes we have shown that the waggle dance use changes with land-use (illustrated in the change in land-use shown in the maps from left to right in Fig 5A and 5C).
+These findings illustrate how our model can be used to estimate reliance on collective behaviour within different landscapes, for exploration within datasets where land-use is systematically chosen by design.
 
-The analysis of waggle dance data can make an important contribution to our understanding of social information use and provides a methodology to further evaluate how honey bees use their unique dance language.
+The analysis of waggle dance data can make an important contribution to our understanding of information use.
 Recently, technological advances have emerged which enable colony metrics to be collected faster, more accurately and over greater time spans than could be gathered by hand, allowing individuals to be tracked within colonies and theories of individual behaviour to be evaluated in more depth than could have been done previously [@Wild2021].
@@
 
-Methods for how the data used for this study was collected can be found in full in Materials and Methods sections 2.1, 2.2 and 2.3 of [@Samuelson2021].
-
-### Land-use preference analysis
-
-Methods for how land-use types were classified can be found in Materials and Methods sections 2.6 of [@Samuelson2021].
-
-### Waggle dance decoding
-
-Methods for waggle dance decoding are fully described in Materials and Methods section 2.4 in [@Samuelson2021].
+Details of data collection , waggle dance decoding and classification of land-use types can be found in full in the Materials and Methods section of [@Samuelson2021].
 
@@
 Each location is then assigned to an instance of a resource object along with a random quality of between 0 and 10.
-This quality is combined with the distance of the resource to the centrally located hive to form a measure of how profitable the resource is (see model, equation?).
+This quality is combined with the distance of the resource to the centrally located hive to form a measure of how profitable the resource is (see model, equations).
 
-One-hundred honeybee objects are created, 20 of which 20 are on scouting trips and the rest recruited to follow scout dances.
+One-hundred honeybee objects are created, 20 of which are on scouting trips and the rest recruited to follow scout dances.
 Scouts leave the environment following a random path through the environment generated by sampling a uniform random step length and angle.
@@
 
-To describe the distribution of dance durations on the dance floor we formulated a generic model the duration of waggle dances.
+To describe the distribution of dance durations on the dance floor we formulated a generic model for the duration of waggle dances.
 In the model resource patches are assumed to be randomly placed in the environment.
@@
 Prior to conducing the PLS we removed any sites in which the models fit was significantly different to the waggle dance durations for that site.
-This resulted in one agri-rural site (ROT) and no urban sites being removed from the analysis.
+This resulted in one agri-rural site and no urban sites being removed from the analysis.
 
@@
 
-All data and code is available on zenodo at....
+All data and code is available on request and will be made available on zenodo prior to publication.
 

### 2022-07-26 analysis/manuscript/paper.Rmd
Updated refs for new version

@@
 ---
-title: "Waggle dance distributions quantify collective foraging in honey bee colonies"
+title: "Honeybees vary collective decision making in different landscapes. "
 # author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent A. A. Jansen"
@@
 always_allow_html: true
-bibliography: paper-refs.bib
+bibliography: reffs.bib
 csl: science.csl
@@
 Such systems are self-organised, relying on the constituents using socially provided information and following behavioural 'rules' to produce a collective outcome [@Couzin2003;@Bonabeau1997].
-However, in addition to information provided by others, individuals also have knowledge of their own (private information) and so can choose to work collectively or individually [@Dall2005;@Reiucau2011;@Gruter2014].
+However, in addition to information provided by others, individuals also have knowledge of their own (private information) and so can choose to work collectively or individually [@Dall2005;@Rieucau2011;@Gruter2014].
 The benefits of working collectively versus independently vary across environmental contexts and individuals often adjust their own use of information accordingly [@Dechaume-Moncharmont2005;@Beekman2008].
@@
 A series of simple rules which determine when and how much bees perform waggle dances (Fig.
-1) mean that choices between feeding sites occur at the level of the group rather than the individual [@Seeley1991;@Seeley2004;@Hasenjager2022].
+1) mean that choices between feeding sites occur at the level of the group rather than the individual [@Seeley1991;@Seeley1994A;@Hasenjager2022].
 Despite this, a wealth of research shows that individual foragers frequently do not use dance information to find forage sites and instead rely on their own knowledge [@Gruter2009].

### 2022-07-27 analysis/manuscript/paper.Rmd
updaed figs for updated version

@@
 **Simulating honey bee foraging**.
-In our simulation model with scouting only (a), foragers leave the hive on a search path (white lines) and continue until they encounter a resource (circles, colours indicate different resource quality).
-When foraging with recruitment (b) foragers continue to identify resources in scouting trips (white lines) and convey this information on the dance floor (brown disc) where foragers can sample dances reporting on scouting and recruiting trips and follow these directions (yellow lines).
-(c) Complementary cumulative frequencies of foraging distances reported from scouting and recruit trips.
+In our simulation model with scouting only (A), foragers leave the hive on a search path (white lines) and continue until they encounter a resource (circles, colours indicate different resource quality).
+When foraging with recruitment (B) foragers continue to identify resources in scouting trips (white lines) and convey this information on the dance floor (brown disc) where foragers can sample dances reporting on scouting and recruiting trips and follow these directions (yellow lines).
+(C) Complementary cumulative frequencies of foraging distances reported from scouting and recruit trips.
 Note the difference in the shape of the distributions.
-The scout distribution is best fit by an exponential (grey fit line), the recruit distribution by a Rayleigh distribution (black line)."}
+The scout distribution is best fit by an exponential (gray fit line), the recruit distribution by a Rayleigh distribution (black line)"}
 knitr::include_graphics("../results/figures/manuscript_figures/fig2.png")
@@
 **The rationale of the foraging model**.
-The distances of resources encountered by scouts are distributed exponentially (a).
-These dances are advertised on the dance floor (b).
-Dances for resources that are closer or higher in quality are repeated more often (c).
-As a consequence, dances for more profitable resource are over-represented and sampling foragers are biased to the more profitable resources (d).
-After successfully visiting advertised resources, recruits also dance for them leading to further amplification of this bias towards the most profitable resource in the vicinity of the hive (e).
-The distances of recruiting trips are than distributed through a Rayleigh distribution (f).
-The distances reported on the dance floor are a mixture of the scout and recruiting trips and can be calculated from the distance distributions of the scouting and recruiting trips, taking the reporting bias into account (see Materials and Methods for detail)."}
+The distances of resources encountered by scouts are distributed exponentially (A).
+These dances are advertised on the dance floor (B).
+Dances for resources that are closer or higher in quality are repeated more often (C).
+As a consequence, dances for more profitable resource are over-represented and sampling foragers are biased to the more profitable resources (D).
+After successfully visiting advertised resources, recruits also dance for them leading to further amplification of this bias towards the most profitable resource in the vicinity of the hive (E).
+The distances of recruiting trips are than distributed through a Rayleigh distribution (F).
+Recruits also report the locations on the dance floor (B) and repeat their runs more often depending on the profitability of the location (C), leading to a distribution of durations of recruit dances (G).
+By taking together the dance distributions for scouts (D) and recruits (G) the distributions of all dances on the dance-floor can be found.
+The distances reported on the dance floor this are a mixture of the scout and recruiting trips and can be calculated from the distance distributions of the scouting and recruiting trips, taking the reporting bias into account (see Materials and Methods for detail)."}
 knitr::include_graphics("../results/figures/manuscript_figures/fig3.png")
@@
 For 16 out of 20 study hives, the collective model provided a better description of the data than the individual model (Fig.
-4a).
+4A).
 In the other 4 sites, despite the collective model having the higher maximum likelihood, the individual model had a higher AIC value and so is more parsimonious (Supplementary Table 1).
 In all but one site, the collective model had a good fit (using a Kolmogorov-Smirnov statistic of $p > 0.05$, see Materials and Methods) to the empirical waggle run durations (Fig.
-4b), whereas the individual model was significantly different to the observed data in 8 sites (Kolmogorov-Smirnov statistic $p < 0.05$, Fig.
+4B), whereas the individual model was significantly different to the observed data in 8 sites (Kolmogorov-Smirnov statistic $p < 0.05$, Fig.
 4b).
-The sites shown in Figs 4c-d are representative examples showing the model fits where the individual (Fig.
-4c) and collective (Fig.
-4d) models fit best.
+The sites shown in Figs 4C-D are representative examples showing the model fits where the individual (Fig.
+4C) and collective (Fig.
+4D) models fit best.
 Note the closeness of the fit to the data, illustrating the overall quality of the model description.
@@
 **The honey bee foraging model fitted to data from 20 hives**.
-(a) Location of study hives in Southern England, shaded area in the main plot indicates Greater London.
-For 16 hives for the collective foraging model provided best explanation (black circles) or for 4 hives the individual search model provided the best explanation(red triangles) as indicated by lowest AIC score.
-(b).
+(A) Location of study hives in Southern England, shaded area in the main plot indicates Greater London.
+For 16 hives for the collective foraging model provided best explanation (black circles) or for 4 hives the individual search model provided the best explanation (red triangles) as indicated by lowest AIC score.
+(B).
 Distribution of goodness of fit confidence values for each model fit to waggle run durations from each site.
@@
 For values exceeding the threshold there is no statistically significant difference between the model and the data, indicating the model provides a good fit.
-For the hive in (c) the individual model (blue line) provided a better fit than the collective foraging model (red line).
-For the hive in (d) the collective foraging model (red line) provided a better fit than the individual model (blue line).
+For the hive in (C) the individual model (blue line) provided a better fit than the collective foraging model (red line).
+For the hive in (D) the collective foraging model (red line) provided a better fit than the individual model (blue line).
+The typical “hump” in the distribution in (D) which is indicative of contribution of recruitment dances (compare to Fig 3H).
 Panels show the compliment cumulative frequencies with binned frequency distributions as inset."}
@@
 
-Further quantification of the use of waggle-dance recruitment within all colonies, as a proportion of all foraging trips, can be achieved by extracting the estimated proportion of scout trips, $p$, for each site.
+Further quantification of the use of waggle-dance recruitment within all colonies, as a proportion of all foraging trips, can be achieved by extracting the estimated proportion of recruit trips, $1-p$, for each site.
 Since our sites varied in land-use characteristics and potentially thus forage distributions (although in this case, not by design for this study), we investigated whether these estimates might correlate with land-use.
 We first classified the different land-use types of the area surrounding each site [@Samuelson2018] to obtain a standardised land-use profile for the urban and agri-rural environments separately as many land-use types present in urban areas do not occur in agri-rural environments and vice versa (see Materials and Methods).
-We then performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see Materials and Methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances within agri-rural and urban environments.
+We then performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see Materials and Methods) to determine the principal components that represent combinations of land-use types which explained the most variation in waggle dance use within agri-rural and urban environments.
 As for one of our sites neither model provided a plausible description, this site was removed from the PLS analysis.
@@
 
-In the agri-rural environments the first principal component is a combination of land use types that explained ~73% of the variation in the proportion of scouts (beta regression: $R^2 = 0.73$, $\phi$ = 4.9, p < 0.05, Fig.
-5a).
+In the agri-rural environments the first principal component is a combination of land use types that explained ~73% of the variation in waggle dance use (beta regression: $R^2 = 0.73$, $\phi$ = 4.9, p < 0.05, Fig.
+5A).
 This principal component correlates positively with arable land (29% of land coverage; Table 1) and negatively with built-up areas (17% of land coverage); note that it also correlates negatively with non-agricultural unmanaged green space and water, but together these represent less than 3% of land-use (Table 1; Fig.
-5b).
+5B).
 These land-use types maintain a significant correlation with the first principal component over the jackknifed PLS (Fig.
-5b), with the exception of non-agricultural unmanaged green space which sits on the boarder, indicating the results are robust.
+5B), with the exception of non-agricultural unmanaged green space which sits on the boarder, indicating the results are robust.
 As arable land increases whilst built-up areas decrease, the proportion of trips that are driven by individual search increases.
@@
 **Collective foraging correlates with land-use**.
-The proportion of scouts for each site against the first principal component derived from a Partial Least Squares analysis of land-use type.
-Beta regression shows the relationship (black line) between first principal component and the proportion of scouts, with 95% CI shown by the grey shaded area for agri-rural (a) and urban landscapes (c).
-The correlations between first principal component and each land-use type are shown for agri-rural (b) and urban landscapes (d).
+Estimated waggle dance use for each site against the first principal component derived from a Partial Least Squares analysis of land-use type.
+Beta regression shows the relationship (black line) between first principal component and waggle dance use, with 95% CI shown by the grey shaded area for agri-rural (A) and urban landscapes (C).
+The correlations between first principal component and each land-use type are shown for agri-rural (B) and urban landscapes (D).
 Correlations outside the shaded area significantly contribute to the first principal component.
@@
 NAMGS) stands for non-agricultural unmanaged (resp.
-managed) green space."}
+managed) green space"}
 knitr::include_graphics("../results/figures/manuscript_figures/fig5.png")
@@
 In the urban environment, our findings were more equivocal.
-The first principal component explained ~73% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.73, $\phi$ = 10.4, p < 0.05, Fig.
-5c).
-This correlated negatively with the dominant land-use type by coverage, sparse residential (land coverage ~35%; Table 1), as would be expected if the proportion of scouts decreases when rich forage is available, since such land includes a high proportion of gardens.
+The first principal component explained ~73% of the variance in waggle dance use (beta regression: $R^2$ = 0.73, $\phi$ = 10.4, p < 0.05, Fig.
+5C).
+This correlated negatively with the dominant land-use type by coverage, sparse residential (land coverage ~35%; Table 1), as would be expected since such land includes a high proportion of gardens.
 However, this relationship is very sensitive to site removal through the jackknife sampling (Supplementary Fig.
 2.), and is not a significant contributor to the variance in the loadings (Fig.
-5c.).
-Furthermore, the first principal component also correlates positively with parks, allotments and cemeteries (8% of land cover; also railways but these constitute ~1% of land cover; Table 1), which would typically be considered forage-rich for honeybees [@Baldock2019] (although note that the amount of land cover for this component is limited and the jackknifed PLS identified substantial variation in the loadings identified for the first principal components; Supplementary Fig.
+5D.).
+The first principal component also correlates positively with railways but these constitute ~1% of land cover (Table 1).
+Furthermore, the first principal component correlates negatively with parks, allotments and cemeteries which would typically be considered forage-rich for honeybees [@Baldock2019] (although note that the amount of land cover for this component is limited and the jackknifed PLS identified substantial variation in the loadings identified for the first principal components; Supplementary Fig.
 3.).
 
-# Discussion New
+# Discussion
 
@@
 
-As our estimated proportion of scouts is continuos on the interval $[0, 1]$ we used the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010].
+As our estimates of waggle dance use is continuos on the interval $[0, 1]$ we used the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010].
 As the betareg package only works on the open interval $(0, 1)$ the data, $x$, was transformed using the following equation: $\frac{x(n-1)+0.5}{n}$, as outlined in the betareg package documentation.

### 2022-07-27 analysis/manuscript/paper.Rmd
fixed typos

@@
 title: "Honeybees vary collective decision making in different landscapes. "
-# author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent A. A. Jansen"
+# author: "Joseph Palmer, Ash Samuelson, Rich Gill, Elli Leadbeater and Vincent A. A. Jansen"
 output:
@@
 
-Honeybee foraging is an extraordinary collective behaviour that is directed by the waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops.
-Here, we present a mathematical framework to quantify how use of the waggle dance shifts in different landscape types.
-By decoding dances inside hives, we infer the proportion that are performed by bees that forage individually, and those that forage collectively.
-By applying this methodology to data from twenty hives, we show our model closely fits real-world honeybee foraging patterns and we demonstrate that colonies vary their use of waggle dance information across different landscapes.
-Our methodology provides a tool to identify the ecological conditions in which honeybee colonies rely on dance communication, opening the door to large-scale experimental exploration of the selection pressures that may have driven the evolution of this remarkable collective behaviour.
+Collective behaviours emerge from the interactions between individuals which vary in their use of different sources of information.
+In social animals, collective behaviours emerge from individuals following behavioural rules which produce a collective level response, yet some conditions make it more profitable for individuals to abandon such structures and use individual rather than collective information.
+In large animal groups, where the ability to track and analyse the interactions between individuals is limited, our understanding of how decisions influence collective processes is largely theoretical.
+Empirically evaluating how individual decisions drive group level behaviours is an important objective with wider consequences for our understanding of the evolution of such complex collective behaviours.
+Here, we explore information use in a model social insect, the honey bee (*Apis* Spp.) and show empirically how variation in individual honey bee's decision making influences collective decision making at a colony level.
+By developing a method to assess a colonies use of the famous waggle dance, we quantify to what extent colonies use the dance to forage collectively and find collective foraging using the dance varies widely over different landscapes.
+These results demonstrate that collective decision making varies in response to environmental conditions.
+Our methods provide new insights into how individual variation in information use shapes an emergent collective process and provides opportunities to further probe the environmental drivers of collective foraging.
 
@@
 Understanding the collective ramifications of individual variation in information use is not only important for deciphering how social animals collectively regulate information use [@Gruter2014], relating this to the environmental context of the behaviour is key to illuminating the ultimate evolutionary drivers of collective processes [@Price2015].
-However, establishing the group level effects of individual variation in information use is not trivial and requires measuring information use across animal groups, an often laborious process which requires monitoring individuals' behaviours and interactions.
-This becomes logistically impossible in very large animal groups, such as in social insects, which display some of the most intricate group decision-making mechanisms and consist of often thousands of individuals, each using a mix of individual and collective information to make decisions [@Bonabeau1997].
+However, establishing the group level effects of individual variation in information use is not trivial and requires measuring information use across animal groups; an often laborious process which requires monitoring individuals' behaviours and interactions.
+This becomes logistically impossible in social insects, which display some of the most intricate group decision-making mechanisms and consist of often thousands of individuals, each using a mix of individual and collective information to make decisions [@Bonabeau1997].
 
@@
 The waggle dance provides a model system to study the group effects of collective and individual information use.
-Although it is likely that individual variation in dance use translates into variations in collective decision making at the colony level [@Donaldson-Matasci2012;@Price2019], this dynamic has not been empirically established.
-Individual based models, which explore the collective effects of individuals following different information streams provide an insight into how individual behaviours scale to the level of the group [@Dornhaus2006;Beekman2008], however, these are difficult to empirically relate to real observations and so are limited to showing what a colony under certain conditions might do, rather than quantifying what they are doing.
+Although it is likely that individual variation in dance use translates into variations in collective decision making at the colony level [@Donaldson-Matasci2012;@Price2019], this has not been empirically established.
+Individual based models, which explore the collective effects of individuals following different information streams provide an insight into how individual behaviours scale to the level of the group [@Dornhaus2006;Beekman2008], however, these are difficult to relate to empirical observations and so are limited to showing what a colony under certain conditions might do, rather than quantifying what they are doing.
 Empirically determining how dance information is used at a colony level requires new methods which can identify the behaviours of individual foragers and integrate them with the mechanisms by which the dance functions to make collective decisions.
@@
 Note the difference in the shape of the distributions.
-The scout distribution is best fit by an exponential (gray fit line), the recruit distribution by a Rayleigh distribution (black line)"}
+The scout distribution is best fit by an exponential (grey fit line), the recruit distribution by a Rayleigh distribution (black line)"}
 knitr::include_graphics("../results/figures/manuscript_figures/fig2.png")
@@
 
-This description intrinsically captures honeybee foraging as collective decision making, whereby the foraging sites represented on the dance floor derive from a mixture of individual search and waggle dance information, modified by the profitability rule that biases recruitment towards closer or richer patches.
+This description intrinsically captures the collective decision making underlying honeybee foraging, whereby the foraging sites represented on the dance floor derive from a mixture of individual search and waggle dance information, modified by the profitability rule that biases recruitment towards closer or richer patches.
 The extent to which collective decision making is used is expressed in the proportion of scout dances, $p$, and the proportion of recruit dances, $1-p$.
@@
 The footage of the dances was decoded manually [@Couvillon2012;@Schurch2019;@Samuelson2021] to extract waggle run durations.
-Using this data, for each site we fit both the collective and individual models and used model selection to determine which provided the better explanation of the data, and (if the collective model provided a better fit) to quantify the relative use of social information through estimating the parameter $p$.
-In each case, We calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004;@Clauset2009].
+Using this data, for each site we fitted both the collective and individual models and used model selection to determine which provided the better explanation of the data, and (if the collective model provided a better fit) to quantify the relative use of social information through estimating the parameter $p$.
+In each case, we calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004;@Clauset2009].
 
@@
 Our findings demonstrate this empirically.
-Individual-based models [@Dornhaus2006;Beekman2008;@Price2019] have outlined how variations at an individual level could drive group level dynamics, however, without verifying this upon empirical data the insights remain theoretical.
+Individual-based models [@Dornhaus2006;@Beekman2008;@Price2019] have outlined how variations at an individual level could drive group level dynamics, however, without verifying this upon empirical data the insights remain theoretical.
 Here, we have shown how an individual based model can be used to better understand a system in a way which allows us to build more informative models which can be empirically verified.
@@
 
-Whilst the benefits of dance communication are known to vary across environments [@Donaldson-Matasci2012;@Gruter2011;@Schurch2014;@Price2019], our results are the first to show colonies vary it's use in response, affording the colony a flexible dance use strategy which they can tailor to different conditions to maximise foraging success.
+Whilst the benefits of dance communication are known to vary across environments [@Donaldson-Matasci2012;@Gruter2011;@Schurch2014;@Price2019], our results show that colonies vary their use in response, affording the colony a flexible dance use strategy which they can tailor to different conditions to maximise foraging success.
 Identifying the circumstances in which dance communication is important have been hampered by the requirement to monitor real-world colony weight [@Sherman2002;@Dornhaus2004;Gruter2013;@Price2019;@Dornhaus2006;@Nurnberger2017;@Kirchner1998] -a noisy proxy of foraging success that is influenced by many other abiotic and biotic factors, including colony health- over ecologically credible time periods [@Schurch2014].
@@
 We thus cannot rule out that longer term effects not captured in the data used in this study, such as resource stability [@Schurch2014], may also have contributed to our estimates of waggle dance use.
-The number of colonies used in our analysis, although comparable to studies in which the benefits of collective foraging have been evaluated [@Sherman2002;@Dornhaus2004;@Price2019;@Nurnberger2017;Kirchner1998], limits the statistical power of our analysis.
+The number of colonies used in our analysis, although comparable to studies in which the benefits of collective foraging have been evaluated [@Sherman2002;@Dornhaus2004;@Price2019;@Nurnberger2017;@Kirchner1998], limits the statistical power of our analysis.
 Despite this, we gain useful insights into the drivers of this relationship by exploring how land-use type influences waggle dance use.
@@
 Our results within the agri-rural sites broadly agree with the findings from other theoretical [@Dornhaus2006] (11) and empirical [@Price2019] studies which examined foraging performance with and without the dance.
-These results suggest that, even though collective foraging is not always beneficial, in environments where high quality resources are present but relatively scarce, collective foraging is being carried out by colonies and under these conditions exchanging social information through the waggle dance confers benefits.
+These results suggest that, even though collective foraging is not always beneficial, in environments where high-quality resources are present but relatively scarce, collective foraging is being carried out by colonies and under these conditions exchanging social information through the waggle dance confers benefits.
 In both agri-rural and urban landscapes we have shown that the waggle dance use changes with land-use (illustrated in the change in land-use shown in the maps from left to right in Fig 5A and 5C).
@@
 Each straight line in the random path is converted to a rectangle with length equal to the path section length and a constant width of ~0.01 to represent an area the scout searches along that path.
-Of all the resources contained in the boxes drawn from the scout's path, the one closest to the colony is selected as the resource patch that the scout will report and will communicate its location if the quality of the resource resource exceeds a minimum threshold.
+Of all the resources contained in the boxes drawn from the scout's path, the one closest to the colony is selected as the resource patch that the scout will report and will communicate its location if the quality of the resource exceeds a minimum threshold.
 Communication is simulated by pooling together all the resource patches found.
@@
 3: upon visiting a resource patch, foragers translate the profitability of a resource into the number of repeats of the dance.
-The number of repeats of the dance is a function of quality an distance.
+The number of repeats of the dance is a function of quality and distance.
 Recruits sample random dances and report the location of successful visits to resource patches on the dance floor.
@@
 
-For the simplified the model we assumed that the number of dances depends weakly on distance and there is a sizable quality differences between resources of a non-negligible size and that there is a sizable intensity of the high quality resource (See Supplementary Material for detailed derivation).
+For the simplified the model we assumed that the number of dances depends weakly on distance and there is a sizable quality differences between resources of a non-negligible size and that there is a sizable intensity of the high-quality resource (See Supplementary Material for detailed derivation).
 Foragers on scouting trips are more likely to report larger distances than foragers on recruiting trips.
@@
 
-Recruit trips will be predominantly to high quality resources.
-Only if the nearest high quality patch is very far away will there be a more profitable patch of lesser quality available, and this happens only rarely if the intensity of the best quality resource is sizable.
+Recruit trips will be predominantly to high-quality resources.
+Only if the nearest high-quality patch is very far away will there be a more profitable patch of lesser quality available, and this happens only rarely if the intensity of the best quality resource is sizable.
 After linearising the function that translates the profitability into the number of waggle dance runs for short dance durations and normalising the distribution of dance durations reported from recruit trips in the simplified model is:
@@
 
-is the normalisation factor, the parameter $a_r$ is the rate with which dances repeats depends on distance for recruit trips and $a_r^2 b_r$ is the intensity of high quality resources reported by recruited foragers.
+is the normalisation factor, the parameter $a_r$ is the rate with which dances repeats depends on distance for recruit trips and $a_r^2 b_r$ is the intensity of high-quality resources reported by recruited foragers.
 
@@
 
-As our estimates of waggle dance use is continuos on the interval $[0, 1]$ we used the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010].
+As our estimates of waggle dance use is continuous on the interval $[0, 1]$ we used the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010].
 As the betareg package only works on the open interval $(0, 1)$ the data, $x$, was transformed using the following equation: $\frac{x(n-1)+0.5}{n}$, as outlined in the betareg package documentation.

### 2022-07-27 analysis/manuscript/paper.Rmd
updated with required changes

@@
 Although it is likely that individual variation in dance use translates into variations in collective decision making at the colony level [@Donaldson-Matasci2012;@Price2019], this has not been empirically established.
-Individual based models, which explore the collective effects of individuals following different information streams provide an insight into how individual behaviours scale to the level of the group [@Dornhaus2006;Beekman2008], however, these are difficult to relate to empirical observations and so are limited to showing what a colony under certain conditions might do, rather than quantifying what they are doing.
+Individual based models, which explore the collective effects of individuals following different information streams provide an insight into how individual behaviours scale to the level of the group [@Dornhaus2006;@Beekman2008], however, these are difficult to relate to empirical observations and so are limited to showing what a colony under certain conditions might do, rather than quantifying what they are doing.
 Empirically determining how dance information is used at a colony level requires new methods which can identify the behaviours of individual foragers and integrate them with the mechanisms by which the dance functions to make collective decisions.
@@
 Thus, in some environments, the majority of foraging trips involve scouting to find new food sites rather than recruitment through dances.
-Note that this does not imply that these bees do not engage in dance following, because bees regularly follow dances but choose not to visit the advertised site [@Gruter2011], although it has also been shown that bees may cease dance following if it is proving unproductive [@Gruter2011; @Price2019].
+Note that this does not imply that these bees do not engage in dance following, because bees regularly follow dances but choose not to visit the advertised site [@Gruter2011], although it has also been shown that bees may cease dance following if it is proving unproductive [@Gruter2011;@Price2019].
 Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individual's foraging behaviour over a landscape.
@@
 As for one of our sites neither model provided a plausible description, this site was removed from the PLS analysis.
-Due to our small sample size (10 urban and 9 agri-rural sites), we used jackknife resampling to evaluate the robustness of our results to influential points (see Materials and Methods, Supplementary Material).
 
-In the agri-rural environments the first principal component is a combination of land use types that explained ~73% of the variation in waggle dance use (beta regression: $R^2 = 0.73$, $\phi$ = 4.9, p < 0.05, Fig.
+Waggle dance use varied widely regardless of environment, ranging from 0 to 1 in both the urban and agri-rural environments whilst showing a significant, positive interaction with waggle dance use.
+In the agri-rural sites, a single principle component explained ~73% of the variation in waggle dance use (beta regression: $R^2 = 0.73$, $\phi$ = 4.9, p < 0.05, Fig.
 5A).
-This principal component correlates positively with arable land (29% of land coverage; Table 1) and negatively with built-up areas (17% of land coverage); note that it also correlates negatively with non-agricultural unmanaged green space and water, but together these represent less than 3% of land-use (Table 1; Fig.
+This finding is also shown in the urban environment which also shows a single principle component explained ~73% of the variation in waggle dance use (beta regression: $R^2$ = 0.73, $\phi$ = 10.4, p < 0.05, Fig.
+5C).
+Probing the individual loadings of these principal components allows us to evaluate the driving mechanisms in more detail.
+As our sample size is limited  (10 urban and 9 agri-rural sites), we used jackknife resampling to evaluate the robustness of our results to influential points (see Materials and Methods, Supplementary Material) and evaluate how these land-use types contribute to the principle component which significantly correlates with waggle dance use.
+
+```{r, fig.cap="Figure 5.
+**Collective foraging correlates with land-use**.
+Estimated waggle dance use for each site against the first principal component derived from a Partial Least Squares analysis of land-use type.
+Beta regression shows the relationship (black line) between first principal component and waggle dance use, with 95% CI shown by the grey shaded area for agri-rural (A) and urban landscapes (C).
+The correlations between first principal component and each land-use type are shown for agri-rural (B) and urban landscapes (D).
+Correlations outside the shaded area significantly contribute to the first principal component.
+Colours correspond to the land use as shown in maps (circular insets) for selected sites.
+NAUMGS (resp.
+NAMGS) stands for non-agricultural unmanaged (resp.
+managed) green space"}
+knitr::include_graphics("../results/figures/manuscript_figures/fig5.png")
+```
+
+In the agri-rural environments the first principal component correlates positively with arable land (29% of land coverage; Table 1) and negatively with built-up areas (17% of land coverage); note that it also correlates negatively with non-agricultural unmanaged green space and water, but together these represent less than 3% of land-use (Table 1; Fig.
 5B).
@@
 Arable land in the UK is typically considered nutritionally poor for bees (note that oilseed rape fields were not included within this category), while there is evidence to suggest that the residential areas that were captured within the “built-up” category are forage-rich hotspots, typically supporting relatively high bee diversity and abundance within gardens [@Baldock2019].
+
+In the urban environment, our findings were more equivocal.
+The first principal component correlated negatively with the dominant land-use type by coverage, sparse residential (land coverage ~35%; Table 1), as would be expected since such land includes a high proportion of gardens.
+However, this relationship is very sensitive to site removal through the jackknife sampling (Supplementary Fig.
+2.), and is not a significant contributor to the variance in the loadings (Fig.
+5D.).
+The first principal component also correlates positively with railways but these constitute ~1% of land cover (Table 1).
+Furthermore, the first principal component correlates negatively with parks, allotments and cemeteries which would typically be considered forage-rich for honeybees [@Baldock2019] (although note that the amount of land cover for this component is limited and the jackknifed PLS identified substantial variation in the loadings identified for the first principal components; Supplementary Fig.
+3.).
 
@@
 
-```{r, fig.cap="Figure 5.
-**Collective foraging correlates with land-use**.
-Estimated waggle dance use for each site against the first principal component derived from a Partial Least Squares analysis of land-use type.
-Beta regression shows the relationship (black line) between first principal component and waggle dance use, with 95% CI shown by the grey shaded area for agri-rural (A) and urban landscapes (C).
-The correlations between first principal component and each land-use type are shown for agri-rural (B) and urban landscapes (D).
-Correlations outside the shaded area significantly contribute to the first principal component.
-Colours correspond to the land use as shown in maps (circular insets) for selected sites.
-NAUMGS (resp.
-NAMGS) stands for non-agricultural unmanaged (resp.
-managed) green space"}
-knitr::include_graphics("../results/figures/manuscript_figures/fig5.png")
-```
-
-In the urban environment, our findings were more equivocal.
-The first principal component explained ~73% of the variance in waggle dance use (beta regression: $R^2$ = 0.73, $\phi$ = 10.4, p < 0.05, Fig.
-5C).
-This correlated negatively with the dominant land-use type by coverage, sparse residential (land coverage ~35%; Table 1), as would be expected since such land includes a high proportion of gardens.
-However, this relationship is very sensitive to site removal through the jackknife sampling (Supplementary Fig.
-2.), and is not a significant contributor to the variance in the loadings (Fig.
-5D.).
-The first principal component also correlates positively with railways but these constitute ~1% of land cover (Table 1).
-Furthermore, the first principal component correlates negatively with parks, allotments and cemeteries which would typically be considered forage-rich for honeybees [@Baldock2019] (although note that the amount of land cover for this component is limited and the jackknifed PLS identified substantial variation in the loadings identified for the first principal components; Supplementary Fig.
-3.).
-
 # Discussion
@@
 Here, we have presented a method to quantify information use at a collective level in a model social insect.
-Using this method, we have shown that collective decision-making is used variably.
+After verifying that this method accurately captures empirical patterns of collective and individual foraging, we have shown how it can be used to evaluate how environmental factors shape collective foraging at the colony level.
 Whilst recruitment is well known to occur through other mechanisms besides the waggle dance, such as through olfactory cues [@Arenas2007], the exceptionally close fit of our model to waggle run durations underlines the importance of the waggle dance in honey bee foraging.
@@
 
-The methods described here provide a framework to start exploring how plasticity in collective decision making can help to buffer the effects of landscape change, and in this case, how flexibility in the use of the waggle dance may help to exploit resources when experiencing novel landscape mosaics, such as in highly urbanized conurbations.
-As we used an existing dataset, we did not systematically choose sites based on forage availability, and instead inferred forage availability based on land-use at each site.
-The dance recordings in our dataset were collected over an extended period of five months, over which time forage availability in the landscape likely changed considerably and non-uniformly across sites.
-We thus cannot rule out that longer term effects not captured in the data used in this study, such as resource stability [@Schurch2014], may also have contributed to our estimates of waggle dance use.
-The number of colonies used in our analysis, although comparable to studies in which the benefits of collective foraging have been evaluated [@Sherman2002;@Dornhaus2004;@Price2019;@Nurnberger2017;@Kirchner1998], limits the statistical power of our analysis.
-Despite this, we gain useful insights into the drivers of this relationship by exploring how land-use type influences waggle dance use.
-
-Our agri-rural results suggest a potential decrease in reliance on waggle dance recruitment as resources become harder to find and foraging trips become longer [@Samuelson2021].
-Our results within the agri-rural sites broadly agree with the findings from other theoretical [@Dornhaus2006] (11) and empirical [@Price2019] studies which examined foraging performance with and without the dance.
+Examining the land-use types which contribute most to the variation in waggle dance use in each of our study environments reveals insights into the drivers of waggle dance use.
+Our agri-rural results suggest a potential decrease in reliance on waggle dance recruitment as resources become harder to find and foraging trips become longer [@Samuelson2021], broadly agreeing with the findings from other theoretical [@Dornhaus2006] (11) and empirical [@Price2019] studies which examined foraging performance with and without the dance.
 These results suggest that, even though collective foraging is not always beneficial, in environments where high-quality resources are present but relatively scarce, collective foraging is being carried out by colonies and under these conditions exchanging social information through the waggle dance confers benefits.
@@
 
-The analysis of waggle dance data can make an important contribution to our understanding of information use.
+Although the number of colonies used in this analysis are comparable to similar studies evaluating collective foraging in honey bees [@Sherman2002;@Dornhaus2004;@Price2019;@Nurnberger2017;@Kirchner1998],  probing the landscape drivers of colony level waggle dance use in more detail requires more colony level data.
+As longer term effects, such as resource stability [@Schurch2014], may also influence waggle dance use, these data need to be gathered over sufficient time frames to ensure these effects are accounted for.
 Recently, technological advances have emerged which enable colony metrics to be collected faster, more accurately and over greater time spans than could be gathered by hand, allowing individuals to be tracked within colonies and theories of individual behaviour to be evaluated in more depth than could have been done previously [@Wild2021].
-By piecing together the behavioural response of individuals and combining these with landscape analyses, we have found an accurate mathematical description of colony foraging which extends our ability to quantify collective behaviour across environments.
+Here, we have presented an accurate mathematical description of colony foraging which extends our ability to quantify collective behaviour across environments.
 With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”.
-Our methodology thus provides a means of analysing such large data sets to inform the debate about the importance of collective decision making, as well as providing useful colony metrics of foraging activity.
+Our methodology thus provides a means of analysing such large data sets and provide a framework to start exploring how plasticity in collective decision making can help to buffer the effects of landscape change, and in this case, how flexibility in the use of the waggle dance may help to exploit resources when experiencing novel landscape mosaics, such as in highly urbanized conurbations.
 

### 2022-08-02 README.Rmd
inital commit to store as feature branch
tidied up analysis code and moved reference to the paper into inst

@@
 
-The goal of wagglefit is to allow for easy waggle dance analysis.
+wagglefit is an R package to analyse waggle run distributions and estimate the resulting proportions of scout and recruit dances, providing an estimate of dance use in undisturbed colonies.
 
@@
 
-You can install the released version of wagglefit from [CRAN](https://CRAN.R-project.org) with:
+You can install the development version from [GitHub](https://github.com/) with:
 
 ``` r
-install.packages("wagglefit")
+install.packages("devtools")
+devtools::install_github("joseph-palmer/wagglefit")
 ```
 
-And the development version from [GitHub](https://github.com/) with:
+when citing wagglefit please use (insert citation for publication).
 
-``` r
-# install.packages("devtools")
-devtools::install_github("joseph-palmer/wagglefit")
-```
-## Example
+## Honey bees vary collective decision making in different landscapes
 
-This is a basic example which shows you how to solve a common problem:
+All code used in the publication (insert citation link) is stored in `inst/publication/`.
 
-```{r example}
-library(wagglefit)
-## basic example code
-```
+## Reporting issues
+
+If you notice any issues with this code please raise an issue in the github repo.

### 2022-08-02 inst/publication/Code/agent_based_sim.Rmd
tidied up analysis code and moved reference to the paper into inst

--- 2022-04-05 analysis/agent_based_sim.Rmd
@@
 
-The code for the honeybee foraging simulation is writen in python and is contained in the `Simulation` directory (specifically `Simulation/Code`).
-The results of the simulation (as output csv files) are stored in `Results/Results`.
+The code for the honeybee foraging simulation is written in python and is contained in the `Simulation` directory (specifically `Simulation/Code`).
+The results of the simulation (as output csv files) are stored in `Results/`.
 Generating the results can be done by running the script `run.py` in `Simulation/Code`.
 Once you have ran the simulation you can run through this Rmd doc to generate the figures.
-To run the python code you will need python 3 installed.
+To run the python code you will need python 3 installed along with the required packages.
 
@@
 recruit <- read.csv(
-  "Simulation/Results/recruit_distribution_v2.csv",
+  "Simulation/Results/recruit_distribution.csv",
   header = FALSE
@@
 scout <- read.csv(
-  "Simulation/Results/scout_distribution_v2.csv",
+  "Simulation/Results/scout_distribution.csv",
   header = FALSE
@@
 ```
-From this analysis we observe that the distribution of foraging distances reported by scouts is better described by an exponential than the minimum of an exponential, whereas the converse is true for foraging distances reported by reruits.
+From this analysis we observe that the distribution of foraging distances reported by scouts is better described by an exponential than the minimum of an exponential, whereas the converse is true for foraging distances reported by recruits.
 These results support our hypothesis for how scout and recruit reported foraging distances occur through different processes and produce different patterns.

### 2022-08-02 inst/publication/Manuscript/paper.Rmd
tidied up analysis code and moved reference to the paper into inst

--- 2022-07-27 analysis/manuscript/paper.Rmd
@@
 ---
-title: "Honeybees vary collective decision making in different landscapes. "
+title: "Honey bees vary collective decision making in different landscapes"
 # author: "Joseph Palmer, Ash Samuelson, Rich Gill, Elli Leadbeater and Vincent A. A. Jansen"

### 2022-08-02 inst/publication/Manuscript/supplementary_material.tex
tidied up analysis code and moved reference to the paper into inst

\documentclass[12pt,letter]{amsart}
%\usepackage[dvips]{graphics}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{caption}
%\usepackage{vruler}

\parindent0.5cm

\oddsidemargin1mm\evensidemargin1mm
 \topmargin-5mm
\textheight240mm\textwidth160mm

\def\baselinestretch{1.1}
\newcommand{\ignore}[1]{}
\newcommand{\df}[0]{{\,\rm d }}
\newcommand{\vectpsi}[0]{\mbox{\boldmath $\psi$}}
\newcommand{\vecttau}[0]{\mbox{\boldmath $\tau$}}
\newcommand{\T}[0]{{\bf T}}
\newcommand{\A}[0]{{\bf A}}
\newcommand{\B}[0]{{\bf B}}
\newcommand{\I}[0]{{\bf I}}
\newcommand{\z}[0]{{\bf z}}
\newcommand{\x}[0]{{\bf x}}
\newcommand{\e}[1]{{\bf e}_#1}
\newcommand{\one}[0]{{\bf 1}}
\newcommand{\diag}[1]{{\rm \bf diag}(#1)}

\renewcommand{\thesection}{\Alph{section}}
\pagestyle{plain}

\DeclareCaptionLabelFormat{sfig}{#1 S#2}
\captionsetup[figure]{labelformat=sfig}

\begin{document}

%\advance\textwidth 3cm

\vspace*{-7mm}

\title{Waggle dance distributions quantify collective foraging in honey bee colonies: Supplementary material}
%\author{}
\maketitle

%\begin{center}
%authors,  Vincent A.A. Jansen
%\vspace*{7mm}
%{\bf title}
%\vspace*{11mm}
 %{\bf \underline{Online Supplementary Text and Methods}}
%\end{center}
%\vspace*{11mm}
%\pagebreak
%\setvruler [][][][][][\textwidth]

\section*{The derivation of the full model for the distribution of distances on the dance floor. }

We will build a simple model to describe the distribution of distances reported on the dance floor. We assume that there are $n$ different resources and they are distributed according to a spatial Poisson point process. The Poisson point process places food sources randomly in the environment of the hive with intensity $\lambda_i$ for resource $i$. The quality of that resource we will denote $q_i.$ For convenience, we will order the resources $q_i$ from small to large such that $q_j \ge q_i$ if $j>i$, and $q_n$ is the resource with the highest quality, which we assume is always present in the environment, hence $\lambda_n>0$, and has a resource quality exceeding unity, hence $q_n>1$. Under this ordering, we define $m$, for which $1\le m\le n$, as the lowest index for which $q_m>1$ hence, $q_i<1$ for all $i=1, \dots, m-1$

\subsection*{The distribution of scout dances}

As a scout travels a small distance $\df x$ its search will cover an area $c_s \df x.$ The constant $c_s$ depends on width of the area scanned, the detection rate and the degree to which a the forager's path covers new ground. If we incorporate these factors into the constant $c_s$ the intensity of discovering a resource over a path of distance $\df x$ is  $c_s \df x$. The probability density of $x$, the distance travelled after which the first source is discovered follows a distribution
$$\lambda_s e^{-\lambda_s x}$$
where $\lambda_s= c_s\sum_{i =1}^n\lambda_i.$

The profitability, which depends on quality and distance, of the source dictates the number of repeats of the dance.
What is the distribution of dances on the dance floor?  Let's assume that there is a function that translates the profitability into the expected number of runs of the waggle dance, $\phi(q_i, x)$,  which is a function of quality an distance. We assume that if this function has a non-zero value it decreases with distance, and increases with quality, hence $\frac{\partial \phi(q, x) }{\partial q}>0$ and $\frac{\partial \phi(q, x) }{\partial  x}<0$ if $\phi(q,x)>0$. There is a minimum quality level of the resource that elicits dancing, and we have set this level to unity, such that if $q \le 1$ then $\phi(q, x)=0$ for $x \ge 0$. It follows that for each $q_i>1$ there is a maximum distance $x_{i, {\rm max}}>0$ for which bees do not dance, that is $\phi(q_i, x)=0$ for $x \ge x_{i, {\rm max}}.$

The expected number of dances for a resource of type $i$ located at distance $x$ is $c_s \lambda_i \phi(q_i, x) e^{-\lambda_s x}.$ The expected total number of scout waggle dances on the dance floor is given by $$M_s=\sum_{i=m}^n \int_{0}^\infty \frac{c_s \lambda_i}{\lambda_s} \phi(q_i, x) \lambda_s e^{-\lambda_s x} \df x
$$ and the distribution of scout dances is
$$\frac{\lambda_s e^{-\lambda_s x}\sum_{i=m}^n  \frac{c_s \lambda_i}{\lambda_s} \phi(q_i, x)  }{M_s}
$$

If the bees go out, scout for resources and dance for them this, and there is no further processing of information, this would be the distribution one would expect.

\subsection*{The distribution of recruit dances}

First, we consider a simple scenario and we assume that all resources are of equal quality $q$ with $q>1$. The scouts that report back on the dance floor cover the surroundings of the hive in their forays. How well they cover, depends on the number of scouts, how well their collective paths cover the area and chance of discovery of an item. Let the fraction of sources that can be discovered by the collection of scouts be $c_r$. We assume $0\le c_r \le 1.$

The chance of discovering food source within a radius of $x$ is $e^{-c_r\lambda_r\pi x^2}$, where $\lambda_r= \sum_{i=1}^n \lambda_i  $ and therefore the chance of having the nearest food source that the recruits report is least a distance $x$ away is $P(\underline x \ge x)=e^{-c_r\lambda_r \pi x^2}.$ The distribution of the nearest food source that recruits report is $$2 c_r\lambda_r \pi x e^{-c_r\lambda_r \pi x^2}.$$ The distribution of recruit dances on the dance floor is:
$$\frac{\phi(q,x)2 c_r\lambda_r \pi x e^{-\lambda_r \pi x^2}}{M_r}$$ where
$$M_r=\int_0^\infty \phi(q,x)2c_r \lambda_r \pi x e^{-c_r\lambda_r \pi x^2} \df x.$$

If the resources differ in quality, the same rationale can be applied. Consider two locations that differ in resource quality and distance, the hive can prefer the further location if it is sufficiently better. Likewise, it can prefer a lesser quality location if it is sufficiently nearer. Given that the hive has found a resource at distance $x$ of quality $q_i$ we define a threshold distance $\xi_{ij}(x)$ at which a resource of type $q_j$ has the same profitability. If  $q_j \ge q_i,$ and $0 \le  x \le x_{i,\max}$ then $\xi_{ij}(x)$ satisfies
\begin{equation*}
\left \{
\begin{array}{ll}
\phi(q_i,x)=\phi(q_j, \xi_{ij}(x)) & \text{if }0 \le  x<x_{i,\max}\\
\xi_{ij}(x_{i,\max})=x_{j,\max} & \text{if } x= x_{i,\max}
\end{array}
\right.
\end{equation*}
If  $q_j<q_i,$ and $x<x_{i,\max}$ then $\xi_{ij}$ satisfies
\begin{equation*}
\left \{
\begin{array}{ll}
\xi_{ij}(x)=0& \text{if } 0 \le x \le \xi_{ji}(0) \\
\phi(q_i,x)=\phi(q_j, \xi_{ij}(x)))& \text{if } \xi_{ji}(0)<x_{i,\max}\\
\xi_{ij}(x_{i,\max})=x_{j,\max} & \text{if } x= x_{i,\max}
\end{array}
\right.
\end{equation*}

This relation defines the threshold distance $\xi_{ij}(x)$ as a function of $x$ for $0 \le x \le x_{i,{\rm max}}$.
%It follows from implicit differentiation that if ??? $\frac{\df \xi_{ij}}{\df q_j}=- \frac{\df \phi(q_j,x)/\df q_j}{\df \phi(q,x)/\df x}>0$ and $\frac{\df \xi_{ij}}{\df %q_i}= \frac{\df \phi(q_i,x)/\df q_i}{\df \phi(q,x)/\df x}<0.$
It follows that $\xi_{ii}(x)=x$. Because $\phi(q_i,x)$ increases with $q$ if $\phi(q_i,x)>0$ it follows that $\phi(q_i,x)>\phi(q_j,x)$ if $q_i>q_j$. As a consequence $\xi_{ij}<x$ if $i>j$ and $\xi_{ij}>x$ and if $i<j.$

The probability to find the nearest resource $i$ at distance $x$ is given by the Rayleigh distribution $2 \pi c_r \lambda_i x e^{-\pi c_r \lambda_i x^2 }.$ For $x \le x_{i, {\rm max}}$ the chance of there not being a source if type $j$ within a radius of $\xi_{ij}(x)$, which is $e^{- \pi c_r \lambda_j \xi_{ij}(x)^2}.$ The probability that resource $i$ at location $x$, $0 \le x \le x_{i,\max}$ is the most profitable is therefore
$$P(\textrm{most profitable resource }i \textrm{ is at distance } x)=2 \pi c_r \lambda_i x e^{- \pi \sum_{j=1}^n\lambda_j c_r \xi_{ij}(x)^2}$$ and 0 if $x > x_{i,\max}$
and the probability of finding the most profitable location at distance $x$ is
$$P(\textrm{most profitable resource is at distance } x )=\sum_{\forall i: {x_{i, \max}>x}} 2 \pi c_r \lambda_i x e^{- \pi \sum_{j=1}^n c_r \lambda_j \xi_{ij}(x)^2}.$$
For $x>x_{n, \max}$ the most profitable resource cannot be determined, as all have zero profitability.

The probability of encountering a recruit reporting a distance $x$ on the dance floor is
$$M_r^{-1}\sum_{i=m}^n  \phi(q_i, x)\lambda_i2 \pi c_r  x e^{- \pi \sum_{j=1}^n c_r \lambda_j \xi_{ij}(x)^2}
$$
where $M_r$ is the normalisation factor $$M_r=\int_0^\infty \sum_{i=m}^n  \phi(q_i, x)\lambda_i2 \pi c_r  x e^{- \pi \sum_{j=1}^n c_r \lambda_j \xi_{ij}(x)^2}\df x.$$

\subsection*{The full model}

The combined distribution of distances of scout and recruit dances on the dance floor is given by
$$P(\underline x= x )=p \frac{\lambda_s e^{-\lambda_s x}\sum_{i=1}^n  \frac{c_s \lambda_i}{\lambda_s} \phi(q_i, x) }{M_s} +(1-p)\frac{\sum_{i=1}^n  \phi(q_i, x)2 \pi c_r \lambda_i  x e^{- \pi \sum_{j=1}^n c_r \lambda_j \xi_{ij}(x)^2}}{M_r}.$$
This is a generic model for the distribution of distances reported on the dance floor.

\section*{The simplified model}

The full model can be used to calculate the likelihood for a given data set, which can be used to estimate parameters. The full model has $2 n+ 3 + k $  parameters: $\lambda_1 \dots \lambda_n$, $q_1 \dots q_n$, $c_s, c_r$ and $p$ plus $k$ parameters needed to specify the function $\phi$. Even though it is possible to reduce the number of parameters through a scaling, e.g. by defining $\lambda_i^r=c_r \lambda_i$ and $c=\frac{c_s}{c_r}$, but even if the number of resources is low, it turned out to be difficult to estimate the parameters even if with a fair sized data sets. To assist in the estimation of the key parameters we therefore formulated a simplified model.

We can simplify the model by assuming that the dependence of the number of dances depends less on the distance and more on quality. We assume that that dependence of the profitability assessment only weakly depends on distance and that there is quality differences between resources of a decent size, so that there are a large threshold distances at which resources have the same profitability. If that is the case then, if a resource is discovered at a small distance from the hive, resources if less quality than this one, are not worth dancing for, whereas there is a large distance over which superior resources can be detected that will have higher profitability. In mathematical terms, under this assumption, for small $x$, $\xi_{ij}(x)=0$ if $j<i$, $\xi_{jj}(x)=x$ and $\xi_{ij}(x)$ is large if $j>i$. We thus have $\sum_{j=1}^n  \lambda_j \xi_{ij}(x)=\sum_{j=i}^n  \lambda_j \xi_{ij}(x)$ for $i<n$ and $\sum_{j=1}^n  \lambda_j \xi_{nj}(x)=\lambda_n x.$ Moreover, as the $\xi_{ij}s$ will be large, $\sum_{j=i}^n  \lambda_j \xi_{ij}(x) \gg\lambda_n x.$  We can approximate
$$\sum_{i=1}^n  \phi(q_i,x)2 \pi c_r \lambda_i  x e^{- \pi \sum_{j=1}^n c_r \lambda_j \xi_{ij}(x)^2}\approx
\phi(q_n,x)2 \pi c_r \lambda_n  x e^{- \pi  c_r \lambda_n x^2}
$$
The recruits on the dance floor are most likely to report short distances. By expanding the function $ \phi(q_n, x)$ around $x=0$ we get $\phi(q_n, x)=\left[\phi(q_n, 0)+x\left. \frac{\partial \phi(q_n, x)}{\partial x}\right|_{x=0}+ O(x^2)\right]_+$ we can approximate the recruit distribution as:
$$M_r^{-1}\sum_{i=m}^n  \phi(q_i, x)\lambda_i2 \pi c_r  x e^{- \pi \sum_{j=1}^n c_r \lambda_j \xi_{ij}(x)^2}\approx  \frac{\left[1- a_r x\right]_+2 \pi a_r^2   b_r x e^{- \pi  b_r  (a_r x)^2}}{1-\frac{{\rm erf}\left(  \sqrt{\pi b_r}\right)}{2 \sqrt{b_r }}}$$
with $a_r=-\phi(q_n, 0)^{-1}\left. \frac{\partial \phi(q_n, x)}{\partial x}\right|_{x=0}$ and $b_r=c_r \gamma_n a_r^{-2}$
and where we used
$$\phi(q_n, 0)\int_{0}^{a_r^{-1}} (1- a_r x )2 \pi a_r^2   b_r x e^{- \pi  b_r  (a_r x)^2}\df x=\phi(q_n, 0)\left(1-\frac{{\rm erf}\left(  \sqrt{\pi b_r}\right)}{2 \sqrt{b_r }}\right)
$$ and cancelled the term $\phi(q_n, 0).$

The scouts on the dance floor are more likely to report large distances. For distances around the largest distance for which scouts dance, $x_{n,\max}$, the $\phi(q_i, x)=0$ for all $i<n$. We therefore expand  $ \phi(q_n, x)$ around $x_{n, \max}$ to get $ \phi(q_n, x)=\left[( x- x_{n, \max})\left.\frac{\phi(q_n, x)}{\partial x}\right|_{x=x_{n, \max}}+ O(x^2)]\right]_+$. With this, we can approximate the scout distribution as:
%$$\frac{\lambda_s e^{-\lambda_s x}  \left[  \frac{q_n - 1}{\alpha}-x\right]_+ }{\frac{q_n-1}{\alpha}-\lambda_s^{-1} \left(1-e^{-\lambda_s\frac{q_n-1}{\alpha }}\right)}$$
$$\frac{\lambda_s e^{-\lambda_s x}\sum_{i=m}^n  \frac{c_s \lambda_i}{\lambda_s} \phi(q_i, x)  }{M_s}\approx \frac{a_s b_s e^{-b_s a_s x}  \left[ 1-a_s x\right]_+ }{1-b_s^{-1} \left(1-e^{-b_s}\right)}$$
with $a_s=x_{n,\max}^{-1}$ and $b_s=\gamma_s a_s^{-1}$
and where we used
$$ \frac{c_s \lambda_i}{\lambda_s}\left.\frac{\phi(q_n, x)}{\partial x}\right|_{x=x_{n, \max}} \int_{0}^{a_s^{-1}} a_s b_s e^{-b_s a_s x}  ( 1-a_s x)  \df x= \frac{c_s \lambda_i}{\lambda_s}\left.\frac{\phi(q_n, x)}{\partial x}\right|_{x=x_{n, \max}} \left( 1-b_s^{-1} \left(1-e^{-b_s}\right)\right)$$ and cancelled the term $\frac{c_s \lambda_i}{\lambda_s}\left.\frac{\phi(q_n, x)}{\partial x}\right|_{x=x_{n, \max}}.$

The distribution then simplifies to
$$P(\underline x= x )=p \frac{\left[ 1-a_s x\right]_+a_s b_s e^{-b_s a_s x}   }{1-b_s^{-1} \left(1-e^{-b_s}\right)}+(1-p) \frac{\left[1- a_r x\right]_+2 \pi a_r^2   b_r x e^{- \pi  b_r  (a_r x)^2}}{1-\frac{{\rm erf}\left(  \sqrt{\pi b_r}\right)}{2 \sqrt{b_r }}}.$$
The simplified model has 5 parameters, and is therefore simpler than the full model for $n \ge2$, which has $2+2n+k $ parameters.

\section*{The simplified distribution with a minimum value}
 Let $m$ be the lowest distance that is recorded, for instance because there is a minimum duration of the waggle dance. We assume that $m<\xi_{ji}(0)$ for all $j<i$. For the simplified recruit distribution we now expanding the function $ \phi(q_n, x)$ around $x=m$ to get $\phi(q_n, x)=\left[\phi(q_n, m)+(x-m)\left. \frac{\partial \phi(q_n, x)}{\partial x}\right|_{x=m}+ O(x^2)\right]_+$. The simplified recruit distribution becomes:
%$$\frac{\left[\frac{q_n - 1}{ \alpha q_n}- x\right]_+2 \pi c_r \lambda_n  x e^{- \pi  c_r \lambda_n x^2}}{\left(\frac{ (q_n-1)}{\alpha  q_n}-m\right)e^{-\pi c_r %\lambda_n m^2 }+\frac{\text{erf}\left(m\sqrt{\pi c_r \lambda_n} \right)-\text{erf}\left(\frac{  (q_n-1)}{\alpha  q_n}\sqrt{\pi c_r \lambda_n}\right)}{2 \sqrt{c_r %\lambda_n}}}$$
$$ \frac{\left[1- a_r x\right]_+2 \pi a_r^2 b_r  x e^{- \pi  b_r (a_r x)^2}}{\left(1-a_r m\right)e^{-\pi b_r (a_r m)^2 }+\frac{\text{erf}\left(a_r \sqrt{\pi b_r}m \right)-\text{erf}\left(\sqrt{\pi b_r}\right)}{2 \sqrt{b_r}}}$$
with $a_r=m-\phi(q_n, m)^{-1}\left. \frac{\partial \phi(q_n, x)}{\partial x}\right|_{x=m}$ and $b_r=c_r \gamma_n a_r^{-2}.$

The scout distribution is only affected through the normalisation factor and becomes:
$$
a_s\frac{ b_s e^{-b_s a_s (x-m)}  \left[ 1-a_s x\right]_+ }
{\left(1-a_s m\right)-b_s^{-1}\left(1-e^{-b_s\left(1-a_s m\right)}\right)}$$

The distribution, for $ x \ge m$ then becomes
\begin{eqnarray*}P(\underline x= x )&=&pa_s\frac{ b_s e^{-b_s a_s (x-m)}  \left[ 1-a_s x\right]_+ }
{\left(1-a_s m\right)-b_s^{-1}\left(1-e^{-b_s\left(1-a_s m\right)}\right)}\\&& \qquad +(1-p) \frac{\left[1- a_r x\right]_+2 \pi a_r^2 b_r  x e^{- \pi  b_r (a_r x)^2}}{\left(1-a_r m\right)e^{-\pi b_r (a_r m)^2 }+\frac{\text{erf}\left(a_r \sqrt{\pi b_r}m \right)-\text{erf}\left(\sqrt{\pi b_r}\right)}{2 \sqrt{b_r}}}.
\end{eqnarray*}
The same function describe the distribution of the dance durations, where dances have a minimum duration. This assumes a linear relation between duration and distance.

\section*{Model fits to all sites}

\captionof{table}{Fit results for the fitting of the waggle dance model to each site.}
\input{../results/figures/SI_figures/fittable.tex}

\vspace*{50px}

\begin{figure}[h]
	\centering
	\includegraphics{../results/figures/SI_figures/allsites_plot.png}
	\caption{\textbf{Cumulative frequency distribution} of waggle dance durations and fits of the collective (red line) and individual (blue line) models for all 20 sites}
\end{figure}

\newpage

\section*{Jackknifed partial least squares analysis}

\begin{figure}[h]
	\centering
	\includegraphics{../results/figures/SI_figures/agrirural_jk_individual_loadings.png}
	\caption{\textbf{Loadings of PLS calculated for each site removed for the agri-rural sites}. Each plot shows the loadings of the first principle component with that site removed from the analysis, showing the individual points making up the overall box plot loadings in Fig 4b.}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics{../results/figures/SI_figures/urban_jk_individual_loadings.png}
	\caption{\textbf{Loadings of PLS calculated for each site removed for the urban sites}. Each plot shows the loadings of the first principle component with that site removed from the analysis, showing the individual points making up the overall box plot loadings in Fig 4d.}
\end{figure}

\end{document}

With the complementary cumulative ($x \ge m$):
\begin{eqnarray*}P(\underline x > x )&=&p\frac{\left(1-a_s x\right)e^{-a_s b_s (x-m)}-b_s^{-1}\left(e^{-a_s b_s(x-m)}-e^{-b_s\left(1-a_s m\right)}\right) }
{\left(1-a_s m\right)-b_s^{-1}\left(1-e^{-b_s\left(1-a_s m\right)}\right)}\\&& \qquad +(1-p) \frac{\left(1-a_r x\right)e^{-\pi b_r (a_r x)^2 }+\frac{\text{erf}\left(a_r \sqrt{\pi b_r}x \right)-\text{erf}\left(\sqrt{\pi b_r}\right)}{2 \sqrt{b_r}}}{\left(1-a_r m\right)e^{-\pi b_r (a_r m)^2 }+\frac{\text{erf}\left(a_r \sqrt{\pi b_r}m \right)-\text{erf}\left(\sqrt{\pi b_r}\right)}{2 \sqrt{b_r}}}.
\end{eqnarray*}

\begin{eqnarray*}P(\underline x > x )&=&p\frac{\left(1-a_s x\right)e^{-a_s b_s x}-b_s^{-1}\left(e^{-a_s b_sx}-e^{-b_s}\right) }
{\left(1-a_s m\right)e^{-a_s b_s m}-b_s^{-1}\left(e^{-a_s b_s m}-e^{-b_s}\right)}\\&& \qquad +(1-p) \frac{\left(1-a_r x\right)e^{-\pi b_r (a_r x)^2 }+\frac{\text{erf}\left(a_r \sqrt{\pi b_r}x \right)-\text{erf}\left(\sqrt{\pi b_r}\right)}{2 \sqrt{b_r}}}{\left(1-a_r m\right)e^{-\pi b_r (a_r m)^2 }+\frac{\text{erf}\left(a_r \sqrt{\pi b_r}m \right)-\text{erf}\left(\sqrt{\pi b_r}\right)}{2 \sqrt{b_r}}}.
\end{eqnarray*}

\section*{Estimating $p$}
The probabilities are of all of the form $P(\underline x= x )=p f(x)+ (1-p) g(x), $ with $f$ and $g$ functions of $x$ that have values between 0 and 1.  The likelihood of the parameter $p$, given a set of data $\x =\{ x_1, \dots, x_n\}$ is
$$L(p|\x)=\prod_{i=1}^n p f(x_i)+ (1-p) g(x_i).$$ If $f(x_i) \ne g(x_i)$ for all $i$, this is a polynomial of degree $n$ (and the degree $n-m$ if there are $m$ elements of $\x$ for which $f(x)= g(x)$. In what follows we will assume that $f(x_i), f(x_i) \ne 0$ and $f(x_i) \ne g(x_i)$ for all $i$. \footnote{It is easy to do the analysis in case there are elements for which $f(x)= g(x)$ by ordering the data points so that for the first $n-m$ elements $f(x_i) \ne g(x_i)$ so that the likelihood can be written as $L(p|\x)=\prod_{i=n-m+1}^n f(x_i) \prod_{i=1}^{n-m} p f(x_i)+ (1-p) g(x_i)$ and follow through the same argument for the $n-m$ roots. } By rewriting the likelihood as
$$L(p|\x)=\prod_{i=1}^n\left(g(x_i)-f(x_i)\right)\prod_{i=1}^n \frac{g(x_i)}{g(x_i)-f(x_i)}-p $$ it is easy to see that the roots of the polynomial are all real valued and take value $\frac{g(x_i)}{g(x_i)-f(x_i)}.$ There are no roots of the polynomial in the interval $[0,1]$: all roots take either negative values, or are larger than 1. This is easy to see: if the root is positive, it requires $g(x)>f(x)$, but this implies that the root must be larger than 1 as $\frac{g(x_i)}{g(x_i)-f(x_i)}=1+\frac{f(x_i)}{g(x_i)-f(x_i)}>1$ if $g(x)>f(x)$.

If $p=0$ the likelihood is given by $L(0|\x)=\prod_{i=1}^n g(x_i)>0. $ Likewise, if $p=1$ the likelihood is given by $L(1|\x)=\prod_{i=1}^n f(x_i)>0. $ As the $L(p|\x)>0$ for all for which $0 \le p \le 1$ the function is continuous, positive on the interval $[0,1].$ The maxima of the likelihood obey $\frac{\df  \mathcal{L}(p|\x)}{\df p}=-\sum_{i=1}^n \left(\frac{g(x_i)}{g(x_i)-f(x_i)}-p\right)^{-1}=0$ where $\mathcal{L}(p|\x)=\ln L(p|\x)$ is the log-likelihood.

The function $\frac{\df  \mathcal{L}(p|\x)}{\df p}$ is a continuously decreasing function of $p$. This follows from the fact that $\frac{\df^2  \mathcal{L}(p|\x)}{\df p^2}=-\sum_{i=1}^n \left(\frac{g(x_i)}{g(x_i)-f(x_i)}-p\right)^{-2}$ which is negative for all values of $p$. As a consequence, on the interval $[0,1]$ there is at most one value for $p$ for which $\frac{\df  \mathcal{L}(p|\x)}{\df p}=0$ and there is a single maximum for the likelihood ${L}(p|\x)$ in $p$. Moreover, if $\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=0}>0$ and $\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=1}\ge 0$ the likelihood function is continuously increasing on $[0,1]$ and the maximum is attained for $p=1$. Likewise, if $\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=0}\le 0$ and $\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=1}<0$ the likelihood function is continuously decreasing on $[0,1]$ and the maximum is attained for $p=0$. If $\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=0}>0$ and $\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=1}<0$ the likelihood function for $p$ attains a single maximum on $(0,1)$.

To find the value of the maximum likelihood, the following methodology can be used:
\begin{itemize}
\item Determine if $f(x_i) \ne g(x_i)$ for any $i$. If so, remove them from the data set and proceed with the remaining $n$ data points.
\item Calculate the derivatives of the log likelihood for $p=0$ and $p=1$. These are given by: $$\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=0}=-\sum_{i=1}^n \left(\frac{g(x_i)}{g(x_i)-f(x_i)}\right)^{-1}=-n+\sum_{i=1}^n \frac{f(x_i)}{g(x_i)}$$
    $$\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=1}=-\sum_{i=1}^n \left(\frac{f(x_i)}{g(x_i)-f(x_i)}\right)^{-1}=n-\sum_{i=1}^n \frac{g(x_i)}{f(x_i)}$$
\item If $\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=0}\le 0$ and $\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=1}<0$ the maximum is attained for $p=0$
\item If $\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=0}>0$ and $\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=1}\ge 0$ the maximum is attained for $p=1.$
\item If $\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=0}>0$ and $\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=1}<0$ the likelihood function for $p$ attains a single maximum on $(0,1)$ and can be determined by numerical means such as Newton's method. An obvious initial point for such a method is $p_0=\frac{\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=0}}{\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=0}-\left.\frac{\df  \mathcal{L}(p|\x)}{\df p}\right|_{p=1}}$ (which is the intersection with the p-axis of a straightline between the values calculated for $p=0$ and $p=1$).
\end{itemize}

\section*{Seeley's profitability function}

To apply this model to data we will need a detailed description of how bees translate the quality $q$ and distance $x$ of a resource into $\phi(q, x)$, their measure of profitability. Tom Seeley conducted detailed experiments how honey bees respond to differences in quality and distance (REFS).

Seeley's work has shown that bees judge the profitability of a resource by the and optimise the energy efficiency, i.e  the net energy gain per amount of energy spend. Seeley (1994) takes the gain to be proportional to the load carried and the sucrose concentration. The cost depends on the different activities undertaken (e.g. flights in and out, collecting food, unloading food) and the respective times these activities took, and the energy spend in undertaking them.

Let the sucrose concentration be given by $q'$ and the travel distance by $x$. The gain is the $G$ is proportional to the quality, and the cost, $C$, increases linearly with the energy spent. We assume here that energy spent is fixed costs and further increases linearly with distance. We can then write the net energy efficiency as
$\frac{G-C}{C}= \frac{q}{1+\alpha  x}-1, $ where $q$ is the quality of the resource and $\alpha$ the energy spend per distance, and both measures are taken relative to the fixed cost of travel.
 If we assume that the bees do not dance if the profitability is negative, we as a measure for the propensity of bees to dance in relation to resource quality and distance the function:
$$\phi(q,x)=\left[\frac{q}{1+\alpha x}-1\right]_+$$
 (we used the notation $[x]_+=\max(x,0)$).
This function has the property that for a fixed $x$, there is a minimum quality $q=1+\alpha  x$ below which the bees don't dance (compare to Fig 5:10, Seeley (1984)). The equivalent quality for two resources at distances $x_1$ and $x_2$ is $\frac{1+\alpha x_2}{1+\alpha x_1}=\frac{ q_2}{q_1}$; the equivalent quality increases linearly with distance (more or less born out by Boch 1956). The critical distance $\xi_{ij}$ is given by $ \xi_{ij}=\left[\frac1 \alpha\frac{ q_j-q_i}{q_i}+ \frac{ q_j}{q_i}x\right]_+.$

The representation of distances on the floor then follows the distribution:
$$P(\underline x= x )=p \frac{\lambda_s e^{-\lambda_s x}\sum_{i=m}^n  \frac{c_s \lambda_i}{\lambda_s} \left[\frac{q_i}{1+\alpha x}-1\right]_+ }{M_s} +(1-p)\frac{\sum_{i=m}^n  \left[\frac{q_i}{1+\alpha x}-1\right]_+2 \pi c_r \lambda_i  x e^{- \pi \sum_{j=1}^n c_r \lambda_j \xi_{ij}(x)^2}}{M_r}$$
with $$M_s=\sum_{i=m}^n \int_{0}^{\frac{q_i-1}{\alpha}} \frac{c_s \lambda_i}{\lambda_s} \left(\frac{q_i}{1+\alpha x}-1\right) \lambda_s e^{-\lambda_s x} \df x
$$ and
$$M_r= \sum_{i=m}^n \int_0^{\frac{q_i-1}{\alpha}} \left(\frac{q_i}{1+\alpha x}-1\right)\lambda_i2 \pi c_r  x e^{- \pi \sum_{j=1}^n c_r \lambda_j \xi_{ij}(x)^2}\df x.$$

This model can be used to calculate the likelihood for a given data set. In the likelihood calculation the  normalisation factors can be calculated numerically for a given set of parameters. As the same normalisation factor applies to all data points, this is not a very computationally intensive procedure.

This model has $2 n+ 4 $ parameters but contains a redundant parameter:
By defining $\lambda_i^r=c_r \lambda_i$ and $c=\frac{c_s}{c_r}$ the number of parameters can be reduced by one. Alternatively, the parameter $c_r$ can be set to 1 without loss of generality to reduce the number of parameters to $2n+ 3. $

The Seeley model provides a biologically grounded description of the distribution of dances on the dance floor. It contains a large number of parameters, and in many practical applications it is challenging to determine all parameters as the model is ill conditioned: the model tends to give the same or very similar behaviour for a certain combinations of parameters. This makes it difficult or impossible to estimate the value parameters of interest with an accuracy. We will next consider a number of other models that are simplifications of the above model. Three simple and straightforward cases are:
\begin{itemize}
\item Scouts only: $p=1$ (tantamount to saying the colony does not act as a superorganism, the information on the dance floor is scouts information pooled, but not processed)
\item Recruits only: $p=0$ (tantamount to saying each forager can take the same decisions as the hive can. The colony does not act as a superorganism because the capabilities of the dance floor do not exceed that of the individual)
\item One resource only: $n=1$
\item No dependence on distance: $\alpha \rightarrow 0$
\end{itemize}
If the profitability becomes independent of distance the pdf reduces to:
$$\lim_{\alpha \rightarrow 0} P(\underline x= x )=p \lambda_s e^{-\lambda_s x} +(1-p)  2 \pi c_r \lambda_n  x e^{- \pi c_r \lambda_n x^2}$$
where we assumed the for the best resource $q_n>1.$ Also in this case there is a redundant parameter and  $c_r$ can be set to 1 without loss of generality.

\section*{The most profitable resource distribution }
 (to be suppressed, but just in case)
The probability of finding the most profitable resource at a distance  $0 \le x \le x_{n, {\rm max}}$ is a proper probability distribution.
This is because $$\int_{0}^\infty  \sum_{\forall i: {x_{i, \max}>x}}^n 2 \pi c_r \lambda_i x e^{- \pi \sum_{j=1, }^n c_r \lambda_j \xi_{ij}(x)^2}\df x=1- \int_{x_{n, \max}}^\infty  2 \pi c_r \lambda_n x e^{- \pi  c_r \lambda_n x^2}\df x$$

To show this, we first establish that \begin{equation}\xi_{ij}(\xi_{ki}(x))=\xi_{kj}(x) \label{phiresult}.\end{equation} Note that $\phi(q_k,x)=\phi(q_i,\xi_{ki}(x))=\phi(q_j,\xi_{ij}(\xi_{ki}(x)))$ From the definition of $\phi$ it follows that $\phi(q_k,x)=\phi(q_j,\xi_{kj}).$ From this (\ref{phiresult}) follows.
As a corollary, $\xi_{ij}(\xi_{ji}(x))=x$ and hence the function $\xi_{ij}(x)$ is the inverse of $\xi_{ji}(x)$.

Because we order the $q_i$ from small to large, and because $\xi_{ij}$ increases with $q_j$, and decreases with $q_i, $ $\xi_{ij}(x)>\xi_{ji}(x)$ if $i<j$.

We evaluate the integral
\begin{eqnarray*}&&\int_{0}^\infty  \sum_{\forall i: {x_{i, \max}>x}}^n 2 \pi c_r \lambda_i x e^{- \pi \sum_{j=1}^n c_r \lambda_j
\xi_{ij}(x)^2}\df x\\
&=& \sum_{i=1}^n \int_{0}^{x_{i, \max}}  2 \pi c_r \lambda_i x_i e^{- \pi \sum_{j=1}^n c_r \lambda_j
\xi_{ij}(x_i)^2}\df x_i\\
\end{eqnarray*}
substitute $x=\xi_{in}(x_i)$, $x_i=\xi_{ni}(x)$, $\df x_i=\xi_{ni}'(x) \df x$
\begin{eqnarray*}
&=& \sum_{i=1}^n \int_{0}^{x_{n, \max}}  2 \pi c_r \lambda_i \xi_{ni}(x)\xi_{ni}'(x) e^{- \pi \sum_{j=1}^n c_r \lambda_j
\xi_{ij}(\xi_{ni}(x))^2} \df x\\
&=& \int_{0}^{x_{n, \max}}  2\pi\sum_{i=1}^n   c_r \lambda_j \xi_{ni}(x)\xi_{ni}'(x) e^{- \pi \sum_{j=1}^n c_r \lambda_j
\xi_{nj}(x)^2} \df x\\
&=&1- \int_{x_{n, \max}}^\infty  2 \pi c_r \lambda_n x e^{- \pi  c_r \lambda_n x^2}\df x
\end{eqnarray*}
q.e.d

------

\end{document}

### 2022-08-02 analysis/Simulation/Code/run.py

@@ -1,193 +1,0 @@
-#!/usr/bin/env python3
-"""Run simulation of honeybee foraging"""
-__appname__  = "run.py"
-__author__   = "Joseph Palmer <joseph.palmer18@imperial.ac.uk>"
-__version__  = "0.0.2"
-__date__     = "02-2020"
-
-import numpy as np
-import pandas as pd
-import scipy.stats
-import matplotlib.pyplot as plt
-import Agents
-import Methods
-import config
-
-def create_flowers(lwr, upr, res, quality_array):
-    # Simulation window parameters
-    r = upr-lwr
-    xx0 = 0
-    yy0 = 0
-
-    areaTotal = np.pi*r**2
-
-    # Point process parameters
-    lambda0 = res
-
-    # Simulate Poisson point process
-    numbPoints = scipy.stats.poisson(lambda0*areaTotal).rvs()
-    theta = 2*np.pi*scipy.stats.uniform.rvs(0,1,((numbPoints,1)))
-    rho = r*np.sqrt(scipy.stats.uniform.rvs(0,1,((numbPoints,1))))
-
-    # Convert from polar to Cartesian coordinates
-    xx = rho * np.cos(theta)
-    yy = rho * np.sin(theta)
-
-    # Shift centre of disk to (xx0,yy0)
-    xx=xx+xx0
-    yy=yy+yy0
-
-    # Set locations as flowers
-    locations = np.column_stack((xx, yy))
-    flowers = np.apply_along_axis(Agents.Flower, 1, locations, quality_array)
-    fdict = {
-        i : flowers[i] for i in range(res)
-    }
-    return fdict
-
-def initialize(lwr, upr, res, nbees, nscouts, quality_array):
-    # create flowers in environment and store globaly.
-    config.FLOWERDICT = create_flowers(lwr, upr, res, quality_array)
-
-    # create honeybee population
-    nrecruits = nbees-nscouts
-    scouts = np.array([Agents.Scout() for i in range(nscouts)])
-    recruits = np.array([Agents.Recruit() for i in range(nrecruits)])
-    return (scouts, recruits)
-
-def flower_turnover(n, lwr, upr, k, quality_array):
-    # select n flowers and remove
-    assert (n <= len(config.FLOWERDICT.keys())), "N > flowers in environment!"
-    rmv_idx = np.random.choice(list(config.FLOWERDICT.keys()), n, replace=False)
[…]

### 2022-08-02 inst/publication/analysis/install-analysis-packages.R
tidied up analysis code and moved reference to the paper into inst

@@ -1,20 +1,0 @@
-library_calls <- purrr::map(
-  list.files("analysis/", pattern = "*.Rmd|.R"),
-  ~ {
-    stringr::str_match(readLines(paste0("analysis/", .x)), "library\\((.*?)\\)")[,2]
-  }
-)
-
-library_calls <- unlist(library_calls)
-packages <- library_calls[!is.na(library_calls)]
-new.packages <- packages[
-  !(packages %in% installed.packages()[,"Package"])
-]
-new.packages <- new.packages[new.packages != "wagglefit"]
-if( length(new.packages)) {
-  message("Installing the following packages:")
-  print(new.packages)
-  install.packages(new.packages)
-} else {
-  message("Nothing to install, all dependencies satisfied")
-}

### 2022-08-02 R/plotting.R
compressed the publication analysis folder

@@ -93,7 +93,7 @@ make_base_plot <- function(x, logit = TRUE) {
 #' along with the model fits to this data.
 #' @importFrom ggplot2 ggplot aes geom_histogram geom_line theme_set theme
 #' theme_classic element_text element_blank element_rect annotation_custom
-#' ggplotGrob scale_colour_manual
+#' ggplotGrob scale_colour_manual scale_x_continuous
 #' @importFrom tibble tibble
 #' @importFrom purrr map map_df
 #' @importFrom rlang .data

### 2022-08-02 R/RcppExports.R
Publication branch (#24)
* Added the lower level fitting functions. next add tests

* Started to make tests. WIP

* Added basic testing functions for models

* added lintr to buildignore

* removed export tags for some functions.

* Try updaing actions yaml to run on pull and prs

* fixed error in actions

* Added readme

* removed action for rendering markdown, lintr should catch this.

* updated checks to run on every branch

* test commit

* Convert to use cpp (#10)

* Converted models to c++ code

* renamed model tests so wornt run. kept code as will be useful when creating tests for c++ m

@@ -1,0 +1,166 @@
+# Generated by using Rcpp::compileAttributes() -> do not edit by hand
+# Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
+
+#' Alter a NumericVector inplace
+#'
+#' @param x NumericArray
+#' @export
+alter_in_place <- function(x) {
+    invisible(.Call('_wagglefit_alter_in_place', PACKAGE = 'wagglefit', x))
+}
+
+#' Model CCDF for scout distribution
+#'
+#' @param x double Foraging distance
+#' @param m double Minimum foraging distance
+#' @param bs double Scout rate
+#' @param as double scout alpha
+#' @return result double the ccdf for that foraging distance (x)
+#' @export
+scout_ccdf <- function(x, m, bs, as) {
+    .Call('_wagglefit_scout_ccdf', PACKAGE = 'wagglefit', x, m, bs, as)
+}
+
+#' Model CCDF for recruit distribution
+#'
+#' @param x double Foraging distance
+#' @param m double Minimum foraging distance
+#' @param br double Recruit rate
+#' @param ar double recruit alpha
+#' @return result double the ccdf for that foraging distance (x)
+#' @export
+recruit_ccdf <- function(x, m, br, ar) {
+    .Call('_wagglefit_recruit_ccdf', PACKAGE = 'wagglefit', x, m, br, ar)
+}
+
+#' Model ccdf function for collective model. Stores results in
+#' given array (y)
+#'
+#' @param x NumericVector foraging distances
+#' @param y NumericVector storage array for the results
+#' @param p double Proportion of scouts (0<=p<=1)
+#' @inheritParams scout_ccdf
+#' @inheritParams recruit_ccdf
+#' @export
+ccdf_model_collective <- function(x, y, p, bs, br, as, ar) {
+    invisible(.Call('_wagglefit_ccdf_model_collective', PACKAGE = 'wagglefit', x, y, p, bs, br, as, ar))
+}
+
+#' Model ccdf function for individual model. Stores results in given array (y)
+#'
+#' @param x NumericVector foraging distances
+#' @param y NumericVector storage array for the results
+#' @inheritParams scout_ccdf
+#' @export
+ccdf_model_individual <- function(x, y, bs, as) {
+    invisible(.Call('_wagglefit_ccdf_model_individual', PACKAGE = 'wagglefit', x, y, bs, as))
+}
+
+#' Get model ccdf for a given model
+#'
+#' @param x NumericVector foraging distances
[…]

### 2022-08-02 R/fitting.R

@@ -1,0 +1,307 @@
+#' Check the upper value to be used meets criteria
+#'
+#' @description Generates upper and lower bounds for all parameters
+#' @param upper double The upper bound for the params
+#' @export
+#' @examples
+#' \dontrun{
+#' check_upper_bound(upper = 10)
+#' }
+#'
+check_upper_bound <- function(upper) {
+  if (length(upper) > 1) {
+    stop("only pass one value as upper")
+  }
+  if (!is.double(upper)) {
+    stop("upper must be of type 'double'")
+  }
+  if (upper < 0) {
+    stop("upper cannot be < 0")
+  }
+}
+
+#' Generates bounds for the collective model
+#'
+#' @description Generates upper and lower bounds for all parameters
+#' @param upper double The upper bound for the params
+#' @return bounds matrix of lower ([,1]) and upper ([,2]) bounds
+#' @export
+#' @examples
+#' \dontrun{
+#' generate_bounds_all()
+#' }
+#'
+get_bounds_collective <- function(upper) {
+  p_bnds <- c(0, 1.0)
+  bs_bnds <- c(1.0e-6, 10)
+  br_bnds <- c(1.0e-6, 10)
+  as_bnds <- c(1.0e-12, upper)
+  ar_bnds <- c(1.0e-12, upper)
+  bounds <- rbind(
+    p_bnds, bs_bnds,
+    br_bnds, as_bnds,
+    ar_bnds
+  )
+  return(bounds)
+}
+
+#' Generates bounds for all parameters of the individual model
+#'
+#' @description Generates upper and lower bounds for all parameters
+#' @param upper double The upper bound for the params
+#' @return bounds matrix of lower ([,1]) and upper ([,2]) bounds
+#' @export
+#' @examples
+#' \dontrun{
+#' generate_bounds_all()
+#' }
+#'
+get_bounds_individual <- function(upper) {
+  upper <- as.double(upper)
+  bs_bnds <- c(1.0e-6, upper)
+  as_bnds <- c(1.0e-6, upper)
+  bounds <- rbind(
+    bs_bnds,
+    as_bnds
+  )
+  return(bounds)
+}
+
+
+#' Generates starting estimates for all numerical optimisation of all parameters
+#' of the collective model
+#'
+#' @description Generates starting estimates for all parameters by taking a
+#' random value sampled from a normal distribution truncated at the upper and
+#' lower bounds. Mean is given as the mid point between the upper and lower
+#' bounds and the sd is set as 1.0, except for the p parameter, which has a mean
[…]

### 2022-08-02 R/plotting.R

@@ -1,0 +1,209 @@
+
+#' Calculates the ccdf for given data
+#'
+#' @description Orders the data smallest to largest and calculates the
+#' probability of sampling a value greater than or equal to a given value.
+#' @param x The data, e.g. foraging distance
+#' @return tibble of the sorted foraging distances and their cumulative
+#' probability
+#' @importFrom purrr map_dbl
+#' @importFrom tibble tibble
+#' @export
+#'
+inverse_ccdf <- function(x) {
+  if (!is.double(x) | length(x) < 2) {
+    stop("x must be a double array")
+  }
+  sd <- -sort(-x)
+  prob <- map_dbl(
+    seq_len(length(x)),
+    ~ {
+      .x / length(x)
+    }
+  )
+  return(tibble(sd, prob))
+}
+
+#' Creates data for plots
+#'
+#' @description Using the optimised paramater estimates, creates the data for
+#' the ccdf plots
+#' @param x doubleArray The foraging distance to plot
+#' @param param_est doubleArray The paramater estimates of the optimisation
+#' @param model characterArray The name of the model used
+#' @param npoints integer The number of points to plot the predicted data for.
+#' Defaults to 100.
+#' @return tibble of model data and cumulative probability
+#' @importFrom tibble tibble
+#' @export
+#'
+make_ccdf_plot_data <- function(x, param_est, model, npoints = 100) {
+  model_n <- model_number_from_model(model)
+  x_seq <- seq(min(x), max(x), length.out = npoints)
+  cumul_ccdf <- model_ccdf(x_seq, param_est, model_n)
+  cumul_ccdf[[1]] <- 1
+  cumul <- tibble(x_seq, cumul_ccdf)
+  return(cumul)
+}
+
+#' Creates ccdf plot of data
+#'
+#' @description Creates a ccdf plot for the data provided
+#' @param x doubleArray The foraging distance to plot
+#' @param logit Bool to log the probabilities or not.
+#' @return ggplot plot of cumulative probability for foraging distances in Km
+#' @importFrom ggplot2 ggplot aes_string geom_point geom_line theme_set theme
+#' theme_classic element_text labs
+#' @importFrom tibble tibble
+#' @importFrom magrittr %>%
+#' @importFrom rlang .data
+#' @export
+#'
+make_base_plot <- function(x, logit = TRUE) {
+  theme_set(
+    theme_classic() +
+      theme(
[…]

### 2022-08-02 R/utils.R

@@ -1,3 +1,37 @@
+#' Get model number from model name
+#'
+#' @description Converts the model name to its number for cpp. E.g. 'all' is 0
+#' @param model characterArray The model to get the number for
+#' @export
+#' @concept utility
+model_number_from_model <- function(model) {
+  model_list <- list(
+    "collective" = 0,
+    "individual" = 1
+  )
+  if (!(model %in% names(model_list))) {
+    stop("Model name not found")
+  }
+  return(model_list[[model]])
+}
+
+#' Wrapper for message
+#'
+#' @description A wrapper for `message` that only prints output when
+#' `verbose = TRUE`.
+#' @param verbose Logical, defaults to `TRUE`. Should verbose processing
+#' messages and warnings be returned.
+#' @param ... Additional arguments passed to `message`.
+#' @concept utility
+message_verbose <- function(verbose = TRUE, ...) {
+  if (verbose) {
+    message(...)
+  }
+  return(invisible(NULL))
+}
+
+#' Caluclates distance from waggle dance
+#'
 #' Calculate distance from waggle run duration
 #' @description Calculates distance in Km from dance duration, using a linear
 #' relationship. Taken from Schürch et al (2019)
@@ -8,9 +42,216 @@
 #' \dontrun{
 #' calc_dist(1.8)
 #' }
-#'
+#' @concept utility
 calc_dist <- function(duration) {
   intercept <- 0.17
   slope <- 1.38
-  return(intercept+(slope*duration))
+  return(intercept + (slope * duration))
+}
+
+#' Draws sample from a normal distribtion trucated between upr and lwr bounds.
+#'
+#' @description Samples from a normal distribution and then checks it falls
+#' within given bounds.
+#' @param mean double Mean for the normal distribution to sample
+#' @param sd double Standard deviation for the normal distribution to sample
+#' @param lwr double Lower bound samples must be >= to
+#' @param upr double Upper bound samples must be <= to
+#' @return x double The sampled value
+#' @importFrom stats rnorm
+#' @export
+#' @examples
+#' \dontrun{
+#' trunc_normal(0, 0.5, -1, 1)
+#' }
+#' @concept utility
+trunc_normal <- function(mean, sd, lwr, upr) {
+  continue <- TRUE
+  while (continue) {
+    x <- rnorm(n = 1, mean = mean, sd = sd)
+    if (x >= lwr & x <= upr) {
+      continue <- FALSE
+    }
+  }
+  return(x)
+}
+
+#' Calculate AIC values for a model
+#'
[…]

### 2022-08-02 R/wagglefit.R

@@ -1,0 +1,4 @@
+#' @useDynLib wagglefit
+#' @importFrom Rcpp sourceCpp
+NULL
+# > NULL

### 2022-08-02 src/RcppExports.cpp

@@ -1,0 +1,231 @@
+// Generated by using Rcpp::compileAttributes() -> do not edit by hand
+// Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
+
+#include <Rcpp.h>
+
+using namespace Rcpp;
+
+#ifdef RCPP_USE_GLOBAL_ROSTREAM
+Rcpp::Rostream<true>&  Rcpp::Rcout = Rcpp::Rcpp_cout_get();
+Rcpp::Rostream<false>& Rcpp::Rcerr = Rcpp::Rcpp_cerr_get();
+#endif
+
+// alter_in_place
+void alter_in_place(NumericVector x);
+RcppExport SEXP _wagglefit_alter_in_place(SEXP xSEXP) {
+BEGIN_RCPP
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< NumericVector >::type x(xSEXP);
+    alter_in_place(x);
+    return R_NilValue;
+END_RCPP
+}
+// scout_ccdf
+double scout_ccdf(double x, double m, double bs, double as);
+RcppExport SEXP _wagglefit_scout_ccdf(SEXP xSEXP, SEXP mSEXP, SEXP bsSEXP, SEXP asSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< double >::type x(xSEXP);
+    Rcpp::traits::input_parameter< double >::type m(mSEXP);
+    Rcpp::traits::input_parameter< double >::type bs(bsSEXP);
+    Rcpp::traits::input_parameter< double >::type as(asSEXP);
+    rcpp_result_gen = Rcpp::wrap(scout_ccdf(x, m, bs, as));
+    return rcpp_result_gen;
+END_RCPP
+}
+// recruit_ccdf
+double recruit_ccdf(double x, double m, double br, double ar);
+RcppExport SEXP _wagglefit_recruit_ccdf(SEXP xSEXP, SEXP mSEXP, SEXP brSEXP, SEXP arSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< double >::type x(xSEXP);
[…]

### 2022-08-02 src/code.cpp

@@ -1,0 +1,44 @@
+/*
+ * Example code for using Rcpp and C++
+ * This is just a series of function which show how to get certain types in and
+ * out of C++ from R. For example, altering an array in place.
+ */
+#include <Rcpp.h>
+using namespace Rcpp;
+
+
+double sum_array(double* x, int n)
+{
+  double res = 0;
+  for (int i = 0; i < n; i++) {
+    res += x[i];
+  }
+  return res;
+}
+
+// Multiply an array by 2 in place
+// These functions (using pointers) cannot be linked directly with R
+// Make sure you write tests for them in C++
+void array_times_two(double* x, int n)
+{
+  for (int i = 0; i < n; i++) {
+    x[i] = x[i]*2;
+  }
+}
+
+//' Alter a NumericVector inplace
+//'
+//' @param x NumericArray
+//' @export
+// [[Rcpp::export]]
+void alter_in_place(NumericVector x) {
+  double* x_ = &x[0];
+
+  // We can now pass the pointer around, e.g.:
+  // pass the array pointer to function which adds them up
+  int n = x.size();
+  double res = sum_array(x_, n);
+  Rcout << res << std::endl;
+
+  array_times_two(x_, n);
+}

### 2022-08-02 src/model_ccdfs.cpp

@@ -1,0 +1,121 @@
+#include <iostream>
+#include <cmath>
+#include <Rcpp.h>
+using namespace Rcpp;
+
+//' Model CCDF for scout distribution
+//'
+//' @param x double Foraging distance
+//' @param m double Minimum foraging distance
+//' @param bs double Scout rate
+//' @param as double scout alpha
+//' @return result double the ccdf for that foraging distance (x)
+//' @export
+// [[Rcpp::export]]
+double scout_ccdf(double x, double m,
+                  double bs, double as)
+{
+  double result;
+  if (x > 1/as) {
+    result = 0;
+  } else {
+    result = ((1-as*x)*exp(-as*bs*x)-pow(bs, -1)*(exp(-as*bs*x)-exp(-bs))) /
+    ((1-as*m)*exp(-as*bs*m)-pow(bs, -1)*(exp(-as*bs*m)-exp(-bs)));
+  }
+  return result;
+}
+
+//' Model CCDF for recruit distribution
+//'
+//' @param x double Foraging distance
+//' @param m double Minimum foraging distance
+//' @param br double Recruit rate
+//' @param ar double recruit alpha
+//' @return result double the ccdf for that foraging distance (x)
+//' @export
+// [[Rcpp::export]]
+double recruit_ccdf(double x, double m,
+                    double br, double ar)
+{
+  double result;
+  if (x > 1/ar) {
+    result = 0;
+  } else {
+    result = ((1-ar*x)*exp(-M_PI*br*pow(ar*x, 2)) +
+      ((erf(ar*sqrt(M_PI*br)*x)-erf(sqrt(M_PI*br)))/(2*sqrt(br)))) /
+      ((1-ar*m)*exp(-M_PI*br*pow(ar*m, 2)) +
+      ((erf(ar*sqrt(M_PI*br)*m)-erf(sqrt(M_PI*br)))/(2*sqrt(br))));
+  }
+  return result;
+}
+
+//' Model ccdf function for collective model. Stores results in
+//' given array (y)
+//'
+//' @param x NumericVector foraging distances
+//' @param y NumericVector storage array for the results
+//' @param p double Proportion of scouts (0<=p<=1)
+//' @inheritParams scout_ccdf
+//' @inheritParams recruit_ccdf
+//' @export
+// [[Rcpp::export]]
+void ccdf_model_collective(NumericVector x, NumericVector y,
+                    double p, double bs, double br,
+                    double as, double ar)
+{
[…]

### 2022-08-02 src/models.cpp

@@ -1,0 +1,391 @@
+#include <iostream>
+#include <cmath>
+#include <Rcpp.h>
+#include <nloptrAPI.h>
+using namespace Rcpp;
+
+/*
+------------------------------- structs and data -- ----------------------------
+*/
+
+
+// struct to hold additional data for optimisation
+//
+// x: The distance data to fit to
+// verbose: To display messages to console
+// fcount: number of times the objective function has been called
+// best_est: the best estimate up to that point
+typedef struct {
+    NumericVector x;
+    bool verbose;
+    int fcount;
+    double best_est;
+} data_struct;
+
+
+/*
+------------------------------- helper functions -------------------------------
+*/
+
+// Check if user has requested abort every 1000 iterations & if so exit
+//
+// @param fcount double The current number of iterations of the objective fun
+void check_user_input(int fcount) {
+    if (fcount % 1000 == 0) {
+        checkUserInterrupt();
+    }
+}
+
+// check optimise exit status and display important info about them to user
+//
+// @param optimise_result Integer The exit code from the nlopt optimisation
+// @param verbose Bool To display messages or not (defaults to TRUE)
+void check_optimise_result(int optimise_result, bool verbose = true) {
+    if (verbose) {
+        if (optimise_result < 0) {
+            Rcout << "NLOPT FAIL! Error code: " << optimise_result << std::endl;
+            if (optimise_result == -4) {
+                Rcout << "Fail is due to round off errors"
+                    "- result may still be usefull" << std::endl;
+            }
+        }
+    }
+}
+
+/*
+------------------------------ distribution functions --------------------------
+*/
+
+//' Model function for scouts
+//'
+//' @param x double Foraging distance
+//' @param m double Minimum foraging distance
+//' @param bs double scout rate
+//' @param as double scout alpha
+//' @export
+// [[Rcpp::export]]
+double scout_dist(double x, double m, double bs, double as)
+{
+    double maxpart = 1-as*x;
+    if (maxpart < 0)
+    {
+        maxpart = 0;
+    }
+    double result =
+        as * (
+            (bs * exp(-bs*as*(x-m))*maxpart) /
+            ((1-as*m) - pow(bs, -1)*(1-exp(-bs*(1-as*m))))
+        );
+    return result;
+}
+
+//' Model function for recruits
+//'
+//' @param br double Recruit rate
[…]

### 2022-08-02 tests/testthat/test-cpp_models.R

@@ -1,0 +1,82 @@
+test_scout_dist <- function(x, m, bs, as) {
+  truth <- 4.15163
+  ans <- scout_dist(x, m, bs, as)
+  test_that("scout_dist gives expected result", {
+    expect_equal(truth, round(ans, 5))
+  })
+}
+
+test_recruit_dist <- function(x, m, br, ar) {
+  truth <- 0.63748
+  ans <- recruit_dist(x, m, br, ar)
+  test_that("recruit_dist gives expected results", {
+    expect_identical(round(truth, 5), round(ans, 5))
+  })
+}
+
+test_loglike_model_collective <- function(x, p, bs, br, as, ar) {
+  truth <- -2.233698
+  ans <- loglike_model_collective(x, p, bs, br, as, ar)
+  test_that("loglike_model_collective gives expected results", {
+    expect_equal(truth, round(ans, 6))
+  })
+}
+
+test_loglike_model_individual <- function(x, br, ar) {
+  truth <- -0.83449
+  ans <- loglike_model_individual(x, br, ar)
+  test_that("loglike_model_individual gives expected results", {
+    expect_equal(truth, round(ans, 5))
+  })
+}
+
+test_optimise_model <- function(x, p, ls, ln, qn, a) {
+  actual_all <- c(
+    1.093483, 4.246401e-01, 1.540067, 1.754589e-06, 1.819189, 4.494388e-01
+  )
+  result_all <- optimise_model(
+    x, c(p, ls, ln, qn, a), c(0, 1e-6, 0, 0, 0), c(1, 5, 5, 5, 5),
+    xtol = 1e-6, model = 0
+  )
+  test_that("optimise_model all returns expected results", {
+    expect_identical(round(actual_all, 5), round(result_all, 5))
+  })
+  actual_scout <- c(
+    0.903998, 0.000001, 2.147276, 1.006386
+  )
+  result_scout <- optimise_model(
+    x, c(ls, qn, a), c(1e-6, 0, 0), c(5, 5, 5),
+    xtol = 1e-6, model = 1
+  )
+  test_that("optimise_model scout returns expected results", {
+    expect_identical(round(actual_scout, 5), round(result_scout, 5))
+  })
+}
+
+model_tests <- function() {
+  x <- c(
+    0.2, 0.5,
+    0.3, 0.6,
+    0.8, 0.5,
+    0.9, 0.6,
+    0.1, 0.48
+  )
+  p <- 0.5
+  bs <- 1.3
[…]

### 2022-08-02 tests/testthat/test-fitting.R

@@ -1,0 +1,187 @@
+# get_bounds
+test_check_upper_bound <- function() {
+  test_that("check_upper_bound gives error if upper < 0", {
+    expect_error(check_upper_bound(-1), "upper cannot be < 0")
+  })
+  test_that("check_upper_bound gives error if upper is not a double", {
+    expect_error(check_upper_bound("1"), "upper must be of type 'double'")
+  })
+  test_that("check_upper_bound gives error if upper is an array", {
+    expect_error(check_upper_bound(c(1, 2)), "only pass one value as upper")
+  })
+}
+
+test_get_bounds_collective <- function() {
+  upper <- 5
+  p_bnds <- c(0, 1.0)
+  bs_bnds <- c(1.0e-6, 10)
+  br_bnds <- c(1.0e-6, 10)
+  as_bnds <- c(1.0e-12, upper)
+  ar_bnds <- c(1.0e-12, upper)
+  actual_bounds <- rbind(
+    p_bnds, bs_bnds,
+    br_bnds, as_bnds,
+    ar_bnds
+  )
+  test_that("get_bounds_collective gets the correct bounds", {
+    expect_identical(actual_bounds, get_bounds_collective(upper))
+  })
+}
+
+test_get_bounds_individual <- function() {
+  upper <- 5
+  bs_bnds <- c(1.0e-6, upper)
+  as_bnds <- c(1.0e-6, upper)
+  actual_bounds <- rbind(
+    bs_bnds,
+    as_bnds
+  )
+  test_that("get_bounds_collective gets the correct bounds", {
+    expect_identical(actual_bounds, get_bounds_individual(upper))
+  })
+}
+
+test_get_starting_est_collective <- function(x) {
+  bnds <- get_bounds_collective(5)
+  result <- get_starting_ests_collective(x, bounds = bnds)
+  test_that("starting paramaters getd are within specified bounds", {
+    purrr::walk(
+      seq_len(length(result)),
+      ~ {
+        expect_true((result[[.x]] > bnds[.x, 1]) & (result[[.x]] < bnds[.x, 2]))
+      }
+    )
+  })
+}
+
+test_get_starting_est_individual <- function(x) {
+  bnds <- get_bounds_individual(5)
+  result <- get_starting_ests_individual(x, bounds = bnds)
+  test_that("starting paramaters getd are within specified bounds", {
+    purrr::walk(
+      seq_len(length(result)),
+      ~ {
[…]

### 2022-08-02 tests/testthat/test-plotting.R

@@ -1,0 +1,107 @@
+test_inverse_cdf <- function(x) {
+  test_that("inverse_cdf works as expected", {
+    expected <- tibble::tibble(
+      sd = c(0.90, 0.80, 0.60, 0.60, 0.50, 0.50, 0.48, 0.30, 0.20, 0.10),
+      prob = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0)
+    )
+    expect_identical(round(inverse_ccdf(x), 3), expected)
+  })
+
+  test_that("inverse_cdf fails if input is not a double array", {
+    x <- tibble::tibble(a = c(0, 4, 2, 6, 8, 3), b = c(0, 5, 1, 8, 3, 9))
+    expect_error(inverse_ccdf(x), "x must be a double array")
+    expect_error(inverse_ccdf(0.5), "x must be a double array")
+  })
+}
+
+test_make_ccdf_plot_data <- function(x, param_est, model, npoints) {
+  expected <- tibble::tibble(
+    x_seq = c(
+      0.100, 0.189, 0.278, 0.367, 0.456, 0.544, 0.633, 0.722, 0.811, 0.900
+    ),
+    cumul_ccdf = c(
+      1.000, 0.713, 0.534, 0.425, 0.362, 0.314, 0.267, 0.221, 0.178, 0.140
+    )
+  )
+  actual <- round(make_ccdf_plot_data(x, param_est, model, npoints), 3)
+  test_that("make_ccdf_plot_data returns expected results", {
+    expect_identical(expected, actual)
+  })
+}
+
+test_make_base_plot <- function(x) {
+  plt <- make_base_plot(x)
+  expected_xlab <- "Waggle dance duration (s)"
+  expected_ylab <- "cumulative frequency"
+  test_that("make_base_plot returns a ggplot object with the correct labels", {
+    expect_true(ggplot2::is.ggplot(plt))
+    expect_equal(expected_xlab, plt$labels$x)
+    expect_equal(expected_ylab, plt$labels$y)
+  })
+}
+
+test_make_full_plot <- function(x, model_result_list) {
+  plt <- make_full_plot(x, model_result_list)
+  dev.off()
+  expected_xlab <- "Waggle dance duration (s)"
+  expected_ylab <- "cumulative frequency"
+  test_that("make_full_plot returns a ggplot object with the correct labels", {
+    expect_true(ggplot2::is.ggplot(plt))
[…]

### 2022-08-02 tests/testthat/test-utils.R

@@ -1,3 +1,87 @@
-test_that("calc_dist works", {
+test_that("calc_dist gives correct answer with double", {
   expect_equal(calc_dist(1.8), 2.654)
+})
+
+test_that("cacl_dist gives correct answer with data frame to 3dp", {
+  duration <- data.frame(A = c(0.2, 0.6, 0.8, 1.3, 1.6, 4))
+  truth <- data.frame(A = c(0.446, 0.998, 1.274, 1.964, 2.378, 5.690))
+  distance <- calc_dist(duration)
+  expect_identical(round(truth, 3), round(distance, 3))
+})
+
+test_that("trunc_normal gives expected results", {
+  check <- purrr::map(
+    seq_len(10000),
+    ~ {
+      trunc_normal(mean = 0.2, sd = 0.5, lwr = 0, upr = 1)
+    }
+  )
+  expect_true(
+    all((check < 1 & check > 0))
+  )
+})
+
+test_that("message_verbose returns a message when asked", {
+  expect_equal(
+    capture.output(message_verbose(TRUE, "hi there"), type = "message"),
+    "hi there"
+  )
+})
+
+test_that("message_verbose does not return a message when asked", {
+  l <- suppressMessages(
+    capture.output(message_verbose(FALSE, "hi there"), type = "message")
+  )
+  expect_equal(l, character(0))
+})
+
+test_that("model_number_from_model returns correct models", {
+  expected <- c(0, 1)
+  results <- purrr::map_dbl(
+    c("collective", "individual"),
+    model_number_from_model
+  )
+  expect_identical(expected, results)
+})
+
+test_that("model_number_from_model returns error if model name is not known", {
+  expect_error(
+    model_number_from_model("amadeupmodel"),
+    "Model name not found"
+  )
+})
+
+test_that("calc_aic works as expected", {
+  expected <- 314
+  result <- calc_aic(2, -155)
+  expect_equal(result, expected)
+})
+
+expect_collective_dstat <- 0.4
+expect_individual_dstat <- 0.7
+x <- c(
+  0.2, 0.5,
+  0.3, 0.6,
+  0.8, 0.5,
+  0.9, 0.6,
+  0.1, 0.48
+)
+p <- 0.5
+bs <- 1.3
+br <- 1.3
+as <- 2.2
+ar <- 0.5
+param_est_collective <- c(p, bs, br, as, ar)
+param_est_individual <- c(bs, as)
+collective_result <- calc_ks_boot(x, param_est_collective, "collective")
[…]

### 2022-08-04 inst/publication/Code/fit_models_durations.Rmd
updated MS with Elli edits

--- 2022-05-03 analysis/fit_models_durations.Rmd
@@
 
-Provides a good fit on the data. All parameters look central and nicely covered.
-
 ```{r, BEL, message = FALSE, warning = FALSE}
@@
 ### BFI
-
-All parameters look central in the likelihood space and a nice fit is returned.
 
@@
 ### BLO
-
-The individual fit isnt great and the Bs paramater is increasing up to the boundry, indicating it can only really take on a straight line / exponential fit.
-The parameters for the collective model are fairly central in the likelihood space, the fit looks very good.
 
@@
 
-Bs and Br go in oposite directions. E.g. Bs approaches 0 whilst Br approaches an every higher number.
-
-Fit looks good.
-
 ```{r, BUR, message = FALSE, warning = FALSE}
@@
 
-The collective model roughly follows the individual model but is able to acheive a slightly improved fit to the shoulder, hence it provides a higher likelihood score.
-The AIC indicates this is overfitting, suggesting the individual model provides a more parsimonious explanation.
-
-$p$ does not approach 1 as one might expect, indicating there is overfitting and so a comparison with an individual model is required.
-
 ```{r, CAD, message = FALSE, warning = FALSE}
@@
 
-The collective model provides the most parsimonious and best fit.
-
 ```{r, GIL, message = FALSE, warning = FALSE}
@@
 ### HER
-
-The collective model provides the best fit to the data, but the proportion of scouts is high.
 
@@
 
-Collective model provides the best fit but falls under the tail and shoulder. The individual model strugles to find a good fit, probably due to the tail.
-
 ```{r, HHS, message = FALSE, warning = FALSE}
@@
 
-Collective model provides a very good fit, whilst the individual model fails to find much traction.
-The proportion of scouts goes very low (~3%) suggesting the majority of the colony are following a small number of scouting individuals.
-
 ```{r, HOR, message = FALSE, warning = FALSE}
@@
 ### MAK
-
-Again the collective model provides a good fit to the data, however, the individual model fits poorly, reducing to an exponential.
 
@@
 
-Collective model provides the best fit, however it misses a large section of the shoulder for the tail.
-The $bs$ parameter wants to go to zero, however when let go bellow 1e-6 the behaviour becomes very erratic and the fit deteriorates.
-
 ```{r, MEL, message = FALSE, warning = FALSE}
@@
 
-Collective provides the best fit.
-
 ```{r, MPA, message = FALSE, warning = FALSE}
@@
 
-Colletive provides the best fit.
-
 ```{r, ROT, message = FALSE, warning = FALSE}
@@
 
-Collective provides the best fit.
-
 ```{r, SAU, message = FALSE, warning = FALSE}
@@
 ### SOM
-
-Collective provides the best fit but misses a large section of the shoulder.
 
@@
 
-Collective provides the best fit.
-
 ```{r, SRA, message = FALSE, warning = FALSE}
@@
 
-Individual provides the best fit. Again, the proportion of scouts does not move towards 1. I.e. the collective model fails to reduce to the individual in the MLE fit.
-
 ```{r, STU, message = FALSE, warning = FALSE}
@@
 ### SWP
-
-Collective provides the best fit.
 
@@
 
-Collective provides the best fit.
-
 ```{r, YAL, message = FALSE, warning = FALSE}
@@
 ### ZSL
-
-Collective provides the best fit.
 
@@
 ```{r, make-plots, message = FALSE, warning = FALSE}
-# save the analysis results for all sites
-# save(all_sites, file = "results/site_fit_results.Rdata")
-
 # group all site results together
@@
 # save results
-# saveRDS(df, file = "results/site_fit_results.Rda")
-
-# all_sites$BFI$fit <- all_sites$BFI$fit +
-#   ylim(0, 6) +
-#   scale_x_continuous(breaks = seq(0, 6, by = 2))
+saveRDS(df, file = "../Results/site_fit_results.Rda")
 
@@
   plot = aic_plot,
-  filename = "results/figures/AIC_plot.png",
+  filename = "../Results/figures/AIC_plot.png",
   width = 90,
@@
   plot = ks_plot_dist,
-  filename = "results/figures/sites_ks.png",
+  filename = "../Results/figures/sites_ks.png",
   width = 86,
@@
   plot = all_sites$STU$fit,
-  filename = "results/figures/STU.png",
+  filename = "../Results/figures/STU.png",
   width = 90,
@@
   plot = all_sites$BFI$fit,
-  filename = "results/figures/BUR.png",
+  filename = "../Results/figures/BUR.png",
   width = 90,
@@
 
-full_data_path <- "data/FullHBForagingData.csv"
+full_data_path <- "../Data/FullHBForagingData.csv"
 data_raw <- tibble(read.csv(full_data_path))
@@
 london <- st_read(
-  "shapefiles/London_Ward.shp"
+  "../Data/shapefiles/London_Ward.shp"
 )
@@
   plot = merge_plot,
-  filename = "results/figures/site_map.png",
+  filename = "../Results/figures/site_map.png",
   width = 90,
@@
   plot = sites_model_plot,
-  filename = "results/figures/sites_model_plot.png",
+  filename = "../Results/figures/sites_model_plot.png",
   width = 183,
@@
 
+model_sites_fits <- plot_grid(
+  merge_plot_2, ks_plot_dist,
+  labels = c("A", "B"), label_size = 24
+)
+
 ggsave(
-  plot = sites_model_plot,
-  filename = "results/figures/sites_model_plot.svg",
+  plot = model_sites_fits,
+  filename = "../Results/figures/model_sites_fits.png",
   width = 183,
@@
 
-ggsave(
-  plot = sites_model_plot,
-  filename = "results/figures/sites_model_plot.pdf",
-  width = 183,
-  height = 190,
-  units = "mm",
-  dpi = 300
-)
-
-model_sites_fits <- plot_grid(
-  merge_plot_2, ks_plot_dist,
-  labels = c("A", "B"), label_size = 24
-)
-
-ggsave(
-  plot = model_sites_fits,
-  filename = "results/figures/model_sites_fits.png",
-  width = 183,
-  height = 190,
-  units = "mm",
-  dpi = 300
-)
-
 stu_zsl_fit <- plot_grid(
@@
   plot = stu_zsl_fit,
-  filename = "results/figures/stu_zsl_fit.svg",
+  filename = "../Results/figures/stu_zsl_fit.svg",
   width = 183,
@@
   plot = merge_plot,
-  filename = "results/figures/site_map.png",
+  filename = "../Results/figures/site_map.png",
   width = 183,
@@
   plot = sites_model_plot,
-  filename = "results/figures/results_model_plot.png",
+  filename = "../Results/figures/results_model_plot.png",
   width = 183,

### 2022-08-04 inst/publication/Code/landuse_effect_jackknife.Rmd
updated MS with Elli edits

--- 2022-06-29 analysis/landuse_effect_jackknife.Rmd
@@
 # load results of model fit (code to generate found in fit_models.Rmd)
-sitedata <- readRDS("results/site_fit_results.Rda")
+sitedata <- readRDS("../Results/site_fit_results.Rda")
 
@@
 # Load land use data taken (taken from Samuelson et al. 2021)
-urban_lu <- read.csv(file = "data/urban-landuse.csv")
-rural_lu <- read.csv(file = "data/agri-rural-landuse.csv")
+urban_lu <- read.csv(file = "../Data/urban-landuse.csv")
+rural_lu <- read.csv(file = "../Data/agri-rural-landuse.csv")
 
@@
 
-ms_mr <- read.csv(file = "data/site_ms_mr.csv")
+ms_mr <- read.csv(file = "../Data/site_ms_mr.csv")
 
@@
 
-# TLDR:
-
-A jackknife was carried out on the PLS beta regression whereby each site was excluded from the analysis and the fits to the others established.
-This was done to assess if a single site biased the model and should be treated as an outlier.
-
-In the urban dataset, site CAD, with a proportion of scouts value of 1 bears the hallmarks of an outlier, the removal of which influences the results of the study more than the removal of any other point.
-No outliers were identified in the agri-rural dataset.
-
-In addition to showing fits with and without outliers for the urban data, for both environments the jackknifed analysis shows how the loadings vary with each site removed and how the model fits look on the average of the PLS beta values from across the jackknifed analyses.
-These results suggest the agri-rural data is fairly robust to removing individual sites, however, the urban data is less robust, exemplifying the bias introduced by site CAD.
-The urban results are overall more volatile than the agri-rural.
-
-By doing this analysis I find more coherent results for the urban sites and demonstrate that more data is needed to draw more firm conclusions.
-The agri-rural data, although only having ten sites, shows a consistent pattern.
-
-In addition, I also plot the loadings to show the amount of area a particular landuse type corresponds to in the total land of all the sites in that environment as a percentage.
-This allows us to see how much land each land-use type covers.
-A linear regression between the loadings of PC1 and the amount of land identified shows no significant relationship between the two variables, indicating the amount of land is not influencing the PLS results.
-
 # Jackknife analysis to investigate PLS sensitivity to outliers
-
-Looking over the results of the PLS it is apparent that in the urban PLS the beta regression analysis appears skewed by a single site, CAD, which has a proportion of scouts much higher than the others.
-It is important to investigate how outliers are influencing our results, particularly in the urban dataset but also with the agri-rural sites.
-To do this I present an analysis to see if there are any outliers and what effect these have on the results of the PLS and beta regression.
-
-Although site CAD appears to be an outlier, it is prudent to formally establish outliers in both datasets to ensure no other outliers are missed and ensure statistical rigour.
-To do this I will perform a jackknifed PLS regression to identify how land-use type influences the proportion of scouts estimated from our model.
-The purpose of the jackknife is to assess how sensitive the results of the PLS are to a given sample and allow for a more robust assessment of any relationship that is skewed by particular sites.
-
-The jackknife will proceed as follows:
-
-A PLS will be be first ran on the full dataset, and then subsets of the data with a single site, $n_i$, removed.
-Running the PLS on this data yields loadings and the $x$ values for a beta regression to explore the relationship between the primary axis of land-use variation and the proportion of scouts.
-
-Here the land-use categories stay the same, but a single row (site and associated proportion of scouts) is removed.
-The PLS is then carried out on the remaining data.
-This results in a loading value for each land-use type and a beta regressor for each site.
-Once the replicates for $n-1$ data points have been collected (that is, once we have loadings and regressors for the data with each site individually removed) I will examine the distribution of slope values identified from the beta regression.
-I will also the plot the distribution of the loading values for each land-use type as a box plot to see how they vary in relation to the included sites.
-We also then take all the regressor values and calculate the median of these values for each site.
-These medians are then compared against the proportion of scouts in a beta regression.
 
@@
 ## Urban jackknife
-
-Starting with the urban dataset, the data is first jackknifed and the metrics reported and discussed in turn.
 
@@
 
-After running the beta regression of the jackknifed data we can formally assess outliers by examining the distribution of slopes identified with each site removed.
-
 ```{r, urban_jk_slope_dist, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
@@
 
-With only 10 points to investigate the quantiles are not particuarly informative, however, from looking at the graph we can see that the data is left skewed and that there is an influencial point in the left hand bar, as including this point casues the slope to be above 0.6.
-The majority of points sit around 0.8 to 1.1, however the other bars standing alone (0.6 and 1.2) are in range of the body of weight.
-Excluding just the 0.4 value is enough to center the data, suggesting these points lie as extremes rather than outliers.
-With such a small dataset, it is therefore not beyond the realm of reason that the gaps between them would fill in.
-There is fundimentally a judgment call to be made here, and for the reasons I have outlined I think 0.4 should be determined as the only outlier.
-
-Now lets see what this slope references to.
-
 ```{r, urban_jk_slope_dist_value, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
@@
 ```
-
-The site missing from this dataset is CAD.
-This is what I would have expected, but it is good to establish this formally.
-The next task is to now look at how the loadings and betaregression fit looks with and without the inclusion of this site.
 
@@
 
-plots A and C show the PLS beta regression of PC1 to the proportion of scouts and the loadings for PC1 respectively for the full data.
-Plots B and D show the same thing but to the data with site CAD removed.
-The first thing to note is the change in loadings when CAD is removed, with water showing a significant negative correlation, whilst sparse residential flips to become a singificant positive correlation.
-The fit beta regression fit with CAD removed remains singificant, but is less significant than with the outlier included: p < 0.05 vs p < 0.01 with CAD indluced.
-Whats more, the amount of variation explained, Psudo-$R^2$ is much lower at ~0.41.
-Overall, this suggests that landuse type explains some, but not a majority of variation in the proportion of scouts identified in our model, indicating other factors are likely contributing, or we simply dont have enough data.
-
-Overall, this result shows the urban dataset is vulnerable to the site CAD, which seems to be acting as an outlier.
-
 ### Examining how the loadings and fit changes through the jackknifed analysis
-
-The jacknife computes a PLS beta regression for the sample with each site removed.
-It is thus possible to observe how the loadings change with each site removed and also to look at the regression fit of the average of the regressors outlined from the beta regressions run over the data.
 
@@
 
-Plot A shows the fit of the beta regression to the median PC1 values (I.e.
-the median of the pc1 values identified at each itteration in the jackknife).
-Plot B shows the distribution of the loadings identified for each itteration in the jackknife.
-What is clear is that, on average, parks allotments and cemeteries remains sigificantly positive whilst sparse residential is singificantly negative and railways boarders on significant negative.
-The points (black dots) show the outliers for all of these and for sparse residential the outlier is in the position identified from conducting the PLS without CAD.
-
-The fit of the model is also better than for the full data, with the psudo $R^2$ value increasing to ~0.63 vs 0.59, suggesting averaging over jackknife itterations helps reduce variability in the beta regression.
-
-In many cases there are multiple outliers identifies in the distributions, such as woodland and water.
-However, drilling down further into these is unlikely to be that informative, as we are dealing with high dimensional spaces with only 9 data points.
-
 ## Further analysis of sensitivity to site removal and landuse proportions: digging into the the distributions of PLS loadings
-
-Here I show the loadings for each jackknife itteration, that is each PLS ran with a specific site removed. The following plot shows abreak down of the boxplots shown previously.
 
@@
   plot = plt,
-  filename = "results/figures/SI_figures/urban_jk_individual_loadings.png",
+  filename = "../Results/figures/SI_figures/urban_jk_individual_loadings.png",
   width = 120,
@@
 
-There is a fair amount of variation in these plots, more than is shown with the boxplots for sure.
-Many landuse types move around in an out of being significant depending on what site is removed, but parks allotments and cemeteries and sparse residential remain relatively consistent in their position, except for when CAD is removed.
-
 ## How does this compare with the proportion of land-coverage for each site?
-
-I now recreate the main plot but using the percentage of total area that covers as the argument to the size parameter. for the plot.
 
@@
 
-As we can see from the graph and the table, sparse residential coveres the most land, around 39%.
-Parks allotments and cemeteries describes around 8% of the land, whilst water explains around 3%.
-Railway, a singificant contributor, accounts just 1% of the total land area.
-
-It is also important to check if there is a correlation between the loadings and the area for each landuse type.
-This was done using a linear model between the loadings for PC1 and the area associated with that landuse type.
-
 ```{r, urban_jk_loadings_landuse_area_correlation, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
@@
 ```
-
-The results show no significant correlation, indicating the amount of land in each landuse type catagory is not influencing the model results.
 
@@
 
-I am not sure if plotting the averages adds anything and it could be making some questionable assumptions if we really dig into this.
-The fact that the PLS slope changes greatly with CAD removed, whereas the others dont as much, suggests that this site should be excluded from the PLS analysis and the result without it used.
-
 ## Agri-rural PLS
-
-Although the agri-rural data doesnt show any ovbious outliers it is prudent to check it with a jackknife to be sure.
-The first thing to do is look at a histogram of the slopes to see if any one site greatly influences it.
 
@@
 )
-save(urban_jk_loadings, rural_jk_loadings, file = "results/pls-loadings.RData")
+save(urban_jk_loadings, rural_jk_loadings, file = "../Results/pls-loadings.RData")
 rural_jk_pls <- do.call(bind_rows, jkplsdata)
@@
 
-The distribution is pretty well centered.
-There is a gap between 0.85 and 0.8, however this is likely due to having a small sample size rather than outliers.
-Given these results it does not seem nesesery to remove any sites from the agri-rural PLS annalysis.
-
 ### Examining how the loadings and fit changes through the jackknifed analysis
-
-Although we did not see any outliers from the jackknife, it would still be usefull to evaluate the distribution of loadings and median model fit from the jackknifed analysis, as we did for the urban dataset.
-This will show us if the results change much accross the jackknifes.
 
@@
 
-Compared to the same plots for the urban site, the distributions are much tighter, and although there are still individual outliers, it seems that the jackknife shows our results don't change much when sites are excluded.
-
 ## Further analysis of sensitivity to site removal and landuse proportions: digging into the the distributions of PLS loadings
-
-Here I show the loadings for each jackknife itteration, that is each PLS ran with a specific site removed. The following plot shows abreak down of the boxplots shown previously.
 
@@
   plot = plt,
-  filename = "results/figures/SI_figures/agrirural_jk_individual_loadings.png",
+  filename = "../Results/figures/SI_figures/agrirural_jk_individual_loadings.png",
   width = 120,
@@
 
-There is a fair amount of variation in these plots, more than is shown with the boxplots for sure.
-Many landuse types move around in an out of being significant depending on what site is removed, but many, such as arable and built up areas, remain consistent in their position.
-
 ## How does this compare with the proportion of land-coverage for each site?
-
-I now recreate the main plot but using the percentage of total area that covers as the argument to the size parameter. for the plot.
 
@@
 
-As we can see from the graph and the table, arable cover the largest percentage of land included in thie study and has a singificant positive association with PC1.
-Built up area, which has a singificant negative association, makes up around 17% of the land, whilst Non-agricultural unmanaged green space accounts for ~3% of the land.
-Water accounts for less than 1% of the total land coverage and so is association is likely to be an artefact of the small sample size used in this analysis.
-
-It is also important to check if there is a correlation between the loadings and the area for each landuse type.
-This was done using a linear model between the loadings for PC1 and the area associated with that landuse type.
-
 ```{r, rural_jk_loadings_landuse_area_correlation, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
@@
 ```
-
-The results show no significant correlation, indicating the amount of land in each landuse type catagory is not influencing the model results.

### 2022-08-04 inst/publication/Code/supplementary_material.Rmd
updated MS with Elli edits

--- 2022-05-03 analysis/manuscript/supplementary_material.Rmd
@@
 title: "Waggle dance distributions quantify collective foraging in honey bee colonies: Supplementary material"
-# author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
+# author: "Joseph Palmer, Ash Samuelson, Rich Gill, Elli Leadbeater and Vincent Jansen"
 output:

### 2022-08-04 inst/publication/Manuscript/paper.Rmd
updated MS with Elli edits

@@
 knitr::opts_chunk$set(echo = FALSE)
+devtools::load_all()
 library(showtext)
@@
 
-Collective behaviours emerge from the interactions between individuals which vary in their use of different sources of information.
-In social animals, collective behaviours emerge from individuals following behavioural rules which produce a collective level response, yet some conditions make it more profitable for individuals to abandon such structures and use individual rather than collective information.
-In large animal groups, where the ability to track and analyse the interactions between individuals is limited, our understanding of how decisions influence collective processes is largely theoretical.
-Empirically evaluating how individual decisions drive group level behaviours is an important objective with wider consequences for our understanding of the evolution of such complex collective behaviours.
-Here, we explore information use in a model social insect, the honey bee (*Apis* Spp.) and show empirically how variation in individual honey bee's decision making influences collective decision making at a colony level.
-By developing a method to assess a colonies use of the famous waggle dance, we quantify to what extent colonies use the dance to forage collectively and find collective foraging using the dance varies widely over different landscapes.
-These results demonstrate that collective decision making varies in response to environmental conditions.
-Our methods provide new insights into how individual variation in information use shapes an emergent collective process and provides opportunities to further probe the environmental drivers of collective foraging.
+Honeybee foraging is an extraordinary collective behaviour which emerges from one of the most sophisticated communication systems in the animal kingdom: the waggle dance.
+Through a series of feedback loops, this elaborate behaviour -unparalleled in even a single other social insect- allows colonies to decide where to forage without requiring any individual to directly compare resources.
+This behaviour is a textbook example of an emergent collective process, yet a growing body of evidence now shows that using the dance does not always improve foraging success and sometimes not having access to the dance even increases it.
+Although the value of dancing varies across environments, it is not known if and to what extent colonies can adjust dance use in response.
+Here, develop a method to quantify dance use and show that colonies vary the use of the waggle dance in different landscapes, with some colonies relying on dance use almost exclusively and others not at all.
+Further probing of the drivers of this variation in dance use found a colonies use of the dance correlates with different landscape characteristics, suggesting colonies adjust collective decision-making in response to environmental conditions.
+Our results provide new evidence of flexible dance use in honey bee colonies and our methodology outlines a means to identify the ecological conditions in which honeybee colonies rely on dance communication, opening the door to large-scale experimental exploration of the selection pressures that may have driven the evolution of this remarkable collective behaviour.
 
@@
 
-In group living animals, many decisions are taken collectively by integrating information from individuals to produce a social behaviour which extends beyond that of the individual units [@Sumpter2006].
-Such systems are self-organised, relying on the constituents using socially provided information and following behavioural 'rules' to produce a collective outcome [@Couzin2003;@Bonabeau1997].
-However, in addition to information provided by others, individuals also have knowledge of their own (private information) and so can choose to work collectively or individually [@Dall2005;@Rieucau2011;@Gruter2014].
-The benefits of working collectively versus independently vary across environmental contexts and individuals often adjust their own use of information accordingly [@Dechaume-Moncharmont2005;@Beekman2008].
-The impact these individual variations in information use have at a group level, however, is less well understood and difficult to empirically evaluate [@Rieucau2011;@Giardina2008;@Delgado2018].
+In group living animals, many decisions are taken collectively by integrating information from multiple individuals to produce a social behaviour which extends beyond that of the individual units [@Sumpter2006].
+Such systems are self-organised, relying on the constituents using socially provided information and following behavioural 'rules' to produce an emergent collective outcome [@Couzin2003;@Bonabeau1997].
+The behavioural architectures which produce these emergent collective behaviours have in some cases, such as in social insects, grown to dazzling levels of complexity.
+However, this complexity is not always beneficial.
+The benefits of working collaboratively versus independently vary across environmental contexts and individuals often adjust their own use of information accordingly [@Dechaume-Moncharmont2005;@Beekman2008].
+Although the benefits of collective behaviours are known to vary across ecological conditions, in many cases we have a more limited understanding of if and how a group can flexibly use collective information to make use of the most profitable strategy under different conditions.
 
-Understanding the collective ramifications of individual variation in information use is not only important for deciphering how social animals collectively regulate information use [@Gruter2014], relating this to the environmental context of the behaviour is key to illuminating the ultimate evolutionary drivers of collective processes [@Price2015].
-However, establishing the group level effects of individual variation in information use is not trivial and requires measuring information use across animal groups; an often laborious process which requires monitoring individuals' behaviours and interactions.
-This becomes logistically impossible in social insects, which display some of the most intricate group decision-making mechanisms and consist of often thousands of individuals, each using a mix of individual and collective information to make decisions [@Bonabeau1997].
-
-Social insects are defined by their highly coordinated collective responses and so provide ideal systems to investigate the impact of individual variation in information use on emergent collective processes.
-Nevertheless, the sheer quantity of individuals in many colonies is a barrier to further understanding these dynamics.
-Honey bees (Apis spp.) provide an opportunity to sidestep this problem of scale as they provide a detailed record of colony foraging activity in the form of their unique waggle dance -unparalleled in even a single other social insect [@Seeley1995].
-Using the dance, individuals share information about profitable resource locations.
-A series of simple rules which determine when and how much bees perform waggle dances (Fig.
-1) mean that choices between feeding sites occur at the level of the group rather than the individual [@Seeley1991;@Seeley1994A;@Hasenjager2022].
-Despite this, a wealth of research shows that individual foragers frequently do not use dance information to find forage sites and instead rely on their own knowledge [@Gruter2009].
-Moreover, studies investigating the value of the dance have shown that when prevented from communicating via dances, honey bee colonies can achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002;@Dornhaus2004;@Gruter2013;@Price2019].
+The extraordinary waggle dance communication system of the honey bee (*Apis* spp.) is a key example of a behavioural mechanism which produces emergent collective decisions [@Seeley1995].
+Within colonies, a series of simple rules which determine when and how much bees perform dances (Fig.
+1) mean that choices between feeding sites are made by the group rather than individuals [@Seeley1991;@Seeley1994A;@Hasenjager2022].
+For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the richer of two equidistant sources [@Seeley1991], or the closer of two equally rich sources [@Hasenjager2022].
+This mechanism allows a colony to pool individual assessments of the landscape and processes them collectively to identify the most profitable resources to focus on, all without the need for any individual to compare resources.
 
@@
 Although bees that follow dances do not specifically interpret this information on an individual level [@Seeley1991], the resulting over-representation of high-quality sites on the dance-floor means that they are more likely to encounter dances that advertise better forage [@Seeley1995], and provides the colony with a mechanism to select the most profitable resources in their environment [@Seeley1994A]."}
-knitr::include_graphics("../results/figures/manuscript_figures/fig1.png")
+knitr::include_graphics("../Results/figures/manuscript_figures/fig1.png")
 ```
 
-The waggle dance provides a model system to study the group effects of collective and individual information use.
-Although it is likely that individual variation in dance use translates into variations in collective decision making at the colony level [@Donaldson-Matasci2012;@Price2019], this has not been empirically established.
-Individual based models, which explore the collective effects of individuals following different information streams provide an insight into how individual behaviours scale to the level of the group [@Dornhaus2006;@Beekman2008], however, these are difficult to relate to empirical observations and so are limited to showing what a colony under certain conditions might do, rather than quantifying what they are doing.
+Despite it's complexity, research has shown that the ***value*** the waggle dance provides to foraging success varies according to environmental context and colonies without access to dance information achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002;@Dornhaus2004;@Gruter2013;@Price2019].
+Interestingly, studies have shown that individuals often ignore the dance when looking for forage sites [@Gruter2009].
+Whilst it likely that individual variation in dance use translates into variations in collective decision making at the colony level [@Donaldson-Matasci2012;@Price2019], this has yet to be empirically established.
+Identifying if and how colonies vary waggle dance ***use*** is not only important for deciphering how honey bees collectively respond to changes in environmental conditions [@Gruter2014], relating this to the environmental context of the behaviour is key to illuminating its ultimate evolutionary drivers [@Price2015].
+
+Quantifying dance use within colonies requires measuring dance use across the foraging workforce [@Seeley1995].
+This has only previously been quantifiable by documenting foragers' individual search history, which requires tracking individual bees and is time -and labour- intensive [@Seeley1995].
+Individual based models, which explore the collective effects of individuals following different behaviours, provide an insight into how individual information use scales to the level of the group [@Dornhaus2006;@Beekman2008], however, these are difficult to relate to empirical observations and so are limited to showing what a colony under certain conditions might do, rather than quantifying what they are doing.
 Empirically determining how dance information is used at a colony level requires new methods which can identify the behaviours of individual foragers and integrate them with the mechanisms by which the dance functions to make collective decisions.
-Previously, this has only been quantifiable by documenting foragers' individual search history, which requires tracking individual bees and is time -and labour- intensive [@Seeley1995].
 
-Here, we investigate how variation in individual information use influences the collective foraging behaviour of honey bees.
-Using a model based on well-established foraging behaviours [@Seeley1986;@Seeley1991;@Seeley1992;@Seeley1994;@Seeley1994A;@Seeley1995], we examine the distribution of waggle run durations and infer from its shape the relative contribution of dances for resources identified by individual search - “scouting”- versus dances for resources found by following dances - “recruitment”- [@Seeley1995].
-This allows us to quantify the extent to which colonies use dance communication and thus forage collectively.
-To investigate how information use varies across environment, we fit our model to waggle dance observations from 20 sites in an agri-rural and urban environment (10 in each) and evaluated the landscape characteristics which best explained the variance in dance use.
+Here, we derive a mathematical model to quantify waggle dance use in colony foraging by identifying the proportion of foraging trips that are made by bees recruited through waggle dances.
+To validate our model, we fit it to waggle dance observations collected from 20 sites across and agri-rural and urban environment (10 in each).
+We then use our model to explore how variations in landscape structure influences dance use in colony foraging.
 
@@
 The scout distribution is best fit by an exponential (grey fit line), the recruit distribution by a Rayleigh distribution (black line)"}
-knitr::include_graphics("../results/figures/manuscript_figures/fig2.png")
+knitr::include_graphics("../Results/figures/manuscript_figures/fig2.png")
 ```
@@
 The distances reported on the dance floor this are a mixture of the scout and recruiting trips and can be calculated from the distance distributions of the scouting and recruiting trips, taking the reporting bias into account (see Materials and Methods for detail)."}
-knitr::include_graphics("../results/figures/manuscript_figures/fig3.png")
+knitr::include_graphics("../Results/figures/manuscript_figures/fig3.png")
 ```
@@
 Using this data, for each site we fitted both the collective and individual models and used model selection to determine which provided the better explanation of the data, and (if the collective model provided a better fit) to quantify the relative use of social information through estimating the parameter $p$.
-In each case, we calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004;@Clauset2009].
+In each case, we calculated the goodness-of-fit using a Kolmogorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004;@Clauset2009].
 
@@
 Panels show the compliment cumulative frequencies with binned frequency distributions as inset."}
-knitr::include_graphics("../results/figures/manuscript_figures/fig4.png")
+knitr::include_graphics("../Results/figures/manuscript_figures/fig4.png")
 ```
@@
 
-Further quantification of the use of waggle-dance recruitment within all colonies, as a proportion of all foraging trips, can be achieved by extracting the estimated proportion of recruit trips, $1-p$, for each site.
-Since our sites varied in land-use characteristics and potentially thus forage distributions (although in this case, not by design for this study), we investigated whether these estimates might correlate with land-use.
+quantifying the use of waggle-dance recruitment within all colonies, as a proportion of all foraging trips, can be achieved by extracting the estimated proportion of recruit trips, $1-p$, for each site.
+Since our sites varied in land-use characteristics, and thus potentially forage distributions (although in this case, not by design for this study), we investigated whether these estimates might correlate with land-use.
 We first classified the different land-use types of the area surrounding each site [@Samuelson2018] to obtain a standardised land-use profile for the urban and agri-rural environments separately as many land-use types present in urban areas do not occur in agri-rural environments and vice versa (see Materials and Methods).
@@
 managed) green space"}
-knitr::include_graphics("../results/figures/manuscript_figures/fig5.png")
+knitr::include_graphics("../Results/figures/manuscript_figures/fig5.png")
 ```
@@
 
-load("../results/area_coverage.RData")
+load("../Results/area_coverage.RData")
 
@@
 
-Goodness of fit is assessed using the two-sample Kolmorgorov-Smirnov (KS) test [@Goldstein2004] and implemented in R using the ks.boot function of the package 'Matching' in R [@Sekhon2011].
+Goodness of fit is assessed using the two-sample Kolmogorov-Smirnov (KS) test [@Goldstein2004] and implemented in R using the ks.boot function of the package 'Matching' in R [@Sekhon2011].
 
@@
 
-All data and code is available on request and will be made available on zenodo prior to publication.
+All of the data and code required to reproduce this article is contained as within the 'wagglefit' which can be found at (link to finalised Zenodo archive finalised)
+
+https://github.com/joseph-palmer/wagglefit
 

### 2022-08-04 inst/publication/README.Rmd
updated MS with Elli edits

---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Honey bees vary collective decision making in different landscapes

---

This directory contains all the code and data needed to reproduce the article:

(insert article reference)

# Instructions

---

The manuscript is an R-markdown document called `paper.Rmd`. This is rendered into a word document called `paper.doc`. The Supplementary material is a tex file, `supplementary_material.tex`.

All generic code to fit the models is defined in the `wagglefit` package whilst specific code is stored in the `Code` directory. The simulation code is written in python and and can be found in `Code/Simulation`.

`agent_based_sim.Rmd` contains code use to make the simulation plots. These results are stored in `Simulation/Results` and were used to create figures 2 and 3 in an external editing program.
`fit_data.R` contains functions to fit the models to the data used in this study (these functions are called in the r-markdown files).
`fit_models_durations.Rmd` contains code to fit the models to the duration of waggle dance runs. The results are stored in `Results/figures` and used to create figures 4 and 5 in an external editing program.
`landuse_effect_jackknife.Rmd` contains code to run the jackknifed PLS analysis. Results are stored in the `Results/figures` directory and used to create figure 5 in an external editing program.
`supplementary_material.Rmd` contains code to create the supplementary figures and tables included in the supplementary material.

Running these analyses creates the raw figures which are then combined in editing software to produce the manuscript figures used in the publication. The final figures can be found in `Results/figures/manuscript_figures/`

You can then recreate the manuscript by kitting the paper.Rmd file.

### 2022-08-04 inst/publication/Code/fit_data.R

@@ -1,0 +1,252 @@
+suppressMessages(library(dplyr))
+library(magrittr)
+library(tibble)
+suppressMessages(library(lubridate))
+library(wagglefit)
+suppressMessages(library(rlang))
+library(ggplot2)
+suppressMessages(library(tidyr))
+devtools::load_all()
+
+#' Get waggle dance data from file and calculate foraging distance
+#' @descriptionFor For specific data in the 'data' directory read it in,
+#' calculate foraging distance and get just the required columns
+#' @importFrom rlang .data
+#' @export
+#'
+get_data <- function() {
+  full_data_path <- "data/FullHBForagingData.csv"
+  data <- tibble(read.csv(full_data_path))
+  data <- data %>%
+    mutate(
+      date = ymd(.data$date),
+      foraging_distance = calc_dist(.data$duration.seconds)
+    ) %>%
+    select(
+      date,
+      site,
+      foraging_distance
+    )
+  return(data)
+}
+
+#' Run my analysis
+run_analysis <- function(data, model = "all", n = 5, upper = 5, bounds = NULL,
+                         xtol = 0, verbose_r = FALSE, verbose = FALSE) {
+  best_result <- fit_mutliple(
+    data$foraging_distance,
+    model = model,
+    n = n,
+    upper = upper,
+    bounds = bounds,
+    verbose_r = verbose_r,
+    verbose = verbose,
+    xtol = xtol
+  )
+  best_result$data_name <- model
+  return(best_result)
+}
+
+#' Run on specific site k times and output likelihood space and plots
+fit_collective_model_to_data <- function(data, bounds, k = 10, n = 10,
+                                         verbose_r = FALSE, verbose = FALSE) {
+  # run optimisation k times and return best one
+  result_data <- purrr::map_df(
+    seq_len(k),
+    ~ {
+      print(paste("Itteration", .x))
+      result <- run_analysis(
+        data,
+        model = "collective", n = n, upper = 10, bounds = bounds,
+        verbose_r = verbose_r, verbose = verbose
+      )
+      return(
+        data.frame(
+          ll = result$fmax,
+          p = result$est[1],
+          bs = result$est[2],
+          br = result$est[3],
+          as = result$est[4],
+          ar = result$est[5]
+        )
+      )
+    }
+  )
+
[…]

### 2022-08-26 inst/publication/Manuscript/paper.Rmd
updated MS

@@
 ---
-title: "Honey bees vary collective decision making in different landscapes"
-# author: "Joseph Palmer, Ash Samuelson, Rich Gill, Elli Leadbeater and Vincent A. A. Jansen"
+title: "Honeybees vary collective decision making in different landscapes"
+# author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater, Richard Gill, and Vincent A. A. Jansen"
 output:
@@
 
-Honeybee foraging is an extraordinary collective behaviour which emerges from one of the most sophisticated communication systems in the animal kingdom: the waggle dance.
-Through a series of feedback loops, this elaborate behaviour -unparalleled in even a single other social insect- allows colonies to decide where to forage without requiring any individual to directly compare resources.
-This behaviour is a textbook example of an emergent collective process, yet a growing body of evidence now shows that using the dance does not always improve foraging success and sometimes not having access to the dance even increases it.
-Although the value of dancing varies across environments, it is not known if and to what extent colonies can adjust dance use in response.
-Here, develop a method to quantify dance use and show that colonies vary the use of the waggle dance in different landscapes, with some colonies relying on dance use almost exclusively and others not at all.
-Further probing of the drivers of this variation in dance use found a colonies use of the dance correlates with different landscape characteristics, suggesting colonies adjust collective decision-making in response to environmental conditions.
-Our results provide new evidence of flexible dance use in honey bee colonies and our methodology outlines a means to identify the ecological conditions in which honeybee colonies rely on dance communication, opening the door to large-scale experimental exploration of the selection pressures that may have driven the evolution of this remarkable collective behaviour.
+Honeybee (*Apis mellifera*) colony foraging decisions arise from the waggle dances of individual foragers, refined and filtered through a series of feedback loops that produce emergent collective behaviour.
+This process is a key example of animal communication at the height of eusociality, but a growing body of evidence suggests that its value for colony foraging success is heavily dependent on local ecology.
+Here, we develop a method to quantify the extent to which a colony forages collectively based on dance-decoding and show how it can be used to relate variation in collective foraging to land-use.
+We show that the extent to which colonies forage collectively varies across landscape types, such that it is sometimes almost entirely dominant but at other times almost negligible.
+By providing a means to identify the ecological conditions in which waggle dances are important on a large scale and without experimental manipulation, our methodology opens the door to exploration of the selection pressures that may have driven the evolution of this remarkable group behaviour.
 
@@
 
-In group living animals, many decisions are taken collectively by integrating information from multiple individuals to produce a social behaviour which extends beyond that of the individual units [@Sumpter2006].
-Such systems are self-organised, relying on the constituents using socially provided information and following behavioural 'rules' to produce an emergent collective outcome [@Couzin2003;@Bonabeau1997].
-The behavioural architectures which produce these emergent collective behaviours have in some cases, such as in social insects, grown to dazzling levels of complexity.
-However, this complexity is not always beneficial.
-The benefits of working collaboratively versus independently vary across environmental contexts and individuals often adjust their own use of information accordingly [@Dechaume-Moncharmont2005;@Beekman2008].
-Although the benefits of collective behaviours are known to vary across ecological conditions, in many cases we have a more limited understanding of if and how a group can flexibly use collective information to make use of the most profitable strategy under different conditions.
-
-The extraordinary waggle dance communication system of the honey bee (*Apis* spp.) is a key example of a behavioural mechanism which produces emergent collective decisions [@Seeley1995].
-Within colonies, a series of simple rules which determine when and how much bees perform dances (Fig.
-1) mean that choices between feeding sites are made by the group rather than individuals [@Seeley1991;@Seeley1994A;@Hasenjager2022].
+In group living animals, many decisions are taken collectively by integrating information from multiple individuals to produce group behaviour that extends beyond that of the individual units [@Sumpter2006].
+Such systems are self-organised, relying on the use of simple behavioural 'rules' that filter social information to produce an emergent collective outcome [@Couzin2003;@Bonabeau1997].
+In honeybee colonies, and other highly eusocial insects, the behavioural architectures that produce these emergent behaviours have become particularly complex.
+Honeybee colony foraging is driven by the extraordinary waggle dances of individual foragers, which famously communicate food source locations.
+A series of rules that determine when and how much bees dance (Fig.
+1) mean that choices between feeding sites are made by the group [@Seeley1991;@Seeley1994A;@Hasenjager2022].
 For example, because the number of dance circuits performed by a bee on returning from a food source reflects the net energetic benefits of the trip, more of the colony's workforce will be recruited to the richer of two equidistant sources [@Seeley1991], or the closer of two equally rich sources [@Hasenjager2022].
-This mechanism allows a colony to pool individual assessments of the landscape and processes them collectively to identify the most profitable resources to focus on, all without the need for any individual to compare resources.
+This mechanism allows a colony to collectively allocate foraging effort adaptively without the need for any individual to compare resources
 
 ```{r, fig.cap="Figure 1.
-**The honeybee waggle dance carries information about the location of a resource.** The duration of the waggle run indicates the distance to the resource and the angle of the dance relative to the vertical indicates the direction of the resource (circle in right panel), relative to the direction of the sun [@Frisch1993].
+**The honeybee waggle dance carries information about the location of a resource.** The angle of the dance relative to the vertical indicates the direction of the resource (circle in right panel), relative to the direction of the sun [@Frisch1993], the duration of the waggle run indicates the distance to the resource.
 Through the observation and decoding of the waggle dance, a colony's dance floor provides a unique opportunity to eavesdrop on the communication and decision making leading to collective foraging decisions.
@@
 
-Despite it's complexity, research has shown that the ***value*** the waggle dance provides to foraging success varies according to environmental context and colonies without access to dance information achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002;@Dornhaus2004;@Gruter2013;@Price2019].
-Interestingly, studies have shown that individuals often ignore the dance when looking for forage sites [@Gruter2009].
-Whilst it likely that individual variation in dance use translates into variations in collective decision making at the colony level [@Donaldson-Matasci2012;@Price2019], this has yet to be empirically established.
-Identifying if and how colonies vary waggle dance ***use*** is not only important for deciphering how honey bees collectively respond to changes in environmental conditions [@Gruter2014], relating this to the environmental context of the behaviour is key to illuminating its ultimate evolutionary drivers [@Price2015].
+Despite its complexity, research has shown that the ***value*** collective foraging provides to foraging success is likely to be realized only in limited ecological contexts, the nature of which has proved hard to identify empirically [@Dornhaus1999;@Sherman2002;@Dornhaus2004;@Dornhaus2006;@Okada2012;@Price2019;@Donaldson-Matasci2012;@Donaldson-Matasci2014@Nurnberger2017].
+Since dancing foragers use the top of a comb to reference the sun's azimuth [@Frisch1993], it is possible to render dances meaningless by placing combs on a horizontal plane, and thus force foragers to search out food individually, preventing collective decisions.
+Initial work following these methods tentatively linked the benefits of collective foraging to landscape heterogeneity [@Sherman2002], but empirical attempts to systematically test this hypothesis have failed to provide support [@Donaldson-Matasci2012;@Nurnberger2017], and dance disruption in challenging environments has even sometimes been associated with higher, rather than lower, foraging success [@Price2019].
+Such mixed evidence has limited progress in identifying the ecological conditions that originally shaped the evolution of the dance communication system, and thus the collective behaviour of these extraordinary eusocial insects [@Price2015].
 
-Quantifying dance use within colonies requires measuring dance use across the foraging workforce [@Seeley1995].
-This has only previously been quantifiable by documenting foragers' individual search history, which requires tracking individual bees and is time -and labour- intensive [@Seeley1995].
-Individual based models, which explore the collective effects of individuals following different behaviours, provide an insight into how individual information use scales to the level of the group [@Dornhaus2006;@Beekman2008], however, these are difficult to relate to empirical observations and so are limited to showing what a colony under certain conditions might do, rather than quantifying what they are doing.
-Empirically determining how dance information is used at a colony level requires new methods which can identify the behaviours of individual foragers and integrate them with the mechanisms by which the dance functions to make collective decisions.
-
-Here, we derive a mathematical model to quantify waggle dance use in colony foraging by identifying the proportion of foraging trips that are made by bees recruited through waggle dances.
-To validate our model, we fit it to waggle dance observations collected from 20 sites across and agri-rural and urban environment (10 in each).
-We then use our model to explore how variations in landscape structure influences dance use in colony foraging.
+Here, we adopt an alternative approach that seeks to identify those environments in which colonies forage collectively and those in which individual search is dominant.
+It is well-established that that bees do not always choose to dance on return from a foraging trip [@Frisch1993], and furthermore, that individuals often ignore the dance when looking for forage sites [@Gruter2009].
+Thus, colonies are likely to vary in their reliance on collective foraging, and identifying the ecological variables that drive this variation offers a means to establish what contexts may have been key to dance evolution.
+To quantify reliance on the dance within colonies, we adopt a modelling approach that draws on the distribution of foraging distances reported on the dance floor of individual hives, thus avoiding the time -and labour- intensive need to document foragers' individual search histories [@Seeley1995].
+Our model provides a description of the distribution of the distances reported on the dance floor and can be used to quantify to what extent colonies rely on collective or individual foraging.
+We apply our model to waggle dance observations collected from 20 sites across and agri-rural and urban environment (10 in each), and find that some colonies forage almost entirely collectively while others are dominated by individual search.
+We then use our model to explore how variations in landscape structure influence dance use in colony foraging within our dataset, demonstrating how our methodology could be used to seek out those ecological circumstances under which collective foraging is important.
 
@@
 
-The patterns of waggle dance-encoded foraging distances might differ under scenarios where colonies rely on individual search versus recruitment.
-To test this, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
+To test whether colonies that rely on individual search might differ from those that rely on recruitment in terms of the distribution of foraging distributions reported by dancing bees, we simulated honeybees foraging in a landscape where resource patches were randomly placed in the environment.
 Foragers could locate these under two different strategies: either acting as a scout and locating resources themselves, or following a recruit strategy and locating resources by following a random dance from the dance floor [@Seeley1992] (Fig.
 2a,b, details in Materials and Methods).
-As it is known in the simulation which individuals in the hive forage under what strategy, we can compare the distributions of foraging distances reported on the dance floor by each type of forager.
+As it is known in the simulation what proportion of individuals in the hive forage under what strategy, we can compare the distributions of foraging distances reported on the dance floor by each type of forager.
 Fig 2c shows that the shapes of the resource distance distributions for bees engaging in the two types of foraging trips are different.
-The distance distribution for the scout trips is akin close to that of an exponential distribution (Fig.
+The distance distribution for the scout trips is close to that of an exponential distribution (Fig.
 2c), which is the nearest neighbour distance distribution for foragers operating in a one-dimensional environment (see Materials and Methods).
@@
 ```{r, fig.cap="Figure 2.
-**Simulating honey bee foraging**.
+**Simulating honeybee foraging**.
 In our simulation model with scouting only (A), foragers leave the hive on a search path (white lines) and continue until they encounter a resource (circles, colours indicate different resource quality).
@@
 
-We constructed two models of foraging: an individual model describing a hypothetical colony that relies only on individual search to find foraging sites, and a collective model which describes a colony using waggle dance communication to exploit resources.
+We constructed two models of foraging: an individual model describing a hypothetical colony that relies only on individual search to find foraging sites, and a collective model which describes a colony in which a proportion of bees follow waggle dances to find resources.
 In the individual model, all trips are "scout" trips, but as recruitment becomes more important, the proportion of “recruit” trips will increase.
@@
 ```{r, fig.cap="Figure 4.
-**The honey bee foraging model fitted to data from 20 hives**.
+**The honeybee foraging model fitted to data from 20 hives**.
 (A) Location of study hives in Southern England, shaded area in the main plot indicates Greater London.
@@
 
-Our results show that, whilst colony-level foraging is mostly comprised of a mixture of scout and recruit foraging trips, in some circumstances, colony foraging can be better described by individual foraging alone.
+Since our model of individual foraging provides a good fit to colony foraging in 11 different sites, and provides a more parsimonious description of foraging than a model of collective foraging in 4 sites, our results indicate that, whilst colony-level foraging is mostly comprised of a mixture of scout and recruit foraging trips, in some circumstances, colony foraging can be better described by individual foraging alone.
 Thus, in some environments, the majority of foraging trips involve scouting to find new food sites rather than recruitment through dances.
-Note that this does not imply that these bees do not engage in dance following, because bees regularly follow dances but choose not to visit the advertised site [@Gruter2011], although it has also been shown that bees may cease dance following if it is proving unproductive [@Gruter2011;@Price2019].
-Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individual's foraging behaviour over a landscape.
-As our results show that our model of individual foraging provides a good fit to colony foraging in 11 different sites, and provides a more parsimonious description of foraging than a model of collective foraging in 4 sites, these findings support the idea that individual decision making can dominate colony foraging and demonstrates further evidence for flexible waggle dance use by honey bee colonies.
+Note that this does not imply that these bees do not engage in dance following, because bees regularly follow dances but choose not to visit the advertised site [@Gruter2011], so it may be the case that dances are followed and ignored, not followed, or not performed.
+However, though the mechanism by which flexibility is achieved at the individual level remains unclear, we have demonstrated that colonies themselves do not always forage collectively.
 
-quantifying the use of waggle-dance recruitment within all colonies, as a proportion of all foraging trips, can be achieved by extracting the estimated proportion of recruit trips, $1-p$, for each site.
+Quantifying the use of waggle-dance recruitment within all colonies, as a proportion of all foraging trips, can be achieved by extracting the estimated proportion of recruit trips, $1-p$, for each site.
 Since our sites varied in land-use characteristics, and thus potentially forage distributions (although in this case, not by design for this study), we investigated whether these estimates might correlate with land-use.
@@
 We then performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see Materials and Methods) to determine the principal components that represent combinations of land-use types which explained the most variation in waggle dance use within agri-rural and urban environments.
-As for one of our sites neither model provided a plausible description, this site was removed from the PLS analysis.
+For one of our sites neither model provided a plausible description, so this site was removed from the PLS analysis.
 
-Waggle dance use varied widely regardless of environment, ranging from 0 to 1 in both the urban and agri-rural environments whilst showing a significant, positive interaction with waggle dance use.
-In the agri-rural sites, a single principle component explained ~73% of the variation in waggle dance use (beta regression: $R^2 = 0.73$, $\phi$ = 4.9, p < 0.05, Fig.
+Waggle dance use varied widely regardless of environment, ranging from absolute ($p = 0$) to none ($p = 1$) in both the urban and agri-rural environments.
+This proportion varied systematically with land-use.
+In the agri-rural sites, a single principal component explained ~73% of the variation in waggle dance use (beta regression: $R^2 = 0.73$, $\phi$ = 4.9, p < 0.05, Fig.
 5A).
-This finding is also shown in the urban environment which also shows a single principle component explained ~73% of the variation in waggle dance use (beta regression: $R^2$ = 0.73, $\phi$ = 10.4, p < 0.05, Fig.
-5C).
+Similarly, in the urban environment, a single principal component explained ~73% of the variation in waggle dance use (beta regression: $R^2$ = 0.73, $\phi$ = 10.4, p < 0.05, Fig.
+5C; note that since the land-use categories differ between the two environments, this single component differs between the two categories).
 Probing the individual loadings of these principal components allows us to evaluate the driving mechanisms in more detail.
-As our sample size is limited  (10 urban and 9 agri-rural sites), we used jackknife resampling to evaluate the robustness of our results to influential points (see Materials and Methods, Supplementary Material) and evaluate how these land-use types contribute to the principle component which significantly correlates with waggle dance use.
+As our sample size is limited  (10 urban and 9 agri-rural sites, reduced from 10), we used jackknife resampling to evaluate the robustness of our results to influential points (see Materials and Methods, Supplementary Material) and evaluate how these land-use types contribute to the principal component which significantly correlates with waggle dance use.
 
@@
 Arable land in the UK is typically considered nutritionally poor for bees (note that oilseed rape fields were not included within this category), while there is evidence to suggest that the residential areas that were captured within the “built-up” category are forage-rich hotspots, typically supporting relatively high bee diversity and abundance within gardens [@Baldock2019].
+Thus, our agri-rural results suggest that collective foraging, driven by waggle dance recruitment, reduces as food becomes sparser.
 
-In the urban environment, our findings were more equivocal.
-The first principal component correlated negatively with the dominant land-use type by coverage, sparse residential (land coverage ~35%; Table 1), as would be expected since such land includes a high proportion of gardens.
+In the urban environment the first principal component correlated negatively with the dominant land-use type by coverage, sparse residential (land coverage ~35%; Table 1), since such land includes a high proportion of gardens, this is in keeping with the hypothesis that collective foraging increases in flower-rich areas.
 However, this relationship is very sensitive to site removal through the jackknife sampling (Supplementary Fig.
@@
 3.).
+Overall, our findings for the urban environments present further support for a relationship between collective foraging reliance and land-use.
 
@@
 
-Understanding how individual variation in information use scales to produce flexible collective processes at the group level is a key challenge in behavioural ecology.
-Here, we have presented a method to quantify information use at a collective level in a model social insect.
-After verifying that this method accurately captures empirical patterns of collective and individual foraging, we have shown how it can be used to evaluate how environmental factors shape collective foraging at the colony level.
-Whilst recruitment is well known to occur through other mechanisms besides the waggle dance, such as through olfactory cues [@Arenas2007], the exceptionally close fit of our model to waggle run durations underlines the importance of the waggle dance in honey bee foraging.
-Further examining our estimates of dance use show that a colonies use of the waggle dance varies across environments and does not appear to be randomly determined.
+By developing a method to quantify the extent to which a colony forages collectively, here we have shown that colony foraging varies in how much it is dominated by the waggle dance recruitment system.
+Furthermore, we have demonstrated how this variation can be compared across land-use types to evaluate how environmental factors shaping collective foraging at the colony level.
+Whilst recruitment is well known to occur through other mechanisms besides the waggle dance, such as through olfactory cues [@Arenas2007], the exceptionally close fit of our model to waggle run durations underlines the importance of the waggle dance in honeybee foraging.
 
-Within both the agri-rural and urban environments, dance use had a significant interaction with land-use type, ranging from total dance use to complete reliance on individual information.
-The notion that individual variation in information use influences the collective effect has only previously been inferred theoretically.
-Our findings demonstrate this empirically.
-Individual-based models [@Dornhaus2006;@Beekman2008;@Price2019] have outlined how variations at an individual level could drive group level dynamics, however, without verifying this upon empirical data the insights remain theoretical.
-Here, we have shown how an individual based model can be used to better understand a system in a way which allows us to build more informative models which can be empirically verified.
-Not only do our findings add support for our understanding of how the dance functions to make collective decisions [@Seeley1995], but they also provide a new dimension to our understanding of how variation at an individual level influences the collective level.
+Within both the agri-rural and urban environments, collective foraging varied with land-use type, ranging from near-total reliance on recruitment to the dominance of individual search.
+Understanding the circumstances under which colonies forage collectively is a step towards identifying those types of environmental feature that may have made collective foraging beneficial for ancestral honeybees [@Price2015].
+Our findings hint that collective foraging may more commonly occur in rich environments, rather than those where food is hard to find.
+However, this suggestion is speculative for two reasons.
+Firstly, we used a pre-existing dataset and did not systematically vary landscape traits that may be important, such as forage heterogeneity.
+We hypothesise that arable land is likely to be more patchy than built-up areas, however, this requires further testing through larger-scale study designed for the purpose.
 
+One of the main outcomes of our study is the development of our methodology to quantifying collective foraging, which can be applied without manipulating colonies through dance disorientation.
+By combining this tool with further dance-decoding datasets, we envisage a picture should emerge of the circumstances under which honeybee colonies forage collectively.
+To this end, a major obstacle has recently been overcome thanks to the development automated decoding protocols [@Wario2017;@Wild2021].
 Whilst the benefits of dance communication are known to vary across environments [@Donaldson-Matasci2012;@Gruter2011;@Schurch2014;@Price2019], our results show that colonies vary their use in response, affording the colony a flexible dance use strategy which they can tailor to different conditions to maximise foraging success.
 Identifying the circumstances in which dance communication is important have been hampered by the requirement to monitor real-world colony weight [@Sherman2002;@Dornhaus2004;Gruter2013;@Price2019;@Dornhaus2006;@Nurnberger2017;@Kirchner1998] -a noisy proxy of foraging success that is influenced by many other abiotic and biotic factors, including colony health- over ecologically credible time periods [@Schurch2014].
-Doing this at a scale that allows inter-colony variation in foraging environments at sufficient replication is a major logistical hurdle (but see [@Nurnberger2017]), particularly given that multiple landscape variables may interact to determine the utility of dance communication.
+Doing this at a scale that allows inter-colony variation in foraging environments at sufficient replication is a major logistical hurdle (but see @Nurnberger2017), particularly given that multiple landscape variables may interact to determine the utility of dance communication.
 Our method circumvents this issue by providing a time and labour efficient methodology to quantify this key collective behaviour and map the environment along the major axis of honeybee information use.
@@
 Examining the land-use types which contribute most to the variation in waggle dance use in each of our study environments reveals insights into the drivers of waggle dance use.
-Our agri-rural results suggest a potential decrease in reliance on waggle dance recruitment as resources become harder to find and foraging trips become longer [@Samuelson2021], broadly agreeing with the findings from other theoretical [@Dornhaus2006] (11) and empirical [@Price2019] studies which examined foraging performance with and without the dance.
+Our agri-rural results suggest a potential decrease in reliance on waggle dance recruitment as resources become harder to find and foraging trips become longer [@Samuelson2021], broadly agreeing with some of the findings from other theoretical [@Dornhaus2006] and empirical [@Price2019] studies which examined foraging performance with and without the dance, but contrasting with others.
 These results suggest that, even though collective foraging is not always beneficial, in environments where high-quality resources are present but relatively scarce, collective foraging is being carried out by colonies and under these conditions exchanging social information through the waggle dance confers benefits.
@@
 
-Although the number of colonies used in this analysis are comparable to similar studies evaluating collective foraging in honey bees [@Sherman2002;@Dornhaus2004;@Price2019;@Nurnberger2017;@Kirchner1998],  probing the landscape drivers of colony level waggle dance use in more detail requires more colony level data.
-As longer term effects, such as resource stability [@Schurch2014], may also influence waggle dance use, these data need to be gathered over sufficient time frames to ensure these effects are accounted for.
-Recently, technological advances have emerged which enable colony metrics to be collected faster, more accurately and over greater time spans than could be gathered by hand, allowing individuals to be tracked within colonies and theories of individual behaviour to be evaluated in more depth than could have been done previously [@Wild2021].
-Here, we have presented an accurate mathematical description of colony foraging which extends our ability to quantify collective behaviour across environments.
-With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”.
-Our methodology thus provides a means of analysing such large data sets and provide a framework to start exploring how plasticity in collective decision making can help to buffer the effects of landscape change, and in this case, how flexibility in the use of the waggle dance may help to exploit resources when experiencing novel landscape mosaics, such as in highly urbanized conurbations.
+Although the number of colonies used in this analysis are comparable to similar studies evaluating collective foraging in honeybees [@Sherman2002;@Dornhaus2004;@Price2019;@Nurnberger2017;@Kirchner1998],  probing the landscape drivers of colony level waggle dance use in more detail requires more colony level data.
+As longer term effects, such as resource stability [@Schurch2014], may also influence waggle dance use, these data need to be gathered over sufficient time to ensure these effects are accounted for.
+Here, we have presented an accurate mathematical description of colony foraging which extends our ability to quantify collective behaviour across environments, providing a methodology to assist in identifying the selection pressures that may have driven the evolution of this remarkable behaviour.
+This will elucidate how plasticity in collective decision making can help to buffer the effects of landscape change, and may help to exploit resources when experiencing novel landscape mosaics, such as in highly urbanized conurbations.
 
@@
 
-All of the data and code required to reproduce this article is contained as within the 'wagglefit' which can be found at (link to finalised Zenodo archive finalised)
-
-https://github.com/joseph-palmer/wagglefit
+All of the data and code required to reproduce this article can be found at:
 

### 2022-08-26 README.Rmd
compressed the publication analysis folder
updated readme

@@
 
-when citing wagglefit please use (insert citation for publication).
+## Honeybees vary collective decision making in different landscapes
 
-## Honey bees vary collective decision making in different landscapes
-
-All code used in the publication (insert citation link) is stored in `inst/publication/`.
+All code used in the publication *Honeybees vary collective decision making in different landscapes* is stored in `inst/publication.tar.gz/`. To extract it use `tar -xzvf publication.tar.gz`.
 

### 2022-08-26 inst/publication/Code/Simulation/Code/Agents.py
removed publication folder and replaced with compression

@@ -1,245 +1,0 @@
-#!/usr/bin/env python3
-"""Classes of agents (workers, scouts, flowers)"""
-__appname__  = "Agents.py"
-__author__   = "Joseph Palmer <joseph.palmer18@imperial.ac.uk>"
-__version__  = "0.0.1"
-__date__     = "12-2019"
-
-import numpy as np
-import matplotlib.pyplot as plt
-import Methods
-import config
-import copy
-from shapely.geometry import Point, Polygon
-import shapely
-
-class Flower:
-    # Attributes #
-    location = (None, None)
-    location_point = None
-    distance = None
-    resource_level = None
-    exists = 0
-
-    # Methods #
-    def __init__(self, location, quality_array):
-        self.location = location
-        self.location_point = Point(*location)
-        self.distance = Methods.getdist(*location)
-        self.resource_level = 2000 #np.random.randint(10, 5000)
-        self.quality_array = quality_array #quality_array[np.random.randint(0, len(quality_array))]
-        self.exists = 1
-        return None
-
-    def remove_flower(self):
-        self.location = (None, None)
-        self.distance = None
-        self.resource_level = None
-        self.exists = 0
-        return None
-
-    def visit(self):
-        if self.resource_level > 0:
-            self.resource_level -= 1
-        else:
-            self.remove_flower()
-        return None
-
-class Honeybee:
-    flower = None
-    distance = None
-    mode = None # -1 is scout, 1 is employed recruit, 0 is unemployed recruit
-
-class Scout(Honeybee):
-    def __init__(self):
-        self.mode = -1
-        return None
-
-    def _drawpath(self, s1, s2, flowers, infa=[], i=0, limmit=10, plot=True):
-        angle = np.random.uniform(0, 360)
-        distance = np.random.uniform(0, 1)
-        base_fa = [
-            tuple(s1),
-            (s1[0], distance+s1[1]),
-            (s2[0], distance+s2[1]),
-            tuple(s2)
-        ]
-        fa = Polygon(base_fa)
-        fa = shapely.affinity.rotate(fa, angle, origin=s1)
-        if plot:
-            x,y = fa.exterior.xy
-            plt.plot(x,y, color="red")
-        a, b = list(fa.exterior.coords)[1:3]
-        infa_idx = [fa.contains(j.location_point) for j in flowers]
-        [infa.append(ii) for ii in np.array([j.distance for j in flowers])[infa_idx]]
[…]

### 2022-08-26 inst/publication/Code/Simulation/Code/Methods.py

@@ -1,42 +1,0 @@
-#!/usr/bin/env python3
-"""Helper functions for honeybee foraging simulation model"""
-__appname__  = "Methods.py"
-__author__   = "Joseph Palmer <joseph.palmer18@imperial.ac.uk>"
-__version__  = "0.0.2"
-__date__     = "02-2020"
-
-import numpy as np
-import matplotlib.pyplot as plt
-
-def getdist(x2, y2):
-    # return distance of resource to hive in 2D plane
-    x1 = 0
-    y1 = 0
-    return np.sqrt(((x1 - x2)**2 + (y1 - y2)**2))
-
-def get_ccdf(data):
-    # Get CCDF for the data
-    sorted_data = np.sort(data)[::-1]
-    prob = np.linspace(0, 1, len(data))
-    return np.column_stack((sorted_data, prob))
-
-def plotccdf(data, title, log=True):
-    # Plot cdf data
-    cdf_dat = get_ccdf(data)
-    if log:
-        plt.plot(cdf_dat[:,0][1:], np.log10(cdf_dat[:,-1][1:]), "o", markersize=2)
-    else:
-        plt.plot(cdf_dat[:,0][1:], cdf_dat[:,-1][1:], "o", markersize=2)
-    plt.ylabel("ln Probability")
-    plt.xlabel("Foraging distance (km)")
-    plt.gca().spines['top'].set_visible(False)
-    plt.gca().spines['right'].set_visible(False)
-    plt.title(title)
-    plt.show()
-    return None
-
-
-def qual2dance(q, x, alpha):
-    res = q/(1+alpha*x)-1
-    res[res<=0] = 0
-    return res

### 2022-08-26 inst/publication/Code/Simulation/Code/config.py

@@ -1,9 +1,0 @@
-#!/usr/bin/env python3
-"""Global variables for honeybee foraging simulation model"""
-__appname__  = "config.py"
-__author__   = "Joseph Palmer <joseph.palmer18@imperial.ac.uk>"
-__version__  = "0.0.2"
-__date__     = "02-2020"
-
-FLOWERDICT = {}
-BROADCASTED_DANCES = {}

### 2022-08-26 inst/publication/Code/Simulation/Code/run.py

@@ -1,193 +1,0 @@
-#!/usr/bin/env python3
-"""Run simulation of honeybee foraging"""
-__appname__  = "run.py"
-__author__   = "Joseph Palmer <joseph.palmer18@imperial.ac.uk>"
-__version__  = "0.0.2"
-__date__     = "02-2020"
-
-import numpy as np
-import pandas as pd
-import scipy.stats
-import matplotlib.pyplot as plt
-import Agents
-import Methods
-import config
-
-def create_flowers(lwr, upr, res, quality_array):
-    # Simulation window parameters
-    r = upr-lwr
-    xx0 = 0
-    yy0 = 0
-
-    areaTotal = np.pi*r**2
-
-    # Point process parameters
-    lambda0 = res
-
-    # Simulate Poisson point process
-    numbPoints = scipy.stats.poisson(lambda0*areaTotal).rvs()
-    theta = 2*np.pi*scipy.stats.uniform.rvs(0,1,((numbPoints,1)))
-    rho = r*np.sqrt(scipy.stats.uniform.rvs(0,1,((numbPoints,1))))
-
-    # Convert from polar to Cartesian coordinates
-    xx = rho * np.cos(theta)
-    yy = rho * np.sin(theta)
-
-    # Shift centre of disk to (xx0,yy0)
-    xx=xx+xx0
-    yy=yy+yy0
-
-    # Set locations as flowers
-    locations = np.column_stack((xx, yy))
-    flowers = np.apply_along_axis(Agents.Flower, 1, locations, quality_array)
-    fdict = {
-        i : flowers[i] for i in range(res)
-    }
-    return fdict
-
-def initialize(lwr, upr, res, nbees, nscouts, quality_array):
-    # create flowers in environment and store globaly.
-    config.FLOWERDICT = create_flowers(lwr, upr, res, quality_array)
-
-    # create honeybee population
-    nrecruits = nbees-nscouts
-    scouts = np.array([Agents.Scout() for i in range(nscouts)])
-    recruits = np.array([Agents.Recruit() for i in range(nrecruits)])
-    return (scouts, recruits)
-
-def flower_turnover(n, lwr, upr, k, quality_array):
-    # select n flowers and remove
-    assert (n <= len(config.FLOWERDICT.keys())), "N > flowers in environment!"
-    rmv_idx = np.random.choice(list(config.FLOWERDICT.keys()), n, replace=False)
[…]

### 2022-08-26 inst/publication/Code/fit_data.R

@@ -1,252 +1,0 @@
-suppressMessages(library(dplyr))
-library(magrittr)
-library(tibble)
-suppressMessages(library(lubridate))
-library(wagglefit)
-suppressMessages(library(rlang))
-library(ggplot2)
-suppressMessages(library(tidyr))
-devtools::load_all()
-
-#' Get waggle dance data from file and calculate foraging distance
-#' @descriptionFor For specific data in the 'data' directory read it in,
-#' calculate foraging distance and get just the required columns
-#' @importFrom rlang .data
-#' @export
-#'
-get_data <- function() {
-  full_data_path <- "data/FullHBForagingData.csv"
-  data <- tibble(read.csv(full_data_path))
-  data <- data %>%
-    mutate(
-      date = ymd(.data$date),
-      foraging_distance = calc_dist(.data$duration.seconds)
-    ) %>%
-    select(
-      date,
-      site,
-      foraging_distance
-    )
-  return(data)
-}
-
-#' Run my analysis
-run_analysis <- function(data, model = "all", n = 5, upper = 5, bounds = NULL,
-                         xtol = 0, verbose_r = FALSE, verbose = FALSE) {
-  best_result <- fit_mutliple(
-    data$foraging_distance,
-    model = model,
-    n = n,
-    upper = upper,
-    bounds = bounds,
-    verbose_r = verbose_r,
-    verbose = verbose,
-    xtol = xtol
-  )
-  best_result$data_name <- model
-  return(best_result)
-}
-
-#' Run on specific site k times and output likelihood space and plots
-fit_collective_model_to_data <- function(data, bounds, k = 10, n = 10,
-                                         verbose_r = FALSE, verbose = FALSE) {
-  # run optimisation k times and return best one
-  result_data <- purrr::map_df(
-    seq_len(k),
-    ~ {
-      print(paste("Itteration", .x))
-      result <- run_analysis(
-        data,
-        model = "collective", n = n, upper = 10, bounds = bounds,
-        verbose_r = verbose_r, verbose = verbose
-      )
-      return(
-        data.frame(
-          ll = result$fmax,
-          p = result$est[1],
-          bs = result$est[2],
-          br = result$est[3],
-          as = result$est[4],
-          ar = result$est[5]
-        )
-      )
-    }
-  )
-
[…]

### 2022-08-26 README.Rmd
added DOI badge to readme

@@
 [![test-coverage](https://github.com/joseph-palmer/wagglefit/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/joseph-palmer/wagglefit/actions/workflows/test-coverage.yaml)
+[![DOI](https://zenodo.org/badge/353413244.svg)](https://zenodo.org/badge/latestdoi/353413244)
 

### 2023-05-02 inst/publication/Code/fit_models_durations.Rmd
Added AIC variation

@@
 devtools::load_all()
-
 knitr::opts_chunk$set(
@@
 library(tibble)
-source("fit_data.R")
+source("Code/fit_data.R")
 library(kableExtra)
-
+print(getwd())
 library(showtext)
@@
   pull()
+
 ```
@@
 # set waggle dance duration in seconds as foraging distance for analysis
-data <- read.csv("data/FullHBForagingData.csv")
+data <- read.csv("Data/FullHBForagingData.csv")
 
@@
 
+### Fits to all urban and all agri-rural sites
+
+```{r, ZSL, message = FALSE, warning = FALSE}
+urban_sites <- c("BLO", "BUR", "CAD", "GIL", "HHS", "HOR", "MAK", "MPA", "SWP", "ZSL")
+rural_sites <- c("BEL", "BFI", "HER", "MEL", "SAU", "SOM", "SRA", "STU", "YAL") # exclude ROT due to bad fit
+
+urban_data <- alldata %>%
+  filter(site %in% urban_sites)
+
+rural_data <- alldata %>%
+  filter(site %in% rural_sites)
+
+# set up bounds for the collective model
+p_bnds <- c(0, 1.0)
+bs_bnds <- c(1.0e-6, 10)
+br_bnds <- c(1.0e-6, 10)
+as_bnds <- c(1.0e-12, 1.5)
+ar_bnds <- c(1.0e-12, 0.5)
+collective_bounds <- rbind(
+  p_bnds, bs_bnds,
+  br_bnds, as_bnds,
+  ar_bnds
+)
+
+# set up bounds for the individual model
+bs_bnds <- c(1.0e-6, 50)
+as_bnds <- c(1.0e-12, 1.3)
+individual_bounds <- rbind(
+  bs_bnds, as_bnds
+)
+
+# set coordinates for histogram subplot
+subplot_coords <- c(0.1, 1.8, -8, -2)
+
+# urban
+all_sites[["urban"]] <- run_wagglefit_analysis(
+  "urban", urban_data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[["urban"]]$fit_result %>%
+  kbl() %>%
+  kable_classic(full_width = F)
+
+# rural
+all_sites[["rural"]] <- run_wagglefit_analysis(
+  "rural", rural_data, collective_bounds, individual_bounds, subplot_coords
+)
+
+all_sites[["rural"]]$fit_result %>%
+  kbl() %>%
+  kable_classic(full_width = F)
+
+```
+
 ## Overall findings
+
+```{r, make-aic-comp, message = FALSE, warning = FALSE}
+# group all site results together
+df <- map(all_sites, 1) %>%
+  bind_rows()
+
+all_sites
+
+# filter for best model
+sitedata <- df %>%
+  group_by(site) %>%
+  slice(which.min(AIC))
+
+sitedata$LU <- NA
+sitedata2 <- sitedata %>%
+  mutate(LU= ifelse(
+    site %in% urban_sites, 'Urban',
+    ifelse(
+      site %in% rural_sites,
+      'Rural',
+      paste0('Total_', site)
+    ))) %>%
+  filter(site != 'ROT')
+
+print(sitedata2, n=22)
+
+sitedata2 %>%
+  group_by(LU) %>%
+  summarise(AIC=sum(AIC))
+
+```
 

### 2023-05-02 inst/publication/Code/landuse_effect_jackknife.Rmd
Added AIC variation

@@
 
+set_ggplot_theme <- function() {
+  theme_set(
+    theme_classic() +
+      theme(
+        text = element_text(family = "URWHelvetica", size = 18),
+        legend.position = "none",
+        # plot.title = element_text(hjust = 0.5),
+        # axis.text.x = element_text(angle = 0,  hjust=1)
+      )
+  )
+}
+
+set_ggplot_theme()
+
+knitr::opts_chunk$set(
+  dev = "png",
+  dpi = 300,
+  collapse = TRUE,
+  comment = "#>",
+  fig.width = 12,
+  fig.height = 8
+)
+
 library(cowplot)
@@
   pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
+
+```
+
+```{r, diff-p-sites}
+
+urban_p <- urban_lu %>% select(site, p)
+rural_p <- rural_lu %>% select(site, p)
+
+urban_p$LU = "Urban"
+rural_p$LU = "Agri-rural"
+
+comb <- rbind(urban_p, rural_p) %>% mutate(p = 1-p)
+comb_ar <- comb %>% filter(LU == "Agri-rural")
+comb_u <- comb %>% filter(LU == "Urban")
+
+plt <- ggplot(comb, aes(y = p, x = LU)) +
+  geom_violin(fill="gray") +
+  # geom_boxplot(width=0.05) +
+  ylab("Waggle dance use") +
+  xlab("Environment")
+
+ggsave(
+  plot = plt,
+  filename = "../Results/figures/SI_figures/urban_rural_violin.png",
+  dpi = 300
+)
+
+ggsave(
+  plot = plt,
+  filename = "../Results/figures/SI_figures/urban_rural_violin_nobox.svg",
+  dpi = 300
+)
+
+res = chisq.test(comb_ar$p, comb_u$p)
+res
 
@@
 
-```{r, urban_jk_outlier_removed, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
+```{r, urban_jk_outlier_removed, message = FALSE, error = FALSE, warning = FALSE, echo = TRUE}
 
@@
 )
+
+print(model_pls)
+
 plsdata <- data.frame(
@@
 
+make_plots_func <- function(urban_jk_loadings) {
   plt <- ggplot(data = urban_jk_loadings, aes(
@@
   )) +
-  geom_point(aes(colour = landuse_type)) +
+    geom_point(aes(colour = landuse_type), size = 8) +
     geom_segment(aes(
       x = landuse_type, xend = landuse_type,
-    y = 0, yend = loadings, colour = landuse_type
+      y = 0, yend = loadings, colour = landuse_type, size = .1
     )) +
@@
       yintercept = 0, color = "black"
-  ) +
-  geom_text(
-    data = urban_loading_text,
-    aes(x = landuse_type, y = result, label = landuse_type),
-    hjust = 0.5,
-    vjust = 0.5,
-    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
     ) +
@@
     ) +
+    ylab("Correlation") +
     coord_flip() +
@@
       legend.position = "None",
-    text = element_text(size = 15),
-    strip.text.x = element_text(size = 18),
-    axis.title.x = element_text(size = 18)
-  ) +
-  scale_fill_manual(
+      axis.title.x = element_text(size = 80),
+      axis.text.x = element_text(size = 60)
+    ) +
+    scale_colour_manual(
       values = c(
-      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
-      "Non.agricultural.unmanaged.green.space" = "#228833",
-      "water" = "#4477AA", other = "#BBBBBB"
-    )
-  ) +
-  facet_wrap(~site_removed, ncol = 2)
+        "sparse.residential" = "#EE6677",
+        "railway" = "#CCBB44", "parks.allotments.cemeteries" = "#228833"
+      )
+    ) +
+    facet_wrap(~site_removed, nrow = 1) +
+    theme(
+      strip.text = element_text(size = 80)
+    )
+  return(plt)
+}
+
+plt <- make_plots_func(urban_jk_loadings %>% filter(site_removed != "None"))
+
 ggsave(
@@
   filename = "../Results/figures/SI_figures/urban_jk_individual_loadings.png",
-  width = 120,
-  height = 200,
+  width = 800,
+  height = 400,
   units = "mm",
@@
 )
-plt
+
+# bur_blo <- urban_jk_loadings %>% filter(site_removed %in% c("BLO", "BUR")) %>%
+#   make_plots_func()
+
+# ggsave(
+#   plot = bur_blo,
+#   filename = "../Results/figures/SI_figures/bur_blo_loadings.png",
+#   width = 120,
+#   height = 200,
+#   units = "mm",
+#   dpi = 300
+# )
+
+# cad_gil <- urban_jk_loadings %>% filter(site_removed %in% c("CAD", "GIL")) %>%
+#   make_plots_func()
+
+# ggsave(
+#   plot = cad_gil,
+#   filename = "../Results/figures/SI_figures/cad_gil_loadings.png",
+#   width = 120,
+#   height = 200,
+#   units = "mm",
+#   dpi = 300
+# )
+
+# hhs_hor <- urban_jk_loadings %>% filter(site_removed %in% c("HHS", "HOR")) %>%
+#   make_plots_func()
+
+# ggsave(
+#   plot = hhs_hor,
+#   filename = "../Results/figures/SI_figures/hhs_hor_loadings.png",
+#   width = 120,
+#   height = 200,
+#   units = "mm",
+#   dpi = 300
+# )
+
+# mak_mpa <- urban_jk_loadings %>% filter(site_removed %in% c("MAK", "MPA")) %>%
+#   make_plots_func()
+
+# ggsave(
+#   plot = mak_mpa,
+#   filename = "../Results/figures/SI_figures/mak_mpa_loadings.png",
+#   width = 120,
+#   height = 200,
+#   units = "mm",
+#   dpi = 300
+# )
+
+# swp_zsl <- urban_jk_loadings %>% filter(site_removed %in% c("SWP", "ZSL")) %>%
+#   make_plots_func()
+
+# ggsave(
+#   plot = swp_zsl,
+#   filename = "../Results/figures/SI_figures/swp_zsl_loadings.png",
+#   width = 120,
+#   height = 200,
+#   units = "mm",
+#   dpi = 300
+# )
+
 ```
@@
 
+make_plots_func2 <- function(rural_jk_loadings){
   plt <- ggplot(data = rural_jk_loadings, aes(
@@
       yintercept = 0, color = "black"
-  ) +
-  geom_text(
-    data = rural_loading_text,
-    aes(x = landuse_type, y = result, label = landuse_type),
-    hjust = 0.5,
-    vjust = 0.5,
-    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
     ) +
@@
       legend.position = "None",
-    text = element_text(size = 15),
-    strip.text.x = element_text(size = 18),
-    axis.title.x = element_text(size = 18)
-  ) +
-  scale_fill_manual(
+      axis.title.x = element_text(size = 80),
+      axis.text.x = element_text(size = 60)
+    ) +
+    scale_colour_manual(
       values = c(
-      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
-      "Non.agricultural.unmanaged.green.space" = "#228833",
-      "water" = "#4477AA", other = "#BBBBBB"
-    )
-  ) +
-  facet_wrap(~site_removed, ncol = 2)
+        "arable" = "#228833", "built.up.area" = "#EE6677",
+        "Non.agricultural.unmanaged.green.space" = "#CCBB44",
+        "water" = "#4477AA"
+      )
+    ) +
+    facet_wrap(~site_removed, nrow = 1) +
+    theme(
+      strip.text = element_text(size = 80)
+    ) +
+    xlab("PC1 Loadings")
+  return(plt)
+}
+
+plt <- make_plots_func2(rural_jk_loadings %>% filter(site_removed != "None"))
+
 ggsave(
   plot = plt,
-  filename = "../Results/figures/SI_figures/agrirural_jk_individual_loadings.png",
-  width = 120,
-  height = 220,
+  filename = "../Results/figures/SI_figures/rural_jk_individual_loadings.png",
+  width = 800,
+  height = 400,
   units = "mm",
@@
 )
-plt
+
+# bel_bfi <- rural_jk_loadings %>% filter(site_removed %in% c("BEL", "BFI")) %>%
+#   make_plots_func2()
+
+# ggsave(
+#   plot = bel_bfi,
+#   filename = "../Results/figures/SI_figures/bel_bfi_loadings.png",
+#   width = 120,
+#   height = 200,
+#   units = "mm",
+#   dpi = 300
+# )
+
+# her_mel <- rural_jk_loadings %>% filter(site_removed %in% c("HER", "MEL")) %>%
+#   make_plots_func2()
+
+# ggsave(
+#   plot = her_mel,
+#   filename = "../Results/figures/SI_figures/her_mel_loadings.png",
+#   width = 120,
+#   height = 200,
+#   units = "mm",
+#   dpi = 300
+# )
+
+# sau_som <- rural_jk_loadings %>% filter(site_removed %in% c("SAU", "SOM")) %>%
+#   make_plots_func2()
+
+# ggsave(
+#   plot = sau_som,
+#   filename = "../Results/figures/SI_figures/sau_som_loadings.png",
+#   width = 120,
+#   height = 200,
+#   units = "mm",
+#   dpi = 300
+# )
+
+# sra_stu_yal <- rural_jk_loadings %>% filter(site_removed %in% c("SRA", "STU", "YAL")) %>%
+#   make_plots_func2()
+
+# ggsave(
+#   plot = sra_stu_yal,
+#   filename = "../Results/figures/SI_figures/sra_stu_yal_loadings.png",
+#   width = 120,
+#   height = 200,
+#   units = "mm",
+#   dpi = 300
+# )
+
 ```

### 2023-05-02 inst/publication/Manuscript/paper.Rmd
Added AIC variation

@@
   set_table_properties(layout = "autofit")
+
+library(xtable)
+
+landuse_area %>%
+  arrange(Environment, desc(`% coverage`)) %>%
+  mutate(`% coverage` = round(`% coverage`, digits = 1)) %>%
+
+  xtable()
+
 ```

### 2023-05-02 inst/publication/Manuscript/supplementary_material.tex
Added AIC variation

@@
 \DeclareCaptionLabelFormat{sfig}{#1 S#2}
+\DeclareCaptionLabelFormat{stab}{#1 S#2}
 \captionsetup[figure]{labelformat=sfig}
+\captionsetup[table]{labelformat=stab}
 
@@
 
-\title{Waggle dance distributions quantify collective foraging in honey bee colonies: Supplementary material}
+\title{Honeybees vary communication and collective decision making across landscapes: Supplementary material}
 %\author{}
@@
 \captionof{table}{Fit results for the fitting of the waggle dance model to each site.}
-\input{../results/figures/SI_figures/fittable.tex}
+\input{../Results/figures/SI_figures/fittable.tex}
 
@@
 	\centering
-	\includegraphics{../results/figures/SI_figures/allsites_plot.png}
+	\includegraphics{../Results/figures/SI_figures/allsites_plot.png}
 	\caption{\textbf{Cumulative frequency distribution} of waggle dance durations and fits of the collective (red line) and individual (blue line) models for all 20 sites}
@@
 
+\clearpage
+
+\captionof{table}{Percentage area covered for each land-use type in the agri-rural and urban environments in the sites studied.}
+\begin{table}[ht]
+	\centering
+	\begin{tabular}{rllr}
+		\hline
+		Environment & Land-use & \% coverage \\ 
+		\hline
+		Agri-rural & Arable & 28.30 \\ 
+		Agri-rural & Pasture & 23.30 \\ 
+		Agri-rural & Woodland & 21.10 \\ 
+		Agri-rural & Built Up Area & 15.00 \\ 
+		Agri-rural & Fruit & 3.10 \\ 
+		Agri-rural & Oilseed Rape & 2.90 \\ 
+		Agri-rural & Non Agricultural Unmanaged Green Space & 2.80 \\ 
+		Agri-rural & Non Agricultural Managed Green Space & 1.80 \\ 
+		Agri-rural & Other Agricultural & 1.50 \\ 
+		Agri-rural & Water & 0.20 \\ 
+		Urban & Sparse Residential & 34.80 \\ 
+		Urban & Continuous Central & 24.30 \\ 
+		Urban & Dense Residential & 21.80 \\ 
+		Urban & Parks Allotments Cemeteries & 7.90 \\ 
+		Urban & Woodland & 4.20 \\ 
+		Urban & Water & 3.40 \\ 
+		Urban & Amenity Grassland & 2.60 \\ 
+		Urban & Railway & 1.00 \\ 
+		\hline
+	\end{tabular}
+\end{table}
+
 \section*{Jackknifed partial least squares analysis}
@@
 	\centering
-	\includegraphics{../results/figures/SI_figures/agrirural_jk_individual_loadings.png}
+	\includegraphics{../Results/figures/SI_figures/agrirural_jk_individual_loadings.png}
 	\caption{\textbf{Loadings of PLS calculated for each site removed for the agri-rural sites}.
 Each plot shows the loadings of the first principle component with that site removed from the analysis, showing the individual points making up the overall box plot loadings in Fig 4b.}
-\end{figure}
-
-\begin{figure}[h]
-	\centering
-	\includegraphics{../results/figures/SI_figures/urban_jk_individual_loadings.png}
-	\caption{\textbf{Loadings of PLS calculated for each site removed for the urban sites}.
-Each plot shows the loadings of the first principle component with that site removed from the analysis, showing the individual points making up the overall box plot loadings in Fig 4d.}
 \end{figure}

### 2023-05-02 inst/publication/README.Rmd
updated readme
Added AIC variation

@@
 
-This directory contains all the code and data needed to reproduce the article:
+This directory contains all the code and data needed to reproduce the article: *Honeybees vary collective decision making in different landscapes*
 
-(insert article reference)
-
-# Instructions
+## Instructions
 
 ---
-
-The manuscript is an R-markdown document called `paper.Rmd`. This is rendered into a word document called `paper.doc`. The Supplementary material is a tex file, `supplementary_material.tex`.
 

### 2023-05-02 inst/publication/Code/Simulation/Code/Agents.py

@@ -1,0 +1,245 @@
+#!/usr/bin/env python3
+"""Classes of agents (workers, scouts, flowers)"""
+__appname__  = "Agents.py"
+__author__   = "Joseph Palmer <joseph.palmer18@imperial.ac.uk>"
+__version__  = "0.0.1"
+__date__     = "12-2019"
+
+import numpy as np
+import matplotlib.pyplot as plt
+import Methods
+import config
+import copy
+from shapely.geometry import Point, Polygon
+import shapely
+
+class Flower:
+    # Attributes #
+    location = (None, None)
+    location_point = None
+    distance = None
+    resource_level = None
+    exists = 0
+
+    # Methods #
+    def __init__(self, location, quality_array):
+        self.location = location
+        self.location_point = Point(*location)
+        self.distance = Methods.getdist(*location)
+        self.resource_level = 2000 #np.random.randint(10, 5000)
+        self.quality_array = quality_array #quality_array[np.random.randint(0, len(quality_array))]
+        self.exists = 1
+        return None
+
+    def remove_flower(self):
+        self.location = (None, None)
+        self.distance = None
+        self.resource_level = None
+        self.exists = 0
+        return None
+
+    def visit(self):
+        if self.resource_level > 0:
+            self.resource_level -= 1
+        else:
+            self.remove_flower()
+        return None
+
+class Honeybee:
+    flower = None
+    distance = None
+    mode = None # -1 is scout, 1 is employed recruit, 0 is unemployed recruit
+
+class Scout(Honeybee):
+    def __init__(self):
+        self.mode = -1
+        return None
+
+    def _drawpath(self, s1, s2, flowers, infa=[], i=0, limmit=10, plot=True):
+        angle = np.random.uniform(0, 360)
+        distance = np.random.uniform(0, 1)
+        base_fa = [
+            tuple(s1),
+            (s1[0], distance+s1[1]),
+            (s2[0], distance+s2[1]),
+            tuple(s2)
+        ]
+        fa = Polygon(base_fa)
+        fa = shapely.affinity.rotate(fa, angle, origin=s1)
+        if plot:
+            x,y = fa.exterior.xy
+            plt.plot(x,y, color="red")
+        a, b = list(fa.exterior.coords)[1:3]
+        infa_idx = [fa.contains(j.location_point) for j in flowers]
+        [infa.append(ii) for ii in np.array([j.distance for j in flowers])[infa_idx]]
[…]

### 2023-05-02 inst/publication/Code/Simulation/Code/Methods.py

@@ -1,0 +1,42 @@
+#!/usr/bin/env python3
+"""Helper functions for honeybee foraging simulation model"""
+__appname__  = "Methods.py"
+__author__   = "Joseph Palmer <joseph.palmer18@imperial.ac.uk>"
+__version__  = "0.0.2"
+__date__     = "02-2020"
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+def getdist(x2, y2):
+    # return distance of resource to hive in 2D plane
+    x1 = 0
+    y1 = 0
+    return np.sqrt(((x1 - x2)**2 + (y1 - y2)**2))
+
+def get_ccdf(data):
+    # Get CCDF for the data
+    sorted_data = np.sort(data)[::-1]
+    prob = np.linspace(0, 1, len(data))
+    return np.column_stack((sorted_data, prob))
+
+def plotccdf(data, title, log=True):
+    # Plot cdf data
+    cdf_dat = get_ccdf(data)
+    if log:
+        plt.plot(cdf_dat[:,0][1:], np.log10(cdf_dat[:,-1][1:]), "o", markersize=2)
+    else:
+        plt.plot(cdf_dat[:,0][1:], cdf_dat[:,-1][1:], "o", markersize=2)
+    plt.ylabel("ln Probability")
+    plt.xlabel("Foraging distance (km)")
+    plt.gca().spines['top'].set_visible(False)
+    plt.gca().spines['right'].set_visible(False)
+    plt.title(title)
+    plt.show()
+    return None
+
+
+def qual2dance(q, x, alpha):
+    res = q/(1+alpha*x)-1
+    res[res<=0] = 0
+    return res

### 2023-05-02 inst/publication/Code/Simulation/Code/config.py

@@ -1,0 +1,9 @@
+#!/usr/bin/env python3
+"""Global variables for honeybee foraging simulation model"""
+__appname__  = "config.py"
+__author__   = "Joseph Palmer <joseph.palmer18@imperial.ac.uk>"
+__version__  = "0.0.2"
+__date__     = "02-2020"
+
+FLOWERDICT = {}
+BROADCASTED_DANCES = {}

### 2023-05-02 inst/publication/Code/Simulation/Code/run.py

@@ -1,0 +1,193 @@
+#!/usr/bin/env python3
+"""Run simulation of honeybee foraging"""
+__appname__  = "run.py"
+__author__   = "Joseph Palmer <joseph.palmer18@imperial.ac.uk>"
+__version__  = "0.0.2"
+__date__     = "02-2020"
+
+import numpy as np
+import pandas as pd
+import scipy.stats
+import matplotlib.pyplot as plt
+import Agents
+import Methods
+import config
+
+def create_flowers(lwr, upr, res, quality_array):
+    # Simulation window parameters
+    r = upr-lwr
+    xx0 = 0
+    yy0 = 0
+
+    areaTotal = np.pi*r**2
+
+    # Point process parameters
+    lambda0 = res
+
+    # Simulate Poisson point process
+    numbPoints = scipy.stats.poisson(lambda0*areaTotal).rvs()
+    theta = 2*np.pi*scipy.stats.uniform.rvs(0,1,((numbPoints,1)))
+    rho = r*np.sqrt(scipy.stats.uniform.rvs(0,1,((numbPoints,1))))
+
+    # Convert from polar to Cartesian coordinates
+    xx = rho * np.cos(theta)
+    yy = rho * np.sin(theta)
+
+    # Shift centre of disk to (xx0,yy0)
+    xx=xx+xx0
+    yy=yy+yy0
+
+    # Set locations as flowers
+    locations = np.column_stack((xx, yy))
+    flowers = np.apply_along_axis(Agents.Flower, 1, locations, quality_array)
+    fdict = {
+        i : flowers[i] for i in range(res)
+    }
+    return fdict
+
+def initialize(lwr, upr, res, nbees, nscouts, quality_array):
+    # create flowers in environment and store globaly.
+    config.FLOWERDICT = create_flowers(lwr, upr, res, quality_array)
+
+    # create honeybee population
+    nrecruits = nbees-nscouts
+    scouts = np.array([Agents.Scout() for i in range(nscouts)])
+    recruits = np.array([Agents.Recruit() for i in range(nrecruits)])
+    return (scouts, recruits)
+
+def flower_turnover(n, lwr, upr, k, quality_array):
+    # select n flowers and remove
+    assert (n <= len(config.FLOWERDICT.keys())), "N > flowers in environment!"
+    rmv_idx = np.random.choice(list(config.FLOWERDICT.keys()), n, replace=False)
[…]

### 2023-05-02 inst/publication/Code/fit_data.R

@@ -1,0 +1,252 @@
+suppressMessages(library(dplyr))
+library(magrittr)
+library(tibble)
+suppressMessages(library(lubridate))
+library(wagglefit)
+suppressMessages(library(rlang))
+library(ggplot2)
+suppressMessages(library(tidyr))
+devtools::load_all()
+
+#' Get waggle dance data from file and calculate foraging distance
+#' @descriptionFor For specific data in the 'data' directory read it in,
+#' calculate foraging distance and get just the required columns
+#' @importFrom rlang .data
+#' @export
+#'
+get_data <- function() {
+  full_data_path <- "Data/FullHBForagingData.csv"
+  data <- tibble(read.csv(full_data_path))
+  data <- data %>%
+    mutate(
+      date = ymd(.data$date),
+      foraging_distance = calc_dist(.data$duration.seconds)
+    ) %>%
+    select(
+      date,
+      site,
+      foraging_distance
+    )
+  return(data)
+}
+
+#' Run my analysis
+run_analysis <- function(data, model = "all", n = 5, upper = 5, bounds = NULL,
+                         xtol = 0, verbose_r = FALSE, verbose = FALSE) {
+  best_result <- fit_mutliple(
+    data$foraging_distance,
+    model = model,
+    n = n,
+    upper = upper,
+    bounds = bounds,
+    verbose_r = verbose_r,
+    verbose = verbose,
+    xtol = xtol
+  )
+  best_result$data_name <- model
+  return(best_result)
+}
+
+#' Run on specific site k times and output likelihood space and plots
+fit_collective_model_to_data <- function(data, bounds, k = 10, n = 10,
+                                         verbose_r = FALSE, verbose = FALSE) {
+  # run optimisation k times and return best one
+  result_data <- purrr::map_df(
+    seq_len(k),
+    ~ {
+      print(paste("Itteration", .x))
+      result <- run_analysis(
+        data,
+        model = "collective", n = n, upper = 10, bounds = bounds,
+        verbose_r = verbose_r, verbose = verbose
+      )
+      return(
+        data.frame(
+          ll = result$fmax,
+          p = result$est[1],
+          bs = result$est[2],
+          br = result$est[3],
+          as = result$est[4],
+          ar = result$est[5]
+        )
+      )
+    }
+  )
+
[…]
